% AbsoluteLayer Selector — Family D Patent (Method, System, Medium)
\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{xcolor}

\title{AbsoluteLayer Selector\\Family D Patent (Method, System, Medium)}
\author{Jonathan Washburn}
\date{\today}

\begin{document}
\maketitle

\begin{center}
{\large Camera-Ready Patent Draft}\\[4pt]
{\normalsize Simultaneous Gate-Identity Solve; Anchor-Free Derivation of $\tau_0,\ \ell_0,\ c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$}
\end{center}

\section*{Field}
This disclosure relates to computer-implemented calibration for instruments and simulators. It provides a method, system, and non-transitory medium for selecting a unique \emph{AbsoluteLayer} by solving simultaneous \emph{dimensionless gate identities} and, from that solution, deriving absolute scales and constants ($\tau_0,\ \ell_0,\ c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$) \emph{without external anchors}.

\section*{Background}
Calibration frameworks commonly rely on external standards or tuned parameters to set absolute scales. These practices introduce drift and reduce reproducibility across devices and sites. Prior approaches do not provide a \emph{general} machinery to: (i) express calibration constraints as dimensionless \emph{gate identities}; (ii) solve them \emph{simultaneously} to produce a \emph{unique} layer; and (iii) emit absolute scales/constants \emph{anchor-free} with auditable evidence.

\section*{Summary}
Disclosed is an \emph{AbsoluteLayer Selector} that:
\begin{enumerate}[leftmargin=*]
  \item constructs a set of \textbf{dimensionless gate identities} $\mathcal{I}=\{\Phi_j(\mathbf{z})=0\}_{j=1}^m$ over factorized displays;
  \item solves the identities \textbf{simultaneously} under declared tolerance bands and convergence criteria (\emph{MeetsBands});
  \item enforces \textbf{uniqueness} (\emph{UniqueCalibration}): a single admissible solution exists; otherwise the calibration fails closed;
  \item \textbf{derives} $\tau_0,\ \ell_0,\ c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$ \textbf{anchor-free} from the solution; and
  \item records signed logs summarizing identities, residuals, solver status, solution values, and outputs.
\end{enumerate}

\section*{Definitions}
\noindent\textbf{Gate identities.} A set of dimensionless constraints that must hold simultaneously at the AbsoluteLayer; examples include K-identities and a Planck-gate identity:
\begin{align*}
\text{K-identities:}\quad &\frac{\tau_{\mathrm{rec}}}{\tau_0} = \frac{\lambda_{\mathrm{kin}}}{\ell_0} = K,\\
\text{Planck gate:}\quad &\frac{c^3\,\lambda_{\mathrm{rec}}^2}{\hbar\,G} = C_{\mathrm{layer}}\ (\text{constant by layer}).
\end{align*}

\noindent\textbf{Simultaneous solve.} A numerical or analytic procedure that enforces all identities together (e.g., root finding, constrained optimization) subject to tolerance bands and convergence criteria (\emph{MeetsBands}).

\noindent\textbf{UniqueCalibration.} A uniqueness condition requiring a single admissible solution in the declared band; multiple or no solutions cause fail-closed behavior.

\noindent\textbf{Anchor-free derivation.} Emitting absolute scales $\tau_0,\ \ell_0$ and constants $c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$ from the identity solution without referencing external standards.

\noindent\textbf{Compliance artifacts.} Signed, hash-addressed logs including identity forms, residuals, solver settings, tolerance bands, solution vectors, and emitted outputs.

\section*{Claims}
\setlist[enumerate]{leftmargin=*,labelsep=0.5em,itemsep=0.6em,topsep=0.4em}

\subsection*{Independent Claims}
\begin{enumerate}
\item \textbf{(Method)} A computer-implemented method for selecting an absolute calibration layer, comprising:
\begin{enumerate}[label=(\alph*)]
  \item constructing a set of dimensionless gate identities $\{\Phi_j(\mathbf{z})=0\}_{j=1}^m$ from factorized displays;\label{step:identities}
  \item solving the identities \emph{simultaneously} subject to declared tolerance bands and convergence criteria (\emph{MeetsBands});\label{step:solve}
  \item verifying \emph{UniqueCalibration} that a single admissible solution exists; otherwise failing closed with diagnostics;\label{step:unique}
  \item deriving $\tau_0,\ \ell_0,\ c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$ \emph{anchor-free} from the solution; and\label{step:derive}
  \item emitting signed, hash-addressed logs comprising identity definitions, residuals, solver status, solution, and emitted outputs.\label{step:logs}
\end{enumerate}

\item \textbf{(System)} A system comprising one or more processors and memory configured to: (i) construct dimensionless gate identities; (ii) solve them simultaneously under tolerance/convergence criteria; (iii) verify uniqueness and fail closed if violated; (iv) derive $\tau_0,\ \ell_0,\ c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$ anchor-free; and (v) record signed logs.

\item \textbf{(Non-transitory medium)} A non-transitory computer-readable medium storing instructions that, when executed, cause a system to perform the method of claim 1.
\end{enumerate}

\subsection*{Dependent Claims}
\begin{enumerate}[resume]
\item (Method) wherein the gate identities include K-identities $\tau_{\mathrm{rec}}/\tau_0 = \lambda_{\mathrm{kin}}/\ell_0$ and a Planck-gate identity $(c^3\,\lambda_{\mathrm{rec}}^2)/(\hbar G)=C_{\mathrm{layer}}$.
\item (Method) wherein additional identities are included to overdetermine the system, and the solver minimizes a residual norm subject to band constraints.
\item (Method) wherein \emph{MeetsBands} comprises numerical tolerances, maximum iterations, and residual thresholds; failure to meet criteria causes fail-closed behavior with a structured diagnostic report.
\item (Method) further comprising sensitivity analysis of the solution to identity perturbations with declared robustness thresholds.
\item (Method) wherein emitted outputs include confidence intervals derived from identity tolerances and solver sensitivity.
\item (System) further comprising a compliance API exposing identity sets, residuals, solver status, solution vectors, outputs, and signatures.
\item (Method) wherein identity construction is policy-governed to prevent mixing of identities from distinct layers in the same solve.
\item (Method) wherein logs are optionally published to a transparency registry for third-party verification.
\item (System) wherein failure modes include: (i) no solution in band; (ii) multiple solutions; (iii) inconsistency between identities; each triggering a fail-closed verdict and diagnostic export.
\item (Method) wherein anchor-free derivation includes emission of $\tau_0,\ \ell_0$ scales and constants $c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$ without reference to external standards or tuned parameters.
\end{enumerate}

\section*{Brief Description of the Drawings}
\begin{enumerate}[leftmargin=*]
  \item Fig. 1: Identity graph and simultaneous solve flow (\emph{UniqueCalibration} $\wedge$ \emph{MeetsBands}).
  \item Fig. 2: Example identities (K-identities, Planck gate) and tolerance bands.
  \item Fig. 3: Anchor-free derivation pipeline for $\tau_0,\ \ell_0,\ c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$.
  \item Fig. 4: Signed logging and compliance API signals.
\end{enumerate}

\section*{Detailed Description}
\subsection*{Identity Construction}
Gate identities are formed as dimensionless equalities from canonically factorized displays, e.g., K-identities equating independent routes to a common invariant and a Planck-gate identity constraining layer-constant ratios. Policies may declare additional identities to overdetermine the system for robustness.

\subsection*{Simultaneous Solve and Uniqueness}
All identities are enforced together via root finding or constrained optimization. Band constraints (\emph{MeetsBands}) specify numeric tolerances and convergence thresholds. A uniqueness check (\emph{UniqueCalibration}) requires exactly one admissible solution; otherwise the calibration fails closed with diagnostics.

\subsection*{Anchor-Free Derivation}
From the solution, absolute scales $\tau_0,\ \ell_0$ and constants $c,\ \hbar,\ \lambda_{\mathrm{rec}},\ G$ are derived purely from internal identities and policies, without external standards or tuned parameters. Confidence intervals are obtained from band widths and solver sensitivities.

\subsection*{Compliance and Logging}
The selector emits signed, hash-addressed logs including identity definitions, residuals, solver diagnostics, tolerance settings, solution vectors, and emitted outputs. A compliance API exposes these artifacts to downstream consumers and auditors; optional registry publication enables transparency.

\section*{Technical Effects}
The selector yields: (i) anchor-free, reproducible absolute calibration across devices and sites; (ii) deterministic verdicts via uniqueness and band criteria; (iii) robustness by overdetermined identity sets; and (iv) audit-ready evidence for conformance and enforcement.

\section*{Enablement and Best Mode}
Reference implementations include identity libraries (K-identities, Planck gate), solvers with band and uniqueness checks, sensitivity calculators, and signed logging/registry support. Benchmarks report solution rates, residual distributions, and robustness to identity noise.

\section*{Alternative Embodiments}
Identities may be domain-specific and extended. Solvers may employ different numerical methods provided simultaneous enforcement and band constraints are preserved. Outputs may include additional derived quantities as applicable.

\section*{Legal Notice}
All embodiments, features, and steps described herein may be combined in any operable permutation. Where means-plus-function is implicated, claims are intended to cover corresponding structures and their equivalents.

\end{document}


