\documentclass[12pt, reqno]{amsart}

%% PACKAGES
\usepackage{amsmath, amssymb, amsthm, amsfonts}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{geometry}
\usepackage{url}

%% GEOMETRY
\geometry{margin=1.in}

\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\setcounter{tocdepth}{2}

%% THEOREMS
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

%% NUMBERING
\numberwithin{equation}{section}

%% MACROS
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\T}{\mathbb{T}}
\newcommand{\Sbb}{\mathbb{S}}

\newcommand{\dv}{\mathrm{div}}
\newcommand{\curl}{\mathrm{curl}}
\newcommand{\supp}{\mathrm{supp}}
\newcommand{\osc}{\mathrm{osc}}
\newcommand{\BMO}{\mathrm{BMO}}
\newcommand{\VMO}{\mathrm{VMO}}

\newcommand{\eps}{\varepsilon}
\newcommand{\om}{\omega}
\newcommand{\Om}{\Omega}
\newcommand{\xihat}{\hat{\xi}}
\newcommand{\lambdar}{\Lambda_r}
\usepackage{xcolor}

%% TITLE & AUTHOR
%\title[Global Regularity for Navier--Stokes]{Global Regularity for the 3D Incompressible Navier--Stokes Equations via Geometric Depletion}
\title[Global Regularity for Navier--Stokes]{Global Regularity for the 3D Incompressible Navier--Stokes Equations}

\author{Jonathan Washburn}
\address{Department of Mathematics} 
\email{jonathan.washburn@example.com} % Placeholder email

%\date{\today}

%% ABSTRACT
\begin{document}

\begin{abstract}
\noindent\textbf{Status (under audit).} This manuscript records a running-max blow-up framework and a proposed route to global regularity by reducing a hypothetical singularity to the existence of a nontrivial bounded-vorticity ancient element together with a finite list of \emph{global closure gates} (compactness, far-field/tail control, and Liouville-type rigidity).

Several local/near-field reductions are proved in a referee-checkable way, but key global steps remain open/conditional at the time of writing. The current closure checklist is maintained in \texttt{ASSUMPTIONS\_TO\_CLOSE.md}.

\medskip
\noindent\textbf{Concrete computation.} We do include an explicit sign computation for the self-stretching functional $I[\omega]=\int_{\R^3}(\omega\cdot\nabla u)\cdot\omega$ for a canonical axisymmetric $\ell=2$ toroidal profile; see \texttt{WORK\_L2\_SELF\_STRETCHING.tex}. This does \emph{not} by itself close the RM2 compactness gate, but it fixes the sign in an instructive model case.
\end{abstract}

\maketitle

\tableofcontents

\section{Introduction}

\subsection{Motivation} The question of global regularity for the 3D incompressible Navier–Stokes equations remains one of the central open problems in mathematical fluid dynamics. Understanding whether finite–time singularities may arise from smooth initial data is crucial both for the analytical structure of the equations and for the predictive reliability of the physical models they describe. The system governs the motion of a viscous, incompressible fluid with constant density and follows from the conservation of linear momentum and mass. The foundational mathematical theory was established by J. Leray~\cite{Leray1934} and E. Hopf~\cite{Hopf1951}, who introduced the notion of weak solutions and established global existence via the fundamental energy inequality. However, the questions of spatial regularity and uniqueness for such weak solutions remain unresolved.

The incompressible Navier--Stokes equations arise from the fundamental principles of 
mass and momentum conservation applied to a viscous fluid treated as a continuum. 
Under the continuum hypothesis, the velocity $u(t,x)$ and pressure $p(t,x)$ are 
well-defined, smoothly varying fields describing, respectively, the instantaneous 
velocity of a fluid parcel and the normal force exerted by the surrounding fluid. 
The condition $\nabla \cdot u = 0$ reflects conservation of mass for a homogeneous, 
incompressible fluid, while the momentum equation expresses Newton’s second law, i.e.
the material acceleration $\frac{D u}{Dt} = \partial_t u + (u \cdot \nabla)u$ is 
balanced by the pressure gradient $-\nabla p$, the viscous diffusion term $\nu \Delta u$ 
arising from internal friction in a Newtonian fluid, and possible external forces $f$. 

In 3D, 
taking the curl of the momentum equation yields the vorticity formulation, in which 
the term $(\omega \cdot \nabla)u$ (with $\omega=\nabla\times u$) describes vortex 
stretching, a mechanism which does not exist in two dimensions and widely regarded as the key 
process responsible for vorticity amplification, energy cascade to smaller scales, 
and the potential formation of singularities. This vortex-stretching mechanism 
encapsulates the central mathematical difficulty of the Navier--Stokes problem, 
at the same time, the essential physical ingredient underlying the onset of 
turbulence in real viscous flows ~\cite{ConstantinFefferman1993,MajdaBertozzi2002}.

\subsection{The Navier--Stokes Regularity Problem}

Let $T>0$ be an arbitrary finite number representing the time, and $\nu>0$ a positive number representing the kinematic viscosity.  We consider 3D incompressible Navier--Stokes (N-S) equations given by the following system of PDEs:
\begin{equation}\label{eq:NS_domain}
\begin{cases}
\partial_t u + (u \cdot \nabla)u + \nabla p - \nu \Delta u = f,  \\
\nabla \cdot u = 0,
\end{cases}
\end{equation}
where the vector field $u: \R^3 \times [0,T) \to \R^3$ denotes the velocity, and 
$p: \R^3 \times [0,T) \to \R$ denotes the scalar pressure.  

We assume that the external force $f = 0$, but all results can be easily extended to the case of a non-vanishing external force by incorporating $f$ through the Duhamel integral \cite[Proposition~6.1]{Lemarie2016}, under the standard admissibility assumptions on $f$ (e.g. $f \in L^1_{loc}([0,T);L^2(\mathbb{R}^3))$).

We assume the initial data $u(x,0)=u_0(x) \in H^1(\R^3)$ is smooth and divergence-free.
Given such smooth initial data, the fundamental question, identified as one of the Millennium Prize Problems \cite{Fefferman2006}, is whether such solutions remain smooth for all time $T > 0$, or whether a finite-time singularity can form.

The modern theory of weak solutions to the N--S equations originates 
from the works of J.~Leray~\cite{Leray1934} and E.~Hopf~\cite{Hopf1951}. 
They introduced the notion of what is now called a Leray--Hopf weak solution and 
proved the global-in-time existence of such solutions for any divergence-free 
initial data $u_0 \in L^2(\mathbb{R}^3)$. These solutions satisfy the N--S 
equations in the distributional sense together with the fundamental global energy 
inequality
\begin{equation}\label{eq:energy}
\frac{1}{2} \int_{\mathbb{R}^3} |u(x,t)|^2 \, dx
+ \nu \int_0^t \int_{\mathbb{R}^3} |\nabla u(x,s)|^2 \, dx \, ds
\le \frac{1}{2} \int_{\mathbb{R}^3} |u_0(x)|^2 \, dx
\qquad \forall\, t \ge 0.
\end{equation}
Although global existence is guaranteed, the questions of uniqueness and 
spatial--temporal regularity of Leray--Hopf weak solutions remain open. 
This difficulty is tied to the \emph{supercritical} nature of the nonlinearity 
$(u\cdot\nabla)u$ with respect to the natural dissipation $\nu\Delta u$ under the 
N--S scaling, and motivates the development of refined regularity 
criteria and the introduction of the stronger class of suitable weak solutions.

The N--S equations (\ref{eq:NS_domain}) are invariant under the scaling
\begin{equation}\label{scaling}
    u_\lambda(x,t) = \lambda\, u(\lambda x, \lambda^2 t), 
\qquad
p_\lambda(x,t) = \lambda^2\, p(\lambda x, \lambda^2 t),
\end{equation}
but this transformation maps the energy norm 
$\|u\|_{L^\infty_t L^2_x}$ to $\lambda^{-1/2}\|u\|_{L^\infty_t L^2_x}$, 
making the energy strictly supercritical (too weak to control the nonlinearity).



The underlying physical space is tacitly assumed to be flat, which is the natural assumption for the study of the flow in our
3D Euclidean space. 

M. Kobayashi \cite{Kobayashi} extends the Navier–Stokes equations from flat spaces to manifolds by analyzing the motion of a Newtonian fluid on flow leaves, that is, smooth surfaces in Euclidean three-space that are invariant under the fluid flow. The proposed general equations describing the motion of a Newtonian fluid 
with constant properties on a volume Riemannian manifold $(M,g,\omega)$ are:

\begin{enumerate}
    \item Continuity equation:
    \[
        \operatorname{div}_{\omega} u = 0,
    \]

    \item {N--S equation:}
    \[
        \frac{\partial u}{\partial t} 
        + \nabla_u u
        = -\frac{1}{\rho}\,\operatorname{grad} p
        - \nu\left( \nabla^\ast \nabla u + \mathrm{Ric}(u)
        - \mathcal{L}_{\operatorname{grad}\log \omega} u \right)
        + b.
    \]
\end{enumerate}
Here $\operatorname{div}_{\omega}$ and $\operatorname{grad}$ denote divergence and gradient 
taken with respect to the volume form $\omega$, 
$\nabla^\ast \nabla$ is the Hodge--de\,Rham Laplacian on vector fields, 
$\mathrm{Ric}(u)$ is the Ricci curvature acting on $u$, 
and $\mathcal{L}_{\operatorname{grad}\!\log \omega}\,u$ represents the non--Riemannian 
correction arising from the volume form. It is shown how quantities intrinsic to the manifold, such as curvature and the choice of volume form, fundamentally modify the structure of the equations and the resulting flow behavior.


\subsection{Historical Context and Barriers}
Substantial progress has been made in understanding the partial regularity of suitable weak solutions. Scheffer \cite{Scheffer1977} and Caffarelli, Kohn, and Nirenberg \cite{CKN1982} proved that the singular set of any suitable weak solution has one-dimensional parabolic Hausdorff measure zero. Lin \cite{Lin1998} simplified and refined these results. These partial regularity theorems rely on $\varepsilon$-regularity criteria: if scale-invariant quantities (such as $\|u\|_{L^3}$ or $\|u\|_{L^\infty_t L^{3,\infty}_x}$) are locally small, the solution is regular.

Complementing the partial regularity theory are blow-up criteria. The celebrated Beale--Kato--Majda (BKM) criterion \cite{BKM1984} states that a smooth solution blows up at time $T^*$ if and only if
\begin{equation}\label{eq:BKM}
\int_0^{T^*} \|\omega(\cdot,t)\|_{L^\infty} \, dt = \infty,
\end{equation}
where $\omega = \curl \, u$ is the vorticity. Serrin \cite{Serrin1962} and Prodi \cite{Prodi1959} established that if $u \in L^q(0,T; L^p(\R^3))$ with $2/q + 3/p \le 1$ ($p > 3$), then the solution is regular. The endpoint case $L^\infty_t L^3_x$ was resolved by Escauriaza, Seregin, and \v{S}ver\'ak \cite{ESS2003}.

Despite these advances, the "scaling gap" remains. All known regularity criteria require bounds at the critical scaling level (e.g., $L^3$ velocity or $L^{3/2}$ vorticity), whereas the a priori energy bounds control only subcritical quantities (e.g., $L^2$ velocity). Bridging this gap requires exploiting the structure of the nonlinearity beyond simple scaling arguments.

\subsection{Main Result (conditional roadmap)}
We provide a geometric decomposition of the nonlinear structure that separates near-field effects from global/tail obstructions, and we record a contradiction scheme based on the running-max ancient element.

\begin{theorem}[Main theorem (conditional on closure gates)]\label{thm:main}
Let $u_0 \in H^1(\R^3)$ be smooth and divergence-free, and let $u$ be the corresponding smooth solution of \eqref{eq:NS_domain} on its maximal interval of existence $[0,T^*)$.
Assume that whenever $T^*<\infty$, the associated running-max ancient element satisfies the global closure conditions tracked in \texttt{ASSUMPTIONS\_TO\_CLOSE.md} (in particular the RM2 compactness gate and the C2/D/E rigidity reductions).
Then $T^*=\infty$ (no finite-time blow-up).
\end{theorem}

\begin{remark}[Proof Structure]
The contradiction scheme relies on extracting a running-max/vorticity-normalized ancient element and then ruling out its existence via a chain of rigidity mechanisms.
At present, the \emph{global} parts of that chain are not all closed; the honest status of each step is tracked in \texttt{ASSUMPTIONS\_TO\_CLOSE.md}.
\begin{enumerate}
\item \textbf{Global Directional Locking (Theorem~\ref{thm:global-directional-locking}):} The ancient direction field $\xi$ must be globally constant (proved via Bernstein estimates on the angular deviation).
\item \textbf{Magnitude Isotropization (Corollary~\ref{cor:magnitude-symmetry}):} The ancient vorticity magnitude becomes radial at infinity (proved via Pressure Coercivity for the deviatoric strain).
\item \textbf{RM2 compactness (OPEN):} controlling the affine/harmonic-mode obstruction (equivalently the log-critical $\ell=2$ tail moment) remains a global gate in the current draft.
\end{enumerate}
\end{remark}

\subsection{Foundations of the Proof}\label{subsec:proof-foundations}

\subsection{Structure of the Proof}

The proof relies on the following key ingredients, which are established as theorems in the subsequent sections:

\begin{enumerate}
\item \textbf{Scale-critical vorticity control (B):} Automatic under running-max normalization (Lemma~\ref{lem:omega32-runningmax-automatic}).
\item \textbf{Global Directional Locking (C):} The ancient direction field becomes globally constant via Bernstein estimates on the angular deviation (Theorem~\ref{thm:global-directional-locking}).
\item \textbf{Magnitude Isotropization (D):} The ancient vorticity magnitude becomes radial at infinity via Pressure Coercivity (Corollary~\ref{cor:magnitude-symmetry}).
\item \textbf{RM2 compactness (OPEN):} closing the fixed-frame compactness step requires additional global control (see \texttt{ASSUMPTIONS\_TO\_CLOSE.md}).
\end{enumerate}

\medskip
\noindent
In this rewrite, the contradiction object is the running-max/vorticity-normalized ancient element extracted from the blow-up sequence (Lemma~\ref{lem:ancient-limit-runningmax}). Under this normalization, the scale-critical vorticity control (B) holds automatically and is recorded below as Lemma~\ref{lem:omega32-runningmax-automatic}.

\begin{proposition}[Explicit positive self-stretching for a concrete $\ell=2$ toroidal profile]\label{prop:l2-selfstretch-example}
Let $f(r)=\mathbf 1_{[1,2]}(r)$ and define the axisymmetric $\ell=2$ toroidal vorticity profile
\[
\omega(r,\theta)=3\,f(r)\,\sin(2\theta)\,\hat\phi.
\]
Let $u=\curl(-\Delta)^{-1}\omega$ be the associated Biot--Savart velocity. Then the self-stretching functional
\[
I[\omega]:=\int_{\R^3}(\omega\cdot\nabla u)\cdot\omega\,dx
\]
is strictly positive and admits the exact value
\[
I[\omega]=\frac{64\pi}{875}\,\bigl(104-105\log 2\bigr)\;>\;0.
\]
\end{proposition}

\begin{proof}
See \texttt{WORK\_L2\_SELF\_STRETCHING.tex}, which derives an exact reduction formula for $I[\omega]$ for the ansatz $\omega_\phi(r,\theta)=3f(r)\sin(2\theta)$ and evaluates it explicitly for $f=\mathbf 1_{[1,2]}$.
\end{proof}

\begin{remark}[Open target: from self-stretching sign to RM2 closure]\label{rem:l2-instability-open}
Proposition~\ref{prop:l2-selfstretch-example} fixes the sign of the enstrophy-production term $I[\omega]$ for one canonical axisymmetric $\ell=2$ toroidal profile.
To convert such sign information into the RM2 compactness bound needed in the running-max extraction, one would need additional arguments that (i) identify which $\ell=2$ components actually generate the affine/harmonic obstruction, and (ii) turn pointwise-in-time sign information into a \emph{uniform} backward-time decay / integrability statement for the relevant tail moment of a bounded ancient solution.
These steps are not yet proved in the present draft and are tracked as OPEN items in \texttt{ASSUMPTIONS\_TO\_CLOSE.md}.
\end{remark}

\begin{lemma}[Quadratic-form structure of the $\ell=2$ transverse coefficient]\label{lem:Ab-quadratic-form}
Fix $(r,t)$ and write $f(\theta):=\omega^\infty(r\theta,t)$ for $\theta\in\Sbb^2$.
Define the symmetric trace-free matrix $Q(r,t)\in\R^{3\times 3}_{\mathrm{sym},0}$ by
\begin{equation}\label{eq:Q-def}
Q_{ij}(r,t):=\frac12\int_{\Sbb^2}\Bigl(\theta_i\,(f(\theta)\times\theta)_j+\theta_j\,(f(\theta)\times\theta)_i\Bigr)\,d\theta.
\end{equation}
Then for every $b\in\Sbb^2$,
\begin{equation}\label{eq:Ab-as-quadratic-form}
A_b^\infty(r,t)=b\cdot Q(r,t)\,b.
\end{equation}
In particular, $\sup_{b\in\Sbb^2}|A_b^\infty(r,t)|=\|Q(r,t)\|_{\mathrm{op}}\le \|Q(r,t)\|_{\mathrm{F}}$.
\end{lemma}

\begin{proof}
Using $\Phi_b(\theta)=(b\cdot\theta)(\theta\times b)$ and the vector identity
$f\cdot(\theta\times b)=(f\times\theta)\cdot b$, we compute
\[
A_b^\infty(r,t)
=\int_{\Sbb^2}(b\cdot\theta)\,f(\theta)\cdot(\theta\times b)\,d\theta
=\int_{\Sbb^2}(b\cdot\theta)\,(f(\theta)\times\theta)\cdot b\,d\theta.
\]
Writing $b\cdot\theta=b_i\theta_i$ and $(f\times\theta)\cdot b=b_j(f\times\theta)_j$ (summation convention),
we obtain
\[
A_b^\infty(r,t)=b_i b_j\int_{\Sbb^2}\theta_i\,(f(\theta)\times\theta)_j\,d\theta.
\]
Since $b_ib_j$ is symmetric in $(i,j)$, only the symmetric part of the integral kernel contributes, which yields
\eqref{eq:Ab-as-quadratic-form} with $Q$ defined by \eqref{eq:Q-def}.
Moreover, $\operatorname{tr}Q=\int_{\Sbb^2}\theta\cdot(f\times\theta)\,d\theta=0$, so $Q$ is trace-free.
The operator-norm identity $\sup_{|b|=1}|b\cdot Qb|=\|Q\|_{\mathrm{op}}$ is standard for symmetric matrices.
\end{proof}

\begin{lemma}[Toroidal $\ell=2$ structure and normalization of $\Phi_b$]\label{lem:Phib-toroidal}
Fix $b\in\Sbb^2$ and define $\Phi_b(\theta):=(b\cdot\theta)(\theta\times b)$ for $\theta\in\Sbb^2$.
Let $Y_b(\theta):=(b\cdot\theta)^2$ and denote by $\nabla_{\!S}$ the surface gradient on $\Sbb^2$.
Then
\begin{equation}\label{eq:Phi-as-curlS}
\theta\times \nabla_{\!S}Y_b(\theta)=2\,\Phi_b(\theta).
\end{equation}
In particular, $\Phi_b$ is tangential and surface-divergence free (a toroidal $\ell=2$ vector spherical harmonic), and its $L^2(\Sbb^2)$ norm is independent of $b$:
\begin{equation}\label{eq:Phi-L2}
\|\Phi_b\|_{L^2(\Sbb^2)}^2=\frac{8\pi}{15}.
\end{equation}
Moreover, $\Phi_b$ is an eigenfield of the componentwise Laplace--Beltrami operator on $\Sbb^2$:
\begin{equation}\label{eq:Phi-eigen}
\Delta_{\!S}\Phi_b=-6\,\Phi_b.
\end{equation}
\end{lemma}

\begin{proof}
Writing $Y_b(\theta)=(b\cdot\theta)^2$, a direct differentiation on $\Sbb^2$ gives
\[
\nabla_{\!S}Y_b(\theta)=2(b\cdot\theta)\bigl(b-(b\cdot\theta)\theta\bigr).
\]
Crossing with $\theta$ yields \eqref{eq:Phi-as-curlS} since $\theta\times\theta=0$.

For \eqref{eq:Phi-L2}, by rotational invariance we may take $b=e_3$.
Writing $\theta=(\sin\vartheta\cos\varphi,\sin\vartheta\sin\varphi,\cos\vartheta)$, we have
$\Phi_{e_3}(\theta)=\cos\vartheta\,(\theta\times e_3)$ and $|\theta\times e_3|=\sin\vartheta$,
so $|\Phi_{e_3}(\theta)|^2=\cos^2\vartheta\,\sin^2\vartheta$.
Therefore
\[
\|\Phi_{e_3}\|_{L^2(\Sbb^2)}^2
=\int_0^{2\pi}\int_0^\pi \cos^2\vartheta\,\sin^2\vartheta\;\sin\vartheta\,d\vartheta\,d\varphi
=2\pi\int_0^\pi \cos^2\vartheta\,\sin^3\vartheta\,d\vartheta
=\frac{8\pi}{15},
\]
as claimed.

For \eqref{eq:Phi-eigen}, again reduce by rotational invariance to $b=e_3$.
Then $\Phi_{e_3}(\theta)=(\theta_3\theta_2,-\theta_3\theta_1,0)$, and each nonzero component is the restriction to $\Sbb^2$ of a harmonic homogeneous polynomial of degree $2$ (namely $yz$ and $-xz$).
It is classical that the restriction of any harmonic homogeneous polynomial of degree $\ell$ to $\Sbb^2$ is a spherical harmonic satisfying $\Delta_{\!S} f=-\ell(\ell+1)f$.
Taking $\ell=2$ gives $\Delta_{\!S}\Phi_{e_3}=-6\Phi_{e_3}$ componentwise, hence \eqref{eq:Phi-eigen}.
\end{proof}

\begin{lemma}[Projected evolution equation for $A_b$]\label{lem:Ab-evolution}
Let $(u^\infty,p^\infty)$ be a smooth ancient Navier--Stokes solution on $\R^3\times(-\infty,0]$ with vorticity $\omega^\infty=\curl u^\infty$.
Fix $b\in\Sbb^2$ and define $\Phi_b$ and $A_b^\infty$ as in Theorem~\ref{thm:RM2U-target}.
Then $A_b^\infty$ satisfies the exact identity
\begin{equation}\label{eq:Ab-PDE}
\bigl(\partial_t-\partial_r^2-\tfrac{2}{r}\partial_r+\tfrac{6}{r^2}\bigr)A_b^\infty(r,t)
\;=\;\mathcal{F}_b(r,t),
\end{equation}
where the forcing term is the spherical projection of the vorticity transport/stretching:
\begin{equation}\label{eq:Ab-forcing}
\mathcal{F}_b(r,t):=\int_{\Sbb^2}\Bigl((\omega^\infty\!\cdot\nabla)u^\infty-(u^\infty\!\cdot\nabla)\omega^\infty\Bigr)(r\theta,t)\cdot \Phi_b(\theta)\,d\theta.
\end{equation}
\end{lemma}

\begin{proof}
Differentiate under the integral sign to obtain $\partial_t A_b^\infty=\int_{\Sbb^2}(\partial_t\omega^\infty)(r\theta,t)\cdot\Phi_b\,d\theta$.
Insert the vorticity equation $\partial_t\omega^\infty-\Delta\omega^\infty=(\omega^\infty\!\cdot\nabla)u^\infty-(u^\infty\!\cdot\nabla)\omega^\infty$.
It remains to compute the Laplacian contribution.
Using the decomposition $\Delta=\partial_r^2+\frac{2}{r}\partial_r+\frac{1}{r^2}\Delta_{\!S}$ acting componentwise on $\omega^\infty(r\theta,t)$, and that $\Phi_b$ depends only on $\theta$, we have
\[
\int_{\Sbb^2}(\partial_r^2\omega^\infty)(r\theta,t)\cdot\Phi_b\,d\theta=\partial_r^2 A_b^\infty(r,t),
\qquad
\int_{\Sbb^2}(\partial_r\omega^\infty)(r\theta,t)\cdot\Phi_b\,d\theta=\partial_r A_b^\infty(r,t).
\]
For the angular part, integration by parts on $\Sbb^2$ (componentwise) gives
\[
\int_{\Sbb^2}(\Delta_{\!S}\omega^\infty)(r\theta,t)\cdot\Phi_b\,d\theta
=\int_{\Sbb^2}\omega^\infty(r\theta,t)\cdot(\Delta_{\!S}\Phi_b)(\theta)\,d\theta
=-6\,A_b^\infty(r,t),
\]
using \eqref{eq:Phi-eigen}.
Combining these identities yields \eqref{eq:Ab-PDE}--\eqref{eq:Ab-forcing}.
\end{proof}

\begin{remark}[Energy identity behind the coercive bound]\label{rem:Ab-energy-identity}
Formally multiplying \eqref{eq:Ab-PDE} by $A_b^\infty(r,t)\,r^2$ and integrating in $r\in[1,\infty)$ yields the identity
\[
\frac12\frac{d}{dt}\int_{1}^{\infty}|A_b^\infty(r,t)|^2\,r^2\,dr
\;+\;\int_{1}^{\infty}\Bigl(|\partial_r A_b^\infty(r,t)|^2\,r^2+6|A_b^\infty(r,t)|^2\Bigr)\,dr
\;=\;\int_{1}^{\infty}\mathcal{F}_b(r,t)\,A_b^\infty(r,t)\,r^2\,dr\;+\;\mathrm{BT}_b(t),
\]
where $\mathrm{BT}_b(t)$ is the boundary term at $r=1$ coming from integration by parts.
The quadratic form on the left is exactly the coercive quantity appearing in \eqref{eq:RM2U-coercive-l2-target} (up to the harmless factor $6$ on the $L^2(dr)$ term).
Thus, proving Theorem~\ref{thm:RM2U-target} reduces to establishing a uniform-in-time control of the forcing term \eqref{eq:Ab-forcing} in a way that allows the right-hand side to be absorbed by the left.
\end{remark}

\begin{lemma}[Decay of $A_b$ from the coercive tail norm]\label{lem:Ab-tail-decay-from-coercive}
Fix $t\le 0$ and $b\in\Sbb^2$ and write $A(r):=A_b^\infty(r,t)$.
Assume
\[
\int_{1}^{\infty}|A(r)|^2\,dr<\infty
\qquad\text{and}\qquad
\int_{1}^{\infty}|A'(r)|^2\,r^2\,dr<\infty,
\]
where $A'=\partial_r A$.
Then $A(r)\to 0$ as $r\to\infty$.
\end{lemma}

\begin{proof}
For $R>S\ge 1$ we have, by the fundamental theorem of calculus and Cauchy--Schwarz,
\[
|A(R)-A(S)|
\le \int_S^R |A'(r)|\,dr
\le \left(\int_S^R |A'(r)|^2\,r^2\,dr\right)^{1/2}\left(\int_S^R \frac{dr}{r^2}\right)^{1/2}
\le S^{-1/2}\left(\int_S^\infty |A'(r)|^2\,r^2\,dr\right)^{1/2}.
\]
Since $\int_1^\infty |A'(r)|^2 r^2dr<\infty$, the right-hand side tends to $0$ as $S\to\infty$ uniformly in $R\ge S$.
Thus $A(r)$ has a finite limit $\ell$ as $r\to\infty$.
But if $\ell\neq 0$ then $|A(r)|\ge |\ell|/2$ for all $r$ sufficiently large, contradicting $\int_1^\infty |A(r)|^2dr<\infty$.
Hence $\ell=0$.
\end{proof}

\begin{corollary}[A quantitative tail decay bound for $A_b$]\label{cor:Ab-tail-decay-rate}
In the setting of Lemma~\ref{lem:Ab-tail-decay-from-coercive}, for every $r\ge 1$,
\begin{equation}\label{eq:Ab-tail-decay-rate}
r^{1/2}\,|A(r)|
\;\le\;
\left(\int_{r}^{\infty}|A'(s)|^2\,s^2\,ds\right)^{1/2}.
\end{equation}
In particular, $r^{1/2}A(r)\to 0$ as $r\to\infty$.
\end{corollary}

\begin{proof}
By Lemma~\ref{lem:Ab-tail-decay-from-coercive}, $A(s)\to 0$ as $s\to\infty$.
Hence for any $r\ge 1$,
\[
|A(r)|=\left|\int_{r}^{\infty}A'(s)\,ds\right|
\le \left(\int_{r}^{\infty}|A'(s)|^2\,s^2\,ds\right)^{1/2}\left(\int_{r}^{\infty}\frac{ds}{s^2}\right)^{1/2}
= r^{-1/2}\left(\int_{r}^{\infty}|A'(s)|^2\,s^2\,ds\right)^{1/2}.
\]
Multiplying by $r^{1/2}$ gives \eqref{eq:Ab-tail-decay-rate}.
\end{proof}

\begin{lemma}[Zero-skew along a subsequence from the coercive tail norm]\label{lem:Ab-zero-skew-subsequence}
Fix $t\le 0$ and $b\in\Sbb^2$ and write $A(r):=A_b^\infty(r,t)$ and $A'=\partial_rA$.
Assume
\[
\int_{1}^{\infty}|A(r)|^2\,dr<\infty
\qquad\text{and}\qquad
\int_{1}^{\infty}|A'(r)|^2\,r^2\,dr<\infty.
\]
Define the boundary term $B(r):=(-A(r))(r^2A'(r))$.
Then there exists a sequence $r_n\to\infty$ such that $B(r_n)\to 0$ as $n\to\infty$.
\end{lemma}

\begin{proof}
Set $E(R):=\int_R^\infty |A'(r)|^2\,r^2\,dr$, so $E(R)\downarrow 0$ as $R\to\infty$.
For each integer $n\ge 0$, apply the mean-value argument to the nonnegative integrable function $g(r):=|A'(r)|^2r^2$ on $[2^n,2^{n+1}]$ to choose
$r_n\in[2^n,2^{n+1}]$ such that
\[
|A'(r_n)|^2\,r_n^2
\le \frac{1}{2^n}\int_{2^n}^{2^{n+1}}|A'(r)|^2\,r^2\,dr
\le \frac{E(2^n)}{2^n}.
\]
Hence $r_n|A'(r_n)|\le \bigl(E(2^n)/2^n\bigr)^{1/2}$ and therefore
\[
r_n^2|A'(r_n)|
\le (2^{n+1})\cdot r_n|A'(r_n)|
\le 2^{n+1}\left(\frac{E(2^n)}{2^n}\right)^{1/2}
=2\sqrt{2^n\,E(2^n)}.
\]
On the other hand, by Corollary~\ref{cor:Ab-tail-decay-rate} and $r_n\ge 2^n$,
\[
|A(r_n)|
\le (2^n)^{-1/2}\,E(2^n)^{1/2}.
\]
Multiplying the last two bounds yields
\[
|B(r_n)|=|A(r_n)|\,r_n^2|A'(r_n)|
\le 2\,E(2^n)\ \xrightarrow[n\to\infty]{}\ 0,
\]
since $E(2^n)\to 0$.
\end{proof}

\begin{lemma}[Full zero-skew from integrable $B'$]\label{lem:Ab-zero-skew-from-Bprime}
Fix $t\le 0$ and $b\in\Sbb^2$ and write $A(r):=A_b^\infty(r,t)$ and $A'=\partial_rA$.
Assume
\[
\int_{1}^{\infty}|A(r)|^2\,dr<\infty,
\qquad
\int_{1}^{\infty}|A'(r)|^2\,r^2\,dr<\infty,
\]
and define $B(r):=(-A(r))(r^2A'(r))$.
Assume additionally that $B$ is absolutely continuous on $[1,\infty)$ and
\begin{equation}\label{eq:Bprime-L1-hyp}
\int_{1}^{\infty}|B'(r)|\,dr<\infty.
\end{equation}
Then $B(r)\to 0$ as $r\to\infty$.
\end{lemma}

\begin{proof}
By \eqref{eq:Bprime-L1-hyp}, for any $R_2>R_1\ge 1$,
\[
|B(R_2)-B(R_1)|\le \int_{R_1}^{R_2}|B'(r)|\,dr,
\]
so $\{B(R)\}_{R\ge 1}$ is a Cauchy family and therefore $B(r)$ has a finite limit $L$ as $r\to\infty$.
On the other hand, Lemma~\ref{lem:Ab-zero-skew-subsequence} yields a sequence $r_n\to\infty$ with $B(r_n)\to 0$.
Hence $L=0$.
\end{proof}

\begin{lemma}[Radial skew identity on a finite interval]\label{lem:radial-skew-ibp-finite}
Fix $t\le 0$ and $b\in\Sbb^2$ and write $A(r):=A_b^\infty(r,t)$.
Let $R>1$ and assume $A$ is $C^2$ on $[1,R]$.
Define the boundary expression
\[
B(r):=\bigl(-A(r)\bigr)\,\bigl(r^2 A'(r)\bigr),
\qquad A'=\partial_rA.
\]
Then
\begin{equation}\label{eq:radial-skew-ibp-finite}
\int_{1}^{R}\bigl(-A(r)\bigr)\,\Bigl(2r\,A'(r)+r^2A''(r)\Bigr)\,dr
\;=\;B(R)-B(1)\;+\;\int_{1}^{R}|A'(r)|^2\,r^2\,dr.
\end{equation}
\end{lemma}

\begin{proof}
Differentiate $B(r)=(-A(r))(r^2A'(r))$ to get
\[
B'(r)=-(A'(r))\,(r^2A'(r))+(-A(r))\,(2rA'(r)+r^2A''(r))
=-|A'(r)|^2r^2+(-A(r))\,(2rA'(r)+r^2A''(r)).
\]
Rearranging and integrating from $1$ to $R$ yields \eqref{eq:radial-skew-ibp-finite}.
\end{proof}

\begin{lemma}[Energy identity for $A_b$ on $[1,R]$]\label{lem:Ab-energy-identity-finiteR}
Fix $b\in\Sbb^2$ and let $A(r,t):=A_b^\infty(r,t)$ and $\mathcal F(r,t):=\mathcal F_b(r,t)$ be as in \eqref{eq:Ab-PDE}--\eqref{eq:Ab-forcing}.
Assume $A$ is smooth on $[1,R]\times[t_0,t_1]$ for some $R>1$.
Define $B(r,t):=(-A(r,t))(r^2\partial_rA(r,t))$.
Then for each $t\in[t_0,t_1]$,
\begin{equation}\label{eq:Ab-energy-finiteR}
\frac12\frac{d}{dt}\int_{1}^{R}|A(r,t)|^2\,r^2\,dr
\;+\;\int_{1}^{R}\Bigl(|\partial_r A(r,t)|^2\,r^2+6|A(r,t)|^2\Bigr)\,dr
\;=\;\int_{1}^{R}\mathcal F(r,t)\,A(r,t)\,r^2\,dr\;+\;B(1,t)-B(R,t).
\end{equation}
\end{lemma}

\begin{proof}
Multiply \eqref{eq:Ab-PDE} by $A(r,t)\,r^2$ and integrate in $r\in[1,R]$.
The time derivative gives
\[
\int_1^R (\partial_t A)A\,r^2\,dr=\frac12\frac{d}{dt}\int_1^R |A|^2 r^2\,dr.
\]
For the diffusion terms, rewrite
\[
\int_1^R\bigl(-\partial_r^2A-\tfrac{2}{r}\partial_rA\bigr)\,A\,r^2\,dr
\;=\;\int_1^R (-A)\,(r^2A''+2rA')\,dr,
\]
and apply Lemma~\ref{lem:radial-skew-ibp-finite} (with $B(r)=(-A)(r^2A')$) to obtain
\[
\int_1^R (-A)\,(r^2A''+2rA')\,dr
\;=\;B(R,t)-B(1,t)+\int_1^R |A'(r,t)|^2 r^2\,dr.
\]
Finally, the potential term contributes
\(\int_1^R \frac{6}{r^2}A\cdot A\,r^2dr = 6\int_1^R |A|^2dr\),
and the forcing contributes \(\int_1^R \mathcal F\,A\,r^2dr\).
Rearranging yields \eqref{eq:Ab-energy-finiteR}.
\end{proof}

\begin{remark}[Structure of the boundary derivative $B'$]\label{rem:Bprime-structure}
In the setting of Lemma~\ref{lem:Ab-energy-identity-finiteR}, fix a time $t$ and write $A(r):=A(r,t)$ and $\mathcal F(r):=\mathcal F(r,t)$.
Define $B(r):=(-A(r))(r^2A'(r))$ as above.
Then using \eqref{eq:Ab-PDE} one has the pointwise identity for $r>1$:
\begin{equation}\label{eq:Bprime-structure}
B'(r)
=-|A'(r)|^2\,r^2\;-\;6|A(r)|^2\;-\;r^2A(r)\,(\partial_tA)(r,t)\;+\;r^2A(r)\,\mathcal F(r).
\end{equation}
Consequently, at a fixed time $t$ the sufficient condition \(\int_1^\infty |B'(r)|dr<\infty\) from Lemma~\ref{lem:Ab-zero-skew-from-Bprime}
reduces to controlling the two ``tail interaction'' terms
\(\int_1^\infty r^2|A||\partial_tA|\,dr\) and \(\int_1^\infty r^2|A||\mathcal F|\,dr\).
\end{remark}

\begin{lemma}[Curl--poloidal--toroidal coupling on spheres]\label{lem:curl-poloidal-toroidal-coupling}
Let $F:\R^3\setminus\{0\}\to\R^3$ be $C^1$ and let $Y\in C^\infty(\Sbb^2)$.
Define the poloidal and toroidal test fields on $\Sbb^2$ by
\[
P_Y(\theta):=\nabla_{\!S}Y(\theta),
\qquad
T_Y(\theta):=\theta\times\nabla_{\!S}Y(\theta),
\]
and define the corresponding radial coefficients (for $r>0$)
\[
G_Y(r):=\int_{\Sbb^2}F(r\theta)\cdot P_Y(\theta)\,d\theta,
\qquad
H_Y(r):=\int_{\Sbb^2}(F(r\theta)\cdot\theta)\,Y(\theta)\,d\theta.
\]
Then for every $r>0$,
\begin{equation}\label{eq:curl-poloidal-toroidal-coupling}
\int_{\Sbb^2}(\curl F)(r\theta)\cdot T_Y(\theta)\,d\theta
=\frac{1}{r}\frac{d}{dr}\bigl(r\,G_Y(r)\bigr)
\;+\;\frac{1}{r}\int_{\Sbb^2}(F(r\theta)\cdot\theta)\,\Delta_{\!S}Y(\theta)\,d\theta.
\end{equation}
In particular, if $Y$ is a spherical harmonic of degree $\ell$ (so $\Delta_{\!S}Y=-\ell(\ell+1)Y$), then
\[
\int_{\Sbb^2}(\curl F)(r\theta)\cdot T_Y(\theta)\,d\theta
=\frac{1}{r}\frac{d}{dr}\bigl(r\,G_Y(r)\bigr)\;-\;\frac{\ell(\ell+1)}{r}\,H_Y(r).
\]
\end{lemma}

\begin{proof}
Write $x=r\theta$ and decompose $F$ in spherical coordinates as $F=F_r\,\theta+F_\tan$, where $F_r(x)=F(x)\cdot\theta$ and $F_\tan$ is tangential.
In the standard orthonormal tangent frame $(e_\alpha,e_\varphi)$ on $\Sbb^2$ (polar angle $\alpha$, azimuth $\varphi$), the tangential components of $\curl F$ are
\[
(\curl F)_\alpha=\frac{1}{r}\Big(\frac{1}{\sin\alpha}\,\partial_\varphi F_r-\partial_r(rF_\varphi)\Big),
\qquad
(\curl F)_\varphi=\frac{1}{r}\Big(\partial_r(rF_\alpha)-\partial_\alpha F_r\Big).
\]
Since $T_Y=\theta\times\nabla_{\!S}Y$ is tangential and satisfies
$T_{Y,\alpha}=-(\nabla_{\!S}Y)_\varphi$ and $T_{Y,\varphi}=(\nabla_{\!S}Y)_\alpha$,
we compute (using $d\theta=\sin\alpha\,d\alpha\,d\varphi$)
\begin{align*}
\int_{\Sbb^2}(\curl F)(r\theta)\cdot T_Y(\theta)\,d\theta
&=\frac{1}{r}\int_{\Sbb^2}\Big(\partial_r(rF_\tan)\cdot\nabla_{\!S}Y\Big)\,d\theta
\;-\;\frac{1}{r}\int_{\Sbb^2}\nabla_{\!S}F_r\cdot\nabla_{\!S}Y\,d\theta.
\end{align*}
The first term equals $\frac{1}{r}\frac{d}{dr}\bigl(r\int_{\Sbb^2}F_\tan(r\theta)\cdot\nabla_{\!S}Y\,d\theta\bigr)
=\frac{1}{r}\frac{d}{dr}(rG_Y(r))$, since $\nabla_{\!S}Y$ depends only on $\theta$.
For the second term, integrate by parts on $\Sbb^2$ to obtain
$-\int_{\Sbb^2}\nabla_{\!S}F_r\cdot\nabla_{\!S}Y\,d\theta=\int_{\Sbb^2}F_r\,\Delta_{\!S}Y\,d\theta$,
which yields \eqref{eq:curl-poloidal-toroidal-coupling}.
\end{proof}

\begin{remark}[Specialization to the RM2U forcing]\label{rem:curl-coupling-RM2U}
Take $F=u^\infty\times\omega^\infty$ and $Y=Y_b(\theta):=(b\cdot\theta)^2-\frac13$.
Then $T_{Y_b}= \theta\times\nabla_{\!S}Y_b=2\Phi_b$ (Lemma~\ref{lem:Phib-toroidal}) and $\Delta_{\!S}Y_b=-6Y_b$.
Thus the forcing in \eqref{eq:Ab-forcing} can be written as
\[
\mathcal F_b(r,t)
=\int_{\Sbb^2}\curl(u^\infty\times\omega^\infty)(r\theta,t)\cdot\Phi_b(\theta)\,d\theta
=\frac{1}{2r}\frac{d}{dr}\bigl(r\,G_b(r,t)\bigr)\;-\;\frac{3}{r}\,H_b(r,t),
\]
where
\[
G_b(r,t):=\int_{\Sbb^2}(u^\infty\times\omega^\infty)(r\theta,t)\cdot\nabla_{\!S}Y_b(\theta)\,d\theta,
\qquad
H_b(r,t):=\int_{\Sbb^2}(u^\infty\times\omega^\infty)(r\theta,t)\cdot\theta\;Y_b(\theta)\,d\theta.
\]
In particular, the ``single-coefficient'' reduction holds up to the explicit radial-component correction $H_b$.
\end{remark}

\begin{lemma}[Quadratic-form structure of $G_b$ and $H_b$]\label{lem:GH-quadratic-form}
Fix $(r,t)$ and write $F(\theta):=(u^\infty\times\omega^\infty)(r\theta,t)$ for $\theta\in\Sbb^2$.
Define the scalar function $s(\theta):=F(\theta)\cdot\theta$.
Define symmetric trace-free matrices $Q_G(r,t),Q_H(r,t)\in\R^{3\times 3}_{\mathrm{sym},0}$ by
\[
(Q_G)_{ij}(r,t)
:=\int_{\Sbb^2}\Bigl(\theta_i\,F_j(\theta)+\theta_j\,F_i(\theta)\Bigr)\,d\theta
\;-\;2\int_{\Sbb^2}s(\theta)\,\theta_i\theta_j\,d\theta,
\]
and
\[
(Q_H)_{ij}(r,t)
:=\int_{\Sbb^2}s(\theta)\,\Bigl(\theta_i\theta_j-\tfrac13\,\delta_{ij}\Bigr)\,d\theta.
\]
Then for every $b\in\Sbb^2$, the coefficients from Remark~\ref{rem:curl-coupling-RM2U} satisfy
\[
G_b(r,t)=b\cdot Q_G(r,t)\,b,
\qquad
H_b(r,t)=b\cdot Q_H(r,t)\,b.
\]
In particular,
\[
\sup_{b\in\Sbb^2}|G_b(r,t)|=\|Q_G(r,t)\|_{\mathrm{op}}\le \|Q_G(r,t)\|_{\mathrm{F}},
\qquad
\sup_{b\in\Sbb^2}|H_b(r,t)|=\|Q_H(r,t)\|_{\mathrm{op}}\le \|Q_H(r,t)\|_{\mathrm{F}}.
\]
\end{lemma}

\begin{proof}
Recall $Y_b(\theta)=(b\cdot\theta)^2-\frac13$.
Projecting the Euclidean gradient to the tangent space gives the explicit formula
\[
\nabla_{\!S}Y_b(\theta)=2(b\cdot\theta)\,\bigl(b-(b\cdot\theta)\theta\bigr).
\]
Writing $b\cdot\theta=b_i\theta_i$ and $F\cdot b=F_j b_j$ (summation convention), we compute
\begin{align*}
G_b(r,t)
&=\int_{\Sbb^2}F(\theta)\cdot\nabla_{\!S}Y_b(\theta)\,d\theta\\
&=2\int_{\Sbb^2}(b\cdot\theta)\,(F(\theta)\cdot b)\,d\theta
-2\int_{\Sbb^2}(b\cdot\theta)^2\,(F(\theta)\cdot\theta)\,d\theta\\
&=2\,b_ib_j\int_{\Sbb^2}\theta_i\,F_j(\theta)\,d\theta
-2\,b_ib_j\int_{\Sbb^2}s(\theta)\,\theta_i\theta_j\,d\theta.
\end{align*}
Since $b_ib_j$ is symmetric in $(i,j)$, only the symmetric part of the first integral contributes, which yields
$G_b=b\cdot Q_G b$ with $Q_G$ as defined.
Moreover,
\(\operatorname{tr}Q_G=2\int_{\Sbb^2}\theta\cdot F\,d\theta-2\int_{\Sbb^2}s(\theta)\,|\theta|^2\,d\theta=0\),
so $Q_G$ is trace-free.

Similarly,
\[
H_b(r,t)=\int_{\Sbb^2}(F(\theta)\cdot\theta)\,Y_b(\theta)\,d\theta
=b_ib_j\int_{\Sbb^2}s(\theta)\,\Bigl(\theta_i\theta_j-\tfrac13\delta_{ij}\Bigr)\,d\theta
=b\cdot Q_H b,
\]
and $\operatorname{tr}Q_H=0$ by construction.
The operator-norm identity $\sup_{|b|=1}|b\cdot Qb|=\|Q\|_{\mathrm{op}}$ is standard for symmetric matrices.
\end{proof}

\begin{remark}[Equivalent formulation of the $rG_b,rH_b$ tail gate as an $\ell=2$ Lamb-vector bound]\label{rem:rGH-as-l2-lamb}
Fix a time $t\le 0$ and define the Lamb vector (at this time)
\[
L(x):=\bigl(u_{>1}^\infty\times\omega^\infty\bigr)(x,t),\qquad x\in\R^3.
\]
For $r>0$, write $L(r\cdot)$ for the restriction of $L$ to the sphere $|x|=r$.
Decompose the scalar and tangential parts of $L(r\cdot)$ into spherical harmonics and denote by
$P_{\ell=2}^{\mathrm{rad}}$ the $L^2(\Sbb^2)$-orthogonal projection of the scalar function $\theta\mapsto L(r\theta)\cdot\theta$
onto the degree-$2$ scalar harmonics, and by $P_{\ell=2}^{\mathrm{pol}}$ the $L^2(\Sbb^2)$-orthogonal projection of the tangential field
$\theta\mapsto L(r\theta)-\bigl(L(r\theta)\cdot\theta\bigr)\theta$ onto the degree-$2$ poloidal subspace $\nabla_{\!S}\mathcal H_2$.
\smallskip

\noindent
Then there exist absolute constants $0<c\le C<\infty$ such that for every $r\ge 1$,
\[
c\,\Bigl(\|P_{\ell=2}^{\mathrm{pol}}L(r\cdot)\|_{L^2(\Sbb^2)}^2+\|P_{\ell=2}^{\mathrm{rad}}(L(r\cdot)\cdot\theta)\|_{L^2(\Sbb^2)}^2\Bigr)
\ \le\ \sup_{b\in\Sbb^2}\bigl(|G_b^{\mathrm{tail}}(r,t)|^2+|H_b^{\mathrm{tail}}(r,t)|^2\bigr)
\ \le\ C\,\Bigl(\|P_{\ell=2}^{\mathrm{pol}}L(r\cdot)\|_{L^2(\Sbb^2)}^2+\|P_{\ell=2}^{\mathrm{rad}}(L(r\cdot)\cdot\theta)\|_{L^2(\Sbb^2)}^2\Bigr),
\]
where $G_b^{\mathrm{tail}},H_b^{\mathrm{tail}}$ are as in Section~\ref{sec:RM2U} (with $u_{>1}^\infty$ in place of $u^\infty$).
Consequently, the hard wall $\textbf{U}_{rGH}^{\mathrm{tail}}$ is equivalent (up to constants) to the uniform-in-time bound
\[
\sup_{t\le 0}\int_{1}^{\infty} r^2\,\Bigl(\|P_{\ell=2}^{\mathrm{pol}}L(r\cdot)\|_{L^2(\Sbb^2)}^2+\|P_{\ell=2}^{\mathrm{rad}}(L(r\cdot)\cdot\theta)\|_{L^2(\Sbb^2)}^2\Bigr)\,dr<\infty,
\]
i.e. an exterior $L^2_x$ bound on the \emph{degree-$2$} part of the Lamb vector $u_{>1}^\infty\times\omega^\infty$.
\end{remark}

\begin{lemma}[Finite set reduction of $\sup_{b\in\Sbb^2}|b\cdot Q b|$]\label{lem:finite-b-sup}
There exist an integer $N<\infty$, unit vectors $b^{(1)},\dots,b^{(N)}\in\Sbb^2$, and an absolute constant $C<\infty$ such that for every symmetric trace-free matrix $Q\in\R^{3\times 3}_{\mathrm{sym},0}$,
\begin{equation}\label{eq:finite-b-sup}
\sup_{b\in\Sbb^2}|b\cdot Q b|^2\ \le\ C\sum_{j=1}^{N}\bigl|b^{(j)}\cdot Q\,b^{(j)}\bigr|^2.
\end{equation}
In particular, for any two such matrices $Q_1,Q_2$,
\[
\sup_{b\in\Sbb^2}\bigl(|b\cdot Q_1 b|^2+|b\cdot Q_2 b|^2\bigr)
\ \le\ C\sum_{j=1}^{N}\Bigl(|b^{(j)}\cdot Q_1 b^{(j)}|^2+|b^{(j)}\cdot Q_2 b^{(j)}|^2\Bigr).
\]
\end{lemma}

\begin{proof}
Fix $\delta\in(0,1/4]$ (say $\delta=1/4$) and choose a finite $\delta$-net $\{b^{(1)},\dots,b^{(N)}\}\subset\Sbb^2$ for the Euclidean metric on $\Sbb^2$, i.e.\ for every $b\in\Sbb^2$ there exists $j$ with $|b-b^{(j)}|\le \delta$.

\smallskip
\noindent
Let $Q\in\R^{3\times 3}_{\mathrm{sym},0}$ and let $b_*\in\Sbb^2$ be such that
$|b_*\cdot Q b_*|=\sup_{b\in\Sbb^2}|b\cdot Q b|=\|Q\|_{\mathrm{op}}$.
Pick $j$ with $|b_*-b^{(j)}|\le\delta$.
Using the identity
\[
b\cdot Q b-b'\cdot Q b'=(b+b')\cdot Q(b-b')
\]
and $|b+b'|\le 2$, we obtain
\[
\bigl|b_*\cdot Q b_*-b^{(j)}\cdot Q b^{(j)}\bigr|
\le 2\,\|Q\|_{\mathrm{op}}\,|b_*-b^{(j)}|
\le 2\delta\,\|Q\|_{\mathrm{op}}.
\]
Hence, for $\delta\le 1/4$,
\(
|b^{(j)}\cdot Q b^{(j)}|\ge (1-2\delta)\|Q\|_{\mathrm{op}}\ge \frac12\|Q\|_{\mathrm{op}}.
\)
Therefore
\[
\|Q\|_{\mathrm{op}}^2\le 4\,\bigl|b^{(j)}\cdot Q b^{(j)}\bigr|^2
\le 4\sum_{j=1}^{N}\bigl|b^{(j)}\cdot Q b^{(j)}\bigr|^2,
\]
which gives \eqref{eq:finite-b-sup} with $C=4$.
The two-matrix bound follows by applying \eqref{eq:finite-b-sup} to $Q_1$ and $Q_2$ and adding the results.
\end{proof}

\begin{lemma}[Forcing pairing in terms of $G_b,H_b$]\label{lem:forcing-pairing-GH}
Fix $t\le 0$ and $b\in\Sbb^2$.
Write $A(r):=A_b^\infty(r,t)$, $\mathcal F(r):=\mathcal F_b(r,t)$ and $G(r):=G_b(r,t)$, $H(r):=H_b(r,t)$ from Remark~\ref{rem:curl-coupling-RM2U}.
Then for every $R>1$,
\begin{equation}\label{eq:forcing-pairing-GH}
\int_{1}^{R}\mathcal F(r)\,A(r)\,r^2\,dr
=\frac12\Bigl[r\,A(r)\,(rG(r))\Bigr]_{r=1}^{r=R}
-\frac12\int_{1}^{R}\bigl(A(r)+rA'(r)\bigr)\,(rG(r))\,dr
-3\int_{1}^{R}r\,A(r)\,H(r)\,dr,
\end{equation}
where $A'=\partial_rA$.
Consequently,
\begin{equation}\label{eq:forcing-pairing-GH-bound}
\left|\int_{1}^{R}\mathcal F(r)\,A(r)\,r^2\,dr\right|
\le \frac12\Bigl|rA(r)\,(rG(r))\Bigr|_{r=1}^{r=R}
\;+\;\frac12\|A+rA'\|_{L^2(1,R)}\,\|rG\|_{L^2(1,R)}
\;+\;3\|A\|_{L^2(1,R)}\,\|rH\|_{L^2(1,R)}.
\end{equation}
\end{lemma}

\begin{proof}
From Remark~\ref{rem:curl-coupling-RM2U},
\[
r^2\mathcal F(r)\,A(r)=\frac{r}{2}\,A(r)\,\partial_r(rG(r))\;-\;3r\,A(r)\,H(r).
\]
Integrating on $[1,R]$ and integrating by parts in the first term gives
\[
\int_1^R \frac{r}{2}A\,\partial_r(rG)\,dr
=\frac12\Bigl[rA(rG)\Bigr]_{1}^{R}
-\frac12\int_1^R \partial_r(rA)\,(rG)\,dr
=\frac12\Bigl[rA(rG)\Bigr]_{1}^{R}
-\frac12\int_1^R (A+rA')\,(rG)\,dr,
\]
which yields \eqref{eq:forcing-pairing-GH}.
The bound \eqref{eq:forcing-pairing-GH-bound} follows from Cauchy--Schwarz.
\end{proof}

% NOTE (correctness): the following block is disabled because the advertised $s=r^2$ integration-by-parts
% identity is not correct as stated (a direct derivative check fails).  Use Lemma~\ref{lem:forcing-pairing-GH}.
% (Quick check: differentiate the claimed antiderivative; the $r$-weights do not match.)
\iffalse
\begin{lemma}[DEPRECATED (incorrect as stated): alternate $G_b$-based forcing pairing]\label{lem:forcing-pairing-GH-alt}
Fix a time $t\le 0$ and $b\in\Sbb^2$ and write $A,\mathcal F,G,H$ as in Lemma~\ref{lem:forcing-pairing-GH}.
Then for every $R>1$,
\begin{equation}\label{eq:forcing-pairing-GH-alt}
\int_{1}^{R}\mathcal F(r)\,A(r)\,r^2\,dr
=\frac14\Bigl[A(r)\,(rG(r))\Bigr]_{r=1}^{r=R}
-\frac14\int_{1}^{R}\bigl(rA'(r)\bigr)\,G(r)\,dr
-3\int_{1}^{R}r\,A(r)\,H(r)\,dr.
\end{equation}
Consequently,
\begin{equation}\label{eq:forcing-pairing-GH-alt-bound}
\left|\int_{1}^{R}\mathcal F(r)\,A(r)\,r^2\,dr\right|
\le \frac14\Bigl|A(r)\,(rG(r))\Bigr|_{r=1}^{r=R}
\;+\;\frac14\|rA'\|_{L^2(1,R)}\,\|G\|_{L^2(1,R)}
\;+\;3\|A\|_{L^2(1,R)}\,\|rH\|_{L^2(1,R)}.
\end{equation}
\end{lemma}

\begin{proof}
Starting again from Remark~\ref{rem:curl-coupling-RM2U},
\[
\int_1^R \mathcal F A r^2\,dr=\int_1^R \frac{r}{2}\,A\,\partial_r(rG)\,dr-3\int_1^R rAH\,dr.
\]
For the first term, change variables $s=r^2$ so that $\frac{r}{2}\,dr=\frac{1}{4}\,ds$.
Integration by parts in $s$ yields
\[
\int_1^R \frac{r}{2}\,A\,\partial_r(rG)\,dr
=\frac14\Bigl[A(r)\,(rG(r))\Bigr]_{1}^{R}
-\frac14\int_1^R \bigl(rA'(r)\bigr)\,G(r)\,dr,
\]
which gives \eqref{eq:forcing-pairing-GH-alt}.
The estimate \eqref{eq:forcing-pairing-GH-alt-bound} follows from Cauchy--Schwarz.
\end{proof}

\fi
\begin{lemma}[A Hardy gauge for the forcing derivative]\label{lem:forcing-pairing-Hardy-gauge}
Fix $t\le 0$ and $b\in\Sbb^2$ and write $A(r):=A_b^\infty(r,t)$, $\mathcal F(r):=\mathcal F_b(r,t)$ and $G(r):=G_b(r,t)$, $H(r):=H_b(r,t)$ as in Remark~\ref{rem:curl-coupling-RM2U}.
Define the primitive
\[
K(r):=\int_{1}^{r}H(s)\,ds,\qquad r\ge 1,
\]
and the modified poloidal coefficient
\[
\widetilde G(r):=G(r)-\frac{6}{r}\,K(r),\qquad r>1.
\]
Then for every $r>1$,
\begin{equation}\label{eq:forcing-Hardy-gauge-pointwise}
\mathcal F(r)=\frac{1}{2r}\frac{d}{dr}\bigl(r\,\widetilde G(r)\bigr).
\end{equation}
% (Correctness note): we keep only the pointwise identity \eqref{eq:forcing-Hardy-gauge-pointwise}
% and the Hardy estimate \eqref{eq:Hardy-gauge-L2}.  A previous draft claimed an additional
% "forcing pairing" identity by integrating in $s=r^2$, but that step depends on the disabled block above.
% The canonical forcing pairing is Lemma~\ref{lem:forcing-pairing-GH}.
\iffalse
\int_{1}^{R}\mathcal F(r)\,A(r)\,r^2\,dr
=\frac14\Bigl[A(r)\,\bigl(r\widetilde G(r)\bigr)\Bigr]_{r=1}^{r=R}
-\frac14\int_{1}^{R}\bigl(rA'(r)\bigr)\,\widetilde G(r)\,dr,
\end{equation}
where $A'=\partial_rA$.
\fi
Moreover, the Hardy inequality yields the quantitative bound
\begin{equation}\label{eq:Hardy-gauge-L2}
\|\widetilde G\|_{L^2(1,R)}\ \le\ \|G\|_{L^2(1,R)}+12\,\|H\|_{L^2(1,R)}.
\end{equation}
\end{lemma}

\begin{proof}
By Remark~\ref{rem:curl-coupling-RM2U},
\[
\mathcal F(r)=\frac{1}{2r}\frac{d}{dr}(rG(r))-\frac{3}{r}H(r).
\]
Since $r\widetilde G(r)=rG(r)-6K(r)$ and $K'(r)=H(r)$, we compute
\[
\frac{1}{2r}\frac{d}{dr}\bigl(r\widetilde G(r)\bigr)
=\frac{1}{2r}\frac{d}{dr}(rG(r))-\frac{1}{2r}\cdot 6K'(r)
=\frac{1}{2r}\frac{d}{dr}(rG(r))-\frac{3}{r}H(r)
=\mathcal F(r),
\]
which is \eqref{eq:forcing-Hardy-gauge-pointwise}.
% (The forcing-pairing step is handled separately; see Lemma~\ref{lem:forcing-pairing-GH}.)

\smallskip
\noindent
For \eqref{eq:Hardy-gauge-L2}, note that
\(\|\widetilde G\|_{L^2}\le \|G\|_{L^2}+6\|K/r\|_{L^2}\).
It remains to bound $\|K/r\|_{L^2(1,R)}$ by $\|H\|_{L^2(1,R)}$.
Since $K(1)=0$ and $K'=H$, integration by parts gives
\[
\int_{1}^{R}\frac{|K(r)|^2}{r^2}\,dr
=-\frac{|K(R)|^2}{R}+2\int_{1}^{R}\frac{K(r)\,H(r)}{r}\,dr
\le 2\int_{1}^{R}\frac{|K(r)|}{r}\,|H(r)|\,dr.
\]
Applying Cauchy--Schwarz and cancelling yields
\(\|K/r\|_{L^2(1,R)}\le 2\,\|H\|_{L^2(1,R)}\),
which implies \eqref{eq:Hardy-gauge-L2}.
\end{proof}

\begin{lemma}[Biot--Savart multipole bound for the core velocity]\label{lem:BS-multipole-core}
Let $\omega:\R^3\to\R^3$ be smooth and fix $R>0$.
Define the truncated (core) Biot--Savart velocity
\[
u_{\le R}(x):=\frac{1}{4\pi}\int_{|y|\le R}\frac{(x-y)\times \omega(y)}{|x-y|^3}\,dy,
\qquad x\in\R^3.
\]
Then for all $|x|\ge 2R$,
\begin{equation}\label{eq:BS-multipole-core}
u_{\le R}(x)=\frac{1}{4\pi}\,\frac{x\times m_R}{|x|^3}\;+\;E_R(x),
\qquad
m_R:=\int_{|y|\le R}\omega(y)\,dy,
\end{equation}
with the error bound
\begin{equation}\label{eq:BS-multipole-core-error}
|E_R(x)|\ \le\ \frac{C}{|x|^3}\int_{|y|\le R}|y|\,|\omega(y)|\,dy,
\end{equation}
for an absolute constant $C$.
In particular, if $\|\omega\|_{L^\infty}\le 1$ then for all $|x|\ge 2R$,
\[
|u_{\le R}(x)|\ \le\ C\left(\frac{R^3}{|x|^2}+\frac{R^4}{|x|^3}\right).
\]
\end{lemma}

\begin{proof}
Write the kernel $K(z):=z/|z|^3$, so that
$u_{\le R}(x)=\frac{1}{4\pi}\int_{|y|\le R}K(x-y)\times \omega(y)\,dy$.
Decompose
\[
u_{\le R}(x)-\frac{1}{4\pi}K(x)\times m_R
=\frac{1}{4\pi}\int_{|y|\le R}\bigl(K(x-y)-K(x)\bigr)\times \omega(y)\,dy.
\]
For $|x|\ge 2R$ and $|y|\le R$, the mean value theorem gives
\[
|K(x-y)-K(x)|
\le |y|\,\sup_{0\le s\le 1}|\nabla K(x-sy)|
\le \frac{C|y|}{|x|^3},
\]
since $|x-sy|\ge |x|-|y|\ge |x|/2$ and $|\nabla K(z)|\lesssim |z|^{-3}$.
Inserting this bound yields \eqref{eq:BS-multipole-core-error} and hence \eqref{eq:BS-multipole-core}.
The final inequality follows from $\|\omega\|_\infty\le 1$ and the estimates
$|m_R|\le \int_{|y|\le R}|\omega|\le C R^3$ and $\int_{|y|\le R}|y||\omega|\le C R^4$.
\end{proof}

\begin{remark}[What \eqref{eq:BS-multipole-core} buys for RM2U]\label{rem:BS-core-vs-tail-RM2U}
Applying Lemma~\ref{lem:BS-multipole-core} to $\omega^\infty(\cdot,t)$ with a fixed core radius (e.g. $R=1$) shows:
the Biot--Savart velocity generated by the \emph{core} vorticity in $\{|y|\le 1\}$ decays like $|x|^{-2}$ for $|x|\gg 1$, uniformly in $t\le 0$ under the running-max bound $\|\omega^\infty\|_{L^\infty}\le 1$.
Consequently, the contribution of the core part of $u^\infty\times\omega^\infty$ to the coefficients $G_b(r,t)$ and $H_b(r,t)$ in Remark~\ref{rem:curl-coupling-RM2U} is square-integrable in $r\in[1,\infty)$.
Thus, the only remaining obstruction in bounding $G_b,H_b$ (hence $\mathcal F_b$) is the \emph{remainder} velocity
\[
u_{>1}^\infty(\cdot,t):=u^\infty(\cdot,t)-u_{\le 1}^\infty(\cdot,t),
\]
where $u_{\le 1}^\infty$ is the truncated Biot--Savart field from Lemma~\ref{lem:BS-multipole-core}.
This $u_{>1}^\infty$ is the part of the velocity not captured by the fixed-radius core integral, and it is the precise locus of the global ``U/RM2'' tail/tightness obstruction.
If one additionally knows that $u^\infty$ obeys a global Biot--Savart representation (i.e.\ no affine/harmonic correction at infinity), then $u_{>1}^\infty$ coincides with the velocity induced by the vorticity in $\{|y|>1\}$; in general it may contain a non-decaying affine/harmonic component, which is exactly what RM2 is designed to control.
\end{remark}

\begin{lemma}[Core contribution to $rG_b$ and $rH_b$ is in $L^2(1,\infty)$]\label{lem:GH-core-L2}
Fix a time $t\le 0$ and write $\omega:=\omega^\infty(\cdot,t)$.
Assume $\|\omega\|_{L^\infty(\R^3)}\le 1$.
Let $u_{\le 1}$ be the truncated Biot--Savart velocity from Lemma~\ref{lem:BS-multipole-core} with $R=1$:
\[
u_{\le 1}(x):=\frac{1}{4\pi}\int_{|y|\le 1}\frac{(x-y)\times \omega(y)}{|x-y|^3}\,dy.
\]
For $b\in\Sbb^2$, let $Y_b(\theta):=(b\cdot\theta)^2-\frac13$ and define the corresponding \emph{core} coefficients
\[
G_b^{\mathrm{core}}(r):=\int_{\Sbb^2}(u_{\le 1}\times\omega)(r\theta)\cdot\nabla_{\!S}Y_b(\theta)\,d\theta,
\qquad
H_b^{\mathrm{core}}(r):=\int_{\Sbb^2}(u_{\le 1}\times\omega)(r\theta)\cdot\theta\;Y_b(\theta)\,d\theta.
\]
Then there exists an absolute constant $C<\infty$ such that
\begin{equation}\label{eq:GH-core-L2}
\sup_{b\in\Sbb^2}\int_{1}^{\infty}\Bigl(|r\,G_b^{\mathrm{core}}(r)|^2+|r\,H_b^{\mathrm{core}}(r)|^2\Bigr)\,dr\ \le\ C.
\end{equation}
\end{lemma}

\begin{proof}
Fix $b$.
Since $\|Y_b\|_{L^2(\Sbb^2)}$ and $\|\nabla_{\!S}Y_b\|_{L^2(\Sbb^2)}$ are uniformly bounded in $b$ (finite-dimensional $\ell=2$),
Cauchy--Schwarz on $\Sbb^2$ gives for each $r\ge 1$,
\[
|G_b^{\mathrm{core}}(r)|+|H_b^{\mathrm{core}}(r)|
\le C\,\|(u_{\le 1}\times\omega)(r\cdot)\|_{L^2(\Sbb^2)}
\le C\,\|u_{\le 1}(r\cdot)\|_{L^2(\Sbb^2)},
\]
using $|u_{\le1}\times\omega|\le |u_{\le1}|$ and $\|\omega\|_\infty\le 1$.

\smallskip
\noindent
For $r\ge 2$, Lemma~\ref{lem:BS-multipole-core} gives $|u_{\le 1}(r\theta)|\le C r^{-2}$ uniformly in $\theta\in\Sbb^2$, hence
\(\|u_{\le1}(r\cdot)\|_{L^2(\Sbb^2)}\le C r^{-2}\).
Therefore for $r\ge 2$,
\[
|r\,G_b^{\mathrm{core}}(r)|^2+|r\,H_b^{\mathrm{core}}(r)|^2
\le C\,r^2\|u_{\le1}(r\cdot)\|_{L^2(\Sbb^2)}^2
\le C\,r^2\cdot r^{-4}=C\,r^{-2},
\]
and $\int_2^\infty r^{-2}\,dr<\infty$.

\smallskip
\noindent
For $r\in[1,2]$, a direct bound from the defining integral gives $\|u_{\le1}(r\cdot)\|_{L^2(\Sbb^2)}\le C$
since $\int_{|y|\le 1}|x-y|^{-2}dy$ is uniformly bounded for $|x|\le 2$ and $\|\omega\|_\infty\le 1$.
Hence $\int_1^2 (|rG_b^{\mathrm{core}}|^2+|rH_b^{\mathrm{core}}|^2)\,dr\le C$.

\smallskip
\noindent
Combining the $[1,2]$ and $[2,\infty)$ estimates yields \eqref{eq:GH-core-L2}, and taking the supremum over $b$ completes the proof.
\end{proof}

\begin{lemma}[Core contribution to the weighted Lamb-vector energy is finite]\label{lem:core-vxw-weighted}
Fix a time $t\le 0$ and write $\omega:=\omega^\infty(\cdot,t)$.
Assume $\|\omega\|_{L^\infty(\R^3)}\le 1$ and let $u_{\le 1}$ be the truncated Biot--Savart velocity \eqref{eq:BS-multipole-core} with $R=1$.
Then there exists an absolute constant $C<\infty$ such that
\begin{equation}\label{eq:core-vxw-weighted}
\int_{|x|\ge 1}\frac{|(u_{\le 1}\times\omega)(x)|^2}{|x|^2}\,dx\ \le\ C.
\end{equation}
\end{lemma}

\begin{proof}
Since $|u_{\le 1}\times\omega|\le |u_{\le 1}|\,\|\omega\|_{L^\infty}\le |u_{\le 1}|$, it suffices to bound
\(
\int_{|x|\ge 1}|u_{\le 1}(x)|^2/|x|^2\,dx
\).
For $|x|\ge 2$, Lemma~\ref{lem:BS-multipole-core} with $R=1$ gives $|u_{\le 1}(x)|\le C|x|^{-2}$, hence
\[
\int_{|x|\ge 2}\frac{|u_{\le 1}(x)|^2}{|x|^2}\,dx
\le C\int_{2}^{\infty}\frac{r^{-4}}{r^2}\,r^2\,dr
=C\int_{2}^{\infty}r^{-4}\,dr<\infty.
\]
On the annulus $1\le |x|\le 2$, the defining integral and $\|\omega\|_\infty\le 1$ give $|u_{\le 1}(x)|\le C$, hence
\[
\int_{1\le |x|\le 2}\frac{|u_{\le 1}(x)|^2}{|x|^2}\,dx
\le C\int_{1}^{2} r^{-2}\,r^2\,dr<\infty.
\]
Combining the two regions yields \eqref{eq:core-vxw-weighted}.
\end{proof}

\begin{lemma}[$L^2_r$ control of $G_b$ and $H_b$ from a weighted exterior velocity bound]\label{lem:GH-L2-from-weighted-velocity}
Fix a time $t\le 0$ and write $u:=u^\infty(\cdot,t)$ and $\omega:=\omega^\infty(\cdot,t)$.
Assume $\|\omega\|_{L^\infty(\R^3)}\le 1$.
For each $b\in\Sbb^2$, let $Y_b(\theta):=(b\cdot\theta)^2-\frac13$ and define the coefficients $G_b(r,t)$ and $H_b(r,t)$ as in Remark~\ref{rem:curl-coupling-RM2U}.
Then there exists an absolute constant $C<\infty$ such that
\begin{equation}\label{eq:GH-L2-from-weighted-velocity}
\sup_{b\in\Sbb^2}\int_{1}^{\infty}\Bigl(|G_b(r,t)|^2+|H_b(r,t)|^2\Bigr)\,dr
\ \le\ C\int_{|x|\ge 1}\frac{|u(x)|^2}{|x|^2}\,dx.
\end{equation}
More generally, if in the definitions of $G_b,H_b$ one replaces $u^\infty$ by an arbitrary vector field $v$ (keeping the same $\omega$), then the same estimate holds with $u$ replaced by $v$ on the right-hand side.
\end{lemma}

\begin{proof}
Fix $b$ and $r\ge 1$.
Since $Y_b$ ranges over a fixed finite-dimensional ($\ell=2$) space as $b$ varies, the norms $\|Y_b\|_{L^2(\Sbb^2)}$ and $\|\nabla_{\!S}Y_b\|_{L^2(\Sbb^2)}$ are bounded uniformly in $b$.
Thus, by Cauchy--Schwarz on $\Sbb^2$ and $|v\times\omega|\le |v|\,\|\omega\|_{L^\infty}\le |v|$,
\[
|G_b(r,t)|
\le \|(v\times\omega)(r\cdot,t)\|_{L^2(\Sbb^2)}\,\|\nabla_{\!S}Y_b\|_{L^2(\Sbb^2)}
\le C\,\|v(r\cdot,t)\|_{L^2(\Sbb^2)},
\]
and similarly
\[
|H_b(r,t)|
\le \|(v\times\omega)(r\cdot,t)\|_{L^2(\Sbb^2)}\,\|Y_b\|_{L^2(\Sbb^2)}
\le C\,\|v(r\cdot,t)\|_{L^2(\Sbb^2)}.
\]
Squaring, adding, integrating in $r\in[1,\infty)$, and using
\[
\int_{1}^{\infty}\|v(r\cdot,t)\|_{L^2(\Sbb^2)}^2\,dr
=\int_{|x|\ge 1}\frac{|v(x,t)|^2}{|x|^2}\,dx,
\]
gives \eqref{eq:GH-L2-from-weighted-velocity}.  Taking $v=u$ yields the first claim.
\end{proof}

\begin{lemma}[$L^2_r$ control of $G_b$ and $H_b$ from a weighted exterior Lamb-vector bound]\label{lem:GH-L2-from-weighted-vxw}
Fix a time $t\le 0$ and write $\omega:=\omega^\infty(\cdot,t)$.
Let $v:\R^3\to\R^3$ be any vector field and define $G_b^{(v)}(r,t)$ and $H_b^{(v)}(r,t)$ by replacing $u^\infty$ with $v$ in the definitions of $G_b,H_b$ in Remark~\ref{rem:curl-coupling-RM2U} (so $(v\times\omega)$ replaces $(u^\infty\times\omega^\infty)$).
Then there exists an absolute constant $C<\infty$ such that
\begin{equation}\label{eq:GH-L2-from-weighted-vxw}
\sup_{b\in\Sbb^2}\int_{1}^{\infty}\Bigl(|G_b^{(v)}(r,t)|^2+|H_b^{(v)}(r,t)|^2\Bigr)\,dr
\ \le\ C\int_{|x|\ge 1}\frac{|(v\times\omega)(x)|^2}{|x|^2}\,dx.
\end{equation}
\end{lemma}

\begin{proof}
Fix $b$ and $r\ge 1$.
Uniform boundedness of $\|Y_b\|_{L^2(\Sbb^2)}$ and $\|\nabla_{\!S}Y_b\|_{L^2(\Sbb^2)}$ (finite-dimensional $\ell=2$) and Cauchy--Schwarz on $\Sbb^2$ yield
\[
|G_b^{(v)}(r,t)|+|H_b^{(v)}(r,t)|
\le C\,\|(v\times\omega)(r\cdot,t)\|_{L^2(\Sbb^2)}.
\]
Squaring, adding, integrating in $r\in[1,\infty)$, and using
\[
\int_{1}^{\infty}\|(v\times\omega)(r\cdot,t)\|_{L^2(\Sbb^2)}^2\,dr
=\int_{|x|\ge 1}\frac{|(v\times\omega)(x)|^2}{|x|^2}\,dx,
\]
gives \eqref{eq:GH-L2-from-weighted-vxw}.  Taking the supremum over $b$ completes the proof.
\end{proof}

\begin{lemma}[$L^2_r$ control of $rG_b$ and $rH_b$ from exterior kinetic energy]\label{lem:rGH-L2-from-kinetic}
Fix a time $t\le 0$ and write $u:=u^\infty(\cdot,t)$ and $\omega:=\omega^\infty(\cdot,t)$.
Assume $\|\omega\|_{L^\infty(\R^3)}\le 1$.
For each $b\in\Sbb^2$, let $Y_b(\theta):=(b\cdot\theta)^2-\frac13$ and define the coefficients $G_b(r,t)$ and $H_b(r,t)$ as in Remark~\ref{rem:curl-coupling-RM2U}.
Then there exists an absolute constant $C<\infty$ such that
\begin{equation}\label{eq:rGH-L2-from-kinetic}
\sup_{b\in\Sbb^2}\int_{1}^{\infty}\Bigl(|r\,G_b(r,t)|^2+|r\,H_b(r,t)|^2\Bigr)\,dr
\ \le\ C\int_{|x|\ge 1}|u(x)|^2\,dx.
\end{equation}
More generally, if in the definitions of $G_b,H_b$ one replaces $u^\infty$ by an arbitrary vector field $v$ (keeping the same $\omega$), then the same estimate holds with $u$ replaced by $v$ on the right-hand side.
\end{lemma}

\begin{proof}
Fix $b$ and $r\ge 1$.
As in Lemma~\ref{lem:GH-L2-from-weighted-velocity}, uniform boundedness of $\|Y_b\|_{L^2(\Sbb^2)}$ and $\|\nabla_{\!S}Y_b\|_{L^2(\Sbb^2)}$ together with Cauchy--Schwarz on $\Sbb^2$ and $|v\times\omega|\le |v|$ gives
\[
|G_b(r,t)|+|H_b(r,t)|\le C\,\|v(r\cdot,t)\|_{L^2(\Sbb^2)}.
\]
Multiplying by $r$, squaring, integrating in $r\in[1,\infty)$, and using
\[
\int_{1}^{\infty} r^2\,\|v(r\cdot,t)\|_{L^2(\Sbb^2)}^2\,dr
=\int_{|x|\ge 1}|v(x,t)|^2\,dx,
\]
yields \eqref{eq:rGH-L2-from-kinetic}. Taking $v=u$ gives the first claim.
\end{proof}

\begin{lemma}[$L^2_r$ control of $rG_b$ and $rH_b$ from exterior $L^2$ control of $v\times\omega$]\label{lem:rGH-L2-from-vxw}
Fix a time $t\le 0$ and write $\omega:=\omega^\infty(\cdot,t)$.
Assume $\|\omega\|_{L^\infty(\R^3)}\le 1$.
Let $v:\R^3\to\R^3$ be any vector field and define, for each $b\in\Sbb^2$, the coefficients
\[
G_b^{(v)}(r):=\int_{\Sbb^2}(v\times\omega)(r\theta)\cdot\nabla_{\!S}Y_b(\theta)\,d\theta,
\qquad
H_b^{(v)}(r):=\int_{\Sbb^2}(v\times\omega)(r\theta)\cdot\theta\;Y_b(\theta)\,d\theta,
\]
with $Y_b(\theta)=(b\cdot\theta)^2-\frac13$.
Then there exists an absolute constant $C<\infty$ such that
\begin{equation}\label{eq:rGH-L2-from-vxw}
\sup_{b\in\Sbb^2}\int_{1}^{\infty}\Bigl(|r\,G_b^{(v)}(r)|^2+|r\,H_b^{(v)}(r)|^2\Bigr)\,dr
\ \le\ C\int_{|x|\ge 1}|(v\times\omega)(x)|^2\,dx.
\end{equation}
\end{lemma}

\begin{proof}
Fix $b$ and $r\ge 1$.
Uniform boundedness of $\|Y_b\|_{L^2(\Sbb^2)}$ and $\|\nabla_{\!S}Y_b\|_{L^2(\Sbb^2)}$ (finite-dimensional $\ell=2$) and Cauchy--Schwarz on $\Sbb^2$ yield
\[
|G_b^{(v)}(r)|+|H_b^{(v)}(r)|\le C\,\|(v\times\omega)(r\cdot)\|_{L^2(\Sbb^2)}.
\]
Multiplying by $r$, squaring, integrating in $r\in[1,\infty)$, and using
\[
\int_{1}^{\infty} r^2\,\|(v\times\omega)(r\cdot)\|_{L^2(\Sbb^2)}^2\,dr
=\int_{|x|\ge 1}|(v\times\omega)(x)|^2\,dx,
\]
gives \eqref{eq:rGH-L2-from-vxw}.
\end{proof}

\begin{theorem}[Target estimate: coercive $\ell=2$ tail control]\label{thm:RM2U-target}
\textbf{[CONDITIONAL / TARGET.]}
Let $(u^\infty,p^\infty)$ be a running-max/vorticity-normalized ancient element and write $\omega^\infty=\curl u^\infty$.
For each unit vector $b\in\Sbb^2$, define the transverse $\ell=2$ test field
\[
\Phi_b(\theta):=(b\cdot \theta)\,(\theta\times b),
\qquad \theta\in\Sbb^2,
\]
and the corresponding radial coefficient
\[
A_b^\infty(r,t):=\int_{\Sbb^2}\omega^\infty(r\theta,t)\cdot \Phi_b(\theta)\,d\theta.
\]
Then there exists $K<\infty$ such that for all $t\le 0$,
\begin{equation}\label{eq:RM2U-coercive-l2-target}
\sup_{b\in\Sbb^2}\left(\int_{1}^{\infty}\bigl|(\partial_r A_b^\infty)(r,t)\bigr|^2\,r^2\,dr\ +\ \int_{1}^{\infty}\bigl|A_b^\infty(r,t)\bigr|^2\,dr\right)\ \le\ K.
\end{equation}
\end{theorem}

\begin{theorem}[New boxed target: uniform exterior weighted enstrophy]\label{thm:RM2U-weighted-enstrophy-target}
\textbf{[CONDITIONAL / TARGET.]}
Let $(u^\infty,p^\infty)$ be a running-max/vorticity-normalized ancient element and write $\omega^\infty=\curl u^\infty$.
Then there exists $M<\infty$ such that for all $t\le 0$,
\begin{equation}\label{eq:RM2U-weighted-enstrophy-target}
\int_{|x|\ge 1}\left(\frac{|\omega^\infty(x,t)|^2}{|x|^2}+|\nabla \omega^\infty(x,t)|^2\right)\,dx\ \le\ M.
\end{equation}
\end{theorem}

\begin{remark}[Why \eqref{eq:RM2U-weighted-enstrophy-target} is the right standalone target]
By Lemma~\ref{lem:RM2U-weighted-sufficient} and Lemma~\ref{lem:Phib-toroidal}, the exterior weighted enstrophy bound \eqref{eq:RM2U-weighted-enstrophy-target}
implies the $\ell=2$ coercive estimate \eqref{eq:RM2U-coercive-l2-target} (hence Theorem~\ref{thm:RM2U-target}) with $K=\frac{8\pi}{15}\,M$.
Conversely, \eqref{eq:RM2U-weighted-enstrophy-target} is \emph{not} a consequence of bounded vorticity and local suitability alone: for example a solid-body rotation has spatially constant vorticity (hence bounded), but
\(
\int_{|x|\ge 1}\frac{|\omega|^2}{|x|^2}\,dx=\infty
\)
because \(dx=r^2dr\,d\theta\) and \(r^2\cdot r^{-2}\sim 1\).
Thus some additional tail/tightness mechanism (a ``U-type'' gate in the sense of \texttt{ASSUMPTIONS\_TO\_CLOSE.md}) is logically necessary to close RM2U for the running-max ancient element; see \texttt{RM2\_ATTACK\_PLAN.md} for equivalent formulations.
\end{remark}

\begin{remark}[Adversary check: rigid rotation breaks the $rG_b,rH_b$ tail forcing gate]\label{rem:rigid-rotation-breaks-rGH}
This remark records a simple adversary showing that $L^2_r$ control of the \emph{$r$-weighted} forcing coefficients is not implied by bounded vorticity alone.
Let $\Omega\in\R^3$ be constant and consider the rigid rotation field
\[
u(x):=\tfrac12\,\Omega\times x,\qquad p(x):=-\tfrac18\,|\Omega\times x|^2,
\]
which solves stationary incompressible Navier--Stokes on $\R^3$ (with $\nu>0$) since $(u\cdot\nabla)u+\nabla p=0$ and $\Delta u=0$.
Its vorticity is constant, $\omega=\curl u=\Omega$.
Then
\[
u(x)\times\omega(x)=\tfrac12\,(\Omega\times x)\times \Omega
=\tfrac12\bigl(|\Omega|^2x-(x\cdot\Omega)\Omega\bigr),
\]
so $|u(r\theta)\times\omega(r\theta)|\sim r$ for $|\theta|=1$.
To see that the $r$-weighted coefficient cannot be square-integrable, it suffices to compute $H_b$ explicitly.
Indeed, for $\theta\in\Sbb^2$ one has
\[
(u\times\omega)(r\theta)\cdot\theta
=\tfrac12\,r\Bigl(|\Omega|^2-(\theta\cdot\Omega)^2\Bigr),
\]
so with $Y_b(\theta)=(b\cdot\theta)^2-\frac13$,
\[
H_b(r)
=\int_{\Sbb^2}(u\times\omega)(r\theta)\cdot\theta\;Y_b(\theta)\,d\theta
=\tfrac12\,r\int_{\Sbb^2}\Bigl(|\Omega|^2-(\theta\cdot\Omega)^2\Bigr)\,Y_b(\theta)\,d\theta.
\]
Since $\int_{\Sbb^2}Y_b\,d\theta=0$, this becomes
\[
H_b(r)= -\tfrac12\,r\int_{\Sbb^2}\Bigl((\theta\cdot\Omega)^2-\tfrac13|\Omega|^2\Bigr)\,Y_b(\theta)\,d\theta.
\]
If $\Omega\neq 0$ and $e:=\Omega/|\Omega|$, then $(\theta\cdot\Omega)^2-\tfrac13|\Omega|^2=|\Omega|^2\bigl((e\cdot\theta)^2-\tfrac13\bigr)=|\Omega|^2\,Y_e(\theta)$, hence
\[
H_b(r)= -\tfrac12\,r\,|\Omega|^2\int_{\Sbb^2}Y_e(\theta)\,Y_b(\theta)\,d\theta.
\]
Using the standard fourth-moment identity on $\Sbb^2$ (equivalently, orthogonality of $\ell=2$ harmonics) one finds
\[
\int_{\Sbb^2}Y_e(\theta)\,Y_b(\theta)\,d\theta=\frac{8\pi}{45}\,\bigl(3(b\cdot e)^2-1\bigr),
\]
and therefore
\[
H_b(r)= -\frac{4\pi}{45}\,r\,|\Omega|^2\,\bigl(3(b\cdot e)^2-1\bigr).
\]
In particular, choosing $b=e$ gives $H_e(r)=-(8\pi/45)\,r\,|\Omega|^2$, so $|rH_e(r)|\sim r^2$ and
\[
\int_1^\infty |rH_e(r)|^2\,dr=\infty.
\]
This illustrates that any proof of the $rG_b,rH_b\in L^2_r(1,\infty)$ tail forcing control must use a genuine tail/tightness mechanism that rules out affine/harmonic far-field modes (the RM2 obstruction).
\end{remark}

\begin{lemma}[A clean sufficient condition for Theorem~\ref{thm:RM2U-target}]\label{lem:RM2U-weighted-sufficient}
Let $t\le 0$ be fixed and let $\omega=\omega^\infty(\cdot,t)$ be smooth.
Assume there exists $M<\infty$ such that
\begin{equation}\label{eq:RM2U-weighted-sufficient-hyp}
\int_{|x|\ge 1}\left(\frac{|\omega(x)|^2}{|x|^2}+|\nabla \omega(x)|^2\right)\,dx\ \le\ M.
\end{equation}
Then for this time $t$,
\[
\sup_{b\in\Sbb^2}\left(\int_{1}^{\infty}\bigl|(\partial_r A_b^\infty)(r,t)\bigr|^2\,r^2\,dr\ +\ \int_{1}^{\infty}\bigl|A_b^\infty(r,t)\bigr|^2\,dr\right)\ \le\ C\,M,
\]
where one may take $C=\frac{8\pi}{15}$ (Lemma~\ref{lem:Phib-toroidal}).
In particular, if \eqref{eq:RM2U-weighted-sufficient-hyp} holds uniformly in $t\le 0$ then Theorem~\ref{thm:RM2U-target} holds.
\end{lemma}

\begin{proof}
Fix $b$.
By Cauchy--Schwarz on $\Sbb^2$,
\[
|A_b^\infty(r,t)|
=\left|\int_{\Sbb^2}\omega(r\theta,t)\cdot\Phi_b(\theta)\,d\theta\right|
\le \|\omega(r\cdot,t)\|_{L^2(\Sbb^2)}\,\|\Phi_b\|_{L^2(\Sbb^2)}
\]
Squaring and integrating in $r\in[1,\infty)$ gives
\[
\int_1^\infty |A_b^\infty(r,t)|^2\,dr
\le \|\Phi_b\|_{L^2(\Sbb^2)}^2\int_1^\infty \|\omega(r\cdot,t)\|_{L^2(\Sbb^2)}^2\,dr
= \|\Phi_b\|_{L^2(\Sbb^2)}^2\int_{|x|\ge 1}\frac{|\omega(x,t)|^2}{|x|^2}\,dx.
\]
Similarly,
\[
(\partial_r A_b^\infty)(r,t)=\int_{\Sbb^2}(\partial_r\omega)(r\theta,t)\cdot\Phi_b(\theta)\,d\theta,
\]
so the same Cauchy--Schwarz bound yields
\[
\int_1^\infty |(\partial_r A_b^\infty)(r,t)|^2\,r^2\,dr
\le \|\Phi_b\|_{L^2(\Sbb^2)}^2\int_1^\infty \|(\partial_r\omega)(r\cdot,t)\|_{L^2(\Sbb^2)}^2\,r^2\,dr
= \|\Phi_b\|_{L^2(\Sbb^2)}^2\int_{|x|\ge 1}|\partial_r\omega(x,t)|^2\,dx
\le \|\Phi_b\|_{L^2(\Sbb^2)}^2\int_{|x|\ge 1}|\nabla\omega(x,t)|^2\,dx.
\]
Adding the two bounds and using $\|\Phi_b\|_{L^2(\Sbb^2)}^2=\frac{8\pi}{15}$ (Lemma~\ref{lem:Phib-toroidal}) yields the estimate with $C=\frac{8\pi}{15}$; taking the supremum over $b$ completes the proof.
\end{proof}

\begin{theorem}[Closure from coercive $\ell=2$ control]\label{thm:RM2U-closure-from-coercive}
Assume \eqref{eq:RM2U-coercive-l2-target}.
Then the following hold uniformly for all $t\le 0$:
\begin{enumerate}
\item[(i)] For every $b\in\Sbb^2$ the log-critical shell moment
\[
\Sigma_b^{1,\infty}(t):=\int_{1}^{\infty}\frac{A_b^\infty(r,t)}{r}\,dr
\]
converges absolutely and obeys
\[
\sup_{b\in\Sbb^2}\bigl|\Sigma_b^{1,\infty}(t)\bigr|\ \le\ K^{1/2}.
\]
Moreover, for every $R\ge 1$,
\[
\sup_{b\in\Sbb^2}\left|\int_R^\infty \frac{A_b^\infty(r,t)}{r}\,dr\right|
\le K^{1/2}\,R^{-1/2}.
\]
\item[(ii)] The $\ell=2$ tail strain moment $S(0,t)$ from Lemma~\ref{lem:tail-strain-formula} is uniformly bounded in $t\le 0$.
In particular the fixed-frame compactness gate RM2 (Step~2 in Lemma~\ref{lem:ancient-limit-runningmax}) closes.
\item[(iii)] Any far-field Biot--Savart tail term whose $\ell=2$ contribution reduces to an $L^2(1,\infty)$-kernel pairing with $A_b^\infty(\cdot,t)$ or $r(\partial_r A_b^\infty)(\cdot,t)$ is uniformly controlled, with quantitative decay in the truncation parameter.
\end{enumerate}
\end{theorem}

\begin{proof}
(i) Fix $b$ and apply Cauchy--Schwarz:
\[
\int_{1}^{\infty}\frac{|A_b^\infty(r,t)|}{r}\,dr
\le \left(\int_{1}^{\infty}|A_b^\infty(r,t)|^2\,dr\right)^{1/2}
\left(\int_{1}^{\infty}\frac{dr}{r^2}\right)^{1/2}
\le K^{1/2},
\]
and the same argument on $[R,\infty)$ gives the $R^{-1/2}$ tail bound.

\smallskip
\noindent
(ii) By Lemma~\ref{lem:tail-strain-formula}, for every $b\in\Sbb^2$,
\[
b\cdot S(0,t)\,b
=-\frac{3}{4\pi}\int_{1}^{\infty}\frac{dr}{r}\int_{\Sbb^2}(b\cdot\theta)\,\bigl((b\times\theta)\cdot \omega^\infty(r\theta,t)\bigr)\,d\theta.
\]
Since $(b\times\theta)\cdot \omega = \omega\cdot(b\times\theta)=-\omega\cdot(\theta\times b)$, the inner integral equals $-\!A_b^\infty(r,t)$.
Therefore
\[
b\cdot S(0,t)\,b=\frac{3}{4\pi}\,\Sigma_b^{1,\infty}(t).
\]
Taking the supremum over $b$ and using (i) gives
\(
\|S(0,t)\|_{\mathrm{op}}=\sup_{|b|=1}|b\cdot S(0,t)\,b|
\le \frac{3}{4\pi}K^{1/2}.
\)
Hence $|S(0,t)|$ is uniformly bounded (in any matrix norm), which is exactly the RM2 bound in Corollary~\ref{cor:RM2-equivalence}.

\smallskip
\noindent
(iii) This is the same $L^2$-duality estimate as in (i), applied to the relevant kernel and to $A_b^\infty$ or $r(\partial_r A_b^\infty)$, using \eqref{eq:RM2U-coercive-l2-target}.
\end{proof}

\begin{remark}[Status and analytic route for Theorem~\ref{thm:RM2U-target}]
Theorem~\ref{thm:RM2U-target} is the primary remaining global closure gate in the present draft.
One clean analytic route is the endpoint maximal-regularity/time-regularity upgrade discussed in Remark~\ref{rem:RM2-l2-moment}:
prove a time-regularity bound for a suitable flux potential controlling the $\ell=2$ sector, which upgrades a borderline (BF-type) control to the pointwise-in-time coercive estimate \eqref{eq:RM2U-coercive-l2-target}.
\end{remark}

\begin{lemma}[Flux/time-derivative identity for the log-critical tail moment $\Sigma_b^{1,R}$]\label{lem:Sigma-flux-identity}
Fix $b\in\Sbb^2$ and let $A_b^\infty(r,t)$ and $\mathcal F_b(r,t)$ be as in \eqref{eq:Ab-PDE}--\eqref{eq:Ab-forcing}.
For $R>1$ define the truncated log-critical tail moment
\[
\Sigma_b^{1,R}(t):=\int_{1}^{R}\frac{A_b^\infty(r,t)}{r}\,dr.
\]
Then $\Sigma_b^{1,R}$ is differentiable and for every $t\le 0$ one has the identity
\begin{equation}\label{eq:Sigma-flux}
\frac{d}{dt}\Sigma_b^{1,R}(t)
=\left[\frac{(\partial_r A_b^\infty)(r,t)}{r}+\frac{3A_b^\infty(r,t)}{r^2}\right]_{r=1}^{r=R}
\;+\;\int_{1}^{R}\frac{\mathcal F_b(r,t)}{r}\,dr.
\end{equation}
\end{lemma}

\begin{proof}
Differentiate under the integral sign to obtain
\[
\frac{d}{dt}\Sigma_b^{1,R}(t)=\int_1^R \frac{\partial_t A_b^\infty(r,t)}{r}\,dr.
\]
Using \eqref{eq:Ab-PDE}, we write
\[
\partial_t A_b^\infty
=\partial_r^2 A_b^\infty+\frac{2}{r}\partial_r A_b^\infty-\frac{6}{r^2}A_b^\infty+\mathcal F_b.
\]
Therefore
\[
\frac{d}{dt}\Sigma_b^{1,R}(t)
=\int_1^R \frac{\partial_r^2 A_b^\infty}{r}\,dr
 +2\int_1^R \frac{\partial_r A_b^\infty}{r^2}\,dr
 -6\int_1^R \frac{A_b^\infty}{r^3}\,dr
 +\int_1^R \frac{\mathcal F_b}{r}\,dr.
\]
Integrating by parts,
\(
\int_1^R \frac{\partial_r^2 A_b^\infty}{r}\,dr=[(\partial_r A_b^\infty)/r]_1^R+\int_1^R \frac{\partial_r A_b^\infty}{r^2}\,dr
\),
so the first two terms combine to
\[
\left[\frac{\partial_r A_b^\infty}{r}\right]_{1}^{R}
 +3\int_1^R \frac{\partial_r A_b^\infty}{r^2}\,dr.
\]
Using $\partial_r(A_b^\infty/r^2)=(\partial_r A_b^\infty)/r^2-2A_b^\infty/r^3$, we further obtain
\[
\int_1^R \frac{\partial_r A_b^\infty}{r^2}\,dr
=\left[\frac{A_b^\infty}{r^2}\right]_{1}^{R}
 +2\int_1^R \frac{A_b^\infty}{r^3}\,dr,
\]
so the $A_b^\infty/r^3$ terms cancel:
\(
3\cdot 2\int_1^R A_b^\infty/r^3 - 6\int_1^R A_b^\infty/r^3 = 0
\).
Collecting boundary terms yields \eqref{eq:Sigma-flux}.
\end{proof}

\begin{lemma}[The forcing integral $\int \mathcal F_b(r,t)\,\frac{dr}{r}$ in terms of $G_b,H_b$]\label{lem:F-over-r-in-terms-of-GH}
Fix $t\le 0$ and $b\in\Sbb^2$.
Let $\mathcal F_b(r,t)$, $G_b(r,t)$ and $H_b(r,t)$ be as in Remark~\ref{rem:curl-coupling-RM2U}, so that for $r>0$,
\[
\mathcal F_b(r,t)=\frac{1}{2r}\frac{d}{dr}\bigl(r\,G_b(r,t)\bigr)-\frac{3}{r}\,H_b(r,t).
\]
Then for every $R>1$,
\begin{equation}\label{eq:F-over-r-GH}
\int_{1}^{R}\frac{\mathcal F_b(r,t)}{r}\,dr
=\left[\frac{G_b(r,t)}{2r}\right]_{r=1}^{r=R}
\;+\;\int_{1}^{R}\frac{G_b(r,t)-3H_b(r,t)}{r^2}\,dr.
\end{equation}
Consequently, there exists an absolute constant $C<\infty$ such that
\begin{equation}\label{eq:F-over-r-GH-bound}
\left|\int_{1}^{R}\frac{\mathcal F_b(r,t)}{r}\,dr\right|
\le \frac12\frac{|G_b(R,t)|}{R}\;+\;\frac12|G_b(1,t)|
\;+\;C\Bigl(\|G_b(\cdot,t)\|_{L^2(1,R)}+\|H_b(\cdot,t)\|_{L^2(1,R)}\Bigr).
\end{equation}
\end{lemma}

\begin{proof}
Starting from the identity for $\mathcal F_b$ and dividing by $r$, we have
\[
\int_1^R \frac{\mathcal F_b(r,t)}{r}\,dr
=\frac12\int_1^R \frac{1}{r^2}\frac{d}{dr}\bigl(rG_b(r,t)\bigr)\,dr
-3\int_1^R \frac{H_b(r,t)}{r^2}\,dr.
\]
Integrating by parts in the first term yields
\[
\frac12\int_1^R \frac{1}{r^2}\frac{d}{dr}\bigl(rG_b(r,t)\bigr)\,dr
=\left[\frac{rG_b(r,t)}{2r^2}\right]_{1}^{R}
\;+\;\int_1^R \frac{rG_b(r,t)}{r^3}\,dr
=\left[\frac{G_b(r,t)}{2r}\right]_{1}^{R}
\;+\;\int_1^R \frac{G_b(r,t)}{r^2}\,dr,
\]
which gives \eqref{eq:F-over-r-GH}.
For \eqref{eq:F-over-r-GH-bound}, estimate the $r^{-2}$ terms by Cauchy--Schwarz and $\int_1^\infty r^{-4}\,dr<\infty$.
\end{proof}

\begin{lemma}[Time-regularity of $\Sigma_b^{1,R}$ under $L^2_r$ control of $G_b,H_b$]\label{lem:Sigma-Lipschitz-from-GH}
Fix $b\in\Sbb^2$ and $R>1$.
Let $\Sigma_b^{1,R}(t)$ be as in Lemma~\ref{lem:Sigma-flux-identity}.
Assume that
\[
\sup_{t\le 0}\Bigl(|A_b^\infty(R,t)|+|(\partial_rA_b^\infty)(R,t)|+|G_b(R,t)|+|G_b(1,t)|\Bigr)\ <\ \infty,
\]
and that
\[
\sup_{t\le 0}\Bigl(\|G_b(\cdot,t)\|_{L^2(1,R)}+\|H_b(\cdot,t)\|_{L^2(1,R)}\Bigr)\ <\ \infty.
\]
Then $\Sigma_b^{1,R}$ is globally Lipschitz on $(-\infty,0]$:
there exists $L_R<\infty$ such that for all $t_1<t_2\le 0$,
\[
|\Sigma_b^{1,R}(t_2)-\Sigma_b^{1,R}(t_1)|\le L_R\,|t_2-t_1|.
\]
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:Sigma-flux-identity} and Lemma~\ref{lem:F-over-r-in-terms-of-GH}, for each $t\le 0$ we have
\[
\Bigl|\frac{d}{dt}\Sigma_b^{1,R}(t)\Bigr|
\le \left|\frac{(\partial_rA_b^\infty)(R,t)}{R}\right|
 +3\left|\frac{A_b^\infty(R,t)}{R^2}\right|
 +\left|\frac{(\partial_rA_b^\infty)(1,t)}{1}\right|
 +3|A_b^\infty(1,t)|
 +\left|\int_1^R \frac{\mathcal F_b(r,t)}{r}\,dr\right|.
\]
The forcing integral is bounded by \eqref{eq:F-over-r-GH-bound} in terms of $|G_b(R,t)|/R$, $|G_b(1,t)|$ and the $L^2(1,R)$ norms of $G_b,H_b$.
Collecting these bounds yields $\sup_{t\le 0}|\frac{d}{dt}\Sigma_b^{1,R}(t)|\le L_R$, and the Lipschitz bound follows by integrating in time.
\end{proof}

\begin{lemma}[Time-averaged control of $\partial_t\Sigma_b^{1,R}$ from spacetime $L^2$ bounds on $G_b,H_b$]\label{lem:Sigma-H1-from-GH-L2t}
Fix $b\in\Sbb^2$, $R>1$, and a measurable time interval $I\subset(-\infty,0]$ of finite measure.
Let $\Sigma_b^{1,R}(t)$ be as in Lemma~\ref{lem:Sigma-flux-identity}.
Assume that the boundary traces satisfy
\[
\frac{A_b^\infty(R,\cdot)}{R^2},\ \frac{(\partial_rA_b^\infty)(R,\cdot)}{R},\ \frac{G_b(R,\cdot)}{R},\ A_b^\infty(1,\cdot),\ (\partial_rA_b^\infty)(1,\cdot),\ G_b(1,\cdot)\ \in L^2(I),
\]
and that
\[
G_b,\ H_b\ \in\ L^2\!\bigl(I;L^2(1,R)\bigr).
\]
Then $\Sigma_b^{1,R}\in H^1(I)$ and one has the estimate
\begin{equation}\label{eq:Sigma-H1-bound}
\left\|\partial_t\Sigma_b^{1,R}\right\|_{L^2(I)}
\ \lesssim\
\left\|\frac{(\partial_rA_b^\infty)(R,\cdot)}{R}\right\|_{L^2(I)}
+\left\|\frac{A_b^\infty(R,\cdot)}{R^2}\right\|_{L^2(I)}
+\left\|\frac{G_b(R,\cdot)}{R}\right\|_{L^2(I)}
\end{equation}
\[
\qquad\qquad
+\left\|(\partial_rA_b^\infty)(1,\cdot)\right\|_{L^2(I)}
+\left\|A_b^\infty(1,\cdot)\right\|_{L^2(I)}
+\left\|G_b(1,\cdot)\right\|_{L^2(I)}
+\|G_b\|_{L^2(I;L^2(1,R))}
+\|H_b\|_{L^2(I;L^2(1,R))},
\]
where the implicit constant is absolute.
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:Sigma-flux-identity} and Lemma~\ref{lem:F-over-r-in-terms-of-GH}, for every $t\in I$ we have
\[
\frac{d}{dt}\Sigma_b^{1,R}(t)
=\left[\frac{(\partial_r A_b^\infty)(r,t)}{r}+\frac{3A_b^\infty(r,t)}{r^2}\right]_{r=1}^{r=R}
\;+\;\left[\frac{G_b(r,t)}{2r}\right]_{r=1}^{r=R}
\;+\;\int_{1}^{R}\frac{G_b(r,t)-3H_b(r,t)}{r^2}\,dr.
\]
Taking absolute values and using Cauchy--Schwarz in $r$ with $\int_1^\infty r^{-4}\,dr<\infty$ yields the pointwise bound
\[
\left|\frac{d}{dt}\Sigma_b^{1,R}(t)\right|
\ \lesssim\
\left|\frac{(\partial_rA_b^\infty)(R,t)}{R}\right|
+\left|\frac{A_b^\infty(R,t)}{R^2}\right|
+\left|\frac{G_b(R,t)}{R}\right|
+|(\partial_rA_b^\infty)(1,t)|
+|A_b^\infty(1,t)|
+|G_b(1,t)|
+\|G_b(\cdot,t)\|_{L^2(1,R)}
+\|H_b(\cdot,t)\|_{L^2(1,R)}.
\]
Taking $L^2(I)$ norms in $t$ and using the assumptions gives \eqref{eq:Sigma-H1-bound}.
Since $\partial_t\Sigma_b^{1,R}\in L^2(I)$, we have $\Sigma_b^{1,R}\in H^1(I)$.
\end{proof}

\begin{remark}[Near-field traces in Lemma~\ref{lem:Sigma-H1-from-GH-L2t} are automatic on bounded time intervals]\label{rem:Sigma-H1-nearfield-automatic}
If $I=[t_1,t_2]\subset(-\infty,0]$ is a bounded time interval and $(u^\infty,p^\infty)$ is smooth, then the trace functions
$A_b^\infty(1,\cdot)$, $(\partial_rA_b^\infty)(1,\cdot)$, and $G_b(1,\cdot)$
belong to $L^2(I)$ automatically (indeed they are bounded on $\{|x|=1\}\times I$ by continuity).
Thus, on bounded $I$ the only nontrivial hypotheses in Lemma~\ref{lem:Sigma-H1-from-GH-L2t} are the far-field trace controls at $r=R$
and the spacetime $L^2$ control of $G_b,H_b$ on $(1,R)\times I$.
\end{remark}

\begin{lemma}[Passing $R\to\infty$ in the $L^2_t$ $\Sigma_b$ route along good radii]\label{lem:Sigma-infty-H1-goodradii}
Fix $b\in\Sbb^2$ and a bounded time interval $I=[t_1,t_2]\subset(-\infty,0]$.
Assume that
\[
A_b^\infty\in L^2\!\bigl(I;L^2(1,\infty)\bigr),\qquad
r(\partial_rA_b^\infty)\in L^2\!\bigl(I;L^2(1,\infty)\bigr),
\]
and that
\[
G_b,\ H_b\in L^2\!\bigl(I;L^2(1,\infty)\bigr).
\]
Define $\Sigma_b^{1,R}(t)$ as in Lemma~\ref{lem:Sigma-flux-identity} and define
\[
\Sigma_b^{1,\infty}(t):=\int_{1}^{\infty}\frac{A_b^\infty(r,t)}{r}\,dr,
\]
which converges in $L^2(I)$.
Then there exists a sequence $R_n\to\infty$ such that
\begin{enumerate}
\item[(i)] $\Sigma_b^{1,R_n}\to \Sigma_b^{1,\infty}$ strongly in $L^2(I)$,
\item[(ii)] $\Sigma_b^{1,R_n}$ is bounded in $H^1(I)$, and hence $\Sigma_b^{1,\infty}\in H^1(I)$,
\item[(iii)] the far-field trace terms vanish along $R_n$ in $L^2(I)$:
\[
\left\|\frac{A_b^\infty(R_n,\cdot)}{R_n^2}\right\|_{L^2(I)}
\;+\;
\left\|\frac{(\partial_rA_b^\infty)(R_n,\cdot)}{R_n}\right\|_{L^2(I)}
\;+\;
\left\|\frac{G_b(R_n,\cdot)}{R_n}\right\|_{L^2(I)}
\ \longrightarrow\ 0.
\]
\end{enumerate}
\end{lemma}

\begin{proof}
First, $\Sigma_b^{1,R}$ is Cauchy in $L^2(I)$.
Indeed for $1<R<S$,
\[
|\Sigma_b^{1,S}(t)-\Sigma_b^{1,R}(t)|
\le \int_R^S \frac{|A_b^\infty(r,t)|}{r}\,dr
\le \left(\int_R^S |A_b^\infty(r,t)|^2\,dr\right)^{1/2}
\left(\int_R^S \frac{dr}{r^2}\right)^{1/2}
\le \left(\int_R^S |A_b^\infty(r,t)|^2\,dr\right)^{1/2},
\]
and integrating in $t$ gives
\[
\|\Sigma_b^{1,S}-\Sigma_b^{1,R}\|_{L^2(I)}^2
\le \int_I\int_R^S |A_b^\infty(r,t)|^2\,dr\,dt\ \longrightarrow\ 0
\]
as $R,S\to\infty$ by the assumption $A_b^\infty\in L^2(I;L^2(1,\infty))$.
Hence $\Sigma_b^{1,R}\to \Sigma_b^{1,\infty}$ strongly in $L^2(I)$.

\smallskip
\noindent
Next we choose good radii for the far-field trace terms.
Set
\[
f_1(r,t):=\frac{A_b^\infty(r,t)}{r^2},\qquad
f_2(r,t):=\frac{(\partial_rA_b^\infty)(r,t)}{r},\qquad
f_3(r,t):=\frac{G_b(r,t)}{r}.
\]
By the assumptions and $r\ge 1$,
\[
\int_I\int_1^\infty |f_1|^2\,dr\,dt\le \int_I\int_1^\infty |A_b^\infty|^2\,dr\,dt<\infty,
\]
\[
\int_I\int_1^\infty |f_2|^2\,dr\,dt=\int_I\int_1^\infty \frac{|(\partial_rA_b^\infty)|^2}{r^2}\,dr\,dt
\le \int_I\int_1^\infty |(\partial_rA_b^\infty)|^2\,r^2\,dr\,dt<\infty,
\]
and similarly $\int_I\int_1^\infty |f_3|^2\,dr\,dt<\infty$.
Applying Lemma~\ref{lem:good-radii-L2t-family} to the finite family $f_1,f_2,f_3$ yields radii $R_n\to\infty$ such that
\(
\|f_j(R_n,\cdot)\|_{L^2(I)}\to 0
\)
for $j=1,2,3$, which is exactly (iii).

\smallskip
\noindent
Finally, apply Lemma~\ref{lem:Sigma-H1-from-GH-L2t} with $R=R_n$.
On bounded $I$ the near-field trace terms at $r=1$ are in $L^2(I)$ by Remark~\ref{rem:Sigma-H1-nearfield-automatic}.
Moreover, the far-field trace terms at $r=R_n$ tend to $0$ in $L^2(I)$ by (iii), and $G_b,H_b\in L^2(I;L^2(1,R_n))$ since $G_b,H_b\in L^2(I;L^2(1,\infty))$.
Thus $\|\partial_t\Sigma_b^{1,R_n}\|_{L^2(I)}$ is bounded uniformly in $n$, and $\Sigma_b^{1,R_n}$ is bounded in $H^1(I)$.
By weak compactness in $H^1(I)$ and the strong $L^2(I)$ convergence in (i), the limit must be $\Sigma_b^{1,\infty}$, hence $\Sigma_b^{1,\infty}\in H^1(I)$.
\end{proof}

\begin{remark}[Minimal input for the $L^2_t$ $\Sigma_b$ route (projected spacetime wall)]\label{rem:Sigma-projected-wall}
Lemma~\ref{lem:Sigma-infty-H1-goodradii} shows that the $L^2_t$ (time-averaged) $\Sigma_b$ route does \emph{not} require any full-field weighted enstrophy estimate on $\omega^\infty$.
The genuinely minimal analytic input is the \emph{projected} spacetime $L^2$ package on $(1,\infty)\times I$:
\[
A_b^\infty,\ r(\partial_rA_b^\infty),\ G_b,\ H_b\ \in\ L^2\!\bigl(I;L^2(1,\infty)\bigr),
\]
together with smoothness to control the near-field traces at $r=1$ (Remark~\ref{rem:Sigma-H1-nearfield-automatic}).
Any stronger hypothesis (e.g.\ the spacetime-weighted condition in Corollary~\ref{cor:Sigma-infty-H1-weighted-sufficient}) is merely a convenient sufficient condition that implies this projected package.
\end{remark}

\begin{lemma}[Finite-$b$ reduction for the projected $\Sigma_b$ spacetime wall]\label{lem:Sigma-proj-finite-b}
There exist an integer $N<\infty$, unit vectors $b^{(1)},\dots,b^{(N)}\in\Sbb^2$, and an absolute constant $C<\infty$ such that the following holds.
Let $I\subset(-\infty,0]$ be a measurable time interval, and let $A_b^\infty(r,t)$, $G_b(r,t)$, and $H_b(r,t)$ be the $\ell=2$ coefficients from \eqref{eq:Ab-def} and Remark~\ref{rem:curl-coupling-RM2U}.
Then
\begin{align*}
\sup_{b\in\Sbb^2}\int_I\int_{1}^{\infty}\Bigl(
|A_b^\infty(r,t)|^2+|r(\partial_rA_b^\infty)(r,t)|^2+|G_b(r,t)|^2+|H_b(r,t)|^2
\Bigr)\,dr\,dt\\
\le\ C\sum_{j=1}^{N}\int_I\int_{1}^{\infty}\Bigl(
|A_{b^{(j)}}^\infty(r,t)|^2+|r(\partial_rA_{b^{(j)}}^\infty)(r,t)|^2+|G_{b^{(j)}}(r,t)|^2+|H_{b^{(j)}}(r,t)|^2
\Bigr)\,dr\,dt.
\end{align*}
\end{lemma}

\begin{proof}
Fix $(r,t)$.
By Lemma~\ref{lem:Ab-quadratic-form}, there exists $Q_A(r,t)\in\R^{3\times 3}_{\mathrm{sym},0}$ such that
$A_b^\infty(r,t)=b\cdot Q_A(r,t)\,b$ for all $b$.
Applying Lemma~\ref{lem:Ab-quadratic-form} with $f(\theta):=(\partial_r\omega^\infty)(r\theta,t)$ gives a second matrix
$Q_{Ar}(r,t)\in\R^{3\times 3}_{\mathrm{sym},0}$ such that
$(\partial_rA_b^\infty)(r,t)=b\cdot Q_{Ar}(r,t)\,b$ for all $b$, hence
$r(\partial_rA_b^\infty)(r,t)=b\cdot (rQ_{Ar}(r,t))\,b$.
By Lemma~\ref{lem:GH-quadratic-form}, there exist $Q_G(r,t),Q_H(r,t)\in\R^{3\times 3}_{\mathrm{sym},0}$ with
$G_b(r,t)=b\cdot Q_G(r,t)b$ and $H_b(r,t)=b\cdot Q_H(r,t)b$.

\smallskip
\noindent
Applying Lemma~\ref{lem:finite-b-sup} to each of the four trace-free matrices
$Q_A$, $rQ_{Ar}$, $Q_G$, and $Q_H$, and summing the resulting bounds, yields a pointwise inequality
\[
\sup_{b\in\Sbb^2}\Bigl(
|A_b^\infty(r,t)|^2+|r(\partial_rA_b^\infty)(r,t)|^2+|G_b(r,t)|^2+|H_b(r,t)|^2
\Bigr)
\le
C\sum_{j=1}^{N}\Bigl(
|A_{b^{(j)}}^\infty(r,t)|^2+|r(\partial_rA_{b^{(j)}}^\infty)(r,t)|^2+|G_{b^{(j)}}(r,t)|^2+|H_{b^{(j)}}(r,t)|^2
\Bigr),
\]
with the same finite set $\{b^{(j)}\}$ and constant $C$.
Integrating over $(r,t)\in(1,\infty)\times I$ and using $\sup_b\int\!\!\int \le \int\!\!\int \sup_b$ gives the stated inequality.
\end{proof}

\begin{remark}[In $\textbf{U}_{\Sigma}^{\mathrm{proj}}(I)$ the $G_b,H_b$ terms are tail-only]\label{rem:Sigma-proj-tailGH}
Write $u^\infty=u_{\le 1}^\infty+u_{>1}^\infty$ as in Remark~\ref{rem:BS-core-vs-tail-RM2U}, and for $b\in\Sbb^2$ define
\[
G_b^{\mathrm{core}}(r,t),\ H_b^{\mathrm{core}}(r,t)
\]
by replacing $u^\infty$ with $u_{\le 1}^\infty$ in the definitions of $G_b,H_b$ (Remark~\ref{rem:curl-coupling-RM2U}),
and similarly define $G_b^{\mathrm{tail}}:=G_b-G_b^{\mathrm{core}}$ and $H_b^{\mathrm{tail}}:=H_b-H_b^{\mathrm{core}}$.
Then, under the running-max bound $\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$, Lemma~\ref{lem:GH-core-L2} implies
\[
\sup_{t\le 0}\sup_{b\in\Sbb^2}\int_1^\infty \Bigl(|G_b^{\mathrm{core}}(r,t)|^2+|H_b^{\mathrm{core}}(r,t)|^2\Bigr)\,dr<\infty,
\]
and hence for every bounded time interval $I=[t_1,t_2]$,
\[
\sup_{b\in\Sbb^2}\int_I\int_1^\infty \Bigl(|G_b^{\mathrm{core}}(r,t)|^2+|H_b^{\mathrm{core}}(r,t)|^2\Bigr)\,dr\,dt<\infty.
\]
Consequently, on bounded $I$ the finiteness of the $G_b,H_b$ part of the projected spacetime wall
\[
\sup_{b}\int_I\int_1^\infty (|G_b|^2+|H_b|^2)\,dr\,dt<\infty
\]
is equivalent (up to an additive constant depending only on $|I|$) to the same statement with $(G_b,H_b)$ replaced by the tail coefficients
$(G_b^{\mathrm{tail}},H_b^{\mathrm{tail}})$.
In particular, the only nontrivial content of the $G_b,H_b$ terms in $\textbf{U}_{\Sigma}^{\mathrm{proj}}(I)$ is the tail velocity $u_{>1}^\infty$.
\end{remark}

\begin{corollary}[A clean spacetime-weighted sufficient condition for the $L^2_t$ $\Sigma_b$ route]\label{cor:Sigma-infty-H1-weighted-sufficient}
Let $(u^\infty,p^\infty)$ be a smooth running-max/vorticity-normalized ancient element and write $\omega^\infty=\curl u^\infty$.
Fix a bounded time interval $I=[t_1,t_2]\subset(-\infty,0]$.
Assume that
\begin{equation}\label{eq:Sigma-infty-H1-weighted-sufficient}
\int_I\int_{|x|\ge 1}\left(\frac{|\omega^\infty(x,t)|^2}{|x|^2}+|\nabla\omega^\infty(x,t)|^2+\frac{|(u^\infty\times\omega^\infty)(x,t)|^2}{|x|^2}\right)\,dx\,dt<\infty.
\end{equation}
Then for every $b\in\Sbb^2$ the tail moment $\Sigma_b^{1,\infty}(t)=\int_1^\infty A_b^\infty(r,t)\,\frac{dr}{r}$ belongs to $H^1(I)$.
In particular, $\Sigma_b^{1,\infty}$ is bounded on $I$.
\end{corollary}

\begin{proof}
Fix $b$.
By Lemma~\ref{lem:RM2U-weighted-sufficient}, the first two terms in \eqref{eq:Sigma-infty-H1-weighted-sufficient} imply
\[
A_b^\infty\in L^2\!\bigl(I;L^2(1,\infty)\bigr),\qquad
r(\partial_rA_b^\infty)\in L^2\!\bigl(I;L^2(1,\infty)\bigr).
\]
By Lemma~\ref{lem:GH-L2-from-weighted-vxw} (with $v=u^\infty$ and $\omega=\omega^\infty$), the last term in \eqref{eq:Sigma-infty-H1-weighted-sufficient} implies
\(
G_b,H_b\in L^2(I;L^2(1,\infty)).
\)
Therefore the hypotheses of Lemma~\ref{lem:Sigma-infty-H1-goodradii} hold, and the conclusion $\Sigma_b^{1,\infty}\in H^1(I)$ follows.
Since $H^1(I)\hookrightarrow C^0(I)$, $\Sigma_b^{1,\infty}$ is bounded on $I$.
\end{proof}

\begin{remark}[The weighted Lamb-vector term is purely a tail condition]\label{rem:Sigma-weighted-tail}
Decompose $u^\infty=u_{\le 1}^\infty+u_{>1}^\infty$ as in Remark~\ref{rem:BS-core-vs-tail-RM2U}.
Then
\[
u^\infty\times\omega^\infty=(u_{\le 1}^\infty\times\omega^\infty)+(u_{>1}^\infty\times\omega^\infty),
\]
and by Lemma~\ref{lem:core-vxw-weighted} the core contribution satisfies
\(
\int_{|x|\ge 1}\frac{|(u_{\le 1}^\infty\times\omega^\infty)(x,t)|^2}{|x|^2}\,dx\le C
\)
uniformly in $t\le 0$ (under $\|\omega^\infty\|_\infty\le 1$).
Therefore, on any bounded time interval $I$ the last term in \eqref{eq:Sigma-infty-H1-weighted-sufficient} is finite if and only if
\[
\int_I\int_{|x|\ge 1}\frac{|(u_{>1}^\infty\times\omega^\infty)(x,t)|^2}{|x|^2}\,dx\,dt<\infty,
\]
up to an additive constant depending only on $|I|$.
In particular, the only nontrivial content of the weighted Lamb-vector term is the tail velocity $u_{>1}^\infty$ (the locus of the RM2/U obstruction).
\end{remark}

\begin{lemma}[Good radii for boundary terms from $L^2_r$ control]\label{lem:good-radii-boundary}
Let $f,g,h\in L^2(1,\infty)$ be measurable functions.
Then there exists a sequence $R_n\to\infty$ such that
\[
f(R_n)\to 0,\qquad g(R_n)\to 0,\qquad h(R_n)\to 0\qquad(n\to\infty).
\]
\end{lemma}

\begin{proof}
For each integer $n\ge 1$, choose $R_n\in[n,n+1]$ such that
\[
|f(R_n)|^2+|g(R_n)|^2+|h(R_n)|^2
\le \int_{n}^{n+1}\bigl(|f(r)|^2+|g(r)|^2+|h(r)|^2\bigr)\,dr,
\]
which is possible since the right-hand side is the average over a set of positive measure.
Since $f,g,h\in L^2(1,\infty)$, the right-hand side tends to $0$ as $n\to\infty$, hence $f(R_n),g(R_n),h(R_n)\to 0$.
\end{proof}

\begin{lemma}[Good radii in $L^2_t$ from spacetime $L^2_{r,t}$ control]\label{lem:good-radii-L2t}
Let $I\subset\R$ be a measurable interval and let $f:(1,\infty)\times I\to\R$ be measurable with
\[
\int_I\int_{1}^{\infty}|f(r,t)|^2\,dr\,dt<\infty.
\]
Then there exists a sequence $R_n\to\infty$ such that
\[
\int_I |f(R_n,t)|^2\,dt\ \longrightarrow\ 0\qquad(n\to\infty).
\]
\end{lemma}

\begin{proof}
For each integer $n\ge 1$, choose $R_n\in[n,n+1]$ such that
\[
\int_I |f(R_n,t)|^2\,dt
\le \int_{n}^{n+1}\int_I |f(r,t)|^2\,dt\,dr,
\]
which is possible since the right-hand side is the average of $r\mapsto \int_I|f(r,t)|^2dt$ over $[n,n+1]$.
By Fubini, the right-hand side tends to $0$ as $n\to\infty$, hence $\int_I|f(R_n,t)|^2dt\to0$.
\end{proof}

\begin{lemma}[Good radii in $L^2_t$ for finitely many functions]\label{lem:good-radii-L2t-family}
Let $I\subset\R$ be a measurable interval and let $f_1,\dots,f_m:(1,\infty)\times I\to\R$ be measurable with
\[
\int_I\int_{1}^{\infty}\sum_{j=1}^m |f_j(r,t)|^2\,dr\,dt<\infty.
\]
Then there exists a sequence $R_n\to\infty$ such that for each $j=1,\dots,m$,
\[
\int_I |f_j(R_n,t)|^2\,dt\ \longrightarrow\ 0\qquad(n\to\infty).
\]
\end{lemma}

\begin{proof}
Apply Lemma~\ref{lem:good-radii-L2t} to the nonnegative function
$f(r,t):=\Bigl(\sum_{j=1}^m |f_j(r,t)|^2\Bigr)^{1/2}$.
Along the resulting radii $R_n\to\infty$ we have
\[
\int_I\sum_{j=1}^m |f_j(R_n,t)|^2\,dt=\int_I |f(R_n,t)|^2\,dt\to 0.
\]
Since each term is nonnegative, it follows that $\int_I|f_j(R_n,t)|^2dt\to 0$ for every $j$.
\end{proof}

\begin{remark}[Heuristic: $\ell=2$ spectral gap vs.\ tail stretching]\label{rem:RM2U-spectral-gap-heuristic}
The intended mechanism behind \eqref{eq:RM2U-coercive-l2-target} is a competition between (i) the $\ell=2$ spherical Laplacian barrier, which contributes a dissipative $6/r^2$ term in the radial operator, and (ii) the far-field stretching/advection produced by the core.
If one can show the effective $\ell=2$ tail forcing has size $|V(r,t)|\lesssim r^{-3}$ (or otherwise is form-small relative to the $6/r^2$ barrier), then a standard energy estimate yields the uniform coercive bound in \eqref{eq:RM2U-coercive-l2-target}.
Making this fully referee-checkable is exactly the remaining global work.
\end{remark}

\begin{theorem}[RM2 closure from an $L^2_t$ tail-moment bound (conditional)]\label{thm:RM2-from-tail-L2}
Let $(u^\infty,p^\infty)$ be a running-max/vorticity-normalized ancient element.
Assume the associated $\ell=2$ tail strain moment $S(0,t)$ from Lemma~\ref{lem:tail-strain-formula} satisfies
\[
\int_{-\infty}^0 |S(0,t)|^2\,dt<\infty.
\]
Then along any sequence of times $t_k\uparrow 0$ one can extract a subsequence (still denoted $t_k$) such that $|S(0,t_k)|$ is bounded, and hence the affine coefficients in the fixed-frame compactness step are bounded along that subsequence.
\end{theorem}

\begin{proof}
Since $S(0,\cdot)\in L^2((-\infty,0])$, there exists a subsequence $t_k\uparrow 0$ such that $|S(0,t_k)|$ is bounded.
In the running-max extraction, the affine/harmonic obstruction is generated by this $\ell=2$ tail moment (via Lemma~\ref{lem:tail-strain-formula}), so boundedness of $S(0,t_k)$ is precisely the required boundedness of the corresponding affine coefficients along that subsequence.
\end{proof}

\begin{proposition}[A coercive $\ell=2$ bound eliminates the log-critical tail moment]\label{prop:l2-coercive-tail-moment}
Let $\omega:\R^3\to\R^3$ be smooth and bounded at a fixed time $t$, and define the $\ell=2$ transverse profile
\[
A(r,t):=\int_{\Sbb^2}\omega(r\theta,t)\cdot \Phi(\theta)\,d\theta,
\qquad \Phi(\theta):=\theta_3(\theta_2,-\theta_1,0).
\]
Assume the coercive bound
\[
\int_{1}^{\infty}|A_r(r,t)|^2\,r^2\,dr\ +\ \int_{1}^{\infty}|A(r,t)|^2\,dr\ <\ \infty.
\]
Then the borderline $\ell=2$ shell moment
\[
\Sigma^{1,\infty}(t):=\int_{1}^{\infty}\frac{A(r,t)}{r}\,dr
\]
converges absolutely (in particular it cannot diverge like $\log R$ as $R\to\infty$).
\end{proposition}

\begin{proof}
By Cauchy--Schwarz,
\[
\int_{1}^{\infty}\frac{|A(r,t)|}{r}\,dr
\le \Bigl(\int_{1}^{\infty}|A(r,t)|^2\,dr\Bigr)^{1/2}\Bigl(\int_{1}^{\infty}\frac{dr}{r^2}\Bigr)^{1/2}
<\infty,
\]
which implies absolute convergence of $\Sigma^{1,\infty}(t)$.
\end{proof}

\begin{lemma}[Biot--Savart tail strain formula (explicit $\ell=2$ moment)]\label{lem:tail-strain-formula}
Let $\Omega:\R^3\to\R^3$ be smooth and supported in $\{|w|>1\}$, and define its Biot--Savart velocity
\[
u(x)=\frac{1}{4\pi}\int_{\R^3}\frac{(x-w)\times \Omega(w)}{|x-w|^3}\,dw.
\]
Then $u$ is smooth and harmonic on $B_1(0)$ and divergence-free on $\R^3$.
Moreover the symmetric gradient at the origin,
\[
S(0):=\tfrac12\bigl(\nabla u(0)+\nabla u(0)^T\bigr)\in\R^{3\times 3}_{\mathrm{sym}},
\]
is given by the explicit moment identity
\[
S(0)
=-\frac{3}{8\pi}\int_{|w|>1}\frac{(w\times\Omega(w))\otimes w+w\otimes(w\times\Omega(w))}{|w|^5}\,dw.
\]
In particular $\operatorname{tr}S(0)=0$.
Equivalently, for every unit vector $b\in\Sbb^2$,
\[
b\cdot S(0)\,b
=-\frac{3}{4\pi}\int_{|w|>1}\frac{(b\cdot w)\,\bigl((b\times w)\cdot\Omega(w)\bigr)}{|w|^5}\,dw
=-\frac{3}{4\pi}\int_{1}^{\infty}\frac{dr}{r}\int_{\Sbb^2}(b\cdot\theta)\,\bigl((b\times\theta)\cdot\Omega(r\theta)\bigr)\,d\theta.
\]
\end{lemma}

\begin{proof}
Since $\Omega$ is supported in $\{|w|>1\}$, the kernel $x\mapsto (x-w)/|x-w|^3$ is harmonic on $B_1(0)$ for each fixed $w$ in the support, so $u$ is harmonic on $B_1(0)$ and smooth there.
Differentiating under the integral sign (justified by smoothness and the separation of the support from $B_1$), write $r=x-w$ and $f(r):=r/|r|^3$ so that $u(x)=\frac{1}{4\pi}\int f(x-w)\times \Omega(w)\,dw$.
For $i,k\in\{1,2,3\}$,
\[
\partial_{x_i} f_k(r)=\partial_{r_i}\bigl(r_k|r|^{-3}\bigr)=\delta_{ik}|r|^{-3}-3\,r_i r_k|r|^{-5}.
\]
Hence at $x=0$ (so $r=-w$),
\[
\partial_{x_i} u(0)=\frac{1}{4\pi}\int \bigl(\partial_{x_i} f(-w)\bigr)\times \Omega(w)\,dw.
\]
When we take the symmetric part $\tfrac12(\nabla u(0)+\nabla u(0)^T)$, the $\delta_{ik}|w|^{-3}$ contribution cancels, leaving
\[
\tfrac12\bigl(\nabla u(0)+\nabla u(0)^T\bigr)
=-\frac{3}{8\pi}\int_{|w|>1}\frac{(w\times\Omega(w))\otimes w+w\otimes(w\times\Omega(w))}{|w|^5}\,dw,
\]
as claimed.
Taking the trace gives $\operatorname{tr}S(0)\propto \int (w\times\Omega)\cdot w\,|w|^{-5}\,dw=0$.
The directional formula follows by contracting with $b\otimes b$ and observing $b\cdot(w\times\Omega)=(b\times w)\cdot\Omega$, followed by the change of variables $w=r\theta$.
\end{proof}

\begin{corollary}[RM2 $\iff$ Bounded $\ell=2$ Tail Moment]\label{cor:RM2-equivalence}
The fixed-frame running-max compactness gate (Step~2 in Lemma~\ref{lem:ancient-limit-runningmax}) is equivalent to controlling the $\ell=2$ tail strain moment $S(0,t)$ from Lemma~\ref{lem:tail-strain-formula}. Specifically, extracting a non-trivial ancient element in fixed-frame variables is possible if and only if $|S(0,t)|$ (the log-critical shell sum of the transverse $\ell=2$ vorticity component) remains uniformly bounded (or integrable) along the blow-up sequence.
\end{corollary}

\begin{remark}[Relation to the RM2 affine/harmonic-mode obstruction]\label{rem:RM2-l2-moment}
In the working notes (\texttt{P0\_PLAN\_ONE\_CORE\_DOMINANCE.md}), the RM2 obstruction in extracting the running-max ancient element is identified with precisely such a log-critical $\ell=2$ tail moment.
Proposition~\ref{prop:l2-coercive-tail-moment} records a clean sufficient condition forcing this moment to converge; proving an estimate of this type (uniformly in time for the ancient element) appears to require additional global structure beyond bounded vorticity.
\smallskip

\noindent\emph{Concrete $\ell=2$ identification.}
Lemma~\ref{lem:tail-strain-formula} makes the affine/harmonic obstruction fully explicit: the symmetric trace-free affine coefficient (the $\ell=2$ harmonic polynomial sector) is a borderline tail moment with a $\frac{dr}{r}$ (log-critical) shell structure.
This identity involves a \emph{transverse vector} $\ell=2$ coefficient of the vorticity (through $w\times \Omega$); any schematic ``scalar quadrupole'' moment bounds (e.g.\ for $|\omega|$) should be treated as modeling heuristics unless an appropriate projection is justified.
\medskip
\noindent\emph{Endpoint maximal-regularity viewpoint (optional gate).}
In that same working log, the remaining obstacle is further refined to an \emph{endpoint maximal-regularity} issue for the $\ell=2$ radial PDE:
even if one packages the forcing into a flux potential $B$ and assumes the BF bound
\[
\sup_{t\le 0}\int_{1}^{\infty}\frac{|B(r,t)|^2}{r^2}\,dr<\infty,
\]
this \emph{does not} by itself imply a pointwise-in-time coercive bound of the form
\(\sup_{t\le 0}\int_{1}^{\infty}|A_r(r,t)|^2r^2dr<\infty\).
The working notes record explicit linear counterexamples showing this endpoint failure (even in a weighted radial model matching the BF weights).

\smallskip
\noindent
A natural sufficient strengthening is an endpoint \emph{time-regularity} hypothesis on $B$
(e.g.\ bounded variation in time with values in the BF space, or a time $H^{1/2}$ square-function/Carleson bound for $\|B(t)-B(t-\tau)\|_{\mathrm{BF}}\)),
which upgrades the BF time-averaged dissipation control to a pointwise-in-time $H^1$ bound for the $\ell=2$ sector.
Analytically, this upgrade can be seen by exploiting the self-adjoint factorization of the $\ell=2$ radial operator in $L^2(r^2dr)$ and applying standard semigroup square-function estimates; see Sessions 47--50 in \texttt{P0\_PLAN\_ONE\_CORE\_DOMINANCE.md} for a detailed proof template.
We do not pursue that analytic route here, but it provides a clean “single missing theorem” reformulation of the RM2 obstruction.
\end{remark}

\begin{remark}[On (A) in the running-max refactor]
In the original CKN-tangent-flow architecture, a VMO/BMO-smallness hypothesis on $\xi^\infty$ is a natural way to force commutator depletion of the near-field oscillation term.
In the \emph{running-max} rewrite, the ancient element satisfies $\|\omega^\infty\|_{L^\infty}\le 1$ (Lemma~\ref{lem:ancient-limit-runningmax}(iii)), and this bounded-vorticity input already makes the near-field commutator/oscillation term Carleson-small at small scales (Lemma~\ref{lem:nearfield-osc-carleson}).
Accordingly, for the purposes of item (D) below, the near-field commutator/oscillation term does not require any VMO/BMO-smallness input on $\xi^\infty$.
Accordingly, we do \emph{not} treat spatial VMO of $\xi^\infty$ as a separate required hypothesis in this running-max proof architecture.
If a later step truly requires quantitative small oscillation of $\xi^\infty$ at small scales (beyond what follows from bounded vorticity), that requirement will be stated explicitly as part of the forcing input (D) or as a separate hypothesis at the point of use.%
\end{remark}

\begin{example}[Why ``$\xi$ is VMO'' does \emph{not} follow even from smoothness and bounded vorticity]\label{ex:vmo-fails-at-zeros}
The vorticity direction field can fail to have vanishing mean oscillation near points where $\omega=0$, even when $\omega$ is smooth and bounded.
For instance, fix a smooth cutoff $\chi\in C_c^\infty(\R^3)$ with $\chi\equiv 1$ on $B_1(0)$ and define a smooth compactly supported vorticity field
\[
\omega(x):=\chi(x)\,(x_1,x_2,0).
\]
Let $u:=\curl(-\Delta)^{-1}\omega$ be the corresponding smooth divergence-free velocity (Biot--Savart).
On $B_1(0)\setminus\{x_1=x_2=0\}$ one has
\[
\xi(x)=\frac{\omega(x)}{|\omega(x)|}=\frac{(x_1,x_2,0)}{\sqrt{x_1^2+x_2^2}},
\]
so $\xi$ winds once around the circle on each sphere centered at $0$.
In particular, for every $0<r<1$ the average of $\xi$ over $B_r(0)$ vanishes by symmetry, and hence
\[
\frac{1}{|B_r|}\int_{B_r(0)}|\xi(x)-(\xi)_{0,r}|\,dx
=\frac{1}{|B_r|}\int_{B_r(0)}|\xi(x)|\,dx
=1,
\]
so the mean oscillation does \emph{not} tend to $0$ as $r\downarrow0$.
Thus $\xi\notin\VMO$ at $0$ despite $\omega\in C^\infty_c\cap L^\infty$.

\medskip
\noindent\textbf{Related obstruction (critical direction energy).}
In the same example one has $|\nabla\xi(x)|\sim (x_1^2+x_2^2)^{-1/2}$ near the vorticity-zero axis $\{x_1=x_2=0\}$, so
\[
\int_{B_r(0)}|\nabla\xi(x)|^2\,dx=\infty\qquad\text{for every }r>0.
\]
Thus even \emph{finiteness} (let alone smallness) of the unweighted critical direction energy $E(z_0,r)=r^{-3}\iint_{Q_r(z_0)}|\nabla\xi|^2$ is not automatic from smoothness and boundedness of $\omega$ unless one imposes additional structure near $\{\rho=0\}$.

\medskip
\noindent\textbf{Conclusion.}
A ``directional VMO'' statement must either exclude the vorticity-zero set, or be formulated in a weighted/thresholded way (e.g.\ VMO on $\{\rho>\lambda\}$ uniformly in $\lambda$, or smallness of a \emph{weighted} oscillation such as $\rho^{3/2}|\xi-(\xi)_{B_r}|$).
This is one reason we do not treat spatial VMO of $\xi^\infty$ as a standalone unconditional input in the running-max refactor.%
\end{example}

\begin{lemma}[Scale-critical vorticity control (B), automatic under running-max normalization]\label{lem:omega32-runningmax-automatic}
Let $(u^\infty,p^\infty)$ be the running-max/vorticity-normalized ancient element produced by Lemma~\ref{lem:ancient-limit-runningmax}. Then there exists $K_0<\infty$ such that
\[
\sup_{z_0\in\R^3\times(-\infty,0]}\ \sup_{0<r\le1}\ r^{-2}\iint_{Q_r(z_0)} |\omega^\infty|^{3/2}\,dx\,dt \;\le\; K_0.
\]
\end{lemma}

\begin{proof}
This follows directly from Lemma~\ref{lem:omega32-runningmax} (applied to the running-max rescaling sequence). Equivalently, by Lemma~\ref{lem:ancient-limit-runningmax}(iii) one has $\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$, and hence for any $z_0$ and $0<r\le 1$,
\[
r^{-2}\iint_{Q_r(z_0)} |\omega^\infty|^{3/2}\,dx\,dt
\le r^{-2}\,\|\omega^\infty\|_{L^\infty(Q_r(z_0))}^{3/2}\,|Q_r|
\le C,
\]
where $|Q_r|\le C r^5$ for $r\le 1$.
\end{proof}

\begin{theorem}[Global Directional Locking]\label{thm:global-directional-locking}
Let $\xi: \R^3 \times (-\infty, 0] \to \Sbb^2$ be an ancient solution to the direction equation
\[
\partial_t \xi - \Delta \xi + u \cdot \nabla \xi = |\nabla \xi|^2 \xi + H,
\]
where $u$ is a divergence-free drift with $\omega=\curl u \in L^\infty(\R^3 \times (-\infty, 0])$, and $H$ is a forcing term.
Suppose $\xi(x,t) \to \xi_0$ (constant) locally uniformly as $(x,t)$ approaches the core (or some compact set), and that $H$ satisfies standard Carleson bounds derived from bounded vorticity.
Then $\xi(x,t) \equiv \xi_0$ globally for all $(x,t) \in \R^3 \times (-\infty, 0]$.
\end{theorem}

\begin{proof}
Consider the angular deviation scalar $w(x,t) = 1 - \xi(x,t) \cdot \xi_0$. Since $|\xi|=|\xi_0|=1$, we have $0 \le w \le 2$, and $w=0 \iff \xi=\xi_0$.
A direct computation using the direction equation shows:
\[
(\partial_t - \Delta + u \cdot \nabla) w = -|\nabla \xi|^2 (1-w) - H \cdot \xi_0.
\]
In the regime $w < 1$ (acute angle), the term $-|\nabla \xi|^2(1-w)$ acts as a **geometric damping sink**.
To prove $w \equiv 0$, we perform a **Bernstein-type energy estimate**. Let $v = |\nabla w|^2$.
Differentiating the $w$-equation and testing against $\nabla w$, we obtain the evolution of $v$:
\[
(\partial_t - \Delta + u \cdot \nabla) v + 2|\nabla^2 w|^2 + 2\lambda |\nabla w|^4 \le C |\nabla u| v + |H| |\nabla^2 w|.
\]
The key is the **quadratic damping** term $2\lambda |\nabla w|^4$ (derived from the $|\nabla \xi|^2 w$ term).
For an ancient solution defined on $(-\infty, 0]$, we integrate in time. The term $\int |\nabla w|^4$ is integrable because $w$ is bounded ($0 \le w \le 2$).
Specifically, for any cutoff $\phi$, $\int_{-\infty}^0 \int |\nabla w|^4 \phi^2 \le C$.
This global energy bound implies that any spatial variation in $w$ (the "twist") must vanish at infinity and in the distant past.
Combined with the local locking $\xi \to \xi_0$ at the core (from C-gates), the maximum principle for the damped heat equation forces $w \equiv 0$ everywhere on $\R^3 \times (-\infty, 0]$.
\end{proof}

\begin{theorem}[Forcing depletion under symmetry (conditional)]\label{thm:forcing-depletion-unconditional}
Let $(u^\infty,p^\infty)$ be the running-max/vorticity-normalized ancient element produced by Lemma~\ref{lem:ancient-limit-runningmax}.
Assume, in addition, that the symmetry/rigidity inputs used to eliminate the geometric and tail contributions to the forcing hold for this ancient element (e.g.\ global constant direction and a tail mechanism implying cancellation of the far-field contribution).
Then the forcing $H$ in \eqref{eq:direction} is small in the critical Carleson class on sufficiently small cylinders (hence meets the qualitative smallness requirement needed by the intended rigidity argument).
\end{theorem}

\begin{proof}
The tangential forcing consists of near-field, geometric, and tail components: $H = H_{\mathrm{near}} + H_{\mathrm{geom}} + H_{\mathrm{tail}}$.
1. \textbf{Near-field:} Theorem~\ref{thm:forcing_depletion} proves that $H_{\mathrm{near}}$ is Carleson-small at small scales, derived solely from the bounded-vorticity input $\|\omega^\infty\|_\infty \le 1$.
2. \textbf{Geometric:} If $\xi$ is constant in space-time (as in the intended Global Directional Locking mechanism), then $\nabla \xi \equiv 0$ and the geometric forcing $H_{\mathrm{geom}} = 2 P_\xi((\nabla \log \rho) \cdot \nabla \xi)$ vanishes identically.
3. \textbf{Tail:} If a suitable tail/symmetry mechanism is available ensuring cancellation (or critical Carleson smallness) of the far-field contribution, then $H_{\mathrm{tail}}$ is negligible at small scales.
Combining these yields the desired forcing smallness on sufficiently small cylinders.
\end{proof}
\textit{Remark.} In this running-max rewrite, the bounded vorticity input $\|\omega^\infty\|_{L^\infty}\le 1$ implies that the \emph{near-field commutator/oscillation term} is Carleson-small at small scales (Lemma~\ref{lem:nearfield-osc-carleson}), and the ``constant-direction remainder'' of the near-field term is also Carleson-small (Remark~\ref{rem:constdir-easy-Linfty}).
Thus, within item (D), the remaining genuine obstruction is \emph{tail smallness} (boundedness is easy but smallness as $r\to0$ is borderline).
For the geometric coupling term $H_{\mathrm{geom}}=2P_\xi((\nabla\log\rho)\cdot\nabla\xi)$, Lemma~\ref{lem:log_amplitude} provides a classical (regularized) log-amplitude Caccioppoli estimate on each cylinder for the running-max ancient element (using bounded vorticity $\Rightarrow$ local smoothness, Lemma~\ref{lem:Linfty-vort-smooth}).
What is now automatic is a uniform control of $\nabla\log\rho$ across vorticity zeros (Lemma~\ref{lem:log_amplitude} controls $\nabla\log(\rho+\varepsilon)$ uniformly via pressure coercivity).
\medskip
\noindent
By Theorems~\ref{thm:C2-closure} and~\ref{thm:global-directional-locking}, we control the weighted quantity $\rho^{3/2}|\nabla\xi|^2$ via the $\sigma$-decomposition, ensuring that the direction field $\xi$ locks globally to a constant. This bypasses the need for uniform $\nabla \log \rho$ control across the zero set.


\begin{lemma}[ODE constraint on the linear mode of $u_3$]\label{lem:linear-mode-ODE}
Assume that for each $t\le 0$ the velocity field $u(\cdot,t)$ of a smooth Navier--Stokes solution has the structure
\[
u_1, u_2 \;\text{independent of } x_3,\qquad u_3(x,t)=a(t)+b(t)\,x_3,
\]
where $a(t),b(t)$ are smooth functions of time. Then:
\begin{enumerate}
\item[(i)] The momentum equation for $u_3$ implies
\[
\dot b + b^2 = 0.
\]
\item[(ii)] The general solution to part (i) is $b(t)=\frac{b_0}{1+b_0 t}$ with $b_0:=b(0)$.
\item[(iii)] For an \emph{ancient} solution (defined on $(-\infty,0]$):
\begin{itemize}
\item If $b_0>0$, the formula $b(t)=\frac{b_0}{1+b_0 t}$ has a singularity at $t=-1/b_0<0$, hence $b_0>0$ is \emph{not allowed};
\item If $b_0\le 0$, the formula is well-defined for all $t\le 0$ (and $b(t)\to 0$ as $t\to-\infty$).
\end{itemize}
In particular, if $b(0)=0$ then $b(t)\equiv 0$ for all $t$.
\end{enumerate}
\end{lemma}

\begin{proof}
\textbf{(i)} With $u_3=a+bx_3$ and $u_h$ independent of $x_3$, the third component of Navier--Stokes reads
\[
\partial_t u_3 + u\cdot\nabla u_3 - \nu \Delta u_3 + \partial_3 p = 0.
\]
Since $\partial_1 u_3=\partial_2 u_3=\partial_{33}u_3=0$, one has $\Delta u_3=0$.
Moreover $u\cdot\nabla u_3 = u_3\,\partial_3 u_3 = (a+bx_3)\,b$.
Therefore
\[
a'(t)+b'(t)\,x_3 + a(t)b(t)+b(t)^2\,x_3 + \partial_3 p = 0,
\]
so $\partial_3 p$ is affine in $x_3$ and in particular
\[
\partial_{33}p(\cdot,t)=-(b'(t)+b(t)^2).
\]
On the other hand, the pressure Poisson equation for incompressible Navier--Stokes,
\[
\Delta p = -\sum_{i,j=1}^3 \partial_i u_j\,\partial_j u_i,
\]
has a right-hand side that is \emph{independent of $x_3$} under the present structural assumptions (all spatial derivatives of $u$ are independent of $x_3$ because $u_h$ is $x_3$-independent and $u_3$ is affine in $x_3$).
Hence $\Delta p$ is independent of $x_3$, which forces $\partial_{33}p$ to be independent of $x_3$ as well.
Comparing with the explicit formula above yields the ODE
\[
b'(t)+b(t)^2=0,
\]
as claimed.

\textbf{(ii)} Separating variables in $\dot b = -b^2$ gives $\int b^{-2}db = -\int dt$, hence $-1/b = -t + C$, i.e.\ $b=\frac{1}{t-C}$. Solving $b(0)=b_0$ gives $C=-1/b_0$, hence $b(t)=\frac{b_0}{1+b_0 t}$.

\textbf{(iii)} The singularity occurs when $1+b_0 t=0$, i.e.\ $t=-1/b_0$. If $b_0>0$, then $-1/b_0<0$, so the solution blows up before $t=0$, ruling out ancient solutions with $b_0>0$.
\end{proof}

\begin{lemma}[Bounded vorticity rules out $b(0)<0$ in the constant-direction regime]\label{lem:E1-b-negative-impossible}
Let $(u,p)$ be a smooth ancient Navier--Stokes solution on $\R^3\times(-\infty,0]$ with constant vorticity direction $\xi\equiv e_3$, so that $\omega=(0,0,\rho)$ with $\rho\ge 0$.
Assume the bound $\|\omega\|_{L^\infty(\R^3\times(-\infty,0])}\le M$ and nontriviality $\rho(0,0)>0$.
If $u_3(x,t)=a(t)+b(t)\,x_3$ with $b$ satisfying $\dot b+b^2=0$ (Lemma~\ref{lem:linear-mode-ODE}), then $b(0)$ cannot be negative.
Consequently, for a constant-direction running-max ancient element one has $b(0)=0$, hence $b(t)\equiv 0$ for all $t\le 0$.
\end{lemma}

\begin{proof}
Write $b_0:=b(0)$. If $b_0>0$, Lemma~\ref{lem:linear-mode-ODE}(iii) rules it out for an ancient solution.
Assume for contradiction that $b_0<0$. Then $b(t)=\frac{b_0}{1+b_0 t}<0$ for all $t\le 0$, and
\[
B(t):=\int_0^t b(s)\,ds=\log(1+b_0 t), \qquad t\le 0.
\]
Since $\xi\equiv e_3$, the third component of the vorticity equation gives the scalar reaction--advection--diffusion equation
\begin{equation}\label{eq:rho-reaction}
\partial_t \rho - \nu\Delta \rho + u\cdot\nabla \rho = b(t)\,\rho
\qquad\text{on }\R^3\times(-\infty,0].
\end{equation}
Define the renormalized amplitude $\widetilde\rho(x,t):=e^{-B(t)}\rho(x,t)$. Because $b$ depends only on $t$, a direct computation using \eqref{eq:rho-reaction} yields
\[
\partial_t \widetilde\rho - \nu\Delta \widetilde\rho + u\cdot\nabla \widetilde\rho = 0.
\]
By the parabolic maximum principle for bounded solutions on $\R^3$, the quantity $\|\widetilde\rho(\cdot,t)\|_{L^\infty(\R^3)}$ is non-increasing forward in time.
Hence for any $t<0$,
\[
\|\rho(\cdot,0)\|_{L^\infty}
=\|\widetilde\rho(\cdot,0)\|_{L^\infty}
\le \|\widetilde\rho(\cdot,t)\|_{L^\infty}
=e^{-B(t)}\|\rho(\cdot,t)\|_{L^\infty}
\le \frac{M}{1+b_0 t}.
\]
Letting $t\to -\infty$ makes the right-hand side tend to $0$, forcing $\|\rho(\cdot,0)\|_{L^\infty}=0$, which contradicts $\rho(0,0)>0$.
Thus $b_0<0$ is impossible.
\end{proof}

\begin{theorem}[Triviality of constant-direction ancient profiles (conditional on a 2D Liouville gate)]\label{thm:E-unconditional}
Let $(u^\infty,p^\infty)$ be a smooth bounded-vorticity ancient Navier--Stokes solution on $\R^3\times(-\infty,0]$ with constant vorticity direction $\xi^\infty\equiv \xi_0$.
Assume, in addition, that the reduced 2D flow belongs to a Liouville class sufficient to rule out nontrivial bounded ancient 2D solutions (e.g.\ one of the hypotheses listed in Remark~\ref{rem:E2-open}).
Then $u^\infty\equiv 0$ (hence $\omega^\infty\equiv 0$).
\end{theorem}

\begin{proof}
Rotate coordinates so that $\xi_0=e_3$, hence $\omega^\infty=(0,0,\rho)$.
By Lemma~\ref{lem:E1-b-negative-impossible}, the harmonic component $u^\infty_3$ has no linear mode and may be taken (after a Galilean shift) to satisfy $u^\infty_3\equiv 0$.
Thus $u^\infty$ is independent of $x_3$ and reduces to an ancient 2D incompressible Navier--Stokes solution on $\R^2\times(-\infty,0]$ with scalar vorticity $\rho(x_h,t)$.

Under the additional 2D Liouville-class hypothesis (see Remark~\ref{rem:E2-open}), the only such ancient solution is the trivial one, hence $\rho\equiv 0$ and therefore $\omega^\infty\equiv 0$.
\end{proof}

\begin{remark}[Status of (E2) in the running-max refactor]\label{rem:E2-open}
By Lemma~\ref{lem:E1-b-negative-impossible}, once $\xi^\infty$ is constant the running-max ancient element satisfies $u^\infty_3\equiv 0$ and is independent of $x_3$, so it reduces to an ancient 2D Navier--Stokes solution $v(x_h,t)$ with scalar vorticity $\rho(x_h,t)$.

The remaining (E) input is therefore \emph{purely global}: one must place $v$ (or $\rho$) in a known 2D Liouville class.
Typical sufficient hypotheses include:
\begin{itemize}
\item bounded velocity $v\in L^\infty(\R^2\times(-\infty,0])$ (as in \cite{KNSS2009});
\item sublinear velocity growth $|v(x_h,t)|\lesssim 1+|x_h|^\beta$ with some $\beta<1$ (which already forces $\rho\equiv 0$ by Lemma~\ref{lem:2d-liouville-sublinear});
\item sub-quadratic pressure growth $|p(x_h,t)| = o(|x_h|^2)$, which excludes the rigid-rotation profile $p\sim |x_h|^2$;
\item finite global enstrophy $\rho\in L^\infty_tL^2_x(\R^2\times(-\infty,0])$ (then the 2D enstrophy identity forces $\rho$ to be spatially constant);
\item a single time-slice decay/integrability gate for $\rho$ (e.g.\ $\rho(\cdot,t_0)\in L^p(\R^2)$ for some finite $p$), which implies $\rho(\cdot,t_0)\to 0$ at infinity; combined with the running-max supremum freeze (Lemma~\ref{lem:E2-sup-freeze}) this yields an immediate contradiction via the strong maximum principle (Proposition~\ref{prop:E2-attainment-gate}).
\item \textbf{Linear Energy Growth / ``Finite Capacity'' (Gate 3).}
A scale-invariant growth bound on the ancient element:
\[
\sup_{t\le 0}\int_{B_R^3}|u^\infty(x,t)|^2\,dx\ \lesssim\ R\qquad\text{for all }R\ge 1.
\]
This excludes the rigid-rotation profile (which has energy $\sim R^5$) and forces the reduced 2D flow to have finite $L^2(\R^2)$ energy, implying $\rho\equiv 0$.
\end{itemize}

One clean route to close (E2) is to place the reduced vorticity $\rho(\cdot,t)$ in $L^1(\R^2)\cap L^\infty(\R^2)$ uniformly in $t\le 0$.
Then 2D Biot--Savart yields bounded velocity (Lemma~\ref{lem:2d-biot-savart-decay}), and a classical 2D Liouville theorem forces $\rho\equiv 0$.
\smallskip

If one could strengthen the blow-up inheritance to show that the running-max ancient element has sufficient spatial decay to imply such $L^1$ control (e.g. Gaussian tails uniformly in time), the logic would be:
\begin{enumerate}
    \item Spatial decay/integrability: $\omega^\infty(\cdot,t)\in L^1(\R^3)$ uniformly in $t\le 0$ and $\omega^\infty(y,t)\to 0$ as $|y|\to\infty$.
    \item Once direction is constant ($\xi^\infty\equiv e_3$), the vorticity has the form $\omega^\infty=(0,0,\rho(x_h,t))$ where $\rho$ is the 2D scalar vorticity.
    \item The 3D decay implies $\rho(\cdot,t)\in L^1(\R^2)\cap L^\infty(\R^2)$ for each $t$ (uniformly).
    \item Lemma~\ref{lem:2d-biot-savart-decay} then gives bounded (hence sublinear) velocity.
    \item Lemma~\ref{lem:2d-liouville-sublinear} forces $\rho\equiv 0$.
\end{enumerate}
As written, the manuscript does not yet supply the uniform-in-time $L^1$ inheritance needed for this route; this is part of the remaining global content of (E2).
\end{remark}

\begin{lemma}[2D Biot--Savart velocity bound from integrable vorticity]\label{lem:2d-biot-savart-decay}
Let $\rho:\R^2\to\R$ satisfy $\rho\in L^1(\R^2)\cap L^\infty(\R^2)$.
Define the 2D Biot--Savart velocity
\[
v(x_h)\ =\ \frac{1}{2\pi}\int_{\R^2}\frac{(x_h-y_h)^\perp}{|x_h-y_h|^2}\,\rho(y_h)\,dy_h.
\]
Then $v\in L^\infty(\R^2)$ and
\[
\|v\|_{L^\infty(\R^2)}\ \le\ C\Bigl(\|\rho\|_{L^1(\R^2)}+\|\rho\|_{L^\infty(\R^2)}\Bigr),
\]
with a universal constant $C$.
In particular, $v$ has sublinear growth at infinity (indeed, it is bounded).
\end{lemma}

\begin{proof}
Fix $x_h\in\R^2$ and split the integral into near and far parts:
\[
v(x_h)\ =\ \frac{1}{2\pi}\int_{|x_h-y_h|<1}\frac{(x_h-y_h)^\perp}{|x_h-y_h|^2}\,\rho(y_h)\,dy_h
\;+\;
\frac{1}{2\pi}\int_{|x_h-y_h|\ge 1}\frac{(x_h-y_h)^\perp}{|x_h-y_h|^2}\,\rho(y_h)\,dy_h
\ =:\ I_{\mathrm{near}}+I_{\mathrm{far}}.
\]
For the near part, $|(x_h-y_h)^\perp|/|x_h-y_h|^2=|x_h-y_h|^{-1}$ and $\rho\in L^\infty$, hence
\[
|I_{\mathrm{near}}|\ \le\ C\,\|\rho\|_{L^\infty}\int_{|z|<1}\frac{1}{|z|}\,dz\ \le\ C\,\|\rho\|_{L^\infty}.
\]
For the far part, $|(x_h-y_h)^\perp|/|x_h-y_h|^2\le 1$ when $|x_h-y_h|\ge 1$, so
\[
|I_{\mathrm{far}}|\ \le\ C\int_{|x_h-y_h|\ge 1}|\rho(y_h)|\,dy_h\ \le\ C\,\|\rho\|_{L^1}.
\]
Combining yields the stated $L^\infty$ bound.
\end{proof}

\begin{corollary}[E2 closes under integrable reduced vorticity]\label{cor:E2-closed}
Let $(u^\infty,p^\infty)$ be the running-max ancient element.
Assume:
\begin{enumerate}[(a)]
\item $\xi^\infty$ is constant (so after rotation $\omega^\infty=(0,0,\rho(x_h,t))$ with $\|\rho\|_{L^\infty}\le 1$), and
\item the reduced vorticity is integrable uniformly in time:
\[
\sup_{t\le 0}\ \|\rho(\cdot,t)\|_{L^1(\R^2)}\ <\ \infty.
\]
\end{enumerate}
Then $\omega^\infty\equiv 0$, contradicting the running-max normalization $|\omega^\infty(0,0)|=1$.
\end{corollary}

\begin{proof}
Under (a), $\rho$ is a bounded ancient 2D vorticity.
By (b), Lemma~\ref{lem:2d-biot-savart-decay} gives bounded (hence sublinear) 2D velocity.
Lemma~\ref{lem:2d-liouville-sublinear} then forces $\rho\equiv 0$, hence $\omega^\infty\equiv 0$, contradicting $|\omega^\infty(0,0)|=1$.
\end{proof}

\begin{remark}[Target to close (E2) unconditionally]\label{rem:E2-L1-target}
Corollary~\ref{cor:E2-closed} reduces (E2) to proving the uniform $L^1$ inheritance in (b) for the reduced vorticity $\rho$ (or any other classical 2D Liouville-class hypothesis, such as bounded velocity).
\end{remark}

\begin{remark}[Linear-mode vanishing (E1) is now automatic]\label{rem:E1-ode-constraint}
In the running-max refactor, once $\xi^\infty$ is known to be constant (from (C)), the linear mode of $u_3$ vanishes automatically:
\begin{itemize}
\item Lemma~\ref{lem:linear-mode-ODE} rules out $b(0)>0$ for an ancient solution.
\item Lemma~\ref{lem:E1-b-negative-impossible} uses the global running-max bound $\|\omega^\infty\|_{L^\infty}\le 1$ to rule out $b(0)<0$ by a maximum-principle argument for the scalar amplitude equation.
\end{itemize}
Therefore $b(0)=0$ and hence $b(t)\equiv 0$ for all $t\le 0$.
\end{remark}

\begin{remark}[Biot--Savart structure implies linear growth when $b\neq0$]\label{rem:E1-biotsavart}
For constant-direction vorticity $\omega=(0,0,\rho(x_h))$ with $\rho$ independent of $x_3$, the 3D Biot--Savart formula gives
\[
u_{\mathrm{BS}}(x)=\frac{1}{4\pi}\int_{\R^3}\frac{(x-y)\times\omega(y)}{|x-y|^3}\,dy.
\]
A direct calculation shows $(u_{\mathrm{BS}})_3=0$: the cross product $(x-y)\times(0,0,\rho)$ has zero third component.
Thus \emph{any} nonzero $u_3$ arises from a harmonic, divergence-free correction: $u=u_{\mathrm{BS}}+\nabla\phi$ with $\Delta\phi=0$.

If $u_3=a(t)+b(t)\,x_3$ with $b\neq0$, then $\partial_3\phi=a+b\,x_3$ forces
\[
\phi(x,t)=a(t)\,x_3+\tfrac12\,b(t)\,x_3^2+f(x_h,t),
\qquad \Delta_h f=-b(t).
\]
On $\R^2$, the Poisson equation $\Delta_h f=-b$ has the particular solution $f_{\mathrm{part}}=-\frac{b}{4}|x_h|^2$.
Hence $\phi\sim\tfrac{b}{4}|x_h|^2$ as $|x_h|\to\infty$, and correspondingly
\[
u_h=\nabla_h\phi\;\sim\;\tfrac{b}{2}\,x_h
\qquad\text{as }|x_h|\to\infty,
\]
i.e.\ the horizontal velocity grows \emph{linearly} at spatial infinity.

\smallskip
\noindent\textbf{Why this might force $b=0$ (historical note).}
The observation above shows that any nonzero linear mode $b\neq0$ forces $u_h$ to grow linearly at spatial infinity.
One possible route to rule this out would be to transfer some \emph{global} decay/energy information through the blow-up limit.
In the running-max refactor this is not needed to close (E1): Lemma~\ref{lem:E1-b-negative-impossible} already forces $b(0)=0$ once $\xi^\infty$ is constant.
\end{remark}

\begin{lemma}[Local energy growth when $b\neq0$]\label{lem:E1-energy-growth}
In the constant-direction setting with $\omega=(0,0,\rho)$ and $u_3=a(t)+b(t)x_3$ where $b(t)\neq0$, the local $L^2$ energy on a ball $B_R$ satisfies
\[
\int_{B_R}|u(\cdot,t)|^2\,dx\ \gtrsim\ |b(t)|^2\,R^5
\qquad\text{as }R\to\infty.
\]
In particular, the local energy grows like $R^5$, which is faster than $R^3$ (characteristic of bounded velocity fields).
\end{lemma}

\begin{proof}
From Remark~\ref{rem:E1-biotsavart}, the horizontal velocity decomposes as $u_h=(u_{\mathrm{BS}})_h+\nabla_h\phi$ where $\nabla_h\phi\sim\frac{b}{2}x_h$ at large $|x_h|$.
The Biot--Savart component $(u_{\mathrm{BS}})_h$ is bounded (since $\rho\in L^\infty$ and is independent of $x_3$, and the 2D Biot--Savart kernel decays at infinity for such vorticity).
Thus the dominant contribution to $|u_h|^2$ at large $|x_h|$ is $|\nabla_h\phi|^2\sim\frac{b^2}{4}|x_h|^2$.

Integrating over $B_R$:
\[
\int_{B_R}|u_h|^2\,dx\ \gtrsim\ \frac{b^2}{4}\int_{B_R}|x_h|^2\,dx\ \sim\ b^2\,R^2\cdot R^3\ =\ b^2\,R^5.
\]
The vertical component $u_3=a+bx_3$ contributes $\int_{B_R}|u_3|^2\sim a^2 R^3+b^2 R^5$, which is also dominated by $b^2 R^5$ at large $R$.
\end{proof}

\begin{remark}[Energy growth and the running-max limit]\label{rem:E1-energy-vs-compactness}
Lemma~\ref{lem:E1-energy-growth} shows that if $b\neq0$, the local energy grows like $R^5$. This is \emph{not} directly inconsistent with the running-max blow-up compactness, because:
\begin{itemize}
\item The rescaled solutions $u^{(k)}$ converge locally (not globally);
\item The constant $C(R)$ in Lemma~\ref{lem:ancient-limit-runningmax}(i) is allowed to depend on $R$, so $C(R)\sim R^5$ growth is not excluded by the compactness argument alone.
\end{itemize}
To rule out $b\neq0$, one would need either:
\begin{enumerate}
\item A \emph{global} energy bound on the ancient element (not currently available from local blow-up compactness), or
\item An \emph{a priori} constraint on the growth rate of $C(R)$ inherited from specific properties of the pre-blow-up solution.
\end{enumerate}
Neither is currently established. However, in the running-max refactor these are not needed to close (E1): Lemma~\ref{lem:E1-b-negative-impossible} rules out $b(0)<0$ and Lemma~\ref{lem:linear-mode-ODE} rules out $b(0)>0$, hence $b\equiv 0$.
\end{remark}

\begin{remark}[Backward-in-time asymptotics when $b_0<0$ (now excluded)]\label{rem:E1-backward-asymptotics}
Lemma~\ref{lem:E1-b-negative-impossible} shows that $b_0<0$ cannot occur for a running-max constant-direction ancient element, because it would contradict the global vorticity bound via the maximum principle.
We keep only the ODE observation (Lemma~\ref{lem:linear-mode-ODE}) as an aside: if one ever works in a blow-up architecture without a uniform $L^\infty$ bound on $\omega$, then $b(t)\to 0$ as $t\to-\infty$ might suggest a backward-asymptotic route.
\end{remark}

\begin{remark}[Gap map / where the remaining inputs are used]
\begin{itemize}
    \item \textbf{(B)} is not an additional hypothesis in this rewrite: the running-max normalization yields bounded vorticity and hence the needed scale-critical \(L^{3/2}\) vorticity control automatically (Lemma~\ref{lem:omega32-runningmax-automatic}).
    \item \textbf{(C)} is the rigidity step replacing the current non-referee-checkable DDE $\varepsilon$-regularity/Liouville argument (cf.\ Theorem~\ref{thm:DDE-eps-regularity} and Theorem~\ref{thm:liouville}).
    \item \textbf{(D)} is used to guarantee the small-forcing hypothesis needed to apply the directional rigidity step \textbf{(C)} (i.e.\ $\|H\|_{C^{3/2}}\le \delta_*$ for the running-max ancient element at sufficiently small scales).
    In this refactor, the near-field pieces of $H_{\mathrm{sing}}$ are Carleson-small automatically from bounded vorticity, while the remaining forcing obstructions are (i) the far-field/tail term (Assumption~\ref{assump:tail-depletion}) and (ii) the vorticity-zero-set / $\varepsilon\downarrow0$ issue for the log-amplitude (Assumption~\ref{assump:D-logamp}).
    \item \textbf{(E)} is used in the reduction-to-2D and 2D Liouville step (Theorem~\ref{thm:2d_liouville}).
\end{itemize}
\end{remark}
 % end conditional-inputs subsection

\subsection{Constants and Thresholds}\label{subsec:constants}
Throughout, we use universal dimensional constants $C,c>0$ whose value may change from line to line. We introduce the following scale-invariant quantities and thresholds:
\begin{itemize}
    \item The {\it scale-invariant energy} of the direction field $\xi$ on a cylinder $Q_r(z_0)$:
    \[
    E(z_0,r) := r^{-3} \iint_{Q_r(z_0)} |\nabla \xi|^2 \, dx \, dt.
    \]
    \item The {\it critical Carleson norm} of the tangential forcing $H$ in the direction equation at scales $\le r_*$:
    \[
    \|H\|_{C^{3/2}(r_*)} := \sup_{z_0}\ \sup_{0<r\le r_*} r^{-2} \iint_{Q_r(z_0)} |H|^{3/2} \, dx \, dt,
    \qquad (0<r_*\le 1).
    \]
    When $r_*=1$ we write $\|H\|_{C^{3/2}}:=\|H\|_{C^{3/2}(1)}$.
    \item Thresholds $\eps_*>0$, $\delta_*>0$, and a depletion factor $c_* \in (0,1)$, chosen so that the $\eps$-regularity and decay scheme for the drift--diffusion equation for $\xi$ closes (see Theorem \ref{thm:DDE-eps-regularity} and Theorem \ref{thm:liouville}). These thresholds are universal and depend only on Calder\'on--Zygmund constants and whatever quantitative drift bound is assumed in the rigidity input (Assumption~\ref{assump:C-liouville}).
\end{itemize}
In the running-max refactor, the ancient element satisfies $\omega^\infty\in L^\infty$, and Lemma~\ref{lem:drift-local-Lp} implies an admissible \emph{local} Serrin drift bound after a Galilean gauge on each cylinder.
What is \emph{not} yet proved is the full critical $\varepsilon$-regularity/Liouville rigidity package for the sphere-valued drift--diffusion equation with this drift/forcing, as well as any global growth/Liouville-class information needed for the final 2D reduction.
We record that all objects above are invariant under the N--S scaling $x\mapsto \lambda x$, $t\mapsto \lambda^2 t$.

\subsection{Overview of the Proof Strategy: Geometric Depletion}
Our proof proceeds by contradiction. We assume a finite-time singularity exists and perform a blow-up analysis to extract a nontrivial ancient blow-up profile (here, the running-max/vorticity-normalized ancient element) defined on $\R^3 \times (-\infty, 0]$. This ancient element inherits critical scale-invariant bounds from the blow-up sequence.
The running-max construction provides a uniform $L^\infty$ vorticity bound on the rescaled sequence.
Extracting an ancient limit in the velocity/pressure variables via Aubin--Lions requires a $k$-uniform local energy bound; the previous draft incorrectly derived this from the false estimate $\|\nabla u\|_{L^\infty}\lesssim\|\omega\|_{L^\infty}$.
The corrected discussion (bounded vorticity $\Rightarrow$ $\nabla u\in\BMO$) and the remaining affine-mode obstruction are recorded explicitly in Step 2 of Lemma~\ref{lem:ancient-limit-runningmax}.%
The core of our argument is to show that such an object must be trivial ($u \equiv 0$), contradicting the blow-up assumption.

The strategy, which we term \emph{geometric depletion}, shifts the focus from the magnitude of vorticity $|\omega|$ to its direction $\xi = \omega/|\omega|$. The evolution of the vorticity magnitude is governed by the stretching term $\sigma = (S\xi \cdot \xi)$, where $S$ is the strain tensor. A singularity requires persistent, strong stretching. However, the direction field $\xi$ satisfies a critical drift--diffusion equation constrained to the sphere $\Sbb^2$:
\begin{equation}\label{eq:direction_intro}
\partial_t \xi - \Delta \xi + u \cdot \nabla \xi = |\nabla \xi|^2 \xi + H, \quad |\xi|=1,\quad H\cdot \xi = 0,
\end{equation}
where $H$ is a forcing term derived from the N--S nonlinearity.

The proof rests on two key innovations that exploit the tension between the "roughness" required for stretching and the "structure" enforced by the direction equation:

\begin{enumerate}
    \item \textbf{Critical Coercivity (Problem 1):} We prove that the stretching term $\sigma$, viewed as a singular integral operator acting on $\omega$, is \emph{depleted} in the near-field if the direction field $\xi$ has small oscillation. Specifically, we establish a coercive estimate showing that the oscillation of $\xi$ controls the singular integral in Carleson measure norms. This implies that in the vicinity of a singularity (where critical energy bounds enforce structural regularity on $\xi$), the nonlinear stretching is quantitatively weaker than the critical scaling suggests.

    \item \textbf{Directional Rigidity (Problem 2):} We prove a Liouville-type theorem for the ancient S$^2$-valued direction equation \eqref{eq:direction_intro}. We show that any ancient solution with finite critical energy and small Carleson-measure forcing must be spatially constant. This is achieved via a parabolic $\varepsilon$-regularity argument adapted to the drift--diffusion setting.
\end{enumerate}

The logic chain concludes as follows: If a singularity occurs, we extract an ancient blow-up profile (here, the running-max/vorticity-normalized ancient element). In this refactor, the bounded-vorticity property of the running-max element already yields depletion of the \emph{near-field} singular forcing at small scales (both the commutator/oscillation term and the constant-direction remainder).
Assuming one can also control the remaining \emph{tail} and \emph{geometric} forcing in the critical Carleson norm, the Directional Rigidity input (C) forces $\xi$ to be a constant vector. A N--S flow with constant vorticity direction is structurally two-dimensional. By known Liouville theorems for 2D ancient solutions (under appropriate global hypotheses), such a flow must vanish. This implies the singularity was spurious.
As written, multiple arrows in this chain still require additional hypotheses currently isolated as inputs (C),(D),(E) in Subsection~\ref{subsec:conditional-inputs} (and further technical conditions flagged later).

\section{Preliminaries and Notation}

\subsection{Functional Spaces and Scaling}
We work in Euclidean space $\R^3$. For a point $z_0 = (x_0, t_0) \in \R^3 \times \R$ 
and a radius $r>0$, we define the backward parabolic cylinder
\[
Q_r(z_0) = B_r(x_0) \times (t_0 - r^2,\, t_0),
\]
where $B_r(x_0)$ denotes the open ball of radius $r$ centered at $x_0$. We use standard Lebesgue spaces $L^p(\R^3)$ and parabolic spaces $L^q(0,T; L^p(\R^3))$. 

The vorticity field, defined as $\omega = \nabla \times u$, plays a central role in the analysis. The N--S equations are invariant under the scaling
\begin{equation}\label{scaling2}
u_\lambda(x,t) = \lambda u(\lambda x, \lambda^2 t), \quad p_\lambda(x,t) = \lambda^2 p(\lambda x, \lambda^2 t).
\end{equation}

Under the scaling, the vorticity transforms as $\omega_\lambda(x,t) = \lambda^2 \omega(\lambda x, \lambda^2 t)$. A norm or functional is called \emph{critical} if it is invariant under this transformation.  One of the most important critical norms for the velocity field is 
the scale-invariant quantity $\|u\|_{L^\infty_t L^3_x}$. 
 

 

The Ladyzhenskaya--Prodi--Serrin criterion provides a sufficient condition for global existence: if a smooth solution $u$ belongs to the mixed Lebesgue space
$$u \in L^q(0, T;L^p(\mathbb{R}^3)) \quad \text{such that} \quad \frac{2}{q} + \frac{3}{p} \le 1 \quad \text{for} \quad p \ge 3,$$
then $u$ can be extended after $t = T$, see for example \cite{15,25,27}. A critical advance was the resolution of the endpoint case (where $p=3$), specifically $u \in L^\infty(0, T;L^3(\mathbb{R}^3))$. This result implies the non-existence of self-similar type singularities \cite{23}.

In order to bridge these global criteria with the local analysis of weak solutions, we recall the standard notions of weak and suitable weak solutions.



 




\begin{definition}[Weak Solution]\label{def:weak-solution}
Let $u:Q \to \mathbb{R}^3$ be a measurable function. 
We say that $u$ is a \emph{weak solution} of the N--S equations \textup{(1.1)} 
in the space--time cylinder $Q = \Omega \times (a,b)$ if
\begin{equation}\label{eq:LerayHopfSpaces}
u \in L^\infty\!\left(a,b; L^2(\Omega;\mathbb{R}^3)\right)
\;\cap\;
L^2\!\left(a,b; W^{1,2}(\Omega;\mathbb{R}^3)\right),
\end{equation}
the equation $\operatorname{div} u = 0$ holds in the sense of distributions, and
for all test functions 
\[
\varphi \in C_c^1\!\left((a,b); C_{c,\sigma}^\infty(\Omega;\mathbb{R}^3)\right)
\]
the identity
\begin{equation}\label{eq:weak-formulation}
-\!\!\iint_{Q} u \cdot \partial_t \varphi \, dx\,dt
+ \iint_{Q} \nabla u : \nabla \varphi \, dx\,dt
- \iint_{Q} (u \otimes u) : \nabla \varphi \, dx\,dt = 0
\end{equation}
holds.
\end{definition}

These solutions exist globally in time and possess the global energy inequality in terms of the initial kinetic energy. 
Such solutions are commonly referred to as \emph{Leray--Hopf weak solutions}.

\smallskip

When studying local and partial regularity of the N--S equations, 
a stronger notion of solution is typically used, the class of 
\emph{suitable weak solutions}. Following Scheffer \cite{Scheffer1977} and Caffarelli, Kohn, and Nirenberg \cite{CKN1982}, we work with the class of suitable weak solutions.  
Here we present a version due to Galdi \cite{6}.

\begin{definition}[Suitable Weak Solution]\label{def:suitable}
Let $u:Q \to \mathbb{R}^3$ and $p:Q \to \mathbb{R}$ be measurable.  
The pair $(u,p)$ is called a \emph{suitable weak solution} of the N--S 
equations \textup{(1.1)} in the cylinder $Q = \Omega \times (a,b)$ if:
\begin{align}
u &\in 
L^\infty\!\left(a,b; L^2(\Omega;\mathbb{R}^3)\right)
\;\cap\;
L^2\!\left(a,b; W^{1,2}(\Omega;\mathbb{R}^3)\right), 
\label{eq:suitable-u}
\\[4pt]
p &\in L^{3/2}(Q), 
\label{eq:suitable-p}
\end{align}
the system \textup{(1.1)} is satisfied in the sense of distributions, and the following
\emph{generalized local energy inequality} holds:

For almost every $t \in (a,b)$ and every non-negative test function 
$\phi \in C_c^\infty(Q)$,
\begin{equation}\label{eq:local-energy-ineq}
\begin{aligned}
\int_{\Omega} |u(t)|^2 \phi(t) \, dx
+ 2 \int_{a}^{t} \!\!\int_{\Omega} |\nabla u|^2 \phi \, dx\,ds
\;\le\;
\int_{a}^{t} \!\!\int_{\Omega} 
u^2 (\partial_t \phi + \Delta \phi)
\, dx\,ds 
\\
+ \int_{a}^{t} \!\!\int_{\Omega} \bigl(|u|^2 + 2p\bigr)\, u \cdot \nabla \phi \, dx\,ds .
\end{aligned}
\end{equation}
\end{definition}





While the Ladyzhenskaya–Prodi–Serrin and endpoint criteria provide global regularity conditions, the local counterpart is given by the $\varepsilon$-regularity theory of Caffarelli–Kohn–Nire\-nberg. 

Standard $\varepsilon$-regularity theory \cite{CKN1982, Lin1998} shows that
smallness of certain scale-invariant quantities on a parabolic cylinder forces
regularity. A fundamental example is the Caffarelli--Kohn--Nirenberg
criterion, based on the dimensionless functional
\[
F(r) := r^{-2}\!\iint_{Q_r(z_0)} \big(|u|^{3} + |p|^{3/2}\big)\,dx\,dt .
\]
There exists a universal constant $\varepsilon_{CKN} > 0$ such that if
\[
F(r) < \varepsilon_{CKN},
\]
then $u$ is bounded (and in fact Hölder continuous) on $Q_{r/2}(z_0)$.
This type of estimate constitutes the first prototype of local
regularity criteria for suitable weak solutions.


\subsection{Blow-up Analysis and Construction of the Running-Max Ancient Element}


 Assume, for contradiction, that the smooth solution develops a finite-time singularity at
$T^* < \infty$. By the Beale–Kato–Majda criterion we know that the vorticity must blow up, so
\[
\limsup_{t \uparrow T^*} \|\omega(\cdot,t)\|_{L^\infty} = \infty.
\]
In order to understand how such a singularity could appear, we rescale the solution near the
points and times where the vorticity is very large, and in this way we obtain a limiting
blow-up profile.



\begin{theorem} [Beale--Kato--Majda (BKM), Euler, \cite{BKM1984}]
Let $u$ be a solution of the incompressible Euler equations (i.e.\ \eqref{eq:NS_domain} with $\nu=0$ and $f=0$), and
suppose there is a time $T^*$ such that the solution cannot be continued in the class $u \in C([0,T]; H^s) \,\cap\, C^1([0,T]; H^{s-1}), \, s \geq 3.$
to $T = T^*$. Assume that $T^*$ is the first such time.
Then
\[
\int_0^{T^*} \|\omega(t)\|_{L^\infty}\, dt = +\infty,
\]
and in particular
\[
\limsup_{t \uparrow T^*} \|\omega(t)\|_{L^\infty} = +\infty.
\]
\end{theorem}


%Lecture notes for Math 256B, Version 2024
%Lenya Ryzhik May 7, 2024
\begin{theorem}[BKM, N-S]\label{thm:BKM-NS}
Let $u_0 \in C^\infty_c(\mathbb{R}^3)$, so that there exists a classical 
solution $u$ to the N-S equations (i.e.\ \eqref{eq:NS_domain} with $f=0$ and viscosity $\nu>0$). 
If for any $T>0$ we have
\begin{equation}\label{eq:BKM-NS-1}
\int_0^T \|\omega(t)\|_{L^\infty}\, dt < +\infty,
\end{equation}
then the smooth solution $u$ exists globally in time.  
If the maximal existence time of the smooth solution is $T < +\infty$, 
then necessarily
\begin{equation}\label{eq:BKM-NS-2}
\int_0^{T} \|\omega(s)\|_{L^\infty}\, ds = +\infty.
\end{equation}
\end{theorem}

\begin{remark}
For the Euler equations the BKM criterion is an equivalence: 
finite--time blow-up occurs if and only if 
$\int_0^{T^*}\|\omega(t)\|_{L^\infty}\,dt=+\infty$. 
For the N-S equations one only has the one--sided continuation 
criterion stated above; the converse implication is not known, nor does it 
hold for weak solutions or suitable weak solutions. 
\end{remark}

The $\varepsilon$--regularity theorem (see Caffarelli--Kohn--Nirenberg \cite{CKN1982})
implies that if no singular point existed at a possible blow\mbox{-}up time $T^{*}$, 
then the solution would remain uniformly bounded in a parabolic neighbourhood of 
the hyperplane $\{t = T^{*}\}$. Combined with the local energy inequality, this 
allows us to extend the solution smoothly past $T^{*}$, contradicting the assumption
that $T^{*}$ is the first blow-up time. F. Lin \cite{Lin1998} later
gave a different proof of this result via a blow-up argument which was expanded upon
and extended by Ladyzhenskaya–Seregin \cite{LG}. The following lemma is a direct consequence of the $\varepsilon$--regularity theory of
Caffarelli–Kohn–Nirenberg (CKN) \cite{CKN1982}.




\begin{remark}[Optional: CKN singular points (not used in the running-max route)]
The running-max/vorticity-normalized construction of the ancient element (Lemmas~\ref{lem:blowup-normalization}--\ref{lem:ancient-limit-runningmax}) does not require a CKN-singular point.
We record the following standard CKN singular-point lemma only to motivate the classical CKN-anchored tangent-flow construction included later for comparison.
\end{remark}

\begin{lemma}\label{lem:singular-point}
Assume that $u$ is a smooth solution of the N-S (\ref{eq:NS_domain}) equations
on $[0,T^*)$ and that $T^*<\infty$ is the first blow-up time.
Then there exists at least one point $x^*\in\R^3$ such that $(x^*,T^*)$ is a singular
point in the sense of CKN.
\end{lemma}


\begin{proof}
Suppose, that no such point exists. Then every $(x,T^*)$ is regular
in the CKN sense. Hence, for each $x\in\R^3$ there exists $r_x>0$ such that 
\[
F(z_0,r) = r^{-2} \iint_{Q_r(z_0)} \bigl(|u|^3 + |p|^{3/2}\bigr)\,dx\,dt
\]
satisfies $F((x,T^*),r_x) < \varepsilon_{\mathrm{CKN}}$.
By the $\varepsilon$-regularity theorem \cite{CKN1982,Lin1998}, this implies that
$u$ is bounded in a smaller parabolic cylinder, there exist constants
$M_x<\infty$ such that
\[
|u(y,s)| \le M_x \quad \text{for all } (y,s)\in
Q_{r_x/2}(x,T^*) = B_{r_x/2}(x)\times(T^*-(r_x/2)^2,T^*].
\]



There exist $R>0$ and consider the compact set $\overline{B_R(0)}\times\{T^*\}$.
Since the balls $B_{r_x/2}(x)$, $x\in\overline{B_R(0)}$, form an open cover of
$\overline{B_R(0)}$, we can extract a finite subcover
\[
\overline{B_R(0)} \subset \bigcup_{i=1}^N B_{r_i/2}(x_i).
\]
Let us define
\[
\delta_R := \min_{1\le i\le N} \frac{r_i^2}{4} > 0,
\qquad
M_R := \max_{1\le i\le N} M_{x_i} < \infty.
\]
Let $(y,s)$ be any point with $|y|\le R$ and $s\in(T^*-\delta_R,T^*]$.
Then there exists $i\in\{1,\dots,N\}$ such that $y\in B_{r_i/2}(x_i)$.
Moreover,  we have
\[
s > T^*-\delta_R \ge T^* - \frac{r_i^2}{4},
\]
so $(y,s)\in Q_{r_i/2}(x_i,T^*)$. Therefore
\[
|u(y,s)| \le M_{x_i} \le M_R.
\]
In other words,
\[
\sup_{|y|\le R,\; s \in (T^*-\delta_R,T^*]} |u(y,s)| \le M_R < \infty.
\]

Thus $u$ is uniformly bounded on $B_R(0)\times(T^*-\delta_R,T^*]$.
Standard local well-posedness and continuation for smooth solutions imply that
$u$ can be smoothly extended beyond $T^*$ on $B_R(0)$.

Since $R>0$ is arbitrary, this shows that $u$ extends smoothly beyond $T^*$ on
all of $\R^3$, contradicting the maximality of $T^*$. Therefore, there exist at least one singular point $(x^*,T^*)$ in the CKN sense.
\end{proof}











\begin{lemma}\label{lem:blowup-normalization}
Let $u_0 \in C_c^\infty(\mathbb{R}^3)$ be divergence-free, and let
$u$ be the unique smooth solution of the N-S equations (\ref{eq:NS_domain})
on its maximal interval of smooth existence $[0,T^*)$. Assume that $T^* < \infty$ is the
first blow-up time.

Then there exist times $t_k \uparrow T^*$, points $x_k \in \mathbb{R}^3$, and scales
$\lambda_k \downarrow 0$ (for instance, $\lambda_k = |\omega(x_k,t_k)|^{-1/2}$) such that,
defining the rescaled velocity fields
\begin{equation}\label{rescaled}
u^{(k)}(y,s)
:=
\lambda_k\, u\!\left(x_k + \lambda_k y,\; t_k + \lambda_k^2 s\right),
\qquad
\;p^{(k)}(y,s)
:=
\lambda_k^2\, p\!\left(x_k + \lambda_k y,\; t_k + \lambda_k^2 s\right),
\qquad
\omega^{(k)} := \curl\, u^{(k)},
\end{equation}
we have the normalization
\[
|\omega^{(k)}(0,0)| = 1 \quad \text{for all } k.
\]
\end{lemma}

\begin{proof}
By the BKM continuation criterion, loss of smoothness at $T^*$ implies that
\[
\limsup_{t \uparrow T^*} \|\omega(\cdot,t)\|_{L^\infty} = +\infty.
\]
Hence we can choose a sequence of times $t_k \uparrow T^*$ such that
\[
M_k := \|\omega(\cdot,t_k)\|_{L^\infty} \to \infty
\quad \text{as } k \to \infty.
\]
One may choose the times $t_k$ so that $\|\omega(\cdot,t)\|_{L^\infty}\le \|\omega(\cdot,t_k)\|_{L^\infty}=M_k$ for all $t\le t_k$
(e.g.\ take $t_k$ to be the first hitting time of a level $L_k\uparrow\infty$). This yields uniform backward-in-time $L^\infty$ control for the rescaled vorticities (see below).
For each $k$, since $\omega(\cdot,t_k)$ is continuous and not identically zero, there exists
a point $x_k \in \mathbb{R}^3$ such that
\[
|\omega(x_k,t_k)| \ge \Bigl(1-\frac1k\Bigr) M_k.
\]
Let us set $
A_k := |\omega(x_k,t_k)|$, then $A_k \ge (1-\frac1k) M_k$, and in particular $A_k \to \infty$ as $k \to \infty$.
Let us define the scaling factors
\[
\lambda_k := A_k^{-1/2}.
\]
Using the rescaling (\ref{rescaled}), by the scaling of the vorticity (\ref{scaling}), we have
\[
\omega^{(k)}(0,0)
= \lambda_k^2\, \omega(x_k,t_k)
= \lambda_k^2 A_k
= 1.
\]
Since $A_k \to \infty$, it follows that $\lambda_k \downarrow 0$.
If the ``running-max'' choice of $t_k$ is made, then for every $s\le 0$ one has $t_k+\lambda_k^2 s\le t_k$ and hence
$\|\omega(\cdot,t_k+\lambda_k^2 s)\|_{L^\infty}\le M_k$.
By scaling this gives the bound
\[
\|\omega^{(k)}(\cdot,s)\|_{L^\infty}\le \frac{M_k}{A_k}\le \Bigl(1-\frac1k\Bigr)^{-1}=: \gamma_k
\qquad\text{for all }s\le 0.
\]
In particular, $\gamma_k\le 2$ for all $k\ge 2$ and $\gamma_k\downarrow 1$ as $k\to\infty$.
In particular, any ancient limit extracted from such a sequence satisfies the scale-critical bound in Lemma~\ref{lem:omega32-runningmax}.
\end{proof}

\begin{lemma}[Running-max vorticity normalization implies a critical \(L^{3/2}\) bound]\label{lem:omega32-runningmax}
Assume the times $t_k\uparrow T^*$ in Lemma~\ref{lem:blowup-normalization} are chosen as \emph{running maxima} for the vorticity:
\[
\|\omega(\cdot,t)\|_{L^\infty}\le \|\omega(\cdot,t_k)\|_{L^\infty}\qquad\text{for all }t\le t_k.
\]
Then the rescaled vorticities $\omega^{(k)}=\curl u^{(k)}$ satisfy the backward-in-time bounds
\[
\|\omega^{(k)}\|_{L^\infty(\R^3\times(-\lambda_k^{-2}t_k,\,0])}\le \gamma_k,
\]
where $\gamma_k:=\frac{M_k}{A_k}\le (1-\frac1k)^{-1}$ and hence $\gamma_k\downarrow 1$.
In particular, any subsequential weak-$\ast$ limit $\omega^\infty$ of $\omega^{(k)}$ in $L^\infty_{\mathrm{loc}}(\R^3\times(-\infty,0])$
obeys the scale-critical estimate (with a universal constant), and satisfies the sharper bound
\[
\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1.
\]
\[
\sup_{z_0\in\R^3\times(-\infty,0]}\ \sup_{0<r\le1}\ r^{-2}\iint_{Q_r(z_0)} |\omega^\infty|^{3/2}\,dx\,dt
\ \le\ C,
\]
where $C>0$ is a universal dimensional constant depending only on the definition of $Q_r$.
\end{lemma}

\begin{proof}
The $L^\infty$ bound on $\omega^{(k)}$ follows from the running-max normalization.
Passing to a subsequence, we may assume $\omega^{(k)}\rightharpoonup^\ast \omega^\infty$ weak-$\ast$ in $L^\infty_{\mathrm{loc}}$ and hence
$\|\omega^\infty\|_{L^\infty_{\mathrm{loc}}}\le \liminf_{k\to\infty}\gamma_k = 1$.
Therefore for any $z_0$ and $0<r\le 1$,
\[
r^{-2}\iint_{Q_r(z_0)} |\omega^\infty|^{3/2}\,dx\,dt
\ \le\ r^{-2}\,\|\omega^\infty\|_{L^\infty(Q_r(z_0))}^{3/2}\,|Q_r|
\ \le\ r^{-2}\,|Q_r|
\ \le\ C,
\]
since $|Q_r|\le C\,r^5$ for $r\le 1$.
\end{proof}

\begin{lemma}\label{lem:domain-rescaled}
Let $u^{(k)}$ be the rescaled sequence defined in \eqref{rescaled}.
Then each $u^{(k)}$ is defined on a time interval of the form
\[
s \in \bigl(-\lambda_k^{-2} t_k,\;0\bigr],
\]
and these intervals exhaust $(-\infty,0]$. It means that for every $R>0$ there exists
$k_0(R)$ such that
\[
(-R^2,0] \subset \bigl(-\lambda_k^{-2} t_k,\;0\bigr]
\quad\text{for all } k \ge k_0(R).
\]
\end{lemma}

\begin{proof} 
The original solution $u$ is defined for $0 \le t < T^*$. Since $u^{(k)}$ be the rescaled by (\ref{rescaled}), for $u^{(k)}$ to be
well-defined at time $s$, we need
\[
0 \le t_k + \lambda_k^2 s < T^*.
\]
The upper bound $t_k + \lambda_k^2 s \le t_k$ corresponds exactly to $s \le 0$.
The lower bound $t_k + \lambda_k^2 s \ge 0$ gives
\[
s \ge -\lambda_k^{-2} t_k.
\]
Hence $u^{(k)}$ is defined on $s \in (-\lambda_k^{-2} t_k,0]$.

Since $t_k \uparrow T^*$ and $\lambda_k \downarrow 0$, we have
$\lambda_k^{-2} t_k \to \infty$ as $k\to\infty$. Therefore, for any fixed $R>0$,
we can choose $k_0(R)$ such that $\lambda_k^{-2} t_k > R^2$ for all $k\ge k_0(R)$.
Finally, for  $k\ge k_0(R)$, we obtain
\[
(-R^2,0] \subset (-\lambda_k^{-2} t_k,0].,
\]
which proves the lemma.
\end{proof}

Lemmas~\ref{lem:blowup-normalization}--\ref{lem:domain-rescaled} construct a \emph{vorticity-normalized} rescaling sequence.
In the running-max variant (choose $t_k$ as running maxima for $\|\omega(\cdot,t)\|_{L^\infty}$), one can extract an ancient limit along this sequence; see Lemma~\ref{lem:ancient-limit-runningmax}.
The CKN-anchored tangent flow of Lemma~\ref{lem:ancient-limit} is retained below for comparison, but the main contradiction chain in this rewrite uses Lemma~\ref{lem:ancient-limit-runningmax}.

\begin{lemma}[Running-max vorticity-normalized ancient element]\label{lem:ancient-limit-runningmax}
Assume the times $t_k\uparrow T^*$ in Lemma~\ref{lem:blowup-normalization} are chosen as \emph{running maxima} for the vorticity (as in Lemma~\ref{lem:omega32-runningmax}), and let $u^{(k)}$ be the corresponding rescaled sequence \eqref{rescaled}.
Then there exists a subsequence (still denoted by $u^{(k)}$) and a pair $(u^\infty,p^\infty)$ such that:
\begin{enumerate}
\item[(i)] For every $R>0$ and $T>0$,
\[
u^{(k)} \to u^\infty \quad\text{strongly in }
L^p(B_R\times(-T,0)) \quad \text{for all } 1\le p<3,
\]
and
\[
u^{(k)} \rightharpoonup u^\infty
\quad \text{weakly in}\quad
L^3_{\mathrm{loc}}(\R^3\times(-\infty,0)).
\]
Moreover,
\[
p^{(k)} \rightharpoonup p^\infty
\quad\text{weakly in } L^{3/2}_{\mathrm{loc}}(\R^3\times(-\infty,0)).
\]
\item[(ii)] The limit $(u^\infty,p^\infty)$ is a suitable weak solution of the N--S equations on $\R^3\times(-\infty,0)$ and satisfies the local energy inequality on every cylinder $B_R\times(-T,0)$.
\item[(iii)] Writing $\omega^\infty=\curl u^\infty$, one has
\[
|\omega^\infty(0,0)|=1,
\qquad
\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1.
\]
In particular, $u^\infty\not\equiv 0$.
\end{enumerate}
We call $u^\infty$ the \emph{running-max ancient element} associated to the blow-up at time $T^*$.
\end{lemma}


\begin{proof}
By Lemma~\ref{lem:domain-rescaled}, for each fixed $R>0$ and $T>0$ the rescaled solutions are well-defined and smooth on $B_R\times(-T,0)$ for all $k$ sufficiently large.

\smallskip
\noindent\textbf{Step 1: Uniform $L^\infty$ vorticity bound.}
By the running-max construction (Lemma~\ref{lem:omega32-runningmax}), for each $k$ we have
\begin{equation}\label{eq:unif-vort-bound}
\|\omega^{(k)}\|_{L^\infty(\R^3\times(-\infty,0])}\ \le\ \gamma_k\ :=\ (1-1/k)^{-1}\ \le\ 2\qquad\text{for }k\ge 2.
\end{equation}
This is the key input that distinguishes the running-max blow-up from generic blow-up sequences.

\smallskip
\noindent\textbf{Step 2: Uniform local energy bounds on cylinders (RM2 gate).}
Fix $R>0$ and $T>0$, and let $Q:=B_{2R}\times(-T-1,0)$. We claim:
\begin{equation}\label{eq:unif-energy-bound}
\sup_{s\in(-T,0)}\int_{B_R}|u^{(k)}(x,s)|^2\,dx\ +\ \iint_{Q_R}|\nabla u^{(k)}|^2\,dx\,ds\ \le\ C(R,T),
\end{equation}
with $C(R,T)$ independent of $k$.
\noindent\textbf{[RM2 note.]}
This fixed-frame local energy compactness is the additional P0 gate isolated as Assumption~\ref{assump:RM2-runningmax}. The discussion below is a proof sketch explaining why bounded vorticity alone does not close it (affine/harmonic-mode obstruction).

\smallskip
\noindent\emph{Derivation.}
Since $u^{(k)}$ is a smooth solution of N--S on $Q$, the local energy inequality holds:
\[
\sup_s\int_{B_R}|u^{(k)}|^2\phi^2\,dx + 2\iint_Q |\nabla u^{(k)}|^2\phi^2
\ \le\ \iint_Q |u^{(k)}|^2\bigl(\partial_t\phi^2 + \Delta\phi^2\bigr) + \iint_Q (|u^{(k)}|^2 + 2p^{(k)})u^{(k)}\cdot\nabla\phi^2
\]
for any non-negative $\phi\in C^\infty_c(Q)$.
Take $\phi\equiv 1$ on $B_R\times(-T,0)$ supported in $B_{2R}\times(-T-1,0)$ with $|\nabla\phi|\le C/R$, $|\partial_t\phi|,|\Delta\phi|\le C$.

The argument above previously used the false implication
$\|\nabla u(\cdot,t)\|_{L^\infty}\lesssim \|\omega(\cdot,t)\|_{L^\infty}$ in 3D.
What is classical from bounded vorticity is only the borderline Calder\'on--Zygmund estimate
$\nabla u(\cdot,t)\in \BMO$ with $\|\nabla u(\cdot,t)\|_{\BMO}\lesssim \|\omega(\cdot,t)\|_{L^\infty}$
(see Lemma~\ref{lem:drift-bmo-from-vorticity}).
John--Nirenberg then yields local $L^p$ control of oscillations of $\nabla u$ on balls, and after subtracting a divergence-free affine approximation one obtains the referee-checkable local $L^p$ drift bound
(Lemma~\ref{lem:drift-local-Lp-affine}).

\smallskip
\noindent
To turn this into the \emph{fixed-frame} $k$-uniform local energy bound \eqref{eq:unif-energy-bound} via the local energy inequality, one must additionally control the spatially constant / affine velocity mode (equivalently: control the ball average and mean strain of $u^{(k)}$ on $B_{2R}$ uniformly in $k$).
This is the same global harmonic/affine-mode obstruction isolated elsewhere (cf.\ Remark~\ref{rem:BS-gauge-justification} and Lemma~\ref{lem:finite-capacity-no-affine}).
Accordingly, Step 2 should be viewed as a proof sketch unless one supplies an independent bound excluding large affine modes (e.g. a finite-capacity bound, or a direct Biot--Savart gauge preservation argument).

Assuming one has the local energy / local $L^3$ control implicit in \eqref{eq:unif-energy-bound}, the Poisson equation
$-\Delta p^{(k)} = \partial_i\partial_j(u^{(k)}_i u^{(k)}_j)$ and standard Calder\'on--Zygmund estimates yield
\begin{equation}\label{eq:unif-pressure-bound}
\|p^{(k)}\|_{L^{3/2}(Q_R)}\ \le\ C(R,T).
\end{equation}

\smallskip
\noindent\textbf{Step 3: Time derivative bound and Aubin--Lions compactness.}
The N--S momentum equation gives
\[
\partial_t u^{(k)}\ =\ \nu\Delta u^{(k)} - (u^{(k)}\cdot\nabla)u^{(k)} - \nabla p^{(k)}.
\]
Using \eqref{eq:unif-energy-bound} and \eqref{eq:unif-pressure-bound}:
\begin{equation}\label{eq:time-deriv-bound}
\|\partial_t u^{(k)}\|_{L^{3/2}((-T,0); W^{-1,3/2}(B_R))}\ \le\ C(R,T).
\end{equation}
By the Aubin--Lions lemma (with $W^{1,2}(B_R)\hookrightarrow\hookrightarrow L^2(B_R)\hookrightarrow W^{-1,3/2}(B_R)$), the sequence $\{u^{(k)}\}$ is precompact in $L^2(Q_R)$.
Extract a subsequence with $u^{(k)}\to u^\infty$ strongly in $L^2(Q_R)$.
By interpolation with the $L^\infty$ bound, convergence holds in $L^p(Q_R)$ for all $p<\infty$.
A diagonal argument over $R_n\uparrow\infty$, $T_n\uparrow\infty$ yields convergence on all of $\R^3\times(-\infty,0)$.

Weak compactness in $L^3_{\mathrm{loc}}$ and $L^{3/2}_{\mathrm{loc}}$ for velocity and pressure gives the weak limits in (i).

\smallskip
\noindent\textbf{Step 4: Passage to the limit and suitability.}
The strong $L^p_{\mathrm{loc}}$ convergence for $p<\infty$ allows passage to the limit in the distributional form of N--S:
\[
\partial_t u^\infty + (u^\infty\cdot\nabla)u^\infty + \nabla p^\infty\ =\ \nu\Delta u^\infty,\qquad \nabla\cdot u^\infty = 0.
\]
For suitability: the local energy inequality for each $u^{(k)}$ is
\[
\int |u^{(k)}|^2\phi\,dx\Big|_{t=s} + 2\int_0^s\!\!\int |\nabla u^{(k)}|^2\phi
\ \le\ \int_0^s\!\!\int |u^{(k)}|^2(\partial_t\phi+\Delta\phi) + (|u^{(k)}|^2+2p^{(k)})u^{(k)}\cdot\nabla\phi
\]
for non-negative test functions $\phi$. Passing to the limit:
\begin{itemize}
\item The left-hand side is lower semicontinuous under strong $L^2$ and weak $H^1$ convergence.
\item The right-hand side converges by strong $L^p$ convergence and weak pressure convergence.
\end{itemize}
Thus $(u^\infty,p^\infty)$ satisfies the local energy inequality on every cylinder, proving (ii).

\smallskip
\noindent\textbf{Step 5: Nontriviality and vorticity bound.}
\emph{(a) $L^\infty$ vorticity bound:} From \eqref{eq:unif-vort-bound} and weak-$\ast$ compactness in $L^\infty$:
\[
\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\ \le\ \liminf_{k\to\infty}\gamma_k\ =\ 1.
\]

\emph{(b) Pointwise nontriviality at $(0,0)$:}
We need $|\omega^\infty(0,0)|=1$. By construction, $|\omega^{(k)}(0,0)|=1$ for all $k$.
The uniform $L^\infty$ vorticity bound \eqref{eq:unif-vort-bound} does \emph{not} by itself yield a $k$-uniform $C^{0,\alpha}_{\mathrm{loc}}$ estimate for $\omega^{(k)}$ without additional control of the drift/stretching coefficients in the vorticity equation. In particular, a referee-checkable uniform H\"older bound on $\omega^{(k)}$ on fixed cylinders can be obtained once one has the fixed-frame local energy/pressure compactness from Step 2--3 (or any substitute that yields the needed local Serrin-type control of $u^{(k)}$ and Calder\'on--Zygmund control of $\nabla u^{(k)}$ on the cylinder).
Assuming such a uniform local regularity input, interior parabolic regularity for the vorticity equation yields uniform $C^{0,\alpha}_{\mathrm{loc}}$ H\"older continuity for $\omega^{(k)}$ on compact cylinders (by interior parabolic regularity for the vorticity equation).
Specifically, for any $\alpha\in(0,1)$:
\[
\|\omega^{(k)}\|_{C^{0,\alpha}(B_1\times(-1,0])}\ \le\ C_\alpha,
\]
with $C_\alpha$ depending on the $L^\infty$ vorticity bound but not on $k$.
By Arzel\`a--Ascoli, a subsequence converges in $C^0(B_1\times(-1,0])$, so in particular $\omega^{(k)}(0,0)\to\omega^\infty(0,0)$.
Therefore $|\omega^\infty(0,0)|=\lim_k|\omega^{(k)}(0,0)|=1$.

This completes the proof of (iii) and the lemma.
\end{proof}

\begin{lemma}[Running-max freezes the vorticity supremum]\label{lem:runningmax-sup-freeze-3d}
Let $(u^\infty,p^\infty)$ be the running-max ancient element from Lemma~\ref{lem:ancient-limit-runningmax}, and write $\omega^\infty=\curl u^\infty$, $\rho^\infty:=|\omega^\infty|$.
Then for every $t\le 0$,
\[
\sup_{x\in\R^3}\rho^\infty(x,t)\ =\ 1.
\]
\end{lemma}

\begin{proof}[Proof sketch]
Lemma~\ref{lem:ancient-limit-runningmax}(iii) gives $\sup_x\rho^\infty(x,t)\le 1$ for all $t\le 0$.
To obtain the reverse inequality, fix $t<0$ and consider the running-max rescaled sequence $u^{(k)}$ from Lemma~\ref{lem:ancient-limit-runningmax}.
By construction, the rescaled vorticity satisfies
\[
\omega^{(k)}(0,0)=1,
\qquad
\omega^{(k)}(0,t)\ =\ \frac{1}{M_k}\,\omega(x_k,\ t_k+t/M_k),
\]
where $M_k:=\|\omega(\cdot,t_k)\|_{L^\infty}$ and $x_k$ is chosen with $|\omega(x_k,t_k)|=M_k$.
Since the pre-blow-up solution is smooth on $[0,T^*)$, the map $s\mapsto \|\omega(\cdot,s)\|_{L^\infty}$ is continuous, hence
\[
\frac{\|\omega(\cdot,t_k+t/M_k)\|_{L^\infty}}{M_k}\ \to\ 1
\qquad\text{as }k\to\infty
\]
for each fixed $t<0$ (because $t/M_k\to 0$).
In particular, $|\omega(x_k,t_k+t/M_k)|/M_k\to 1$, so $|\omega^{(k)}(0,t)|\to 1$.
Passing to the limit along the subsequence in Lemma~\ref{lem:ancient-limit-runningmax}(i) yields $|\omega^\infty(0,t)|=1$, and hence $\sup_x\rho^\infty(x,t)\ge 1$.
Combining with $\sup_x\rho^\infty(x,t)\le 1$ gives the claim.

This passage relies on the local compactness asserted in Lemma~\ref{lem:ancient-limit-runningmax}.
As noted in the correction inside Step 2 of that proof, obtaining the required $k$-uniform local compactness from the sole vorticity bound involves a nontrivial control of the affine/harmonic velocity mode.
Accordingly, the $C^{0,\alpha}_{\mathrm{loc}}$ convergence claim here should be read as conditional on closing that Step 2 gate.
\end{proof}

\begin{remark}[Running-max constraint on stretching injection at the top vorticity level]\label{rem:runningmax-injection-constraint}
Lemma~\ref{lem:runningmax-sup-freeze-3d} is the precise classical form of the running-max ``finite budget'' constraint: the vorticity amplitude never exceeds the normalized budget $1$ at any time $t\le 0$.
Consequently, at any time $t$ and any point $x_t$ where $\rho^\infty(\cdot,t)$ attains its supremum $1$, one has $\partial_t\rho^\infty(x_t,t)\le 0$ (otherwise $\sup_x\rho^\infty(\cdot,t)$ would increase above $1$ for slightly later times).
Evaluating the amplitude equation \eqref{eq:amplitude} at such a maximum point (where $\nabla\rho=0$ and $\Delta\rho\le 0$) gives the pointwise constraint
\[
\sigma(x_t,t)\ \le\ |\nabla\xi(x_t,t)|^2\ -\ \Delta\rho(x_t,t).
\]
Thus, \emph{positive stretching at the top vorticity level} can only occur if it is paid for by either:
\begin{itemize}
\item large direction-coherence cost $|\nabla\xi|^2$ (the RS ``recognition strain''), or
\item strong concavity $-\Delta\rho$ (a sharp spatial peak in vorticity magnitude).
\end{itemize}
This is the most direct ``next inch'' constraint toward C2: persistent positive injection $\rho^{3/2}\sigma$ in regions where $\rho\approx 1$ is not free; it must be balanced by a compensating cost.
Upgrading this pointwise constraint into a uniform \emph{integral} control of $\iint\rho^{3/2}\sigma_+$ (and hence of $\mathcal E_\omega$) is achieved in Theorem~\ref{thm:C2-closure} via the $\sigma$-decomposition, conditional on the global spatial-decay gate (Assumption~\ref{assump:U-decay}).
\end{remark}

\begin{lemma}[Quantitative thick maximum at a running-max time slice]\label{lem:thick-maximum}
Let $t\le 0$ and let $\rho(\cdot,t):\R^3\to[0,1]$ be $C^2$ with
\[
\sup_{x\in\R^3}\rho(x,t)=1
\]
attained at some point $x_t$ (so $\rho(x_t,t)=1$ and $\nabla\rho(x_t,t)=0$).
Set
\[
A(t):=-\Delta\rho(x_t,t)\ \ge\ 0.
\]
Then for every $\eta\in(0,\tfrac14]$ there exists a radius $r_\eta(t)>0$ such that
\begin{equation}\label{eq:thick-max}
\rho(x,t)\ \ge\ 1-\eta\qquad\text{for all }x\in B_{r_\eta(t)}(x_t),
\end{equation}
and
\begin{equation}\label{eq:thick-max-radius}
r_\eta(t)\ \ge\ c\,\sqrt{\frac{\eta}{A(t)+1}},
\end{equation}
where $c\in(0,1)$ is a universal dimensional constant.
In particular, the superlevel set $\{\rho(\cdot,t)\ge 1-\eta\}$ has nontrivial measure:
\[
\bigl|\{x\in\R^3:\rho(x,t)\ge 1-\eta\}\cap B_{r_\eta(t)}(x_t)\bigr|
\ =\ |B_{r_\eta(t)}|.
\]
\end{lemma}

\begin{proof}[Proof sketch]
Since $\rho(\cdot,t)$ is $C^2$, its Hessian $D^2\rho(\cdot,t)$ is continuous.
At the maximizer $x_t$, $D^2\rho(x_t,t)$ is negative semidefinite, so $\|D^2\rho(x_t,t)\|_{\mathrm{op}}\le A(t)$.
By continuity, there exists $r_0(t)>0$ such that
\[
\sup_{x\in B_{r_0(t)}(x_t)}\|D^2\rho(x,t)\|_{\mathrm{op}}\ \le\ 2(A(t)+1).
\]
Then for any $x\in B_{r_0(t)}(x_t)$, Taylor's theorem with remainder gives
\[
\rho(x,t)\ \ge\ \rho(x_t,t)\ -\ (A(t)+1)\,|x-x_t|^2
\ =\ 1-(A(t)+1)|x-x_t|^2.
\]
Choose $r_\eta(t):=\min\bigl\{r_0(t),\,\sqrt{\eta/(A(t)+1)}\bigr\}$ to obtain \eqref{eq:thick-max}.
The lower bound \eqref{eq:thick-max-radius} follows by taking $c:=\min\{1,\,r_0(t)\sqrt{A(t)+1}\}^{-1}$ and noting that for fixed $t$ one has $r_0(t)\sqrt{A(t)+1}>0$.

The lemma is a purely local $C^2$ fact at each fixed time $t$.
For the running-max ancient element, local smoothness on compact cylinders ensures such an $r_0(t)$ exists at each time, but obtaining a \emph{uniform} radius bound in $t$ (or uniformly over all maximizers) would require additional global controls on higher derivatives.
\end{proof}

\begin{lemma}[From max-point stretching to a positive-measure injection region]\label{lem:maxpoint-to-injection-region}
In the setting of Lemma~\ref{lem:thick-maximum}, assume in addition that the stretching scalar $\sigma(\cdot,t)$ is $C^1$ in a neighborhood of $x_t$ and that
\[
\sigma(x_t,t)\ \ge\ s_0
\qquad\text{for some }s_0>0.
\]
Fix $\eta\in(0,\tfrac14]$ and let $r_\eta(t)$ be the radius from Lemma~\ref{lem:thick-maximum} so that $\rho(\cdot,t)\ge 1-\eta$ on $B_{r_\eta(t)}(x_t)$.
Let
\[
L_\eta(t)\ :=\ \|\nabla\sigma(\cdot,t)\|_{L^\infty(B_{r_\eta(t)}(x_t))}.
\]
Then for
\[
r_*(t)\ :=\ \min\Bigl\{r_\eta(t),\ \frac{s_0}{2(L_\eta(t)+1)}\Bigr\},
\]
one has
\[
\rho(\cdot,t)\ \ge\ 1-\eta
\quad\text{and}\quad
\sigma(\cdot,t)\ \ge\ s_0/2
\qquad\text{on }B_{r_*(t)}(x_t),
\]
and hence the time-slice weighted stretching obeys the quantitative lower bound
\[
\int_{B_{r_*(t)}(x_t)} \rho(x,t)^{3/2}\,\sigma_+(x,t)\,dx
\ \ge\ (1-\eta)^{3/2}\,\frac{s_0}{2}\,|B_{r_*(t)}|.
\]
\end{lemma}

\begin{proof}
The $\rho$ bound on $B_{r_*(t)}$ is immediate since $r_*(t)\le r_\eta(t)$.
For $\sigma$, if $x\in B_{r_*(t)}(x_t)$ then by the mean-value theorem
\[
\sigma(x,t)\ \ge\ \sigma(x_t,t)-L_\eta(t)\,|x-x_t|
\ \ge\ s_0 - L_\eta(t)\,r_*(t)
\ \ge\ s_0/2,
\]
by the definition of $r_*(t)$.
The integral lower bound follows since $\rho^{3/2}\ge (1-\eta)^{3/2}$ and $\sigma_+\ge s_0/2$ on $B_{r_*(t)}(x_t)$.
\end{proof}

\begin{remark}[From pointwise cap to integral cost (what this lemma enables)]\label{rem:thick-max-to-integral}
Lemma~\ref{lem:thick-maximum} is the first step in turning the running-max cap into an \emph{integral} statement.
If, on a time interval $I\subset(-\infty,0]$, the stretching injection at maximizers were persistently positive, e.g.\ $\sigma(x_t,t)\ge c_0>0$ for $t\in I$, then by continuity there would exist space-time cylinders $Q_{r_\eta(t)}(x_t,t)$ on which $\rho\ge 1-\eta$ and $\sigma\ge c_0/2$ on a nontrivial subset.
On such cylinders, the $\rho^{3/2}$ identity \eqref{eq:rho32} forces a comparable amount of \emph{damping} through $\rho^{3/2}|\nabla\xi|^2$ (and/or $\nabla\rho^{3/4}$), which is exactly the C2 ``integral cost'' mechanism suggested by the RS ``finite budget over infinite history'' intuition.

\smallskip
\noindent
This uniform and quantitative control is achieved by Theorem~\ref{thm:C2-closure}: the $\sigma$-decomposition (Lemma~\ref{lem:sigma-decomposition}) combined with the global tail/decay gate (Assumption~\ref{assump:U-decay}; equivalently a relative tail depletion bound along the blow-up sequence as in Lemma~\ref{lem:decay-inheritance}) shows that $\iint\rho^{3/2}\sigma_+\to 0$ at small scales.
\end{remark}

\begin{lemma}[Spacetime lower bound from persistent max-point stretching (conditional quantitative form)]\label{lem:spacetime-injection-lower}
Let $(u^\infty,p^\infty)$ be the running-max ancient element and write $\rho=|\omega^\infty|$, $\xi=\omega^\infty/|\omega^\infty|$ on $\{\rho>0\}$.
Fix a time interval $I=[t_1,t_2]\subset(-\infty,0]$.
Assume that for each $t\in I$ there exists a maximizer $x_t$ with $\rho(x_t,t)=1$ such that
\[
\sigma(x_t,t)\ \ge\ s_0
\qquad\text{for some fixed }s_0>0.
\]
Assume moreover that there exist \emph{uniform} bounds on the local curvature and local $\sigma$-Lipschitz modulus at these maximizers:
\[
-\Delta\rho(x_t,t)\ \le\ A_0,
\qquad
\|\nabla\sigma(\cdot,t)\|_{L^\infty(B_{r_\eta}(x_t))}\ \le\ L_0
\qquad\text{for all }t\in I,
\]
where $r_\eta=r_\eta(t)$ is as in Lemma~\ref{lem:thick-maximum} for some fixed $\eta\in(0,1/4]$.
Then there exists a radius $r_*>0$ (depending only on $\eta,s_0,A_0,L_0$) such that for all $t\in I$,
\[
\rho(\cdot,t)\ge 1-\eta
\quad\text{and}\quad
\sigma(\cdot,t)\ge s_0/2
\qquad\text{on }B_{r_*}(x_t),
\]
and consequently one has the spacetime lower bound
\[
\int_{t_1}^{t_2}\int_{\R^3}\rho^{3/2}(x,t)\,\sigma_+(x,t)\,dx\,dt
\ \ge\ (t_2-t_1)\,(1-\eta)^{3/2}\,\frac{s_0}{2}\,|B_{r_*}|.
\]
\end{lemma}

\begin{proof}[Proof sketch]
By Lemma~\ref{lem:thick-maximum} and the uniform Laplacian bound $-\Delta\rho(x_t,t)\le A_0$, one may take $r_\eta(t)\ge c\sqrt{\eta/(A_0+1)}$.
By Lemma~\ref{lem:maxpoint-to-injection-region} and the uniform Lipschitz bound $L_\eta(t)\le L_0$, one may choose
\[
r_*:=\min\Bigl\{c\sqrt{\frac{\eta}{A_0+1}},\ \frac{s_0}{2(L_0+1)}\Bigr\},
\]
which is independent of $t\in I$.
Then the stated pointwise bounds hold on $B_{r_*}(x_t)$ for every $t\in I$, and the integral estimate follows by integrating the time-slice lower bound from Lemma~\ref{lem:maxpoint-to-injection-region}.
\end{proof}

\begin{remark}[Status of Lemma~\ref{lem:spacetime-injection-lower} (conditional)]\label{rem:spacetime-injection-lower-status}
Lemma~\ref{lem:spacetime-injection-lower} formalizes the intended ``continuity + thick maximum'' mechanism, but it is \emph{conditional} on uniform control of the curvature $-\Delta\rho$ and the spatial Lipschitz modulus of $\sigma$ along the moving set of maximizers $x_t$.
For the running-max ancient element we have the global cap $0\le\rho\le 1$ and local smoothness on each compact cylinder, but the manuscript does not currently supply \emph{uniform} bounds on $\nabla\sigma$ (or on $-\Delta\rho$) \emph{uniformly in $t$ and in the maximizer location}.
Without such uniformity, ``$\sigma$ positive at maximizers infinitely often'' could occur on a set of times of arbitrarily small measure and/or with arbitrarily small spatial neighborhoods, yielding no quantitative spacetime lower bound.
\end{remark}

\begin{lemma}[Injection--damping balance from the $\rho^{3/2}$ identity (localized)]\label{lem:injection-damping-balance}
Let $\rho=|\omega|$ and $\xi=\omega/|\omega|$ on $\{\rho>0\}$ for a smooth Navier--Stokes solution on $Q_{2r}(z_0)$.
Let $\phi\in C_c^\infty(Q_{2r}(z_0))$ satisfy $\phi\equiv 1$ on $Q_r(z_0)$ and $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
Then the $\rho^{3/2}$ identity \eqref{eq:rho32} implies the estimate
\begin{align}\label{eq:inj-damp}
\frac{3}{2}\iint_{Q_r(z_0)}\rho^{3/2}|\nabla\xi|^2
\;+\;\frac{4}{3}\iint_{Q_r(z_0)}|\nabla(\rho^{3/4})|^2
\ \ge\ \frac{3}{2}\iint_{Q_r(z_0)}\rho^{3/2}\,\sigma\,dx\,dt
\ -\ C\,r^{-2}\iint_{Q_{2r}(z_0)}\rho^{3/2}
\ -\ C\,\sup_{t\in(t_0-(2r)^2,t_0)}\int_{B_{2r}(x_0)}\rho^{3/2}(\cdot,t),
\end{align}
with a universal constant $C$.
\end{lemma}

\begin{proof}[Proof sketch]
Multiply \eqref{eq:rho32} by $\phi^2$ and integrate over $Q_{2r}(z_0)$.
Integrate by parts in time and space for the transport/diffusion terms (using $\nabla\cdot u=0$), and use the cutoff bounds
$|\partial_t\phi|\lesssim r^{-2}$, $|\Delta(\phi^2)|\lesssim r^{-2}$.
The $\sigma$ term yields $\iint\rho^{3/2}\sigma\,\phi^2$.
Move the damping terms $\rho^{3/2}|\nabla\xi|^2$ and $|\nabla(\rho^{3/4})|^2$ to the left-hand side; since $\phi\equiv 1$ on $Q_r(z_0)$, the left-hand side dominates the corresponding integrals over $Q_r(z_0)$.
The remaining drift/diffusion/time terms are cutoff/time-boundary errors; bounding them in absolute value by
$Cr^{-2}\iint_{Q_{2r}}\rho^{3/2}$ and $C\sup_t\int_{B_{2r}}\rho^{3/2}$ yields \eqref{eq:inj-damp}.
\end{proof}

\begin{corollary}[Positive injection forces positive damping on a cylinder]\label{cor:positive-injection-forces-damping}
In the setting of Lemma~\ref{lem:injection-damping-balance}, suppose moreover that $\sigma\ge 0$ on $Q_r(z_0)$.
Then
\[
\iint_{Q_r(z_0)}\rho^{3/2}|\nabla\xi|^2
\ \ge\ c\iint_{Q_r(z_0)}\rho^{3/2}\sigma
\ -\ C\Bigl(r^{-2}\iint_{Q_{2r}(z_0)}\rho^{3/2}
\ +\ \sup_{t\in(t_0-(2r)^2,t_0)}\int_{B_{2r}(x_0)}\rho^{3/2}(\cdot,t)\Bigr),
\]
with universal constants $c,C>0$.
In particular, if the signed injection $\iint_{Q_r}\rho^{3/2}\sigma$ dominates the cutoff/time-boundary errors, then the weighted direction-coherence cost $\iint_{Q_r}\rho^{3/2}|\nabla\xi|^2$ is quantitatively positive.
\end{corollary}

\begin{proof}
Under $\sigma\ge 0$, the right-hand side of \eqref{eq:inj-damp} is bounded below by the injection integral on $Q_r(z_0)$ minus the error terms.
Dropping the nonnegative term $\iint_{Q_r}|\nabla(\rho^{3/4})|^2$ and dividing by $\tfrac32$ yields the claim.
\end{proof}

\begin{lemma}[Superlevel-set selection for the weighted injection (no $\nabla\sigma$)]\label{lem:superlevel-selection}
Fix $\eta\in(0,1/4]$.
Let $\chi_\eta:[0,1]\to[0,1]$ be a Lipschitz cutoff satisfying
\[
\chi_\eta(s)=0\ \text{for }s\le 1-2\eta,\qquad
\chi_\eta(s)=1\ \text{for }s\ge 1-\eta,\qquad
0\le \chi_\eta'(s)\le \frac{2}{\eta}.
\]
Let $(u,p)$ be a smooth Navier--Stokes solution on $Q_{2r}(z_0)$, with $\omega=\rho\xi$ on $\{\rho>0\}$ and $\sigma=(S\xi\cdot\xi)$.
Let $\phi\in C_c^\infty(Q_{2r}(z_0))$ satisfy $\phi\equiv 1$ on $Q_r(z_0)$ and $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
Then
\begin{align}\label{eq:superlevel-injection}
\iint_{Q_r(z_0)\cap\{\rho\ge 1-\eta\}} \rho^{3/2}\,\sigma\,dx\,dt
\ \le\ &
\iint_{Q_{2r}(z_0)} \rho^{3/2}\,|\nabla\xi|^2\,\chi_\eta(\rho)\,\phi^2\,dx\,dt
\ +\ \iint_{Q_{2r}(z_0)} |\nabla(\rho^{3/4})|^2\,\phi^2\,dx\,dt \\
&\ +\ C\,\eta^{-1}\iint_{Q_{2r}(z_0)\cap\{1-2\eta<\rho<1-\eta\}} |\nabla(\rho^{3/4})|^2\,dx\,dt
\ +\ C\,r^{-2}\iint_{Q_{2r}(z_0)} \rho^{3/2}\,dx\,dt \nonumber\\
&\ +\ C\,\sup_{t\in(t_0-(2r)^2,t_0)}\int_{B_{2r}(x_0)}\rho^{3/2}(\cdot,t)\,dx,
\nonumber
\end{align}
with a universal constant $C$.
\end{lemma}

\begin{proof}[Proof sketch]
Multiply the $\rho^{3/2}$ identity \eqref{eq:rho32} by $\chi_\eta(\rho)\phi^2$ and integrate over $Q_{2r}(z_0)$.
The right-hand side contains
\(
\frac{3}{2}\iint \rho^{3/2}\sigma\,\chi_\eta(\rho)\phi^2
\)
and
\(
-\frac{3}{2}\iint \rho^{3/2}|\nabla\xi|^2\,\chi_\eta(\rho)\phi^2.
\)
On the left-hand side, integration by parts of the diffusion term produces (besides cutoff terms) a positive contribution involving $\chi_\eta'(\rho)\,|\nabla(\rho^{3/2})|^2$.
Since $\chi_\eta'$ is supported on $\{1-2\eta<\rho<1-\eta\}$ and $|\chi_\eta'|\lesssim \eta^{-1}$, and since
$|\nabla(\rho^{3/2})|^2\lesssim |\nabla(\rho^{3/4})|^2$ for $\rho\in[0,1]$, this yields the band term in \eqref{eq:superlevel-injection}.
All remaining transport/time/cutoff contributions are bounded in absolute value by the standard $r^{-2}\iint\rho^{3/2}$ and time-boundary terms (as in Lemma~\ref{lem:injection-damping-balance}).
Finally, since $\chi_\eta(\rho)\equiv 1$ on $\{\rho\ge 1-\eta\}$ and $\phi\equiv 1$ on $Q_r(z_0)$, we obtain \eqref{eq:superlevel-injection}.
\end{proof}

\begin{corollary}[Superlevel-set selection with simplified band payment (running-max)]\label{cor:superlevel-selection-simplified}
In the setting of Lemma~\ref{lem:superlevel-selection}, assume additionally that $(u,p)=(u^\infty,p^\infty)$ is the running-max ancient element, so that $0\le \rho\le 1$ on $\R^3\times(-\infty,0]$.
Then for every $\eta\in(0,1/4]$ and every $0<r\le 1$,
\begin{align}\label{eq:superlevel-injection-simplified}
\iint_{Q_r(z_0)\cap\{\rho\ge 1-\eta\}} \rho^{3/2}\,\sigma\,dx\,dt
\ \le\ &
\mathcal E_\omega(z_0,2r)
\ +\ \iint_{Q_{2r}(z_0)} |\nabla(\rho^{3/4})|^2\,dx\,dt
\ +\ C_\eta\,r^3 \nonumber\\
&\ +\ C_\eta\iint_{Q_{4r}(z_0)}\bigl(|\sigma|+|\nabla\xi|^2\bigr)\,dx\,dt,
\end{align}
where $\mathcal E_\omega(z_0,2r)=\iint_{Q_{2r}(z_0)}\rho^{3/2}|\nabla\xi|^2\,dx\,dt$ and $C_\eta$ depends only on $\eta$.
\end{corollary}

\begin{proof}[Proof sketch]
Start from \eqref{eq:superlevel-injection}.
The time-boundary and cutoff terms are $O(r^3)$ since $\rho^{3/2}\le 1$ (Remark~\ref{rem:rho32-errors-small-scale}).
The damping term is bounded by $\mathcal E_\omega(z_0,2r)$ since $\chi_\eta(\rho)\phi^2\le 1$.
For the band term, apply Corollary~\ref{cor:band-payment-simplified} (which yields $\eta^{-1}\iint_{\mathrm{band}}|\nabla(\rho^{3/4})|^2\lesssim_\eta r^3+\iint_{Q_{4r}}(|\sigma|+|\nabla\xi|^2)$).
Collect the bounds.
\end{proof}

\begin{remark}[Superlevel selection is signed (remaining sign obstruction)]\label{rem:superlevel-selection-block}
Lemma~\ref{lem:superlevel-selection} achieves the requested ``superlevel-set selection'' step: it converts the stretching injection on $\{\rho\ge 1-\eta\}$ into a bound involving only
\(\rho^{3/2}|\nabla\xi|^2\),
\(|\nabla(\rho^{3/4})|^2\),
and cutoff/time-boundary errors, \emph{without any use of} $\nabla\sigma$ or maximizer tracking.

\smallskip
\noindent
However, the bound is for the \emph{signed} integral of $\rho^{3/2}\sigma$ on the superlevel set.
To bound the desired positive part $\iint_{\{\rho\ge 1-\eta\}}\rho^{3/2}\sigma_+$, one still needs an additional mechanism controlling the negative part of $\sigma$ on the same set (or a structural reason why $\sigma$ cannot oscillate in sign there).
This is the remaining obstruction to converting the superlevel-set estimate into a uniform quantitative bound on $\iint\rho^{3/2}\sigma_+$.
\end{remark}

\begin{remark}[Superlevel time-fraction attempt fails (signed injection)]\label{rem:superlevel-time-fraction-sign}
It is tempting to turn \eqref{eq:superlevel-injection-simplified} into a ``large injection occurs only on a small fraction of times'' statement by defining
\(
\mathsf J(t):=\int_{B_r(x_0)\cap\{\rho(\cdot,t)\ge 1-\eta\}}\rho^{3/2}\sigma
\)
and applying Markov/Chebyshev.
However, $\mathsf J(t)$ is \emph{signed} (since $\sigma$ changes sign), so Markov's inequality applies only to $\mathsf J_+(t)$ (or $|\mathsf J(t)|$).
Thus, obtaining a genuine time-fraction bound for large \emph{positive} injection requires an \emph{additional} mechanism controlling $\int \mathsf J_+(t)\,dt$ (equivalently, controlling the positive part $\iint_{\{\rho\ge 1-\eta\}}\rho^{3/2}\sigma_+$ or at least $\iint_{\{\rho\ge 1-\eta\}}\rho^{3/2}|\sigma|$).
This is another concrete way to see why the remaining C2 task is to control $\sigma_+$ on high-vorticity superlevel sets.
\end{remark}

\begin{remark}[Heuristic consequence: negative stretching on $\{\rho\approx 1\}$ must be paid for by diffusion]\label{rem:sigma-minus-paid-by-diffusion}
Lemma~\ref{lem:superlevel-selection} can be read as a PDE version of the RS ``finite budget'' intuition on the \emph{top-level} superlevel set.
Roughly: if $\sigma$ were strongly negative on $\{\rho\ge 1-\eta\}$ for a long time, then the signed injection
\(
\iint_{Q_r\cap\{\rho\ge 1-\eta\}}\rho^{3/2}\sigma
\)
would be very negative.
To maintain the running-max cap $\sup_x\rho(\cdot,t)=1$ for all $t$ (Lemma~\ref{lem:runningmax-sup-freeze-3d}), the set $\{\rho\ge 1-\eta\}$ cannot disappear entirely.
The only way to prevent collapse of this superlevel set in the presence of negative reaction is to replenish it through diffusion/transport across the transition band $\{1-2\eta<\rho<1-\eta\}$.
Quantitatively, Lemma~\ref{lem:superlevel-selection} shows that such replenishment necessarily incurs a cost in the band-gradient term
\[
\eta^{-1}\iint_{Q_{2r}\cap\{1-2\eta<\rho<1-\eta\}}|\nabla(\rho^{3/4})|^2,
\]
and in the global cutoff/time-boundary errors.

\smallskip
\noindent
This provides a precise dichotomy: cancellation of $\sigma_+$ by $\sigma_-$ on the top-level superlevel set is only possible if one pays a compensating diffusion/transition cost.
Turning this dichotomy into a \emph{uniform} bound on $\iint\rho^{3/2}\sigma_+$ (or into a global smallness of $\mathcal E_\omega$) would require additional large-scale control of the boundary terms and/or a mechanism preventing the transition-band cost from concentrating on vanishingly small spacetime regions.%
\end{remark}

\begin{lemma}[Crude inherited bound on the critical vorticity mass on balls]\label{lem:rho32-ball-bound}
Let $(u^\infty,p^\infty)$ be the running-max ancient element from Lemma~\ref{lem:ancient-limit-runningmax} and write $\rho^\infty:=|\omega^\infty|$.
Then for every $R>0$ and every $t\le 0$,
\[
\int_{B_R}\bigl(\rho^\infty(x,t)\bigr)^{3/2}\,dx\ \le\ |B_R|\ \le\ C\,R^3.
\]
In particular,
\[
\sup_{t\le 0}\int_{B_R}\bigl(\rho^\infty(x,t)\bigr)^{3/2}\,dx\ \le\ C\,R^3.
\]
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:ancient-limit-runningmax}(iii), $\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$, hence $0\le \rho^\infty\le 1$ pointwise.
Therefore $(\rho^\infty)^{3/2}\le 1$ and the bound follows by integrating over $B_R$.
\end{proof}

\begin{remark}[Historical: why this lemma alone does not close C2]\label{rem:rho32-ball-not-enough}
Lemma~\ref{lem:rho32-ball-bound} provides a universal (but coarse) growth bound on the critical vorticity mass on balls.
This is sufficient to control the time-boundary term in localized identities (e.g.\ Lemma~\ref{lem:injection-damping-balance}) when $r\ll 1$ (since then $R\sim r$ and $R^3$ is small).
However, it does not control the \emph{diffusion budget} $\iint |\nabla(\rho^{3/4})|^2$ on cylinders in any uniform way.

\textbf{[CONDITIONAL.]} The C2 closure is achieved in Theorem~\ref{thm:C2-closure} via the third option above: the $\sigma$-decomposition combined with the spatial-decay gate (Assumption~\ref{assump:U-decay}) controls $\iint\rho^{3/2}\sigma_+$ without requiring diffusion budget control.
\end{remark}

\begin{remark}[Small-scale behavior of the cutoff/time-boundary errors]\label{rem:rho32-errors-small-scale}
For the running-max ancient element, $\rho\le 1$ implies that the cutoff/time-boundary errors appearing in the localized $\rho^{3/2}$ identities scale like $r^3$ as $r\downarrow 0$.
For example, in Lemma~\ref{lem:injection-damping-balance} and Lemma~\ref{lem:superlevel-selection} the error terms of the form
\[
r^{-2}\iint_{Q_{2r}(z_0)}\rho^{3/2}\,dx\,dt
\qquad\text{and}\qquad
\sup_{t\in(t_0-(2r)^2,t_0)}\int_{B_{2r}(x_0)}\rho^{3/2}(\cdot,t)\,dx
\]
are bounded by $C r^3$ and $C r^3$, respectively, since $\rho^{3/2}\le 1$ and $|Q_{2r}|\sim r^5$, $|B_{2r}|\sim r^3$.
Thus, at \emph{sufficiently small scales}, the only genuinely nontrivial contribution in the superlevel-set selection mechanism is the transition-band diffusion budget
\(
\eta^{-1}\iint_{Q_{2r}\cap\{1-2\eta<\rho<1-\eta\}}|\nabla(\rho^{3/4})|^2.
\)
\end{remark}

\begin{lemma}[Band-gradient control from the log-amplitude estimate on high-vorticity sets]\label{lem:band-gradient-from-logamp}
Let $(u^\infty,p^\infty)$ be the running-max ancient element and write $\omega^\infty=\rho\,\xi$ on $\{\rho>0\}$ with $\rho:=|\omega^\infty|$.
Fix $\eta\in(0,1/4]$ and a cylinder $Q_{2r}(z_0)$ with $0<r\le 1$.
Then on the high-vorticity set $\{\rho\ge 1-2\eta\}\cap Q_{2r}(z_0)$ one has the pointwise comparison
\[
|\nabla(\rho^{3/4})|^2\ \le\ C_\eta\,|\nabla\log\rho|^2,
\qquad
C_\eta\sim (1-2\eta)^{3/2},
\]
and hence for every $\varepsilon\in(0,1)$,
\[
\iint_{Q_{2r}(z_0)\cap\{\rho\ge 1-2\eta\}} |\nabla(\rho^{3/4})|^2
\ \le\ C_\eta \iint_{Q_{2r}(z_0)} |\nabla\log(\rho+\varepsilon)|^2.
\]
\end{lemma}

\begin{proof}
On $\{\rho\ge 1-2\eta\}$ we have $(\rho+\varepsilon)^{-1}\le (1-2\eta)^{-1}$ and $\rho^{-1/2}\le (1-2\eta)^{-1/2}$.
Since $\nabla(\rho^{3/4})=\frac{3}{4}\rho^{-1/4}\nabla\rho$ and $\nabla\log(\rho+\varepsilon)=(\rho+\varepsilon)^{-1}\nabla\rho$, we obtain
\[
|\nabla(\rho^{3/4})|^2
=\frac{9}{16}\rho^{-1/2}|\nabla\rho|^2
\le C_\eta\,(\rho+\varepsilon)^{-2}|\nabla\rho|^2
= C_\eta\,|\nabla\log(\rho+\varepsilon)|^2,
\]
with $C_\eta$ depending only on the lower bound $1-2\eta$.
Integrate over the stated set.
\end{proof}

\begin{remark}[Band-gradient control reduces to a global budget obstruction]
Lemma~\ref{lem:band-gradient-from-logamp} shows that, on the high-vorticity region $\{\rho\ge 1-2\eta\}$, the band diffusion cost appearing in Lemma~\ref{lem:superlevel-selection} can be controlled by the log-amplitude gradient.
However, Lemma~\ref{lem:log_amplitude} bounds $\iint|\nabla\log(\rho+\varepsilon)|^2$ in terms of \emph{(i)} the stretching magnitude $|\sigma|$, \emph{(ii)} the direction energy $|\nabla\xi|^2$, and \emph{(iii)} a drift term (now written using the divergence-free affine gauge $\ell_{x_0,r}$, which removes the curl-free affine obstruction but still leaves a scale-critical budget term).
Thus, without an additional inherited global estimate that controls these quantities (uniformly in basepoint/scale), one cannot close the loop and produce a scale-uniform bound on the band diffusion budget.

\smallskip
\noindent
In other words: the current PDE reductions are now sharp enough that the remaining obstruction is purely a \emph{global budget} question for the running-max ancient element (exactly the C2 gate).%
\end{remark}

\begin{lemma}[Local-in-time bound on the band payment from local energy and CZ drift control]\label{lem:band-payment-local-time}
Let $(u^\infty,p^\infty)$ be the running-max ancient element from Lemma~\ref{lem:ancient-limit-runningmax} and write $\omega^\infty=\rho\,\xi$ on $\{\rho>0\}$ with $\rho:=|\omega^\infty|$.
Fix $\eta\in(0,1/4]$, a basepoint $z_0=(x_0,t_0)$, and a scale $0<r\le 1$.
Then there exists a constant $C_\eta<\infty$ such that for every $\varepsilon\in(0,1)$,
\begin{align}\label{eq:band-payment-local}
\eta^{-1}\iint_{Q_{2r}(z_0)\cap\{1-2\eta<\rho<1-\eta\}} |\nabla(\rho^{3/4})|^2
\ \le\ &
C_\eta\,r^3
\ +\ C_\eta\iint_{Q_{4r}(z_0)} |\sigma|
\ +\ C_\eta\iint_{Q_{4r}(z_0)} |\nabla\xi|^2 \nonumber\\
&\ +\ C_\eta\,r^{-2}\iint_{Q_{4r}(z_0)}|u^\infty-\ell_{x_0,4r}(\cdot,t)|^2\,dx\,dt,
\end{align}
where $\sigma=(S\xi\cdot\xi)$ and $\ell_{x_0,4r}(\cdot,t)$ denotes the divergence-free affine approximation
\[
\ell_{x_0,4r}(x,t):=u_{B_{4r}(x_0)}(t)\ +\ \bigl(\nabla u\bigr)_{B_{4r}(x_0)}(t)\,(x-x_0).
\]
All integrals are over spacetime.
\end{lemma}

\begin{proof}[Proof sketch]
On the band $\{1-2\eta<\rho<1-\eta\}$ we have $\rho\ge 1-2\eta$, hence Lemma~\ref{lem:band-gradient-from-logamp} yields
\[
|\nabla(\rho^{3/4})|^2\ \le\ C_\eta\,|\nabla\log(\rho+\varepsilon)|^2.
\]
Therefore the left-hand side of \eqref{eq:band-payment-local} is bounded by
$C_\eta\eta^{-1}\iint_{Q_{2r}(z_0)}|\nabla\log(\rho+\varepsilon)|^2$.
Now apply Lemma~\ref{lem:log_amplitude} at scale $2r$ and use that the cutoff/time-boundary errors are $O(r^3)$ when $\rho\le 1$ (Remark~\ref{rem:rho32-errors-small-scale}).
The affine-gauged drift contribution coming from Lemma~\ref{lem:log_amplitude} yields the last term in \eqref{eq:band-payment-local}.
\end{proof}

\begin{corollary}[Affine-gauged oscillation is lower order under bounded vorticity]\label{cor:affine-gauged-osc-lower}
In the setting of Lemma~\ref{lem:band-payment-local-time}, the affine-gauged term is lower order at small scales:
if $\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$, then for $0<r\le 1$,
\[
r^{-2}\iint_{Q_{4r}(z_0)}|u^\infty-\ell_{x_0,4r}|^2\,dx\,dt\ \le\ C\,r^5.
\]
In particular, it can be absorbed into the $C_\eta r^3$ term in \eqref{eq:band-payment-local} for $r\le 1$.
\end{corollary}

\begin{proof}[Proof sketch]
Fix $t$ and write $\ell:=\ell_{x_0,4r}(\cdot,t)$.
Since $\nabla(u^\infty-\ell)=\nabla u^\infty-(\nabla u^\infty)_{B_{4r}}(t)$, Poincar\'e gives
\[
\|u^\infty(\cdot,t)-\ell\|_{L^2(B_{4r})}\ \lesssim\ r\,\|\nabla u^\infty(\cdot,t)-(\nabla u^\infty)_{B_{4r}}(t)\|_{L^2(B_{4r})}.
\]
By Lemma~\ref{lem:drift-bmo-from-vorticity}, $\|\nabla u^\infty(\cdot,t)\|_{\BMO}\lesssim \|\omega^\infty(\cdot,t)\|_{L^\infty}\le 1$ for a.e.\ $t$.
John--Nirenberg then yields $\|\nabla u^\infty-(\nabla u^\infty)_{B_{4r}}\|_{L^2(B_{4r})}\lesssim r^{3/2}$, hence
$\|u^\infty-\ell\|_{L^2(B_{4r})}^2\lesssim r^5$.
Integrating over a time interval of length $(4r)^2$ gives $\iint_{Q_{4r}}|u^\infty-\ell|^2\lesssim r^7$, and dividing by $r^2$ yields the claim.
\end{proof}

\begin{corollary}[Simplified band-payment bound for the running-max ancient element]\label{cor:band-payment-simplified}
In the setting of Lemma~\ref{lem:band-payment-local-time} for the running-max ancient element (so $\|\omega^\infty\|_{L^\infty}\le 1$), combining \eqref{eq:band-payment-local} with Corollary~\ref{cor:affine-gauged-osc-lower} yields the simplified bound
\[
\eta^{-1}\iint_{Q_{2r}(z_0)\cap\{1-2\eta<\rho<1-\eta\}} |\nabla(\rho^{3/4})|^2
\ \le\ C_\eta\,r^3
\ +\ C_\eta\iint_{Q_{4r}(z_0)} \bigl(|\sigma|+|\nabla\xi|^2\bigr)\,dx\,dt
\]
for all $0<r\le 1$.
\end{corollary}

\begin{remark}[Status of the band-payment estimate (global uniformity is still open)]\label{rem:band-payment-local-audit}
Lemma~\ref{lem:band-payment-local-time} provides a \emph{local-in-time} control of the transition-band diffusion cost in terms of scale-critical quantities (notably $\iint |\sigma|$ and $\iint |\nabla\xi|^2$) and an affine-gauged oscillation term.
However, turning this into a \emph{uniform} bound in basepoint/scale (as would be needed for a global C2 contradiction) requires additional global control on the running-max ancient element.
In particular, RM2-type control of the affine/harmonic obstruction alone does not automatically yield the required uniform bounds on $\iint |\sigma|$ or $\iint |\nabla\xi|^2$ across all cylinders.
This uniformity issue is part of the remaining global C2/RM2 closure work tracked in \texttt{ASSUMPTIONS\_TO\_CLOSE.md}.
\end{remark}

\begin{lemma}[Local Hodge control of $\nabla u$ by vorticity and velocity oscillation]\label{lem:local-hodge-grad-u}
Let $u(\cdot,t)\in H^1_{\mathrm{loc}}(\R^3)$ be divergence-free and set $\omega(\cdot,t)=\curl u(\cdot,t)$.
Fix $x_0\in\R^3$ and $r>0$, and let
\[
c(t):=c_{x_0,2r}(t)=\frac{1}{|B_{2r}|}\int_{B_{2r}(x_0)}u(x,t)\,dx.
\]
Then for a.e.\ $t$,
\[
\int_{B_r(x_0)}|\nabla u(x,t)|^2\,dx
\ \le\ C\int_{B_{2r}(x_0)}|\omega(x,t)|^2\,dx
\ +\ C r^{-2}\int_{B_{2r}(x_0)}|u(x,t)-c(t)|^2\,dx,
\]
with a universal constant $C$.
\end{lemma}

\begin{proof}[Proof sketch]
Let $\phi\in C_c^\infty(B_{2r}(x_0))$ satisfy $\phi\equiv 1$ on $B_r(x_0)$ and $|\nabla\phi|\lesssim r^{-1}$.
Apply the vector-calculus identity
\(
\|\nabla(\phi v)\|_{L^2}^2
=\|\curl(\phi v)\|_{L^2}^2+\|\div(\phi v)\|_{L^2}^2
\)
to $v:=u-c(t)$ (so $\div v=0$ and $\curl v=\omega$).
Expanding $\curl(\phi v)$ and $\div(\phi v)$ and using $|\nabla\phi|\lesssim r^{-1}$ yields
\[
\int |\nabla(\phi v)|^2
\ \lesssim\ \int \phi^2|\omega|^2\ +\ \int |\nabla\phi|^2|v|^2.
\]
Since $\phi\equiv 1$ on $B_r(x_0)$, $\int_{B_r}|\nabla u|^2\le \int|\nabla(\phi v)|^2$, which gives the claim.
\end{proof}

\begin{corollary}[Local dissipation bound from vorticity and the (harmonic/affine) velocity mode]\label{cor:local-dissipation-from-hodge}
Let $(u,p)$ be smooth on $Q_{2r}(z_0)$ and assume $\|\omega\|_{L^\infty(Q_{2r}(z_0))}\le M$.
Let $c(t):=c_{x_0,2r}(t)$ be the ball average of $u(\cdot,t)$ on $B_{2r}(x_0)$.
Then
\[
r^{-2}\iint_{Q_r(z_0)}|\nabla u|^2\,dx\,dt
\ \le\ C\,M^2\,r^3
\ +\ C\,r^{-4}\iint_{Q_{2r}(z_0)}|u-c(t)|^2\,dx\,dt,
\]
with a universal constant $C$.
\end{corollary}

\begin{proof}[Proof sketch]
Integrate Lemma~\ref{lem:local-hodge-grad-u} over $t\in(t_0-r^2,t_0)$.
The vorticity term is bounded by $M^2|Q_{2r}|\lesssim M^2 r^5$.
Dividing by $r^2$ yields the stated bound.
\end{proof}

\begin{remark}[Why local dissipation control still sees a curl-free mode]
Corollary~\ref{cor:local-dissipation-from-hodge} shows that local dissipation is controlled by vorticity \emph{and} by the local kinetic oscillation $u-c(t)$.
The latter term cannot in general be bounded purely by $\|\omega\|_{L^\infty}$ without an additional normalization that rules out nontrivial curl-free (harmonic/affine) components of $u$.
This issue is well-known: a divergence-free, curl-free field on $\R^3$ can be nonconstant (e.g.\ affine fields), yet has $\omega\equiv 0$.
\end{remark}

\begin{remark}[C2 sub-gate: suppressing the curl-free affine mode of $u$]\label{rem:C2-affine-mode-gate}
The ``curl-free affine mode'' of the velocity is invisible to vorticity (it is a low-frequency degree of freedom in blow-up compactness), so any estimate that tries to control $u$ \emph{purely} from $\omega$ must either fix a gauge or accept an additional global normalization input.

\smallskip
\noindent
\textbf{Route 2 (implemented locally).}
In the present C2 bookkeeping, we eliminate this obstruction at the level of local cutoff estimates by using a \emph{divergence-free affine gauge}:
in Lemma~\ref{lem:log_amplitude} and Lemma~\ref{lem:band-payment-local-time} the drift contribution is written in terms of $u-\ell_{x_0,r}$, where $\ell_{x_0,r}$ is the divergence-free affine approximation of $u$ on $B_r(x_0)$.
This quantity vanishes identically for a purely affine divergence-free field and is controlled by the \emph{oscillation} of $\nabla u$ (hence by $\|\nabla u\|_{\BMO}$, which is controlled by $\|\omega\|_{L^\infty}$ up to constants).

\smallskip
\noindent
\textbf{Route 1 (global, still open).}
A stronger alternative is to rule out affine modes globally by an inherited ``finite capacity'' (linear energy growth) bound
\[
\sup_{t\le 0}\int_{B_R}|u^\infty(x,t)|^2\,dx\ \lesssim\ R\qquad(R\ge 1),
\]
which excludes any nontrivial affine mode (energy $\sim R^5$).
However, as discussed in the closure document (Gate~3 / ``cutoff drift''), such a global bound is not an immediate consequence of finite energy alone and appears to require additional structure (e.g.\ filament geometry after (C)).
\end{remark}

\begin{corollary}[Large band payment can only occur on a small fraction of times]\label{cor:band-payment-time-fraction}
In the setting of Lemma~\ref{lem:band-payment-local-time}, define for $t\in(t_0-4r^2,t_0)$ the instantaneous band payment
\[
\mathsf B(t):=\int_{B_{2r}(x_0)\cap\{1-2\eta<\rho(\cdot,t)<1-\eta\}}|\nabla(\rho^{3/4})(\cdot,t)|^2\,dx.
\]
Then for every $\Lambda>0$,
\[
|\{t\in(t_0-4r^2,t_0):\ \mathsf B(t)\ge \Lambda\}|
\ \le\ \frac{1}{\Lambda}\,
\iint_{Q_{2r}(z_0)\cap\{1-2\eta<\rho<1-\eta\}}|\nabla(\rho^{3/4})|^2\,dx\,dt,
\]
and hence, by Lemma~\ref{lem:band-payment-local-time}, the right-hand side is bounded by an explicit expression involving the scale-critical terms $\iint|\sigma|$ and $\iint|\nabla\xi|^2$ and the affine-gauged oscillation term $r^{-2}\iint|u-\ell_{x_0,4r}|^2$.
\end{corollary}

\begin{remark}[Good-time selection: band payment vs signed injection]
For the transition-band payment, $\mathsf B(t)\ge 0$ and Markov's inequality yields the robust time-fraction estimate in Corollary~\ref{cor:band-payment-time-fraction}.
For the top-set injection, the natural quantity
\(
\mathsf J(t)=\int_{B_r\cap\{\rho\ge 1-\eta\}}\rho^{3/2}\sigma
\)
is \emph{signed}, so the analogous good-time selection for injection would require control of $\int \mathsf J_+(t)\,dt$ (or $|\mathsf J(t)|$), i.e.\ a mechanism controlling $\sigma_+$ (or $|\sigma|$) on $\{\rho\approx 1\}$.
This sign obstruction is recorded in Remark~\ref{rem:superlevel-time-fraction-sign} and is one of the concrete remaining sub-gates in C2.
\end{remark}

\begin{remark}[Scaling of the diffusion budget for $\rho^{3/4}$ under vorticity normalization]\label{rem:rho34-budget-scaling}
Let $u^{(k)}$ be the vorticity-normalized rescaling \eqref{rescaled} with factor $\lambda_k=A_k^{-1/2}$.
Write $\rho^{(k)}:=|\omega^{(k)}|$ and $\rho:=|\omega|$ for the original solution.
Then for any cylinder $B_R\times(-T,0)$ in rescaled variables one has the exact scaling relation
\[
\int_{-T}^0\int_{B_R}\bigl|\nabla_y\bigl((\rho^{(k)})^{3/4}\bigr)\bigr|^2\,dy\,ds
\ =\ \int_{t_k-\lambda_k^2 T}^{t_k}\int_{B_{\lambda_k R}(x_k)}
\bigl|\nabla_x(\rho^{3/4})\bigr|^2\,dx\,dt.
\]
In particular, the rescaled diffusion budget on a fixed unit cylinder corresponds to the original diffusion budget on a \emph{shrinking} cylinder at physical scale $\lambda_k$.

\smallskip
\noindent
Thus, any attempt to obtain a \emph{uniform} bound on
\(\int_{-T}^0\int_{B_R}|\nabla((\rho^\infty)^{3/4})|^2\)
for the running-max ancient element must come from \emph{uniform control} of the original diffusion budget on arbitrarily small cylinders near blow-up times.
This is not provided by the global energy inequality alone and is exactly why converting the ``band cost'' mechanism into an unconditional contradiction is difficult.%
\end{remark}

\begin{remark}[What this does and does not give for C2]\label{rem:inj-damp-interpretation}
Lemma~\ref{lem:injection-damping-balance} formalizes the ``finite budget $\Rightarrow$ cost'' mechanism: on any cylinder, the weighted stretching injection $\rho^{3/2}\sigma$ can only persist if it is balanced by damping through \(\rho^{3/2}|\nabla\xi|^2\) and \(|\nabla(\rho^{3/4})|^2\), up to cutoff/time-boundary errors.

\smallskip
\noindent
However, for C2 one needs a \emph{global} conclusion uniform in $z_0,r$.
Without additional large-scale control of the boundary terms in \eqref{eq:inj-damp} (or a mechanism forcing the signed injection $\iint\rho^{3/2}\sigma$ to be small), this identity alone does not yield a uniform smallness (or vanishing) of the weighted direction coherence $\mathcal E_\omega$.
\end{remark}

\begin{remark}[Optional: CKN-anchored tangent flow (not used in the running-max route)]
The main contradiction chain in this manuscript uses the running-max ancient element of Lemma~\ref{lem:ancient-limit-runningmax}.
For completeness and comparison with the classical partial-regularity framework, we record below the standard CKN-anchored tangent-flow construction at a CKN singular point.
\end{remark}

\begin{lemma}\label{lem:ancient-limit}
Let $u_0\in C_c^\infty(\R^3)$ be divergence-free, let $u$ be the corresponding
smooth solution of the N-S equations \eqref{eq:NS_domain} on its
maximal interval of existence $[0,T^*)$, and assume that $T^*<\infty$ is the
first blow-up time. Let $x^*\in\R^3$ be a CKN-singular point at time $T^*$ as in Lemma~\ref{lem:singular-point}.
Let $r_k\downarrow 0$ be any sequence and define the CKN rescalings
\begin{equation}\label{eq:ckn-rescaled}
\tilde u^{(k)}(y,s):=r_k\,u(x^*+r_k y,\;T^*+r_k^2 s),
\qquad
\tilde p^{(k)}(y,s):=r_k^2\,p(x^*+r_k y,\;T^*+r_k^2 s),
\qquad s<0.
\end{equation}

Then there exists a subsequence (still denoted by $\tilde u^{(k)},\tilde p^{(k)}$) 
and a pair $(u^\infty,p^\infty)$ such that:

\begin{enumerate}

\item[(i)] For every $R>0$ and $T>0$,
\[
\tilde u^{(k)} \to u^\infty \quad\text{strongly in } 
L^p(B_R\times(-T,0)) \quad \text{for all } 1\le p<3,
\]
and
\[
\tilde u^{(k)} \rightharpoonup u^\infty 
\quad \text{weakly in}\quad
L^3_{\mathrm{loc}}(\R^3\times(-\infty,0)).
\]
Moreover,
\[
\tilde p^{(k)} \rightharpoonup p^\infty
\quad\text{weakly in } L^{3/2}_{\mathrm{loc}}(\R^3\times(-\infty,0)).
\]

\item[(ii)]
The limit $(u^\infty,p^\infty)$ is a suitable weak solution of the
N-S equations on $\R^3\times(-\infty,0)$ and satisfies the
local energy inequality on every parabolic cylinder
$B_R\times(-T,0)$.

\item[(iii)] The limit $u^\infty$ is an ancient solution, defined for all $t\le 0$, and it is
non-trivial.  More precisely, there exist $r>0$ and $c>0$ such that
\[
\int_{Q_r(0,0)} |u^\infty(x,t)|^3 \,dx\,dt \;\ge\; c > 0,
\]
where $Q_r(0,0)=B_r(0)\times(-r^2,0)$.
In particular, $u^\infty \not\equiv 0$.
\end{enumerate}

We call $u^\infty$ an \emph{ancient tangent flow} associated to the
blow-up at time $T^*$.
\end{lemma}


\begin{proof}
\textbf{[ADDED PROOF / closure of Lemma~\ref{lem:ancient-limit} (compactness + nontriviality).]}
We outline the standard compactness argument for suitable weak solutions, and we make explicit
the missing nontriviality mechanism.

\medskip
\noindent\textbf{Step 1: Uniform local bounds on cylinders.}
Fix $R>0$. For $k$ sufficiently large, the CKN rescalings \eqref{eq:ckn-rescaled} are well-defined on
$Q_R:=B_R\times(-R^2,0)$ since $T^*+r_k^2 s<T^*$ for all $s\in(-R^2,0)$ and $r_k^2R^2<T^*$ for $k$ large.
Since $u$ is smooth on $[0,T^*)$, each rescaled pair $(\tilde u^{(k)},\tilde p^{(k)})$ is smooth on $Q_R$
and in particular is a suitable weak solution there; hence it satisfies the local energy inequality
(cf.\ Definition~\ref{def:suitable}), with constants independent of $k$ after scaling.
Using standard cutoff functions supported in $B_{2R}$, one obtains a bound of the form
\begin{equation}\label{eq:uniform_local_energy_rescaled}
\sup_{s\in(-R^2,0)}\int_{B_R}|\tilde u^{(k)}(x,s)|^2\,dx
\;+\;\int_{Q_R}|\nabla \tilde u^{(k)}|^2\,dx\,ds
\;\le\; C(R),
\end{equation}
where $C(R)$ is independent of $k$.
By interpolation (Ladyzhenskaya + Sobolev) and \eqref{eq:uniform_local_energy_rescaled} we also get
\begin{equation}\label{eq:uniform_L3_rescaled}
\iint_{Q_R}|\tilde u^{(k)}|^3\,dx\,ds \le C(R).
\end{equation}
Finally, the pressure satisfies the standard local estimate (via
$-\Delta \tilde p^{(k)}=\partial_i\partial_j(\tilde u^{(k)}_i\tilde u^{(k)}_j)$ and Calder\'on--Zygmund),
which yields
\begin{equation}\label{eq:uniform_p32_rescaled}
\|\tilde p^{(k)}\|_{L^{3/2}(Q_R)} \le C(R)
\end{equation}
after fixing the additive-in-time constant of the pressure (see, e.g., \cite{CKN1982,Seregin2012}).

\medskip
\noindent\textbf{Step 2: Compactness (Aubin--Lions).}
From the Navier--Stokes system on $Q_R$,
\[
\partial_s \tilde u^{(k)}=\Delta \tilde u^{(k)}-\nabla \tilde p^{(k)}-(\tilde u^{(k)}\cdot\nabla)\tilde u^{(k)},
\]
the bounds \eqref{eq:uniform_local_energy_rescaled}--\eqref{eq:uniform_p32_rescaled} imply
that $\partial_s \tilde u^{(k)}$ is bounded in a negative Sobolev space on $Q_R$
uniformly in $k$ (e.g.\ in $L^{3/2}(-R^2,0;W^{-2,3/2}(B_R))$).
Therefore, by the Aubin--Lions compactness lemma, after passing to a subsequence we have
\[
\tilde u^{(k)}\to u^\infty \quad\text{strongly in }L^2(Q_R).
\]
Combining strong $L^2$ convergence with the uniform $L^3$ bound \eqref{eq:uniform_L3_rescaled}
and interpolation yields strong convergence in $L^p(Q_R)$ for every $1\le p<3$.
Using a diagonal subsequence over $R\in\N$ gives (i).
Similarly, by \eqref{eq:uniform_p32_rescaled} we may extract a subsequence with
$\tilde p^{(k)}\rightharpoonup p^\infty$ weakly in $L^{3/2}_{\mathrm{loc}}$, proving the pressure part of (i).

\medskip
\noindent\textbf{Step 3: Passage to the limit; suitable weak limit.}
The strong convergence of $\tilde u^{(k)}$ in $L^2_{\mathrm{loc}}$ and the weak convergence of $\nabla \tilde u^{(k)}$
in $L^2_{\mathrm{loc}}$ imply $\tilde u^{(k)}\otimes \tilde u^{(k)}\to u^\infty\otimes u^\infty$ in distributions,
so we may pass to the limit in the N--S equations on each $Q_R$.
Lower semicontinuity passes the local energy inequality to the limit, so $(u^\infty,p^\infty)$
is a suitable weak solution on $\R^3\times(-\infty,0)$, proving (ii).

\medskip
\noindent\textbf{Step 4: Nontriviality (how to close (iii) rigorously).}
Nontriviality follows from the CKN-singularity of $(x^*,T^*)$.
By the contrapositive of CKN $\varepsilon$-regularity, there exists a universal $\varepsilon_{\mathrm{CKN}}>0$ such that
for all sufficiently small $r>0$,
\[
r^{-2}\iint_{Q_r(x^*,T^*)}\bigl(|u|^3+|p|^{3/2}\bigr)\,dx\,dt \;\ge\; \varepsilon_{\mathrm{CKN}}.
\]
Taking $r=r_k$ and using the scale invariance of the CKN functional under \eqref{eq:ckn-rescaled} gives
\[
\iint_{Q_1(0,0)}\bigl(|\tilde u^{(k)}|^3+|\tilde p^{(k)}|^{3/2}\bigr)\,dy\,ds \;\ge\; \varepsilon_{\mathrm{CKN}}
\quad\text{for all }k.
\]
Passing to the limit and using lower semicontinuity yields
\[
\iint_{Q_1(0,0)}|u^\infty|^3\,dy\,ds \;\ge\; c_0>0
\]
for a universal $c_0$, proving (iii) (with $r=1$ and $c=c_0$).

\medskip
\noindent\textit{Remark.} If one prefers the vorticity normalization of Lemma~\ref{lem:blowup-normalization} for later
geometric arguments, one can re-center/renormalize the CKN blow-up sequence at a point of large vorticity
inside $Q_1$; the essential point for (iii) is that the construction must preserve a scale-invariant
lower bound (such as the CKN functional), so that triviality of the limit is ruled out.
\end{proof}






\section{The Vorticity Direction Equation}

\subsection{Derivation of the Coupled System}

Let $u$ be a sufficiently smooth divergence-free solution of the incompressible N–S equations with unit viscosity and $\omega = \curl\, u$ be the vorticity field. In the region $\{\omega \neq 0\}$,
we decompose the vorticity into its magnitude $\rho = |\omega|$ and its direction
$\xi = \omega/|\omega| \in \mathbb{S}^2$. The vorticity equation  can be written in
vector form as
\begin{equation}
\partial_t \omega + (u \cdot \nabla)\omega - \Delta \omega = (\omega \cdot \nabla)u.
    \end{equation}
Substituting $\omega = \rho \xi$ yields
\[
(\partial_t \rho + u \cdot \nabla \rho - \Delta \rho)\xi
+ \rho (\partial_t \xi + u \cdot \nabla \xi - \Delta \xi)
- 2 (\nabla \rho \cdot \nabla) \xi
= \rho (S\xi),
\]
where $S = \tfrac{1}{2}(\nabla u + (\nabla u)^T)$ is the strain tensor. We take the inner product with $\xi$ to isolate the amplitude equation.Using the identities $|\xi|^2=1$, $\xi \cdot \partial_t \xi = 0$, and $\xi \cdot \Delta \xi = -|\nabla \xi|^2$, we obtain:
\begin{equation}\label{eq:amplitude}
\partial_t \rho + u \cdot \nabla \rho - \Delta \rho = \rho (\sigma - |\nabla \xi|^2),
\end{equation}
where $\sigma = (S\xi \cdot \xi)$ is the vortex stretching scalar.

\smallskip
\noindent
\textbf{C2 bridge (direction coherence weight).}
The damping term $-\rho\,|\nabla\xi|^2$ in \eqref{eq:amplitude} shows that direction oscillation suppresses vorticity growth.
At the scale-critical exponent $3/2$, this damping produces the natural vorticity-weighted direction-coherence density $\rho^{3/2}|\nabla\xi|^2$.

\begin{lemma}[The $\rho^{3/2}$ equation and weighted direction coherence]\label{lem:rho32-equation}
On the set $\{\rho>0\}$, the quantity $\rho^{3/2}$ satisfies
\begin{equation}\label{eq:rho32}
\partial_t(\rho^{3/2}) + u\cdot\nabla(\rho^{3/2}) - \Delta(\rho^{3/2})
\;+\; \frac{4}{3}\,|\nabla(\rho^{3/4})|^2
\;=\; \frac{3}{2}\,\rho^{3/2}\,\sigma\;-\;\frac{3}{2}\,\rho^{3/2}\,|\nabla\xi|^2.
\end{equation}
\end{lemma}

\begin{proof}
This is a direct computation from \eqref{eq:amplitude} using the chain rule.
Set $f(s)=s^{3/2}$, so $f'(s)=\frac{3}{2}s^{1/2}$ and $f''(s)=\frac{3}{4}s^{-1/2}$ for $s>0$.
Then
\[
\partial_t(\rho^{3/2})+u\cdot\nabla(\rho^{3/2})-\Delta(\rho^{3/2})
\;=\; f'(\rho)\bigl(\partial_t\rho+u\cdot\nabla\rho-\Delta\rho\bigr)\;-\;f''(\rho)|\nabla\rho|^2.
\]
Substituting \eqref{eq:amplitude} gives
\[
\partial_t(\rho^{3/2})+u\cdot\nabla(\rho^{3/2})-\Delta(\rho^{3/2})
\;=\;\frac{3}{2}\rho^{3/2}(\sigma-|\nabla\xi|^2)\;-\;\frac{3}{4}\rho^{-1/2}|\nabla\rho|^2.
\]
Finally, since $\nabla(\rho^{3/4})=\frac{3}{4}\rho^{-1/4}\nabla\rho$, we have
$|\nabla(\rho^{3/4})|^2=\frac{9}{16}\rho^{-1/2}|\nabla\rho|^2$, i.e.\ $\frac{3}{4}\rho^{-1/2}|\nabla\rho|^2=\frac{4}{3}|\nabla(\rho^{3/4})|^2$.
Rearranging yields \eqref{eq:rho32}.
\end{proof}

\begin{definition}[Vorticity-weighted direction coherence]\label{def:weighted-coherence}
For a cylinder $Q_r(z_0)$, define the vorticity-weighted (scale-invariant) direction-coherence functional by
\[
\mathcal E_\omega(z_0,r)\ :=\ \iint_{Q_r(z_0)} \rho^{3/2}\,|\nabla\xi|^2\,dx\,dt.
\]
\end{definition}

\begin{lemma}[Localized bound for $\mathcal E_\omega$ (reduction to weighted stretching)]\label{lem:weighted-coherence-bound}
Let $u$ be smooth on $Q_{2r}(z_0)$ and let $\rho=|\omega|$ and $\xi=\omega/|\omega|$ on $\{\rho>0\}$.
Let $\phi\in C_c^\infty(Q_{2r}(z_0))$ satisfy $\phi\equiv 1$ on $Q_r(z_0)$ and $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
Then
\begin{equation}\label{eq:weighted-coherence-local}
\iint_{Q_r(z_0)} \rho^{3/2}|\nabla\xi|^2
\ \le\ C\iint_{Q_{2r}(z_0)} \rho^{3/2}\,\sigma_+(x,t)\,dx\,dt
\ +\ C r^{-2}\iint_{Q_{2r}(z_0)} \rho^{3/2}\,dx\,dt
\ +\ C\sup_{t\in(t_0-(2r)^2,t_0)}\int_{B_{2r}(x_0)} \rho^{3/2}(x,t)\,dx,
\end{equation}
with a universal constant $C$.
\end{lemma}

\begin{proof}[Proof sketch]
Multiply \eqref{eq:rho32} by $\phi^2$ and integrate over $Q_{2r}(z_0)$.
Integrate by parts in time for the $\partial_t(\rho^{3/2})$ term and in space for the transport/diffusion terms (using $\nabla\cdot u=0$ for the drift).
The positive term $\frac{4}{3}\iint |\nabla(\rho^{3/4})|^2\phi^2$ is dropped.
The diffusion and drift cutoffs produce the $r^{-2}\iint \rho^{3/2}$ term (via $|\partial_t\phi|\lesssim r^{-2}$ and $|\Delta(\phi^2)|\lesssim r^{-2}$), and the time integration by parts produces the supremum-in-time boundary term.
Since $\rho^{3/2}\ge 0$, the integral of the stretching term satisfies
\(
\iint \rho^{3/2}\sigma\,\phi^2 \le \iint \rho^{3/2}\sigma_+\,\phi^2
\),
so the negative part of $\sigma$ only improves the upper bound on $\iint \rho^{3/2}|\nabla\xi|^2$.
Rearranging yields \eqref{eq:weighted-coherence-local}.
\end{proof}

\begin{remark}[Link to the critical vorticity $L^{3/2}$ balance]\label{rem:rho32-vort-balance}
The same critical weight $\rho^{3/2}$ appears in the classical $L^{3/2}$ vorticity balance.
Formally, testing the vorticity equation against $|\omega|^{-1/2}\omega$ yields an evolution identity for $\int \rho^{3/2}$ whose diffusion term contains
$\rho^{3/2}|\nabla\xi|^2$ (since $|\nabla\omega|^2=|\nabla\rho|^2+\rho^2|\nabla\xi|^2$ by orthogonality).
Thus, any mechanism that controls the \emph{growth} of the critical vorticity mass $\int \rho^{3/2}$ on the running-max ancient element (globally or locally) has the potential to control the integrated direction-coherence density $\rho^{3/2}|\nabla\xi|^2$.

\smallskip
\noindent
\textbf{[CONDITIONAL.]} The C2 obstruction is closed in Theorem~\ref{thm:C2-closure} via the $\sigma$-decomposition, conditional on the spatial-decay gate (Assumption~\ref{assump:U-decay}).%
\end{remark}

\begin{remark}[C2 reduces to controlling weighted positive stretching]\label{rem:C2-stretch-reduction}
Lemma~\ref{lem:weighted-coherence-bound} shows that any attempt to prove a \emph{global} smallness (or vanishing) mechanism for the weighted direction coherence $\mathcal E_\omega$ must ultimately control the \emph{weighted stretching} integral
\[
\iint_{Q_{2r}(z_0)} \rho^{3/2}\,\sigma_+(x,t)\,dx\,dt,
\qquad \sigma=(S\xi\cdot\xi).
\]
The remaining terms in \eqref{eq:weighted-coherence-local} are lower order:
\begin{itemize}
\item the cutoff term $r^{-2}\iint_{Q_{2r}}\rho^{3/2}$ is controlled by bounded vorticity for $r\le 1$ (and is small when $r\ll 1$),
\item the time-boundary term $\sup_t\int_{B_{2r}}\rho^{3/2}$ is a scale-critical quantity that does not automatically vanish without additional large-scale information.
\item the affine-gauged velocity oscillation term appearing in the band-payment estimate (Lemma~\ref{lem:band-payment-local-time}) is lower order at small scales for the running-max ancient element (Corollary~\ref{cor:affine-gauged-osc-lower}), so the curl-free affine mode does not obstruct the local budget identities at $r\ll 1$.
\end{itemize}
Thus, \textbf{C2 reduces to finding a mechanism that prevents persistent positive weighted stretching} in the running-max ancient element.
This is exactly the ``finite budget over infinite history'' intuition: if $\rho^{3/2}\sigma$ injects vorticity mass at a scale-critical rate, then an ancient bounded profile must compensate via the damping $\rho^{3/2}|\nabla\xi|^2$.
This is made quantitative and uniform in Theorem~\ref{thm:C2-closure}.
\end{remark}

\begin{lemma}[Conditional closure of C2 from scale-uniform control of weighted positive stretching]\label{lem:C2-closure-from-stretch}
Let $(u^\infty,p^\infty)$ be the running-max ancient element from Lemma~\ref{lem:ancient-limit-runningmax} and write $\rho=|\omega^\infty|$, $\xi=\omega^\infty/|\omega^\infty|$ on $\{\rho>0\}$.
Assume that the weighted positive stretching is \emph{vanishing at small scales} in the following scale-invariant sense:
there exists a modulus $\alpha:(0,1]\to[0,\infty)$ with $\alpha(r)\to 0$ as $r\downarrow 0$ such that for every $z_0\in\R^3\times(-\infty,0]$ and every $0<r\le 1$,
\begin{equation}\label{eq:C2-stretch-hyp}
\iint_{Q_r(z_0)} \rho(x,t)^{3/2}\,\sigma_+(x,t)\,dx\,dt\ \le\ \alpha(r).
\end{equation}
Then the vorticity-weighted direction coherence is uniformly vanishing at small scales:
there exists a constant $C$ such that for every $0<r\le 1$,
\begin{equation}\label{eq:C2-coherence-vanish}
\sup_{z_0}\ \mathcal E_\omega(z_0,r)\ \le\ C\,\alpha(2r)\ +\ C\,r^3,
\end{equation}
and in particular $\lim_{r\downarrow 0}\sup_{z_0}\mathcal E_\omega(z_0,r)=0$.

\smallskip
\noindent
The same conclusion holds (with $\alpha$ replaced by an upper bound for $\iint \rho^{3/2}|\sigma|$) if one assumes a scale-uniform control of $\iint \rho^{3/2}|\sigma|$ instead of \eqref{eq:C2-stretch-hyp}.
\end{lemma}

\begin{proof}
Fix $z_0$ and $0<r\le 1$.
Apply Lemma~\ref{lem:weighted-coherence-bound} at scale $2r$ to obtain
\[
\mathcal E_\omega(z_0,r)
\le C\iint_{Q_{4r}(z_0)}\rho^{3/2}\sigma_+
\ +\ C r^{-2}\iint_{Q_{4r}(z_0)}\rho^{3/2}
\ +\ C\sup_{t\in(t_0-(4r)^2,t_0)}\int_{B_{4r}(x_0)}\rho^{3/2}(\cdot,t).
\]
By \eqref{eq:C2-stretch-hyp} applied with radius $4r$ we have
$\iint_{Q_{4r}(z_0)}\rho^{3/2}\sigma_+\le \alpha(4r)\le \alpha(2r)$ after redefining $\alpha$ to be nondecreasing (replace $\alpha(r)$ by $\sup_{s\le r}\alpha(s)$).
For the remaining terms, bounded vorticity gives $0\le\rho\le 1$, hence
\[
r^{-2}\iint_{Q_{4r}(z_0)}\rho^{3/2}\ \le\ r^{-2}|Q_{4r}|\ \lesssim\ r^{-2}\cdot r^5\ =\ O(r^3),
\]
and similarly
\(
\sup_t\int_{B_{4r}}\rho^{3/2}\le |B_{4r}|\lesssim r^3.
\)
Combining yields \eqref{eq:C2-coherence-vanish}.
\end{proof}

\begin{remark}[Status of the C2 stretch hypothesis (what remains open)]\label{rem:C2-stretch-hyp-status}
Lemma~\ref{lem:C2-closure-from-stretch} isolates a single scale-invariant missing input: the vanishing of the weighted positive stretching integral \eqref{eq:C2-stretch-hyp}.
At present, the manuscript contains several \emph{partial} mechanisms that fall short of proving \eqref{eq:C2-stretch-hyp} unconditionally:
\begin{itemize}
\item \textbf{Max-point control (running-max cap).}
Remark~\ref{rem:runningmax-injection-constraint} gives a pointwise inequality at maximizers $\rho=1$.
Lemmas~\ref{lem:thick-maximum}--\ref{lem:spacetime-injection-lower} show how to propagate this to a positive-measure region \emph{if} one has uniform control of $-\Delta\rho$ and $\nabla\sigma$ along maximizers.
No such uniform-in-time/maximizer bounds are currently derived for the running-max ancient element.
\item \textbf{Superlevel selection (no $\nabla\sigma$) but signed.}
Lemma~\ref{lem:superlevel-selection} and Corollary~\ref{cor:superlevel-selection-simplified} control the \emph{signed} injection
$\iint_{\{\rho\ge 1-\eta\}}\rho^{3/2}\sigma$ by \(\mathcal E_\omega\), the band diffusion budget, and lower-order errors.
Converting this into control of the positive part $\rho^{3/2}\sigma_+$ requires an additional sign/cancellation mechanism on $\{\rho\approx 1\}$ (Remark~\ref{rem:superlevel-selection-block}) or a scale-uniform diffusion-budget bound (Remark~\ref{rem:rho32-errors-small-scale}), neither of which is currently available.
\item \textbf{Geometric depletion for the \emph{tangential} forcing.}
The near-field commutator machinery controls the \emph{tangential} singular forcing $H_{\mathrm{sing}}=P_\xi(S\xi)$ in the direction equation (Lemma~\ref{lem:nearfield-osc-carleson}).
However, $\sigma=(S\xi\cdot\xi)$ is the \emph{normal} component of $S\xi$ and is not directly controlled by Carleson smallness of $H_{\mathrm{sing}}$.
\end{itemize}
Accordingly, \eqref{eq:C2-stretch-hyp} should be viewed as the current sharp formulation of the remaining C2 gate: one needs a new argument that either (i) controls $\sigma_+$ on high-vorticity regions, or (ii) forces enough signed cancellation plus diffusion payment to make $\iint\rho^{3/2}\sigma_+$ small uniformly on small cylinders.
\end{remark}

%%%%

To isolate the evolution of the direction field $\xi$, we apply the
orthogonal projection $P_\xi = I - \xi \otimes \xi$ onto the tangent space
$T_\xi \mathbb{S}^2$.  
Since $P_\xi \xi = 0$, all terms parallel to $\xi$, including the
amplitude component $(\partial_t \rho + u\cdot\nabla\rho - \Delta\rho)\xi$, 
are eliminated after projection. Thus, to derive the direction equation, we project the vorticity decomposition onto
$T_\xi \mathbb{S}^2$, which yields
\[
\rho (\partial_t \xi + u \cdot \nabla \xi - \Delta \xi)
- 2 P_\xi (\nabla \rho \cdot \nabla) \xi
= \rho P_\xi (S\xi).
\]
Dividing by $\rho$ (where $\rho > 0$) we obtain
\begin{equation}\label{eq:direction_intermediate}
\partial_t \xi + u \cdot \nabla \xi - \Delta \xi = P_\xi(S\xi) + 2 P_\xi\bigl( (\nabla \log\rho) \cdot \nabla \xi \bigr).
\end{equation}

The projection step yields a \emph{tangential} diffusion operator.  Using the identity
$P_\xi(\Delta \xi)=\Delta \xi + |\nabla\xi|^2\xi$ (equivalently $\Delta \xi = P_\xi(\Delta\xi)-|\nabla\xi|^2\xi$),
we may rewrite \eqref{eq:direction_intermediate} in the standard harmonic-map form:
\begin{equation}\label{eq:direction}
\partial_t \xi + u \cdot \nabla \xi - \Delta \xi  = |\nabla\xi|^2\,\xi + H,
\end{equation}
where the forcing $H$ is given by
\[
H = H_{\mathrm{sing}} + H_{\mathrm{geom}}.
\]
Here, $H_{\mathrm{sing}} = P_\xi (S\xi)$ represents the projection of the vortex stretching term, and $H_{\mathrm{geom}}$ collects the geometric coupling terms:
\begin{equation}\label{hgeom}
H_{\mathrm{geom}} = 2 P_\xi \bigl( (\nabla \log \rho) \cdot \nabla \xi \bigr).
\end{equation}
By construction, the singular term $H_{\mathrm{sing}} = P_\xi(S\xi)$ and the
tangential component of $H_{\mathrm{geom}}$ lie in the tangent space
$T_\xi \mathbb{S}^2$.  
The normal component on the right-hand side of \eqref{eq:direction} is the curvature term $|\nabla\xi|^2\xi$.

\begin{remark}[Tangentiality of the geometric coupling term]
Since $|\xi|=1$, one has $\xi\cdot\partial_i\xi=\frac12\partial_i(|\xi|^2)=0$ for each spatial derivative $\partial_i$.
Therefore $(\nabla\log\rho)\cdot\nabla\xi=\sum_i(\partial_i\log\rho)\,\partial_i\xi$ is automatically orthogonal to $\xi$, and hence already lies in $T_\xi\mathbb S^2$.
In particular, the projection in \eqref{hgeom} is redundant:
\[
P_\xi\big((\nabla\log\rho)\cdot\nabla\xi\big)=(\nabla\log\rho)\cdot\nabla\xi.
\]
\end{remark}










%%%%%%%%%%


 \subsection{The Singular Stretching Term}

The term \( H_{\mathrm{sing}} = P_\xi (S\xi) \) encodes the non‑local nonlinearity 
of the N--S equations. 

Strictly speaking, $S$ is a \emph{matrix} field obtained from $\omega$ by a matrix of Calder\'on--Zygmund operators (Riesz transforms).
One convenient way to write Biot--Savart at this level is componentwise:
\[
S_{ij}(x)=\mathrm{p.v.}\int_{\R^3}\mathcal{K}_{ij\ell}(x-y)\,\omega_\ell(y)\,dy,
\]
where $\mathcal{K}$ is a tensor kernel homogeneous of degree $-3$ with cancellation.
Consequently, for each unit vector $e\in\mathbb{S}^2$ there exists a vector-valued Calder\'on--Zygmund kernel $K_e$ (depending linearly on $e$) such that
$(S e)(x)=\mathrm{p.v.}\int_{\R^3} K_e(x-y)\,\omega(y)\,dy$.

\begin{equation}\label{eq:H_sing_integral}

H_{\mathrm{sing}}(x)
=P_{\xi(x)}\bigl(S(x)\xi(x)\bigr)
=P_{\xi(x)}\left(\mathrm{p.v.}\int_{\R^3} K_{\xi(x)}(x-y)\,\omega(y)\,dy\right)
=P_{\xi(x)}\left(\mathrm{p.v.}\int_{\R^3} K_{\xi(x)}(x-y)\,\rho(y)\xi(y)\,dy\right).
\end{equation}

\begin{lemma}[Biot--Savart identity for vortex stretching]\label{lem:biot-savart-stretching}
Let $u$ be smooth, divergence-free on $\R^3$ at a fixed time, with vorticity $\omega=\curl u$.
Then for each $x\in\R^3$,
\[
(\omega\cdot\nabla)u(x)
=
\frac{1}{4\pi}\,\mathrm{p.v.}\int_{\R^3}
\left(
\frac{\omega(x)\times\omega(y)}{|x-y|^3}
\;+\;3\,\frac{(\omega(x)\cdot(x-y))\,(\omega(y)\times(x-y))}{|x-y|^5}
\right)\,dy.
\]
\end{lemma}

\begin{proof}
This follows by differentiating the Biot--Savart law
$u(x)=\frac{1}{4\pi}\int_{\R^3}\frac{(x-y)\times\omega(y)}{|x-y|^3}\,dy$
in the $\omega(x)$ direction and using the identities
$(\omega(x)\cdot\nabla_x)(x-y)=\omega(x)$ and
$(\omega(x)\cdot\nabla_x)|x-y|^{-3}=-3(\omega(x)\cdot(x-y))|x-y|^{-5}$.
\end{proof}

Writing $\omega=\rho\xi$, the first term in Lemma~\ref{lem:biot-savart-stretching} contains the factor
$\omega(x)\times\omega(y)=\rho(x)\rho(y)\,\xi(x)\times\xi(y)$ and therefore vanishes when directions align.
In particular, since $\xi(x)\times\xi(x)=0$, one may rewrite that part using the direction difference $\xi(y)-\xi(x)$.
The second term requires additional cancellation (e.g.\ via $\nabla\cdot\omega=0$ and/or a refined symmetric representation) and is part of what must be made referee-checkable in the ``near-field commutator'' step.

\begin{lemma}[$(\xi\cdot\nabla)u$ as a singular integral]\label{lem:xi-derivative}
Let $u$ be smooth and divergence-free on $\R^3$ at a fixed time, with vorticity $\omega=\curl u$.
For any $x$ with $\omega(x)\neq 0$, set $\xi(x):=\omega(x)/|\omega(x)|$. Then
\[
(\xi(x)\cdot\nabla)u(x)
=\frac{1}{4\pi}\,\mathrm{p.v.}\int_{\R^3}
\left(
\frac{\xi(x)\times\omega(y)}{|x-y|^3}
\;-\;3\,\frac{(\xi(x)\cdot(x-y))\,((x-y)\times\omega(y))}{|x-y|^5}
\right)\,dy.
\]
\end{lemma}

\begin{proof}
Differentiate the Biot--Savart law
$u(x)=\frac{1}{4\pi}\int_{\R^3}\frac{(x-y)\times\omega(y)}{|x-y|^3}\,dy$
in the (constant) direction $\xi(x)$ at the point $x$.
\end{proof}

Since $\xi\parallel\omega$, the antisymmetric part of $\nabla u$ annihilates $\xi$, so $(\xi\cdot\nabla)u=S\xi$ and hence
$H_{\mathrm{sing}}=P_\xi(S\xi)=P_\xi((\xi\cdot\nabla)u)$.
The first term in Lemma~\ref{lem:xi-derivative} is already tangential and equals $\rho(y)\,\xi(x)\times\xi(y)/|x-y|^3$.
The second term does not display a direction-difference factor directly and is one of the main technical obstacles in turning the schematic commutator step into a complete proof.

\begin{lemma}[Scalar stretching as a Biot--Savart singular integral (exhibiting direction-difference cancellation)]\label{lem:sigma-singint}
Let $u$ be smooth and divergence-free on $\R^3$ at a fixed time, with vorticity $\omega=\curl u$.
Assume $u$ is represented by the (full-space) Biot--Savart law so that Lemma~\ref{lem:xi-derivative} applies.
For any $x$ with $\omega(x)\neq 0$, set $\rho(x):=|\omega(x)|$ and $\xi(x):=\omega(x)/|\omega(x)|$ and write $r:=x-y$.
Then the vortex-stretching scalar
\(
\sigma(x)=(S(x)\xi(x)\cdot\xi(x))
\)
admits the singular-integral representation
\begin{equation}\label{eq:sigma-singint}
\sigma(x)
\;=\;-\frac{3}{4\pi}\,\mathrm{p.v.}\int_{\R^3}\frac{(\xi(x)\cdot r)\,((\xi(x)\times r)\cdot \omega(y))}{|r|^5}\,dy.
\end{equation}
Equivalently, writing $\omega(y)=\rho(y)\,\xi(y)$,
\begin{equation}\label{eq:sigma-singint-dir-diff}
\sigma(x)
\;=\;-\frac{3}{4\pi}\,\mathrm{p.v.}\int_{\R^3}\frac{(\xi(x)\cdot r)\,\rho(y)\,((\xi(x)\times r)\cdot (\xi(y)-\xi(x)))}{|r|^5}\,dy.
\end{equation}
In particular, the Biot--Savart component of $\sigma$ vanishes whenever the vorticity direction is locally constant (since $(\xi(x)\times r)\cdot \xi(x)=0$), so \eqref{eq:sigma-singint-dir-diff} is the natural “normal-component commutator” analogue of the $H_{\mathrm{sing}}$ oscillation structure.
\end{lemma}

\begin{proof}
Dot the identity in Lemma~\ref{lem:xi-derivative} with $\xi(x)$.
Since $\xi(x)\cdot(\xi(x)\times\omega(y))=0$, only the second term contributes, giving \eqref{eq:sigma-singint}.
For \eqref{eq:sigma-singint-dir-diff}, write $\omega(y)=\rho(y)\xi(y)$ and use the identity
$(\xi(x)\times r)\cdot \xi(y)=(\xi(x)\times r)\cdot(\xi(y)-\xi(x))$ because $(\xi(x)\times r)\cdot\xi(x)=0$.
\end{proof}

\begin{remark}[Biot--Savart gauge caveat for the scalar stretching representation]
The representation \eqref{eq:sigma-singint} is an identity for the Biot--Savart component of the velocity (i.e.\ after fixing a gauge that excludes curl-free harmonic/affine modes of $u$).
In the running-max blow-up compactness, such modes are a known obstruction: one may add a divergence-free, curl-free affine field to $u$ without changing $\omega$, but it changes $S$ (hence $\sigma$).
Accordingly, any unconditional use of \eqref{eq:sigma-singint} in C2 must either (i) justify an inherited global normalization that removes these modes, or (ii) localize the estimate in a way that is insensitive to them (as done for cutoff drift terms via the affine gauge $\ell_{x_0,r}$).
\end{remark}

Writing $\omega=\rho\,\xi$ in Lemma~\ref{lem:xi-derivative} yields the decomposition
\[
H_{\mathrm{sing}}(x)=I_{\mathrm{null}}(x)+I_{\mathrm{const}}(x)+I_{\mathrm{osc}}(x),
\]
where (with $r:=x-y$)
\[
I_{\mathrm{null}}(x):=\frac{1}{4\pi}\,\mathrm{p.v.}\int_{\R^3}\frac{\rho(y)\,\xi(x)\times\xi(y)}{|r|^3}\,dy,
\qquad
I_{\mathrm{const}}(x):=-\frac{3}{4\pi}\,\mathrm{p.v.}\int_{\R^3}\frac{(\xi(x)\cdot r)\,\rho(y)\,(r\times\xi(x))}{|r|^5}\,dy,
\]
and
\[
I_{\mathrm{osc}}(x):=-\frac{3}{4\pi}\,P_{\xi(x)}\,\mathrm{p.v.}\int_{\R^3}\frac{(\xi(x)\cdot r)\,\rho(y)\,(r\times(\xi(y)-\xi(x)))}{|r|^5}\,dy.
\]
In particular, $I_{\mathrm{null}}$ vanishes pointwise when $\xi(y)=\xi(x)$, while $I_{\mathrm{const}}$ is a fixed Calder\'on--Zygmund operator on $\rho$ depending only on the frozen direction $\xi(x)$, and equals
$I_{\mathrm{const}}(x)=\xi(x)\times\nabla\bigl((\xi(x)\cdot\nabla)(-\Delta)^{-1}\rho\bigr)(x)$.
If $\xi$ is exactly constant and $\nabla\cdot\omega=0$ (so $\xi\cdot\nabla\rho=0$), then $I_{\mathrm{const}}\equiv 0$ and hence $H_{\mathrm{sing}}\equiv 0$ as required.

To separate the singular local interaction from the smoother far‑field contribution, 
we fix a (small) radius \( r > 0 \) and decompose the integral into a near‑field 
part and a tail:
\[
H_{\mathrm{sing}} = H_{\mathrm{near}} + H_{\mathrm{tail}},
\]
where
\[
\begin{aligned}
H_{\mathrm{near}}(x) &= P_{\xi(x)}\Bigl( \mathrm{p.v.} \int_{B_r(x)} K_{\xi(x)}(x-y) \rho(y) \xi(y) \, dy \Bigr), \\[2mm]
H_{\mathrm{tail}}(x)  &= P_{\xi(x)}\Bigl( \int_{\mathbb{R}^3 \setminus B_r(x)} K_{\xi(x)}(x-y) \rho(y) \xi(y) \, dy \Bigr).
\end{aligned}
\]
For fixed $r$, the operator $f\mapsto \int_{\R^3\setminus B_r(x)} K_{\xi(x)}(x-y)f(y)\,dy$ is a standard Calder\'on--Zygmund truncation (up to the frozen-direction dependence).
Thus, from scale-critical $L^{3/2}$ bounds on $\rho=|\omega|$ one can obtain \emph{boundedness} of the tail contribution in the critical Carleson norm.
However, \emph{smallness as $r\to0$ does not follow} from scale-critical control alone; it requires additional input (e.g.\ vanishing-Carleson hypotheses or a separate far-field depletion mechanism).
See \texttt{NS\_Unconditional\_Closures\_A\_to\_E.tex}, \S\texttt{subsec:D-tail}.
Here $K_{\xi(x)}$ denotes the vector-valued Calder\'on--Zygmund kernel appearing in \eqref{eq:H_sing_integral}.
For readability, the dependence on $\xi(x)$ is often suppressed later in the text; any use of CRW/commutator estimates
must account for this dependence.

The dependence of $K_{\xi(x)}$ on the frozen direction is \emph{linear} in $\xi(x)$ for the Biot--Savart-derived formula in Lemma~\ref{lem:xi-derivative}.
Consequently, for any fixed $a\in S^2$, the difference operator $(T_{\xi(x)}-T_a)$ has kernel bounded by $C|\xi(x)-a|/|x-y|^3$ and is a Calder\'on--Zygmund operator with $L^p$ operator norm $\lesssim |\xi(x)-a|$.
On a small ball where $\xi$ has small mean oscillation (VMO/BMO$_{\le r}$ small), one can choose $a$ to be the local average direction and ``freeze'' the kernel to $T_a$, paying an error controlled by the oscillation of $\xi$.
This is the natural analytic precursor to any referee-checkable CRW commutator estimate in the presence of $x$-dependent frozen kernels.

The analysis of \( H_{\mathrm{near}} \) is central to our method. A key observation (e.g. see 
\cite{ConstantinFefferman1993}), is that the near‑field term decomposes into:
(i) a \emph{constant-direction} part (obtained by freezing $\xi(y)$ to $\xi(x)$) and
(ii) an \emph{oscillation} part (carrying $\xi(y)-\xi(x)$).
Explicitly, write \( \xi(y) = \xi(x) + (\xi(y) - \xi(x)) \); then
\[
H_{\mathrm{near}}(x) = P_{\xi(x)}\Bigl( 
\int_{B_r(x)} K(x-y)\rho(y)\,\xi(x)\,dy 
+ \mathrm{p.v.} \int_{B_r(x)} K(x-y)\rho(y)\bigl(\xi(y)-\xi(x)\bigr)dy 
\Bigr).
\]
The cancellation properties of the ``constant-direction'' contribution
$P_{\xi(x)}\!\left(\int_{B_r(x)} K(x-y)\rho(y)\,\xi(x)\,dy\right)$
depend on the \emph{exact} Biot--Savart representation of $P_\xi(S\xi)$.
As discussed in the kernel-consistency note leading to \eqref{eq:H_sing_integral}, the operator involves the contraction with $\xi(x)$ and the projection,
so a referee-checkable depletion argument requires an explicit identity showing that $H_{\mathrm{near}}$ can be rewritten \emph{purely} in terms of the oscillation
$\xi(y)-\xi(x)$ (a true commutator form), so that constant $\xi$ yields $H_{\mathrm{near}}\equiv 0$.
This derivation is not supplied in the current manuscript and must be added (or stated as an explicit hypothesis).

Lemma~\ref{lem:xi-derivative} shows that there is a nontrivial ``constant-direction'' contribution hiding inside the second term:
if one freezes $\xi(y)$ to $\xi(x)$ in that term (i.e.\ replaces $\omega(y)$ by $\rho(y)\,\xi(x)$), then the resulting vector field equals
$\xi(x)\times\nabla((\xi(x)\cdot\nabla)(-\Delta)^{-1}\rho)(x)$ (up to universal constants), which is a fixed Calder\'on--Zygmund operator on $\rho$.
In the \emph{ideal} constant-direction case, $\omega=\rho\,\xi$ with $\xi$ constant and $\nabla\cdot\omega=0$ forces $(\xi\cdot\nabla)\rho=0$, and then
$(\xi\cdot\nabla)(-\Delta)^{-1}\rho\equiv 0$ (Fourier support has $\xi\cdot k=0$), so this term vanishes as it must.
Moreover, using $\nabla\cdot\omega=0$ one has for any fixed $a\in S^2$ the exact identity
$a\cdot\nabla\rho=\nabla\cdot(\rho a-\omega)$, and therefore
\[
a\times\nabla\bigl((a\cdot\nabla)(-\Delta)^{-1}\rho\bigr)
\;=\;a\times\nabla(-\Delta)^{-1}\nabla\cdot(\rho a-\omega).
\]
Taking $a=\xi(x)$ shows that this ``constant-direction'' term can be rewritten as a CZ operator applied to the \emph{direction error} $\rho(\xi(x)-\xi)$.
The remaining issue is to make this cancellation \emph{quantitative} (small in the critical Carleson norm) under the hypotheses available for the running-max ancient element.

\begin{lemma}[Constant-direction remainder as a CZ operator on the direction error]\label{lem:constdir-remainder}
Let $u$ be smooth and divergence-free on $\R^3$ at a fixed time, with vorticity $\omega=\curl u$. Write $\omega=\rho\,\xi$ on $\{\omega\neq0\}$ and extend $\rho:=|\omega|$ by $0$ on $\{\omega=0\}$. Fix a constant unit vector $a\in\Sbb^2$.
Then, in the sense of distributions on $\R^3$,
\[
a\times\nabla\bigl((a\cdot\nabla)(-\Delta)^{-1}\rho\bigr)
\;=\;a\times\nabla(-\Delta)^{-1}\nabla\cdot(\rho a-\omega).
\]
In particular, since $\rho a-\omega=\rho(a-\xi)$, the left-hand side is a Calder\'on--Zygmund operator applied to the direction error $\rho(a-\xi)$.
\end{lemma}

\begin{proof}
Since $\nabla\cdot\omega=0$, we have $\nabla\cdot(\rho a-\omega)=a\cdot\nabla\rho$ in distributions. Therefore
\[
(a\cdot\nabla)(-\Delta)^{-1}\rho=(-\Delta)^{-1}(a\cdot\nabla\rho)=(-\Delta)^{-1}\nabla\cdot(\rho a-\omega),
\]
and applying $a\times\nabla$ to both sides yields the claim.
\end{proof}

\begin{lemma}[Quantitative consequence: constant-direction term is controlled by a weighted direction error]\label{lem:constdir-weighted-error}
Fix $a\in\Sbb^2$ and define the constant-direction Calder\'on--Zygmund operator on scalars
\[
(T_a f)(x):=a\times\nabla\bigl((a\cdot\nabla)(-\Delta)^{-1}f\bigr)(x).
\]
Then for every $1<p<\infty$ there exists $C_p<\infty$ such that for all vector fields $F:\R^3\to\R^3$,
\[
\|a\times\nabla(-\Delta)^{-1}\nabla\cdot F\|_{L^p(\R^3)}\le C_p\,\|F\|_{L^p(\R^3)}.
\]
In particular, if $\omega=\rho\,\xi$ with $\nabla\cdot\omega=0$, then for each fixed $a\in\Sbb^2$,
\[
\|T_a \rho\|_{L^p(\R^3)} \;=\; \|a\times\nabla(-\Delta)^{-1}\nabla\cdot(\rho(a-\xi))\|_{L^p(\R^3)}
\;\le\; C_p\,\|\rho(a-\xi)\|_{L^p(\R^3)}.
\]
\end{lemma}

\begin{proof}
Each component of $a\times\nabla(-\Delta)^{-1}\nabla\cdot$ is a finite linear combination of Riesz transforms, hence a Calder\'on--Zygmund operator bounded on $L^p$ for $1<p<\infty$.
The final estimate follows from Lemma~\ref{lem:constdir-remainder} with $F=\rho(a-\xi)$.
\end{proof}

Lemma~\ref{lem:constdir-weighted-error} shows that the remaining ``constant-direction'' contribution is quantitatively controlled by the \emph{weighted direction error} $\rho(a-\xi)$.
Thus, in general one needs a mechanism that makes $\rho(\xi-\text{local frozen direction})$ small in a scale-invariant $L^{3/2}$ sense on shrinking cylinders.
\emph{In the running-max setting}, boundedness of $\rho=|\omega^\infty|$ already provides this automatically (Remark~\ref{rem:constdir-easy-Linfty}).

\begin{remark}[Running-max bonus: bounded vorticity makes the constant-direction remainder Carleson-small]\label{rem:constdir-easy-Linfty}
Let $(u^\infty,p^\infty)$ be the running-max ancient element from Lemma~\ref{lem:ancient-limit-runningmax}, and write $\omega^\infty=\rho^\infty\xi^\infty$ on $\{\omega^\infty\neq 0\}$.
Then $\|\rho^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$ by Lemma~\ref{lem:ancient-limit-runningmax}(iii). Since $|a-\xi^\infty|\le 2$ for any unit vector $a$,
\[
r^{-2}\iint_{Q_r(z_0)} |\rho^\infty(a-\xi^\infty)|^{3/2}
\le (2)^{3/2}\,r^{-2}\,|Q_r|
\le C\,r^{3}\qquad(0<r\le 1),
\]
and hence $\lim_{r_*\to0}\|\rho^\infty(a-\xi^\infty)\|_{C^{3/2}(r_*)}=0$.
Combined with Lemma~\ref{lem:constdir-weighted-error} (with $p=3/2$ and localization to balls), this yields smallness of the constant-direction remainder in the critical Carleson norm at sufficiently small scales.
\end{remark}

At the level of the truncated near-field operator, one has the exact algebraic split
\[
H_{\mathrm{near}}(x)=\frac{1}{4\pi}\,\mathcal T_{\xi(x),r}(\rho(\cdot)\xi(x))(x)
\;+\;P_{\xi(x)}\Bigl(\frac{1}{4\pi}\,\mathcal T_{\xi(x),r}\bigl(\rho(\cdot)(\xi(\cdot)-\xi(x))\bigr)(x)\Bigr),
\]
where $\mathcal T_{\xi(x),r}$ denotes the Biot--Savart-derived truncated singular integral in Lemma~\ref{lem:xi-derivative}.
Using $\nabla\cdot\omega=0$, the \emph{full-space} version of the first term is a CZ operator applied to the direction error $\rho(\xi(x)-\xi)$; truncation introduces an explicit tail remainder, so the near-field commutator reduction is now a precise, referee-checkable target.

\begin{lemma}[Commutator representation of the near-field oscillation forcing]\label{lem:nearfield-osc-commutator}
Let $u$ be smooth and divergence-free on $\R^3$ at a fixed time, with vorticity $\omega=\curl u$.
Write $\omega=\rho\,\xi$ on $\{\omega\neq 0\}$, where $\rho=|\omega|$ and $|\xi|=1$.
Fix a truncation scale $r>0$ and define the truncated Biot--Savart differential operator (from Lemma~\ref{lem:xi-derivative})
\[
(\mathcal T_{a,r}F)(x)\ :=\ \mathrm{p.v.}\int_{B_r(x)}
\left(
\frac{a\times F(y)}{|x-y|^3}
\;-\;3\,\frac{(a\cdot(x-y))\,((x-y)\times F(y))}{|x-y|^5}
\right)\,dy,
\]
for any fixed vector $a\in\R^3$ and any vector field $F$.
For $m,j\in\{1,2,3\}$ define the fixed Calder\'on--Zygmund kernels (with $z\in\R^3\setminus\{0\}$)
\[
k_{m,j}(z)\ :=\ \frac{e_m\times e_j}{|z|^3}\;-\;3\,\frac{z_m\,(z\times e_j)}{|z|^5},
\qquad
(T_{m,j,r}f)(x)\ :=\ \mathrm{p.v.}\int_{B_r(x)} k_{m,j}(x-y)\,f(y)\,dy.
\]
Then, for every $x$ with $\omega(x)\neq 0$,
\begin{equation}\label{eq:nearfield-osc-commutator}
P_{\xi(x)}\Bigl(\mathcal T_{\xi(x),r}\bigl(\rho(\cdot)\,(\xi(\cdot)-\xi(x))\bigr)(x)\Bigr)
\;=\;
P_{\xi(x)}\sum_{m,j=1}^3 \xi_m(x)\,[T_{m,j,r},\xi_j]\,\rho\,(x),
\end{equation}
where $[T,b]f:=T(bf)-b\,Tf$ and $\xi_m:=\xi\cdot e_m$, $\xi_j:=\xi\cdot e_j$.
\end{lemma}

\begin{proof}
Fix $x$ with $\omega(x)\neq 0$ and write $\xi(x)=\sum_{m=1}^3 \xi_m(x)\,e_m$.
Also expand $\xi(y)-\xi(x)=\sum_{j=1}^3(\xi_j(y)-\xi_j(x))\,e_j$.
By bilinearity of the cross product and the definition of $\mathcal T_{\xi(x),r}$,
\[
\mathcal T_{\xi(x),r}\bigl(\rho(\cdot)(\xi(\cdot)-\xi(x))\bigr)(x)
=\sum_{m,j=1}^3 \xi_m(x)\,\mathrm{p.v.}\int_{B_r(x)} k_{m,j}(x-y)\,\rho(y)\,(\xi_j(y)-\xi_j(x))\,dy.
\]
Recognizing the integral as $T_{m,j,r}(\rho\,\xi_j)(x)-\xi_j(x)\,T_{m,j,r}\rho(x)$ yields
\[
\mathcal T_{\xi(x),r}\bigl(\rho(\cdot)(\xi(\cdot)-\xi(x))\bigr)(x)
=\sum_{m,j=1}^3 \xi_m(x)\,[T_{m,j,r},\xi_j]\,\rho\,(x).
\]
Applying the tangential projection $P_{\xi(x)}$ to both sides gives \eqref{eq:nearfield-osc-commutator}.
\end{proof}

Lemma~\ref{lem:sigma-singint} shows that (in a Biot--Savart gauge) the scalar stretching
\(
\sigma=(S\xi\cdot\xi)
\)
is also a singular integral that vanishes when $\xi$ is locally constant.
Expanding \eqref{eq:sigma-singint-dir-diff} in components yields an explicit commutator form at the truncated near-field level.
Define the standard traceless Calder\'on--Zygmund kernels
\[
\kappa_{b,d}(r)\ :=\ \frac{3\,r_b r_d-\delta_{bd}|r|^2}{|r|^5},
\qquad
(\mathcal R_{b,d,r} f)(x)\ :=\ \mathrm{p.v.}\int_{B_r(x)} \kappa_{b,d}(x-y)\,f(y)\,dy,
\]
and write $\varepsilon_{jab}$ for the Levi--Civita symbol.
Then the truncated Biot--Savart contribution to $\sigma$ admits the explicit commutator form
\[
\sigma_{\mathrm{near}}^{\mathrm{BS}}(x)
\;=\;
-\frac{1}{4\pi}\sum_{j,a,b,d=1}^3 \varepsilon_{jab}\,\xi_a(x)\,\xi_d(x)\,[\mathcal R_{b,d,r},\xi_j]\,\rho\,(x),
\]
which is the near-field truncation of the full-space identity obtained by expanding \eqref{eq:sigma-singint-dir-diff} and using the traceless kernel $\kappa_{b,d}$ (the isotropic $\delta_{bd}|r|^{-3}$ part cancels against $\varepsilon_{jab}\xi_a\xi_b=0$).
In particular, $\sigma_{\mathrm{near}}^{\mathrm{BS}}$ is a finite linear combination of commutators with \emph{fixed} truncated CZ operators, with bounded multipliers $\xi_a\xi_d$.

If one can justify that the Biot--Savart gauge is the relevant one for the running-max ancient element (or otherwise control the harmonic/affine ambiguity in Remark~\ref{rem:sigma-harmonic-mode}),
then CRW estimates applied to $[\mathcal R_{b,d,r},\xi_j]\rho$ yield that $\sigma_{\mathrm{near}}^{\mathrm{BS}}$ is \emph{Carleson-small in $L^{3/2}$} at small scales under the bounded-vorticity input $\rho^\infty\in L^\infty$.
This makes the near-field part of the C2 stretching budget negligible at sufficiently small scales, reducing the remaining obstruction to the tail/large-scale component.%


The dangerous part that can become large is precisely the second term, 
involving the difference \( \xi(y)-\xi(x) \). If the direction field \( \xi \) 
varies slowly (e.g. is Lipschitz with a moderate constant), this term remains 
controllable. Rapid oscillations of \( \xi \), on the other hand, can interact 
with the singular kernel to produce uncontrolled amplification, the mechanism 
that could potentially lead to a finite‑time blow‑up.  

Hence, the geometric regularity criterion can be phrased as follows:  
singular vortex stretching can be tamed provided the vorticity direction does not 
oscillate too violently in regions of intense vorticity.


\subsection{The Geometric Forcing Term}

By analyzing the singular stretching term \( H_{\mathrm{sing}} \), we now turn to 
the geometric contributions on the right-hand side of \eqref{eq:direction}.  Geometrically, these arise from the constraint \( |\xi| = 1 \) and 
the coupling between the amplitude \( \rho \) and the direction \( \xi \). They consist 
of two distinct parts:
\begin{enumerate}
    \item The harmonic map tension term \( |\nabla \xi|^2 \xi \), which is
          normal to the sphere \( \mathbb{S}^2 \). In the equation for \( \xi \), 
          it appears as a Lagrange multiplier such that $|\xi|=1$.
    \item The cross‑term \( 2 P_\xi (\nabla \log \rho \cdot \nabla \xi) \), which 
          is tangential and connects the geometry of the direction field to the 
          gradient of the log‑amplitude \( \log\rho \).
\end{enumerate}

Both geometric contributions (the curvature term \( |\nabla \xi|^2 \xi \) and the tangential coupling term \(H_{\mathrm{geom}}\) from \eqref{hgeom}) involve first derivatives and are
bilinear or quadratic in gradients. Under the  scaling (\ref{scaling}), both terms have the same 
homogeneity as the diffusion term $-\Delta\xi$, placing them at the critical 
dimensional threshold. Unlike the 
nonlocal stretching term \(H_{\mathrm{sing}}\), these geometric contributions are 
purely local and, in analytical practice, can often be controlled through energy 
estimates or interpolation inequalities, provided suitable a priori bounds are 
available on \(\nabla\xi\) and \(\nabla\log\rho\). Nevertheless, their critical 
scaling means that they cannot be treated as negligible error terms in a 
blow-up scenario and must be handled with care in any critical or supercritical 
regularity framework.



\section{Critical Coercivity of the Stretching Term}

\subsection{Regularity structure of the direction field}
In the original CKN-tangent-flow route, a VMO/BMO-smallness hypothesis on $\xi^\infty$ is a natural way to force commutator depletion of the near-field oscillation term.
In the running-max rewrite, bounded vorticity already yields near-field oscillation depletion for the commutator/oscillation term (Lemma~\ref{lem:nearfield-osc-carleson}).
We therefore do not treat a directional VMO hypothesis as a separate conditional input in this running-max rewrite.
If a later step truly requires quantitative small oscillation of $\xi^\infty$ (beyond bounded vorticity), that requirement should be stated explicitly at the point of use.

\subsection{The CRW Commutator Estimate}
The key to controlling the singular stretching term lies in the structure of $H_{near}$. 
The ``commutator'' representation below is \emph{schematic} and does not follow from $P_{\xi(x)}\xi(x)=0$ alone,
since the kernel acts before the projection (and the correct Biot--Savart kernel for $S\xi$ depends on $\xi(x)$ as noted in \eqref{eq:H_sing_integral}).
To use CRW rigorously, one must supply a derivation that reduces $H_{near}$ to a Calder\'on--Zygmund commutator with multiplier $\xi$ (or else assume such a representation).
In the present manuscript, the \emph{oscillation} component of $H_{\mathrm{near}}$ has already been reduced to a finite sum of commutators with \emph{fixed} truncated Calder\'on--Zygmund operators; see the explicit identity in the audit block
\textbf{[CHECKABLE COMMUTATOR FORM FOR THE OSCILLATION TERM]} preceding this subsection (cf.\ \eqref{eq:H_sing_integral} and Lemma~\ref{lem:xi-derivative}).

We now record the classical commutator bound that converts small BMO oscillation of $\xi$ into smallness of these commutator terms.

\begin{lemma}[CRW Commutator Estimate]\label{lem:crw}
Let $T$ be a Calder\'on--Zygmund operator on $\R^3$ and let $T_r$ denote a standard truncation at scale $r>0$
(e.g.\ $T_r f(x)=\mathrm{p.v.}\int_{|x-y|<r}K(x-y)f(y)\,dy$ for a CZ kernel $K$).
Then for every $1<p<\infty$ there exists $C_p<\infty$ (depending only on $p$ and CZ constants of $T$) such that for all $r>0$,
\[
\|[T_r,b]f\|_{L^p(\R^3)}\le C_p\,\|b\|_{\BMO(\R^3)}\,\|f\|_{L^p(\R^3)},
\]
where $[T_r,b]f:=T_r(bf)-b\,T_r f$.
\end{lemma}

\begin{proof}
This is the classical Coifman--Rochberg--Weiss commutator theorem \cite{CRW1976}. The dependence on the truncation scale $r$ is uniform.
\end{proof}

\begin{remark}[How \ref{lem:crw} is used here]
Lemma~\ref{lem:crw} is applied to the fixed truncated kernels $T_{m,j,r}$ introduced in the commutator identity
\(
P_{\xi(x)}\sum_{m,j}\xi_m(x)\,[T_{m,j,r},\xi_j]\rho
\)
(see the earlier audit block).
In the running-max setting, $\rho^\infty=|\omega^\infty|$ is \emph{bounded} (Lemma~\ref{lem:ancient-limit-runningmax}(iii)), and $\xi$ is bounded by $1$.
Since $L^\infty\subset\BMO$, the commutator estimate yields a uniform $L^{3/2}$ bound on $H_{\mathrm{near}}^{\mathrm{osc}}$ on each small cylinder, and the parabolic Carleson normalization then forces \emph{smallness as $r\to0$}.
\end{remark}

\begin{lemma}[Near-field commutator/oscillation term is small in the critical Carleson norm]\label{lem:nearfield-osc-carleson}
Let $(u^\infty,p^\infty)$ be the running-max ancient element of Lemma~\ref{lem:ancient-limit-runningmax}, and write $\omega^\infty=\rho^\infty\xi^\infty$ on $\{\omega^\infty\neq0\}$.
Fix $0<r\le 1$ and, for a.e.\ time $t$, define the truncated Calder\'on--Zygmund operators
\[
(T_{m,j,r}f)(x,t)\ :=\ \mathrm{p.v.}\int_{B_r(x)} k_{m,j}(x-y)\,f(y,t)\,dy,
\qquad
k_{m,j}(z):=\frac{e_m\times e_j}{|z|^3}-3\,\frac{z_m\,(z\times e_j)}{|z|^5}.
\]
Define the near-field oscillation forcing (at truncation scale $r$) by the commutator formula from Lemma~\ref{lem:nearfield-osc-commutator}:
\[
H_{\mathrm{near}}^{\mathrm{osc}}(x,t;r)
\;:=\;
\frac{1}{4\pi}\,P_{\xi^\infty(x,t)}\sum_{m,j=1}^3 \xi^\infty_m(x,t)\,[T_{m,j,r},\xi^\infty_j(\cdot,t)]\,\rho^\infty(\cdot,t)\,(x).
\]
Then for every $\varepsilon>0$ there exists $r_0>0$ such that for all $0<r\le r_0$,
\[
\sup_{z_0}\ r^{-2}\iint_{Q_r(z_0)} |H_{\mathrm{near}}^{\mathrm{osc}}(\cdot,\cdot;r)|^{3/2}\,dx\,dt\ \le\ \varepsilon.
\]
\end{lemma}

\begin{proof}
Fix $z_0=(x_0,t_0)$ and $0<r\le 1$. For a.e.\ $t\in(t_0-r^2,t_0)$, by the commutator representation in Lemma~\ref{lem:nearfield-osc-commutator} and the Coifman--Rochberg--Weiss bound (Lemma~\ref{lem:crw} with $p=3/2$), we have
\[
\|H_{\mathrm{near}}^{\mathrm{osc}}(\cdot,t)\|_{L^{3/2}(B_r(x_0))}
\ \le\ C\,\|\xi(\cdot,t)\|_{\BMO(\R^3)}\,\|\rho(\cdot,t)\|_{L^{3/2}(B_{2r}(x_0))}.
\]
Since $|\xi|\le 1$, one has $\|\xi(\cdot,t)\|_{\BMO(\R^3)}\le 2$.
Moreover, by Lemma~\ref{lem:ancient-limit-runningmax}(iii) we have $\|\rho^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$, hence for each $t$,
\(
\|\rho(\cdot,t)\|_{L^{3/2}(B_{2r}(x_0))}\le C\,\|\rho\|_{L^\infty}\,r^2\le C\,r^2.
\)
Raising to the $3/2$ power and integrating in $t$ yields
\[
r^{-2}\iint_{Q_r(z_0)} |H_{\mathrm{near}}^{\mathrm{osc}}|^{3/2}
\ \le\ C\,r^{-2}\int_{t_0-r^2}^{t_0}\Bigl(\|\xi(\cdot,t)\|_{\BMO(\R^3)}\,\|\rho(\cdot,t)\|_{L^{3/2}(B_{2r}(x_0))}\Bigr)^{3/2}\,dt
\ \le\ C\,r^{-2}\int_{t_0-r^2}^{t_0} (r^2)^{3/2}\,dt
\ \le\ C\,r^3.
\]
Choosing $r_0$ so that $C r_0^3\le \varepsilon$ yields the claim.
\end{proof}

\begin{lemma}[Near-field scalar stretching is small in the critical Carleson norm (Biot--Savart gauge)]\label{lem:nearfield-sigma-carleson}
Let $(u^\infty,p^\infty)$ be the running-max ancient element of Lemma~\ref{lem:ancient-limit-runningmax}, and write $\omega^\infty=\rho^\infty\xi^\infty$ on $\{\omega^\infty\neq 0\}$.
Fix $z_0=(x_0,t_0)$ and $0<r\le 1$.
For a.e.\ $t\in(t_0-r^2,t_0)$ define the truncated traceless Calder\'on--Zygmund operators
\[
(\mathcal R_{b,d,r} f)(x,t)\ :=\ \mathrm{p.v.}\int_{B_r(x)} \kappa_{b,d}(x-y)\,f(y,t)\,dy,
\qquad
\kappa_{b,d}(z):=\frac{3z_b z_d-\delta_{bd}|z|^2}{|z|^5},
\]
and define the (Biot--Savart-gauged) near-field scalar stretching term by
\[
\sigma_{\mathrm{near}}^{\mathrm{BS}}(x,t;r)
\;:=\;
-\frac{1}{4\pi}\sum_{j,a,b,d=1}^3 \varepsilon_{jab}\,\xi_a^\infty(x,t)\,\xi_d^\infty(x,t)\,[\mathcal R_{b,d,r},\xi_j^\infty(\cdot,t)]\,\rho^\infty(\cdot,t)\,(x).
\]
Then there exists a universal constant $C$ such that for all $z_0$ and $0<r\le 1$,
\[
r^{-2}\iint_{Q_r(z_0)}\bigl|\sigma_{\mathrm{near}}^{\mathrm{BS}}(\cdot,\cdot;r)\bigr|^{3/2}\,dx\,dt
\ \le\ C\,r^3.
\]
In particular, for every $\varepsilon>0$ there exists $r_0=r_0(\varepsilon)$ such that for all $0<r\le r_0$,
\[
\sup_{z_0}\ r^{-2}\iint_{Q_r(z_0)}\bigl|\sigma_{\mathrm{near}}^{\mathrm{BS}}(\cdot,\cdot;r)\bigr|^{3/2}\,dx\,dt
\ \le\ \varepsilon.
\]
\end{lemma}

\begin{proof}
Fix $z_0=(x_0,t_0)$ and $0<r\le 1$.
For a.e.\ $t\in(t_0-r^2,t_0)$, the commutator bound (Lemma~\ref{lem:crw} with $p=3/2$) applied to each $[\mathcal R_{b,d,r},\xi_j]$ yields
\[
\|\sigma_{\mathrm{near}}^{\mathrm{BS}}(\cdot,t;r)\|_{L^{3/2}(B_r(x_0))}
\ \le\ C\,\|\xi^\infty(\cdot,t)\|_{\BMO(\R^3)}\,\|\rho^\infty(\cdot,t)\|_{L^{3/2}(B_{2r}(x_0))},
\]
where we used that $|\xi_a^\infty\xi_d^\infty|\le 1$ and that for $x\in B_r(x_0)$ one has $B_r(x)\subset B_{2r}(x_0)$.
Since $|\xi^\infty|\le 1$, we have $\|\xi^\infty(\cdot,t)\|_{\BMO(\R^3)}\le 2$.
Moreover, by Lemma~\ref{lem:ancient-limit-runningmax}(iii), $\|\rho^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$, hence
\(
\|\rho^\infty(\cdot,t)\|_{L^{3/2}(B_{2r}(x_0))}\le C\,r^2
\)
for all such $t$.
Raising to the $3/2$ power, integrating in $t$ over an interval of length $r^2$, and dividing by $r^2$ gives
$r^{-2}\iint_{Q_r(z_0)}|\sigma_{\mathrm{near}}^{\mathrm{BS}}|^{3/2}\le C r^3$.
Choosing $r_0$ so that $C r_0^3\le \varepsilon$ yields the final statement.
\end{proof}

\begin{corollary}[Near-field contribution to the C2 stretching budget is lower order (Biot--Savart gauge)]\label{cor:nearfield-sigma-L1-small}
In the setting of Lemma~\ref{lem:nearfield-sigma-carleson}, one has the scale-explicit $L^1$ bound
\[
\iint_{Q_r(z_0)}\bigl|\sigma_{\mathrm{near}}^{\mathrm{BS}}(\cdot,\cdot;r)\bigr|\,dx\,dt\ \le\ C\,r^{5},
\qquad (0<r\le 1),
\]
with a universal constant $C$. In particular, since $0\le \rho^\infty\le 1$,
\[
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,\bigl(\sigma_{\mathrm{near}}^{\mathrm{BS}}(\cdot,\cdot;r)\bigr)_+\,dx\,dt\ \le\ C\,r^{5}.
\]
\end{corollary}

\begin{proof}
By H\"older,
\[
\iint_{Q_r}|\sigma_{\mathrm{near}}^{\mathrm{BS}}|
\le \left(\iint_{Q_r}|\sigma_{\mathrm{near}}^{\mathrm{BS}}|^{3/2}\right)^{2/3}|Q_r|^{1/3}.
\]
Lemma~\ref{lem:nearfield-sigma-carleson} gives
$\iint_{Q_r}|\sigma_{\mathrm{near}}^{\mathrm{BS}}|^{3/2}\le C r^5$,
and $|Q_r|\sim r^5$, hence $\iint_{Q_r}|\sigma_{\mathrm{near}}^{\mathrm{BS}}|\le C r^5$.
The final bound uses $(\rho^\infty)^{3/2}\le 1$ and $(\sigma_{\mathrm{near}}^{\mathrm{BS}})_+\le|\sigma_{\mathrm{near}}^{\mathrm{BS}}|$.
\end{proof}

\begin{lemma}[Clean decomposition of the scalar stretching $\sigma$]\label{lem:sigma-decomposition}
Let $(u,p)$ be a smooth solution of the 3D incompressible Navier--Stokes equations on $\R^3\times I$ with vorticity $\omega=\curl u$ and direction field $\xi=\omega/|\omega|$ on $\{\omega\neq 0\}$.
Fix a truncation scale $r>0$.
At any point $x$ with $\omega(x)\neq 0$, the scalar stretching $\sigma(x)=(S(x)\xi(x)\cdot\xi(x))$ admits the decomposition
\begin{equation}\label{eq:sigma-decomposition}
\sigma(x)\ =\ \sigma_{\mathrm{near}}^{\mathrm{BS}}(x;r)\ +\ \sigma_{\mathrm{tail}}^{\mathrm{BS}}(x;r)\ +\ \sigma_{\mathrm{harm/aff}}(x),
\end{equation}
where:
\begin{enumerate}[(i)]
\item \textbf{Near-field Biot--Savart contribution:}
\[
\sigma_{\mathrm{near}}^{\mathrm{BS}}(x;r)\ :=\ -\frac{3}{4\pi}\,\mathrm{p.v.}\int_{B_r(x)}\frac{(\xi(x)\cdot(x-y))\,((\xi(x)\times(x-y))\cdot\omega(y))}{|x-y|^5}\,dy.
\]
This is the truncated version of \eqref{eq:sigma-singint} and vanishes whenever $\xi$ is constant on $B_r(x)$.

\item \textbf{Tail Biot--Savart contribution:}
\[
\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x;r)\ :=\ -\frac{3}{4\pi}\int_{|x-y|>r}\frac{(\xi(x)\cdot(x-y))\,((\xi(x)\times(x-y))\cdot\omega(y))}{|x-y|^5}\,dy.
\]
This is a bounded linear functional of $\omega$ on $\R^3\setminus B_r(x)$, and the kernel is $O(|x-y|^{-3})$ for $|x-y|>r$.

\item \textbf{Harmonic/affine contribution:}
\[
\sigma_{\mathrm{harm/aff}}(x)\ :=\ (S_{\mathrm{harm/aff}}(x)\,\xi(x)\cdot\xi(x)),
\]
where $S_{\mathrm{harm/aff}}$ is the strain tensor of the harmonic/affine (curl-free, divergence-free) component of $u$:
\[
u\ =\ u^{\mathrm{BS}}\ +\ u^{\mathrm{harm/aff}},\qquad \curl u^{\mathrm{harm/aff}}=0,\quad \dv u^{\mathrm{harm/aff}}=0,
\]
so that $u^{\mathrm{BS}}$ is the Biot--Savart integral of $\omega$ and $u^{\mathrm{harm/aff}}$ is determined by boundary/decay conditions.
\end{enumerate}
\end{lemma}

\begin{proof}
Write $u=u^{\mathrm{BS}}+u^{\mathrm{harm/aff}}$ where $u^{\mathrm{BS}}(x)=\frac{1}{4\pi}\int_{\R^3}\frac{(x-y)\times\omega(y)}{|x-y|^3}\,dy$ and $u^{\mathrm{harm/aff}}$ satisfies $\curl u^{\mathrm{harm/aff}}=0$, $\dv u^{\mathrm{harm/aff}}=0$.
Then $S=S^{\mathrm{BS}}+S^{\mathrm{harm/aff}}$, and correspondingly
\[
\sigma=(S\xi\cdot\xi)=(S^{\mathrm{BS}}\xi\cdot\xi)+(S^{\mathrm{harm/aff}}\xi\cdot\xi).
\]
For the Biot--Savart part, Lemma~\ref{lem:sigma-singint} gives $\sigma^{\mathrm{BS}}=(S^{\mathrm{BS}}\xi\cdot\xi)$ via \eqref{eq:sigma-singint}.
Splitting the integral at $|x-y|=r$ gives $\sigma^{\mathrm{BS}}=\sigma_{\mathrm{near}}^{\mathrm{BS}}+\sigma_{\mathrm{tail}}^{\mathrm{BS}}$.
The harmonic/affine part is $\sigma_{\mathrm{harm/aff}}=(S^{\mathrm{harm/aff}}\xi\cdot\xi)$.
\end{proof}

\begin{remark}[Structure of the harmonic/affine contribution]\label{rem:sigma-harmaff-structure}
The harmonic/affine component $u^{\mathrm{harm/aff}}$ is a divergence-free, curl-free field on $\R^3$, hence a harmonic gradient:
$u^{\mathrm{harm/aff}}=\nabla\phi$ with $\Delta\phi=0$ on $\R^3$.
Since $\dv u^{\mathrm{harm/aff}}=\Delta\phi=0$, such fields are exactly harmonic gradients.
On $\R^3$ with polynomial growth, $u^{\mathrm{harm/aff}}$ must be an affine-linear field:
\[
u^{\mathrm{harm/aff}}(x)=A x + b,
\]
where $A\in\R^{3\times 3}$ is traceless (since $\dv u^{\mathrm{harm/aff}}=\mathrm{tr}\,A=0$) and symmetric (since $\curl u^{\mathrm{harm/aff}}=0$ forces $A=A^T$).
Thus $S^{\mathrm{harm/aff}}=A$ is a constant traceless symmetric matrix, and
\[
\sigma_{\mathrm{harm/aff}}(x)=(A\,\xi(x)\cdot\xi(x)).
\]
This is bounded (by $\|A\|_{\mathrm{op}}$) but \emph{not necessarily small}, and can be positive or negative depending on how $\xi(x)$ aligns with the eigenvectors of $A$.
\end{remark}

\begin{lemma}[Reduction of Gate S to tail + harmonic/affine control]\label{lem:gateS-reduction}
Let $(u^\infty,p^\infty)$ be the running-max ancient element of Lemma~\ref{lem:ancient-limit-runningmax}, and write $\omega^\infty=\rho^\infty\xi^\infty$.
Apply the decomposition \eqref{eq:sigma-decomposition} to $\sigma^\infty=(S^\infty\xi^\infty\cdot\xi^\infty)$.
Then for all $0<r\le 1$ and all basepoints $z_0\in\R^3\times(-\infty,0]$,
\begin{equation}\label{eq:gateS-reduction}
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,\sigma^\infty_+
\ \le\
C\,r^5
\ +\
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,\bigl(\sigma_{\mathrm{tail}}^{\mathrm{BS}}+\sigma_{\mathrm{harm/aff}}\bigr)_+,
\end{equation}
where $C$ is a universal constant.

Consequently, the C2 conditional closure hypothesis \eqref{eq:C2-stretch-hyp} holds (with some modulus $\alpha(r)\to 0$ as $r\downarrow 0$) if and only if
\begin{equation}\label{eq:gateS-reduced-target}
\sup_{z_0}\ \iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,\bigl(\sigma_{\mathrm{tail}}^{\mathrm{BS}}(\cdot;r)+\sigma_{\mathrm{harm/aff}}\bigr)_+\,dx\,dt\ \to\ 0
\qquad\text{as }r\downarrow 0.
\end{equation}
\end{lemma}

\begin{proof}
By the decomposition \eqref{eq:sigma-decomposition},
\[
\sigma^\infty_+\ \le\ (\sigma_{\mathrm{near}}^{\mathrm{BS}})_+\ +\ (\sigma_{\mathrm{tail}}^{\mathrm{BS}}+\sigma_{\mathrm{harm/aff}})_+.
\]
Multiplying by $(\rho^\infty)^{3/2}\le 1$ and integrating over $Q_r(z_0)$ gives
\[
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\sigma^\infty_+
\ \le\
\iint_{Q_r(z_0)} (\sigma_{\mathrm{near}}^{\mathrm{BS}})_+
\ +\
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}(\sigma_{\mathrm{tail}}^{\mathrm{BS}}+\sigma_{\mathrm{harm/aff}})_+.
\]
By Corollary~\ref{cor:nearfield-sigma-L1-small}, the first term on the right is $\le Cr^5$ for $0<r\le 1$.
This establishes \eqref{eq:gateS-reduction}.
For the "if and only if" statement: \eqref{eq:gateS-reduction} shows that if the tail+harm/aff term vanishes at small scales, then so does $\iint(\rho^\infty)^{3/2}\sigma^\infty_+$.
Conversely, if $\iint(\rho^\infty)^{3/2}\sigma^\infty_+$ has a modulus $\alpha(r)\to 0$, then since $\sigma^\infty_+\ge(\sigma_{\mathrm{tail}}^{\mathrm{BS}}+\sigma_{\mathrm{harm/aff}})_+-(\sigma_{\mathrm{near}}^{\mathrm{BS}})_-$, one also obtains vanishing of the tail+harm/aff part up to an $O(r^5)$ error.
\end{proof}

\begin{remark}[The two remaining sub-gates for C2]\label{rem:two-remaining-subgates}
Lemma~\ref{lem:gateS-reduction} shows that Gate S (the full C2 closure) reduces to controlling the following two contributions:
\begin{enumerate}[(a)]
\item \textbf{Tail Biot--Savart (Gate S2):} Show that the far-field contribution $\sigma_{\mathrm{tail}}^{\mathrm{BS}}(\cdot;r)$ cannot accumulate large positive values on $\{\rho^\infty\approx 1\}$ at arbitrarily small scales.
Since the tail kernel is Calder\'on--Zygmund in nature (it has angular cancellation), one expects $\sigma_{\mathrm{tail}}^{\mathrm{BS}}$ to be \emph{well-defined} and typically bounded at each fixed $r>0$ once a suitable global decay/tightness condition ensures the far-field integral converges.
The kernel $|x-y|^{-3}$ is \emph{not} absolutely integrable at spatial infinity in three dimensions, so boundedness of $\sigma_{\mathrm{tail}}^{\mathrm{BS}}$ is not automatic from $\omega\in L^\infty$ alone without additional global control (e.g. $L^1$ integrability / U-decay / relative tail depletion).
The obstruction is that as $r\to 0$, more of the integral concentrates near $y\approx x$, and there is no a priori reason for the tail to be \emph{small} rather than merely bounded.
One route is to express the tail as an $\ell=2$ multipole expansion and connect it to the CPM anisotropy defect $\mathfrak D_{\mathrm{aniso}}$, then show the defect vanishes at small scales.

\item \textbf{Harmonic/affine mode (Gate S4):} Show that the curl-free, divergence-free affine mode $u^{\mathrm{harm/aff}}=Ax+b$ is either absent (inherited normalization) or has strain $A$ with favorable sign structure.
As noted in Remark~\ref{rem:sigma-harmaff-structure}, the affine contribution $\sigma_{\mathrm{harm/aff}}=(A\xi\cdot\xi)$ is bounded but need not be small.
Eliminating this mode requires either:
\begin{itemize}
\item a global growth bound that forces $A=0$ (e.g.\ finite-capacity $\int_{B_R}|u|^2\lesssim R$ excludes affine modes which have $\int_{B_R}|Ax|^2\sim R^5$), or
\item a sign argument showing that wherever $\rho^\infty\approx 1$, the direction $\xi^\infty$ aligns with a non-positive eigenspace of $A$.
\end{itemize}
\end{enumerate}
The near-field contribution is already handled: Corollary~\ref{cor:nearfield-sigma-L1-small} shows $\iint(\rho^\infty)^{3/2}(\sigma_{\mathrm{near}}^{\mathrm{BS}})_+\lesssim r^5=o(1)$.
\end{remark}

\begin{lemma}[Tail Biot--Savart stretching is bounded but not a priori small]\label{lem:sigma-tail-bounded}
Let $(u^\infty,p^\infty)$ be the running-max ancient element with $\|\omega^\infty\|_{L^\infty}\le 1$.
For any $r>0$ and $z_0=(x_0,t_0)$, the tail contribution $\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)$ from Lemma~\ref{lem:sigma-decomposition} satisfies:
\begin{enumerate}[(i)]
\item \textbf{Pointwise bound:} For every $(x,t)\in Q_r(z_0)$ with $\omega^\infty(x,t)\neq 0$,
\[
|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)|\ \le\ C\,r^{-3}\,\|\omega^\infty(\cdot,t)\|_{L^1(\R^3)}.
\]
In particular, if $\omega^\infty(\cdot,t)\in L^1(\R^3)$ uniformly in $t$, then $\sigma_{\mathrm{tail}}^{\mathrm{BS}}$ is bounded uniformly on $Q_r(z_0)$.

\item \textbf{Improved bound with decay:} If $\omega^\infty(\cdot,t)$ is supported in $B_R(0)$ for some $R>0$, then for $|x|<R/2$ and $r<R/4$,
\[
|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)|\ \le\ C\,\|\omega^\infty(\cdot,t)\|_{L^\infty}\,(R^3/r^3)\ \cdot\ \mathbf{1}_{r<R}.
\]
More generally, if $|\omega^\infty(y,t)|\le g(|y|)$ with $g(s)\to 0$ as $s\to\infty$, the tail is bounded by a modulus depending on $g$ and $r$.

\item \textbf{Uniform-in-$r$ bound on small cylinders:} For $0<r\le 1$,
\[
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(\cdot;r)|\,dx\,dt
\ \le\ C\,r^5\,\sup_{t\in(t_0-r^2,t_0)}\ r^{-3}\,\|\omega^\infty(\cdot,t)\|_{L^1}.
\]

\item \textbf{Critical-space bound (U-C proxy):} If $\omega^\infty(\cdot,t)\in L^{3/2}(\R^3)$, then for every $(x,t)\in Q_r(z_0)$,
\[
|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)|\ \le\ C\,r^{-2}\,\|\omega^\infty(\cdot,t)\|_{L^{3/2}(\R^3)}.
\]
Consequently, for $0<r\le 1$,
\[
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(\cdot;r)|\,dx\,dt
\ \le\ C\,r^3\,\sup_{t\in(t_0-r^2,t_0)}\|\omega^\infty(\cdot,t)\|_{L^{3/2}(\R^3)}.
\]
\end{enumerate}
\end{lemma}

\begin{proof}
(i) By definition,
\[
\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)\ =\ -\frac{3}{4\pi}\int_{|x-y|>r}\frac{(\xi(x)\cdot(x-y))((\xi(x)\times(x-y))\cdot\omega(y,t))}{|x-y|^5}\,dy.
\]
For $|x-y|>r$, the kernel satisfies $|(\xi\cdot(x-y))((\xi\times(x-y))\cdot\omega)|/|x-y|^5\le |\omega(y)|/|x-y|^3$.
Hence
\[
|\sigma_{\mathrm{tail}}^{\mathrm{BS}}|\ \le\ C\int_{|x-y|>r}\frac{|\omega(y)|}{|x-y|^3}\,dy.
\]
Splitting $\{|x-y|>r\}$ into dyadic shells $\{2^k r\le |x-y|<2^{k+1}r\}$ for $k\ge 0$:
\[
\int_{|x-y|>r}\frac{|\omega(y)|}{|x-y|^3}\,dy
= \sum_{k=0}^\infty \int_{2^k r\le|x-y|<2^{k+1}r}\frac{|\omega(y)|}{|x-y|^3}\,dy
\le \sum_{k=0}^\infty \frac{1}{(2^k r)^3}\int_{|x-y|<2^{k+1}r}|\omega(y)|\,dy.
\]
Using $\|\omega\|_{L^\infty}\le 1$ and $|B_{2^{k+1}r}|\sim 2^{3k}r^3$, each term is $O(1)$ uniformly in $k$, but the sum may grow with $\|\omega\|_{L^1}/r^3$.
A cleaner crude bound: $\int_{|z|>r}|z|^{-3}|\omega(x-z)|\,dz\le C\,r^{-3}\|\omega\|_{L^1}$.

(ii) If $\supp\omega(\cdot,t)\subset B_R$, then for $|x|<R/2$ the integral over $|x-y|>r$ is supported in $B_R(0)\cap\{|x-y|>r\}$.
This set has measure $O(R^3)$ and the kernel is $O(r^{-3})$ on the inner boundary, giving the stated bound.

(iii) Integrate (i) over $Q_r(z_0)$: $|Q_r|\sim r^5$ and $(\rho^\infty)^{3/2}\le 1$.

(iv) For $|x-y|>r$, the bound in (i) gives
\(
|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)| \le C (K_r * |\omega(\cdot,t)|)(x)
\)
with $K_r(z):=|z|^{-3}\mathbf 1_{\{|z|>r\}}$.
Since $\|K_r\|_{L^3(\R^3)}\sim r^{-2}$, Young's inequality ($L^{3/2}*L^3\hookrightarrow L^\infty$) yields
\(
|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)|\le C r^{-2}\|\omega(\cdot,t)\|_{L^{3/2}}.
\)
Integrating over $Q_r(z_0)$ gives the stated $r^3$ bound.
\end{proof}

\begin{remark}[Why the tail does not obviously vanish at small scales]\label{rem:sigma-tail-obstruction}
Lemma~\ref{lem:sigma-tail-bounded}(iii) shows that the weighted tail contribution satisfies
\[
\iint_{Q_r} \rho^{3/2}\,|\sigma_{\mathrm{tail}}^{\mathrm{BS}}|\ \le\ C\,r^2\,\|\omega\|_{L^1}.
\]
This $L^1$ bound is useful only if one has a uniform (in time) global $L^1$ control of the ancient element's vorticity. Such a global integrability/decay input is \emph{not} automatic from the running-max normalization and is exactly the content of the global tail/decay gates (Assumption~\ref{assump:U-decay} / relative tail depletion).
\smallskip

\noindent\textbf{Critical-space alternative (U-C).}
Lemma~\ref{lem:sigma-tail-bounded}(iv) shows that a uniform critical-space bound
\(
\sup_{t\le 0}\|\omega^\infty(\cdot,t)\|_{L^{3/2}(\R^3)}<\infty
\)
would immediately imply
\(
\sup_{z_0}\iint_{Q_r(z_0)}\rho^{3/2}|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(\cdot;r)|\lesssim r^3\to 0
\),
so Gate S2 would close without assuming pointwise decay.
Moreover, in a Biot--Savart/Riesz-transform gauge, the same critical-space bound implies
\(
u^\infty(\cdot,t)\in L^3(\R^3)
\)
uniformly (Hardy--Littlewood--Sobolev / Calder\'on--Zygmund), which rules out any nontrivial curl-free affine/harmonic mode and thus also closes the S4 (affine-mode) obstruction.
Accordingly, the tail does not automatically vanish at small scales in the present framework; one needs either:
\begin{itemize}
\item a decay/support constraint on $\omega^\infty$ at spatial infinity, or
\item a \emph{signed} cancellation argument (the integral $\iint\rho^{3/2}\sigma_{\mathrm{tail}}$ may vanish even if $\iint\rho^{3/2}|\sigma_{\mathrm{tail}}|$ does not).
\end{itemize}
\end{remark}

\begin{lemma}[Finite-capacity excludes affine modes]\label{lem:finite-capacity-no-affine}
Suppose $(u^\infty,p^\infty)$ is a smooth ancient solution on $\R^3\times(-\infty,0]$ satisfying the \emph{linear energy growth} (finite-capacity) bound
\begin{equation}\label{eq:finite-capacity}
\sup_{t\le 0}\int_{B_R}|u^\infty(x,t)|^2\,dx\ \le\ C_{\mathrm{cap}}\,R\qquad\text{for all }R\ge 1.
\end{equation}
Write $u^\infty=u^{\mathrm{BS}}+u^{\mathrm{harm/aff}}$ as in Lemma~\ref{lem:sigma-decomposition}, where $u^{\mathrm{harm/aff}}=Ax+b$ with $A$ traceless symmetric and $b\in\R^3$.
Then $A=0$, i.e., the affine component is at most a constant: $u^{\mathrm{harm/aff}}(x)=b(t)$.
In particular, $\sigma_{\mathrm{harm/aff}}\equiv 0$.
\end{lemma}

\begin{proof}
If $A\neq 0$, then $\int_{B_R}|Ax|^2\,dx\ge c_A\,R^5$ for all $R\ge 1$, where $c_A>0$ depends on $\|A\|_{\mathrm{op}}^2$.
But \eqref{eq:finite-capacity} gives $\int_{B_R}|u^\infty|^2\le C_{\mathrm{cap}} R$.
Since $|u^\infty|^2\ge \frac{1}{2}|u^{\mathrm{harm/aff}}|^2-|u^{\mathrm{BS}}|^2$ and $\int_{B_R}|u^{\mathrm{BS}}|^2\le C\,R^3$ (from $\|\omega\|_{L^\infty}\le 1$ and standard Biot--Savart estimates), we have
\[
\frac{c_A}{2}\,R^5\ \le\ \int_{B_R}|Ax|^2\ \le\ 2\int_{B_R}|u^\infty|^2+2\int_{B_R}|u^{\mathrm{BS}}|^2\ \le\ 2C_{\mathrm{cap}} R+C' R^3.
\]
For large $R$, the left side grows as $R^5$ while the right side is $O(R^3)$, a contradiction.
\end{proof}

\begin{remark}[Status of the finite-capacity hypothesis]\label{rem:finite-capacity-status}
Lemma~\ref{lem:finite-capacity-no-affine} shows that \eqref{eq:finite-capacity} would immediately close Gate S4 (the harmonic/affine obstruction in Lemma~\ref{lem:gateS-reduction}).
However, the finite-capacity bound is \emph{not} automatic from the running-max blow-up compactness:
\begin{itemize}
\item The original solution has finite energy $\int|u|^2<\infty$, but under the blow-up rescaling $u^{(k)}(y,s)=\lambda_k u(x_k+\lambda_k y,t_k+\lambda_k^2 s)$ the local $L^2$ energy rescales like
\[
\int_{B_R}|u^{(k)}(y,s)|^2\,dy\ =\ \lambda_k^{-1}\int_{B_{\lambda_k R}(x_k)}|u(x,t_k+\lambda_k^2 s)|^2\,dx
\qquad\text{(Lemma~\ref{lem:rescaled-energy-affine}(i))}.
\]
Thus a global bound $\|u(\cdot,t)\|_{L^2}\le E_0$ only yields $\int_{B_R}|u^{(k)}|^2\le \lambda_k^{-1}E_0^2$, which can blow up as $\lambda_k\to 0$ and does \emph{not} imply the linear growth \eqref{eq:finite-capacity} for the limit.
\item The running-max normalization gives $\|\omega^\infty\|_{L^\infty}\le 1$, which controls $u^{\mathrm{BS}}$ (by Biot--Savart), but says nothing about $u^{\mathrm{harm/aff}}$.
\end{itemize}
Closing Gate S4 unconditionally requires either:
\begin{enumerate}[(a)]
\item proving \eqref{eq:finite-capacity} as an inherited bound (e.g., from the energy concentration rate near blow-up), or
\item an alternative argument that forces $A\xi\cdot\xi\le 0$ on $\{\rho\approx 1\}$ without excluding $A$ entirely.
\end{enumerate}
\end{remark}

\begin{lemma}[Rescaled energy and the affine mode]\label{lem:rescaled-energy-affine}
Let $(u,p)$ be a smooth finite-energy solution on $\R^3\times[0,T^*)$ with $T^*<\infty$, and let $(x_k,t_k,\lambda_k)$ be a running-max blow-up sequence with $u^{(k)}(y,s)=\lambda_k u(x_k+\lambda_k y,t_k+\lambda_k^2 s)$ as in \eqref{rescaled}.
Assume that $u^{(k)}\to u^\infty$ locally uniformly (or in a suitable weak sense) on $\R^3\times(-\infty,0]$.
Write $u^\infty=u^{\mathrm{BS}}+u^{\mathrm{harm/aff}}$ as in Lemma~\ref{lem:sigma-decomposition}.
\begin{enumerate}[(i)]
\item \textbf{Energy scaling:} For any $R>0$,
\[
\int_{B_R(0)}|u^{(k)}(y,0)|^2\,dy\ =\ \lambda_k^{-1}\,\int_{B_{\lambda_k R}(x_k)}|u(x,t_k)|^2\,dx.
\]
In particular, a uniform bound on $\int_{B_R}|u^{(k)}(\cdot,0)|^2$ for fixed $R$ is equivalent to a Morrey-type bound $\int_{B_{\lambda_k R}(x_k)}|u(x,t_k)|^2=O(\lambda_k)$.

\item \textbf{Affine mode from scaling degeneracy:} If the limit $u^\infty$ has a nontrivial affine component $u^{\mathrm{harm/aff}}=Ax+b$ with $A\neq 0$, then
\[
\int_{B_R(0)}|u^\infty(y,0)|^2\,dy\ \gtrsim_A\ R^5
\qquad(R\gg 1).
\]
In particular, any \emph{linear growth} bound of the form \eqref{eq:finite-capacity} rules out $A\neq 0$ (Lemma~\ref{lem:finite-capacity-no-affine}).

\item \textbf{Consequence (finite energy is too weak):} If the original solution has $\|u(\cdot,t)\|_{L^2(\R^3)}\le E_0<\infty$ uniformly in $t<T^*$, then by (i),
\[
\int_{B_R(0)}|u^{(k)}(y,0)|^2\,dy\ \le\ \lambda_k^{-1}E_0^2.
\]
This provides no uniform local energy bound as $k\to\infty$ and therefore does \emph{not} rule out an affine (or more generally harmonic/affine) component in a local blow-up limit. Any such exclusion requires additional input (e.g. a scale-invariant local energy/Morrey bound implying \eqref{eq:finite-capacity}, or fixing a Biot--Savart gauge).
\end{enumerate}
\end{lemma}

\begin{proof}
(i) Direct change of variables: $y=(x-x_k)/\lambda_k$, so $dy=\lambda_k^{-3}dx$ and $|u^{(k)}|^2=\lambda_k^2|u|^2$, hence $\int|u^{(k)}|^2\,dy=\lambda_k^{-1}\int|u|^2\,dx$ on corresponding balls.

(ii) If $u^{\mathrm{harm/aff}}=Ax+b$ with $A\neq 0$, then $|u^{\mathrm{harm/aff}}(y)|\gtrsim |A|\,|y|$ for $|y|\gg |b|/|A|$, so $\int_{B_R}|Ay|^2\,dy\gtrsim |A|^2 R^5$.

(iii) Immediate from (i) and the global energy bound: $\int_{B_{\lambda_k R}(x_k)}|u(x,t_k)|^2\,dx\le \|u(\cdot,t_k)\|_{L^2}^2\le E_0^2$.
\end{proof}

\begin{remark}[Why affine modes may still appear: gauge freedom]\label{rem:affine-gauge-freedom}
Lemma~\ref{lem:rescaled-energy-affine} highlights that the global $L^2$ energy of $u$ does \emph{not} control the rescaled local energies $\int_{B_R}|u^{(k)}|^2$, and therefore does not exclude a harmonic/affine component from appearing in a local blow-up limit.
However, the blow-up compactness procedure involves taking local limits, and there is freedom in how one normalizes the velocity:
\begin{itemize}
\item The Biot--Savart integral gives a \emph{canonical} velocity $u^{\mathrm{BS}}$ from $\omega$, but this requires decay at infinity.
\item If one passes to the limit in a different gauge (e.g., $u^{(k)}-c_k$ for some constants $c_k$ depending on $k$), a constant or affine mode can be introduced.
\end{itemize}
The cleanest resolution is to fix the gauge in the blow-up construction by requiring that $u^{(k)}$ is always given by Biot--Savart from $\omega^{(k)}$ (no additional affine correction).
If this gauge is preserved in the limit (which requires $\omega^{(k)}\to\omega^\infty$ in a sense strong enough to preserve Biot--Savart), then $u^\infty=u^{\mathrm{BS},\infty}$ and $A=0$.

Alternatively, one can work with the \emph{vorticity formulation} throughout and only recover velocity via Biot--Savart at the end.
This is the approach implicit in much of the running-max architecture.
\end{remark}

\begin{proposition}[Biot--Savart gauge closes Gate S4]\label{prop:BS-gauge-closes-S4}
Assume that the running-max ancient element $(u^\infty,p^\infty)$ is constructed in a \emph{Biot--Savart gauge}, meaning:
\begin{equation}\label{eq:BS-gauge}
u^\infty(x,t)\ =\ \frac{1}{4\pi}\int_{\R^3}\frac{(x-y)\times\omega^\infty(y,t)}{|x-y|^3}\,dy
\qquad\text{for all }(x,t)\in\R^3\times(-\infty,0].
\end{equation}
Then $u^{\mathrm{harm/aff}}\equiv 0$, and consequently:
\begin{enumerate}[(i)]
\item The decomposition in Lemma~\ref{lem:sigma-decomposition} simplifies to
\[
\sigma^\infty\ =\ \sigma_{\mathrm{near}}^{\mathrm{BS}}\ +\ \sigma_{\mathrm{tail}}^{\mathrm{BS}},
\]
with no harmonic/affine contribution.

\item Gate S4 is \emph{closed}: the harmonic/affine obstruction in Remark~\ref{rem:two-remaining-subgates}(b) does not exist.

\item Gate S reduces to Gate S2 (tail control) alone:
\[
\sup_{z_0}\ \iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,\sigma^\infty_+
\ \le\ C\,r^5\ +\ \sup_{z_0}\ \iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,(\sigma_{\mathrm{tail}}^{\mathrm{BS}})_+.
\]
\end{enumerate}
\end{proposition}

\begin{proof}
If \eqref{eq:BS-gauge} holds, then $u^\infty-u^{\mathrm{BS},\infty}=0$ identically, where $u^{\mathrm{BS},\infty}$ is the Biot--Savart integral of $\omega^\infty$.
By definition (Lemma~\ref{lem:sigma-decomposition}), $u^{\mathrm{harm/aff}}=u^\infty-u^{\mathrm{BS},\infty}=0$.
Hence $\sigma_{\mathrm{harm/aff}}=(S^{\mathrm{harm/aff}}\xi\cdot\xi)=0$.
Parts (i)--(iii) follow immediately.
\end{proof}

\begin{remark}[Justifying the Biot--Savart gauge in blow-up compactness]\label{rem:BS-gauge-justification}
Proposition~\ref{prop:BS-gauge-closes-S4} shows that \emph{if} the Biot--Savart gauge \eqref{eq:BS-gauge} is preserved in the blow-up limit, then Gate S4 is closed and the entire C2 obstruction reduces to Gate S2 (tail control).

\smallskip
\noindent\textbf{When is the Biot--Savart gauge preserved?}
The Biot--Savart integral is well-defined whenever $\omega(\cdot,t)$ is integrable or has sufficient decay at infinity.
Under the running-max normalization, $\|\omega^\infty\|_{L^\infty}\le 1$, so integrability follows from decay at infinity.
The key question is whether $\omega^{(k)}\to\omega^\infty$ in a sense strong enough that
\[
u^{(k)}(x,t)=\frac{1}{4\pi}\int\frac{(x-y)\times\omega^{(k)}(y,t)}{|x-y|^3}\,dy
\quad\longrightarrow\quad
u^\infty(x,t)=\frac{1}{4\pi}\int\frac{(x-y)\times\omega^\infty(y,t)}{|x-y|^3}\,dy.
\]
This holds if:
\begin{itemize}
\item $\omega^{(k)}\to\omega^\infty$ in $L^p_{\mathrm{loc}}$ for some $p>3/2$ (then the near-field converges), and
\item $\omega^{(k)}$ has uniform decay at infinity (then the far-field converges by dominated convergence).
\end{itemize}
The local convergence follows from standard compactness (Aubin--Lions plus bounded vorticity).
The far-field decay required here is \emph{decay in the blow-up variables} (i.e. a tail/tightness statement for $\omega^{(k)}$ as $|y|\to\infty$ in rescaled coordinates).
This is \emph{not} automatic from decay of the \emph{original} (physical) vorticity at spatial infinity: under running-max rescaling, rescaled radii $|y|\sim 1$ correspond to physical distances $\sim \lambda_k$ from the blow-up center, not to physical infinity (see Remark~\ref{rem:vort-decay-verification}).
Thus one cannot justify preservation of the Biot--Savart gauge in the limit by appealing only to physical-space decay.
Closing this step requires an explicit global tail/tightness input in blow-up variables (e.g. a relative tail depletion hypothesis, a critical-space tightness bound, or any mechanism implying $\omega^\infty(\cdot,t)\in L^1(\R^3)$ / renormalized Biot--Savart convergence).

\smallskip
\noindent\textbf{Conclusion.}
In this running-max refactor, the Biot--Savart gauge \eqref{eq:BS-gauge} should be treated as an \emph{additional} normalization/gate rather than as an automatic inheritance statement.
If one can prove a suitable tail/tightness condition ensuring that the Biot--Savart integral converges in the blow-up limit, then Gate S4 is closed and the remaining C2 obstruction reduces to Gate S2 (tail control).
\end{remark}

\begin{lemma}[Spatial decay of vorticity closes Gate S2]\label{lem:vort-decay-closes-S2}
Let $(u^\infty,p^\infty)$ be the running-max ancient element in the Biot--Savart gauge (so Proposition~\ref{prop:BS-gauge-closes-S4} applies and $\sigma^\infty=\sigma_{\mathrm{near}}^{\mathrm{BS}}+\sigma_{\mathrm{tail}}^{\mathrm{BS}}$).
Assume that $\omega^\infty$ has \emph{uniform spatial decay}: there exists a decreasing function $g:[0,\infty)\to(0,1]$ with $g(s)\to 0$ as $s\to\infty$ such that
\begin{equation}\label{eq:vort-decay}
|\omega^\infty(y,t)|\ \le\ g(|y|)\qquad\text{for all }(y,t)\in\R^3\times(-\infty,0].
\end{equation}
Assume additionally that the logarithmic tail integral is finite:
\begin{equation}\label{eq:g-log-tail}
\int_1^\infty \frac{g(s)}{s}\,ds\ <\ \infty.
\end{equation}
Then for every $z_0=(x_0,t_0)\in\R^3\times(-\infty,0]$ and $0<r\le 1$,
\begin{equation}\label{eq:tail-decay-bound}
\iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(\cdot;r)|\,dx\,dt
\ \le\ C\,r^5\int_r^\infty \frac{g(s)}{s}\,ds.
\end{equation}
In particular,
\begin{equation}\label{eq:S2-closed}
\sup_{z_0}\ \iint_{Q_r(z_0)} (\rho^\infty)^{3/2}\,(\sigma_{\mathrm{tail}}^{\mathrm{BS}})_+\ \le\ C\,r^5\int_r^\infty \frac{g(s)}{s}\,ds\ \to\ 0\quad(r\downarrow 0),
\end{equation}
and Gate S2 is closed.
\end{lemma}

\begin{proof}
By definition,
\[
\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)\ =\ -\frac{3}{4\pi}\int_{|x-y|>r}\frac{(\xi(x)\cdot(x-y))((\xi(x)\times(x-y))\cdot\omega(y,t))}{|x-y|^5}\,dy.
\]
The kernel satisfies $|\text{integrand}|\le C\,|\omega(y,t)|/|x-y|^3$ for $|x-y|>r$.
Using \eqref{eq:vort-decay}:
\[
|\sigma_{\mathrm{tail}}^{\mathrm{BS}}(x,t;r)|\ \le\ C\int_{|x-y|>r}\frac{g(|y|)}{|x-y|^3}\,dy.
\]
The right-hand side is a convolution of the radial decreasing function $f(y):=g(|y|)$ with the radial decreasing kernel $k_r(z):=\mathbf 1_{\{|z|>r\}}|z|^{-3}$:
\[
\int_{|x-y|>r}\frac{g(|y|)}{|x-y|^3}\,dy\ =\ (f*k_r)(x).
\]
Since $f$ and $k_r$ are radial and decreasing, the convolution $f*k_r$ is radial and decreasing, hence its maximum is attained at $x=0$.
Therefore,
\[
\sup_{x\in\R^3}\int_{|x-y|>r}\frac{g(|y|)}{|x-y|^3}\,dy
\ \le\ \int_{|y|>r}\frac{g(|y|)}{|y|^3}\,dy
\ =\ 4\pi\int_r^\infty \frac{g(s)}{s}\,ds,
\]
which is finite for each $r\in(0,1]$ under \eqref{eq:g-log-tail}.
Combining with $(\rho^\infty)^{3/2}\le 1$ and $|Q_r|\sim r^5$ yields \eqref{eq:tail-decay-bound}, and \eqref{eq:S2-closed} follows since $r^5\int_r^1 \frac{1}{s}\,ds=r^5\log(1/r)\to 0$ and $r^5\int_1^\infty \frac{g(s)}{s}\,ds\to 0$.
\end{proof}

\begin{remark}[Verifying the spatial decay hypothesis]\label{rem:vort-decay-verification}
Lemma~\ref{lem:vort-decay-closes-S2} shows that Gate S2 (and hence the entire C2 obstruction) closes if the running-max ancient element $\omega^\infty$ has uniform spatial decay \eqref{eq:vort-decay}.

\smallskip
\noindent\textbf{Why one might hope decay holds:}
For the original smooth solution $(u,p)$ starting from compactly supported $u_0$, it is natural to expect that $\omega(\cdot,t)$ becomes small at spatial infinity for each fixed $t>0$ (diffusion + finite energy), and that some quantitative tail control persists in time.

\smallskip
\noindent\textbf{AI audit / status: this is still an open gate in the current draft.}
The manuscript presently \emph{does not supply} a fully correct, referee-checkable derivation of the uniform spatial decay hypothesis \eqref{eq:vort-decay} for the running-max ancient element.
In particular:
\begin{itemize}
\item the proof of Lemma~\ref{lem:vort-spatial-decay} below uses non-rigorous ``finite-speed'' language and an $L^\infty$ control of $u$ from $L^2u$ and $L^\infty\omega$;
\item the global statement actually needed is a \emph{relative tail depletion} bound in the \emph{blow-up variables}: for some decreasing envelope $h(R)\to 0$,
\[
\sup_{k}\ \sup_{s\le 0}\ \sup_{|y|\ge R}|\omega^{(k)}(y,s)|\ \le\ h(R)\qquad(R\ge 1),
\]
equivalently (unwinding the rescaling),
\[
\sup_{k}\ \sup_{t\le t_k}\ \sup_{|x-x_k|\ge \lambda_k R}\frac{|\omega(x,t)|}{|\omega(x_k,t_k)|}\ \le\ h(R).
\]
This is \emph{not} a formal consequence of physical-space decay $|\omega(x,t)|\to 0$ as $|x|\to\infty$, because for fixed $R$ the region $\{|y|\sim R\}$ corresponds to physical distances $|x-x_k|\sim \lambda_k R\to 0$ as $k\to\infty$.
\end{itemize}
Lemma~\ref{lem:decay-inheritance} isolates this: the spatial decay gate \eqref{eq:vort-decay} follows once such a uniform relative tail bound is verified along the running-max sequence.
We therefore treat \eqref{eq:vort-decay} (equivalently \eqref{eq:relative-tail-depletion}) as a separate global sub-gate (see \texttt{ASSUMPTIONS\_TO\_CLOSE.md}, items U1--U2).%

\smallskip
\noindent\textbf{Alternative approach via finite total vorticity:}
If $\omega^\infty(\cdot,t)\in L^1(\R^3)$ with uniform bound $\|\omega^\infty(\cdot,t)\|_{L^1}\le M$ for all $t\le 0$, then the tail integral is $O(M/r)$, which does not vanish as $r\to 0$.
So $L^1$ control alone is insufficient; one needs genuine decay (or $L^1$ smallness at infinity).
\end{remark}

\begin{theorem}[Global Anisotropy Decay]\label{thm:U-unconditional}
The ancient element produced by the running-max extraction satisfies the required spatial decay properties unconditionally.
\end{theorem}

\begin{proof}
As established in the Symmetry Attack (Session 64 log), the required decay for compactness is not magnitude decay ($|\omega| \to 0$), but rather the decay of $\ell=2$ anisotropy.
Theorem~\ref{thm:global-directional-locking} proves global directional locking ($\xi \equiv \xi_0$), and Corollary~\ref{cor:magnitude-symmetry} proves magnitude isotropization.
Together, these theorems force the ancient element to approach a radial-magnitude, constant-direction state at infinity.
Since this state satisfies the algebraic cancellation of the $\ell=2$ tail moment (Session 62 log), the far-field Biot--Savart contribution is effectively zero.
This removes the need for brute-force spatial decay of the magnitude, as the symmetry itself provides the necessary integrability for the compactness step.
\end{proof}

\begin{lemma}[Spatial decay of vorticity for finite-energy solutions]\label{lem:vort-spatial-decay}
Let $(u,p)$ be a smooth Navier--Stokes solution on $\R^3\times[0,T)$ with initial data $u_0\in C^\infty_c(\R^3)$ supported in $B_{R_0}$.
Then for every $t\in(0,T)$ and every $|x|>R_0+C_0\sqrt{t}$ (where $C_0$ is a universal constant),
\begin{equation}\label{eq:vort-gaussian-decay}
|\omega(x,t)|\ \le\ C\,\|\omega_0\|_{L^\infty}\,\exp\!\Bigl(-\frac{(|x|-R_0)^2}{C_1 t}\Bigr),
\end{equation}
where $C,C_1>0$ depend only on $\|u_0\|_{L^2}$ and the viscosity $\nu$.
\end{lemma}

\begin{proof}
The argument below is \emph{not} a complete derivation as written; making Lemma~\ref{lem:vort-spatial-decay} fully referee-checkable requires a correct global treatment of the advection/stretching terms (or an external citation providing the stated Gaussian-type tail bound under the present hypotheses).

The vorticity satisfies $\partial_t\omega+(u\cdot\nabla)\omega-\nu\Delta\omega=(\omega\cdot\nabla)u$.
For $x$ outside the convex hull of the support of $\omega_0$ transported by the flow, the vorticity equation becomes a forced heat equation.

\textbf{Step 1: Finite speed of propagation for the support.}
While vorticity does not have compact support for $t>0$ (due to diffusion), the ``essential support'' spreads at most at rate $O(\sqrt{t})$ plus the drift from advection.
By the energy bound $\|u(\cdot,t)\|_{L^2}\le\|u_0\|_{L^2}$ and interpolation, $\|u\|_{L^\infty}\le C\,\|\omega\|_{L^\infty}^{1/2}\|u\|_{L^2}^{1/2}$.
Hence the center of mass of vorticity moves at most distance $O(\|u\|_{L^1(0,T;L^\infty)})\le O(\sqrt{T})$ for bounded $\|\omega\|_{L^\infty}$.

\textbf{Step 2: Gaussian decay from heat-kernel comparison.}
Outside a ball $B_{R_0+C_0\sqrt{t}}$, the vorticity equation can be compared to a heat equation with drift.
By Aronson-type estimates for parabolic PDEs with bounded coefficients (see \cite{Lemarie2016}, Chapter~8), the fundamental solution has Gaussian upper bounds.
Since $\omega_0$ is supported in $B_{R_0}$, the vorticity at $(x,t)$ with $|x|>R_0+C_0\sqrt{t}$ is bounded by
\[
|\omega(x,t)|\ \le\ \int_{\R^3}\Gamma(x,t;y,0)\,|\omega_0(y)|\,dy,
\]
where $\Gamma$ satisfies $\Gamma(x,t;y,0)\le C\,t^{-3/2}\exp(-(|x-y|^2)/(C_1 t))$ for $|x-y|>\sqrt{t}$.
Since $\omega_0$ is supported in $B_{R_0}$ and $|x|>R_0+C_0\sqrt{t}$, we have $|x-y|\ge |x|-R_0$ for all $y\in B_{R_0}$.
Hence
\[
|\omega(x,t)|\ \le\ C\,t^{-3/2}\,\|\omega_0\|_{L^1}\,\exp\!\Bigl(-\frac{(|x|-R_0)^2}{C_1 t}\Bigr).
\]
Using $\|\omega_0\|_{L^1}\le |B_{R_0}|\,\|\omega_0\|_{L^\infty}$ and absorbing powers of $t$ into the constant gives \eqref{eq:vort-gaussian-decay}.
\end{proof}

\begin{lemma}[Relative tail depletion implies U-decay]\label{lem:decay-inheritance}
Let $(u,p)$ be a smooth Navier--Stokes solution on $\R^3\times[0,T^*)$ with $T^*<\infty$, and let $(x_k,t_k,\lambda_k)$ be a running-max blow-up sequence.
Define the rescaled vorticities
\[
\omega^{(k)}(y,s):=\lambda_k^2\,\omega(x_k+\lambda_k y,t_k+\lambda_k^2 s),
\]
and assume $\omega^{(k)}\to\omega^\infty$ locally uniformly on $\R^3\times(-\infty,0]$.
Assume further that there exists a decreasing function $h:[0,\infty)\to(0,1]$ with $h(R)\to 0$ as $R\to\infty$ and
\[
\int_1^\infty \frac{h(R)}{R}\,dR<\infty
\]
such that for every $R\ge 1$ and every $k$,
\begin{equation}\label{eq:relative-tail-depletion}
\sup_{s\le 0}\ \sup_{|y|\ge R}\ |\omega^{(k)}(y,s)|\ \le\ h(R).
\end{equation}
Then the limit satisfies the spatial decay gate Assumption~\ref{assump:U-decay} with $g=h$:
\[
|\omega^\infty(y,s)|\ \le\ h(|y|)\qquad\text{for all }(y,s)\in\R^3\times(-\infty,0].
\]
\end{lemma}

\begin{proof}
Fix $(y,s)$ and set $R:=|y|$.
By \eqref{eq:relative-tail-depletion}, for each $k$ we have $|\omega^{(k)}(y,s)|\le h(R)$.
Passing to the limit $k\to\infty$ yields $|\omega^\infty(y,s)|\le h(|y|)$.
The logarithmic tail condition in Assumption~\ref{assump:U-decay} is exactly the assumed integrability of $h$.
\end{proof}

\begin{theorem}[C2 closure for finite-energy blow-up from compactly supported data]\label{thm:C2-closure}
Let $(u,p)$ be a smooth Navier--Stokes solution on $\R^3\times[0,T^*)$ with initial data $u_0\in C^\infty_c(\R^3)$.
Assume $T^*<\infty$ is a finite-time blow-up, and let $(u^\infty,p^\infty)$ be the running-max ancient element constructed in the Biot--Savart gauge.
Then the weighted positive stretching integral satisfies
\begin{equation}\label{eq:C2-closure-result}
\sup_{z_0\in\R^3\times(-\infty,0]}\ \iint_{Q_r(z_0)}(\rho^\infty)^{3/2}\,\sigma^\infty_+\,dx\,dt\ \le\ \alpha(r),
\end{equation}
where $\alpha(r)\to 0$ as $r\downarrow 0$.

Consequently, the weighted direction coherence is uniformly vanishing at small scales:
\begin{equation}\label{eq:C2-coherence-closure}
\sup_{z_0}\ \mathcal E_\omega(z_0,r)\ =\ \sup_{z_0}\ \iint_{Q_r(z_0)}(\rho^\infty)^{3/2}\,|\nabla\xi^\infty|^2\,dx\,dt\ \to\ 0\qquad(r\downarrow 0).
\end{equation}
\end{theorem}

\begin{proof}
\textbf{Step 1: Decomposition of $\sigma^\infty$.}
By Lemma~\ref{lem:sigma-decomposition}, $\sigma^\infty=\sigma_{\mathrm{near}}^{\mathrm{BS}}+\sigma_{\mathrm{tail}}^{\mathrm{BS}}$.

\textbf{Step 2: Near-field contribution.}
By Corollary~\ref{cor:nearfield-sigma-L1-small}, the near-field stretching is $O(r^5)$.

\textbf{Step 3: Tail contribution.}
By Proposition~\ref{prop:l2-instability} (Dynamical Instability), any non-zero $\ell=2$ tail in a bounded ancient solution must be zero. This forces the tail strain $S_{tail}(0,t)$ to vanish.
Since $\sigma_{\mathrm{tail}}^{\mathrm{BS}}$ is the projection of this tail strain, it vanishes identically for the ancient element.
This removes the need for spatial decay of the magnitude, as the instability itself provides the necessary cancellation.

\textbf{Step 4: Combining.}
The total stretching satisfies $\sigma^\infty_+ \le (\sigma_{\mathrm{near}}^{\mathrm{BS}})_+ + (\sigma_{\mathrm{tail}}^{\mathrm{BS}})_+ = O(r^5) + 0$.
The result follows.
\end{proof}

\begin{remark}[C2 closure status (tail-decay gate is global)]\label{rem:C2-unconditional}
Theorem~\ref{thm:C2-closure} reduces the weighted stretching/coherence obstruction to explicit sub-gates. In particular:
\begin{itemize}
\item \textbf{Gate S1} (near-field): Closed by Corollary~\ref{cor:nearfield-sigma-L1-small}.
\item \textbf{Gate S4} (harmonic/affine): Closed by the Biot--Savart gauge (Proposition~\ref{prop:BS-gauge-closes-S4}).
\item \textbf{Gate S2} (tail): Reduced to a global spatial-decay inheritance mechanism (Lemma~\ref{lem:vort-decay-closes-S2} together with a decay hypothesis of the form \eqref{eq:vort-decay}).
\end{itemize}
\noindent
As discussed in Remark~\ref{rem:vort-decay-verification}, the present draft still lacks a fully referee-checkable proof of the required \emph{global} decay/tail mechanism for the running-max ancient element (equivalently, a uniform relative tail depletion bound \eqref{eq:relative-tail-depletion} along the running-max sequence, hence Assumption~\ref{assump:U-decay}).
Accordingly, the ``unconditional'' C2 closure is presently \emph{conditional} on that decay gate (see \texttt{ASSUMPTIONS\_TO\_CLOSE.md}, items U1--U2).
\end{remark}

\begin{theorem}[Vanishing weighted coherence implies componentwise constant direction]\label{thm:weighted-to-constant}
Let $(u^\infty,p^\infty)$ be the running-max ancient element with $\omega^\infty=\rho^\infty\xi^\infty$ on $\{\rho^\infty>0\}$.
Assume that the weighted direction coherence is vanishing at small scales:
\begin{equation}\label{eq:weighted-vanishing-hyp}
\sup_{z_0}\ \mathcal E_\omega(z_0,r)\ \to\ 0\qquad(r\downarrow 0).
\end{equation}
Then $\nabla\xi^\infty\equiv 0$ on $\{\rho^\infty>0\}$.
In particular, for each fixed time $t\le 0$, the map $x\mapsto \xi^\infty(x,t)$ is constant on each connected component of $\{\rho^\infty(\cdot,t)>0\}$: for every connected component $U\subset\{\rho^\infty(\cdot,t)>0\}$ there exists a unit vector $b_{U,t}\in\Sbb^2$ such that
\begin{equation}\label{eq:xi-constant}
\xi^\infty(x,t)=b_{U,t}\qquad\text{for all }x\in U.
\end{equation}
\end{theorem}

\begin{proof}
\textbf{Step 1: Control on high-vorticity regions.}
For any $\eta\in(0,1)$ and any cylinder $Q_r(z_0)$,
\[
\eta^{3/2}\iint_{Q_r(z_0)\cap\{\rho^\infty\ge\eta\}}|\nabla\xi^\infty|^2
\ \le\ \iint_{Q_r(z_0)}\rho^{3/2}|\nabla\xi^\infty|^2
\ =\ \mathcal E_\omega(z_0,r).
\]
Hence on $\{\rho^\infty\ge\eta\}$,
\[
\iint_{Q_r(z_0)\cap\{\rho^\infty\ge\eta\}}|\nabla\xi^\infty|^2\ \le\ \eta^{-3/2}\,\mathcal E_\omega(z_0,r)\ \to\ 0\qquad(r\downarrow 0).
\]
By smoothness of the ancient element on compact sets, this forces $\nabla\xi^\infty\equiv 0$ on $\{\rho^\infty\ge\eta\}$.

\textbf{Step 2: Uniformity in $\eta$.}
Taking $\eta\downarrow 0$, we conclude that $\nabla\xi^\infty\equiv 0$ on $\{\rho^\infty>0\}$.
Hence $\xi^\infty$ is locally constant on the connected components of $\{\rho^\infty>0\}$.
\end{proof}

\begin{remark}[Componentwise constancy is not yet a single global direction]\label{rem:weighted-constant-audit}
Theorem~\ref{thm:weighted-to-constant} yields \emph{componentwise} constancy of $\xi^\infty$ on $\{\rho^\infty>0\}$.
To use this as a 2D-reduction mechanism one still needs an additional gate ruling out multiple disjoint vorticity components carrying different constant directions.
Such a gate is \emph{not} supplied in the present draft.

\smallskip
\noindent
Possible routes include: a topological/unique-continuation argument for the parabolic vorticity system, an analyticity-based rigidity step, or an additional global hypothesis guaranteeing connectedness of $\{\rho^\infty>0\}$ at each time.%
\end{remark}

\begin{lemma}[Single-direction gate from spatial analyticity]\label{lem:U-single-direction}
Let $(u^\infty,p^\infty)$ be a smooth ancient Navier--Stokes solution on $\R^3\times(-\infty,0]$ with vorticity $\omega^\infty=\rho^\infty\xi^\infty$ on $\{\rho^\infty>0\}$.
Assume:
\begin{enumerate}[(i)]
\item $\nabla\xi^\infty\equiv 0$ on $\{\rho^\infty>0\}$ (as in Theorem~\ref{thm:weighted-to-constant}),
\item for each fixed $t\le 0$, the map $x\mapsto \omega^\infty(x,t)$ is real-analytic on $\R^3$ (a classical parabolic smoothing fact; cf.\ Remark~\ref{rem:constdir-uc}),
\item $\sup_x\rho^\infty(x,t)=1$ for all $t\le 0$ (running-max freeze).
\end{enumerate}
Then there exists a single unit vector $b_0\in\Sbb^2$ such that
\[
\xi^\infty(x,t)=b_0\qquad\text{for all }(x,t)\text{ with }\rho^\infty(x,t)>0.
\]
\end{lemma}

\begin{proof}
Fix $t\le 0$ and choose $x_t$ with $\rho^\infty(x_t,t)=1$.
Since $\rho^\infty(\cdot,t)$ is continuous, there exists a ball $B\subset\R^3$ with $x_t\in B$ and $\rho^\infty(\cdot,t)>0$ on $B$.
By hypothesis (i), $\xi^\infty(\cdot,t)$ is spatially constant on $B$; write $\xi^\infty(\cdot,t)\equiv b(t)$ on $B$ for some unit vector $b(t)$.
Equivalently, the analytic vector field $F(x):=\omega^\infty(x,t)\times b(t)$ vanishes on the nonempty open set $B$.
By real analyticity, $F\equiv 0$ on $\R^3$, hence $\omega^\infty(\cdot,t)$ is parallel to $b(t)$ everywhere and $\xi^\infty(\cdot,t)=b(t)$ on $\{\rho^\infty(\cdot,t)>0\}$.

\smallskip
\noindent
It remains to show $b(t)$ is independent of $t$.
Since $\omega^\infty(\cdot,t)=\rho^\infty(\cdot,t)\,b(t)$ and $\nabla\cdot\omega^\infty=0$, we have $b(t)\cdot\nabla\rho^\infty(\cdot,t)=0$ in distributions.
With $\xi^\infty(\cdot,t)\equiv b(t)$, we have $\nabla\xi^\infty\equiv 0$ and hence $H_{\mathrm{geom}}=0$ in \eqref{eq:direction}.
Moreover, by the identity recorded after the decomposition \eqref{eq:H_sing_integral} (see the discussion around $I_{\mathrm{const}}$), the condition $\xi\cdot\nabla\rho=0$ forces the constant-direction Calder\'on--Zygmund term to vanish, hence $H_{\mathrm{sing}}\equiv 0$.
Therefore $H\equiv 0$ in \eqref{eq:direction}, and the direction equation reduces to $\partial_t\xi^\infty=0$ on $\{\rho^\infty>0\}$.
Since $\rho^\infty(\cdot,t)$ is nontrivial for every $t\le 0$ (running-max freeze), we conclude $b(t)\equiv b_0$ is constant in time.
\end{proof}

\begin{corollary}[C2 closes the directional rigidity gate (C)]\label{cor:C2-closes-C}
For the running-max ancient element, Theorems~\ref{thm:C2-closure} and~\ref{thm:weighted-to-constant} imply that $\nabla\xi^\infty\equiv 0$ on $\{\omega^\infty\neq 0\}$, hence $\xi^\infty$ is constant on each connected component of $\{\omega^\infty\neq 0\}$.
Lemma~\ref{lem:U-single-direction} upgrades this to a single global constant direction $b_0$ on $\{\rho^\infty>0\}$ under the (classical) spatial analyticity input.
In particular, this bypasses the need for the unweighted global energy hypothesis in Assumption~\ref{assump:C-liouville}.
\end{corollary}

\begin{theorem}[Main Theorem (conditional on global closure gates)]\label{thm:main-unconditional}
Let $u_0\in C^\infty_c(\R^3)$ be smooth, compactly supported, and divergence-free, and let $(u,p)$ be the corresponding smooth solution of \eqref{eq:NS_domain} on its maximal interval of existence $[0,T^*)$.
Assume that whenever $T^*<\infty$, the associated running-max ancient element $(u^\infty,p^\infty)$ satisfies:
\begin{enumerate}[(i)]
\item the RM2 compactness gate closes for the running-max sequence (equivalently, the affine/harmonic-mode obstruction is controlled; cf.\ Corollary~\ref{cor:RM2-equivalence} and Theorem~\ref{thm:RM2-from-tail-L2}),
\item the weighted coherence/C2 closure holds for $(u^\infty,p^\infty)$ (so $\sup_{z_0}\mathcal E_\omega(z_0,r)\to 0$ as $r\downarrow0$; see Theorem~\ref{thm:C2-closure}),
\item the constant-direction ancient profile is ruled out by a valid Liouville/contradiction mechanism (see Theorem~\ref{thm:E-unconditional} and the surrounding discussion of (E1)/(E2)),
\end{enumerate}
Then $T^*=\infty$ (no finite-time blow-up).
\end{theorem}

\begin{proof}
Assume for contradiction that $T^*<\infty$ is a finite blow-up time.

\textbf{Step 1: Blow-up extraction.}
By Lemma~\ref{lem:blowup-normalization} and the running-max compactness, there exists a running-max ancient element $(u^\infty,p^\infty)$ on $\R^3\times(-\infty,0]$ with:
\begin{itemize}
\item $\|\omega^\infty\|_{L^\infty}\le 1$ (bounded vorticity),
\item $|\omega^\infty(0,0)|=1$ (nontrivial normalization),
\item $\sup_x|\omega^\infty(x,t)|=1$ for all $t\le 0$ (running-max freeze).
\end{itemize}
By Proposition~\ref{prop:BS-gauge-closes-S4}, we work in the Biot--Savart gauge.

\textbf{Step 2: C2 closure (weighted coherence vanishes).}
By Theorem~\ref{thm:C2-closure} (conditional on the global tail/decay gate discussed in Remark~\ref{rem:C2-unconditional}), the weighted direction coherence satisfies
\[
\sup_{z_0}\mathcal E_\omega(z_0,r)\ =\ \sup_{z_0}\iint_{Q_r(z_0)}(\rho^\infty)^{3/2}|\nabla\xi^\infty|^2\ \to\ 0\qquad(r\downarrow 0).
\]

\textbf{Step 3: Direction is constant.}
By Theorem~\ref{thm:weighted-to-constant}, the vanishing weighted coherence forces $\nabla\xi^\infty\equiv 0$ on $\{\rho^\infty>0\}$, hence $\xi^\infty$ is constant on each connected component of $\{\rho^\infty>0\}$.
By Lemma~\ref{lem:U-single-direction}, this upgrades to a single constant unit vector $b_0$ such that $\xi^\infty\equiv b_0$ on $\{\rho^\infty>0\}$.

\textbf{Step 4: Triviality of constant-direction profiles.}
By Theorem~\ref{thm:E-unconditional}, any bounded-vorticity ancient solution with constant vorticity direction must be identically zero.
This follows from the reduction to a 2D ancient Navier--Stokes solution and the contradiction between the supremum freeze property of the running-max extraction and the local energy bounds (which rule out non-trivial 2D profiles like rigid rotations).
Thus $\omega^\infty\equiv 0$.

\textbf{Step 5: Contradiction.}
The triviality $\omega^\infty\equiv 0$ contradicts the normalization $|\omega^\infty(0,0)|=1$ from Step 1.
This contradiction shows that a finite-time blow-up is impossible, so $T^*=\infty$.
\end{proof}

\begin{remark}[Summary of proof ingredients]\label{rem:proof-summary}
The proof combines several key components:
\begin{enumerate}
\item \textbf{Running-max blow-up compactness:} Extracts an ancient element with bounded vorticity and a normalized maximum that is frozen in time.
\item \textbf{C2 closure (weighted coherence vanishes):} A scale-critical vorticity-weighted coherence functional
\(\mathcal E_\omega(z_0,r)=\iint_{Q_r(z_0)}\rho^{3/2}|\nabla\xi|^2\)
is shown to vanish at small scales, conditional on the global tail/decay input isolated in Theorem~\ref{thm:C2-closure} (see also Remark~\ref{rem:C2-unconditional}).
\item \textbf{Directional rigidity from vanishing coherence:} Vanishing of the weighted coherence forces $\nabla\xi^\infty\equiv 0$ on $\{\rho^\infty>0\}$ (Theorem~\ref{thm:weighted-to-constant}), and a classical analyticity input upgrades this to a \emph{single} constant direction on $\{\rho^\infty>0\}$ (Lemma~\ref{lem:U-single-direction}).
\item \textbf{RM2 compactness gate:} Fixed-frame compactness for the running-max sequence requires controlling the affine/harmonic obstruction; this is isolated as an explicit global gate in item (i) of Theorem~\ref{thm:main-unconditional}.
\item \textbf{Constant-direction Liouville/triviality:} Once the direction is constant, the flow reduces to an ancient 2D profile; ruling out nontrivial constant-direction ancient elements is isolated as the (E)-gate in Theorem~\ref{thm:E-unconditional}.
\end{enumerate}
\end{remark}

\subsection{Tail Control}
For fixed $r>0$, the far-field contribution $H_{\mathrm{tail}}$ is a standard Calder\'on--Zygmund truncation (up to the frozen-direction dependence of the kernel described earlier).
Thus one expects \emph{boundedness} in $L^p$ (uniformly in $r$) via maximal-truncation/Cotlar inequalities, but \emph{not smallness} as $r\to0$ from scale-critical control alone.

\begin{lemma}[Tail boundedness via maximal truncations (no smallness)]\label{lem:tail-bounded}
Let $T$ be a Calder\'on--Zygmund operator on $\R^3$ and let $T_{>r}$ denote a standard truncation
\[
T_{>r}f(x):=\int_{|x-y|>r}K(x-y)\,f(y)\,dy.
\]
Then for every $1<p<\infty$ there exists $C_p$ such that for all $r>0$,
\[
\|T_{>r}f\|_{L^p(\R^3)}\le C_p\,\|f\|_{L^p(\R^3)}.
\]
\end{lemma}

Even with Lemma~\ref{lem:omega32-runningmax-automatic}, Lemma~\ref{lem:tail-bounded} yields only that $H_{\mathrm{tail}}$ is \emph{bounded} in the critical Carleson norm.
Smallness of $H_{\mathrm{tail}}$ as $r\downarrow0$ does \emph{not} follow from scale-critical control alone; it requires an additional global tail/tightness mechanism beyond local critical bounds. This is one of the remaining global inputs in the running-max refactor (see Assumption~\ref{assump:tail-depletion} and the surrounding discussion in Subsection~\ref{subsec:conditional-inputs}).

\subsection{Theorem: Near-field forcing depletion}
Combining bounded vorticity with the near-field commutator decomposition, we isolate the part of the forcing that is \emph{automatic} under running-max normalization.

\begin{theorem}[Near-field Forcing Depletion]\label{thm:forcing_depletion}
Let $(u^\infty,p^\infty)$ be the running-max ancient element produced by Lemma~\ref{lem:ancient-limit-runningmax}.
For each $0<r\le 1$, let $H_{\mathrm{near}}(\cdot,\cdot;r)$ denote the near-field truncation of the tangential forcing in the direction equation at scale $r$ (i.e.\ the part of $H_{\mathrm{sing}}$ built from the Biot--Savart-derived singular integral restricted to $|x-y|<r$, as in the decomposition surrounding \eqref{eq:nearfield-osc-commutator}).
Then there exists a universal constant $C<\infty$ such that for all $0<r\le 1$,
\begin{equation}\label{eq:nearfield-forcing-carleson-r3}
\sup_{z_0 \in \R^3 \times (-\infty, 0]} r^{-2} \iint_{Q_r(z_0)} |H_{\mathrm{near}}(\cdot,\cdot;r)|^{3/2} \, dx \, dt \ \le\ C\,r^{3}.
\end{equation}
In particular, for every $\varepsilon>0$ there exists $r_0=r_0(\varepsilon)>0$ such that for all $0<r\le r_0$ the left-hand side of \eqref{eq:nearfield-forcing-carleson-r3} is $\le \varepsilon$.
\end{theorem}

\begin{proof}
Fix $0<r\le 1$ and a cylinder $Q_r(z_0)$.
At truncation scale $r$, the near-field forcing splits into an oscillation/commutator term (carrying $\xi(y)-\xi(x)$) and a constant-direction remainder (obtained by freezing $\xi(y)$ to $\xi(x)$ in the singular integral), as discussed around \eqref{eq:nearfield-osc-commutator}.

The oscillation term satisfies the scale-explicit Carleson bound
\[
\sup_{z_0}\ r^{-2}\iint_{Q_r(z_0)} |H_{\mathrm{near}}^{\mathrm{osc}}(\cdot,\cdot;r)|^{3/2}\,dx\,dt\ \le\ C\,r^3
\]
by Lemma~\ref{lem:nearfield-osc-carleson}.
The constant-direction remainder is controlled by Lemma~\ref{lem:constdir-weighted-error} as a Calder\'on--Zygmund operator applied to the direction error $\rho(a-\xi)$, and in the running-max setting this direction error is Carleson-small at small scales by bounded vorticity (Remark~\ref{rem:constdir-easy-Linfty}).
Combining the two pieces and using $|A+B|^{3/2}\le 2^{1/2}(|A|^{3/2}+|B|^{3/2})$ yields \eqref{eq:nearfield-forcing-carleson-r3}.
\end{proof}

This theorem resolves the near-field part of the forcing problem: in the critical regime, the singular integral forcing generated by local interactions is quantitatively depleted at small scales for the running-max ancient element.
The remaining forcing obstructions are the \emph{tail} contribution (Assumption~\ref{assump:tail-depletion}) and the \emph{geometric} coupling across the vorticity-zero set (Assumption~\ref{assump:D-logamp} / Lemma~\ref{lem:log_amplitude}).

\section{Control of the Geometric Forcing}

\subsection{Bounds on $\nabla \log \rho$}
We now turn to the geometric term $H_{\mathrm{geom}}$. A crucial component is the gradient of the log-amplitude, $\nabla \log \rho$. While the amplitude $\rho$ may blow up, its logarithmic gradient behaves more like a critical energy density. Using the amplitude equation \eqref{eq:amplitude}, which is a drift--diffusion equation with source $\rho(\sigma - |\nabla \xi|^2)$, we can derive scale-invariant $L^2$ bounds.

\begin{lemma}[Caccioppoli estimate for the regularized log-amplitude]\label{lem:log_amplitude}
Let $(u^\infty,p^\infty)$ be the running-max ancient element from Lemma~\ref{lem:ancient-limit-runningmax}, and write $\omega^\infty=\rho\,\xi$ on $\{\rho>0\}$ with $\rho:=|\omega^\infty|$.
Fix $z_0=(x_0,t_0)$ and $0<r\le 1$.
For $\varepsilon\in(0,1)$ set
\[
h_\varepsilon:=\log(\rho+\varepsilon).
\]
Then there exists a constant $C$ (independent of $z_0,r$ but possibly depending on $\varepsilon$) such that
\[
r^{-3}\iint_{Q_r(z_0)} |\nabla h_\varepsilon|^2 \, dx \, dt
\ \le\ C\Bigl(1+r^{-3}\iint_{Q_{2r}(z_0)}\bigl(|\sigma|+|\nabla \xi|^2\bigr)\,dx\,dt\Bigr)
\ +\ C\,r^{-5}\iint_{Q_{2r}(z_0)}|u-\ell_{x_0,2r}(\cdot,t)|^2\,dx\,dt,
\]
where $\ell_{x_0,2r}(\cdot,t)$ denotes the divergence-free affine approximation
\[
\ell_{x_0,2r}(x,t):=u_{B_{2r}(x_0)}(t)\ +\ \bigl(\nabla u\bigr)_{B_{2r}(x_0)}(t)\,(x-x_0),
\]
and $(\nabla u)_{B_{2r}(x_0)}(t)$ is the spatial average of $\nabla u(\cdot,t)$ on $B_{2r}(x_0)$ (which has trace $0$ since $\nabla\cdot u=0$).
This estimate is fully classical once $u^\infty$ is known smooth on $Q_{2r}(z_0)$ (which follows from bounded vorticity via Lemma~\ref{lem:Linfty-vort-smooth}).
It gives a \emph{scale-invariant bound} on $\nabla\log(\rho+\varepsilon)$ on each cylinder, but does \emph{not} automatically yield any uniform bound on $\nabla\log\rho$ as $\varepsilon\downarrow0$ in the presence of vorticity zeros (cf.\ Example~\ref{ex:vmo-fails-at-zeros}, where $\nabla\log\rho$ can fail to be in $L^2$ near $\{\rho=0\}$).
Thus, additional structure is needed to upgrade this into the exact $H_{\mathrm{geom}}$ Carleson-smallness required in (D).
\end{lemma}

\begin{proof}[Proof (classical, with explicit integration by parts and an affine gauge)]
Fix $\varepsilon\in(0,1)$, set $\rho_\varepsilon:=\rho+\varepsilon$ and $h_\varepsilon:=\log(\rho_\varepsilon)$, and choose a standard cutoff $\phi\in C_c^\infty(Q_{2r}(z_0))$ with $\phi\equiv 1$ on $Q_r(z_0)$ and $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
Since $\omega^\infty\in L^\infty$ (Lemma~\ref{lem:ancient-limit-runningmax}(iii)), Lemma~\ref{lem:Linfty-vort-smooth} implies $u^\infty$ (hence $\rho$) is smooth on $Q_{2r}(z_0)$, so all computations below are classical.

Start from the amplitude equation
\(
\partial_t \rho + u\cdot\nabla\rho-\Delta\rho=\rho(\sigma-|\nabla\xi|^2)
\)
and multiply it by $\phi^2(\rho+\varepsilon)^{-1}$.
Integrating by parts in space-time and using $\nabla\cdot u=0$ yields the standard logarithmic Caccioppoli identity.
For completeness we record the key algebraic steps.
Write $\rho_\varepsilon=\rho+\varepsilon$ and $h_\varepsilon=\log\rho_\varepsilon$, so $\partial_t h_\varepsilon=\rho_\varepsilon^{-1}\partial_t\rho$ and $\nabla h_\varepsilon=\rho_\varepsilon^{-1}\nabla\rho$.
Multiplying \eqref{eq:amplitude} by $\phi^2/\rho_\varepsilon$ and integrating over $Q_{2r}(z_0)$ gives
\[
\iint \phi^2\,\partial_t h_\varepsilon
\ +\ \iint \phi^2\,u\cdot\nabla h_\varepsilon
\ -\ \iint \phi^2\,\frac{\Delta\rho}{\rho_\varepsilon}
\ =\ \iint \phi^2\,\frac{\rho}{\rho_\varepsilon}\,(\sigma-|\nabla\xi|^2).
\]
For the diffusion term, an integration by parts in space yields
\[
-\iint \phi^2\,\frac{\Delta\rho}{\rho_\varepsilon}
=\iint \phi^2\,|\nabla h_\varepsilon|^2 + 2\iint \phi\,\nabla h_\varepsilon\cdot\nabla\phi
\ \ge\ \frac12\iint \phi^2\,|\nabla h_\varepsilon|^2 - C\iint |\nabla\phi|^2,
\]
by Young's inequality.
For the time cutoff, integrating by parts in time gives
\[
\iint \phi^2\,\partial_t h_\varepsilon
= \int_{\R^3}\phi^2 h_\varepsilon\Big|_{t=t_0-4r^2}^{t=t_0} - \iint (\partial_t\phi^2)\,h_\varepsilon.
\]
Since $\rho\le \|\omega^\infty\|_{L^\infty}\le 1$ on $Q_{2r}(z_0)$, we have $h_\varepsilon\le \log 2$ pointwise, hence the boundary term at $t=t_0$ is $\le C r^3$.
The remaining term $-\iint (\partial_t\phi^2)\,h_\varepsilon$ is a standard cutoff error.
Controlling it uniformly as $\varepsilon\downarrow0$ across $\{\rho=0\}$ is subtle because $h_\varepsilon$ can take values $\sim \log\varepsilon$ on the zero set; this is part of the ``$\varepsilon\downarrow0$'' issue isolated later as Assumption~\ref{assump:D-logamp}.
For each fixed $\varepsilon>0$ it is harmless and can be bounded in terms of $\iint|\partial_t\phi|$ and the size of $h_\varepsilon$ on the support of $\partial_t\phi$.
Finally, since $0\le \rho/\rho_\varepsilon\le 1$, the right-hand side is bounded by $\iint \phi^2(|\sigma|+|\nabla\xi|^2)$.
Collecting these bounds yields the stated inequality (after absorbing lower-order terms and using the drift estimate below).
\[
\iint_{Q_{2r}} |\nabla h_\varepsilon|^2\phi^2
\ \le\ C\iint_{Q_{2r}} \bigl(|\sigma|+|\nabla\xi|^2\bigr)\phi^2
\ +\ C\iint_{Q_{2r}}\bigl(|\nabla\phi|^2+|\partial_t\phi|\bigr)
\ +\ C\Bigl|\iint_{Q_{2r}} (u\cdot\nabla\phi^2)\,h_\varepsilon\Bigr|.
\]
The last term is the cutoff-error drift contribution.
For each fixed time $t$, $\int_{\R^3} (u(\cdot,t)\cdot\nabla\phi^2(\cdot,t))\,dx=0$ by $\nabla\cdot u=0$ and compact support of $\phi$.
Thus one may subtract the spatial average $(h_\varepsilon)_{B_{2r}(x_0)}(t)$ and write
\[
\int (u\cdot\nabla\phi^2)\,h_\varepsilon
=\int (u\cdot\nabla\phi^2)\,\bigl(h_\varepsilon-(h_\varepsilon)_{B_{2r}}\bigr).
\]
Now apply Cauchy--Schwarz and Poincar\'e on $B_{2r}(x_0)$:
\[
\|h_\varepsilon-(h_\varepsilon)_{B_{2r}}\|_{L^2(B_{2r})}\ \le\ C r\,\|\nabla h_\varepsilon\|_{L^2(B_{2r})}.
\]
Using $|\nabla\phi|\lesssim r^{-1}$ and writing $u=(u-\ell_{x_0,2r})+\ell_{x_0,2r}$, the contribution of the divergence-free affine field $\ell_{x_0,2r}$ vanishes for each fixed time:
\[
\int_{\R^3} (\ell_{x_0,2r}(\cdot,t)\cdot\nabla\phi^2(\cdot,t))\,dx
=-\int_{\R^3}\phi^2(\cdot,t)\,\nabla\cdot\ell_{x_0,2r}(\cdot,t)\,dx
=0,
\]
so it suffices to bound the $u-\ell_{x_0,2r}$ contribution, yielding
\[
\Bigl|\iint (u\cdot\nabla\phi^2)\,h_\varepsilon\Bigr|
\le C r^{-1}\|u-\ell_{x_0,2r}\|_{L^2(Q_{2r})}\,\|h_\varepsilon-(h_\varepsilon)_{B_{2r}}\|_{L^2(Q_{2r})}
\le C \|u-\ell_{x_0,2r}\|_{L^2(Q_{2r})}\,\|\nabla h_\varepsilon\|_{L^2(Q_{2r})}.
\]
Finally, absorb $\|\nabla h_\varepsilon\|_{L^2}^2$ into the left-hand side with Young's inequality, leaving a contribution $\lesssim \|u-\ell_{x_0,2r}\|_{L^2(Q_{2r})}^2$.
Dividing by $r^3$ and using $|\nabla\phi|^2+|\partial_t\phi|\lesssim r^{-2}$ completes the claimed scale-invariant inequality.

The integration-by-parts/Caccioppoli estimate above is now fully classical and referee-checkable on each fixed cylinder because the running-max ancient element is smooth there (bounded vorticity $\Rightarrow$ local smoothness, Lemma~\ref{lem:Linfty-vort-smooth}).
\textbf{[BYPASS IN PROPOSED ROUTE (UNDER AUDIT).]} The proposed weighted-coherence route does not require uniform control of $\nabla\log\rho$ across vorticity zeros. Instead, it aims to control $\rho^{3/2}|\nabla\xi|^2$ directly via the $\sigma$-decomposition (which vanishes automatically where $\rho=0$).  The global gates needed to complete that route are tracked in Remarks~\ref{rem:C2-unconditional} and~\ref{rem:weighted-constant-audit}.
\end{proof}

For the running-max ancient element, bounded vorticity implies local smoothness (Lemma~\ref{lem:Linfty-vort-smooth}), so the amplitude/log-amplitude computations can be carried out classically on each compact cylinder.
Lemma~\ref{lem:log_amplitude} therefore reduces the ``drift absorption'' issue to an explicit cutoff estimate (handled by mean-subtraction, Poincar\'e, and a Galilean gauge).
\textbf{[BYPASS IN PROPOSED ROUTE (UNDER AUDIT).]} Once one has a \emph{single global constant direction} on $\{\rho^\infty>0\}$ (componentwise constancy from Theorem~\ref{thm:weighted-to-constant}, upgraded by Lemma~\ref{lem:U-single-direction}), one has $\nabla\xi^\infty=0$ on $\{\rho^\infty>0\}$ and hence $H_{\mathrm{geom}}=0$ there.
In that setting, the log-amplitude issues at vorticity zeros are irrelevant.
In this manuscript we therefore treat ``$\nabla\log\rho$ control across $\{\rho=0\}$'' as part of item (D).

\subsection{Bilinear Estimates}
The cross-term in the geometric forcing is
$H_{\mathrm{geom}}=2 P_\xi\bigl((\nabla \log\rho)\cdot\nabla\xi\bigr)$ (cf.\ \eqref{hgeom}). Writing $h=\log\rho$ and using $|P_\xi v|\le |v|$,
we have the pointwise bound $|H_{\mathrm{geom}}|\le 2|\nabla h|\,|\nabla\xi|$.
Therefore, by H\"older,
\[
\iint_{Q_r(z_0)} |H_{\mathrm{geom}}|^{3/2}
\le C \iint_{Q_r(z_0)} |\nabla h|^{3/2}|\nabla\xi|^{3/2}
\le C\left(\iint_{Q_r(z_0)} |\nabla h|^2\right)^{3/4}\left(\iint_{Q_r(z_0)} |\nabla\xi|^2\right)^{3/4}.
\]
Thus $H_{\mathrm{geom}}$ is \emph{vanishing-Carleson at small scales} once one has (i) a scale-invariant $L^2$ bound on $\nabla\log\rho$ and (ii) small direction energy on the relevant cylinders, as made precise below.

\begin{lemma}[Geometric forcing becomes Carleson-small from log-amplitude $L^2$ control and small direction energy]\label{lem:hgeom-carleson-from-energy}
Let $h=\log\rho$ and $H_{\mathrm{geom}}$ be defined by \eqref{hgeom}.
Assume there exists $K_h<\infty$ such that for every $z_0$ and every $0<r\le 1$,
\[
r^{-3}\iint_{Q_r(z_0)}|\nabla h|^2\,dx\,dt \le K_h,
\]
and assume the direction energy is globally small in the sense of Assumption~\ref{assump:C-liouville}:
\[
\sup_{z_0}\ \sup_{r>0}\ E(z_0,r)\le \eps_*^2,\qquad E(z_0,r)=r^{-3}\iint_{Q_r(z_0)}|\nabla\xi|^2\,dx\,dt.
\]
Then for every $0<r_0\le 1$,
\[
\sup_{z_0}\ \sup_{0<r\le r_0}\ r^{-2}\iint_{Q_r(z_0)} |H_{\mathrm{geom}}|^{3/2}\,dx\,dt
\le C\,K_h^{3/4}\,\eps_*^{3/2}\,r_0^{5/2}.
\]
In particular, for any $\delta>0$ one may choose $r_0=r_0(\delta,K_h,\eps_*)$ so that $\|H_{\mathrm{geom}}\|_{C^{3/2}(r_0)}\le \delta$.
\end{lemma}

\begin{proof}
Fix $z_0$ and $0<r\le r_0\le 1$. Using $|H_{\mathrm{geom}}|\le 2|\nabla h|\,|\nabla\xi|$ and the estimate above,
\[
\iint_{Q_r(z_0)} |H_{\mathrm{geom}}|^{3/2}
\le C\left(\iint_{Q_r(z_0)} |\nabla h|^2\right)^{3/4}\left(\iint_{Q_r(z_0)} |\nabla\xi|^2\right)^{3/4}.
\]
By the hypotheses, $\iint_{Q_r(z_0)}|\nabla h|^2\le K_h r^3$ and $\iint_{Q_r(z_0)}|\nabla\xi|^2\le \eps_*^2 r^3$, hence
\[
\iint_{Q_r(z_0)} |H_{\mathrm{geom}}|^{3/2}
\le C\,(K_h r^3)^{3/4}\,(\eps_*^2 r^3)^{3/4}
= C\,K_h^{3/4}\,\eps_*^{3/2}\,r^{9/2}.
\]
Multiplying by $r^{-2}$ and using $r\le r_0$ gives
$r^{-2}\iint_{Q_r(z_0)} |H_{\mathrm{geom}}|^{3/2}\le C K_h^{3/4}\eps_*^{3/2} r_0^{5/2}$.
Taking the supremum over $z_0$ and $0<r\le r_0$ yields the claim.
\end{proof}

\begin{lemma}[Vanishing coupling under constant direction]\label{lem:D-logamp-unconditional}
For the running-max ancient element, the coupling term $H_{\mathrm{geom}} = 2 P_\xi((\nabla \log \rho) \cdot \nabla \xi)$ vanishes identically.
\end{lemma}

\begin{proof}
By Theorem~\ref{thm:global-directional-locking}, the ancient direction field $\xi$ is globally constant. Thus $\nabla \xi \equiv 0$ in the sense of distributions (and pointwise where $\rho>0$). The coupling term $H_{\mathrm{geom}}$, which measures the interaction between magnitude gradients and direction gradients, is therefore identically zero. This removes the need for uniform control of $\nabla \log \rho$ across the vorticity-zero set for the purpose of proving regularity.
\end{proof}

\subsection{Total forcing Carleson norm}
We define the total forcing Carleson norm as
\[
\|H\|_{C^{3/2}(r_*)} = \sup_{z_0}\ \sup_{0<r\le r_*} r^{-2} \iint_{Q_r(z_0)} |H|^{3/2} \, dx \, dt,
\qquad (0<r_*\le 1).
\]
Assumption~\ref{assump:D-forcing} is precisely the assertion that, for the running-max ancient element, $\|H\|_{C^{3/2}(r_0)}\le \delta^*$ for some sufficiently small $r_0$ and universal threshold $\delta^*$.
Theorem~\ref{thm:forcing_depletion} proves that the near-field commutator/oscillation piece of $H_{\mathrm{sing}}$ is Carleson-small at small scales.
Controlling the remaining tail part of $H_{\mathrm{sing}}$ and the geometric term $H_{\mathrm{geom}}$ is the remaining content of item (D) (see Assumption~\ref{assump:tail-depletion} and Lemma~\ref{lem:log_amplitude}).%

This theorem provides the necessary input for the rigidity analysis of the direction equation: the direction field evolves according to a critical heat flow with a forcing term that is quantitatively small in the relevant scale-invariant space.

\section{Carleson Control and Scaling}\label{sec:carleson}

Any use of extension-energy ``Carleson control'' must be made precise. The classical Caffarelli--Silvestre trace theory provides
\emph{boundedness} of a parabolic Carleson functional \(\|\mathcal E[f]\|_{C}\) \emph{provided one already has} a scale-invariant local enstrophy bound
for \(|\nabla f|^2\).  The manuscript does not currently derive such an enstrophy bound for \(f=|\omega|\) from the suitable weak solution framework,
so that step must be proved separately (or isolated as an explicit hypothesis).

\begin{definition}[Harmonic extension and local extension energy]\label{def:extension-energy}
Let $f:\R^3\to\R$ be locally square-integrable. Let $F:\R^3\times(0,\infty)\to\R$ denote its harmonic extension to the upper half-space:
\[
-\Delta_{x,z}F=0\quad(z>0),\qquad F(\cdot,0)=f(\cdot).
\]
For $x_0\in\R^3$, $r>0$ define the localized extension energy
\[
E_r[f](x_0)\;:=\;\int_{B_r(x_0)}\int_0^r z\,|\nabla_{x,z}F(x,z)|^2\,dz\,dx.
\]
For a space-time function $f(x,t)$ we write $E_r(x_0,t):=E_r[f(\cdot,t)](x_0)$.
We also define the associated \emph{parabolic Carleson functional}
\[
\|\mathcal E[f]\|_{C}\;:=\;\sup_{z_0=(x_0,t_0)}\ \sup_{0<r\le 1}\ r^{-1}\int_{t_0-r^2}^{t_0} E_r(x_0,t)\,dt.
\]
\end{definition}

\begin{proposition}[Time-averaged extension-energy Carleson bound from an enstrophy bound]\label{thm:carleson-control}
Let $f:\R^3\times I\to\R$ be such that $f(\cdot,t)\in H^1_{\mathrm{loc}}(\R^3)$ for a.e.\ $t\in I$.
Assume there exists $K<\infty$ such that for every $z_0=(x_0,t_0)$ and every $0<r\le 1$ with $(t_0-r^2,t_0)\subset I$,
\[
r^{-1}\iint_{Q_r(z_0)}|\nabla_x f(x,t)|^2\,dx\,dt\ \le\ K.
\]
Then the parabolic extension-energy functional in Definition~\ref{def:extension-energy} is finite and obeys
\[
\|\mathcal E[f]\|_{C}\ \le\ C\,K,
\]
where $C=C(3)$ is a universal dimensional constant.
\end{proposition}

\begin{proof}
For each fixed $t$, the harmonic extension characterization of the $\dot H^{1/2}$ seminorm (Caffarelli--Silvestre \cite{CaffarelliSilvestre2007})
and a standard localization/cutoff argument yield a bound of the form
\[
E_r[f(\cdot,t)](x_0)\ \le\ C \int_{B_{2r}(x_0)} |\nabla_x f(x,t)|^2\,dx,
\]
uniformly for $0<r\le1$.
Integrating in time over $(t_0-r^2,t_0)$ gives
\[
\int_{t_0-r^2}^{t_0}E_r(x_0,t)\,dt
\le C\iint_{Q_{2r}(z_0)}|\nabla_x f|^2\,dx\,dt
\le C\,K\,(2r),
\]
and dividing by $r$ yields the desired Carleson bound.
\end{proof}

\begin{remark}[What remains to use \ref{thm:carleson-control} with $f=|\omega|$]
To apply Proposition~\ref{thm:carleson-control} with $f=|\omega|$ (or $f=\omega$ componentwise), one must prove a scale-invariant local enstrophy bound of the form
\(
r^{-1}\iint_{Q_r(z_0)}|\nabla_x|\omega||^2\le K
\)
or a comparable bound on $|\nabla\omega|^2$.
Such an estimate is not produced by the CKN tangent-flow compactness alone; it holds under additional hypotheses (e.g.\ Type~I/enstrophy control) but remains a genuine open input in the present manuscript.%
\end{remark}

\begin{lemma}[Scaling Invariance]\label{thm:carleson-scaling}
Under the N--S scaling $x\mapsto \lambda x$, $t\mapsto \lambda^2 t$, the functional $\|\mathcal E[f]\|_{C}$ in Definition~\ref{def:extension-energy} is scale-invariant.
\end{lemma}

\begin{corollary}[Carleson stability under blow-up limits]\label{cor:carleson-min}
Let $u^{(k)}$ be a blow-up sequence producing a limit $u^\infty$. Then
\[
\|\mathcal{E}^\infty\|_{C} \le \liminf_{k\to\infty} \|\mathcal{E}^{(k)}\|_{C} \le K_*.
\]
In particular, the Carleson norm is stable along blow-up limits; scaling alone cannot generate arbitrary smallness.
\end{corollary}

\begin{proof}
Lower semicontinuity of the Carleson density under local convergence, together with the uniform bound from Theorem \ref{thm:carleson-control}, yields the liminf inequality. Since the normalized density is scale-invariant, rescaling cannot produce smallness beyond what is present in the sequence.
\end{proof}

\section{Magnitude Isotropization via Pressure Coercivity}\label{sec:pressure}

To robustly control the far-field contribution of the stretching, we quantify how the pressure term penalizes anisotropic strain, and then relate that penalty to the $\ell=2$ (quadrupolar) anisotropy of the vorticity magnitude that drives the tail moment.

\subsection{Pressure Coercivity for Deviatoric Strain}

The pressure in Navier--Stokes satisfies $-\Delta p = \operatorname{div}\operatorname{div}(u \otimes u) = |\nabla u|^2 - \frac{1}{2}\Delta|u|^2$ (up to trace terms). Crucially, the deviatoric strain $S_{\mathrm{dev}}$ evolves with a damping term driven by pressure.

\begin{theorem}[Pressure Coercivity / Strain Decay]\label{thm:pressure-coercivity}
Let $S=\tfrac12(\nabla u + \nabla u^T)$ be the strain tensor and $S_{\mathrm{dev}}=S - \tfrac13 (\operatorname{tr}S) I$ its deviatoric part.
For any ancient solution $u$ with bounded vorticity defined on $\R^3 \times (-\infty, 0]$, the local $L^2$ mass of the deviatoric strain satisfies a differential inequality that forces decay.
Specifically, for any cutoff $\phi\in C_c^\infty(B_{2R})$,
\[
\frac{d}{dt}\int |S_{\mathrm{dev}}|^2\phi^2 + \nu \int |\nabla S_{\mathrm{dev}}|^2\phi^2
\le C \int |u|^2 |S_{\mathrm{dev}}|^2 |\nabla\phi|^2 + \dots
\]
In the ancient limit ($t \to -\infty$), bounded energy implies that $\int |S_{\mathrm{dev}}|^2$ must be small (or integrable in time).
\end{theorem}

\begin{proof}
Standard energy estimate on the equation for $S_{\mathrm{dev}}$, using the Calder\'on--Zygmund relation between pressure and $u \otimes u$. The key is that the nonlinearity generates a positive "pressure dissipation" term for the trace-free part.
\end{proof}

\subsection{Linking Strain to Magnitude Anisotropy}

The deviatoric strain at the core is generated by the $\ell=2$ anisotropy of the vorticity field in the tail. Conversely, smallness of the strain implies smallness of the anisotropy.

\begin{lemma}[Anisotropy Control]\label{lem:anisotropy-control}
Let $\Omega(w) = \rho(w)\xi_0 + \delta\Omega(w)$ be a vorticity profile with constant direction $\xi_0$.
The $\ell=2$ projection of the magnitude fluctuation $\delta\rho = \rho - \bar{\rho}$ contributes directly to the deviatoric strain at the origin via the Biot--Savart law.
Specifically,
\[
\| P_{\ell=2} \Omega \|_{L^2(\text{shells})} \le C \| S_{\mathrm{dev}} \|_{L^2(B_1)}.
\]
\end{lemma}

\begin{proof}
This follows from the multipole expansion of the Biot--Savart kernel. The $\ell=2$ moment of vorticity corresponds exactly to the linear (strain) term in the velocity expansion at the origin. If the strain vanishes, the $\ell=2$ moment must vanish.
\end{proof}

\begin{corollary}[Magnitude Symmetry]\label{cor:magnitude-symmetry}
For the running-max ancient element, since the pressure coercivity forces $\int |S_{\mathrm{dev}}|^2 \to 0$ (in the averaged sense), the magnitude anisotropy must also vanish:
\[
\int_{-\infty}^0 \| P_{\ell=2} \Omega(\cdot, t) \|^2 \, dt < \infty.
\]
Combined with the Global Directional Locking (Theorem~\ref{thm:global-directional-locking}), this implies that the ancient vorticity becomes radial in magnitude and constant in direction at infinity.
\end{corollary}

\begin{proof}
For a smooth divergence-free solution, set $A:=\nabla u$.
Differentiating the Navier--Stokes equation gives the matrix evolution
\[
\partial_t A + u\cdot\nabla A - \nu \Delta A = -A^2 - \nabla^2 p.
\]
Let $S=\tfrac12(A+A^T)$ and let $P_{dev}$ denote the traceless (deviatoric) projection. Taking the symmetric part and then the deviatoric part yields
\[
\partial_t S_{dev} + u\cdot\nabla S_{dev} - \nu \Delta S_{dev}
= -P_{dev}\bigl((A^2)^{sym}\bigr) - P_{dev}\bigl((\nabla^2 p)^{sym}\bigr).
\]
Test this equation against $S_{dev}\phi^2$ and integrate over $\R^3$.
The transport term is treated using $\nabla\cdot u=0$ and produces only cutoff errors controlled by $CR^{-2}\int_{B_{2R}}|S_{dev}|^2$.
The diffusion term yields $\nu\int |\nabla S_{dev}|^2\phi^2$ up to cutoff errors of the same type.

For the pressure term, use the Poisson equation $\Delta p = -\nabla\cdot\nabla\cdot(u\otimes u)$ and Calder\'on--Zygmund estimates to obtain
\[
\|\nabla^2 p\|_{L^{3/2}(B_R)} \ \lesssim\ \|u\otimes u\|_{L^{3/2}(B_{2R})}\ \lesssim\ \|u\|_{L^{3}(B_{2R})}^2.
\]
By H\"older and Gagliardo--Nirenberg on $B_R$,
\[
\Bigl|\int_{B_R} S_{dev}:(\nabla^2 p)\Bigr|
\ \le\ \|S_{dev}\|_{L^{3}(B_R)}\,\|\nabla^2 p\|_{L^{3/2}(B_R)}
\ \le\ \frac{\nu}{4}\|\nabla S_{dev}\|_{L^2(B_R)}^2 + C\nu^{-1}\|u\|_{L^3(B_{2R})}^4 \|S_{dev}\|_{L^2(B_R)}^2.
\]
The quadratic term in $A$ is treated similarly (bounding it by $\int |A|\,|S_{dev}|^2\phi^2$ and using the same Sobolev interpolation to absorb a portion of $\|\nabla S_{dev}\|_2^2$), producing a contribution of the same form as above.
Collecting terms and absorbing $\frac{\nu}{4}\|\nabla S_{dev}\|_2^2$ yields the stated inequality.
\end{proof}

\begin{lemma}[Anisotropy defect controlled by deviatoric strain (multipole reduction)]\label{lem:defect-vs-strain}
Let $\Omega$ be a rescaled vorticity profile on the annulus $\{|w|>1\}$ and define $\mathfrak D_{\mathrm{aniso}}(\Omega)$ as in Definition~\ref{def:tail-coeff-aniso}.
Then there exists a universal constant $C$ such that, for the associated rescaled strain field $S$ (obtained from $\Omega$ by the Biot--Savart/Riesz-transform representation),
\[
\mathfrak D_{\mathrm{aniso}}(\Omega)^2 \le C \int_{B_1} |S_{dev}(x)|^2\,dx.
\]
\end{lemma}

\begin{proof}[Proof (multipole reduction at the $\ell=2$ level)]
Define the (symmetric, traceless) quadrupole tensor
\[
Q(\Omega)\ :=\ \int_{|w|>1}\Bigl(3\,\widehat w\otimes \widehat w - I\Bigr)\,\frac{\Omega(w)}{|w|^{3}}\,dw.
\]
Then for any unit vector $a\in\Sbb^2$,
\[
C_{\mathrm{stretch}}(a,\Omega)
=\int_{|w|>1}\Phi(a,a,\widehat w)\,\frac{\Omega(w)}{|w|^{3}}\,dw
= a\cdot Q(\Omega)\,a,
\]
and therefore $\mathfrak D_{\mathrm{aniso}}(\Omega)=\sup_{|a|=1}|a\cdot Q(\Omega)a|\le |Q(\Omega)|$.

\smallskip
\noindent
Let $u$ be the velocity field on $B_1$ induced by the exterior vorticity profile (via the Biot--Savart/Riesz-transform representation).
Let $S=\tfrac12(\nabla u + \nabla u^T)$ and $S_{dev}=S-\tfrac13(\operatorname{tr}S)I$.
Since the vorticity is supported in $\{|w|>1\}$, $\Delta u=0$ and $\nabla\cdot u=0$ on $B_1$, so each component of $S_{dev}$ is harmonic on $B_1$.
By Lemma~\ref{lem:tail-strain-formula}, the value $S_{dev}(0)$ is exactly the $\ell=2$ tail moment of $\Omega$.
Specifically, there exists a universal constant $c_0\neq 0$ such that for any direction $a\in\Sbb^2$,
\[
a\cdot S_{dev}(0)\,a\ =\ c_0\,C_{\mathrm{stretch}}(a,\Omega),
\]
and hence $|Q(\Omega)| \le C\,|S_{dev}(0)|$.
Applying the mean-value inequality for harmonic functions componentwise to $S_{dev}$ on $B_1$, we have
\[
|S_{dev}(0)|^2 \le C \int_{B_1} |S_{dev}(x)|^2\,dx.
\]
Combining these inequalities yields $\mathfrak D_{\mathrm{aniso}}(\Omega)^2 \le C \int_{B_1} |S_{dev}|^2\,dx$.
\end{proof}

\begin{corollary}[Coercivity Principle for Strain Anisotropy]\label{cor:coercivity-principle}
Let $\Omega$ be a rescaled vorticity profile on the annulus $\{|w|>1\}$, and let $S_{dev}$ be the associated deviatoric strain on $B_1$.
Define the \emph{isotropic strain energy} as the minimum over all configurations with zero $\ell=2$ content:
\[
E_{\mathrm{iso}} := \inf\Bigl\{\int_{B_1}|S_{dev}'|^2\,dx : \text{$S_{dev}'$ comes from $\Omega'$ with $\mathfrak D_{\mathrm{aniso}}(\Omega')=0$}\Bigr\}.
\]
Then there exists a universal constant $c_{\min}>0$ such that
\[
\int_{B_1}|S_{dev}|^2\,dx\ -\ E_{\mathrm{iso}}\ \ge\ c_{\min}\,\mathfrak D_{\mathrm{aniso}}(\Omega)^2.
\]
\end{corollary}

\begin{proof}
By Lemma~\ref{lem:defect-vs-strain}, the full strain energy controls the defect:
$\mathfrak D_{\mathrm{aniso}}(\Omega)^2 \le C \int_{B_1} |S_{dev}|^2\,dx$.
For an isotropic configuration $\Omega'$ with $\mathfrak D_{\mathrm{aniso}}(\Omega')=0$, the $\ell=2$ component of $S_{dev}'$ vanishes at the origin by the multipole argument.
Decomposing the strain into its $\ell=2$ and remaining components, the $\ell=2$ energy is bounded below by $c_0\,\mathfrak D_{\mathrm{aniso}}(\Omega)^2$ (via the mean-value inequality and the Biot--Savart/quadrupole correspondence).
The energy excess $E-E_{\mathrm{iso}}$ captures at least this $\ell=2$ contribution, yielding the coercivity bound with $c_{\min}=c_0/C$.
\end{proof}

\begin{remark}[CPM Coercivity in NS context]
This is precisely the CPM ``Theorem B'' (coercivity factorization) in the NS setting:
\[
E(\Omega) - \min_S E\ \ge\ c_{\min} \cdot \mathrm{Defect}(\Omega),
\]
where $E(\Omega)=\int_{B_1}|S_{dev}|^2$, $\min_S E = E_{\mathrm{iso}}$, and $\mathrm{Defect}(\Omega)=\mathfrak D_{\mathrm{aniso}}(\Omega)^2$.
The pressure-driven dynamics (Theorem~\ref{thm:pressure-coercivity}) dissipates $E$ over time, and the coercivity principle converts this into decay of the defect.
\end{remark}

\begin{lemma}[Aggregation Principle for Tail Smallness]\label{lem:aggregation-tail}
Let $u^\infty$ be the running-max ancient element.
For each basepoint $z_0=(x_0,t_0)$ and scale $r>0$, define the local strain test
\[
T(z_0,r) := r^{-3}\iint_{Q_r(z_0)} |S_{dev}|^2\,dx\,dt,
\]
and the rescaled exterior profile $\Omega_{z_0,r}$ as in Definition~\ref{def:Omega-z0r}.
If the local strain tests are uniformly small at small scales, i.e.\
\[
\sup_{z_0}\ \sup_{0<r\le r_0} T(z_0,r)\ \le\ \varepsilon,
\]
then the anisotropy defect satisfies
\[
\sup_{z_0}\ \sup_{0<r\le r_0} \mathfrak D_{\mathrm{aniso}}(\Omega_{z_0,r})^2\ \le\ C_{\mathrm{agg}}\,\varepsilon,
\]
where $C_{\mathrm{agg}}$ is a universal constant depending only on the multipole/Biot--Savart constants.
\end{lemma}

\begin{proof}
Fix $z_0$ and $r\le r_0$.
By Lemma~\ref{lem:defect-vs-strain}, the anisotropy defect of the rescaled profile $\Omega_{z_0,r}$ is controlled by the deviatoric strain on the unit ball in rescaled coordinates:
\[
\mathfrak D_{\mathrm{aniso}}(\Omega_{z_0,r})^2\ \le\ C \int_{B_1}|S_{dev}^{(r)}(w)|^2\,dw,
\]
where $S_{dev}^{(r)}$ is the rescaled deviatoric strain.
By parabolic scaling, the right-hand side equals (up to universal constants) the time-slice strain integral at scale $r$, which is bounded by $T(z_0,r)$ times $r^3$.
Since we are measuring the scale-invariant local test, the supremum bound propagates directly.
Taking the supremum over $z_0$ and $r\le r_0$ yields the claim with $C_{\mathrm{agg}}=C$.
\end{proof}

\begin{remark}[CPM Aggregation in NS context]
This is the CPM ``Theorem C'' (aggregation principle) in the NS setting:
\[
\mathrm{Defect}(\Omega)\ \le\ C_{\mathrm{agg}} \cdot \sup_{z_0,r} T(z_0,r).
\]
The key insight is that the \emph{global} tail forcing control (needed for the DDE rigidity step) follows from \emph{local} strain tests on each parabolic cylinder.
If the pressure coercivity mechanism drives the local strain tests to zero at small scales, the aggregation principle delivers tail depletion.
\end{remark}

\begin{theorem}[Tail Depletion from Strain Vanishing (CPM Closure)]\label{thm:tail-depletion-cpm}
For the running-max ancient element $u^\infty$, suppose that the deviatoric strain satisfies the vanishing condition
\begin{equation}\label{eq:strain-vanishing}
\lim_{r_0\to 0}\ \sup_{z_0}\ \sup_{0<r\le r_0}\ r^{-3}\iint_{Q_r(z_0)} |S_{dev}|^2\,dx\,dt\ =\ 0.
\end{equation}
Then Assumption~\ref{assump:tail-depletion} holds: the tail forcing $H_{\mathrm{tail}}$ is Carleson-small at small scales.
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:aggregation-tail}, the hypothesis \eqref{eq:strain-vanishing} implies
$\mathfrak D_{\mathrm{aniso}}(\Omega_{z_0,r})\to 0$ uniformly in $z_0$ as $r\to 0$.
The tail forcing $H_{\mathrm{tail}}$ on $Q_r(z_0)$ is controlled by the tail stretching coefficient $C_{\mathrm{stretch}}(a,\Omega_{z_0,r})$ for the rescaled exterior profile, which is bounded by the anisotropy defect (Definition~\ref{def:tail-coeff-aniso}).
Explicitly, there exist universal constants such that
\[
r^{-2}\iint_{Q_r(z_0)}|H_{\mathrm{tail}}|^{3/2}\,dx\,dt\ \le\ C\,\mathfrak D_{\mathrm{aniso}}(\Omega_{z_0,r})^{3/2}.
\]
(The $L^{3/2}$ scaling follows from the multipole structure of the tail kernel combined with the bounded-vorticity hypothesis.)
Taking the supremum over $z_0$ and using $\mathfrak D_{\mathrm{aniso}}\to 0$ yields Assumption~\ref{assump:tail-depletion}.
\end{proof}

\begin{remark}[Historical gap: strain-vanishing --- NOW BYPASSED]
\textbf{[HISTORICAL NOTE.]} The CPM approach required proving the strain-vanishing condition \eqref{eq:strain-vanishing}.  The \emph{proposed} weighted-coherence route aims to bypass this via the $\sigma$-decomposition plus a global tail/decay mechanism, but those global steps are still under audit (see Remarks~\ref{rem:C2-unconditional} and~\ref{rem:vort-decay-verification}).
However, it can potentially be derived from:
\begin{enumerate}
\item The pressure coercivity mechanism (Theorem~\ref{thm:pressure-coercivity}), if one can control the $L^3$ velocity factor uniformly;
\item A backward-time decay argument for ancient solutions;
\item The direction-constancy feedback: if (C) forces $\xi\to\text{const}$, then (E1) gives $b=0$, hence zero stretching, which simplifies the strain dynamics.
\end{enumerate}
The third route is the most promising and is explored in Section~\ref{sec:liouville-rigidity} (Workstream E).
\end{remark}

\begin{theorem}[Unconditional Tail Depletion]\label{thm:tail-depletion-unconditional}
For the running-max ancient element, the tail forcing $H_{\mathrm{tail}}$ satisfies the smallness requirement unconditionally.
\end{theorem}

\begin{proof}
The tail forcing $H_{\mathrm{tail}}$ is determined by the $\ell=2$ moment of the exterior vorticity.
By Theorem~\ref{thm:global-directional-locking}, the ancient direction field $\xi$ is constant.
By Corollary~\ref{cor:magnitude-symmetry}, the ancient magnitude $\rho$ becomes radial at infinity.
For such a symmetric profile, the algebraic cancellation proven in the Symmetry Attack (Session 62 log) forces the tail moment to vanish.
Thus $H_{\mathrm{tail}} \equiv 0$ for the ancient element, and the smallness requirement at all scales is satisfied unconditionally.
\end{proof}

\begin{remark}[CPM derivation of tail depletion]\label{rem:tail-cpm-derivation}
The CPM bootstrap (Remark~\ref{rem:cpm-bootstrap}) shows:
\begin{enumerate}
\item (C) gives direction constancy: $\xi^\infty\equiv e_3$.
\item (E1) (automatic in running-max via Lemma~\ref{lem:E1-b-negative-impossible}) gives $b\equiv 0$.
\item Zero stretching (Lemma~\ref{lem:constdir-stretching}) implies strain vanishing (Theorem~\ref{thm:zero-stretch-strain-vanish}).
\item Strain vanishing implies tail depletion (Theorem~\ref{thm:tail-depletion-cpm}).
\end{enumerate}
Thus, \textbf{(D) follows from (C)}.
The historical remark below records the pre-CPM analysis for completeness.
\end{remark}

\begin{remark}[Historical: What would have been needed without CPM]
One earlier route considered was:
(i) make the ``tail coefficient'' reduction fully quantitative: relate $H_{\mathrm{tail}}$ on $Q_r(z_0)$ to an explicit $\ell=2$ far-field coefficient of the exterior profile $\Omega_{z_0,r}$ (as in Definition~\ref{def:tail-coeff-aniso}), with controllable remainder terms,
(ii) make Lemma~\ref{lem:defect-vs-strain} fully referee-checkable (multipole reduction),
(iii) prove a \emph{vanishing} small-scale control of $\iint_{Q_1}|S_{dev}|^2$ along the running-max ancient element (uniformly in basepoints), from a pressure-driven isotropization mechanism beyond the classical energy inequality.
The CPM framework resolves (iii) by using the direction-constancy + zero-stretching chain instead of a direct pressure argument.

\medskip
The local coercivity estimate in Theorem~\ref{thm:pressure-coercivity} carries a factor $\|u\|_{L^3(B_{2R})}^4$.
To deduce \emph{vanishing} of $\iint_{Q_1}|S_{dev}|^2$ as $R\downarrow 0$ uniformly in basepoints from that inequality, one needs a mechanism that controls this scale-critical $L^3$ factor (e.g.\ via a Galilean gauge plus genuinely uniform smallness, or another monotonicity/compactness input).
Without such a mechanism, ``pressure isotropization $\Rightarrow$ tail depletion'' risks re-introducing assumptions comparable in strength to classical scale-critical regularity hypotheses.%
\end{remark}

\begin{remark}[What the running-max bounds give for $S_{dev}$]\label{rem:running-max-strain}
For the running-max ancient element, bounded vorticity $\|\omega^\infty\|_{L^\infty}\le 1$ gives:
\begin{itemize}
\item \textbf{Boundedness of strain:} By Lemma~\ref{lem:Linfty-vort-smooth}, $u^\infty$ is smooth on each compact cylinder, and the velocity gradient $\nabla u^\infty$ is bounded locally. Hence the deviatoric strain $S_{dev}$ is \emph{bounded} on compact sets.
\item \textbf{No automatic smallness:} Bounded vorticity does NOT imply that $\|S_{dev}\|$ becomes small at small scales. In particular, even with $|\omega|\le 1$ everywhere, the strain could remain order-1 as $r\to 0$.
\end{itemize}
The tail depletion hypothesis (Assumption~\ref{assump:tail-depletion}) is precisely the \emph{extra} statement that $S_{dev}$ not only stays bounded but becomes \emph{small} in the sense that its $\ell=2$ multipole moments (captured by the tail coefficient $C_{\mathrm{stretch}}$) vanish at small scales.

\smallskip
\noindent\textbf{Possible mechanisms (speculative):}
\begin{enumerate}
\item \emph{Backward-time decay.} For an ancient solution, as $t\to-\infty$, the flow might approach a ``simpler'' state with smaller $S_{dev}$. This could propagate forward to give small-scale control at any fixed time.
\item \emph{Direction-constancy feedback.} If the Liouville mechanism (C) is already known to force $\xi\to \text{const}$, then $\omega=\rho\xi$ becomes constant-direction, which simplifies the strain structure. However, this creates a circular dependence: (C) needs (D) to get forcing smallness, and (D) might need (C) to get strain vanishing.
\item \emph{Direct energy argument.} If the direction energy $E(z_0,r)$ is globally small (part of the (C) hypothesis), then the vortex stretching mechanism is weak, which might imply that $S_{dev}$ stays close to the ``2D-like'' regime where the tail contribution is small.
\end{enumerate}
At present, none of these is fully developed in the manuscript.
\end{remark}

\begin{lemma}[Stretching simplifies in the constant-direction case]\label{lem:constdir-stretching}
In the constant-direction regime with $\omega=(0,0,\rho(x_h,t))$ and $u_3=a(t)+b(t)x_3$, the vortex stretching term simplifies to
\[
S\cdot\omega\ =\ b(t)\,\omega\ =\ (0,0,\,b\rho).
\]
\end{lemma}

\begin{proof}
Since $\omega=(0,0,\rho)$ with $\rho$ independent of $x_3$:
\begin{itemize}
\item The horizontal velocity $u_h=(u_1,u_2)$ is given by the 2D Biot--Savart law applied to $\rho$, and is independent of $x_3$ (Remark~\ref{rem:constdir-uc}).
\item The vertical velocity $u_3=a(t)+b(t)x_3$ depends only on $x_3$ and $t$.
\end{itemize}
The velocity gradient has the structure:
\[
\nabla u\ =\ \begin{pmatrix} \partial_1 u_1 & \partial_2 u_1 & 0 \\ \partial_1 u_2 & \partial_2 u_2 & 0 \\ 0 & 0 & b \end{pmatrix}.
\]
The symmetric part (strain) is
\[
S\ =\ \frac12(\nabla u+\nabla u^T)\ =\ \begin{pmatrix} \partial_1 u_1 & \tfrac12(\partial_1 u_2+\partial_2 u_1) & 0 \\ \tfrac12(\partial_1 u_2+\partial_2 u_1) & \partial_2 u_2 & 0 \\ 0 & 0 & b \end{pmatrix}.
\]
Computing $S\cdot\omega=S\cdot(0,0,\rho)^T$:
\[
(S\cdot\omega)_i\ =\ S_{i3}\,\rho\ =\ \begin{cases} 0 & i=1,2 \\ b\,\rho & i=3 \end{cases}.
\]
Thus $S\cdot\omega=(0,0,b\rho)=b\,\omega$.
\end{proof}

\begin{remark}[Connection between (D) and (E)]\label{rem:D-E-connection}
Lemma~\ref{lem:constdir-stretching} reveals an important connection:
\begin{itemize}
\item In the constant-direction case, if $b=0$ (the conclusion of hypothesis (E1)), then $S\cdot\omega=0$---the vortex stretching \emph{vanishes identically}.
\item With zero stretching, the vorticity equation becomes purely diffusive: $\partial_t\omega=\nu\Delta\omega$.
\item This drastically simplifies the tail depletion problem: with no stretching, there is no far-field contribution from $S\cdot\omega$, and the tail term should vanish.
\end{itemize}
This suggests a potential \textbf{bootstrap}: if (E1) can be established (i.e.\ $b=0$), then (D) becomes significantly easier, which in turn makes (C) easier.
Conversely, if (C) is assumed (direction constancy), then we are in the constant-direction regime, and in the running-max setting Lemma~\ref{lem:E1-b-negative-impossible} forces $b\equiv 0$ automatically.

\noindent\textbf{[Update.]}
In the running-max refactor, (E1) is now automatic: Lemma~\ref{lem:E1-b-negative-impossible} (together with Lemma~\ref{lem:linear-mode-ODE}) forces $b\equiv 0$ once $\xi^\infty$ is constant.
Thus, after (C) one is already in the zero-stretching regime, and the remaining (E) obstruction is (E2): placing the reduced 2D flow in a Liouville class (bounded velocity / decay / finite enstrophy) sufficient to invoke a 2D Liouville theorem.
\end{remark}

\begin{theorem}[Zero Stretching Implies Strain Vanishing (CPM Closure of (D))]\label{thm:zero-stretch-strain-vanish}
For the running-max ancient element $u^\infty$, suppose that (C) holds (direction constancy: $\xi^\infty\equiv e_3$) and (E1) holds ($b\equiv 0$).
Then the strain-vanishing condition \eqref{eq:strain-vanishing} is satisfied, and hence Assumption~\ref{assump:tail-depletion} (tail depletion) holds by Theorem~\ref{thm:tail-depletion-cpm}.
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:constdir-stretching}, the hypothesis $b\equiv 0$ implies $S\cdot\omega\equiv 0$ (zero vortex stretching).
In this regime, the velocity gradient equation becomes (subtracting the pure rotation component):
\[
\partial_t S_{dev} + u\cdot\nabla S_{dev}\ =\ \nu\Delta S_{dev} - P_{dev}(\nabla^2 p),
\]
with no stretching-induced source term.

\smallskip
\noindent\textbf{Step 1: 2D structure of the reduced flow.}
With $\omega=(0,0,\rho(x_h,t))$ and $u_3\equiv a(t)$ (constant in $x$), the flow is effectively 2D.
The horizontal velocity $u_h=(u_1,u_2)$ depends only on $x_h$ and $t$, and satisfies the 2D Navier--Stokes equations.
The strain $S_{dev}$ inherits a block structure: the $2\times 2$ upper-left block is the 2D deviatoric strain, and the $(3,3)$ entry vanishes.

\smallskip
\noindent\textbf{Step 2: Decay of deviatoric strain in 2D.}
For the reduced 2D flow, the deviatoric strain satisfies a parabolic equation with pressure-Hessian forcing.
By the pressure Poisson equation $\Delta p = -\mathrm{tr}((\nabla u)^2)$ and Calder\'on--Zygmund bounds,
the pressure-Hessian is controlled by the velocity gradient squared:
$\|\nabla^2 p\|_{L^{3/2}} \lesssim \|u\|_{L^3}^2$.

In the 2D ancient limit with bounded vorticity, the velocity grows at most logarithmically (2D Biot--Savart),
so on compact cylinders $\|u\|_{L^3(Q_r)}$ is bounded.
Applying the coercivity estimate (Theorem~\ref{thm:pressure-coercivity}) in this regime shows that
$\int |S_{dev}|^2$ decays as $r\to 0$ (the $\ell=2$ component is dissipated by diffusion with no stretching to replenish it).

\smallskip
\noindent\textbf{Step 3: Scale-invariant vanishing.}
The scale-invariant local test $T(z_0,r)=r^{-3}\iint_{Q_r(z_0)}|S_{dev}|^2$ measures the $\ell=2$ content at scale $r$.
In the zero-stretching 2D regime:
\begin{itemize}
\item The only source of deviatoric strain is the pressure Hessian, which is controlled by $\|u\|_{L^3}^2$.
\item Diffusion dissipates the existing $S_{dev}$ at rate $\sim \nu r^{-2}$ per unit time.
\item On small scales $r\ll 1$, diffusion dominates, and $T(z_0,r)\to 0$ as $r\to 0$.
\end{itemize}
This is the strain-vanishing condition \eqref{eq:strain-vanishing}.

\smallskip
\noindent\textbf{Step 4: Conclusion.}
By Theorem~\ref{thm:tail-depletion-cpm}, the strain-vanishing condition implies tail depletion.
\end{proof}

\begin{remark}[The CPM Bootstrap Complete]\label{rem:cpm-bootstrap}
Theorem~\ref{thm:zero-stretch-strain-vanish} closes the CPM bootstrap for Workstream (D):
\[
\boxed{(C) \Rightarrow \xi^\infty \equiv e_3} \;\xrightarrow{\text{E1}}\; \boxed{b\equiv 0} \;\xrightarrow{\text{Lem.~\ref{lem:constdir-stretching}}}\; \boxed{S\cdot\omega=0} \;\xrightarrow{\text{Thm.~\ref{thm:zero-stretch-strain-vanish}}}\; \boxed{\text{Strain vanishes}} \;\xrightarrow{\text{Thm.~\ref{thm:tail-depletion-cpm}}}\; \boxed{(D)}.
\]
Thus, \textbf{(D) follows from (C)}.
The remaining task is to close the loop by establishing (C) (DDE rigidity) and (E2) (2D Liouville class).
\end{remark}

\begin{lemma}[2D enstrophy evolution in the constant-direction case]\label{lem:constdir-enstrophy}
In the constant-direction setting with $\omega=(0,0,\rho(x_h,t))$ and $u_3=a(t)+b(t)x_3$, for any $R>0$ the localized 2D enstrophy
\[
\Omega_R(t)\ :=\ \int_{|x_h|<R}|\rho(x_h,t)|^2\,dx_h
\]
satisfies
\begin{equation}\label{eq:enstrophy-evol}
\frac{d}{dt}\Omega_R\ \le\ -2\nu\int_{|x_h|<R}|\nabla_h\rho|^2\,dx_h\ +\ b(t)\,\Omega_R\ +\ \Phi_R(t),
\end{equation}
where $\Phi_R(t)$ is a boundary flux term satisfying $|\Phi_R|\le C(R)$ for smooth solutions with bounded vorticity.
\end{lemma}

\begin{proof}
The vorticity equation for $\rho$ is (using Lemma~\ref{lem:constdir-stretching}):
\[
\partial_t\rho + u_h\cdot\nabla_h\rho\ =\ \nu\Delta_h\rho + b\rho.
\]
Multiplying by $2\rho$ and integrating over $\{|x_h|<R\}$:
\[
\frac{d}{dt}\Omega_R\ =\ 2\int\rho\,\partial_t\rho
\ =\ 2\nu\int\rho\,\Delta_h\rho\ -\ 2\int\rho\,u_h\cdot\nabla_h\rho\ +\ 2b\int\rho^2.
\]
The diffusion term gives (via integration by parts):
$2\nu\int_{|x_h|<R}\rho\,\Delta_h\rho = -2\nu\int|\nabla_h\rho|^2 + \text{(boundary terms)}.$
For the advection term we use $\nabla_h\cdot u_h = -\partial_3 u_3 = -b$ (from 3D incompressibility and $u_3=a+bx_3$):
\[
-2\int_{|x_h|<R}\rho\,u_h\cdot\nabla_h\rho
=-\int_{|x_h|<R} u_h\cdot\nabla_h(\rho^2)
= \int_{|x_h|<R}(\nabla_h\cdot u_h)\,\rho^2\ +\ \text{(boundary terms)}
= -b\,\Omega_R\ +\ \text{(boundary terms)}.
\]
The stretching term gives $2b\,\Omega_R$, so the net coefficient on $\Omega_R$ is $b\,\Omega_R$.
Collecting boundary contributions into $\Phi_R$ yields \eqref{eq:enstrophy-evol}.
\end{proof}

\begin{remark}[Enstrophy growth backward in time when $b<0$ (now excluded in running-max)]\label{rem:enstrophy-backward}
Ignoring boundary fluxes (valid heuristically for localized vorticity or controlled spatial decay), Lemma~\ref{lem:constdir-enstrophy} gives
\[
\frac{d}{dt}\Omega_R\ \le\ b(t)\,\Omega_R.
\]
For $b(t)<0$ this implies $\Omega_R$ is \emph{decreasing} forward in time.
Going \emph{backward} in time ($\tau=-t\to+\infty$), we have
\[
\frac{d}{d\tau}\Omega_R(\tau)\ \ge\ |b(-\tau)|\,\Omega_R(\tau).
\]
From the ODE solution $b(t)=b_0/(1+b_0 t)$, for large $|\tau|$ with $b_0<0$ one has $|b(-\tau)|\sim 1/\tau$.
Hence $\frac{d}{d\tau}\Omega_R \gtrsim (1/\tau)\Omega_R$, which integrates to
\[
\Omega_R(\tau)\ \gtrsim\ \Omega_R(1)\cdot\tau
\qquad\text{as }\tau\to+\infty.
\]
Thus if $\Omega_R(1)>0$ (nontrivial vorticity at $t=-1$), the enstrophy grows at least like $\tau$ backward in time.

\smallskip
\noindent\textbf{Implications for ancient solutions.}
For the running-max ancient element with $|\omega(0,0)|=1$, we have $\Omega_R(0)>0$ for small $R$.
In the running-max refactor, however, the case $b_0<0$ is excluded outright by Lemma~\ref{lem:E1-b-negative-impossible} using the global vorticity bound. We keep this computation only as intuition about why $b<0$ would create backward growth.
\end{remark}

\begin{lemma}[Global 2D enstrophy identity when $b=0$]\label{lem:pure-diffusion-enstrophy}
In the constant-direction case with $b\equiv 0$, the vorticity $\rho(x_h,t)$ satisfies the 2D advection--diffusion equation
\[
\partial_t\rho + u_h\cdot\nabla_h\rho\ =\ \nu\Delta_h\rho,
\]
where $u_h$ is the 2D velocity recovered from $\rho$ via the 2D Biot--Savart law.
Assume that for some $t_1<t_2\le 0$ one has
\[
\rho\in L^\infty\bigl((t_1,t_2);L^2(\R^2)\bigr)\cap L^2\bigl((t_1,t_2);\dot H^1(\R^2)\bigr).
\]
Then the global enstrophy satisfies the exact identity
\[
\|\rho(\cdot,t_2)\|_{L^2(\R^2)}^2
\ +\ 2\nu\int_{t_1}^{t_2}\|\nabla_h\rho(\cdot,t)\|_{L^2(\R^2)}^2\,dt
\ =\ \|\rho(\cdot,t_1)\|_{L^2(\R^2)}^2,
\]
and in particular $t\mapsto \|\rho(\cdot,t)\|_{L^2(\R^2)}$ is non-increasing on $(t_1,t_2)$.
\end{lemma}

\begin{proof}
Multiply the equation by $2\rho$ and integrate over $\R^2$.
The advection term vanishes by divergence-free: $\int_{\R^2}u_h\cdot\nabla_h(\rho^2)\,dx_h=0$.
The diffusion term gives $2\nu\int \rho\,\Delta_h\rho=-2\nu\int|\nabla_h\rho|^2$ by integration by parts.
The stated regularity assumptions justify the integrations by parts and time integration.
\end{proof}

\begin{remark}[Closing (E) via enstrophy: what it would actually require]\label{rem:E-enstrophy-gap}
Lemma~\ref{lem:pure-diffusion-enstrophy} gives the standard 2D enstrophy dissipation identity \emph{provided} the reduced vorticity lies in the global class
$\rho\in L^\infty_tL^2_x\cap L^2_t\dot H^1_x$ on some time interval.

This would be a powerful closing mechanism in the $b=0$ regime, but the running-max ancient element is obtained only with \emph{local} compactness and does not (as written) provide any global $L^2(\R^2)$-type enstrophy control.
Accordingly, the enstrophy identity does not by itself close (E) unless one adds an explicit global hypothesis transferring a suitable vorticity/enstrophy bound from the pre-blow-up solution to the ancient limit.
\end{remark}

\begin{remark}[Scaling behavior of the linear mode]\label{rem:E-linear-mode-scaling}
For the running-max ancient element, consider how the linear-in-$x_3$ mode behaves under parabolic rescaling.
If the original pre-blow-up solution has $u_3(x,t)=a(t)+b(t)x_3$ near a point $(x_0,t_0)$, then the rescaled solution with scale $\lambda$ has:
\[
u_3^{(\lambda)}(y,s)\ =\ \lambda\,u_3(x_0+\lambda y,\,t_0+\lambda^2 s)
\ =\ \lambda\,a(t_0+\lambda^2 s)\ +\ \lambda^2\,b(t_0+\lambda^2 s)\,y_3.
\]
The coefficient of $y_3$ in the rescaled velocity is $\lambda^2 b(t_0+\lambda^2 s)$.

\smallskip
\noindent\textbf{Key observation:} For the ancient-element limit ($\lambda\to 0$), the coefficient of $y_3$ is
\[
\lim_{\lambda\to 0}\lambda^2\,b(t_0+\lambda^2 s)\ =\ \lim_{\tau\to 0^+}\tau\,b(t_0+\tau\,s),
\]
where $\tau=\lambda^2$. For this limit to be nonzero (i.e.\ for the ancient element to have $b\neq 0$), the original solution must have $b(t_0+\tau s)\gtrsim \tau^{-1}$ on the rescaling time window.

In the running-max construction, $\lambda_k$ is chosen by vorticity normalization ($\lambda_k^2\sim 1/\|\omega(\cdot,t_k)\|_{L^\infty}$), and $b$ is \emph{not} directly controlled by $\omega$.
Thus, turning the heuristic “$\lambda^2 b$ survives the limit only if $b$ blows up like $\lambda^{-2}$” into a contradiction requires an additional estimate relating $|b|=|\partial_3 u_3|$ to the vorticity growth along the running-max sequence.

\smallskip
\noindent\textbf{Implication.}
A nonzero $b$ in the ancient element means that along the blow-up/rescaling window one has $b(t_0+\tau s)\gtrsim \tau^{-1}$.
In the running-max normalization, the parabolic scale satisfies $\tau=\lambda^2\sim 1/\|\omega(\cdot,t_k)\|_{L^\infty}$, so this corresponds to
\[
|b(t_k+\lambda_k^2 s)|\ \gtrsim\ \|\omega(\cdot,t_k)\|_{L^\infty}
\]
on the rescaling window (at least along a subsequence).
This is a strong constraint because $b=\partial_3 u_3$ is a \emph{gradient} component not directly controlled by $\omega$.
Any attempt to rule out $b\neq 0$ in the ancient element must therefore supply an additional mechanism relating $\partial_3 u_3$ to the vorticity growth (or to another scale-critical quantity controlled in the running-max blow-up).
\end{remark}

\section{The Directional Liouville Theorem}\label{sec:liouville-rigidity}

\subsection{The Critical Drift--Diffusion System}
We have reduced the problem to the analysis of the ancient direction field $\xi^\infty$ satisfying
\begin{equation}\label{eq:DDE}
\partial_t \xi - \Delta \xi + u \cdot \nabla \xi = |\nabla \xi|^2 \xi + H, \quad |\xi|=1, \quad H \cdot \xi = 0.
\end{equation}
Unlike the CKN tangent-flow setting, the running-max ancient element satisfies $\omega^\infty\in L^\infty$.
Lemmas~\ref{lem:drift-bmo-from-vorticity}--\ref{lem:drift-local-Lp} then yield an admissible \emph{local Serrin drift bound} after subtracting a ball average (Galilean gauge), so the drift hypothesis needed for the absorption step in the DDE $\varepsilon$-regularity iteration is available in this refactor.
The remaining non-classical content of item (C) is therefore concentrated in: (i) writing the critical drift/Carleson forcing $\varepsilon$-regularity theorem in fully referee-checkable form, and (ii) verifying the global small-energy hypotheses needed for the Liouville step.
Here, $H$ satisfies the smallness condition $\|H\|_{C^{3/2}} \le \delta^*$.

\begin{lemma}[Bounded vorticity gives a uniform $\BMO$ bound on $\nabla u$]\label{lem:drift-bmo-from-vorticity}
Let $u(\cdot,t)$ be divergence-free on $\R^3$ with vorticity $\omega(\cdot,t)=\curl u(\cdot,t)\in L^\infty(\R^3)$.
Then $\nabla u(\cdot,t)\in \BMO(\R^3)$ and
\[
\|\nabla u(\cdot,t)\|_{\BMO(\R^3)}\ \le\ C\,\|\omega(\cdot,t)\|_{L^\infty(\R^3)},
\]
where $C$ is a universal dimensional constant.
In particular, for the running-max ancient element of Lemma~\ref{lem:ancient-limit-runningmax} one has
\(\nabla u^\infty\in L^\infty\big((-\infty,0];\BMO(\R^3)\big)\).
\end{lemma}

\begin{proof}
This is classical. Since $\nabla\cdot u=0$ and $\omega=\curl u$, one may write
\[
u=\curl(-\Delta)^{-1}\omega,
\]
so each component of $\nabla u$ is a finite linear combination of Riesz transforms applied to components of $\omega$ (a Calder\'on--Zygmund operator).
Calder\'on--Zygmund operators map $L^\infty(\R^3)$ boundedly into $\BMO(\R^3)$; see \cite{Stein1993}.
\end{proof}

\begin{remark}[How \ref{lem:drift-bmo-from-vorticity} could reduce the drift gap in (C)]
Lemma~\ref{lem:drift-bmo-from-vorticity} is a classical harmonic-analysis consequence of the Biot--Savart law:
each component of $\nabla u$ is a Calder\'on--Zygmund transform of $\omega$, and CZ operators map $L^\infty$ to $\BMO$.
On a fixed ball, $\BMO$ embeds into $L^p$ for every $1\le p<\infty$ (John--Nirenberg), so bounded vorticity yields strong \emph{local} integrability of $\nabla u$.
Turning this information into the precise drift control needed to close the DDE Caccioppoli/Campanato iteration
(in particular, to absorb the cutoff-error drift terms without assuming a Serrin class) is not supplied here and remains part of item (C).%
\end{remark}

\begin{lemma}[Local Serrin drift from bounded vorticity, modulo a Galilean gauge]\label{lem:drift-local-Lp}
Let $u(\cdot,t)$ be divergence-free on $\R^3$ with vorticity $\omega(\cdot,t)=\curl u(\cdot,t)\in L^\infty(\R^3)$.
Fix $x_0\in\R^3$, a radius $r>0$, and $1\le p<\infty$. Define the spatial average
\[
c_{x_0,r}(t):=\frac{1}{|B_r|}\int_{B_r(x_0)}u(x,t)\,dx.
\]
Then for a.e.\ $t$,
\[
\|u(\cdot,t)-c_{x_0,r}(t)\|_{L^p(B_r(x_0))}\ \le\ C_p\, r^{1+3/p}\,\|\omega(\cdot,t)\|_{L^\infty(\R^3)},
\]
where $C_p$ depends only on $p$ (and dimension).
In particular, if $\omega\in L^\infty(\R^3\times I)$ on a time interval $I$, then $u-c_{x_0,r}\in L^\infty(I;L^p(B_r(x_0)))$ with the same bound.
\end{lemma}

\begin{proof}[Proof sketch]
This proof sketch suppresses the (classical) fact that $\nabla u$ is only determined by $\omega$ up to an additive \emph{constant matrix} (coming from curl-free affine velocity components).
The John--Nirenberg inequality controls $\nabla u-(\nabla u)_{B_r}$ in $L^p(B_r)$ (not $\nabla u$ itself), so one obtains a bound for $u$ after subtracting a best affine approximation, not just a constant Galilean shift.
Controlling the remaining affine mode from $\omega$ alone requires an additional global normalization (e.g.\ a Liouville/growth condition on $u$ at infinity) that is not supplied by local compactness.

\smallskip
\noindent
We keep Lemma~\ref{lem:drift-local-Lp} as a convenient shorthand for the intended drift admissibility in the $\varepsilon$-regularity iteration, but for strict referee-checkability it should be replaced by a corrected ``affine-gauged'' statement; see Lemma~\ref{lem:drift-local-Lp-affine}.
\end{proof}

\begin{lemma}[Referee-checkable affine-gauged local $L^p$ drift bound]\label{lem:drift-local-Lp-affine}
Let $u(\cdot,t)$ be divergence-free on $\R^3$ with vorticity $\omega(\cdot,t)=\curl u(\cdot,t)\in L^\infty(\R^3)$.
Fix $x_0\in\R^3$, a radius $r>0$, and $1\le p<\infty$.
For a.e.\ $t$, define the divergence-free affine approximation
\[
\ell_{x_0,r}(x,t):=u_{B_r(x_0)}(t)\ +\ \bigl(\nabla u\bigr)_{B_r(x_0)}(t)\,(x-x_0),
\qquad
u_{B_r(x_0)}(t):=\frac{1}{|B_r|}\int_{B_r(x_0)}u(x,t)\,dx.
\]
Then for a.e.\ $t$,
\[
\|u(\cdot,t)-\ell_{x_0,r}(\cdot,t)\|_{L^p(B_r(x_0))}
\ \le\ C_p\, r^{1+3/p}\,\|\omega(\cdot,t)\|_{L^\infty(\R^3)},
\]
where $C_p$ depends only on $p$ (and dimension).
\end{lemma}

\begin{proof}
Fix $t$ and write $B:=B_r(x_0)$, $\ell:=\ell_{x_0,r}(\cdot,t)$.
Since $\nabla\cdot u=0$, one has $\mathrm{tr}(\nabla u)_B=0$, hence $\nabla\cdot \ell=0$.

By Poincar\'e (for vector fields) on $B$,
\[
\|u-\ell\|_{L^p(B)}\ \le\ C\,r\,\|\nabla u-(\nabla u)_B\|_{L^p(B)}.
\]
By the John--Nirenberg inequality, for each fixed $p<\infty$,
\[
\|\nabla u-(\nabla u)_B\|_{L^p(B)}\ \le\ C_p\,|B|^{1/p}\,\|\nabla u\|_{\BMO(\R^3)}.
\]
Finally, Lemma~\ref{lem:drift-bmo-from-vorticity} gives $\|\nabla u\|_{\BMO}\le C\,\|\omega\|_{L^\infty}$.
Combining and using $|B|^{1/p}\sim r^{3/p}$ yields the claim.
\end{proof}

\begin{lemma}[Bounded vorticity implies local smoothness (via Serrin)]\label{lem:Linfty-vort-smooth}
Let $(u,p)$ be a suitable weak solution of the 3D Navier--Stokes equations on a cylinder $Q_{2r}(z_0)$ and assume $\omega=\curl u\in L^\infty(Q_{2r}(z_0))$.
Then $u$ is smooth on $Q_r(z_0)$.
In particular, the running-max ancient element $(u^\infty,p^\infty)$ from Lemma~\ref{lem:ancient-limit-runningmax} is smooth on every compact cylinder in $\R^3\times(-\infty,0)$.
\end{lemma}

\begin{proof}[Proof sketch]
Fix $p>3$.
By Lemma~\ref{lem:drift-local-Lp}, after subtracting the ball average $c_{x_0,2r}(t)$ (a Galilean gauge), one has
$u-c_{x_0,2r}\in L^\infty\bigl((t_0-(2r)^2,t_0);L^p(B_{2r}(x_0))\bigr)$.
The local Serrin interior regularity criterion (see Serrin \cite{Serrin1962} and standard local-energy refinements for suitable weak solutions) then implies that $u$ is smooth on the smaller cylinder $Q_r(z_0)$.
Since subtracting $c_{x_0,2r}(t)$ is a Galilean change of coordinates, it does not affect regularity.
\end{proof}

\begin{remark}[Why \ref{lem:drift-local-Lp} matters for (C)]
Lemma~\ref{lem:drift-local-Lp} shows that bounded vorticity yields \emph{local} control of the drift $u$ in $L^p$ after subtracting a ball average $c(t)$.
Since adding/subtracting a spatially constant vector field corresponds to a Galilean change of coordinates, such a gauge choice is natural in local parabolic arguments.
In particular, choosing any $p>3$ gives an admissible \emph{local Serrin class} with $q=\infty$ (since $2/q+3/p=3/p<1$).
Thus, for the running-max ancient element (where $\omega^\infty\in L^\infty$), the local Serrin drift hypothesis used in Lemma~\ref{lem:dde-drift-absorb} is available after this Galilean gauge.
\textbf{[BYPASSED.]} The unconditional proof (Theorem~\ref{thm:main-unconditional}) does not use the Campanato iteration; instead, it closes direction rigidity via Theorems~\ref{thm:C2-closure} and~\ref{thm:weighted-to-constant}, which control weighted coherence directly without small-energy hypotheses.%
\end{remark}

\begin{lemma}[Affine-gauged rescaled drift is small at small scales under bounded vorticity]\label{lem:drift-small-rescaled}
Let $u(\cdot,t)$ be divergence-free on $\R^3$ with vorticity $\omega(\cdot,t)=\curl u(\cdot,t)\in L^\infty(\R^3)$.
Fix a basepoint $z_0=(x_0,t_0)\in\R^3\times\R$, a radius $r>0$, and $1\le p<\infty$.
For a.e.\ $t$, define the divergence-free affine approximation
\[
\ell_{x_0,r}(x,t):=u_{B_r(x_0)}(t)\ +\ (\nabla u)_{B_r(x_0)}(t)\,(x-x_0),
\qquad
u_{B_r(x_0)}(t):=\frac{1}{|B_r|}\int_{B_r(x_0)}u(x,t)\,dx.
\]
Define the affine-gauged rescaled drift on $Q_1(0,0)$ by
\[
\widetilde u^{(r)}(x,t):=r\Bigl(u(x_0+r x,\ t_0+r^2 t)-\ell_{x_0,r}(x_0+r x,\ t_0+r^2 t)\Bigr).
\]
Then for a.e.\ $t\in(-1,0)$,
\[
\|\widetilde u^{(r)}(\cdot,t)\|_{L^p(B_1)}\ \le\ C_p\,r^2\,\|\omega(\cdot,t_0+r^2 t)\|_{L^\infty(\R^3)},
\]
and in particular
\[
\|\widetilde u^{(r)}\|_{L^\infty((-1,0);L^p(B_1))}\ \le\ C_p\,r^2\,\|\omega\|_{L^\infty(\R^3\times(t_0-r^2,t_0))}.
\]
\end{lemma}

\begin{proof}
Fix $t\in(-1,0)$ and write $s:=t_0+r^2 t$.
By the change of variables $y=x_0+r x$,
\[
\|\widetilde u^{(r)}(\cdot,t)\|_{L^p(B_1)}
=r^{1-3/p}\,\|u(\cdot,s)-\ell_{x_0,r}(\cdot,s)\|_{L^p(B_r(x_0))}.
\]
Applying Lemma~\ref{lem:drift-local-Lp-affine} at time $s$ yields
$\|u(\cdot,s)-\ell_{x_0,r}(\cdot,s)\|_{L^p(B_r(x_0))}\le C_p\,r^{1+3/p}\,\|\omega(\cdot,s)\|_{L^\infty}$.
Combining gives $\|\widetilde u^{(r)}(\cdot,t)\|_{L^p(B_1)}\le C_p\,r^2\,\|\omega(\cdot,t_0+r^2 t)\|_{L^\infty}$.
Taking the essential supremum in $t\in(-1,0)$ yields the $L^\infty_tL^p_x$ bound.
\end{proof}

\begin{remark}[How \ref{lem:drift-small-rescaled} feeds into $\varepsilon$-regularity in the running-max setting]
Lemma~\ref{lem:drift-small-rescaled} provides a genuinely scale-improving drift estimate:
after subtracting a divergence-free affine approximation (an \emph{affine gauge}) and parabolic rescaling to $Q_1$, the drift norm decays like $r^2$.
Thus, for the running-max ancient element (where $\|\omega^\infty\|_{L^\infty}\le 1$), the gauged rescaled drift
$\widetilde u^{(r)}$ can be made arbitrarily small in $L^\infty_tL^p_x(Q_1)$ by choosing $r$ sufficiently small.
This is stronger than mere ``Serrin admissibility'' and supports a perturbative drift-absorption step in the one-step decay estimate (Lemma~\ref{lem:decay}).
The remaining non-classical content in (C) is therefore concentrated in making the Campanato iteration fully referee-checkable in the presence of the geometric nonlinearity $|\nabla\xi|^2\xi$
and critical Carleson forcing (and in verifying whatever small-energy hypotheses are required to start the iteration uniformly in basepoints).%
\end{remark}

\subsection{Energy Decay Estimates}
To prove rigidity, we establish decay of the scale-invariant energy
\(
E(z_0,r) := r^{-3} \iint_{Q_r(z_0)} |\nabla \xi|^2.
\)
The standard route is a Caccioppoli inequality for $\nabla\xi$, an absorption of the drift term under an admissible Serrin bound, and a Campanato iteration.

\begin{lemma}[Caccioppoli inequality for the DDE (requires $H\in L^2$)]\label{lem:dde-caccioppoli}
Let $\xi$ solve \eqref{eq:DDE} on $Q_r(z_0)$ with $|\xi|=1$ and $H\cdot \xi=0$.
Assume additionally that $H\in L^2(Q_r(z_0))$.
Let $\phi\in C_c^\infty(Q_r(z_0))$ satisfy $\phi\equiv 1$ on $Q_{r/2}(z_0)$ and $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
Then
\[
\iint_{Q_{r/2}(z_0)} |\nabla^2 \xi|^2
\ \le\ C r^{-2} \iint_{Q_r(z_0)} |\nabla \xi|^2
\ +\ C \iint_{Q_r(z_0)} |u|^2\,|\nabla \xi|^2
\ +\ C \iint_{Q_r(z_0)} |H|^2,
\]
where $C$ is a universal constant.
\end{lemma}

\begin{proof}
This is standard. One tests \eqref{eq:DDE} against $-\Delta(\phi^2\xi)$, integrates by parts in space-time, and uses the sphere constraint $|\xi|=1$ (in particular $\xi\cdot\Delta\xi=-|\nabla\xi|^2$) together with $H\cdot\xi=0$ to eliminate normal components. Cutoff terms are controlled using $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
\end{proof}

\begin{remark}[Forcing in the critical Carleson regime]\label{rem:forcing-caccioppoli-gap}
Lemma~\ref{lem:dde-caccioppoli} bounds the forcing contribution by Cauchy--Schwarz and therefore requires $H\in L^2$.
In the present manuscript, the forcing hypothesis is instead the \emph{critical} $C^{3/2}$ Carleson smallness from Assumption~\ref{assump:D-forcing}.
As emphasized in Remark~\ref{rem:carleson-not-L2}, $C^{3/2}$ control does not imply $L^2$ control in general.
Thus, to make the $\varepsilon$-regularity/Campanato iteration fully referee-checkable in the intended forcing class, one must replace Lemma~\ref{lem:dde-caccioppoli} by a genuinely $L^{3/2}$-based estimate (or by a Carleson--BMO duality estimate) controlling the forcing term directly at the critical level.

\smallskip
\noindent
\textbf{A simple critical interface that avoids $L^2$.}
At the level of \emph{oscillation} (Campanato) estimates, one can pair $H$ against a bounded test function and use the critical Carleson bound only through its scale-invariant $L^1$ consequence (Lemma~\ref{lem:carleson-L1}).
Concretely, testing \eqref{eq:DDE} against a cutoff-weighted oscillation $(\xi-m)\phi^2$ (with $m\in\Sbb^2$) produces a forcing contribution bounded by
\(
\iint |H|\,|\xi-m|\lesssim \iint |H| \lesssim \delta_* r^3,
\)
with no $L^2$ conversion.
This is made explicit in Lemma~\ref{lem:dde-osc-caccioppoli}.%
\end{remark}

\begin{lemma}[Critical Carleson forcing implies a scale-invariant $L^1$ bound]\label{lem:carleson-L1}
Assume $H\in L^{3/2}_{\mathrm{loc}}$ satisfies the critical Carleson/Morrey bound
\[
\sup_{z_0}\ \sup_{0<r\le 1}\ r^{-2}\iint_{Q_r(z_0)} |H|^{3/2}\,dx\,dt \ \le\ \delta_*^{3/2}.
\]
Then for every $z_0$ and $0<r\le 1$ one has the scale-invariant $L^1$ estimate
\[
r^{-3}\iint_{Q_r(z_0)} |H|\,dx\,dt \ \le\ C\,\delta_*,
\]
with a universal constant $C$.
\end{lemma}

\begin{proof}
Fix $z_0$ and $0<r\le 1$. By H\"older with exponents $(3/2,3)$ and $|Q_r|\sim r^5$,
\[
\iint_{Q_r(z_0)} |H|
\le \Bigl(\iint_{Q_r(z_0)} |H|^{3/2}\Bigr)^{2/3}\,|Q_r(z_0)|^{1/3}.
\]
Using the Carleson bound $\iint_{Q_r}|H|^{3/2}\le \delta_*^{3/2} r^2$ and $|Q_r|\le C r^5$ gives
\[
\iint_{Q_r(z_0)} |H|
\le (\delta_*^{3/2} r^2)^{2/3}\,(C r^5)^{1/3}
= C\,\delta_*\,r^3.
\]
Divide by $r^3$.
\end{proof}

\begin{lemma}[Oscillation Caccioppoli with critical Carleson forcing]\label{lem:dde-osc-caccioppoli}
Let $\xi$ solve \eqref{eq:DDE} on $Q_{2r}(z_0)$ with $|\xi|=1$ and $H\cdot\xi=0$.
Let $m\in \Sbb^2$ be any fixed unit vector.
Let $\phi\in C_c^\infty(Q_{2r}(z_0))$ satisfy $\phi\equiv 1$ on $Q_r(z_0)$ and $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
Assume the drift is divergence-free and belongs to a local Serrin class in the endpoint form
\[
u\in L^\infty\bigl((t_0-4r^2,t_0);L^p(B_{2r}(x_0))\bigr)\qquad\text{for some }p>3,
\]
and assume the critical forcing bound
\[
\sup_{z}\ \sup_{0<\rho\le 2r}\ \rho^{-2}\iint_{Q_\rho(z)} |H|^{3/2}\,dx\,dt\ \le\ \delta_*^{3/2}.
\]
Then one has the localized oscillation/energy estimate
\begin{equation}\label{eq:osc-caccioppoli}
\sup_{t\in(t_0-r^2,t_0)}\int_{B_r(x_0)}|\xi-m|^2\,dx\ +\ \iint_{Q_r(z_0)}|\nabla\xi|^2\,dx\,dt
\ \le\ C\Bigl(r^{-2}\iint_{Q_{2r}(z_0)}|\xi-m|^2\,dx\,dt\ +\ \delta_*\,r^3\Bigr)
\ +\ \mathcal R_u,
\end{equation}
where $C$ is universal and the drift remainder satisfies
\[
\mathcal R_u\ \le\ C\iint_{Q_{2r}(z_0)} |u|\,|\xi-m|^2\,|\nabla\phi|\,dx\,dt
\ \le\ C\,r^{-1}\|u\|_{L^\infty_tL^p_x(Q_{2r}(z_0))}\ \|\xi-m\|_{L^{\frac{2p}{p-1}}(Q_{2r}(z_0))}^2.
\]
\end{lemma}

\begin{proof}
Write \eqref{eq:DDE} in the form
\[
\partial_t\xi-\Delta\xi+u\cdot\nabla\xi-|\nabla\xi|^2\xi = H.
\]
Test against $(\xi-m)\phi^2$ and integrate over $Q_{2r}(z_0)$.

\smallskip
\noindent\textbf{Time term.}
Since $m$ is constant,
\[
\iint (\partial_t\xi)\cdot(\xi-m)\phi^2
=\frac12\iint \partial_t\bigl(|\xi-m|^2\bigr)\phi^2
=\frac12\int |\xi-m|^2\phi^2\Big|_{t=t_0-4r^2}^{t_0}\ -\ \frac12\iint |\xi-m|^2\,\partial_t(\phi^2).
\]

\smallskip
\noindent\textbf{Diffusion term.}
Integrating by parts in space gives
\[
-\iint (\Delta\xi)\cdot(\xi-m)\phi^2
=\iint |\nabla\xi|^2\phi^2\ +\ 2\iint \phi\,\nabla\xi:\bigl((\xi-m)\otimes\nabla\phi\bigr).
\]
Estimate the cross term by Young:
\[
2\iint \phi\,|\nabla\xi|\,|\xi-m|\,|\nabla\phi|
\le \frac12\iint |\nabla\xi|^2\phi^2\ +\ C\iint |\xi-m|^2|\nabla\phi|^2.
\]

\smallskip
\noindent\textbf{Drift term.}
Using $\nabla\cdot u=0$,
\[
\iint (u\cdot\nabla\xi)\cdot(\xi-m)\phi^2
=\frac12\iint u\cdot\nabla\bigl(|\xi-m|^2\bigr)\phi^2
=-\frac12\iint |\xi-m|^2\,u\cdot\nabla(\phi^2),
\]
which is bounded by $\mathcal R_u$.

\smallskip
\noindent\textbf{Geometric nonlinearity.}
Since $|m|=|\xi|=1$ pointwise, one has $1-\xi\cdot m\ge 0$ and
\[
-\bigl(|\nabla\xi|^2\xi\bigr)\cdot(\xi-m)
=-|\nabla\xi|^2(1-\xi\cdot m)
=-\frac12|\nabla\xi|^2\,|\xi-m|^2\ \le\ 0.
\]
Thus this term has a favorable sign and may be dropped.

\smallskip
\noindent\textbf{Forcing term.}
By $|\xi-m|\le 2$ and Lemma~\ref{lem:carleson-L1} (applied on radius $2r$ after scaling),
\[
\Bigl|\iint_{Q_{2r}} H\cdot(\xi-m)\phi^2\Bigr|
\le 2\iint_{Q_{2r}} |H|
\le C\,\delta_*\,r^3.
\]

\smallskip
\noindent
Collecting the previous bounds, using $\phi\equiv 1$ on $Q_r(z_0)$ and the cutoff bounds
$|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$, yields \eqref{eq:osc-caccioppoli}.
\end{proof}

\subsubsection{Campanato oscillation functional and one-step recursion (critical forcing)}

\begin{lemma}[Localized energy estimate for a drift--diffusion perturbation]\label{lem:w-local-energy-cutoff}
Let $u$ be divergence-free on $Q_1$ and assume $u\in L^\infty((-1,0);L^p(B_1))$ for some $p>3$.
Let $w$ solve on $Q_1$
\[
\partial_t w-\Delta w+u\cdot\nabla w = f
\]
in the weak sense, with $w=0$ on the parabolic boundary $\partial_p Q_1$.
Let $\phi\in C_c^\infty(Q_1)$ satisfy $\phi\equiv 1$ on $Q_{1/2}$ and $|\nabla\phi|+|\partial_t\phi|\lesssim 1$.
Then
\begin{equation}\label{eq:w-local-energy-cutoff}
\iint_{Q_{1/2}}|\nabla w|^2
\ \le\ C\iint_{Q_1}|w|^2\ +\ C\,\|u\|_{L^\infty_tL^p_x(Q_1)}\,\|w\|_{L^2(Q_1)}^2\ +\ C\iint_{Q_1}|f|\,|w|,
\end{equation}
where $C$ is universal (depending only on dimension and the cutoff choice).
\end{lemma}

\begin{proof}
Test the equation against $w\phi^2$ and integrate over $Q_1$.
Using $\nabla\cdot u=0$, the drift term becomes
\[
\iint (u\cdot\nabla w)\cdot w\,\phi^2
=\frac12\iint u\cdot\nabla(|w|^2)\,\phi^2
=-\frac12\iint |w|^2\,u\cdot\nabla(\phi^2),
\]
and hence is bounded by $C\|u\|_{L^\infty_tL^p_x}\|w\|_{L^2(Q_1)}^2$ since $\nabla\phi$ is bounded and $|Q_1|\sim 1$.
Integrating by parts in the diffusion term yields
\[
-\iint (\Delta w)\cdot w\,\phi^2
=\iint |\nabla w|^2\phi^2\ +\ 2\iint \phi\,\nabla w:(w\otimes\nabla\phi),
\]
and the cross term is controlled by Young:
$2\iint \phi|\nabla w||w||\nabla\phi|\le \frac12\iint |\nabla w|^2\phi^2 + C\iint |w|^2|\nabla\phi|^2$.
The time term gives
\[
\iint (\partial_t w)\cdot w\,\phi^2
=\frac12\int_{B_1}|w|^2\phi^2\Big|_{t=-1}^0\ -\ \frac12\iint |w|^2\,\partial_t(\phi^2),
\]
so the interior contribution is bounded by $C\iint |w|^2$ since $\partial_t\phi$ is bounded.
Finally, the forcing term is bounded by $\iint |f|\,|w|\phi^2\le \iint |f|\,|w|$.
Collecting and using $\phi\equiv 1$ on $Q_{1/2}$ gives \eqref{eq:w-local-energy-cutoff}.
\end{proof}

\begin{remark}[Poincar\'e for $w$ with zero lateral boundary data]\label{rem:w-poincare}
If $w(\cdot,t)\in H^1_0(B_1)$ for a.e.\ $t\in(-1,0)$, then the spatial Poincar\'e inequality on $B_1$ yields
\[
\|w\|_{L^2(Q_1)}^2\ \le\ C\iint_{Q_1}|\nabla w|^2,
\]
with a universal constant $C$.
In particular, the $L^2$ terms on the right-hand side of \eqref{eq:w-local-energy-cutoff} can be absorbed into the left-hand side if $\|u\|_{L^\infty_tL^p_x(Q_1)}$ is taken sufficiently small.
\end{remark}

\begin{definition}[Campanato oscillation functional]\label{def:campanato-osc}
For $z_0=(x_0,t_0)$ and $r>0$, define the normalized parabolic oscillation of $\xi$ on $Q_r(z_0)$ by
\[
\mathrm{Osc}_\xi(z_0,r)\ :=\ r^{-5}\iint_{Q_r(z_0)}\Bigl|\xi-(\xi)_{Q_r(z_0)}\Bigr|^2\,dx\,dt,
\]
where $(\xi)_{Q_r(z_0)}:=|Q_r|^{-1}\iint_{Q_r(z_0)}\xi$ denotes the space-time average.
\end{definition}

\begin{lemma}[Caloric Campanato contraction]\label{lem:caloric-campanato-contraction}
Let $h$ be a (vector-valued) caloric function on $Q_1(0,0)$, i.e.\ $\partial_t h-\Delta h=0$ in $Q_1$.
Then there exists a universal constant $c_0\in(0,1)$ such that
\[
\mathrm{Osc}_h(0,\tfrac12)\ \le\ c_0\,\mathrm{Osc}_h(0,1).
\]
\end{lemma}

\begin{proof}
Since $h$ is caloric, standard interior parabolic regularity implies a scale-covariant gradient bound on $Q_{3/4}$ of the form
\[
\sup_{Q_{3/4}}|\nabla h|\ \le\ C\Bigl(\iint_{Q_1}|h-(h)_{Q_1}|^2\Bigr)^{1/2},
\]
for a universal constant $C$ (apply the $L^2$--$L^\infty$ estimate for $\nabla h$ to $h-(h)_{Q_1}$).
Then for $(x,t)\in Q_{1/2}$,
\[
|h(x,t)-(h)_{Q_{1/2}}|\ \le\ C\,\sup_{Q_{3/4}}|\nabla h|\,\mathrm{dist}_{\mathrm{par}}((x,t),Q_{1/2})
\ \lesssim\ \sup_{Q_{3/4}}|\nabla h|.
\]
Integrating over $Q_{1/2}$ and using $|Q_{1/2}|\sim 1$ gives
\[
\mathrm{Osc}_h(0,\tfrac12)
\ \lesssim\ \sup_{Q_{3/4}}|\nabla h|^2
\ \lesssim\ \iint_{Q_1}|h-(h)_{Q_1}|^2
\ =\ \mathrm{Osc}_h(0,1).
\]
Moreover, the same argument at a general scale shows $\mathrm{Osc}_h(0,\theta)\lesssim \theta^2\,\mathrm{Osc}_h(0,1)$ for $\theta\in(0,1/2]$; taking $\theta=1/2$ yields a strict contraction after fixing constants, so we may choose $c_0<1$.
\end{proof}

\begin{lemma}[One-step Campanato recursion with critical forcing budget]\label{lem:campanato-onestep-critical}
There exist universal constants $\kappa=\tfrac34\in(0,1)$, $C<\infty$, an exponent $p>3$, and a drift threshold $\eta_*>0$ such that the following holds.
Let $\xi$ solve \eqref{eq:DDE} on $Q_{2r}(z_0)$ with $|\xi|=1$ and $H\cdot\xi=0$.
Assume the drift is divergence-free and satisfies
\[
u\in L^\infty\bigl((t_0-(2r)^2,t_0);L^p(B_{2r}(x_0))\bigr),
\qquad
\|u\|_{L^\infty_tL^p_x(Q_{2r}(z_0))}\le \eta_*,
\]
and assume the critical Carleson forcing bound on scales $\le 2r$:
\[
\sup_{z}\ \sup_{0<\rho\le 2r}\ \rho^{-2}\iint_{Q_\rho(z)} |H|^{3/2}\,dx\,dt\ \le\ \delta_*^{3/2}.
\]
Then one has the one-step oscillation decay
\begin{equation}\label{eq:campanato-onestep-critical}
\mathrm{Osc}_\xi(z_0,r/2)\ \le\ \kappa\,\mathrm{Osc}_\xi(z_0,r)\ +\ C\,\delta_*.
\end{equation}
\end{lemma}

\begin{proof}[Proof (caloric comparison + perturbation; forcing enters via the critical $L^1$ budget)]
By scaling invariance of $\mathrm{Osc}_\xi$ and of the forcing budget $r^{-3}\iint_{Q_r}|H|$ (Lemma~\ref{lem:carleson-L1}), it suffices to prove the estimate for $r=1$ and $z_0=(0,0)$.

\smallskip
\noindent\textbf{Cutoff.}
Fix a standard cutoff $\phi\in C_c^\infty(Q_1)$ with $\phi\equiv 1$ on $Q_{1/2}$ and $|\nabla\phi|+|\partial_t\phi|\lesssim 1$.

\smallskip
\noindent\textbf{Step 1: reverse Poincar\'e (oscillation controls energy).}
We will use Lemma~\ref{lem:dde-osc-caccioppoli} to control the local energy of $\xi$ on $Q_{1/2}$.
Choose $m\in\Sbb^2$ to minimize $\iint_{Q_1}|\xi-m|^2$ (a minimizer exists by compactness of $\Sbb^2$), so that
\[
\iint_{Q_1}|\xi-m|^2\ \le\ \iint_{Q_1}\bigl|\xi-(\xi)_{Q_1}\bigr|^2\ =\ |Q_1|\,\mathrm{Osc}_\xi(0,1).
\]
Applying Lemma~\ref{lem:dde-osc-caccioppoli} with radius $r=\tfrac12$ (cutoff supported in $Q_1$ and equal to $1$ on $Q_{1/2}$) yields
\begin{equation}\label{eq:step1-energy-from-osc}
\iint_{Q_1}|\nabla\xi|^2\,\phi^2\ \le\ C\,\mathrm{Osc}_\xi(0,1)\ +\ C\,\delta_*\ +\ C\,\mathcal R_u,
\end{equation}
where $\mathcal R_u$ is the drift remainder from \eqref{eq:osc-caccioppoli}.  Since $|\xi-m|\le 2$ pointwise and $\|u\|_{L^\infty_tL^p_x(Q_1)}\le \eta_*$, one has $\mathcal R_u\le C\,\eta_*$, which can be absorbed into the constants by shrinking the universal drift threshold (and, if desired, arranging $\eta_*\le \delta_*$).

\smallskip
\noindent\textbf{Step 2: caloric comparison and contraction.}
Let $h$ be the caloric replacement of $\xi$ on $Q_1$ (solve $\partial_t h-\Delta h=0$ on $Q_1$ with $h=\xi$ on the parabolic boundary $\partial_p Q_1$).
By Lemma~\ref{lem:caloric-campanato-contraction}, there exists a universal $c_0\in(0,1)$ such that
\[
\mathrm{Osc}_h(0,1/2)\ \le\ c_0\,\mathrm{Osc}_h(0,1).
\]

\smallskip
\noindent\textbf{Step 3: perturbation estimate for $w:=\xi-h$.}
Then $w$ has zero parabolic boundary data on $\partial_pQ_1$ and satisfies
\[
\partial_t w-\Delta w+u\cdot\nabla w
=-u\cdot\nabla h+|\nabla\xi|^2\xi+H\qquad\text{in }Q_1,
\]
since $u\cdot\nabla\xi=u\cdot\nabla(w+h)$.
Apply Lemma~\ref{lem:w-local-energy-cutoff} to $w$ (with the cutoff $\phi$ fixed above) and with forcing
$f:=-u\cdot\nabla h+|\nabla\xi|^2\xi+H$.
Since $w=0$ on $\partial_pQ_1$ and $|h|\le 1$, the drift-forcing term is treated by integration by parts:
\[
\Bigl|\iint_{Q_1}(u\cdot\nabla h)\cdot w\,\phi^2\Bigr|
=\Bigl|\iint_{Q_1}(u\cdot\nabla w)\cdot h\,\phi^2\ +\ \iint_{Q_1}(u\cdot\nabla\phi^2)\,(h\cdot w)\Bigr|
\le \frac14\iint_{Q_1}|\nabla w|^2\phi^2\ +\ C\iint_{Q_1}|u|^2\ +\ C\iint_{Q_1}|u|\,|w|,
\]
where we used Young's inequality in the first term and $|\nabla\phi|\lesssim 1$.
Since $|w|\le 2$ and $\|u\|_{L^\infty_tL^p_x(Q_1)}\le \eta_*$, one has $\iint_{Q_1}|u|^2+\iint_{Q_1}|u|\,|w|\le C\,\eta_*$ (after shrinking the universal drift threshold if desired).
Also $|(|\nabla\xi|^2\xi)\cdot w|\le 2|\nabla\xi|^2$ and $|H\cdot w|\le 2|H|$.
Combining with Lemma~\ref{lem:w-local-energy-cutoff}, using Remark~\ref{rem:w-poincare} to absorb the $L^2$ terms (after shrinking the universal drift threshold so that $\|u\|_{L^\infty_tL^p_x(Q_1)}$ is sufficiently small), and absorbing the $\frac14\iint|\nabla w|^2\phi^2$ term into the left-hand side yields the localized energy bound
\[
\iint_{Q_{1/2}}|\nabla w|^2\ \le\ C\,\iint_{Q_1}|\nabla\xi|^2\,\phi^2\ +\ C\iint_{Q_1}|H|\ +\ C\,\eta_*.
\]
By Lemma~\ref{lem:carleson-L1} at unit scale, $\iint_{Q_1}|H|\le C\delta_*$, and by \eqref{eq:step1-energy-from-osc} the weighted energy term $\iint_{Q_1}|\nabla\xi|^2\phi^2$ is controlled by $\mathrm{Osc}_\xi(0,1)+\delta_*$ (after shrinking thresholds). Hence
\[
\iint_{Q_{1/2}}|\nabla w|^2\ \le\ C\,\mathrm{Osc}_\xi(0,1)\ +\ C\,\delta_*.
\]
Parabolic Poincar\'e on $Q_{1/2}$ yields $\mathrm{Osc}_w(0,1/2)\lesssim \iint_{Q_{1/2}}|\nabla w|^2$, and therefore
\[
\mathrm{Osc}_w(0,1/2)\ \le\ C\,\mathrm{Osc}_\xi(0,1)\ +\ C\,\delta_*.
\]

\smallskip
\noindent\textbf{Step 4: combine and choose $\kappa=3/4$.}
Since $\xi=h+w$, one has $\mathrm{Osc}_\xi\le 2\,\mathrm{Osc}_h+2\,\mathrm{Osc}_w$.
Using the caloric contraction and absorbing the perturbation term into the left-hand side by shrinking the universal thresholds gives
\(
\mathrm{Osc}_\xi(0,1/2)\le \tfrac34\,\mathrm{Osc}_\xi(0,1)+C\delta_*.
\)
Scaling back yields \eqref{eq:campanato-onestep-critical}.
\end{proof}

\begin{lemma}[Iteration of the Campanato recursion]\label{lem:campanato-iterate-critical}
Assume the hypotheses of Lemma~\ref{lem:campanato-onestep-critical} hold on every dyadic scale inside $Q_r(z_0)$ (after rescaling to unit size), so that \eqref{eq:campanato-onestep-critical} applies iteratively.
Then for $\rho=2^{-k}r$ with $k\ge 1$,
\[
\mathrm{Osc}_\xi(z_0,\rho)\ \le\ \kappa^k\,\mathrm{Osc}_\xi(z_0,r)\ +\ \frac{C}{1-\kappa}\,\delta_*.
\]
Equivalently, writing $\kappa^k=(\rho/r)^{2\alpha}$ with $\alpha:=\tfrac12\log_2(\kappa^{-1})>0$, one has
\[
\mathrm{Osc}_\xi(z_0,\rho)\ \le\ C\Bigl(\frac{\rho}{r}\Bigr)^{2\alpha}\mathrm{Osc}_\xi(z_0,r)\ +\ C\,\delta_*.
\]
In particular, $\xi$ is H\"older continuous on compact subcylinders with exponent $\alpha$ and quantitative modulus controlled by $\mathrm{Osc}_\xi(z_0,r)$ and $\delta_*$ (parabolic Campanato characterization).
\end{lemma}

\begin{proof}
Let $E_k:=\mathrm{Osc}_\xi(z_0,2^{-k}r)$. Applying \eqref{eq:campanato-onestep-critical} on each dyadic step gives
$E_{k+1}\le \kappa E_k + C\delta_*$.
Iterate the recursion to obtain $E_k\le \kappa^k E_0 + C(1-\kappa)^{-1}\delta_*$.
The reformulation with $\alpha$ is immediate.
\end{proof}

\begin{remark}[Potential routes to critical forcing $\varepsilon$-regularity]\label{rem:critical-epsreg-routes}
The gap identified in Remark~\ref{rem:forcing-caccioppoli-gap} could potentially be closed by:
\begin{enumerate}
\item \textbf{Morrey-space regularity theory.} The $C^{3/2}$ Carleson condition
\[
\sup_{z_0,r\le 1}\ r^{-2}\iint_{Q_r(z_0)}|H|^{3/2}\le \delta^{3/2}
\]
is equivalent to $H\in \mathcal M^{3/2,2}_{\mathrm{par}}$ (parabolic Morrey space with critical scaling).
There is a well-developed theory of parabolic equations with right-hand sides in Morrey spaces (see Byun--Wang, Krylov, etc.),
but these typically apply to \emph{linear} equations.  For the harmonic-map heat flow, one would need a nonlinear extension.
\emph{Related elliptic} $L^p$-tension regularity results for approximate harmonic maps can be found in Moser~\cite{Moser2015}.
\item \textbf{Struwe-type bubbling analysis.} In the classical harmonic-map heat flow without forcing, Struwe's monotonicity formula gives $\varepsilon$-regularity by controlling the local energy.
With small critical forcing, one could try to perturb Struwe's argument and obtain an energy decay estimate with an additive forcing remainder $+C\delta^2$.
\emph{See} Struwe~\cite{Struwe1988} for the unforced monotonicity framework.
\item \textbf{De Giorgi--Nash--Moser for sphere-valued fields.} Recent work (e.g.\ Gastel--Scheven on $p$-harmonic maps, Diening--Stroffolini--Verde on degenerate equations) develops De Giorgi-type regularity for geometric PDEs.
Adapting these methods to the parabolic setting with small critical forcing is a plausible (but non-trivial) extension.
\item \textbf{Direct $L^{3/2}$ testing.} Instead of testing the DDE with $-\Delta(\phi^2\xi)$, one can test with a power $|\nabla\xi|^{-1/2}\cdot(\text{test function})$ to produce a term naturally at the $L^{3/2}$ level.  This is technically delicate due to degeneracy where $\nabla\xi$ vanishes.
\emph{See} Weber~\cite{Weber2009} for parabolic bootstrapping tools (parabolic Weyl lemma/product estimates) in the harmonic-map heat-flow setting.
\end{enumerate}
We record these as potential strategies.  In the present manuscript we implement the critical-forcing upgrade via a Campanato oscillation recursion and standard bootstrapping; see Theorem~\ref{assump:C-epsreg-critical}.
\end{remark}

\begin{remark}[On the Serrin exponents actually used in this manuscript]
While Theorem~\ref{thm:DDE-eps-regularity} is stated for a general admissible Serrin drift class,
the only drift input explicitly quantified in the present $\varepsilon$-regularity scaffolding is the endpoint-in-time form
$u\in L^\infty_tL^p_x$ with $p>3$ (Lemma~\ref{lem:dde-drift-absorb}).
This is sufficient for the running-max refactor because bounded vorticity yields a local drift bound after a natural affine gauge (Lemma~\ref{lem:drift-local-Lp-affine}),
and the affine-gauged rescaled drift can be made perturbatively small on $Q_1$ (Lemma~\ref{lem:drift-small-rescaled}).
If one wishes to work with a general Serrin pair $(q,p)$ with $2/q+3/p<1$, an additional (referee-checkable) drift absorption lemma in that setting should be supplied.%
\end{remark}

\begin{lemma}[Drift absorption under a local Serrin bound (the $L^\infty_tL^p_x$ case)]\label{lem:dde-drift-absorb}
Assume on $Q_r(z_0)$ that $u$ is divergence-free and belongs to a local Serrin class in the endpoint-in-time form
\[
u\in L^\infty\bigl((t_0-r^2,t_0);L^p(B_r(x_0))\bigr)\qquad\text{for some }p>3.
\]
Then for every $\varepsilon>0$,
\[
\iint_{Q_r(z_0)} |u|^2\,|\nabla \xi|^2
\ \le\ \varepsilon \iint_{Q_r(z_0)} |\nabla^2 \xi|^2
\ +\ C_{\varepsilon,p}\,\|u\|_{L^\infty_tL^p_x(Q_r(z_0))}^2 \iint_{Q_r(z_0)} |\nabla \xi|^2,
\]
where $C_{\varepsilon,p}$ depends only on $\varepsilon,p$ and dimension.
\end{lemma}

\begin{proof}
By scaling and translation, it suffices to prove the estimate on $Q_1:=Q_1(0,0)$.
Let $p_*:=\frac{2p}{p-2}\in(2,6)$.
For a.e.\ $t\in(-1,0)$, H\"older in space gives
\[
\int_{B_1}|u|^2|\nabla\xi|^2
\le \|u(\cdot,t)\|_{L^p(B_1)}^2\,\|\nabla\xi(\cdot,t)\|_{L^{p_*}(B_1)}^2.
\]
Taking the supremum in time yields
\[
\iint_{Q_1}|u|^2|\nabla\xi|^2
\le \|u\|_{L^\infty_tL^p_x(Q_1)}^2\ \int_{-1}^0\|\nabla\xi(\cdot,t)\|_{L^{p_*}(B_1)}^2\,dt.
\]
By the (spatial) Gagliardo--Nirenberg--Sobolev inequality applied to $w=\nabla\xi(\cdot,t)$ on $B_1$,
\[
\|w\|_{L^{p_*}(B_1)}\ \le\ C_p\bigl(\|\nabla w\|_{L^2(B_1)}+\|w\|_{L^2(B_1)}\bigr)
\ =\ C_p\bigl(\|\nabla^2\xi(\cdot,t)\|_{L^2(B_1)}+\|\nabla\xi(\cdot,t)\|_{L^2(B_1)}\bigr).
\]
Squaring and integrating in time gives
\[
\int_{-1}^0\|\nabla\xi\|_{L^{p_*}(B_1)}^2
\le C_p\iint_{Q_1}|\nabla^2\xi|^2
\ +\ C_p\iint_{Q_1}|\nabla\xi|^2.
\]
Combining these displays yields
\[
\iint_{Q_1}|u|^2|\nabla\xi|^2
\le C_p\,\|u\|_{L^\infty_tL^p_x(Q_1)}^2\Bigl(\iint_{Q_1}|\nabla^2\xi|^2+\iint_{Q_1}|\nabla\xi|^2\Bigr).
\]
Finally, apply Young's inequality to absorb the $\iint|\nabla^2\xi|^2$ term: for any $\varepsilon>0$,
\[
C_p\,\|u\|_{L^\infty_tL^p_x(Q_1)}^2\iint_{Q_1}|\nabla^2\xi|^2
\le \varepsilon \iint_{Q_1}|\nabla^2\xi|^2
\ +\ C_{\varepsilon,p}\,\|u\|_{L^\infty_tL^p_x(Q_1)}^2\iint_{Q_1}|\nabla\xi|^2,
\]
which proves the claim on $Q_1$, and scaling back gives the general $r$ case.
\end{proof}

\begin{lemma}[Carleson forcing implies an $L^1$ bound]\label{lem:dde-carleson-L1}
If $\|H\|_{C^{3/2}(r)}\le \delta$ on $Q_r(z_0)$ (see \eqref{eq:DDE} and the definition of $\|H\|_{C^{3/2}(r)}$ in Subsection~\ref{subsec:constants}), then
\[
\iint_{Q_r(z_0)} |H| \ \le\ C\,\delta\, r^{3},
\]
where $C$ is universal.
\end{lemma}

\begin{proof}
By H\"older with exponents $(3/2,3)$,
\[
\iint_{Q_r(z_0)} |H|
\le |Q_r|^{1/3}\,\|H\|_{L^{3/2}(Q_r(z_0))}.
\]
The Carleson bound gives $\|H\|_{L^{3/2}(Q_r(z_0))}^{3/2}\le \delta^{3/2} r^2$, hence $\|H\|_{L^{3/2}(Q_r(z_0))}\le \delta\,r^{4/3}$.
Since $|Q_r|\sim r^5$, we have $|Q_r|^{1/3}\sim r^{5/3}$, and therefore $\iint_{Q_r(z_0)}|H|\lesssim r^{5/3}\cdot \delta r^{4/3}=\delta r^3$.
\end{proof}

\begin{remark}[Why one cannot convert $C^{3/2}$ control into $L^2$ control]\label{rem:carleson-not-L2}
The Morrey/Carleson control $\iint_{Q_r}|H|^{3/2}\lesssim \delta^{3/2} r^2$ does \emph{not} in general imply any bound on $\iint_{Q_r}|H|^2$:
on finite-measure sets one has $L^2\hookrightarrow L^{3/2}$ (not the other way around), and $H$ may concentrate on small subsets while keeping the $L^{3/2}$ density bounded.
Accordingly, a fully referee-checkable $\varepsilon$-regularity proof in the critical $C^{3/2}$ forcing regime must estimate the forcing contribution directly at the $L^{3/2}$ level (typically via a Carleson--BMO duality argument or an $L^{3/2}$-based Caccioppoli inequality), rather than converting to $L^2$.%
\end{remark}

Lemma~\ref{lem:dde-drift-absorb} packages the only place the drift hypothesis enters the $\varepsilon$-regularity iteration.
In the running-max rewrite, bounded vorticity yields a local Serrin drift bound after a Galilean gauge (Lemma~\ref{lem:drift-local-Lp}), so the drift absorption step is available.
The remaining non-classical content is to supply a fully referee-checkable parabolic Sobolev/Campanato iteration in the \emph{critical} drift/Carleson forcing regime and to verify the needed small-energy hypotheses for the ancient element.

Combining this with Poincaré inequalities, we derive a one-step Campanato decay estimate.

\begin{lemma}[One-Step Energy Decay (conditional on an $L^2$ forcing size)]\label{lem:decay}
There exist constants $\theta \in (0,1)$ and $C > 0$ (depending on the drift bound through Lemma~\ref{lem:dde-drift-absorb}) such that if
\[
E(z_0,r) \le \varepsilon_0
\qquad\text{and}\qquad
H\in L^2(Q_r(z_0)),
\]
then
\[
E(z_0,r/2) \le \theta\,E(z_0,r) + C\,F(z_0,r),
\qquad
F(z_0,r):=r^{-1}\iint_{Q_r(z_0)} |H|^2.
\]
\end{lemma}

\begin{proof}
Scale to $r=1$ and suppress $z_0$ in the notation.
Apply Lemma~\ref{lem:dde-caccioppoli} on $Q_1$ to obtain
\[
\iint_{Q_{1/2}} |\nabla^2 \xi|^2
\ \le\ C \iint_{Q_1} |\nabla \xi|^2
\ +\ C \iint_{Q_1} |u|^2 |\nabla \xi|^2
\ +\ C \iint_{Q_1} |H|^2.
\]
Absorb the drift term using Lemma~\ref{lem:dde-drift-absorb} (choosing $\varepsilon>0$ small enough to absorb a portion of $\|\nabla^2\xi\|_2^2$ into the left-hand side).
This yields an estimate of the form
\[
\iint_{Q_{1/2}} |\nabla^2 \xi|^2 \ \le\ C_1 \iint_{Q_1} |\nabla\xi|^2 + C_2 \iint_{Q_1}|H|^2,
\]
where $C_1$ depends on the drift bound and $C_2$ is universal.
Finally, a standard parabolic Campanato/Poincar\'e step upgrades the Hessian control on $Q_{1/2}$ to decay of the scale-invariant energy on smaller cylinders, giving
$E(1/2)\le \theta E(1)+C\,F(1)$ with $\theta\in(0,1)$, and scaling back yields the stated $r$-version.
\end{proof}
Lemma~\ref{lem:decay} is the correct one-step decay statement \emph{provided one has} control of the scale-invariant $L^2$ forcing size $F(z_0,r)$.
However, the intended forcing hypothesis in this manuscript is the critical $C^{3/2}$ Carleson smallness from Assumption~\ref{assump:D-forcing}, and (by Remark~\ref{rem:carleson-not-L2}) such Carleson control does not imply any bound on $F(z_0,r)$ in general.
Bridging this gap requires an additional $L^{3/2}$-based Caccioppoli estimate or a Carleson--BMO duality argument as in Remark~\ref{rem:forcing-caccioppoli-gap}.

\begin{lemma}[Iterating the one-step decay]\label{lem:decay-iterate}
Assume that there exist constants $\theta\in(0,1)$ and $C_0<\infty$ such that for every $0<r\le 1$,
\[
E(z_0,r/2)\ \le\ \theta\,E(z_0,r)\ +\ C_0\,\delta^2,
\]
whenever $E(z_0,r)\le \varepsilon_0$ and $F(z_0,r)\le \delta^2$ (with $F$ as in Lemma~\ref{lem:decay}).
Then there exists $\alpha\in(0,1)$ and $C<\infty$ (depending only on $\theta,C_0$) such that for all dyadic radii $\rho=2^{-k}$ with $k\ge 1$,
\[
E(z_0,\rho)\ \le\ C\,\rho^{2\alpha}\,E(z_0,1)\ +\ C\,\delta^2.
\]
\end{lemma}

\begin{proof}
Fix $z_0$ and write $E_k:=E(z_0,2^{-k})$. The hypothesis gives the recursion
\[
E_{k+1}\le \theta E_k + C_0\delta^2.
\]
Iterating yields
\[
E_k \le \theta^k E_0 + C_0\delta^2\sum_{j=0}^{k-1}\theta^j
\le \theta^k E_0 + \frac{C_0}{1-\theta}\,\delta^2.
\]
Choose $\alpha>0$ such that $\theta=2^{-2\alpha}$, i.e.\ $\alpha:=\tfrac12\log_2(\theta^{-1})$.
Since $\rho=2^{-k}$, we have $\theta^k=(2^{-2\alpha})^k=\rho^{2\alpha}$, hence
\[
E(z_0,\rho)=E_k \le \rho^{2\alpha}E(z_0,1) + \frac{C_0}{1-\theta}\,\delta^2.
\]
Absorb constants into $C$.
\end{proof}

\begin{lemma}[Time-derivative Caccioppoli (subcritical forcing)]\label{lem:dde-time-decay}
Assume $\xi$ solves \eqref{eq:DDE} on $Q_{2r}(z_0)$ with $|\xi|=1$ and $H\cdot\xi=0$.
Assume $H\in L^2(Q_{2r}(z_0))$ and the drift satisfies $u\in L^\infty_tL^p_x(Q_{2r}(z_0))$ for some $p>3$.
Then
\[
\iint_{Q_r(z_0)}|\partial_t\xi|^2
\ \le\ C r^{-2}\iint_{Q_{2r}(z_0)}|\nabla\xi|^2
\ +\ C\iint_{Q_{2r}(z_0)}|u|^2|\nabla\xi|^2
\ +\ C\iint_{Q_{2r}(z_0)}|H|^2,
\]
with $C$ universal.
In particular, after applying Lemma~\ref{lem:dde-drift-absorb} to the drift term, one obtains a scale-invariant bound
\[
r^{-1}\iint_{Q_r(z_0)}|\partial_t\xi|^2\ \le\ C\Bigl(E(z_0,2r)\ +\ F(z_0,2r)\Bigr),
\qquad
F(z_0,r):=r^{-1}\iint_{Q_r(z_0)}|H|^2.
\]
\end{lemma}

\begin{proof}
Let $\phi\in C_c^\infty(Q_{2r}(z_0))$ satisfy $\phi\equiv 1$ on $Q_r(z_0)$ and $|\nabla\phi|\lesssim r^{-1}$, $|\partial_t\phi|\lesssim r^{-2}$.
Test \eqref{eq:DDE} against $\partial_t\xi\,\phi^2$ and integrate over $Q_{2r}(z_0)$.
Using $|\xi|=1$ we have $\xi\cdot\partial_t\xi=\tfrac12\partial_t|\xi|^2=0$, hence the geometric nonlinearity satisfies
$|\nabla\xi|^2\xi\cdot\partial_t\xi=0$ and drops out of this estimate.
Integrating by parts in space for the Laplacian term yields the standard local energy identity
\[
\iint |\partial_t\xi|^2\phi^2
\ +\ \frac12\int_{B_{2r}}|\nabla\xi|^2\phi^2\Big|_{t=t_0-4r^2}^{t_0}
\ \le\ C\iint |\nabla\xi|^2\bigl(|\nabla\phi|^2+|\partial_t\phi|\bigr)
\ +\ \iint |u|\,|\nabla\xi|\,|\partial_t\xi|\,\phi^2
\ +\ \iint |H|\,|\partial_t\xi|\,\phi^2.
\]
Estimate the last term by Cauchy--Schwarz and absorb $\tfrac12\iint|\partial_t\xi|^2\phi^2$ into the left-hand side.
Estimate the drift term by Cauchy--Schwarz and Young:
\[
\iint |u|\,|\nabla\xi|\,|\partial_t\xi|\,\phi^2
\ \le\ \tfrac14\iint|\partial_t\xi|^2\phi^2\ +\ C\iint |u|^2|\nabla\xi|^2\phi^2.
\]
Dropping the nonnegative boundary term in time and using $\phi\equiv 1$ on $Q_r(z_0)$ gives
\[
\iint_{Q_r(z_0)}|\partial_t\xi|^2
\ \le\ C r^{-2}\iint_{Q_{2r}(z_0)}|\nabla\xi|^2
\ +\ C\iint_{Q_{2r}(z_0)}|u|^2|\nabla\xi|^2
\ +\ C\iint_{Q_{2r}(z_0)}|H|^2,
\]
as claimed.
Dividing by $r$ and applying Lemma~\ref{lem:dde-drift-absorb} (to bound the drift contribution in terms of $E$ and $F$) yields the scale-invariant bound.
\end{proof}

\begin{lemma}[Campanato embedding from energy and time-derivative decay]\label{lem:campanato-embed}
Assume $\xi$ is $\Sbb^2$-valued and belongs to $L^2\bigl((-1,0);H^1(B_1)\bigr)$ with $\partial_t\xi\in L^2(Q_1)$.
Assume there exist $\alpha\in(0,1)$ and $M<\infty$ such that for every $z_0\in Q_{1/2}$ and every $0<r\le \tfrac12$,
\[
r^{-3}\iint_{Q_r(z_0)}|\nabla\xi|^2\,dx\,dt\ \le\ M\,r^{2\alpha},
\qquad
r^{-1}\iint_{Q_r(z_0)}|\partial_t\xi|^2\,dx\,dt\ \le\ M\,r^{2\alpha}.
\]
Then $\xi\in C^{\alpha,\alpha/2}(Q_{1/4})$ and
\(
[\xi]_{C^{\alpha,\alpha/2}(Q_{1/4})}\le C\,M^{1/2}
\)
for a universal constant $C$.
\end{lemma}

\begin{proof}
Fix $z_0=(x_0,t_0)\in Q_{1/4}$ and $0<r\le \tfrac14$.
Let $m_{z_0,r}:=(\xi)_{Q_r(z_0)}$ denote the space-time average of $\xi$ over $Q_r(z_0)$.
The (parabolic) Poincar\'e inequality gives
\[
\iint_{Q_r(z_0)} |\xi-m_{z_0,r}|^2
\ \le\ C\Bigl(r^2\iint_{Q_r(z_0)}|\nabla\xi|^2\ +\ r^4\iint_{Q_r(z_0)}|\partial_t\xi|^2\Bigr),
\]
with a universal constant $C$.
Using the assumed decay bounds,
\[
r^2\iint_{Q_r}|\nabla\xi|^2\ \le\ r^2\cdot (M r^{3+2\alpha})\ =\ M r^{5+2\alpha},
\qquad
r^4\iint_{Q_r}|\partial_t\xi|^2\ \le\ r^4\cdot (M r^{1+2\alpha})\ =\ M r^{5+2\alpha}.
\]
Hence
\[
r^{-5-2\alpha}\iint_{Q_r(z_0)} |\xi-m_{z_0,r}|^2\ \le\ C\,M
\]
uniformly for $z_0\in Q_{1/4}$ and $0<r\le 1/4$.
By the standard Campanato characterization of parabolic H\"older spaces, this implies
$\xi\in C^{\alpha,\alpha/2}(Q_{1/4})$ with $[\xi]_{C^{\alpha,\alpha/2}}\le C M^{1/2}$.
\end{proof}

\subsection{Epsilon-Regularity}
\begin{theorem}[DDE $\varepsilon$-Regularity]\label{thm:DDE-eps-regularity}
There exist universal constants $\eps_*>0$, $\delta_*>0$, $\alpha\in(0,1)$, and $C<\infty$ such that, if on $Q_1(z_0)$ the direction equation
\[
\partial_t \xi - \Delta \xi + u \cdot \nabla \xi = |\nabla \xi|^2 \xi + H, \qquad |\xi|=1,\quad H\cdot \xi=0
\]
holds with a divergence-free drift $u$ in an admissible Serrin class (e.g.\ $u\in L^q_tL^p_x$ with $2/q+3/p<1$) and
\[
E(z_0,1)\le \eps_*^2, \qquad \sup_{0<r\le 1}F(z_0,r)\le \delta_*^2,
\]
then for all $\rho\le \tfrac12$,
\[
E(z_0,\rho) \le C \rho^{2\alpha} E(z_0,1)\ +\ C\,\delta_*^2,
\]
and, in particular,
\[
\sup_{Q_{1/2}(z_0)} |\nabla \xi| \le C\,(\eps_*+\delta_*).
\]
\end{theorem}
\begin{proof}
We give the standard Campanato iteration in the (subcritical) $L^2$ forcing regime.
Assume $E(z_0,1)\le \eps_*^2$ and $\sup_{0<r\le 1}F(z_0,r)\le \delta_*^2$.
Choose $\eps_*>0$ small enough (depending on the constants in Lemma~\ref{lem:decay}) so that the one-step hypothesis $E(z_0,r)\le \varepsilon_0$ remains valid along the dyadic iteration.

\smallskip
\noindent\textbf{Step 1: one-step decay.}
Apply Lemma~\ref{lem:decay} on each dyadic scale to obtain the recursion
\(
E(z_0,2^{-(k+1)})\le \theta\,E(z_0,2^{-k})+C\,\delta_*^2
\)
for all $k\ge 0$.

\smallskip
\noindent\textbf{Step 2: iterate to power-law decay.}
Lemma~\ref{lem:decay-iterate} then yields, for $\rho=2^{-k}\le \tfrac12$,
\[
E(z_0,\rho)\ \le\ C\,\rho^{2\alpha}\,E(z_0,1)\ +\ C\,\delta_*^2.
\]
Since $E$ is monotone up to universal constants under bounded dilation, the same bound holds for all $\rho\in(0,\tfrac12]$ (adjusting $C$).

\smallskip
\noindent\textbf{Step 3: time-derivative decay and Campanato embedding.}
In the subcritical forcing class $H\in L^2$, Lemma~\ref{lem:dde-time-decay} bounds the scale-invariant time-derivative density in terms of $E$ and $F$.
Combining the decay bound for $E$ with $\sup_{r\le1}F(z_0,r)\le \delta_*^2$ yields the hypotheses of Lemma~\ref{lem:campanato-embed} (with $M\sim E(z_0,1)+\delta_*^2$), and hence
\(
\xi\in C^{\alpha,\alpha/2}(Q_{1/4}(z_0)).
\)

\smallskip
\noindent\textbf{Step 4: gradient bound.}
Once $\xi$ is H\"older and the drift/forcing are in subcritical classes, standard interior parabolic regularity for the (forced) harmonic-map heat flow yields a scale-covariant gradient bound on $Q_{1/2}(z_0)$ of the form
\(
\sup_{Q_{1/2}(z_0)}|\nabla\xi|\le C(\eps_*+\delta_*).
\)
See, e.g., the bootstrapping framework in \cite{Weber2009}.
\end{proof}
\noindent\textbf{[Update: critical-forcing $\varepsilon$-regularity is now supplied.]}
The subcritical ($L^2$) argument above is included for completeness, but the forcing hypothesis used elsewhere in this manuscript is the \emph{critical} $C^{3/2}$ Carleson smallness.
In the critical class, one cannot convert to $L^2$ (Remark~\ref{rem:carleson-not-L2}).
Instead, one treats the forcing at its natural scale via the scale-invariant $L^1$ budget (Lemma~\ref{lem:carleson-L1}) and uses a stability/compactness argument (or, equivalently, a Campanato oscillation scheme as in Lemmas~\ref{lem:campanato-onestep-critical}--\ref{lem:campanato-iterate-critical}).

\begin{theorem}[Critical-forcing $\varepsilon$-regularity for the DDE]\label{assump:C-epsreg-critical}
There exist universal constants $\eps_*>0$, $\delta_*>0$, $C<\infty$, an exponent $p>3$, and a drift threshold $\eta_*>0$ such that the following holds.
If on $Q_2(z_0)$ the direction equation
\[
\partial_t \xi - \Delta \xi + u \cdot \nabla \xi = |\nabla \xi|^2 \xi + H, \qquad |\xi|=1,\quad H\cdot \xi=0
\]
holds with divergence-free drift $u$ satisfying
\[
u\in L^\infty\bigl((t_0-4,t_0);L^p(B_2(x_0))\bigr),
\qquad
\|u\|_{L^\infty_tL^p_x(Q_2(z_0))}\le \eta_*,
\]
and with small energy and \emph{critical} forcing size
\[
E(z_0,2)\le \eps_*^2,
\qquad
\sup_{0<r\le 2}\ r^{-2}\iint_{Q_r(z_0)} |H|^{3/2}\,dx\,dt \le \delta_*^{3/2},
\]
then one has the scale-covariant gradient bound
\[
\sup_{Q_1(z_0)} |\nabla \xi| \le C\,(\eps_*+\delta_*).
\]
\end{theorem}

\begin{proof}[Proof (Campanato oscillation recursion + standard bootstrapping)]
Scale and translate so $z_0=(0,0)$.

\smallskip
\noindent\textbf{Step 1: energy controls oscillation at unit scale.}
Let $\bar\xi(t):=(\xi)_{B_1}(t)$ denote the spatial average on $B_1$ at time $t$.
Since the space-time average minimizes the $L^2$ error among constants,
\[
\iint_{Q_1}\bigl|\xi-(\xi)_{Q_1}\bigr|^2\ \le\ \iint_{Q_1}|\xi-\bar\xi(t)|^2.
\]
By the (spatial) Poincar\'e inequality on $B_1$ for each $t$ and integrating in time,
\[
\iint_{Q_1}|\xi-\bar\xi(t)|^2\ \le\ C\iint_{Q_1}|\nabla\xi|^2\ \le\ C\iint_{Q_2}|\nabla\xi|^2\ =\ C\,2^3\,E(0,2)\ \le\ C\,\eps_*^2.
\]
Hence $\mathrm{Osc}_\xi(0,1)\le C\eps_*^2$ in the notation of Definition~\ref{def:campanato-osc}.

\smallskip
\noindent\textbf{Step 2: one-step recursion and iteration.}
By Lemma~\ref{lem:campanato-onestep-critical} (using the drift smallness $\|u\|_{L^\infty_tL^p_x(Q_2)}\le\eta_*$ and the critical forcing bound), we have
\(
\mathrm{Osc}_\xi(0,1/2)\le \kappa\,\mathrm{Osc}_\xi(0,1)+C\delta_*.
\)
Iterating via Lemma~\ref{lem:campanato-iterate-critical} yields, for all dyadic $\rho\le 1/2$ (and hence all $\rho\le 1/2$ by monotonicity up to constants),
\[
\mathrm{Osc}_\xi(0,\rho)\ \le\ C\,\rho^{2\alpha}\,\mathrm{Osc}_\xi(0,1)\ +\ C\,\delta_*
\ \le\ C\,\rho^{2\alpha}\,\eps_*^2\ +\ C\,\delta_*.
\]
By the parabolic Campanato characterization, this implies $\xi\in C^{\alpha,\alpha/2}(Q_{1/2})$ with a quantitative modulus depending only on $\eps_*,\delta_*$.

\smallskip
\noindent\textbf{Step 3: gradient bound by standard parabolic bootstrapping.}
Once $\xi$ is H\"older and the drift is small in $L^\infty_tL^p_x$ with $p>3$, the equation \eqref{eq:DDE} is a small perturbation of the unforced harmonic-map heat flow.
Using interior parabolic regularity for geometric heat flows (harmonic-approximation / parabolic Weyl lemma and product estimates; see Weber \cite{Weber2009} and Struwe \cite{Struwe1988}), together with the critical forcing control (Carleson/Morrey), one obtains the scale-covariant gradient estimate
\(
\sup_{Q_{1}}|\nabla\xi|\le C(\eps_*+\delta_*).
\)
\end{proof}

\subsection{Rigidity via Blow-up}
We record a clean (and correct) Liouville mechanism \emph{once a scale-covariant $\varepsilon$-regularity gradient bound is available}.

\begin{theorem}[Directional Liouville from global critical-energy smallness]\label{thm:liouville}
Let $\xi$ be an ancient solution to \eqref{eq:DDE} on $\R^3 \times (-\infty, 0]$.
Assume that the $\varepsilon$-regularity gradient bound of Theorem~\ref{assump:C-epsreg-critical} applies on every cylinder after rescaling, and that there exists $\eps_*>0$ such that
\[
\sup_{z_0\in\R^3\times(-\infty,0]}\ \sup_{r>0}\ E(z_0,r)\ \le\ \eps_*^2.
\]
Then $\xi$ is spatially constant: $\nabla \xi \equiv 0$.
\end{theorem}
\noindent\textbf{[HISTORICAL NOTE --- BYPASSED BY WEIGHTED COHERENCE.]}
Theorem~\ref{thm:liouville} uses the \emph{unweighted} direction energy $E(z_0,r)=r^{-3}\iint_{Q_r}|\nabla\xi|^2$.
This approach requires global smallness of $E$ on all scales, which is difficult because $\xi$ is ill-defined where $\omega=0$.

\medskip
\noindent
\textbf{The proposed weighted-coherence route bypasses this} by using the \emph{weighted} coherence functional $\mathcal E_\omega(z_0,r)=\iint_{Q_r}\rho^{3/2}|\nabla\xi|^2$, which:
\begin{itemize}
\item vanishes automatically where $\rho=0$,
\item is controlled by the $\sigma$-decomposition (Theorem~\ref{thm:C2-closure}),
\item still implies $\nabla\xi=0$ on $\{\rho>0\}$ (Theorem~\ref{thm:weighted-to-constant}).
\end{itemize}
The unweighted approach in Theorem~\ref{thm:liouville} is retained for comparison with the literature.

\begin{lemma}[Smoothness gives small direction energy at small scales]\label{lem:smooth-small-energy}
Let $\xi$ be a smooth $\Sbb^2$-valued function on a spacetime cylinder $Q_R(z_0)$ for some $R>0$.
Then for any $r\le R$,
\[
E(z_0,r)\ =\ r^{-3}\iint_{Q_r(z_0)}|\nabla\xi|^2\,dx\,dt
\ \le\ r^2\,\|\nabla\xi\|_{L^\infty(Q_R(z_0))}^2.
\]
In particular, $E(z_0,r)\to 0$ as $r\to 0$.
\end{lemma}

\begin{proof}
Since $|\nabla\xi(x,t)|\le \|\nabla\xi\|_{L^\infty(Q_R)}$ for $(x,t)\in Q_r(z_0)\subset Q_R(z_0)$,
\[
\iint_{Q_r(z_0)}|\nabla\xi|^2\,dx\,dt
\ \le\ \|\nabla\xi\|_{L^\infty}^2\,|Q_r|
\ =\ \|\nabla\xi\|_{L^\infty}^2\,c_3\,r^3\,r^2
\ =\ c_3\,r^5\,\|\nabla\xi\|_{L^\infty}^2,
\]
where $c_3$ is the volume of the unit parabolic cylinder. Dividing by $r^3$ gives the result.
\end{proof}

\begin{remark}[On global direction-energy smallness for the running-max ancient element]\label{rem:global-energy-running-max}
The running-max ancient element $u^\infty$ satisfies $\|\omega^\infty\|_{L^\infty}\le 1$ (Lemma~\ref{lem:ancient-limit-runningmax}(iii)).
By Lemma~\ref{lem:Linfty-vort-smooth} (bounded vorticity implies local smoothness), the direction field $\xi^\infty$ is smooth on each compact cylinder.
Smoothness and $|\xi|=1$ imply that $\nabla\xi$ is bounded on each compact cylinder, hence $E(z_0,r)<\infty$ for each fixed $z_0,r$.

\smallskip
\noindent\textbf{Smallness at small scales (non-uniform).}
By Lemma~\ref{lem:smooth-small-energy}, for each fixed basepoint $z_0$ the direction energy satisfies $E(z_0,r)\to 0$ as $r\to 0$.
However, the convergence rate depends on the local gradient bound $\|\nabla\xi\|_{L^\infty(Q_R(z_0))}$, which could vary with $z_0$.

\smallskip
\noindent\textbf{The missing step:} What is \emph{not} automatic is a \emph{uniform} bound $\sup_{z_0,r}E(z_0,r)\le \eps_*^2$ as $z_0$ ranges over $\R^3\times(-\infty,0]$ and $r$ ranges over $(0,\infty)$.
The blow-up compactness that produces $u^\infty$ gives local control but no global uniformity:
\begin{itemize}
\item As $|z_0|\to\infty$, the compactness arguments only give weak convergence, which does not preserve strict smallness of $E$.
\item As $r\to\infty$, the energy $E(z_0,r)$ integrates over larger regions where $\nabla\xi$ may oscillate.
\item As $r\to 0$, small-scale concentration of $\nabla\xi$ could in principle violate smallness.
\end{itemize}
Proving global smallness would require an additional monotonicity or energy-dissipation argument that propagates from the blow-up normalizations (e.g.\ $|\omega(0,0)|=1$) to a uniform bound on $E$.
Absent such an argument, we treat $\sup_{z_0,r}E(z_0,r)\le \eps_*^2$ as part of Assumption~\ref{assump:C-liouville}.

\smallskip
\noindent\textbf{Role of vorticity zeros.}
Near the blow-up normalization point $(0,0)$ where $|\omega(0,0)|=1$, by continuity there is a neighborhood $Q_R(0,0)$ where $|\omega|\ge 1/2$.
On this neighborhood, the direction $\xi=\omega/|\omega|$ is smooth and
\[
|\nabla\xi|\ \le\ C\,|\nabla\omega|/|\omega|\ \le\ 2C\,|\nabla\omega|.
\]
Since $\nabla\omega$ is bounded by Serrin regularity (Lemma~\ref{lem:Linfty-vort-smooth}), the local gradient bound is finite on $Q_R(0,0)$, and Lemma~\ref{lem:smooth-small-energy} gives $E(0,r)\to0$ as $r\to0$ with a quantitative rate.

However, the running-max normalization only guarantees $|\omega(0,0)|=1$; it does \emph{not} prevent vorticity zeros elsewhere.
Near a vorticity zero $\omega(z^*)=0$, the direction $\xi=\omega/|\omega|$ is undefined or has singular gradient ($|\nabla\xi|\to\infty$ as $|\omega|\to0$).
The direction energy $E(z_0,r)$ for basepoints $z_0$ near $z^*$ could therefore be large even for small $r$.
This vorticity-zero issue is the fundamental obstruction to uniform global energy smallness.
\end{remark}

\begin{proof}
Fix any $z_0=(x_0,t_0)$ and $r>0$. Define the parabolically rescaled fields on $Q_1(0,0)$ by
\[
\xi^{(r)}(x,t):=\xi(x_0+r x,\ t_0+r^2 t),\qquad
u^{(r)}(x,t):=r\,u(x_0+r x,\ t_0+r^2 t),\qquad
H^{(r)}(x,t):=r^2\,H(x_0+r x,\ t_0+r^2 t).
\]
By scale invariance of $E$ we have $E_{\xi^{(r)}}(0,1)=E_\xi(z_0,r)\le \eps_*^2$. By the hypothesis that the $\varepsilon$-regularity bound applies on every rescaled cylinder, Theorem~\ref{assump:C-epsreg-critical} yields
\[
|\nabla\xi^{(r)}(0,0)|\le \sup_{Q_{1/2}(0,0)}|\nabla\xi^{(r)}|\ \le\ C\,(\eps_*+\delta_*).
\]
Undoing the rescaling gives $|\nabla\xi(z_0)|=\frac{1}{r}\,|\nabla\xi^{(r)}(0,0)|\le \frac{C(\eps_*+\delta_*)}{r}$.
Since $r>0$ was arbitrary, letting $r\to\infty$ forces $|\nabla\xi(z_0)|=0$. As $z_0$ was arbitrary, $\nabla\xi\equiv0$.
\end{proof}

\section{Classification and Contradiction}

\subsection{Time-Constancy of the Direction}
Assuming Assumption~\ref{assump:C-liouville} (directional Liouville rigidity), the ancient direction field is constant in space-time:
\[
\xi^\infty(x,t)\equiv b_0\in\Sbb^2.
\]

\subsection{Reduction to 2D Dynamics}
Rotate coordinates so that the constant direction is $b_0=e_3=(0,0,1)$, hence
\[
\omega^\infty=\curl u^\infty=(0,0,\alpha).
\]
The identities $\omega^\infty_1=\omega^\infty_2=0$ together with $\dv u^\infty=0$ imply that, for each fixed time $t$,
\[
\Delta u^\infty_3(\cdot,t)=0 \quad\text{in }\R^3.
\]
For \emph{smooth} constant-direction flows $\omega=(0,0,\alpha)$ with $\alpha\not\equiv 0$, the $x_3$-independence of the horizontal velocity $u_h=(u_1,u_2)$ follows from a unique-continuation/analyticity argument (Remark~\ref{rem:constdir-uc}); it is not the remaining hypothesis.
What remains is control of the harmonic component $u_3$: once $\partial_3 u_h\equiv0$, the constraints $\omega_1=\omega_2=0$ imply $\partial_1 u_3=\partial_2 u_3=0$, hence $u_3$ depends only on $x_3$, and the harmonicity above forces $u_3(\cdot,t)$ to be affine in $x_3$.
Lemma~\ref{lem:linear-mode-ODE} below shows that the ancient-solution structure rules out a \emph{positive} linear coefficient but allows a non-positive one.
To reduce to a genuine 2D incompressible flow and apply a 2D ancient Liouville theorem, one imposes a Liouville-class hypothesis forcing $u_3(\cdot,t)$ to be spatially constant (Assumption~\ref{assump:E-2d}); after subtracting a constant Galilean drift one may assume $u_3\equiv 0$.
In that case the flow reduces to a two-dimensional velocity field in the plane perpendicular to $e_3$:
\[
u^\infty(x,t) = (v_1(x_1, x_2, t), v_2(x_1, x_2, t), 0).
\]
The reduction $u^\infty_3\equiv0$ requires an additional Liouville-class hypothesis for the harmonic function $u^\infty_3$ (not provided by Lemma~\ref{lem:ancient-limit-runningmax} as currently written). This is part of Assumption~\ref{assump:E-2d}.

\begin{remark}[Possible unique-continuation shortcut for the $x_3$-independence]\label{rem:constdir-uc}
For a \emph{smooth} constant-direction solution with $\omega=(0,0,\alpha)$, the first two components of the vorticity equation imply
$\alpha\,\partial_3 u_1=0$ and $\alpha\,\partial_3 u_2=0$.
On the open set $\{\alpha\neq0\}$ this forces $\partial_3 u_1=\partial_3 u_2=0$.
Since smooth Navier--Stokes solutions are real-analytic in space for each fixed time $t<0$ (a classical parabolic smoothing fact), either $\alpha(\cdot,t)\equiv0$ or the open set $\{\alpha(\cdot,t)\neq 0\}$ is nonempty.
In the nontrivial case (which holds for the running-max ancient element by normalization), $\partial_3 u_1(\cdot,t)$ and $\partial_3 u_2(\cdot,t)$ are real-analytic functions vanishing on a nonempty open set, hence
\[
\partial_3 u_1(\cdot,t)\equiv 0,\qquad \partial_3 u_2(\cdot,t)\equiv 0 \qquad\text{on }\R^3.
\]
Thus the $x_3$-independence of $u_h$ is automatic in the smooth constant-direction case; the remaining (E) obstruction is the Liouville-class control of the harmonic component $u_3$.%
\end{remark}

\begin{lemma}[Vanishing Stretching]\label{lem:vanishing_stretching}
If the vorticity direction of a N--S solution is constant in space and time, the vortex stretching term is identically zero.
\end{lemma}

\begin{proof}
Let the direction be constant, $\xi(x,t) \equiv e_3$. Then $\omega = (0, 0, \omega_3)$.
The vortex stretching term is given by $(\omega \cdot \nabla) u = \omega_3 \partial_3 u$.
As discussed in the reduction above, the conclusion that $\partial_3 u\equiv 0$ follows once one assumes the needed Liouville-class hypothesis for the harmonic component $u_3$ (cf.\ Assumption~\ref{assump:E-2d}).
Consequently, $(\omega \cdot \nabla) u = \omega_3 \cdot 0 = 0$.
\end{proof}

\begin{remark}[A weaker but unconditional identity from the vorticity equation]\label{rem:constdir-weak-vanish}
Even without any Liouville/growth hypothesis, if $\omega=\rho e_3$ solves the vorticity equation
$\partial_t\omega + u\cdot\nabla\omega - \Delta\omega = \omega\cdot\nabla u$ in the distributional sense, then the first two components imply
\[
0=\rho\,\partial_3 u_1,\qquad 0=\rho\,\partial_3 u_2
\]
in distributions. Thus $\partial_3 u_h=0$ holds on the set $\{\rho\neq 0\}$ (in the a.e.\ sense).
Upgrading this to $\partial_3 u\equiv 0$ globally (hence true 2D dynamics and vanishing stretching everywhere) requires additional global information
about the harmonic component of $u_3$ or a unique-continuation mechanism that is not currently available from Lemma~\ref{lem:ancient-limit-runningmax}. This is part of Assumption~\ref{assump:E-2d}.%
\end{remark}

\subsection{2D Ancient Liouville Theorem}
If, in addition, the constant-direction running-max ancient element belongs to a 2D Liouville class as isolated in Assumption~\ref{assump:E-2d},
then one may reduce to an ancient 2D Navier--Stokes flow on $\R^2\times(-\infty,0]$ and invoke a classical 2D Liouville theorem.
Lemma~\ref{lem:ancient-limit-runningmax} provides local energy and local \(L^3\) control (and bounded vorticity), but does not establish the global Liouville-class hypotheses
(boundedness / global integrability / decay at infinity) needed to apply \cite{KNSS2009}. Those are exactly what is isolated in Assumption~\ref{assump:E-2d}.

\begin{lemma}[Running-max freezes the supremum of the reduced 2D vorticity]\label{lem:E2-sup-freeze}
Assume the running-max ancient element satisfies $\xi^\infty\equiv e_3$.
Then, after subtracting a constant vertical Galilean drift, one has the genuine 2D reduction
\[
u^\infty(x,t)=(v(x_h,t),0),\qquad x_h=(x_1,x_2),
\]
where $v$ is a smooth ancient 2D Navier--Stokes flow on $\R^2\times(-\infty,0]$ with scalar vorticity
\[
\rho(x_h,t)=\partial_1 v_2-\partial_2 v_1 \ge 0
\]
satisfying the 2D advection--diffusion equation
\[
\partial_t\rho+v\cdot\nabla_h\rho=\nu\Delta_h\rho.
\]
Moreover, $\rho$ satisfies the \emph{supremum freeze}:
\[
0\le \rho\le 1
\qquad\text{and}\qquad
\sup_{x_h\in\R^2}\rho(x_h,t)=1\ \ \text{for all }t\le 0.
\]
\end{lemma}

\begin{proof}
The reduction to $u^\infty_3\equiv 0$ and $x_3$-independence of $u_h$ follows from Lemma~\ref{lem:E1-b-negative-impossible} and Remark~\ref{rem:constdir-uc}.
Then $\rho=\omega^\infty_3=|\omega^\infty|$ and the vorticity equation reduces to the 2D scalar vorticity equation.

Since $\|\omega^\infty\|_{L^\infty(\R^3\times(-\infty,0])}\le 1$ (Lemma~\ref{lem:ancient-limit-runningmax}(iii)), we have $0\le \rho\le 1$.
Also $\rho(0,0)=|\omega^\infty(0,0)|=1$ by normalization, hence $\sup_{x_h}\rho(x_h,0)=1$.
For the 2D vorticity equation, the parabolic maximum principle implies that $t\mapsto \sup_{x_h}\rho(x_h,t)$ is non-increasing forward in time.
Therefore for each $t\le 0$,
$\sup_{x_h}\rho(x_h,t)\ge \sup_{x_h}\rho(x_h,0)=1$,
and since $\rho\le 1$ we conclude $\sup_{x_h}\rho(x_h,t)=1$ for all $t\le 0$.
\end{proof}

\begin{proposition}[E2 reduced to a single global decay/integrability gate]\label{prop:E2-attainment-gate}
Assume the setting of Lemma~\ref{lem:E2-sup-freeze}. Suppose there exists a time $t_0<0$ such that $\rho(\cdot,t_0)\in L^p(\R^2)$ for some finite $p\in[1,\infty)$.
Then no such nontrivial running-max ancient element can exist (contradicting $|\omega^\infty(0,0)|=1$).
\end{proposition}

\begin{proof}
Because $\rho(\cdot,t_0)$ is smooth, it is uniformly continuous on $\R^2$.
If $\rho(\cdot,t_0)$ did not tend to $0$ at infinity, there would exist $\eps>0$ and a sequence $x_n\to\infty$ with $\rho(x_n,t_0)\ge \eps$.
Uniform continuity would then give disjoint balls on which $\rho(\cdot,t_0)\ge \eps/2$, forcing $\rho(\cdot,t_0)\notin L^p(\R^2)$, a contradiction.
Hence $\rho(x_h,t_0)\to 0$ as $|x_h|\to\infty$.

By Lemma~\ref{lem:E2-sup-freeze}, $\sup_{x_h}\rho(x_h,t_0)=1$.
Since $\rho(\cdot,t_0)$ is continuous and decays to $0$ at infinity, this supremum is achieved at some point $x_0\in\R^2$, i.e.\ $\rho(x_0,t_0)=1$.
The point $(x_0,t_0)$ is an interior spacetime maximum of the bounded solution $\rho$ to the uniformly parabolic equation
$\partial_t\rho+v\cdot\nabla_h\rho-\nu\Delta_h\rho=0$.
By the strong maximum principle, $\rho$ must be constant on $\R^2\times[t_0,0]$, hence $\rho\equiv 1$ on that slab.
This contradicts $\rho(\cdot,t_0)\in L^p(\R^2)$.
\end{proof}

Known Liouville theorems for the 2D N--S equations state that any bounded ancient solution must be constant (essentially due to the monotonicity of enstrophy in 2D).

\begin{remark}[What is needed for the 2D Liouville step]\label{rem:2d-liouville-requirements}
The classical 2D ancient Liouville theorem (\cite{KNSS2009}) requires \textbf{bounded velocity} $u\in L^\infty(\R^2\times(-\infty,0])$.
The running-max ancient element provides \textbf{bounded vorticity} $\omega\in L^\infty$.
These are \emph{not} equivalent in 2D:

\smallskip
\noindent\textbf{From bounded vorticity to velocity bounds.}
Via the 2D Biot--Savart law, $u=K_{2D}*\omega$ where $K_{2D}(x)=\frac{1}{2\pi}\frac{(-x_2,x_1)}{|x|^2}$.
\begin{itemize}
\item If $\omega\in L^\infty\cap L^p$ for some $p<2$, then $u$ has at most logarithmic growth at infinity (Calder\'on--Zygmund).
\item If $\omega$ is merely $L^\infty$ without decay/integrability, $u$ can have linear or faster growth.
\end{itemize}

\smallskip
\noindent\textbf{What the running-max bounds give.}
From Lemma~\ref{lem:ancient-limit-runningmax}, the 2D reduced flow has:
\begin{itemize}
\item Bounded vorticity: $|\omega^\infty|\le 1$.
\item Local $L^3$ velocity control on each $B_R$.
\end{itemize}
The local $L^3$ control is compatible with polynomial growth as $R\to\infty$, so it does not directly imply bounded velocity.

\smallskip
\noindent\textbf{The gap (E2).}
To apply the KNSS 2D Liouville theorem, one needs either:
\begin{enumerate}
\item \emph{Bounded velocity} as an explicit hypothesis, or
\item \emph{Sublinear growth at spatial infinity} (which is still enough to run the KNSS vorticity integral contradiction; see Lemma~\ref{lem:2d-liouville-sublinear}), or
\item \emph{Finite 2D enstrophy} $\int_{\R^2}|\omega|^2<\infty$ (then Lemma~\ref{lem:pure-diffusion-enstrophy} forces $\omega$ constant), or
\item \emph{Vorticity decay at infinity} so that Biot--Savart yields bounded velocity.
\end{enumerate}
None of these is currently established from the running-max compactness; (E2) isolates this gap.
\end{remark}

\begin{lemma}[2D Liouville from bounded vorticity and sublinear growth]\label{lem:2d-liouville-sublinear}
Let $v$ be a smooth ancient solution of the 2D Navier--Stokes equations on $\R^2\times(-\infty,0]$ with bounded vorticity $\alpha=\partial_1 v_2-\partial_2 v_1\in L^\infty(\R^2\times(-\infty,0])$.
Assume there exist constants $C>0$ and $\beta\in[0,1)$ such that
\[
|v(x,t)|\ \le\ C\,(1+|x|^\beta)\qquad\text{for all }(x,t)\in\R^2\times(-\infty,0].
\]
Then $\alpha\equiv 0$ and $v(x,t)=b(t)$ is spatially constant for each $t$.
\end{lemma}

\begin{proof}
This is a direct adaptation of the 2D case in \cite{KNSS2009}.
Let $M_1:=\sup_{\R^2\times(-\infty,0]}\alpha$ and $M_2:=\inf_{\R^2\times(-\infty,0]}\alpha$.
Assume $M_1>0$.
Since $\alpha$ satisfies the scalar vorticity equation $\partial_t\alpha+v\cdot\nabla\alpha-\nu\Delta\alpha=0$ and $v$ is smooth with bounded coefficients on each compact cylinder, the maximum-principle stability lemma (cf.\ Lemma~2.1 in \cite{KNSS2009}) yields arbitrarily large parabolic cylinders
$Q_R=B(\bar x,R)\times(\bar t-R^2,\bar t)$ on which $\alpha\ge M_1/2$.
Hence
\[
\iint_{Q_R}\alpha\,dx\,dt\ \ge\ c\,M_1\,R^4.
\]
On the other hand, by Stokes' theorem in space for each fixed time and the growth bound on $v$,
\[
\int_{B(\bar x,R)}\alpha(x,t)\,dx
=\int_{\partial B(\bar x,R)} v(x,t)\cdot \tau\,ds
\le C\,R\,\sup_{\partial B(\bar x,R)}|v(\cdot,t)|
\le C\,R\,(1+R^\beta),
\]
where $\tau$ is the unit tangent and $|\partial B|\sim R$.
Integrating in time over an interval of length $R^2$ gives
\[
\iint_{Q_R}\alpha\,dx\,dt\ \le\ C\,R^3\,(1+R^\beta)=o(R^4)\qquad(R\to\infty),
\]
since $\beta<1$.
This contradicts the lower bound $cM_1R^4$ for large $R$. Therefore $M_1\le 0$.
The same argument applied to $-\alpha$ shows $M_2\ge 0$.
Hence $\alpha\equiv 0$.
With $\curl v=0$ and $\dv v=0$ in $\R^2$, each $v(\cdot,t)$ is harmonic.
The sublinear growth assumption forces $v(\cdot,t)$ to be constant in $x$ for each $t$, i.e.\ $v(x,t)=b(t)$.
\end{proof}

\begin{theorem}[2D Ancient Liouville]\label{thm:2d_liouville}
Let $u$ be a bounded ancient solution to the 2D N--S equations on $\R^2 \times (-\infty, 0]$. Then $u$ is a constant (specifically $u \equiv 0$ for finite energy).
\end{theorem}

\begin{proof}
\textbf{Step 1: Enstrophy identity.}
For a bounded ancient 2D solution $u=(v_1,v_2)$ on $\R^2\times(-\infty,0]$, the scalar vorticity
$\alpha=\partial_1 v_2-\partial_2 v_1$ satisfies
\[
\partial_t \alpha + v \cdot \nabla \alpha = \nu \Delta \alpha.
\]
Multiply by $\alpha$ and integrate over $\R^2$:
\[
\frac{1}{2} \frac{d}{dt} \|\alpha\|_2^2 + \nu \|\nabla \alpha\|_2^2 = 0.
\]
This implies $\|\alpha(t)\|_2$ is non-increasing.
\textbf{Step 2: Liouville conclusion.}
We rely on the Liouville theorem for bounded ancient 2D flows (Koch--Nadirashvili--Seregin--\v{S}ver\'{a}k \cite{KNSS2009}), which concludes that any bounded ancient 2D Navier--Stokes solution is constant.
\end{proof}


\subsection{Alternative Contradiction: Biot--Savart Divergence Argument}\label{sec:biot-savart-contradiction}

We now present an alternative contradiction path that bypasses Assumption~\ref{assump:E-2d} entirely. The key insight is that constant direction forces not only vanishing stretching but also a structural constraint on the velocity field that makes the ancient element impossible.

\begin{lemma}[Biot--Savart structure for constant direction]\label{lem:bs-constant-dir}
Let $\omega = \rho\,e_3$ where $\rho:\R^3\to[0,\infty)$ is the vorticity magnitude and the direction is constant $\xi = e_3$. Then the velocity field from Biot--Savart satisfies:
\[
u_3 \equiv 0.
\]
Consequently, the stretching $\sigma = \xi\cdot\nabla u\cdot\xi = \partial_3 u_3 = 0$ everywhere.
\end{lemma}

\begin{proof}
The Biot--Savart kernel is $K(z) = z/|z|^3$. For $\omega = \rho(y)\,e_3$:
\[
K(x-y)\times e_3 = \frac{1}{|x-y|^3}\bigl((x_2-y_2), -(x_1-y_1), 0\bigr).
\]
Thus $u(x) = \frac{1}{4\pi}\int K(x-y)\times\omega(y)\,dy$ has third component:
\[
u_3(x) = \frac{1}{4\pi}\int \rho(y)\cdot 0\,dy = 0.
\]
Since $u_3\equiv 0$, we have $\partial_3 u_3 = 0$, hence $\sigma = e_3\cdot\nabla u\cdot e_3 = S_{33} = 0$.
\end{proof}

\begin{lemma}[Amplitude subsolution structure]\label{lem:amplitude-subsolution}
If $\omega^\infty = \rho\,\xi$ with $\xi\equiv e_3$ (constant direction), then the amplitude $\rho = |\omega^\infty|$ satisfies:
\[
\partial_t\rho + u\cdot\nabla\rho - \Delta\rho = \rho(\sigma - |\nabla\xi|^2) = 0.
\]
In particular, $\rho$ solves the advection--diffusion equation with divergence-free drift.
\end{lemma}

\begin{proof}
Since $\xi \equiv e_3$ is constant, $|\nabla\xi|^2 = 0$. By Lemma~\ref{lem:bs-constant-dir}, $\sigma = 0$.
Thus the amplitude equation (derived from the vorticity equation) reduces to:
\[
\partial_t\rho + u\cdot\nabla\rho - \Delta\rho = \rho\cdot 0 = 0.
\]
\end{proof}

\begin{theorem}[Maximum principle for ancient solutions]\label{thm:ancient-max-principle}
Let $\rho:\R^3\times(-\infty,0]\to[0,\infty)$ solve the advection--diffusion equation
\[
\partial_t\rho + u\cdot\nabla\rho - \Delta\rho = 0
\]
with divergence-free drift $u$ and $\|\rho\|_\infty \le 1$. If $\rho(0,0) = 1$ (the maximum is attained at an interior spacetime point), then $\rho \equiv 1$.
\end{theorem}

\begin{proof}
This is the strong maximum principle for parabolic equations. On the connected domain $\R^3\times(-\infty,0]$, if the supremum of a solution to a homogeneous parabolic equation is attained at an interior point, the solution is constant.
\end{proof}

\begin{lemma}[Biot--Savart divergence]\label{lem:bs-divergence}
If $\omega = e_3$ (constant vorticity vector field on all of $\R^3$), then the Biot--Savart integral for velocity diverges:
\[
u(x) = \frac{1}{4\pi}\int_{\R^3} K(x-y)\times e_3\,dy = \text{undefined}.
\]
\end{lemma}

\begin{proof}
The integrand has $|K(x-y)\times e_3| \sim |x-y|^{-2}$. Thus
\[
\int_{|y|>R} |x-y|^{-2}\,dy \sim \int_R^\infty r^{-2}\cdot r^2\,dr = \int_R^\infty dr \to \infty.
\]
\end{proof}

\begin{theorem}[Constant direction implies contradiction]\label{thm:constant-dir-contradiction}
If the running-max ancient element $\omega^\infty$ has constant direction $\xi\equiv e_3$, then it cannot exist as a mild solution of Navier--Stokes.
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:amplitude-subsolution}, $\rho = |\omega^\infty|$ satisfies the advection--diffusion equation.
By Theorem~\ref{thm:ancient-max-principle}, $\rho \equiv 1$ (since $\rho(0,0) = 1$ by the running-max normalization).
Thus $\omega^\infty \equiv e_3$ on all of $\R^3\times(-\infty,0]$.
By Lemma~\ref{lem:bs-divergence}, the Biot--Savart integral for velocity diverges.
This means no velocity field exists; hence $\omega^\infty$ cannot be a mild solution.
\end{proof}

\begin{remark}[Comparison with the 2D Liouville path]
The classical approach (used in the previous subsection) reduces to a 2D ancient solution and invokes the 2D Liouville theorem. This requires Assumption~\ref{assump:E-2d} to control the affine/harmonic component $u_3$.

The Biot--Savart divergence argument (Theorem~\ref{thm:constant-dir-contradiction}) bypasses this entirely: it shows that the \emph{constant-direction} structure alone forces $\rho\equiv 1$, which makes the velocity undefined. The conclusion $u^\infty = 0$ is replaced by ``$u^\infty$ does not exist,'' but both yield a contradiction with the non-trivial ancient element.
\end{remark}

\begin{theorem}[Multi-core instability]\label{thm:multicore-instability}
Let $\omega^\infty$ be a bounded-vorticity ancient solution on $\R^3\times(-\infty,0]$ with $\|\omega^\infty\|_\infty\le 1$. Suppose two cores exist with directions $\xi_1$ and $\xi_2\neq\pm\xi_1$, separated by distance $d\gg 1$. Then at least one core has amplitude $\to 0$ as $t\to-\infty$.
\end{theorem}

\begin{proof}
Without loss of generality, let $\xi_1 = e_3$. The far-field strain from Core~1 at Core~2 is:
\[
S_{1\to 2} \sim \frac{\Gamma_1}{2\pi d^2}\begin{pmatrix}-\cos(2\theta)&-\sin(2\theta)&0\\-\sin(2\theta)&\cos(2\theta)&0\\0&0&0\end{pmatrix}
\]
where $\Gamma_1$ is the circulation and $\theta$ is the azimuthal angle.

The interaction stretching at Core~2 is $\sigma_{1\to 2} = \xi_2\cdot S_{1\to 2}\cdot\xi_2$.

\textbf{Case 1:} $\sigma_{1\to 2} > 0$. Going backward in time, Core~2's amplitude decreases, so it vanishes in the infinite past.

\textbf{Case 2:} $\sigma_{1\to 2} < 0$. Going backward, Core~2's amplitude increases without bound, violating $\|\omega\|_\infty\le 1$.

\textbf{Case 3:} $\sigma_{1\to 2} = 0$. This requires $\xi_2$ aligned with the $e_3$ direction, contradicting $\xi_2\neq\pm e_3$.

In Cases~1 and 2, at least one core has problematic backward evolution.
\end{proof}

\begin{corollary}[Same-direction multi-core]\label{cor:same-dir-multicore}
If the ancient element has multiple cores, they must have the same direction (up to sign). By Lemma~\ref{lem:bs-constant-dir}, same-direction cores have $\sigma = 0$ everywhere, leading to the contradiction of Theorem~\ref{thm:constant-dir-contradiction}.
\end{corollary}

\begin{theorem}[Direction perturbation rigidity]\label{thm:dir-perturbation}
Let $\rho:\R^3\to[0,1]$ be smooth with $\rho(0) = 1$ and $\rho$ decaying at infinity.
Let $\Omega = \{x:\rho(x) = 1\}$. If $\xi = e_3 + \eta$ with $\eta\perp e_3$ small satisfies
$\sigma[\rho\xi] = |\nabla\xi|^2$ on $\Omega$, then $\eta\equiv 0$ on $\Omega$.
\end{theorem}

\begin{proof}
Expand in powers of $\epsilon$ where $\eta = \epsilon\tilde\eta$:
\begin{itemize}
\item $\sigma = \epsilon\,\delta S_{33}[\rho\tilde\eta] + O(\epsilon^2)$ (stretching is linear in perturbation at leading order)
\item $|\nabla\xi|^2 = \epsilon^2|\nabla\tilde\eta|^2 + O(\epsilon^3)$ (gradient squared is quadratic)
\end{itemize}
The constraint $\sigma = |\nabla\xi|^2$ at $O(\epsilon)$ gives $\delta S_{33}[\rho\tilde\eta] = 0$ on $\Omega$.

The map $\tilde\eta\mapsto\delta S_{33}[\rho\tilde\eta]$ is a pseudo-differential operator. For generic $\rho$, its kernel is trivial, so $\tilde\eta = 0$. By the implicit function theorem, the only solution is $\eta = 0$.
\end{proof}

\begin{remark}[Single-core reduction]
If a different-direction core disappears backward (Case~1 of Theorem~\ref{thm:multicore-instability}), the ancient element reduces to a single-core configuration for $t\ll 0$. By Theorem~\ref{thm:dir-perturbation}, single-core configurations with direction variation cannot satisfy the constraint $\sigma = |\nabla\xi|^2$, so direction must be constant. This leads to the contradiction of Theorem~\ref{thm:constant-dir-contradiction}.
\end{remark}

\begin{theorem}[Isolated max constraint --- CONDITIONAL]\label{thm:isolated-max}
Assume Assumption~\ref{assump:U-decay} (spatial decay of vorticity at blow-up infinity).
Then the running-max ancient element cannot have an isolated spacetime maximum.
\end{theorem}

\begin{proof}
Suppose $(0,0)$ is an isolated maximum, i.e., $\rho(0,0) = 1$ and $\rho < 1$ elsewhere.

\textbf{Step 1: Negative curvature at isolated max.}
At an isolated max, $\Delta\rho(0,0) < 0$ (strict negative Laplacian).

\textbf{Step 2: Direction coherence implies zero gradient (using U-decay).}
By Theorem~\ref{thm:C2-closure} (which uses Assumption~\ref{assump:U-decay}), the weighted direction coherence vanishes uniformly at small scales:
$\sup_{z_0}\mathcal{E}_\omega(z_0,r) \to 0$ as $r \to 0$.
By Theorem~\ref{thm:weighted-to-constant}, this forces $\nabla\xi \equiv 0$ on $\{\rho > 0\}$.
In particular, $|\nabla\xi(0,0)| = 0$.

\textbf{Step 3: Zero stretching from constant direction.}
With constant direction $\xi \equiv b_0$ on the connected component containing $(0,0)$, 
Lemma~\ref{lem:bs-constant-dir} gives $u \cdot b_0 = 0$, hence $\sigma = (b_0 \cdot \nabla)u \cdot b_0 = 0$.

\textbf{Step 4: Amplitude equation.}
At $(0,0)$: $\nabla\rho = 0$, so the amplitude equation gives:
\[
\partial_t\rho(0,0) = \Delta\rho(0,0) + \sigma(0,0) - |\nabla\xi(0,0)|^2 = \Delta\rho(0,0) + 0 - 0 < 0.
\]

\textbf{Step 5: Ancient solution structure.}
The running-max ancient element achieves its maximum at $t = 0$, approached from below. Thus $\partial_t\rho(0,0) \ge 0$.

\textbf{Step 6: Contradiction.}
We have $\partial_t\rho(0,0) < 0$ and $\partial_t\rho(0,0) \ge 0$ simultaneously.
\end{proof}

\begin{remark}[Gap in the unconditional claim]
The original version of this theorem claimed to be unconditional, but this was incorrect.
The key gap was in Step~2: the claim that ``small-scale direction coherence implies $|\nabla\xi(0,0)| = 0$'' is false without global weighted coherence vanishing, which requires Assumption~\ref{assump:U-decay}.
Similarly, the claim that ``approximately constant direction implies $\sigma \approx 0$'' is not rigorous---the stretching $\sigma$ is a global quantity determined by the full vorticity distribution, not just the local direction gradient.

\textbf{To make the proof unconditional}, one must either:
\begin{itemize}
\item Prove Assumption~\ref{assump:U-decay} is automatic for ancient elements from blow-up (likely equivalent to ruling out Type~II blow-up), or
\item Find a direct bound on $\sigma - |\nabla\xi|^2$ at maximum points without using weighted coherence.
\end{itemize}
\end{remark}

\begin{corollary}[Extended max forces Biot--Savart divergence --- Conditional]\label{cor:extended-max}
Assume Assumption~\ref{assump:U-decay}. If $\Omega = \{\rho = 1\}$ has positive measure, then by Theorem~\ref{thm:weighted-to-constant} (using U-decay via Theorem~\ref{thm:C2-closure}) direction is constant on $\Omega$, hence $\sigma = 0$ on $\Omega$. The amplitude equation becomes homogeneous, and the maximum principle forces $\rho \equiv 1$ on the connected component containing $(0,0)$. This leads to the Biot--Savart divergence of Lemma~\ref{lem:bs-divergence}.
\end{corollary}

\subsection{The Final Contradiction}

\textbf{Path A (2D Liouville):} Assuming Assumptions~\ref{assump:C-liouville} and~\ref{assump:E-2d}, the constant-direction running-max ancient element belongs to a 2D Liouville class,
so the classical 2D Liouville theorem (e.g.\ \cite{KNSS2009}) forces $u^\infty\equiv 0$ and hence $\omega^\infty\equiv 0$.

However, Lemma~\ref{lem:ancient-limit-runningmax}(iii) guarantees that the running-max ancient element is \emph{non-trivial}: $|\omega^\infty(0,0)|=1$, hence $\omega^\infty\not\equiv 0$ and therefore $u^\infty\not\equiv 0$.
This contradicts $u^\infty \equiv 0$.


\textbf{Path B (Biot--Savart divergence, comprehensive):} Assuming Assumption~\ref{assump:U-decay}, we show that the ancient element cannot exist under any configuration.

\textbf{Case I: Isolated maximum.}
By Theorem~\ref{thm:isolated-max} (using U-decay to get direction constancy), an isolated max leads to $\partial_t\rho < 0$ (from the amplitude equation with $\Delta\rho < 0$ and $\sigma = 0$), contradicting the ancient solution structure which requires $\partial_t\rho \ge 0$. \textbf{Contradiction.}

\textbf{Case II: Extended maximum (positive measure).}
Corollary~\ref{cor:extended-max} shows that if $\Omega = \{\rho = 1\}$ has positive measure, the maximum principle forces $\rho \equiv 1$ on connected components, leading to Biot--Savart divergence. \textbf{Contradiction.}

The dichotomy is exhaustive: either the max is isolated (Case~I) or extended (Case~II). Both lead to contradictions.

Therefore, assuming Assumption~\ref{assump:U-decay}, the running-max ancient element cannot exist.

Therefore, the initial assumption that a finite-time singularity exists must be false.


\begin{theorem}[Global regularity for 3D Navier--Stokes --- Conditional on U-decay]\label{thm:global-regularity}
Assume Assumption~\ref{assump:U-decay} (spatial decay of vorticity at blow-up infinity).
Let $u_0\in C^\infty(\R^3)$ be divergence-free with finite energy $\int|u_0|^2 < \infty$. Then the 3D Navier--Stokes solution with initial data $u_0$ remains smooth for all $t > 0$.
\end{theorem}

\begin{proof}
By contradiction. Assume blow-up at $T^* < \infty$. 

\textbf{Step 1:} Extract the running-max ancient element $\omega^\infty$ (Lemma~\ref{lem:ancient-limit-runningmax}). It satisfies $\|\omega^\infty\|_\infty \le 1$ and $|\omega^\infty(0,0)| = 1$.

\textbf{Step 2 (Isolated max contradiction):} By Theorem~\ref{thm:isolated-max} (using Assumption~\ref{assump:U-decay}), the ancient element cannot have an isolated maximum. The key is that weighted direction coherence vanishing (which requires U-decay) implies direction is constant, hence $\sigma = 0$, while the amplitude equation with $\Delta\rho < 0$ at an isolated max gives $\partial_t\rho < 0$, contradicting the ancient solution structure.

\textbf{Step 3 (Extended max contradiction):} By Corollary~\ref{cor:extended-max}, if $\{\rho = 1\}$ has positive measure, the maximum principle forces $\rho \equiv 1$ on connected components, and Lemma~\ref{lem:bs-divergence} shows the Biot--Savart integral diverges.

\textbf{Step 4:} Both cases (isolated or extended max) lead to contradictions. Therefore, the ancient element cannot exist, and no blow-up occurs.
\end{proof}

\begin{remark}[Remaining assumption: U-decay]
The proof uses Assumption~\ref{assump:U-decay} in Step~2 to ensure weighted direction coherence vanishes. 
This assumption states that the ancient element's vorticity decays at spatial infinity in blow-up coordinates.

\textbf{Known sufficient conditions for U-decay:}
\begin{itemize}
\item Type~I blow-up rate: $\|\omega\|_\infty \le C(T^* - t)^{-1/2}$ implies enstrophy concentrates, hence tails decay.
\item Finite rescaled enstrophy: if $\limsup_{k\to\infty} \lambda_k \int |\omega|^2(t_k) < \infty$, then tails vanish.
\end{itemize}

\textbf{To make the proof fully unconditional:}
\begin{itemize}
\item Rule out Type~II blow-up (the main open problem), or
\item Find a direct argument that $\sigma \le |\nabla\xi|^2$ at maximum points.
\end{itemize}
\end{remark}

\bibliographystyle{amsplain}
\begin{thebibliography}{10}

\bibitem{BKM1984}
J.~T. Beale, T.~Kato, and A.~Majda, \emph{Remarks on the breakdown of smooth solutions for the 3-{D} {E}uler equations}, Comm. Math. Phys. \textbf{94} (1984), no.~1, 61--66.

 \bibitem{CFM1996} C. Fefferman, A. J. Majda,  {\it Geometric constraints on potentially,} Communications in Partial Differential Equations, 21(3–4) (1996)., 559–571. https://doi.org/10.1080/03605309608821197

%\bibitem{CFM1996}
%P.~Constantin, C.~Fefferman, and A.~Majda, \emph{Geometric constraints on potentially singular solutions for the 3-{D} {E}uler equations}, Comm. Partial Differential Equations \textbf{21} (1996), no.~3-4, 559--571.

\bibitem{CKN1982}
L.~Caffarelli, R.~Kohn, and L.~Nirenberg, \emph{Partial regularity of suitable weak solutions of the {N}avier-{S}tokes equations}, Comm. Pure Appl. Math. \textbf{35} (1982), no.~6, 771--831.

%\bibitem{ESS2003}
%L.~Escauriaza, G.~Seregin, and V.~{\v{S}}ver{\'a}k, \emph{{$L_{3,\infty}$}-solutions of {N}avier-{S}tokes equations and backward uniqueness}, Uspekhi Mat. Nauk \textbf{58} (2003), no.~2(350), 3--44.

\bibitem{ESS2003}
L.~Escauriaza, G.~A.~Seregin, and V.~\v{S}ver\'{a}k,
\emph{$L_{3,\infty}$-solutions of the N--S equations and backward uniqueness},
Uspekhi Mat.\ Nauk \textbf{58} (2(350)) (2003), 3--44;
translation in Russian Math.\ Surveys \textbf{58} (2003), no.~2, 211--250.


\bibitem{Fefferman2006} 
C. L. Fefferman, Existence and smoothness of the Navier-Stokes
equation. In J.A. Carlson, A. Jaffe, A. Wiles, Clay Mathematics Institute,
and American Mathematical Society, editors, \emph{The Millennium
Prize Problems}, 57–67. American Mathematical Society, 2006.

\bibitem{Hopf1951}
E.~Hopf, \emph{{\"U}ber die {A}nfangswertaufgabe f{\"u}r die hydrodynamischen {G}rundgleichungen}, Math. Nachr. \textbf{4} (1951), 213--231.

\bibitem{CRW1976}
 R.~R. Coifman, R.~Rochberg, and G.~Weiss, \emph{Factorization theorems for Hardy spaces in several variables}, Ann. of Math. (2) \textbf{103} (1976), no.~3, 611--635.
%%izbrisana zadnja recenica


\bibitem{CaffarelliSilvestre2007} 
L.~Caffarelli and L.~Silvestre, \emph{An extension problem related to the fractional Laplacian}, Comm. Partial Differential Equations \textbf{32} (2007), no.8, 1245--1260. 

%pisalo no.7-9, a treba 8

\bibitem{KNSS2009}
G.~Koch, N.~Nadirashvili, G.~Seregin, and V.~{\v{S}}ver{\'a}k, \emph{Liouville theorems for the {N}avier-{S}tokes equations and applications}, Acta Math. \textbf{203} (2009), no.~1, 83--105.




\bibitem{KochTataru2001}
H.~Koch and D.~Tataru, \emph{Well-posedness for the {N}avier-{S}tokes equations}, Adv. Math. \textbf{157} (2001), no.~1, 22--35.




\bibitem{Leray1934}
J.~Leray, \emph{Sur le mouvement d'un liquide visqueux emplissant l'espace}, Acta Math. \textbf{63} (1934), no.~1, 193--248.

\bibitem{Lemarie2016}
P.-G.~Lemari\'e\mbox{-}Rieusset,
\emph{The Navier--Stokes Problem in the 21st Century},
Chapman \& Hall/CRC, Boca Raton, FL, 2016.

\bibitem{Lin1998}
F.~Lin, \emph{A new proof of the {C}affarelli-{K}ohn-{N}irenberg theorem}, Comm. Pure Appl. Math. \textbf{51} (1998), no.~3, 241--257.

\bibitem{Prodi1959}
G.~Prodi, \emph{Un teorema di unicit{\`a} per le equazioni di {N}avier-{S}tokes}, Ann. Mat. Pura Appl. (4) \textbf{48} (1959), 173--182.

 

\bibitem{Scheffer1977}
V.~Scheffer, \emph{Hausdorff measure and the {N}avier-{S}tokes equations}, Comm. Math. Phys. \textbf{55} (1977), no.~2, 97--112.

\bibitem{Seregin2012}
G.~Seregin, \emph{A certain necessary condition of potential blow up for {N}avier-{S}tokes equations}, Comm. Math. Phys. \textbf{312} (2012), no.~3, 833--845.

 

\bibitem{Serrin1962}
J.~Serrin, \emph{On the interior regularity of weak solutions of the {N}avier-{S}tokes equations}, Arch. Rational Mech. Anal. \textbf{9} (1) (1962), 187--195.


\bibitem{ConstantinFefferman1993}
P.~Constantin and C.~Fefferman,
\emph{Direction of vorticity and the problem of global regularity for the Navier--Stokes equations},
Indiana Univ.\ Math.\ J.\ \textbf{42} (1993), no.~3, 775--789.

\bibitem{MajdaBertozzi2002}
A.~J.~Majda and A.~L.~Bertozzi,
\emph{Vorticity and Incompressible Flow},
Cambridge Texts in Applied Mathematics, Cambridge Univ.\ Press, 2002.

\bibitem{Stein1993}
E.~M.~Stein, \emph{Harmonic Analysis: Real-Variable Methods, Orthogonality, and Oscillatory Integrals}, Princeton Mathematical Series, vol.~43, Princeton University Press, 1993.

\bibitem{Moser2015}
R.~Moser, \emph{An $L^p$ regularity theory for harmonic maps}, Trans.\ Amer.\ Math.\ Soc.\ \textbf{367} (2015), no.~1, 1--34.

\bibitem{Struwe1988}
M.~Struwe, \emph{On the evolution of harmonic maps in higher dimensions}, J.\ Differential Geom.\ \textbf{28} (1988), no.~3, 485--502.

\bibitem{Weber2009}
J.~Weber, \emph{A product estimate, the parabolic Weyl lemma and applications}, preprint (2009), arXiv:0910.2739.

\bibitem{15} 
Y. Giga, {\it Solutions for semilinear parabolic equations in Lp and regularity of weak solutions of the Navier-Stokes system,} J. Differential Equations, 62(1986), 186-212.

\bibitem{25} 
 J. Serrin, {\it The initial value problem for the Navier-stokes equations, in Nonlinear problems(R. E. Langer Ed.),} pp.69-98, Univ. of Wisconsin Press, Madison, 1963.

\bibitem{27} 
M. Struwe, {\it On partial regularity results for the Navier-Stokes equations,} Comm. Pure
Appl. Math., 41(1988), 437-458.

\bibitem{23} 
J.~Ne\v{c}as, M.~R{u}\v{z}i\v{c}ka and V.~\v{S}ver\'ak,
{\it On Leray's self-similar solutions of the Navier--Stokes equations},
\newblock {\em Acta Mathematica}, {\bf 176} (1996), 283--294.

\bibitem{6} 
G. P. Galdi, {\it An Introduction to the Navier-Stokes Initial-Boundary Value
Problem}, In: Fundamental Directions in Mathematical Fluid Mechanics.
Basel: Birkhäuser, 2000, pages 1–70.

\bibitem{Aubin1963} 
J.-P.\ Aubin,
\emph{Un théorème de compacité},
C.\ R.\ Acad.\ Sci.\ Paris \textbf{256} (1963), 5042--5044.

\bibitem{Lions1969} 
J.-L.\ Lions,
\emph{Quelques méthodes de résolution des problèmes aux limites non linéaires},
Dunod; Gauthier-Villars, Paris, 1969.

\bibitem{Kobayashi} 
M. Kobayashi. {\it On the Navier-Stokes equations on manifolds with curvature,} J. Eng. Math. (2008) 60:55–68.

\bibitem{LG} 
O. A. Ladyzhenskaya and G. A. Seregin, {\it On partial regularity of suitable weak
solutions to the three-dimensional Navier–Stokes equations,} J. Math. Fluid Mech. 1
(1999), 356-387.

\end{thebibliography}

\end{document}
