\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{hyperref}

\title {Step 2 - referee\_package\_2026-02-03}
\author{}
\date{2026-02-03}

\begin{document}
\maketitle

\section*{Follow up with Author}
 in the uploaded
\texttt{referee\_package\_2026-02-03.zip}. It is \emph{theory-only}: no scripts or JSON artifacts were executed here.

\section*{What appears resolved (green responses)}
\begin{itemize}[leftmargin=*]
\item \textbf{(AC$_\mu$) removal claim.} Multiple green blocks state that Theorem~\texttt{\textbackslash ref\{thm:Pplus\}}
is now proved \emph{directly} as a Lebesgue-a.e.\ wedge statement in Appendix~\texttt{\textbackslash ref\{app:pplus-proof\}},
so no $\mu$-a.e.\ upgrade and no (AC$_\mu$) hypothesis is used.
\item \textbf{Citations added.} Green blocks add citations for the $\det_2$ product formula and for standard half-plane
outer-function and Schur-transport facts (e.g.\ Garnett; Rosenblum--Rovnyak; Simon).
\item \textbf{Gauge/domain sanity.} Green block addresses the $(s-1)/s$ compensator and regularity at $s=1$ on $\Omega$.
\item \textbf{Blocker-mapping narrative.} The manuscript now contains an explicit map from the two analytic inputs needed for
Lemma~\texttt{\textbackslash ref\{lem:F-boundary-admissible\}} to concrete Appendix lemmas (determinant $L^1$ control,
outer stability, Carleson/Whitney control, etc.).
\end{itemize}

\section*{Still inconsistent statements (must fix for a clean ``unconditional'' story)}
\begin{enumerate}[leftmargin=*]
\item \textbf{Main Theorem statement still asserts (AC$_\mu$).}\\
In \texttt{paper1\_farfield.tex}, Theorem~\texttt{\textbackslash ref\{thm:farfield\}} still contains a blue sentence saying
the proof relies on an Appendix $\mu$-construction and the domination hypothesis (AC$_\mu$).
This directly contradicts the green claim that (AC$_\mu$) is no longer used.
\item \textbf{One paragraph says JSON artifacts are ``used in the proof''.}\\
In the ``Audit posture'' paragraph (near the repository contents), it says the JSON artifacts recording maxima are ``used in the proof''.
This contradicts the Abstract and the ``Conclusion and limitations'' section, which state artifacts are \emph{not} used in the proof.
\item \textbf{Obsolete conditionality blocks remain inline.}\\
Several blue blocks still describe a $\mu$-a.e.\ $\Rightarrow$ Lebesgue-a.e.\ upgrade as a required step, and cite (AC$_\mu$).
A green block says these are ``historical referee commentary'', but they remain in the main logical flow.
For a journal-ready unconditional draft, these should be removed or moved to a separate ``Referee history'' appendix.
\end{enumerate}

\section*{Remaining theory gaps that are \emph{not} just editorial}
\begin{enumerate}[leftmargin=*]
\item \textbf{Lemma \texttt{lem:F-boundary-admissible} is still phrased as an assumption.}\\
It begins with ``Assume the analytic inputs established in Appendix...'', then lists two inputs (boundary trace/log-modulus and a harmonic
majorant/box control). The manuscript \emph{does} provide an Appendix mapping to specific lemmas, but the main text should be updated
to a standard form:
\begin{itemize}
\item either: state the lemma unconditionally and give a short proof that explicitly cites the exact Appendix lemmas that discharge (i)--(ii), or
\item: promote those Appendix items to a named theorem and cite that theorem here.
\end{itemize}
Until this is rewritten, a referee can reasonably read the result as ``assumed'' rather than proved in-paper.

\item \textbf{The distributional phase--velocity lemma is only a ``proof sketch''.}\\
In \texttt{paper1\_pplus\_proof.tex}, Lemma~\texttt{lem:pv-distributional} has a proof labeled ``Proof sketch and what must be checked''.
This should be upgraded to either (a) a complete proof, or (b) a precise citation (theorem/lemma number) in a standard reference
(e.g.\ Duren or Garnett) with hypotheses matching exactly what is used later.

\item \textbf{Remove/neutralize unused $\mu$-infrastructure if it is truly unused.}\\
The Appendix still contains (AC$_\mu$) and $\mu$-to-Lebesgue discussion, including ``what must be checked for the present $\mu$''.
If the final (P+) proof chain no longer uses it, keep it only as an optional historical note clearly marked ``not used''; otherwise,
a referee may interpret it as part of the logical dependency chain.
\end{enumerate}

\section*{Minimal patch list to make the paper unambiguously unconditional}
\begin{itemize}[leftmargin=*]
\item Delete or rewrite the (AC$_\mu$) sentence inside Theorem~\texttt{\textbackslash ref\{thm:farfield\}}; replace with a single forward reference
to the Lebesgue-a.e.\ Appendix proof of Theorem~\texttt{\textbackslash ref\{thm:Pplus\}}.
\item In ``Audit posture'': change ``used in the proof'' to ``for independent cross-check / auditability only'' (or, if artifacts are load-bearing,
then remove the Abstract/Conclusion statements that say they are not used).
\item Rewrite Lemma~\texttt{\textbackslash ref\{lem:F-boundary-admissible\}} so it is no longer an ``Assume ...'' lemma: explicitly discharge
(i)--(ii) by citing the exact Appendix lemmas.
\item Upgrade Lemma~\texttt{lem:pv-distributional} from ``proof sketch'' to a complete proof or a precise theorem-number citation.
\item Move all obsolete blue conditionality blocks into a separate ``Referee log'' appendix (or delete for the submission version).
\end{itemize}

\end{document}
