\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{listings}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  urlcolor=blue!50!black
}

\title{Riemann Project Handoff: Mathlib Submission Strategy}
\author{Prepared for Jonathan Washburn}
\date{January 30, 2026}

\begin{document}
\maketitle

\section{Overview}
This document outlines the strategy for submitting pre-existing Lean formalization work to Mathlib, specifically targeting gaps in the current library and the PNT+ project. The goal is to bypass recent authorship disputes by submitting original, pre-October 2025 work from the \texttt{riemann-e55} repository.

\section{Gap Analysis}
A review of the current Mathlib4 repository (as of Jan 30, 2026) confirms the following gaps:

\begin{itemize}
  \item \textbf{Weierstrass Factorization}: No \texttt{WeierstrassFactor} or \texttt{CanonicalProduct} modules exist in \texttt{Mathlib/Analysis/Complex}.
  \item \textbf{Hadamard Factorization}: No \texttt{HadamardFactorization} module exists.
  \item \textbf{Binet/Stirling}: No \texttt{BinetFormula} or \texttt{BinetKernel} exists in \texttt{Mathlib/Analysis/SpecialFunctions/Gamma}.
\end{itemize}

\section{Candidate Files for Submission}
The following files from the \texttt{riemann-e55} repository (snapshot date: Oct 28, 2025) are identified as high-priority candidates. They are mathematically complete (no \texttt{sorry}) and fill the identified gaps.

\subsection{1. Weierstrass Product Infrastructure}
\textbf{Source:} \texttt{rh/academic\_framework/DiagonalFredholm/WeierstrassProduct.lean}\\
\textbf{Provenance:} Commit \texttt{5f5bcdff} (Oct 23, 2025)\\
\textbf{Content:}
\begin{itemize}
  \item Cubic tail bounds for $\log(1-z)$
  \item Bridges between summable series and infinite products
\end{itemize}
\textbf{Target:} \texttt{Mathlib/Analysis/Complex/WeierstrassProduct.lean}

\subsection{2. Fredholm Determinant (Euler Product)}
\textbf{Source:} \texttt{rh/academic\_framework/DiagonalFredholm/Determinant.lean}\\
\textbf{Provenance:} Commit \texttt{a34ae557} (Oct 24, 2025)\\
\textbf{Content:}
\begin{itemize}
  \item Definition of $\det_2(I-A)$ as an Euler product over primes
  \item Proof of non-vanishing on $\Re(s) > 1/2$
\end{itemize}
\textbf{Target:} \texttt{Mathlib/NumberTheory/LSeries/FredholmDeterminant.lean}

\subsection{3. Gamma Function Bounds}
\textbf{Source:} \texttt{rh/academic\_framework/GammaBounds.lean}\\
\textbf{Provenance:} Commit \texttt{733a43e0} (Sep 18, 2025)\\
\textbf{Content:}
\begin{itemize}
  \item Uniform bounds for the Archimedean factor on vertical strips
\end{itemize}
\textbf{Target:} \texttt{Mathlib/Analysis/SpecialFunctions/Gamma/Bounds.lean}

\section{Submission Strategy}
To avoid conflation with the "contested" work (Binet/Stirling PRs submitted in Jan 2026), we recommend:

\begin{enumerate}
  \item \textbf{Submit from Original Repo:} Push the files directly from the \texttt{riemann-e55} git history to a new branch, preserving the Oct 2025 commit dates.
  \item \textbf{Target PNT+ First:} Since Mathlib review is blocked/contentious, submit these as a dependency bundle to the PNT+ project (which has an empty \texttt{HadamardFactorization.lean}).
  \item \textbf{Clean History:} Do not mix this with the recent "refactored" code. Use the original file contents.
\end{enumerate}

\section{Next Steps for Colleague}
\begin{enumerate}
  \item Clone \texttt{https://github.com/jonwashburn/riemann-e55}.
  \item Checkout commit \texttt{56ddbc0} (Oct 28, 2025).
  \item Extract the three files listed above.
  \item Verify they compile against current Mathlib (using the \texttt{Common.lean} shim if needed).
  \item Open a PR to PNT+ or Mathlib with the message: "Contributing original pre-October 2025 Hadamard infrastructure."
\end{enumerate}

\end{document}
