% Dual-Route Invariant Equalization + Single-Inequality Audit — Family C Patent
\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{xcolor}

\title{Dual-Route Invariant Equalization and Single-Inequality Audit\\Family C Patent (Method, System, Medium)}
\author{Jonathan Washburn}
\date{\today}

\begin{document}
\maketitle

\begin{center}
{\large Camera-Ready Patent Draft}\\[4pt]
{\normalsize Computing $\lambda_{\mathrm{kin}}$, $\lambda_{\mathrm{rec}}$ with Combined Uncertainty; Single-Inequality Acceptance; Fail-Closed Signaling}
\end{center}

\section*{Field}
This disclosure relates to computer-implemented acceptance auditing of calibration invariants. It provides a method, system, and non-transitory medium for computing independent, dimensionless route invariants and their combined uncertainty (with correlation), applying a single-inequality acceptance test, and emitting fail-closed accept/reject signals and signed logs.

\section*{Background}
Calibration acceptance often relies on multi-threshold heuristics or post-hoc statistics that are difficult to audit and tune across devices. There is a need for a \emph{single}, interpretable inequality that (i) compares independent routes to the same invariant, (ii) accounts for uncertainty and correlation, (iii) yields a deterministic accept/reject outcome, and (iv) generates machine-verifiable evidence for audit and enforcement.

\section*{Summary}
Disclosed is an acceptance audit that:
\begin{enumerate}[leftmargin=*]
  \item computes independent route invariants $\lambda_{\mathrm{kin}}$ and $\lambda_{\mathrm{rec}}$ (dimensionless);
  \item computes combined uncertainty with correlation,
  \[
    u_{\mathrm{comb}}\;=\;\sqrt{\,u(\lambda_{\mathrm{kin}})^2 + u(\lambda_{\mathrm{rec}})^2 - 2\,\rho\,u(\lambda_{\mathrm{kin}})\,u(\lambda_{\mathrm{rec}})\,}\,;\]
  \item applies a single acceptance inequality,
  \[
    \frac{\,\lvert\lambda_{\mathrm{kin}}-\lambda_{\mathrm{rec}}\rvert\,}{\lambda_{\mathrm{rec}}}\;\le\;k\,u_{\mathrm{comb}}\,;
  \]
  \item emits fail-closed accept/reject signals and logs with signatures.
\end{enumerate}

\section*{Definitions}
\noindent\textbf{Route invariants.} $\lambda_{\mathrm{kin}}$ and $\lambda_{\mathrm{rec}}$ are independent, dimensionless derivations of the same physical invariant (e.g., kinematic vs recognition routes) computed from factorized displays.

\noindent\textbf{Uncertainty model.} $u(\cdot)$ denotes standard uncertainty; $\rho\in[-1,1]$ is the correlation coefficient between route estimates; $u_{\mathrm{comb}}$ is the combined uncertainty per the above expression.

\noindent\textbf{Acceptance threshold.} $k>0$ is a declared scalar (risk/tier dependent). The inequality compares normalized route discrepancy to combined uncertainty.

\noindent\textbf{Fail-closed signaling.} A control policy that denies acceptance if the inequality is not satisfied or if uncertainty/correlation cannot be established within declared bounds.

\noindent\textbf{Compliance artifacts.} Signed, hash-addressed logs that include route values, uncertainties, correlation estimate, threshold $k$, verdict, and metadata.

\section*{Claims}
\setlist[enumerate]{leftmargin=*,labelsep=0.5em,itemsep=0.6em,topsep=0.4em}

\subsection*{Independent Claims}
\begin{enumerate}
\item \textbf{(Method)} A computer-implemented method for acceptance auditing of calibration invariants, comprising:
\begin{enumerate}[label=(\alph*)]
  \item computing first and second dimensionless route invariants $\lambda_{\mathrm{kin}}$ and $\lambda_{\mathrm{rec}}$ for a common physical invariant;
  \item computing standard uncertainties $u(\lambda_{\mathrm{kin}})$, $u(\lambda_{\mathrm{rec}})$ and a correlation coefficient $\rho$;
  \item computing a combined uncertainty $u_{\mathrm{comb}}=\sqrt{u(\lambda_{\mathrm{kin}})^2 + u(\lambda_{\mathrm{rec}})^2 - 2\rho\,u(\lambda_{\mathrm{kin}})u(\lambda_{\mathrm{rec}})}$;
  \item testing an acceptance inequality $\lvert\lambda_{\mathrm{kin}}-\lambda_{\mathrm{rec}}\rvert/\lambda_{\mathrm{rec}} \le k\,u_{\mathrm{comb}}$ with a declared threshold $k>0$; and
  \item emitting a fail-closed accept/reject signal and a signed, hash-addressed log including route values, uncertainties, correlation, $k$, and the verdict.
\end{enumerate}

\item \textbf{(System)} A system comprising one or more processors and memory configured to: (i) compute route invariants $\lambda_{\mathrm{kin}}$, $\lambda_{\mathrm{rec}}$; (ii) compute uncertainties and correlation; (iii) compute $u_{\mathrm{comb}}$; (iv) test the acceptance inequality; and (v) emit fail-closed verdicts and signed logs.

\item \textbf{(Non-transitory medium)} A non-transitory computer-readable medium storing instructions that, when executed, cause a system to perform the method of claim 1.
\end{enumerate}

\subsection*{Dependent Claims}
\begin{enumerate}[resume]
\item (Method) wherein the uncertainties and correlation are estimated from windowed samples, bootstrapped replicates, or spectral cross-estimators with declared confidence bounds.
\item (Method) wherein $k$ is selected from a policy tier: safety-critical, regulated, or research, each with predefined numeric values.
\item (Method) wherein acceptance requires $u(\lambda_{\mathrm{kin}})$, $u(\lambda_{\mathrm{rec}})$, and $\rho$ to be within declared bounds; otherwise the verdict is reject.
\item (Method) further comprising equalization by adjusting internal displays or parameters to minimize $\lvert\lambda_{\mathrm{kin}}-\lambda_{\mathrm{rec}}\rvert$ subject to constraints, followed by re-testing the inequality.
\item (Method) wherein logs include timestamps, device IDs, run identifiers, masks or regions-of-interest, and the acceptance decision signed with an instrument key.
\item (System) further comprising a compliance API that exposes route values, uncertainties, $\rho$, $k$, inequality residual, and verdict, with cryptographic signatures.
\item (Method) wherein the decision may be computed per-window and aggregated over blocks with a declared aggregation rule (e.g., worst-case, median, or bounded mean) before final verdict.
\item (Method) further comprising certificate integration, wherein acceptance logs are included in a calibration certificate for gating downstream operation or data ingestion.
\item (Method) wherein correlation $\rho$ is bounded by priors or physical constraints and is clamped to a declared range if estimation uncertainty exceeds a threshold.
\item (System) wherein failure cases emit structured diagnostics indicating dominant contributors to inequality violation and recommended adjustments.
\end{enumerate}

\section*{Brief Description of the Drawings}
\begin{enumerate}[leftmargin=*]
  \item Fig. 1: Route computation of $\lambda_{\mathrm{kin}}$ and $\lambda_{\mathrm{rec}}$.
  \item Fig. 2: Uncertainty and correlation estimation pipeline.
  \item Fig. 3: Acceptance inequality test and decision logic.
  \item Fig. 4: Fail-closed signaling and signed log emission.
\end{enumerate}

\section*{Detailed Description}
\subsection*{Route Invariants}
Two independent routes derive the same dimensionless invariant: a kinematic route ($\lambda_{\mathrm{kin}}$) and a recognition route ($\lambda_{\mathrm{rec}}$). Routes are computed from factorized, unit-invariant displays; masks and regions-of-interest are held constant per policy.

\subsection*{Uncertainty and Correlation}
Standard uncertainties $u(\lambda_{\mathrm{kin}})$ and $u(\lambda_{\mathrm{rec}})$ are estimated from repeated or windowed measurements; correlation $\rho$ accounts for common-mode contributions. Estimation uses bootstraps, cross-spectral estimators, or model-based variance propagation with declared confidence bounds.

\subsection*{Acceptance Inequality}
The inequality compares normalized discrepancy to combined uncertainty. If $\lvert\lambda_{\mathrm{kin}}-\lambda_{\mathrm{rec}}\rvert/\lambda_{\mathrm{rec}} \le k\,u_{\mathrm{comb}}$ holds, acceptance is granted; otherwise the verdict is reject. Policies may require re-test after equalization adjustments; failure to meet bounds remains reject.

\subsection*{Fail-Closed Policy and Logging}
If inputs are missing or uncertainty/correlation cannot be established within bounds, the verdict is reject. Logs include all decision inputs and are signed and hash-addressed; optional registry publication supports audit.

\section*{Technical Effects}
The method yields deterministic acceptance with a single interpretable inequality; improved robustness by accounting for correlation; reproducibility via unit-invariant route computation and fixed masks; and audit-ready, signed logs for compliance and enforcement.

\section*{Enablement and Best Mode}
Implementations include route-computation modules, uncertainty/correlation estimators (bootstrap and cross-spectrum), an inequality test engine with policy-tier $k$ values, and a compliance API with signed logging. Benchmarks report true/false acceptance rates and sensitivity to $k$.

\section*{Alternative Embodiments}
Extensions include more than two routes aggregated by a generalized inequality; dynamic $k$ based on operating regime; per-window vs aggregated decisions. Certificates can embed acceptance logs for downstream gating.

\section*{Legal Notice}
All embodiments, features, and steps described herein may be combined in any operable permutation. Where means-plus-function is implicated, claims are intended to cover corresponding structures and their equivalents.

\end{document}


