% Eight-Tick Neutral Scheduler — Family E Patent (Method, System, Medium)
\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{xcolor}

\title{Eight-Tick Neutral Scheduler\\Family E Patent (Method, System, Medium)}
\author{Jonathan Washburn}
\date{\today}

\begin{document}
\maketitle

\begin{center}
{\large Camera-Ready Patent Draft}\\[4pt]
{\normalsize Gray-Aligned 8-Tick Commits; Neutrality Enforcement ($\sum\delta_i=0$); J Block-Sum Minimization for Alias Suppression}
\end{center}

\section*{Field}
This disclosure relates to real-time scheduling for instruments and multi-actuator control systems. It provides a method, system, and non-transitory medium for scheduling commits in Gray-aligned \emph{eight-tick} windows, enforcing neutrality ($\sum_{i=1}^{8}\delta_i=0$ with mirrored halves), and minimizing a J-kernel block-sum to suppress alias and phase errors in measurements and control.

\section*{Background}
Periodic operations (sensing, integration, actuator updates) are susceptible to aliasing and modal cross-interference when update timing aligns with plant resonances or when residuals accumulate. Equal-spaced or co-phased schedules often reinforce low-order modes and produce drift. There is a need for a scheduler that: (i) uses \emph{eight-tick} minimality (in $D=3$) with Gray alignment; (ii) enforces \emph{neutrality} per block; and (iii) \emph{minimizes a J block-sum} objective to reduce aliasing.

\section*{Summary}
Disclosed is an Eight-Tick Neutral Scheduler that:
\begin{enumerate}[leftmargin=*]
  \item partitions each period into eight windows with Gray-aligned commits and mirrored halves (mid-cycle FLIP);
  \item enforces neutrality per block: $\sum_{i=1}^{8}\delta_i=0$ and $\delta_{i+4}=-\delta_i$ for $i=1,\dots,4$;
  \item minimizes a block-sum of the convex symmetric J-kernel $J(x)=\tfrac12(x+1/x)-1$ over windowed observables to suppress alias and phase errors; and
  \item emits compliance logs with timing, neutrality metrics, J block-sums, and signatures.
\end{enumerate}

\section*{Definitions}
\noindent\textbf{Eight-tick window.} A block of eight sequential windows $(W_1,\dots,W_8)$ that spans the minimal spatially complete period in $D=3$.

\noindent\textbf{Gray-aligned commits.} Window transitions and commits are arranged to change one timing state at a time, minimizing switching artifacts and ensuring deterministic phase behavior.

\noindent\textbf{Neutrality.} For per-window signed updates $\delta_i$, neutrality requires $\sum_{i=1}^{8}\delta_i=0$ and mirrored halves $\delta_{i+4}=-\delta_i$ ($i=1..4$).

\noindent\textbf{J-kernel block-sum.} For a per-window observable or residual sequence $x_i>0$, the scheduler minimizes $\sum_{i=1}^{8} J(x_i)$, where $J(x)=\tfrac12(x+1/x)-1$, to suppress alias/phase errors.

\noindent\textbf{Compliance API.} Interface exposing timing parameters, neutrality observables, J block-sum metrics, and signed logs for audit.

\section*{Claims}
\setlist[enumerate]{leftmargin=*,labelsep=0.5em,itemsep=0.6em,topsep=0.4em}

\subsection*{Independent Claims}
\begin{enumerate}
\item \textbf{(Method)} A computer-implemented scheduling method for periodic measurement or control, comprising:
\begin{enumerate}[label=(\alph*)]
  \item partitioning each period into eight windows with Gray-aligned commits and mirrored halves, including a mid-cycle FLIP;
  \item enforcing neutrality per block by constraining per-window signed updates $\delta_i$ to satisfy $\sum_{i=1}^{8}\delta_i=0$ and $\delta_{i+4}=-\delta_i$ for $i=1..4$;
  \item minimizing a J-kernel block-sum $\sum_{i=1}^{8} J(x_i)$ over windowed observables $x_i>0$ with $J(x)=\tfrac12(x+1/x)-1$; and
  \item emitting signed, hash-addressed logs including timing, neutrality metrics, block-sum values, and configuration.
\end{enumerate}

\item \textbf{(System)} A scheduling system comprising one or more processors, memory, and timing circuitry configured to: (i) generate eight-window Gray-aligned schedules with mirrored halves; (ii) enforce neutrality constraints per block; (iii) compute and minimize a J block-sum over windowed observables; and (iv) emit compliance logs with signatures.

\item \textbf{(Non-transitory medium)} A non-transitory computer-readable medium storing instructions that, when executed, cause a system to perform the method of claim 1.
\end{enumerate}

\subsection*{Dependent Claims}
\begin{enumerate}[resume]
\item (Method) wherein timing edges use raised-cosine ($C^1$) envelopes to reduce spectral leakage while preserving neutrality constraints.
\item (Method) wherein jitter is bounded by $\varepsilon_j$ and verified by an external sensor; violations trigger fail-closed behavior.
\item (Method) wherein per-window observables $x_i$ are derived from canonically factorized, dimensionless displays.
\item (Method) wherein neutrality is validated from synchronized sensing (e.g., EL/PL, impedance, IR) by a block-sum neutrality observable $Z=\sum_{i=1}^{4}(M_i-M_{i+4})$.
\item (Method) further comprising per-window aggregation rules (median, trimmed mean) for robust J block-sum computation.
\item (System) further comprising an FPGA or SoC timing engine generating eight-tick schedules with bounded jitter and a streaming J-kernel core computing $J(e^{u})=\cosh(u)-1$ in log-domain with a block-sum accumulator.
\item (Method) wherein alias suppression is quantified by reduction in a declared alias/phase error metric relative to an equal-spaced baseline.
\item (Method) wherein mirrored halves are constructed as exact inverses in signed updates subject to dwell and slew constraints.
\item (System) wherein compliance logs include timing ratios, jitter statistics, neutrality observables, J block-sum values, and acceptance verdicts.
\item (Method) further comprising a compliance API exposing schedule state (BeginWindow($\ell$), EndWindow($\ell$), WindowIndex()), neutrality flags, and block-sum metrics with signatures.
\end{enumerate}

\section*{Brief Description of the Drawings}
\begin{enumerate}[leftmargin=*]
  \item Fig. 1: Eight-tick Gray-aligned schedule with mirrored halves and mid-cycle FLIP.
  \item Fig. 2: Neutrality constraints and block-sum neutrality observable $Z$.
  \item Fig. 3: J-kernel block-sum computation over windowed observables.
  \item Fig. 4: Timing engine (FPGA/SoC) and compliance logging interface.
\end{enumerate}

\section*{Detailed Description}
\subsection*{Scheduler Core}
Each period is partitioned into eight windows $(W_1..W_8)$. Gray alignment ensures only one timing state changes at a time. Mirrored halves enforce $\delta_{i+4}=-\delta_i$ with a mid-cycle FLIP. Dwell and slew constraints are enforced to guarantee realizability.

\subsection*{Neutrality Enforcement}
Per-window signed updates $\delta_i$ (e.g., integrated exposure, actuator postings) satisfy $\sum_{i=1}^{8}\delta_i=0$. Neutrality may be verified from synchronized sensing via a block-sum neutrality observable $Z=\sum_{i=1}^{4}(M_i-M_{i+4})$ expected to be zero-mean within declared bands.

\subsection*{J Block-Sum Minimization}
For positive per-window observables $x_i$, a convex symmetric cost $J(x)=\tfrac12(x+1/x)-1$ is minimized over the block to reduce alias/phase errors. Robust aggregation (median/trimmed mean) may be used against outliers. Alias suppression is evidenced by improved metrics relative to equal-spaced baselines.

\subsection*{Compliance and Logging}
The scheduler records timing parameters, jitter, neutrality metrics, J block-sum values, and acceptance verdicts in signed, hash-addressed logs. A compliance API exposes current window state, neutrality flags, and block-sum metrics for audit and downstream gating.

\section*{Technical Effects}
The scheduler yields: (i) reduced alias/phase error via neutrality and J minimization; (ii) deterministic periodic invariance with mirrored halves; (iii) robustness to timing perturbations through Gray alignment and jitter bounds; and (iv) auditability via signed logs and compliance API.

\section*{Enablement and Best Mode}
Reference implementations include an FPGA/SoC timing engine, raised-cosine window envelopes, a streaming J-kernel core with a block-sum accumulator, and synchronized sensing for neutrality verification. Benchmarks quantify alias reduction vs equal-spaced schedules.

\section*{Alternative Embodiments}
Window envelopes may vary (provided $C^1$ smoothness or policy criteria). Jitter budgets, dwell/slew constraints, and aggregation rules can be tuned per modality. Additional windows ($>8$) may be used while preserving mirrored neutrality.

\section*{Legal Notice}
All embodiments, features, and steps described herein may be combined in any operable permutation. Where means-plus-function is implicated, claims are intended to cover corresponding structures and their equivalents.

\end{document}


