% Reality Bridge Calibration OS — Family B Patent (Method, System, Medium)
\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{xcolor}

\title{Reality Bridge Calibration OS (Core)\\Family B Patent (Method, System, Medium)}
\author{Jonathan Washburn}
\date{\today}

\begin{document}
\maketitle

\begin{center}
{\large Camera-Ready Patent Draft}\\[4pt]
{\normalsize Units-Quotient Factorization, AbsoluteLayer Selection, K-Route Equalization, Layer Non-Mixing}
\end{center}

\section*{Field}
This disclosure relates to calibration and measurement control for instruments and simulators. It provides a computer-implemented operating system (OS) that performs units-quotient factorization of displays, selects a unique \emph{AbsoluteLayer} by solving simultaneous dimensionless gate identities, equalizes independent calibration routes (K-identities), and enforces layer non-mixing (Planck vs IR) to produce anchor-free, units-invariant calibration outputs.

\section*{Background}
Calibration pipelines typically depend on external anchors, manually tuned parameters, and per-device unit conventions. These practices introduce drift, reduce reproducibility, and obscure provenance. Prior methods do not: (i) factor displays through a \emph{units quotient} to eliminate unit gaming; (ii) select a unique absolute layer by solving multiple dimensionless identities concurrently; (iii) enforce \emph{route equalization} across independent derivation paths; and (iv) enforce \emph{non-mixing} between fundamentally distinct layers (e.g., Planck vs IR) within a single calibration solution.

\section*{Summary}
Disclosed is a calibration OS that:
\begin{enumerate}[leftmargin=*]
  \item performs \textbf{units-quotient factorization} to compute dimensionless displays invariant under unit rescaling;
  \item computes dimensionless \textbf{gate identities} and solves them simultaneously to select a unique \textbf{AbsoluteLayer} (UniqueCalibration \$\wedge$\ MeetsBands);
  \item computes \textbf{K-route} invariants for independent routes and \textbf{equalizes} them under combined-uncertainty tolerance; and
  \item enforces \textbf{layer non-mixing} (Planck vs IR) so that anchors and identities from distinct layers are not combined in a single solution.
\end{enumerate}
Embodiments include method, system, and non-transitory medium claims, with optional compliance APIs and logs.

\section*{Definitions}
\noindent\textbf{Units-quotient factorization.} A process mapping displays (unitful quantities) to a quotient space of dimensionless displays invariant under unit transformations. Let $[y]$ denote the unit class of $y$; the factorization produces canonical, dimensionless $\tilde y$.

\noindent\textbf{Gate identities.} Dimensionless equalities that must hold concurrently at the AbsoluteLayer. Examples: \emph{K-identities} and \emph{Planck gate}. 
\begin{align*}
\text{K-identities:}\quad &\frac{\tau_{\mathrm{rec}}}{\tau_0} = \frac{\lambda_{\mathrm{kin}}}{\ell_0} = K,\\
\text{Planck gate:}\quad &\frac{c^3\,\lambda_{\mathrm{rec}}^2}{\hbar\,G} = C_{\mathrm{layer}}\ (\text{constant by layer}).
\end{align*}

\noindent\textbf{AbsoluteLayer.} The unique layer (solution) produced by solving the simultaneous gate identities within declared tolerance bands (\emph{UniqueCalibration \$\wedge$\ MeetsBands}).

\noindent\textbf{K-route equalization.} Enforcement that independent derivation routes yield equal dimensionless invariants within combined uncertainty; otherwise the calibration fails closed or is adjusted.

\noindent\textbf{Layer non-mixing.} A policy and constraint set prohibiting concurrent use of anchors or identities from fundamentally distinct layers (e.g., Planck vs IR) within a single calibration solution.

\noindent\textbf{Compliance API.} An interface exposing factorized displays, identity residuals, equalization metrics, non-mixing decisions, and run logs with signatures.

\section*{Claims}
\setlist[enumerate]{leftmargin=*,labelsep=0.5em,itemsep=0.6em,topsep=0.4em}

\subsection*{Independent Claims}
\begin{enumerate}
\item \textbf{(Method)} A computer-implemented method for calibration of an instrument or simulator, comprising:
\begin{enumerate}[label=(\alph*)]
  \item performing \emph{units-quotient factorization} to compute dimensionless displays from unitful observables;
  \item computing a plurality of \emph{dimensionless gate identities} and solving them \emph{simultaneously} to select a unique \emph{AbsoluteLayer} within declared tolerance bands;
  \item computing \emph{K-route} invariants for independent derivation routes and \emph{equalizing} the routes within a combined-uncertainty tolerance; and
  \item enforcing \emph{layer non-mixing} such that identities and anchors from distinct layers (including Planck vs IR) are not combined in a single calibration solution;
  \item outputting calibrated scales and constants and recording calibration logs.
\end{enumerate}

\item \textbf{(System)} A calibration system comprising one or more processors, memory, and I/O configured to: (i) factor displays through a units quotient; (ii) compute and solve simultaneous gate identities to select a unique AbsoluteLayer; (iii) compute K-route invariants and equalize independent routes within combined uncertainty; (iv) enforce layer non-mixing; and (v) output calibrated scales/constants and logs.

\item \textbf{(Non-transitory medium)} A non-transitory computer-readable medium storing instructions that, when executed by one or more processors, cause performance of the method of claim 1.
\end{enumerate}

\subsection*{Dependent Claims}
\begin{enumerate}[resume]
\item (Method) wherein units-quotient factorization comprises canonicalization of unitful quantities to dimensionless displays via a quotient mapping that is invariant under unit transformations.
\item (Method) wherein the gate identities include K-identities $\tau_{\mathrm{rec}}/\tau_0 = \lambda_{\mathrm{kin}}/\ell_0$ and a Planck-gate identity $(c^3\,\lambda_{\mathrm{rec}}^2)/(\hbar G)=C_{\mathrm{layer}}$.
\item (Method) wherein simultaneous solving comprises a constrained optimization or root-finding with tolerance bands and convergence criteria (MeetsBands).
\item (Method) wherein K-route equalization uses a combined-uncertainty tolerance model with correlation.
\item (Method) wherein layer non-mixing is enforced by policy constraints that block any calibration solution combining Planck-layer and IR-layer anchors or identities.
\item (Method) further comprising emitting an acceptance verdict and logs containing identity residuals, equalization metrics, and non-mixing decisions.
\item (System) further comprising a compliance API exposing factorized displays, identity residuals, K-route metrics, and non-mixing flags with cryptographic signatures.
\item (Method) wherein AbsoluteLayer selection emits absolute scales $(\tau_0,\ell_0)$ and constants $(c,\hbar,\lambda_{\mathrm{rec}},G)$ internally without external anchors.
\item (Method) wherein failure to satisfy simultaneous identities or K-route equalization triggers a fail-closed outcome and a structured diagnostic report.
\item (System) wherein calibration logs are signed, hash-addressed, and optionally published to a transparency registry.
\end{enumerate}

\section*{Brief Description of the Drawings}
\begin{enumerate}[leftmargin=*]
  \item Fig. 1: Units-quotient factorization mapping from unitful displays to dimensionless displays.
  \item Fig. 2: Simultaneous gate-identity solve producing a unique AbsoluteLayer (identity graph).
  \item Fig. 3: K-route equalization flow with combined-uncertainty model.
  \item Fig. 4: Layer non-mixing policy diagram (Planck vs IR separation).
  \item Fig. 5: Compliance API signals and signed calibration logs.
\end{enumerate}

\section*{Detailed Description}
\subsection*{Units-Quotient Factorization}
The system identifies unit classes and forms a quotient mapping that sends each display to a canonical, dimensionless representative. Canonicalization removes unit ambiguity and prevents unit gaming. Factorization precedes any identity evaluation and is deterministic under declared unit equivalences.

\subsection*{Simultaneous Gate-Identity Solve (AbsoluteLayer)}
Dimensionless gate identities are constructed from factorized displays. The solver enforces \emph{UniqueCalibration \$\wedge$\ MeetsBands}: a single solution exists within specified tolerance bands and convergence criteria. Example identities include K-identities and a Planck-gate constraint. Solutions outside tolerance fail closed with diagnostics.

\subsection*{K-Route Equalization}
Independent derivation routes (e.g., kinematic vs recognition) produce dimensionless invariants. The OS equalizes routes within a combined-uncertainty tolerance accounting for correlation. If equalization fails, the calibration either iterates (adjusting internal displays) or fails closed with a structured report.

\subsection*{Layer Non-Mixing (Planck vs IR)}
The OS enforces layer separation by policy and constraints: identities and anchors from distinct layers (e.g., Planck-level invariants vs IR-level invariants) cannot be combined in a single calibration solution. The policy is checked mechanically during identity solving and equalization. Violations result in blocking the calibration and emitting non-mixing diagnostics.

\subsection*{Compliance API and Logs}
Implementations provide a compliance API exposing factorized displays, identity residuals, K-route metrics, non-mixing flags, and acceptance verdicts, with cryptographic signatures and content-addressable logs for audit.

\section*{Technical Effects}
The OS yields: (i) units-invariant, anchor-free calibration; (ii) deterministic AbsoluteLayer selection with reproducibility across environments; (iii) improved accuracy via route equalization; (iv) safety via fail-closed policy under non-mixing and identity mismatch; and (v) auditable logs enabling conformance and enforcement.

\section*{Enablement and Best Mode}
Reference implementations include: (i) factorization libraries for units quotient; (ii) identity solvers with tolerance bands and convergence criteria; (iii) K-route equalization modules with combined-uncertainty models; (iv) non-mixing policy engine; and (v) compliance API with signed logs. Benchmarks quantify reproducibility and identity residuals.

\section*{Alternative Embodiments}
Additional identities may be included depending on modality. Equalization can be extended to more than two routes. Logs may be on-device or cloud-registered. The OS may interoperate with certificate engines for external issuance.

\section*{Legal Notice}
All embodiments, features, and steps described herein may be combined in any operable permutation. Where means-plus-function is implicated, claims are intended to cover corresponding structures and their equivalents.

\end{document}


