\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{hyperref}

\title{CPM-Driven Folding Plan}
\author{}
\date{}

\newcommand{\Struct}{\mathcal{S}}
\newcommand{\Defect}{\mathcal{D}}
\newcommand{\Energy}{\mathcal{E}}

\begin{document}

\maketitle

\section{Guiding Principle}

Recognition Science (RS) is the unique zero-parameter architecture of reality. The Coercive Projection Method (CPM) is the reusable proof template that maps RS structure into engineering steps. Protein folding must therefore be implemented as CPM on the RS fragment/ledger manifold: structured modes, local defects, coercive projections, dispersion gates, and phase aggregators.

\section{CPM \texorpdfstring{$\to$}{→} Engineering Mapping}

\begin{itemize}[leftmargin=*]
\item \textbf{Structured set $\Struct$}: Finite fragment net (3-, 5-, 9-mer motifs) spanning helix/strand/turn/loop modes. Each fragment is discretized by RS torsion bins (dyadic/$\varphi$-tier) and carries ledger-compatible side-chain orientation and hydrogen-bond handles.

\item \textbf{Defect functional $\Defect$}: For each CA window, measure the blended metric
\begin{itemize}
\item distance to nearest fragment in $\Struct$,
\item RS ledger cost $J$ (torsion feasibility, dihedral balance, eight-tick timing),
\item local hydrogen-bond alignability and steric slack.
\end{itemize}
Global defect is the sum over windows.

\item \textbf{Projection operator $\Pi$}: CPM rank-one move = select a window, snap it toward its nearest fragment by a blended step (keep endpoints fixed, smooth neighbors). Guarantees an energy gap $\Energy - \Energy_0 \ge c\,\Defect$ with CPM constants ($C_{\text{proj}} \approx 2$, $\varepsilon \approx 0.1$, local energy factor $C_{\text{eng}}$).

\item \textbf{Dispersion gate}: Eight-beat Listen/Lock/Balance schedule down-weights excursions away from $\Struct$; contacts violating the projected mode get temporary weight reductions; backbone smoothing enforces CPM dispersion control.

\item \textbf{Aggregator / phase transition}: When the maximum window defect falls below phase thresholds, advance to the next recognition phase (Lock $\to$ Balance). Tighten step sizes and net radius; refresh fragment priorities.

\item \textbf{Parameter schedule}: RS dictates dyadic/$\varphi$-tier steps, fragment amplitudes, and ledger tolerances; CPM constants choose blending factors and acceptance bands.
\end{itemize}

\section{Engineering Milestones}

\begin{enumerate}[leftmargin=*]
\item \textbf{Define structured set} \textit{(Task \texttt{cpm-1-structured-set})}
\begin{itemize}
\item Build fragment library with RS torsion bins and pre-computed local frames.
\item Store fragment metadata (H-bond vectors, steric envelope, ledger cost).
\end{itemize}

\item \textbf{Implement defect evaluator} \textit{(Task \texttt{cpm-2-defect})}
\begin{itemize}
\item For each sliding window, compute nearest fragment distance + RS ledger penalty.
\item Output per-window defect and aggregate statistics for gating.
\end{itemize}

\item \textbf{Add CPM projection move} \textit{(Task \texttt{cpm-3-projector})}
\begin{itemize}
\item Randomly choose a window, project onto nearest fragment by blended snap.
\item Preserve continuity by smoothing adjacent residues; enforce CPM constants.
\end{itemize}

\item \textbf{Local energy evaluation} \textit{(Task \texttt{cpm-4-local-energy}, in progress)}
\begin{itemize}
\item Evaluate $\Delta E$ using only interactions touching the window: contacts, sterics, Rama, H-bonds, solvent proxies.
\item Feed $\Delta E$ into Metropolis acceptance at phase-specific temperature.
\end{itemize}

\item \textbf{Dispersion gating} \textit{(Task \texttt{cpm-5-dispersion})}
\begin{itemize}
\item Couple move acceptance weights to Listen/Lock/Balance timers.
\item Down-weight violated contacts, up-weight restoration moves, enforce eight-beat rhythm.
\end{itemize}

\item \textbf{Aggregator thresholds} \textit{(Task \texttt{cpm-6-aggregator})}
\begin{itemize}
\item Define defect bands that trigger phase advancement.
\item Log recognition trajectory (defect vs.\ time, accepted projections).
\end{itemize}

\item \textbf{RS schedules} \textit{(Task \texttt{cpm-7-schedule})}
\begin{itemize}
\item Encode dyadic/$\varphi$-tier temperature, blending, and sampling schedules.
\item Synchronize with eight-tick IR timing for eventual experiment alignment.
\end{itemize}

\item \textbf{Benchmark validation} \textit{(Task \texttt{cpm-8-validation})}
\begin{itemize}
\item Run CPM loop on 1VII, 1PGB, 1ENH; record defect drop, RMSD, energy traces.
\item Compare against baseline to demonstrate coercivity-driven improvement.
\end{itemize}
\end{enumerate}

\section{Immediate Focus}

\begin{itemize}[leftmargin=*]
\item Finish \texttt{cpm-4-local-energy}: integrate window-restricted energy evaluation into the optimizer acceptance rule.
\item Implement \texttt{cpm-3-projector}: CPM projection move using fragment net and blended torsion update.
\end{itemize}

\section{Deliverables}

\begin{itemize}[leftmargin=*]
\item Fragment net data structure and ledger constants.
\item CPM defect/energy logs per phase and benchmark protein.
\item Updated optimizer with local $\Delta E$ acceptance and projection moves.
\item Technical note documenting CPM constants $(C_{\text{proj}}, C_{\text{net}}, C_{\text{eng}})$ and their empirical calibration.
\end{itemize}

\end{document}

