\documentclass[11pt]{article}

\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{setspace}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!70!black,
  citecolor=blue!70!black,
  urlcolor=blue!70!black
}

\newcommand{\phig}{\varphi}
\newcommand{\Ecoh}{E_{\mathrm{coh}}}
\newcommand{\Fgap}{\mathcal F}

\setstretch{1.15}

\title{\textbf{Why Particle Masses Have Structure}\\[0.3em]
\large A plain-language explanation of the Recognition Science mass framework}
\author{}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\noindent
This paper explains, in plain language, where the mass formula comes from and why each piece exists.
The goal is not to convince you the framework is correct---it is to show you what the framework \emph{claims},
why those claims are logically forced once you accept the starting point, and how you can falsify or verify them.
We focus on the conceptual story: what problem each ingredient solves, how it connects to physics, and why alternatives would fail.
\end{abstract}

\tableofcontents
\newpage

%===========================================
\section{The Central Question}
%===========================================

The Standard Model of particle physics contains roughly two dozen free parameters.
Among these are the masses of quarks and leptons: the electron, muon, tau, up, down, strange, charm, bottom, and top.
These masses span an enormous range---from about half a MeV for the electron to 173 GeV for the top quark, a ratio of over 300,000 to 1.
And yet the Standard Model offers no explanation for \emph{why} these particular values.
They are simply measured, inserted, and used.

The Recognition Science mass framework proposes that these masses are not arbitrary.
It claims they follow a \emph{structural law}: a discrete pattern built from simple integers, a single irrational number (the golden ratio), and geometric reasoning about what it means for something to be stable.
This paper explains that structure piece by piece.

%===========================================
\section{The Foundational Idea: Stability Requires Closure}
%===========================================

At the heart of the framework is an idea that sounds almost tautological but has surprising consequences: \textbf{anything that persists must close}.

What does ``close'' mean? Consider a particle. It exists over time. It interacts with fields. It could, in principle, radiate away, decay, dissolve. Why doesn't it? Standard physics answers: conservation laws, gauge symmetries, quantum numbers. These are correct, but they don't explain \emph{where the specific mass values come from}.

The Recognition framework takes a different angle. It says: a stable object is one whose internal updating process returns to a consistent state. Think of it like a clock. If every tick of the clock changes the internal state, then the only way the object can be ``the same thing'' over time is if the sequence of changes eventually comes back around. The states must form a cycle. The update rule must \emph{close}.

This is not a metaphor; it is meant literally. The framework models stable particles as closures of discrete update sequences. The size of the closure---how many ticks before it returns---determines the ``weight'' of the object. More ticks in the closure cycle means more structure, which means more mass.

%===========================================
\section{Why Eight Ticks? The Minimal Closure in Three Dimensions}
%===========================================

If stable things are closures, the next question is: what is the minimum number of steps required for closure in a three-dimensional space?

The answer is eight, and here is why.

Three dimensions means three independent directions: call them $x$, $y$, $z$. At each tick, the update rule must ``address'' at least one of these dimensions---otherwise, nothing happens. But it must also eventually address all three, or else one direction would be left out of the closure.

Now, the minimal way to cover three independent bits of information (which direction?) is with a 3-bit address space. A 3-bit space has $2^3 = 8$ states. The claim is that a closure must visit all 8 combinations to be complete. If it visited only 4, it would leave some combination unaddressed; the closure would be partial, and the object would be unstable in some direction.

This gives us the ``Octave''---an 8-tick cycle that is the fundamental rhythm of stable structure. It is not chosen arbitrarily; it is the smallest complete cycle in 3D.

The framework makes a further claim: the 8-tick cycle is not arbitrary in its \emph{ordering} either. There is a special ordering (related to Gray codes and cube geometry) that minimizes the ``strain'' of transitioning from one state to the next. But for understanding the mass formula, the key point is simply: 8 is the fundamental closure number.

%===========================================
\section{The Golden Ratio: Not Mysticism, but Self-Similarity}
%===========================================

The golden ratio, $\phig = \frac{1+\sqrt{5}}{2} \approx 1.618$, appears throughout the framework. This is not numerology or mysticism. It is a consequence of a specific requirement: \textbf{self-similarity under discrete iteration}.

Here is the idea. Suppose you have a system that updates in discrete steps, and you want the system to ``scale'' gracefully---meaning that zooming in or out by one step looks the same as the original. What scaling factor has this property?

Let $x$ be the scaling factor per step. After one step, the system scales by $x$. After two steps, it scales by $x^2$. For the system to be self-similar, the ``sum'' of one-step and two-step contributions should equal the total structure. This gives the equation:
\[
x^2 = x + 1.
\]
The positive solution is $x = \phig$.

So the golden ratio is not an accident or a choice. It is the \emph{unique} scaling factor that makes discrete, step-by-step iteration self-similar. If stable objects are built from discrete closures, and if those closures must scale consistently, then the golden ratio is forced.

This means the mass ladder is built in powers of $\phig$. Moving up or down one ``rung'' multiplies or divides the mass by $\phig$. The entire spectrum of masses is organized as integer multiples of $\log_\phig$.

%===========================================
\section{The Coherence Unit: Where the Scale Begins}
%===========================================

If masses are organized as powers of $\phig$, we need a reference point: a ``unit'' of mass-energy that sets the scale. The framework calls this the \textbf{coherence energy}, denoted $\Ecoh$.

Where does it come from? The framework derives it from the closure structure itself. An 8-tick closure has a natural ``energy cost'' associated with completing the cycle. That cost, expressed in the self-similar $\phig$-units, turns out to be:
\[
\Ecoh = \phig^{-5}.
\]
In conventional units (with appropriate conversion factors), this sets a characteristic energy around the electroweak scale.

Why $-5$? The exponent arises from the geometry of the 8-tick closure interacting with the 3D structure of space. The details involve counting how many ``passive'' degrees of freedom are present at each tick versus how many are ``active.'' The number 5 is not arbitrary; it is $8 - 3 = 5$, the number of ticks in an 8-cycle that are not directly aligned with a coordinate axis.

You do not need to accept this derivation to understand the framework. The key point is: there is a single, fixed energy unit, derived from the closure structure, and all masses are expressed as integer-offset powers of $\phig$ times this unit.

%===========================================
\section{Sector Yardsticks: Why Quarks and Leptons Have Different Baselines}
%===========================================

If all particles lived on the same $\phig$-ladder with the same starting point, we would expect quarks and leptons to interleave smoothly. They do not. Quarks are generally heavier than leptons at comparable generation. The up quark is heavier than the electron; the top quark is much heavier than the tau.

The framework explains this with \textbf{sector yardsticks}: each sector (leptons, up-type quarks, down-type quarks) has its own baseline, its own ``zero'' on the ladder.

Why? The framework's answer is that sectors differ in their \emph{charge structure}, and charge structure affects how the closure cycle couples to the external field. A quark has color charge; a lepton does not. A quark has fractional electric charge; a lepton has integer charge. These differences shift the effective starting point of the ladder.

Mathematically, the sector yardstick takes the form:
\[
A_{\text{sector}} = 2^B \cdot \Ecoh \cdot \phig^{r_0},
\]
where $B$ and $r_0$ are integers specific to each sector. The factor $2^B$ is a binary shift (the framework interprets this as related to ``how many dimensions of the closure are 'active' versus 'passive' for this sector''). The factor $\phig^{r_0}$ is a $\phig$-offset that places the sector on the right part of the ladder.

Crucially, $B$ and $r_0$ are \emph{sector-global}. They are not adjusted per particle. All leptons share one yardstick; all up-type quarks share another. This is a strong constraint that the framework imposes on itself.

%===========================================
\section{The Integer Charge Index: From Charge to Band}
%===========================================

Within each sector, individual particles are distinguished by their \emph{charge}. The electron has charge $-1$; the muon has charge $-1$; the tau has charge $-1$. They all have the same charge, so how do they differ?

The framework says they differ in \emph{generation}, which corresponds to different ``rungs'' on the ladder. But before we get to rungs, we need to understand how charge itself affects the mass.

The framework defines an \textbf{integer charge index} $Z$ for each particle. The formula is:
\[
Z = (6Q)^2 + (6Q)^4,
\]
where $Q$ is the electric charge. For quarks, there is an additional offset:
\[
Z = 4 + (6Q)^2 + (6Q)^4.
\]
Why this formula? The factor of 6 makes all charges integral: $6 \times \frac{2}{3} = 4$, $6 \times \frac{1}{3} = 2$, $6 \times 1 = 6$. The squares and fourth powers capture how charge ``couples'' in different ways (quadratic for leading electromagnetic effects, quartic for higher-order structure).

The result is that each particle gets a specific integer $Z$. For example:
\begin{center}
\begin{tabular}{lcc}
\toprule
Particle & $Q$ & $Z$ \\
\midrule
Electron/Muon/Tau & $-1$ & $6^2 + 6^4 = 36 + 1296 = 1332$ \\
Up/Charm/Top & $+\frac{2}{3}$ & $4 + 4^2 + 4^4 = 4 + 16 + 256 = 276$ \\
Down/Strange/Bottom & $-\frac{1}{3}$ & $4 + 2^2 + 2^4 = 4 + 4 + 16 = 24$ \\
\bottomrule
\end{tabular}
\end{center}

%===========================================
\section{The Band Map: Converting Integers to Exponents}
%===========================================

Now we have an integer $Z$ for each particle. How does this affect the mass? The framework uses a \textbf{band map}:
\[
\Fgap(Z) = \frac{\ln(1 + Z/\phig)}{\ln \phig}.
\]
This function converts the integer $Z$ into a shift in the $\phig$-exponent.

Why this particular function? It is the simplest, parameter-free way to turn an integer into a multiplicative factor in the $\phig$-basis. Notice that:
\[
\phig^{\Fgap(Z)} = 1 + \frac{Z}{\phig}.
\]
So the band map says: ``your charge index $Z$ shifts your mass by a factor of $1 + Z/\phig$.'' This is linear in $Z$ for small $Z$ and logarithmic for large $Z$---a natural interpolation.

The band map is \emph{not fitted}. It has no adjustable parameters. It is a fixed, closed-form function of the integer $Z$.

%===========================================
\section{Rungs: The Large-Scale Ladder}
%===========================================

We now have: a coherence unit, a sector yardstick, and a band map from charge to fine structure. But we still have not explained \emph{generations}. Why is the muon 200 times heavier than the electron? Why is the tau 17 times heavier than the muon?

The framework's answer is \textbf{rungs}. Each particle sits on an integer rung $r$ of the $\phig$-ladder. The mass formula is:
\[
m = A_{\text{sector}} \cdot \phig^{r + \Fgap(Z) - 8},
\]
where the $-8$ is an \emph{octave-closure reference} we will explain shortly.

For the leptons:
\begin{center}
\begin{tabular}{lc}
\toprule
Particle & Rung $r$ \\
\midrule
Electron & 2 \\
Muon & 13 \\
Tau & 19 \\
\bottomrule
\end{tabular}
\end{center}

The differences are: 11 rungs from electron to muon, and 6 rungs from muon to tau. Where do these numbers come from?

\paragraph{The generation torsion: why 11 and 6.}
The numbers 11 and 6 are not fitted. They come directly from the cube integers:
\begin{itemize}
\item $E_{\text{passive}} = 11$: the number of passive edges of a cube (12 edges minus 1 active edge per tick).
\item $F_{\text{cube}} = 6$: the number of faces of a cube.
\end{itemize}

So the generation steps have a dominant integer backbone:
\begin{align}
\Delta r_{e \to \mu} &\approx E_{\text{passive}} = 11, \\
\Delta r_{\mu \to \tau} &\approx F_{\text{cube}} = 6.
\end{align}

But the actual lepton predictions used in the repo refine these integers by small, fully specified corrections (still with no per-species fitting):
\begin{align}
S_{e \to \mu} &= E_{\text{passive}} + \frac{1}{4\pi} - \alpha^2, \\
S_{\mu \to \tau} &= F_{\text{cube}} - \frac{2W+3}{2}\,\alpha,
\end{align}
where $W=17$ is the wallpaper-group count and $\alpha$ is the fine-structure constant (derived from the same integer layer).

\paragraph{Why this matters numerically.}
Starting from the electron, the pure integer step $\phig^{11}$ gives a muon baseline of about $101.7$ MeV. The refined step adds about $0.079$ in the exponent (a factor $\phig^{0.079}\approx 1.039$), producing the $\approx 105.7$ MeV prediction. This $\sim 4\%$ lift is \emph{not} coming from RG transport; it is part of the framework's derived generation-step model.

The total span from electron to tau is $11 + 6 = 17$, which equals the number of wallpaper groups. This is not a coincidence in the framework's view: generation structure \emph{is} topological structure, and topology is counted by these classification integers.

\paragraph{Why generation steps are topological.}
Moving from one generation to the next is not like ``making the particle heavier by some amount.'' It is like adding a \emph{topological feature} to the closure structure---an additional loop, handle, or face that the 8-tick cycle must traverse. The number of rungs added corresponds to how much additional ``structure'' must be closed over. This is why the steps are integers (11 edges, 6 faces) rather than continuous parameters.

%===========================================
\section{The Octave Reference: Why $-8$ Appears in the Formula}
%===========================================

You may have noticed an explicit $-8$ in the mass formula. This is not an arbitrary normalization. It is the framework's way of saying: \textbf{the mass scale is referenced to an octave-closure boundary}.

\paragraph{What is an octave-closure boundary?}
In the 8-tick model, a stable object completes one full cycle every 8 ticks. At tick 0, the object is in some state. At tick 8, it returns to a compatible state---the cycle closes. This closure point is special: it is where the object ``confirms its existence'' by successfully completing a full cycle.

The $-8$ offset in the exponent anchors the mass scale to this closure point. Without a canonical cycle length, the ``zero'' of the rung coordinate would be arbitrary---you could shift all rungs by any constant without changing the physics. But \emph{with} a canonical 8-tick cycle, there is a preferred reference: the completion of one full octave.

\paragraph{Why this matters.}
In the Standard Model, the overall scale of masses is set by the Higgs vacuum expectation value, which is a continuous parameter. In the Recognition framework, the scale is set by \emph{closure}---specifically, by how many octave boundaries the particle's structure spans. This is a qualitatively different kind of explanation: discrete rather than continuous, structural rather than parametric.

\paragraph{Machine-certified Gray adjacency.}
The claim that ``8 is special'' is not just an assertion. It is backed by a machine-checked proof in Lean. The proof establishes:
\begin{enumerate}
\item There exists a specific 8-cycle over 3-bit patterns (the Gray code: $000 \to 001 \to 011 \to 010 \to 110 \to 111 \to 101 \to 100 \to 000$).
\item This cycle is \emph{surjective}: it visits all 8 possible 3-bit patterns.
\item Consecutive phases differ by \emph{exactly one bit} (Gray adjacency).
\end{enumerate}

The Gray adjacency property is the formal version of ``atomic update'': each tick changes only one thing. This is not a loose metaphor; it is a certified mathematical fact. If you accept that fundamental updates must be atomic (change one bit at a time), then 8 is the minimal cycle length that covers all 3-dimensional states.

%===========================================
\section{The Fine-Structure Constant: A Derived Quantity}
%===========================================

In the standard presentation of physics, $\alpha \approx 1/137$ is a measured constant with no deeper explanation. The framework claims to derive it.

The derivation uses the cube integers and the closure structure:
\[
\alpha^{-1} = 4\pi \cdot E_{\text{passive}} - (f_{\text{gap}} + \delta_\kappa),
\]
where $f_{\text{gap}}$ is a weight associated with the 8-tick closure (computable from a discrete Fourier transform on the closure cycle) and $\delta_\kappa$ is a small seam-counting correction.

The key ingredients are:
\begin{itemize}
\item $E_{\text{passive}} = 11$: the passive edges of the cube.
\item $4\pi$: the solid angle of a sphere, representing ``full coverage.''
\item Corrections from the discrete structure of the 8-tick cycle.
\end{itemize}

The resulting value matches the measured fine-structure constant to several significant figures. Whether this is a genuine derivation or a coincidence is an empirical question. The framework makes the claim explicit and testable.

%===========================================
\section{Putting It Together: The Mass Formula}
%===========================================

We can now write down the full mass formula:
\[
\boxed{m_i = A_{\text{sector}} \cdot \phig^{r_i + \Fgap(Z_i) - 8},}
\]
where $A_{\text{sector}} = 2^{B_{\text{sector}}} \cdot \phig^{-5} \cdot \phig^{r_{0,\text{sector}}}$ is the sector yardstick.

Each piece has a meaning:
\begin{itemize}
\item $\phig^{-5}$: the coherence unit, derived from closure geometry.
\item $2^{B_{\text{sector}}} \cdot \phig^{r_{0,\text{sector}}}$: the sector yardstick, determined by charge type.
\item $\phig^{r_i}$: the rung, an integer determined by generation and topology.
\item $\phig^{\Fgap(Z_i)}$: the band coordinate, a zero-parameter function of the charge index.
\item $\phig^{-8}$: the octave-closure reference, anchoring the scale to one complete 8-tick cycle.
\end{itemize}

The formula has no free parameters in the sense of ``knobs tuned to match data.'' The integers are either mathematical constants (cube edges, wallpaper groups) or sector-global values derived from those constants. The only irrational number is $\phig$, which is forced by self-similarity.

\paragraph{What the $-8$ buys us.}
Without the octave reference, you could shift all rungs by an arbitrary constant without changing predictions. The $-8$ removes this arbitrariness: it says ``zero on the rung scale means one complete octave closure.'' This is analogous to choosing absolute zero for temperature rather than an arbitrary reference point. The octave provides a \emph{natural} zero.

%===========================================
\section{The Ledger Bridge: From Atomic Updates to Gray Codes}
%===========================================

There is a deeper layer to the ``why 8'' story. The framework does not just assert that 8-tick cycles are special; it proves a bridge theorem connecting two seemingly unrelated ideas:

\paragraph{The claim.}
If updates are ``atomic'' in the ledger sense (one posting per tick), then the parity of the ledger state changes by exactly one bit per tick.

\paragraph{What this means.}
A ``ledger'' is a model where there are two sides (debit and credit) and discrete accounts. A ``posting'' is a single entry that changes one account on one side. The claim is: if every tick makes exactly one posting, then the ``parity vector'' (which sides have odd vs.\ even totals) changes in exactly one coordinate.

This is a mathematical theorem, not an empirical claim. Given the definition of ``single posting per tick,'' the Gray adjacency follows by proof. The Lean formalization provides:
\begin{itemize}
\item \texttt{postingStep\_implies\_grayAdj}: a single posting implies one-bit parity change.
\item \texttt{legalAtomicTick\_implies\_grayAdj}: atomic ticks imply Gray adjacency.
\end{itemize}

\paragraph{Why this matters for physics.}
The bridge theorem connects abstract accounting (``one update at a time'') to concrete geometry (``traverse a Gray code''). If you believe that physical updates are atomic---that you cannot change two independent things simultaneously---then you are forced into Gray-code dynamics. And Gray-code dynamics over 3 dimensions forces an 8-tick cycle.

This is the sense in which 8 is ``derived'': it follows from atomicity + dimensionality, not from fitting to data.

%===========================================
\section{Non-Circularity: Avoiding the Tautology Trap}
%===========================================

A valid criticism of any mass formula is: ``Did you just rearrange the data you claim to predict?''

This is a serious concern, and the framework addresses it explicitly with a \textbf{non-circular protocol}:

\begin{enumerate}
\item \textbf{No per-species fitting.} The sector yardsticks, rung steps, and band map are fixed globally. You cannot adjust them separately for each particle.
\item \textbf{Structural prediction is computed without using the target mass.} When we predict the electron mass, we use the sector yardstick (derived from cube integers and wallpaper groups), the band map (a fixed function), and the generation-step formulas (derived from $\alpha$ and cube geometry). We do not use the measured electron mass anywhere in the calculation.
\item \textbf{Hold-out testing.} The repository includes a script that holds out one particle from each sector, calibrates sector-level offsets using only the remaining particles, and then predicts the held-out mass. This turns ``it looks close'' into a falsifiable procedure.
\end{enumerate}

The lepton chain (electron $\to$ muon $\to$ tau) is the cleanest example. Here are the predicted versus measured masses:

\input{../out/masses/lepton_chain_pred_vs_pdg.tex}

The predictions use \emph{only} the cube integers (12 edges, 6 faces, 11 passive edges, 17 wallpaper groups), the golden ratio, and the derived $\alpha$. No fitting to the measured masses occurs.

%===========================================
\section{RG Transport: Comparing Apples to Apples}
%===========================================

A subtlety: particle masses ``run'' with energy scale in quantum field theory. The electron mass at low energy is not quite the same as the electron mass at 100 GeV. This is renormalization group (RG) flow.

The framework handles this by separating two concerns:
\begin{enumerate}
\item \textbf{Structural law}: the discrete, $\phig$-based pattern described above. This is the framework's claim about ``where masses come from.''
\item \textbf{RG transport}: the standard quantum field theory machinery for relating masses at different scales. This is not the framework's claim; it is borrowed from the Standard Model.
\end{enumerate}

When comparing predictions to measured values, we use RG transport to bring everything to a common reference scale. The framework picks a specific scale (around 182 GeV, related to the electroweak sector) where the pattern is clearest. But the \emph{existence} of the pattern is not created by RG transport; it is a claim about the underlying structure that RG transport merely reveals or obscures depending on the scale.

Some earlier drafts confused these two roles, using the same symbol for ``structural band coordinate'' and ``RG transport residue.'' This led to apparent contradictions. The corrected framework keeps them strictly separate.

%===========================================
\section{What Is Derived, What Is Assumed, What Is Open}
%===========================================

Let us be honest about the status of each piece:

\paragraph{Derived and machine-certified.}
\begin{itemize}
\item \textbf{The 8-tick closure}: Lean proves that 8 is the period of a complete Gray cycle over 3-bit patterns, and that consecutive phases differ by exactly one bit.
\item \textbf{The ledger-to-Gray bridge}: Lean proves that ``single posting per tick'' implies one-bit parity adjacency.
\item \textbf{The cube integers}: $E = 12$ edges, $F = 6$ faces, $E_{\text{passive}} = 11$ passive edges. Verified in Lean.
\item \textbf{The wallpaper group count}: $W = 17$. This is a classification theorem from mathematics.
\item \textbf{The self-similarity of $\phig$}: $\phig^2 = \phig + 1$. Verified in Lean.
\item \textbf{The band map $\Fgap(Z)$}: defined as a closed-form function with no fitting.
\item \textbf{Phase alignment preservation}: Lean proves that if two systems satisfy ``step advances phase by 1,'' alignment is preserved under iteration.
\end{itemize}

\paragraph{Derived but with certified numeric interfaces.}
\begin{itemize}
\item \textbf{The $\alpha$ formula}: the expression is derived from cube integers and closure structure, and the numeric value is certified to match experiment within stated precision.
\item \textbf{The gap weight $w_8$}: defined via a discrete Fourier transform on the 8-tick cycle. The full derivation is scaffolded but the final numeric value is currently injected as a constant.
\item \textbf{The sector yardsticks}: expressed in terms of cube integers and wallpaper groups; the derivation is in Lean but uses the constants as inputs.
\end{itemize}

\paragraph{Open (not yet fully certified).}
\begin{itemize}
\item \textbf{The ``word $\to$ rung'' constructor}: why each particle has its specific rung integer. The ribbon/braid machinery is scaffolded but not yet the certified surface.
\item \textbf{SM RG integrals}: the framework uses standard QFT results, not independently derived.
\item \textbf{Full atomicity hypothesis}: the framework assumes that physical updates are atomic (one change per tick); this is a physical hypothesis, not a theorem.
\end{itemize}

The goal is to close all gaps, but we report them honestly.

%===========================================
\section{Falsifiability}
%===========================================

A theory that cannot be wrong is not a theory. Here is how to falsify this framework:

\begin{enumerate}
\item \textbf{Predict a new mass.} The framework makes predictions for particles not yet measured precisely, or for mass ratios. If those predictions fail, the framework fails.

\item \textbf{Break the 8-tick structure.} The framework claims that atomic updates in 3D force an 8-tick Gray cycle. If someone finds a physical system where atomic updates produce a different cycle length, or where stability arises from non-Gray dynamics, the structural foundation is wrong.

\item \textbf{Show the integers are wrong.} The cube integers (12 edges, 6 faces, 11 passive edges) and wallpaper count (17) are mathematical facts. But if the framework \emph{uses them incorrectly}---if the mapping from geometry to generation steps is wrong---the predictions will fail. In particular: if future measurements show that generation mass ratios are \emph{not} close to $\phig^{11}$ and $\phig^{6}$, the torsion hypothesis fails.

\item \textbf{Find a better anchor scale.} The framework claims the pattern is clearest at a specific scale (around 182 GeV). If a different scale reveals a \emph{different} pattern with equal or better coherence, the framework's claim about ``why this scale'' is undermined.

\item \textbf{Show the $-8$ is arbitrary.} If the octave-closure reference is just a fitting trick rather than a structural necessity, this should be demonstrable: an equally good model would work with $-7$ or $-9$ or any other offset.

\item \textbf{Demonstrate circularity.} If someone shows that the structural predictions actually depend (implicitly) on the data they claim to predict, the framework is tautological.
\end{enumerate}

We invite scrutiny on all fronts. The framework is designed to be broken if it is wrong.

%===========================================
\section{Conclusion}
%===========================================

The Recognition Science mass framework claims that particle masses are not arbitrary. It proposes:
\begin{itemize}
\item Stable objects are closures of discrete update rules, with 8-tick cycles as the minimal 3D closure.
\item The 8-tick structure is forced by atomicity: if updates change one thing at a time, Gray-code dynamics over 3 dimensions requires period 8. This is a machine-certified theorem.
\item The golden ratio is forced by self-similarity under discrete iteration.
\item Masses are organized as powers of $\phig$ times a coherence unit, with sector-global yardsticks and integer rungs.
\item The $-8$ offset in the mass formula anchors the scale to an octave-closure boundary, removing arbitrary normalization.
\item The fine-structure constant is derived from cube geometry and the 8-tick closure.
\item Generation steps (11 rungs electron$\to$muon, 6 rungs muon$\to$tau) come from cube integers: 11 passive edges and 6 faces.
\end{itemize}

The framework is explicit about what is certified (Gray adjacency, phase alignment, cube counts) versus what remains open (the word$\to$rung constructor, full atomicity hypothesis). This separation is intentional: it lets you evaluate the structural claims independently of the empirical hypotheses.

Whether these claims are correct is an empirical and mathematical question. This paper has tried to explain \emph{what} is being claimed and \emph{why}, so that you can evaluate it for yourself.

\end{document}
