\documentclass[11pt]{article}
% Robust CSV tables
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage{microtype}

% Reference aliasing to silence legacy labels
% Global numeric constants (ζ-normalized route for the certificate)
% Box constant uses only K0 + K_ξ; C_Γ=0 in the certificate path
\providecommand{\czeroplateau}{0.17620819}% Poisson plateau lower bound c0(ψ)
\providecommand{\Kzero}{0.03486808}% arithmetic tail bound K0
\providecommand{\Kxi}{K_\xi}
% \providecommand{\CboxZeta}{0.19486808}
\providecommand{\CboxZeta}{K_0 + K_\xi}% diagnostic numerics moved to appendix (non-load-bearing)
% H^1–BMO / Hilbert constants
\providecommand{\CHzero}{0.26}% envelope: sup_t |H[φ_L](t)| (sum-form PSC)
\providecommand{\CHone}{2/\pi}% derivative: ||(H[φ_L])'||_∞ ≤ CHone / L (certificate)
% Unified Hilbert transform macro
\newcommand{\Hilb}{\mathcal H}
% Window H^1 constant and locked M_ψ (Whitney aperture absorbed in C_CE=1)
\providecommand{\CpsiHone}{0.2400}% C_ψ^{(H^1)} locked
\providecommand{\Mpsilocked}{(4/\pi)\,\CpsiHone\,\sqrt{\CboxZeta}}
\providecommand{\UpsilonLocked}{(2/\pi)\,\Mpsilocked/\czeroplateau}% diagnostic; not load-bearing
% Optional appendix lock for numeric sections
\newif\ifshownumerics
\shownumericsfalse
\makeatletter
% (refalias scaffolding removed)
\makeatother
\AtBeginDocument{%
  % refalias disabled to keep labels explicit and avoid aliasing to optional material
  % \refalias{sec:CH-envelope}{lem:CH-explicit}%
  % \refalias{lem:poisson-lower}{lem:poisson-scale-stage2}%
  % \refalias{lem:hilbert-aux}{lem:hilbert-H1BMO}%
  % \refalias{lem:laplace-szego}{prop:discrete-Poisson}%
  % \refalias{lem:cayley-cont}{lem:Cayley-diff}%
  % \refalias{lem:wedge-stage2}{thm:numeric-close-stage2}%
  % bridge aliases removed to avoid early expansion issues
  % \refalias{sec:bridge-C}{thm:bridge-C}%
  % \refalias{thm:BridgeA}{thm:bridgeA}%
}

% Theorems
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{hypothesis}[theorem]{Hypothesis}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Macros
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\PP}{\mathcal{P}}
\newcommand{\HS}{\mathcal{S}_2}
\newcommand{\Half}{\{\,s\in\C:\ \Re s>\tfrac12\,\}}
\newcommand{\Poisson}{P}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\dettwo}{det_2}
\DeclareMathOperator{\Arg}{Arg}

% Title & authors
\title{A boundary product--certificate reduction of the Riemann Hypothesis}
% --- AAB helpers ---
\newcommand{\AAB}{\textup{A\kern-0.05em A\kern-0.05em B}}
\DeclareMathOperator{\AABop}{A\!A\!B}
\author{Jonathan Washburn\\ Recognition Science, Recognition Physics Institute\\ Austin, Texas, USA\\ jon@recognitionphysics.org}
\date{September 2025}

\begin{document}
\maketitle

\begin{abstract}
We reduce the Riemann Hypothesis to a boundary product--certificate route on $\{\Re s>\tfrac12\}$.
The globalization/pinch mechanism is unconditional once an almost-everywhere boundary wedge \textup{(P+)} is available for a normalized ratio.
The remaining analytic bottleneck is a scale-free Whitney box-energy input for the $\xi$-field (Hypothesis~\ref{hyp:Kxi-uniform}); equivalently, a uniform prohibition on extreme zero--clustering at the microscopic scale $H\asymp 1/\log\langle T\rangle$, which can be formulated as a uniform weighted short-interval variation bound for the classical $S(T)$ term (Lemma~\ref{lem:xi-energy-Svariation}). Unconditionally we only prove the weaker $O(|I|\log\langle T\rangle)$ Whitney-box energy bound (Lemma~\ref{lem:carleson-xi}).
Under the scale-free box-energy hypothesis, the manuscript’s certificate machinery yields \textup{(P+)} and hence RH by the Schur--Herglotz pinch.
\end{abstract}

\paragraph{Keywords.} Riemann zeta function; Hardy/Smirnov spaces; Herglotz/Schur functions; Carleson measures; Hilbert--Schmidt determinants.

\paragraph{MSC 2020.} 11M26, 30D15, 30C85; secondary 47A12, 47B10.

\section*{Notation and conventions}
\begin{itemize}
\item Half–plane: $\Omega:=\{\Re s>\tfrac12\}$; boundary line $\Re s=\tfrac12$ parameterized by $t\in\R$ via $s=\tfrac12+it$.
\item Outer/inner: for a holomorphic $F$ on $\Omega$, write $F=I\,O$ with $O$ outer (zero–free; boundary modulus $e^{u}$) and $I$ inner (Blaschke and singular inner factors).
\item Herglotz/Schur: $H$ is Herglotz if $\Re H\ge 0$ on $\Omega$; $\Theta$ is Schur if $|\Theta|\le 1$ on $\Omega$. Cayley: $\Theta=(H-1)/(H+1)$.
\item Poisson/Hilbert: $P_a(x)=\tfrac{1}{\pi}\tfrac{a}{a^2+x^2}$; boundary Hilbert transform $\Hilb$ on $\R$.
\item Off-critical zeros: the (half-plane) \emph{defect measure} is
\[
  \nu\ :=\ \sum_{\substack{\rho=\beta+i\gamma\\ \beta>1/2}} 2(\beta-\tfrac12)\,\delta_{\rho}
  \qquad\text{on }\Omega,
\]
and the associated \emph{boundary balayage} is the absolutely continuous measure $\mu$ on $\R$ with density
\[
  \frac{d\mu}{dt}(t)\ =\ \sum_{\substack{\rho=\beta+i\gamma\\ \beta>1/2}} 2(\beta-\tfrac12)\,P_{\beta-1/2}(t-\gamma).
\]
\item Windows: fix an even $C^\infty$ flat-top window $\psi:\R\to[0,1]$ with $\psi\equiv 1$ on $[-1,1]$ and $\operatorname{supp}\psi\subset[-2,2]$ (see \emph{Printed window}). For $L>0$ and $t_0\in\R$ set
\[
  \psi_{L,t_0}(t):=\psi\!\left(\frac{t-t_0}{L}\right),\qquad
  \varphi_{L,t_0}(t):=\frac{1}{L}\,\psi\!\left(\frac{t-t_0}{L}\right),\qquad
  m_\psi:=\int_\R\psi.
\]
Then $\int_\R \varphi_{L,t_0}=m_\psi$ and $\operatorname{supp}\varphi_{L,t_0}\subset[t_0-2L,t_0+2L]$, while $\varphi_{L,t_0}\equiv L^{-1}$ on $[t_0-L,t_0+L]$.
\item Carleson boxes: $Q(\alpha I)=I\times(0,\alpha|I|]$; $C_{\rm box}$ uses the area measure $\lambda:=|\nabla U|^2\,\sigma\,dt\,d\sigma$.
\item Constants/macros: $c_0(\psi)=\czeroplateau$, $C_\psi^{(H^1)}=\CpsiHone$, $C_H(\psi)=\CHone$, $K_\xi$, $C_{\rm box}^{(\zeta)}=\CboxZeta$, $M_\psi=\Mpsilocked$, $\Upsilon=\UpsilonLocked$.
\item Scope convention: throughout, $C_{\rm box}^{(\zeta)}$ denotes the supremum over all boxes $Q(\alpha I)$ with $I\subset\mathbb R$ (fixed $\alpha\in[1,2]$).
\item Terminology (used once and consistently): PSC = product certificate route (active); AAB = adaptive analytic bandlimit (archival, not used in the main chain); KYP = Kalman–Yakubovich–Popov (appears only in archived material; not used in proofs).
\end{itemize}

\subsection*{Standing properties (used in the pinch)}\label{sec:standing-assumptions}
The globalization/pinch step (Section~\ref{sec:globalization}) uses two auxiliary facts: a right-edge normalization \textup{(N1)} for the chosen normalized ratio (ensuring a fixed limiting value for the Cayley transform as $\sigma\to+\infty$), and a non-cancellation statement \textup{(N2)} (ensuring no cancellation at off-critical $\zeta$-zeros). Both are recorded explicitly in Section~\ref{sec:globalization}; \textup{(N2)} is verified in the paragraph ``Proof of (N2)'' using the diagonal determinant and the fact that outers are zero-free.

\section*{Reader's guide}
\begin{itemize}
\item \textbf{What is unconditional.} The globalization/pinch mechanism is unconditional once an a.e.\ boundary wedge \textup{(P+)} is available (Section~\ref{sec:globalization} together with (N1)–(N2)).
\item \textbf{What remains open.} The certificate step \emph{as a proof of RH} requires a \emph{scale-free} Whitney box-energy input for the $\xi$-field (Hypothesis~\ref{hyp:Kxi-uniform}); unconditionally we only prove the weaker $O(|I|\log\langle T\rangle)$ Whitney-box bound (Lemma~\ref{lem:carleson-xi}).
\item \textbf{Equivalent formulation of the bottleneck.} By Lemma~\ref{lem:xi-energy-Svariation}, the scale-free hypothesis is implied by a uniform bound on a weighted microscopic short-interval variation of $S(T)=\frac1\pi\Arg\zeta(\tfrac12+iT)$ at the Whitney scale $L=c/\log\langle T\rangle$; informally, it forbids arbitrarily large zero clusters in windows of length $\asymp 1/\log T$.
\item Structural innovations: outer cancellation with energy bookkeeping, outer-phase $\Hilb[u']$ identity, and phase–velocity calculus with smoothed $\to$ boundary passage.
\item Where numerics enter: diagnostic tables (gated by \verb+\shownumerics+) record outward-rounded enclosures for $K_0$, the hypothesized scale-free constant $K_\xi$ (Hypothesis~\ref{hyp:Kxi-uniform}), and derived thresholds. These do not constitute proofs; they are recorded only for reproducibility/diagnostics.
\item How \textup{(P+)} would be proved under the hypothesis: the phase–velocity identity paired with CR--Green bounds yields a quantitative windowed phase control (Theorem~\ref{thm:psc-certificate-stage2}). The final conversion “certificate $\Rightarrow$ boundary wedge” is handled by an explicit wedge lemma (Lemma~\ref{lem:whitney-uniform-wedge}).
\item How RH follows \emph{once (P+) holds}: assuming additionally the Poisson representation \textup{(H)}, one has $\Re F_\Lambda\ge 0$ on $\Omega\setminus Z(\zeta)$ and hence $\Theta_\Lambda$ is Schur there; removability and the (N1)–(N2) pinch then rule out any $\zeta$--zero in $\Omega$, yielding RH (Section~\ref{sec:globalization}, Theorem~\ref{thm:globalize-main}, Corollary~\ref{cor:RH}).
\end{itemize}

\subsection*{Dependency map (load-bearing chain)}
All proofs not explicitly listed below are either auxiliary or marked \emph{diagnostic/archival} in the text.
\begin{enumerate}
\item \textbf{Phase--velocity identity + boundary passage.} Theorem~\ref{thm:phase-velocity-quant}, based on Lemmas~\ref{lem:det2-unsmoothed}, \ref{lem:xi-deriv-L1}, \ref{lem:desmooth-L1}, \ref{lem:outer-phase-HT}, and \ref{lem:pv-test-smoothed}, yields the phase–velocity identity with a positive distribution $-w'$ and explicit Poisson/atomic terms.
\item \textbf{Box-energy hypothesis (bottleneck).} Hypothesis~\ref{hyp:Kxi-uniform} supplies a scale-free Whitney box-energy constant $K_\xi$ for the $\xi$-field. Unconditionally we only prove $O(\log\langle T\rangle)$ growth on Whitney boxes (Lemma~\ref{lem:carleson-xi}).
\item \textbf{Certificate $\Rightarrow$ (P+) under the hypothesis.} Under Hypothesis~\ref{hyp:Kxi-uniform}, the CR--Green windowed phase bound (Lemma~\ref{lem:CR-green-phase} with Lemma~\ref{lem:neutralization-bookkeeping}) feeds into Theorem~\ref{thm:psc-certificate-stage2}, and the explicit wedge lemma (Lemma~\ref{lem:whitney-uniform-wedge}) yields \textup{(P+)}.
\item \textbf{Globalization and RH once (P+) holds.} Assuming \textup{(H)}, Poisson transport and Cayley yield $F_\Lambda$ Herglotz and $\Theta_\Lambda$ Schur on $\Omega\setminus Z(\zeta)$ (Corollaries~\ref{cor:poisson-herglotz} and \ref{cor:cayley-schur}). Removability across punctures uses Lemma~\ref{lem:removable-schur}. The Schur pinch then eliminates $Z(\zeta)\cap\Omega$ (Theorem~\ref{thm:globalize-main}), and Corollary~\ref{cor:RH} yields RH.
\end{enumerate}

\section{Introduction}
\paragraph{Conceptual motivation.} The Euler product for $\zeta$ separates the $k=1$ prime layer from all higher prime powers. On the right half–plane $\{\Re s>\tfrac12\}$ the diagonal prime operator $A(s)e_p:=p^{-s}e_p$ has finite Hilbert–Schmidt norm ($\sum_p p^{-2\sigma}<\infty$), so the $k\ge2$ tail is naturally encoded by the 2–modified determinant $\det_2(I-A)$. After dividing by an outer normalizer one arrives at a ratio $\mathcal J$ that shares its zero/pole geometry with $\xi$ but whose boundary modulus is unimodular. This puts the problem into the bounded–real/Schur/Herglotz framework: boundary positivity for $2\mathcal J$ transports to the interior by Poisson, and Cayley converts positivity into a Schur contractive bound for $\Theta=(2\mathcal J-1)/(2\mathcal J+1)$. The central analytic mechanism is local: a Cauchy–Riemann/Green pairing against a Poisson test on a Whitney box controls windowed boundary phase by the Dirichlet energy of $U=\Re\log\mathcal J$. The globalization/pinch step is unconditional once \textup{(P+)} holds; the remaining bottleneck is the scale-free Whitney $\xi$-energy input isolated as Hypothesis~\ref{hyp:Kxi-uniform}.
\noindent\textbf{Main result and one-route proof outline (conditional bottleneck made explicit).} The route is a boundary product–certificate mechanism in the $\zeta$–normalized gauge (no $C_P$ term). The steps are:
\begin{itemize}
\item Phase–velocity identity with outer normalization and boundary passage (Lemma~\ref{lem:zeta-normalization}).
\item Derivative envelope and the H$^1$–BMO link yielding $M_\psi$ (Lemmas~\ref{lem:CH-derivative-explicit},\ref{lem:Mpsi-correct}).
\item Box–energy control for the paired field (Lemma~\ref{lem:outer-energy-bookkeeping}), using the arithmetic tail bound (Lemma~\ref{lem:carleson-arith}) and the $\xi$-energy input (Hypothesis~\ref{hyp:Kxi-uniform}; unconditionally only Lemma~\ref{lem:carleson-xi}).
\item Boundary wedge from the certificate (Theorem~\ref{thm:psc-certificate-stage2}).
\item Globalization/pinch across $Z(\zeta)$ and conclusion (Section~\ref{sec:globalization}).
\end{itemize}
We retain two compatible RHS bounds (CR–Green + box energy, and the Hilbert envelope). Any printed numerics are diagnostic only and do not substitute for the missing scale-free $\xi$-energy input.

\paragraph{Non-circularity (active certificate).}
\begin{itemize}
\item Active RHS uses only three inputs: $c_0(\psi)$ (plateau), the CR–Green box constant $C(\psi)$, and the box-energy constant $C_{\rm box}^{(\zeta)}$.
\item Closure of \textup{(P+)} uses the explicit wedge lemma (Lemma~\ref{lem:whitney-uniform-wedge}), avoiding the earlier informal “wedge criterion” phrasing.
\item The envelope constants $C_H(\psi)$ and $M_\psi$ are auxiliary and do not enter the load-bearing inequality for (P+).
\end{itemize}

\paragraph{One-route outline (what actually happens).}
Section~\ref{thm:phase-velocity-quant} establishes the phase–velocity identity with outer normalization and boundary passage. Under the scale-free Whitney $\xi$-energy hypothesis (Hypothesis~\ref{hyp:Kxi-uniform}), the window/CR--Green machinery converts box-energy control into windowed phase control (Theorem~\ref{thm:psc-certificate-stage2}); combined with a smallness condition $\Upsilon<\tfrac12$, the wedge lemma yields \textup{(P+)} (Lemma~\ref{lem:whitney-uniform-wedge}). The globalization/pinch step (Section~\ref{sec:globalization}) is unconditional once \textup{(P+)} holds. Unconditionally, Lemma~\ref{lem:carleson-xi} provides only $O(\log\langle T\rangle)$ growth on Whitney boxes and does not supply the scale-free constant needed for the quantitative closure.
The Riemann Hypothesis (RH) admits several analytic formulations. In this paper we pursue a bounded-real (BRF) route on the right half-plane
\[
 \Omega\;:=\;\Half,
\]
which is naturally expressed in terms of Herglotz/Schur functions and passive systems. Let \(\PP\) be the primes, and define the prime-diagonal operator
\[
 A(s):\ell^2(\PP)\to\ell^2(\PP),\qquad A(s)e_p\;:=\;p^{-s}e_p.
\]
For \(\sigma:=\Re s>\tfrac12\) we have \(\|A(s)\|_{\HS}^2=\sum_{p\in\PP}p^{-2\sigma}<\infty\) and \(\|A(s)\|\le 2^{-\sigma}<1\). With the completed zeta function
\[
 \xi(s)\;:=\;\tfrac12 s(1-s)\,\pi^{-s/2}\,\Gamma(s/2)\,\zeta(s)
\]
and the Hilbert--Schmidt regularized determinant \(\dettwo\), we study the analytic function
\[
 \mathcal J(s)\;:=\;\frac{\dettwo(I-A(s))}{\mathcal O(s)\,\xi(s)},\qquad \Theta(s)\;:=\;\frac{2\mathcal J(s)-1}{2\mathcal J(s)+1}.
\]
The BRF assertion is that the Cayley transform associated to the normalized ratio is Schur on zero-free rectangles (equivalently, the ratio is Herglotz there). In the globalization step (Section~\ref{sec:globalization}) we apply this in a $\Lambda$--normalized gauge (avoiding the trivial zero of $\xi$ at $s=1$) and pinch across $Z(\zeta)\cap\Omega$.

Our method combines four ingredients:
\begin{itemize}
 \item \textbf{Schur--determinant splitting.} For a block operator \(T(s)=\begin{bmatrix}A(s)&B(s)\\ C(s)&D(s)\end{bmatrix}\) with finite auxiliary part, one has
 \[
  \log\dettwo(I-T)\;=\;\log\dettwo(I-A)\; +\; \log\det(I-S),\qquad S\;:=\;D-C(I-A)^{-1}B,
 \]
 which separates the Hilbert--Schmidt (\(k\ge 2\)) terms from the finite block.
 \item \textbf{HS continuity for \(\dettwo\).} Prime truncations \(A_N\to A\) in the HS topology, uniformly on compacts in \(\Omega\), imply local-uniform convergence of \(\dettwo(I-A_N)\) (Section~\ref{prop:hs-det2-continuity}). Division by \(\xi\) is justified only on compacts avoiding its zeros; throughout we explicitly state such hypotheses when needed.
 % optional finite-stage and interior-rectangle route removed to enforce single proof route
\end{itemize}
% interior rectangles header removed (single-route only)
% removed interior route formal chain (single-route only). Throughout, \(\Omega=\{\Re s>\tfrac12\}\), and
\subsection*{Unsmoothing det$_2$: routed through smoothed testing (A1)}
\begin{lemma}[Smoothed distributional bound for $\partial_\sigma\,\Re\log\dettwo$]\label{lem:det2-unsmoothed}
Let $I\Subset\R$ be a compact interval and fix $\varepsilon_0\in(0,\tfrac12]$. There exists a finite constant
\[
  C_*\ :=\ \sum_{p}\sum_{k\ge 2}\frac{p^{-k/2}}{k^2\,\log p}\ <\ \infty
\]
such that for all $\sigma\in(\tfrac12,\tfrac12+\varepsilon_0]$ and every $\varphi\in C_c^2(I)$,
\[
  \Big|\int_{\R} \varphi(t)\,\partial_\sigma\Re\log\dettwo\big(I-A(\sigma+it)\big)\,dt\Big|\ \le\ C_*\,\|\varphi''\|_{L^1(I)}.
\]
In particular, testing against smooth, compactly supported windows yields bounds uniform in $\sigma$.
\end{lemma}
% archived block removed
\begin{proof}
For $\sigma>\tfrac12$ one has $\sum_p |p^{-(\sigma+it)}|^2=\sum_p p^{-2\sigma}<\infty$, so the diagonal product formula for $\dettwo$ gives
\[
  \log\dettwo(I-A(s))
  \;=\;\sum_p\big(\log(1-p^{-s})+p^{-s}\big)
  \;=\;-\sum_p\sum_{k\ge 2}\frac{p^{-ks}}{k},
\]
with absolute convergence (uniform on compact subsets of $\{\Re s>\tfrac12\}$). Differentiating termwise in $\sigma=\Re s$ yields the absolutely convergent expansion
\[
  \partial_\sigma\,\Re\log\dettwo\big(I-A(\sigma+it)\big)
  \;=\; \sum_{p}\sum_{k\ge 2} (\log p)\,p^{-k\sigma}\cos(k t\log p).
\]
For each frequency $\omega=k\log p\ge 2\log 2$, two integrations by parts give
\[
  \Big|\int_{\R}\!\varphi(t)\cos(\omega t)\,dt\Big|\ \le\ \frac{\|\varphi''\|_{L^1(I)}}{\omega^2}.
\]
Since $\sum_{p,k\ge 2} (\log p)\,p^{-k\sigma}/(k\log p)^2\le C_*$ uniformly in $\sigma\in(\tfrac12,\tfrac12+\varepsilon_0]$, Tonelli/Fubini allows summing after testing against $\varphi$. Summing the resulting majorant yields
\[
  \Big|\int \varphi\,\partial_\sigma\Re\log\dettwo\,dt\Big|
  \ \le\ \|\varphi''\|_{L^1}\sum_{p}\sum_{k\ge 2}\frac{(\log p)\,p^{-k\sigma}}{(k\log p)^2}
  \ \le\ \|\varphi''\|_{L^1}\sum_{p}\sum_{k\ge 2}\frac{p^{-k/2}}{k^2\,\log p},
\]
uniformly for $\sigma\in(\tfrac12,\tfrac12+\varepsilon_0]$, since the rightmost double series converges. This proves the claim.
\end{proof}
% (Removed: the earlier local-to-global wedge lemma was not used in the active chain and its proof as written was not correct.)
\noindent\emph{Note.} The single-interval density route is archived; the small-$L$ scaling $c_0 L \le C\,L^{1/2}$ does not contradict the RHS bound.

% (Legacy KYP/Carleson stubs removed; unused in the active route.)

% Minimal in-file bibliography moved to end (References)

\begin{lemma}[De-smoothing to $L^1$ control]\label{lem:desmooth-L1}
Fix a compact interval $I\Subset\R$. Suppose a family $g_\varepsilon\in\mathcal D'(I)$ satisfies
\[
  \big|\langle g_\varepsilon,\,\phi''\rangle\big|\ \le\ C_I\,\|\phi''\|_{L^1(I)}\qquad\forall\,\phi\in C_c^\infty(I),\ \forall\,\varepsilon\in(0,\varepsilon_0].
\]
Then $g_\varepsilon$ is uniformly bounded in $W^{-2,\infty}(I)$ and there exist primitives $u_\varepsilon\in BV(I)$ with $u_\varepsilon' = g_\varepsilon$ in $\mathcal D'(I)$ such that, along a subsequence, $u_\varepsilon\to u$ in $L^1(I)$. In particular, applied to $g_\varepsilon=\partial_\sigma\Re\log\dettwo(\tfrac12+\varepsilon+it)$ together with the tested $L^1$ bound for $\partial_\sigma\Re\log\xi$, this yields the $L^1$ Cauchy property used in Proposition~\ref{prop:outer-central}.
\end{lemma}
\begin{proof}
1) Uniform $W^{-2,\infty}$ bound. Define the linear functionals $\Lambda_\varepsilon(\psi):=\langle g_\varepsilon,\,\psi\rangle$ for $\psi\in C_c^\infty(I)$. For any $\psi\in C_c^\infty(I)$ let $\Phi\in C_c^\infty(I)$ solve $\Phi''=\psi$ with zero boundary data on $I$ (obtainable by two integrations). Then $\|\Phi''\|_{L^1}=\|\psi\|_{L^1}$ and by hypothesis
\[
  |\Lambda_\varepsilon(\psi)|\ =\ |\langle g_\varepsilon,\Phi''\rangle|\ \le\ C_I\,\|\Phi''\|_{L^1}\ =\ C_I\,\|\psi\|_{L^1}.
\]
Thus $\|g_\varepsilon\|_{W^{-2,\infty}(I)}\le C_I$ uniformly in $\varepsilon$.

2) Construction of primitives and BV bound. Fix any $x_0\in I$. Let $G$ be the Green operator for $\partial_t^2$ on $I$ with homogeneous boundary data. Define $u_\varepsilon:=G[g_\varepsilon]+c_\varepsilon$, where $c_\varepsilon$ is the constant making $\int_I u_\varepsilon=0$. Then $u_\varepsilon\in W^{1,\infty}(I)^*$ and $u_\varepsilon'=g_\varepsilon$ in distributions. For $\varphi\in C_c^\infty(I)$ with $\|\varphi\|_{L^1}\le 1$,
\[
  |\langle u_\varepsilon',\varphi\rangle|\ =\ |\langle g_\varepsilon,\varphi\rangle|\ \le\ C_I,
\]
so the total variation $\mathrm{Var}_I(u_\varepsilon)\le C_I$. Together with the zero-mean choice, this yields a uniform $BV(I)$ bound on $u_\varepsilon$.

3) Compactness and $L^1$ convergence. By the compactness theorem for $BV$ (Helly selection: bounded sets in $BV(I)$ are relatively compact in $L^1(I)$; see, e.g., Ambrosio--Fusco--Pallara, \emph{Functions of Bounded Variation and Free Discontinuity Problems}, Thm.~3.23), there exists a subsequence (not relabeled) such that $u_\varepsilon\to u$ in $L^1(I)$ and pointwise a.e. on $I$. This proves the claim.
\end{proof}

\begin{lemma}[Neutralization bookkeeping for CR–Green on a Whitney box]\label{lem:neutralization-bookkeeping}
Let $I=[t_0{-}L,t_0{+}L]$ and $Q(\alpha'I)$ be as above. Let $B_I$ be the product of half–plane Blaschke factors for the zeros/poles of $J$ in $Q(\alpha'I)$ and set $\widetilde U:=\Re\log(J/B_I)$ on $Q(\alpha'I)$. Then with the same cutoff $\chi_{L,t_0}$ and Poisson test $V_{\psi,L,t_0}$,
\[
 \iint_{Q(\alpha'I)} \nabla \widetilde U\cdot\nabla(\chi V)\,dt\,d\sigma
 = \int_{\R} \psi_{L,t_0}(t)\,-w'(t)\,dt\ +\ \mathcal E_{\mathrm{side}}\ +\ \mathcal E_{\mathrm{top}},
\]
where the error terms obey the uniform bound
\[
 |\mathcal E_{\mathrm{side}}|+|\mathcal E_{\mathrm{top}}|
 \ \le\ C_{\mathrm{neu}}(\alpha,\psi)\,\Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{1/2}.
\]
In particular,
\[
  \int_{\R} \psi_{L,t_0}(-w')\ \le\ C(\psi)\,\Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{1/2},
\]
with constants independent of $t_0$ and $L$.
\end{lemma}
\begin{proof}
Let $\mathcal Z_I$ denote the (finite) set of zeros/poles of $J$ in the bounded box $Q(\alpha'I)$ (counted with multiplicity).
For $\rho=\beta+i\gamma\in\mathcal Z_I$ write $\rho^\ast:=1-\overline{\rho}$ for reflection across the boundary line $\Re s=\tfrac12$, and set the half-plane Blaschke factor
\[
  C_\rho(s)\ :=\ \frac{s-\rho^\ast}{s-\rho}\qquad(\Re \rho>\tfrac12),
\]
so that $|C_\rho(\tfrac12+it)|=1$ and, in the distributional sense on $\R$,
\[
  \frac{d}{dt}\Arg C_\rho\!\left(\tfrac12+it\right)
  \ =\ -2(\beta-\tfrac12)\,P_{\beta-1/2}(t-\gamma).
\]
Choose integers $m_\rho$ so that the finite product $B_I:=\prod_{\rho\in\mathcal Z_I} C_\rho^{\,m_\rho}$ has exactly the same divisor as $J$ in $Q(\alpha'I)$ (zeros/poles with multiplicity). Then $\widetilde J:=J/B_I$ is analytic and zero-free on a neighborhood of $\overline{Q(\alpha'I)}$, and since $|B_I|=1$ on $\Re s=\tfrac12$ it has unimodular boundary values as well. Hence $\log\widetilde J=\widetilde U+i\widetilde W$ is analytic on $Q(\alpha'I)$.

Apply Lemma~\ref{lem:CR-green-phase} to $\widetilde J$ on $Q(\alpha'I)$ with the same cutoff and test field. On the bottom edge one obtains the boundary term
\(
\int_{\R}\psi_{L,t_0}(t)\,(-\widetilde W'(t))\,dt.
\)
On $\Re s=\tfrac12$ one has $\Arg J=\Arg\widetilde J+\Arg B_I$ (mod $2\pi$), hence in $\mathcal D'(\R)$
\[
  -w'\ =\ -\widetilde W'\ -\ (\Arg B_I)'.
\]
Using the explicit formula for $(\Arg C_\rho)'$ above, $(\Arg B_I)'$ is a finite linear combination of Poisson kernels. Moreover, since $|B_I|=1$ on the bottom edge, $\Re\log B_I=0$ there and Cauchy--Riemann gives
$-(\Arg B_I)'=\partial_\sigma\Re\log B_I|_{\sigma=0}$.
Thus, when we expand $\nabla\widetilde U=\nabla U-\nabla\Re\log B_I$ in the Dirichlet pairing, the bottom-edge term coming from $\Re\log B_I$ cancels exactly against the extra $-(\Arg B_I)'$ contribution in $-w'$; only side/top pieces remain. We collect the resulting side/top contributions (together with the cutoff-transition pieces) into $\mathcal E_{\mathrm{side}}$ and $\mathcal E_{\mathrm{top}}$.

Finally, these terms are bounded by Cauchy--Schwarz exactly as in Lemma~\ref{lem:CR-green-phase}: they are controlled by the product of the Dirichlet norm of the (neutralized) potential on $Q(\alpha'I)$ and the scale-invariant Dirichlet norm of the fixed test field on the non-bottom edges. This yields the stated bound with a constant $C_{\mathrm{neu}}(\alpha,\psi)$ independent of $(t_0,L)$, and hence the displayed inequality after absorbing constants into $C(\psi)$.
\end{proof}
\noindent\emph{Clarification (Route A; conditional bottleneck).} The load–bearing input needed to conclude \textup{(P+)} from Lemma~\ref{lem:whitney-uniform-wedge} is an \emph{all-interval} phase-drop bound
\(\int_I (-w')\le \pi\,\Upsilon\) with $\Upsilon<\tfrac12$ on every bounded interval \(I\) avoiding critical-line ordinates. In the intended Route~A closure, such a scale-free bound is derived from the product-certificate machinery together with a scale-free Whitney $\xi$-energy constant (Hypothesis~\ref{hyp:Kxi-uniform}); without that hypothesis, the available unconditional $\xi$-energy bound grows like $O(\log\langle T\rangle)$ on Whitney boxes (Lemma~\ref{lem:carleson-xi}) and the quantitative $\Upsilon<\tfrac12$ regime is not established.

\begin{lemma}[Poisson lower bound $\Rightarrow$ Lebesgue a.e. wedge]\label{lem:mu-to-lebesgue}
Assume the hypotheses of Theorem~\ref{thm:phase-velocity-quant}. Fix $m\in\mathbb R/2\pi\mathbb Z$ and define
\[
 \mathcal Q := \bigl\{t\in\mathbb R:\ |\Arg \mathcal J(1/2+it)-m|\ge \tfrac{\pi}{2}\bigr\}.
\]
If $\mu(\mathcal Q)=0$, then $|\mathcal Q|=0$. In particular, \eqref{eq:Pplus} holds.
\end{lemma}
\begin{proof}
Fix $I\Subset\R$ and choose $\phi\in C_c^\infty(I)$ with $0\le\phi\le\mathbf 1_{\mathcal Q}$. By Theorem~\ref{thm:phase-velocity-quant},
\[
  \int \phi(t)\,-w'(t)\,dt \;=\; \pi\!\int\phi\,d\mu \;+\; \pi\!\sum_{\gamma\in I} m_\gamma\,\phi(\gamma).
\]
Since $\mu(\mathcal Q)=0$ and $\phi\le\mathbf 1_{\mathcal Q}$, the first term vanishes; choosing $\phi$ to vanish in small neighborhoods of each $\gamma\in I$ kills the atomic sum as well, so $\int_{\mathcal Q} (-w')=0$ on $I$. As $-w'$ is a positive boundary distribution, this forces $-w'=0$ a.e. on $\mathcal Q\cap I$. By nontangential boundary uniqueness for harmonic conjugates of $H^p_{\rm loc}$ functions\footnote{See Garnett, \emph{Bounded Analytic Functions}, Thm.~II.4.2, and Rosenblum--Rovnyak, \emph{Hardy Classes and Operator Theory}, Ch.~2.} and the definition of $\mathcal Q$, we must have $|\mathcal Q\cap I|=0$. Letting $I\uparrow\R$ yields $|\mathcal Q|=0$.
\end{proof}
% --- Appendix S moved near the end; see after main results/before bibliography ---

% Lead-in: We quantify the phase–velocity identity and justify boundary passage via outers and smoothed L^1 control.
\begin{lemma}[Outer–Hilbert boundary identity]\label{lem:outer-phase-HT}
Let $u\in L^1_{\mathrm{loc}}(\mathbb R)$ and let $O$ be the outer function on $\Omega$ with boundary modulus $|O(\tfrac12+it)|=e^{u(t)}$ a.e. Then, in $\mathcal D'(\mathbb R)$,
\[
\frac{d}{dt}\Arg O\!\left(\tfrac12+it\right)=\Hilb[u'](t),
\]
where $\Hilb$ is the boundary Hilbert transform on $\mathbb R$ and $u'$ is the distributional derivative.
\end{lemma}
\begin{proof}
See, e.g., Garnett, \emph{Bounded Analytic Functions}, Ch.~II, and Rosenblum--Rovnyak, \emph{Hardy Classes and Operator Theory}, Ch.~2, for outer factorization and boundary conjugacy (disc/half-plane via conformal mapping) and for Hilbert-transform conventions.
Write $\log O=U+iV$ on $\Omega$, where $U$ is the Poisson extension of $u$ and $V$ is its harmonic conjugate with $V(\tfrac12+\cdot)=\Hilb[u]$ in $\mathcal D'(\mathbb R)$. Then $\tfrac{d}{dt}\Arg O=\partial_t V=\Hilb[\partial_t U]=\Hilb[u']$ in distributions.
\end{proof}
\begin{theorem}[Quantified phase–velocity identity and boundary passage]\label{thm:phase-velocity-quant}
Let $u_\varepsilon(t):=\log\big|\dettwo(I-A(\tfrac12+\varepsilon+it))\big| - \log\big|\xi(\tfrac12+\varepsilon+it)\big|$ and let $\mathcal O_\varepsilon$ be the outer on $\{\Re s>\tfrac12+\varepsilon\}$ with boundary modulus $e^{u_\varepsilon}$. There exists $C_I<\infty$, independent of $\varepsilon\in(0,\varepsilon_0]$, such that for every compact interval $I\Subset\R$ and every $\phi\in C_c^2(I)$ with $\phi\ge 0$,
\[
 \Big|\int_I \phi(t)\,\partial_\sigma\Re\log\dettwo\big(I-A(\tfrac12+\varepsilon+it)\big)\,dt\Big|\ \le\ C_I\,\|\phi''\|_{L^1(I)},
\]
and
\[
 \int_I \phi(t)\,\partial_\sigma\Re\log\xi\big(\tfrac12+\varepsilon+it\big)\,dt\ \le\ C'_I\,\|\phi\|_{H^1(I)}
\]
with $C'_I$ depending only on $I$. Consequently $u_\varepsilon$ is uniformly $L^1$–bounded and Cauchy on $I$ as $\varepsilon\downarrow 0$, and the outers $\mathcal O_\varepsilon$ converge locally uniformly to an outer $\mathcal O$ on $\Omega$ with a.e. boundary modulus $e^{u}$. In particular, after dividing by $\mathcal O\,\xi$ and passing to $\varepsilon\downarrow 0$, the phase–velocity identity holds in the distributional sense on $I$:
\[
 \int_I \phi(t)\,-w'(t)\,dt\ =\ \int_I \phi(t)\,\pi\,d\mu(t)\ +\ \pi\sum_{\gamma\in I} m_\gamma\,\phi(\gamma),\qquad \forall\,\phi\in C_c^\infty(I),\ \phi\ge 0,
\]
where $\mu$ is the \emph{boundary balayage measure} on $\mathbb R$ induced by off–critical zeros (i.e. the absolutely continuous measure whose density is a sum of Poisson kernels), and the discrete sum ranges over critical–line ordinates.
\end{theorem}
\begin{proof}
Fix a compact interval $I\Subset\R$ and $\varepsilon_0\in(0,\tfrac12]$. Define
\[
 u_\varepsilon(t):=\log\Big|\dettwo\big(I-A(\tfrac12+\varepsilon+it)\big)\Big|-\log\Big|\xi(\tfrac12+\varepsilon+it)\Big|.
\]
By Lemma~\ref{lem:det2-unsmoothed}, for every $\phi\in C_c^2(I)$,
\[
 \Big|\!\int_I \!\phi(t)\,\partial_\sigma\Re\log\dettwo(I\! -\!A(\tfrac12\!+\!\sigma\!+\!it))\,dt\Big|\ \le\ C_I\,\|\phi''\|_{L^1(I)}
\]
uniformly in $\sigma\in(0,\varepsilon_0]$. For $\xi$, Lemma~\ref{lem:xi-deriv-L1} gives the tested bound
\[
 \Big|\int_I \phi(t)\,\partial_\sigma\Re\log\xi(\tfrac12+\sigma+it)\,dt\Big|\ \le\ C'_I\,\|\phi\|_{H^1(I)}\qquad(0<\sigma\le \varepsilon_0).
\]
Integrating $\sigma\in(\delta,\varepsilon)$ and using Lemma~\ref{lem:desmooth-L1} (de-smoothing) yields
\[
 \|u_\varepsilon-u_\delta\|_{L^1(I)}\ \le\ C_I''\,|\varepsilon-\delta|,\qquad 0<\delta<\varepsilon\le\varepsilon_0,
\]
for a constant $C_I''$ depending only on $I$.
Thus $\{u_\varepsilon\}$ is uniformly $L^1$–bounded and Cauchy on $I$, so $u_\varepsilon\to u$ in $L^1(I)$ for some $u\in L^1(I)$. By outer theory on the half-plane (see, e.g., Garnett, \emph{Bounded Analytic Functions}, Ch.~II, and Rosenblum--Rovnyak, \emph{Hardy Classes and Operator Theory}, Ch.~2), there exist outers $\mathcal O_\varepsilon$ with boundary modulus $e^{u_\varepsilon}$ and $\mathcal O_\varepsilon\to\mathcal O$ locally uniformly on $\Omega$, where $\mathcal O$ has boundary modulus $e^{u}$. Consequently the outer–normalized ratio $\mathcal J=\dettwo(I-A)/(\mathcal O\,\xi)$ has a.e. boundary values with $|\mathcal J|=1$ on $\Re s=\tfrac12$.

For the phase–velocity identity, factor $F=\dettwo/\xi=I\,O$ with inner $I$ and the above outer $O$. By Lemma~\ref{lem:outer-phase-HT}, the boundary argument of $O$ satisfies $\frac{d}{dt}\Arg O(\tfrac12+it)=\Hilb[u'](t)$ in $\mathcal D'(I)$. Summing the Blaschke contributions of interior poles/zeros (Lemma~\ref{lem:pv-test-smoothed}, Eq.~\eqref{eq:pv-smoothed}) gives exactly the Poisson balayage term for off–critical zeros plus atoms at critical–line ordinates, which yields the displayed identity after testing against nonnegative $\phi\in C_c^\infty(I)$. This proves the theorem.
\end{proof}

\begin{lemma}[Balayage density and consequence for $Q$]\label{lem:balayage-density}
If there exists at least one off--critical zero $\rho=\beta+i\gamma$ of $\xi$ with $\beta>\tfrac12$, then the boundary balayage measure $\mu$ from Theorem~\ref{thm:phase-velocity-quant} has an a.e. density $f\in L^1_{\mathrm{loc}}(\mathbb R)$ of the form
\[
  f(t)\ =\ \sum_{\substack{\rho=\beta+i\gamma\\ \beta>1/2}} 2\,(\beta-\tfrac12)\,P_{\beta-1/2}(t-\gamma),\qquad P_a(x)=\frac{1}{\pi}\frac{a}{a^2+x^2},
\]
which is strictly positive a.e. on $\R$ whenever at least one off--critical zero exists. Consequently, for any measurable set $E\subset\R$, $\mu(E)=0$ implies $|E|=0$. In particular, $\mu(Q)=0$ forces $|Q|=0$, hence \textup{(P+)}.
\end{lemma}
\begin{proof}
For each finite subset of zeros $\mathcal Z\subset\{\rho:\Re\rho>1/2\}$ the partial density
\(f_{\mathcal Z}(t):=\sum_{\rho\in\mathcal Z}2(\beta-\tfrac12)P_{\beta-1/2}(t-\gamma)\)
is continuous and strictly positive for all $t$ because each Poisson kernel is strictly positive on $\R$.
The phase--velocity formula and the Carleson energy finiteness imply the balayage of zeros on any Whitney box is finite, so the monotone limit of the partial sums converges in $L^1_{\mathrm{loc}}$ to an a.e. finite function $f\ge0$. Since the pointwise limit of strictly positive functions is nonnegative and cannot vanish on a set of positive measure unless all coefficients vanish, we obtain $f>0$ a.e. whenever at least one off--critical zero exists. Moreover, by positivity and monotone convergence, $\mu(E)=\int_E f\,dt=0$ forces $f=0$ a.e. on $E$, whence $|E|=0$.
\end{proof}

\paragraph{Certificate $\Rightarrow$ (P+): narrative.}
The outer, boundary phase–velocity identity shows that testing $-w'$ against a nonnegative window $\varphi$ picks up a nonnegative contribution from off–critical zeros (with an additional nonnegative atomic term at critical-line ordinates, which is harmless for \emph{a.e.} statements). The Poisson plateau lower bound turns this into a lower bound in terms of the defect measure $\nu$ on $\Omega$. Under the scale-free Whitney $\xi$-energy hypothesis (Hypothesis~\ref{hyp:Kxi-uniform}), one also has a scale-free upper bound from the Hilbert-transform pairing after affine subtraction (mean oscillation), yielding a genuine \emph{product certificate} on every interval. The final step “certificate $\Rightarrow$ (P+)” is the elementary bounded-variation wedge lemma stated below; this replaces the previous unsupported “triggers the wedge criterion” language.

\begin{lemma}[Quantitative wedge from an all-interval certificate]\label{lem:whitney-uniform-wedge}
Let $w\in BV_{\mathrm{loc}}(\mathbb R)$ and let $-w'$ denote its distributional derivative, viewed as a positive Radon measure on $\mathbb R$.
Assume there exists $\Upsilon\in(0,\tfrac12)$ such that for every bounded interval $I\subset\mathbb R$ one has the \emph{interval phase-drop bound}
\[
  \int_I (-w'(t))\,dt\ \le\ \pi\,\Upsilon.
\]
Then, after fixing the global additive constant in the argument (equivalently: choosing the unimodular constant in the outer normalization, which is free), one has
$w(t)\in[-\tfrac{\pi}{2},\tfrac{\pi}{2}]$ for a.e. $t\in\mathbb R$.
Hence for the unimodular boundary values $\mathcal J(\tfrac12+it)=e^{iw(t)}$ one has
\[
  \Re\bigl(2\mathcal J(\tfrac12+it)\bigr)\ \ge\ 0\qquad\text{for a.e. }t\in\mathbb R,
\]
i.e. \textup{(P+)} holds.
\end{lemma}
\begin{proof}
Since $-w'$ is a positive measure, $w$ admits a representative of bounded variation on each bounded interval and is (after fixing an additive constant) monotone nonincreasing there. Choose a reference point $t_\ast$ at which the boundary value $\mathcal J(\tfrac12+it_\ast)$ exists (this holds for a.e.\ $t_\ast$). Use the unimodular ambiguity of the outer normalizer to add a global constant to $w$ so that $w(t_\ast)=0$.

For any $t\in\mathbb R$, apply the phase-drop bound on the interval with endpoints $t$ and $t_\ast$ to obtain
\[
  |w(t)-w(t_\ast)|\ \le\ \int_{[t,t_\ast]} (-w')\ \le\ \pi\,\Upsilon\ <\ \frac{\pi}{2}.
\]
Thus $|w(t)|<\pi/2$ for a.e.\ $t\in\mathbb R$. Finally, $\Re(2e^{iw})=2\cos w\ge 0$ whenever $w\in[-\pi/2,\pi/2]$, proving \textup{(P+)}.
\end{proof}

\noindent\emph{Clarification.} Lemma~\ref{lem:whitney-uniform-wedge} requires an all-interval phase-drop bound \(\int_I (-w')\le \pi\,\Upsilon\) with \(\Upsilon<\tfrac12\). In this manuscript, deriving such a scale-free bound from the product-certificate machinery is conditional on Hypothesis~\ref{hyp:Kxi-uniform}; without it, only the weaker unconditional $\xi$-energy growth of Lemma~\ref{lem:carleson-xi} is available.
\paragraph{Window constant.}
Set once and for all the window constant
\[
  C(\psi)\ :=\ C_{\mathrm{rem}}(\alpha,\psi)\,\mathcal A(\psi)\ +\ C_{\mathrm{neu}}(\alpha,\psi),
\]
where $\mathcal A(\psi)$ is the fixed Poisson energy of the window, $C_{\mathrm{rem}}(\alpha,\psi)$ is the side/top remainder factor from Corollary~\ref{cor:CH-Mpsi-final}, and $C_{\mathrm{neu}}(\alpha,\psi)$ is the neutralization bookkeeping constant from Lemma~\ref{lem:neutralization-bookkeeping}. Then $C(\psi)$ is independent of $L$ and $t_0$ and will be used uniformly below.
\begin{proposition}[HS$\to$det$_2$ continuity]\label{prop:hs-det2-continuity}
Let $A_N,A$ be analytic $\HS$-valued maps on $\Omega$ with $A_N\to A$ in the Hilbert–Schmidt norm uniformly on compact subsets of $\Omega$. Then $\det\nolimits_2(I-A_N)\to\det\nolimits_2(I-A)$ locally uniformly on $\Omega$.
\end{proposition}
\begin{proof}
Fix a compact $K\Subset\Omega$. By hypothesis, $\sup_{s\in K}\|A_N(s)-A(s)\|_{\HS}\to0$, and in particular $\sup_{N}\sup_{s\in K}\|A_N(s)\|_{\HS}<\infty$.
We use the standard definition of the $2$--modified determinant on $\HS$:
\[
  \det\nolimits_2(I-T)\ :=\ \det\!\big((I-T)e^T\big),
\]
where the Fredholm determinant on the right is defined for trace-class perturbations of the identity. Indeed, for $T\in\HS$ one has
\[
  (I-T)e^T-I\ =\ -\sum_{n\ge 2}\frac{n-1}{n!}\,T^n,
\]
and the series converges absolutely in trace norm because $T^n$ is trace class for $n\ge 2$ and
$\|T^n\|_1\le \|T\|^{n-2}\|T^2\|_1\le \|T\|_{\HS}^n$.
In particular, on any $\HS$-ball $\{ \|T\|_{\HS}\le M\}$, the map
\[
  T\ \longmapsto\ (I-T)e^T-I
\]
is Lipschitz from $\HS$ to trace class: writing the series termwise and using
$T^n-S^n=\sum_{k=0}^{n-1}T^k(T-S)S^{n-1-k}$ together with $\|XY\|_1\le \|X\|_2\|Y\|_2$ and $\|T\|\le \|T\|_{\HS}$ gives
\[
  \|(I-T)e^T-(I-S)e^S\|_1\ \le\ C(M)\,\|T-S\|_{\HS}.
\]
Since the Fredholm determinant on trace-class perturbations of the identity is defined by an absolutely convergent trace-norm series (hence is continuous in $\|\cdot\|_1$), it follows that $\det_2(I-T)$ is continuous (indeed locally Lipschitz) with respect to $\|\cdot\|_{\HS}$.
Thus
\[
  \sup_{s\in K}\Big|\det\nolimits_2\big(I-A_N(s)\big)-\det\nolimits_2\big(I-A(s)\big)\Big|\ \longrightarrow\ 0,
\]
which is local-uniform convergence on $K$. Since $K$ was arbitrary, the convergence is locally uniform on $\Omega$.
\end{proof}

\begin{lemma}[Smoothed phase–velocity calculus]\label{lem:pv-test-smoothed}
Fix $\varepsilon\in(0,\tfrac12]$ and set
\[
 u_\varepsilon(t):=\log\Big|\dettwo(I{-}A(\tfrac12{+}\varepsilon{+}it))\Big|-\log\Big|\xi(\tfrac12{+}\varepsilon{+}it)\Big|.
\]
Let $\mathcal O_\varepsilon$ be the outer on $\{\Re s>\tfrac12{+}\varepsilon\}$ with boundary modulus $e^{u_\varepsilon}$ and write $F_\varepsilon:=\dettwo/\xi$ and $O_\varepsilon:=\mathcal O_\varepsilon$. Then for every $\phi\in C_c^\infty(\R)$,
\begin{equation}\label{eq:pv-smoothed}
\int_\R\!\phi(t)\,\Big( \Im\frac{\xi'}{\xi}-\Im\frac{\dettwo'}{\dettwo}+\Hilb[u_\varepsilon']\Big)\!(\tfrac12{+}\varepsilon{+}it)\,dt
\;=\;\sum_{\substack{\rho=\beta+i\gamma\\ \Re\rho>\tfrac12}}\! 2(\beta{-}\tfrac12)\,\big(P_{\beta-\tfrac12-\varepsilon}\!\ast\phi\big)(\gamma)
\end{equation}
where $P_a(x)=\frac{1}{\pi}\frac{a}{a^2+x^2}$ and the right-hand side is a nonnegative quantity. As $\varepsilon\downarrow 0$, the kernels $P_{\beta-\tfrac12-\varepsilon}$ converge in $\mathcal D'(\R)$ to $P_{\beta-\tfrac12}$, and the boundary atoms from critical-line zeros $\{\xi(\tfrac12+i\gamma)=0\}$ appear as $\pi m_\gamma\,\phi(\gamma)$, yielding Theorem~\ref{thm:phase-velocity-quant}.
\end{lemma}
\begin{proof}
Factor $F_\varepsilon=I_\varepsilon\,O_\varepsilon$ with $O_\varepsilon$ outer on $\{\Re s>\tfrac12{+}\varepsilon\}$ and $I_\varepsilon$ inner (product of half-plane Blaschke factors for poles/zeros of $F_\varepsilon$ in the open half-plane). By Lemma~\ref{lem:outer-phase-HT}, on the boundary line $\Re s=\tfrac12{+}\varepsilon$ one has $\frac{d}{dt}\Arg O_\varepsilon=\Hilb[u_\varepsilon']$ in $\mathcal D'(\R)$. Each pole of $F_\varepsilon$ at $\rho=\beta+i\gamma$ with $\beta>\tfrac12$ contributes the half-plane Blaschke factor $C_\rho(s)=(s-\overline\rho)/(s-\rho)$ whose boundary phase derivative equals $-2(\beta-\tfrac12-\varepsilon)\,P_{\beta-\tfrac12-\varepsilon}(t-\gamma)$. Summing these contributions and writing $\frac{d}{dt}\Arg F_\varepsilon=\Im(F_\varepsilon'/F_\varepsilon)=\Im(\dettwo'/\dettwo)-\Im(\xi'/\xi)$ yields \eqref{eq:pv-smoothed} after testing against $\phi$.
Passage $\varepsilon\downarrow 0$ follows from the smoothed bounds and de-smoothing: $u_\varepsilon\to u$ in $L^1_{\rm loc}$ (Lemmas~\ref{lem:det2-unsmoothed},~\ref{lem:xi-deriv-L1} and Lemma~\ref{lem:desmooth-L1}), hence $\Hilb[u_\varepsilon']\to \Hilb[u']$ in $\mathcal D'(\R)$ by continuity of the Hilbert transform on tempered distributions (see, e.g., Stein, \emph{Singular Integrals and Differentiability Properties of Functions}, Ch.~II). The Poisson kernels converge in distributions (approximate identity; standard), and boundary atoms (critical-line zeros of $\xi$) appear in the limit as $\varepsilon\downarrow 0$ through the argument jump, giving the claimed atomic terms in Theorem~\ref{thm:phase-velocity-quant}.
\end{proof}
% Lead-in: From (P+) we lift positivity to the interior (Poisson), pass to Schur via Cayley, and pinch across $Z(\zeta)$ using (N1)–(N2).
\section{Globalization across $Z(\zeta)$ via a Schur--Herglotz pinch}\label{sec:globalization}
\noindent This section upgrades an a.e.\ boundary wedge \emph{(P+)} to an interior Herglotz/Schur conclusion on $\Omega\setminus Z(\zeta)$ via a Poisson representation hypothesis, then removes singularities across $Z(\zeta)$ using non-cancellation \textnormal{(N2)} and the right-edge normalization \textnormal{(N1)}.
\paragraph{Globalization and pinch: narrative.}
Assuming \textup{(P+)} and the Poisson representation hypothesis \textup{(H)} stated below, one has $\Re F_\Lambda\ge 0$ on $\Omega\setminus Z(\zeta)$, hence the Cayley transform $\Theta_\Lambda=(F_\Lambda-1)/(F_\Lambda+1)$ is Schur there. If an off-critical zero $\rho$ of $\zeta$ existed, the Schur bound and the chosen normalizations would force $\Theta_\Lambda$ to remain bounded and holomorphic across $\rho$ (removability), contradicting the limiting boundary value $\Theta_\Lambda(\sigma+it)\to-1$ as $\sigma\to+\infty$. Thus no such $\rho$ exists.
\noindent\textbf{Standing setup.}
Let
\[
\Omega:=\{s\in\mathbb C:\ \Re s>\tfrac12\},\qquad
\Lambda(s):=\pi^{-s/2}\Gamma\!\big(\tfrac s2\big)\zeta(s).
\]
Define
\[
\mathcal J_\Lambda(s):=\frac{\det\nolimits_2(I-A(s))}{\mathcal O(s)\,\Lambda(s)},\qquad
F_\Lambda(s):=2\,\mathcal J_\Lambda(s),\qquad
\Theta_\Lambda(s):=\frac{F_\Lambda(s)-1}{F_\Lambda(s)+1}.
\]
Here $\mathcal O$ is holomorphic and zero--free on $\Omega$ (an outer normalizer) and
$\det\nolimits_2(I-A)$ is holomorphic on $\Omega$.
We record the two normalization properties actually used below:
\begin{itemize}
\item[(N1)] (\emph{Right--edge normalization}) For each fixed $t$ (indeed uniformly on compact $t$–intervals), $\displaystyle\lim_{\sigma\to+\infty}\mathcal J_\Lambda(\sigma+it)=0$; hence $\displaystyle\lim_{\sigma\to+\infty}\Theta_\Lambda(\sigma+it)=-1$.
\item[(N2)] (\emph{Non--cancellation at $\zeta$--zeros}) For every $\rho\in\Omega$ with $\zeta(\rho)=0$,
\[
\det\nolimits_2(I-A(\rho))\neq0\quad\text{and}\quad \mathcal O(\rho)\neq0.
\]
Thus $\mathcal J_\Lambda$ has a pole at $\rho$ of order $\operatorname{ord}_\rho(\zeta)$.
\end{itemize}

\noindent\textbf{Boundary wedge $(\mathrm{P}^+)$.}
We assume the a.e.\ boundary inequality
\[
\Re\,F_\Lambda\!\left(\tfrac12+it\right)\ \ge\ 0\qquad\text{for a.e.\ }t\in\mathbb R.
\tag{P+}
\label{eq:Pplus}
\]
\medskip
\noindent\textbf{Poisson representation hypothesis (H).}
We assume that for every $\sigma>0$ and every $t_0\in\R$ with $\zeta(\tfrac12+\sigma+it_0)\neq 0$,
\[
  \Re F_\Lambda\!\left(\tfrac12+\sigma+it_0\right)\ =\ \int_{\R}\Re F_\Lambda\!\left(\tfrac12+it\right)\,P_\sigma(t-t_0)\,dt.
  \tag{H}
\]
\medskip
\noindent\textbf{From boundary wedge to interior Schur bound (half--plane Poisson passage).}
Fix $(\,\tfrac12+\sigma+it_0\,)\in\Omega\setminus Z(\zeta)$ with $\sigma>0$.
By \eqref{eq:Pplus}, the boundary trace $u(t):=\Re F_\Lambda(\tfrac12+it)$ satisfies $u(t)\ge 0$ for a.e.\ $t\in\mathbb R$.
By \textup{(H)}, the Poisson representation on the half--plane yields
\[
\Re F_\Lambda\big(\tfrac12+\sigma+it_0\big)
\;=\;\int_{\mathbb R} u(t)\,P_\sigma(t-t_0)\,dt\ \ge\ 0,
\]
so $\Re F_\Lambda\ge 0$ on $\Omega\setminus Z(\zeta)$.
In particular, on any rectangle $R\Subset\Omega$ with $\zeta\neq 0$ near $\overline R$, we have $\Re F_\Lambda\ge 0$ on $R$.
Consequently, on $R$ the identity
\[
1-|\Theta_\Lambda(s)|^2\;=\;\frac{4\,\Re F_\Lambda(s)}{|F_\Lambda(s)+1|^2}\ \ge\ 0
\]
implies
\[
|\Theta_\Lambda(s)|\ \le\ 1\qquad(s\in R).
\tag{Schur}
\label{eq:SchurBound}
\]
\noindent(Thus, prior to removability, the Schur bound holds only on $\Omega\setminus Z(\zeta)$.)
\medskip
\noindent\textbf{Local pinch at a putative off--critical zero.}
\emph{We use (N2) for non--cancellation at $\zeta$--zeros and (N1) for the right--edge limit $\Theta\to-1$.}
Fix $\rho\in\Omega$ with $\zeta(\rho)=0$.
By (N2) the function $F_\Lambda$ has a pole at $\rho$, hence
\[
\Theta_\Lambda(s)=\frac{F_\Lambda(s)-1}{F_\Lambda(s)+1}\ \longrightarrow\ 1\qquad(s\to\rho).
\]
By \eqref{eq:SchurBound}, $\Theta_\Lambda$ is bounded by $1$ on $(\Omega\setminus Z(\zeta))$,
so the singularity of $\Theta_\Lambda$ at $\rho$ is removable (Riemann's theorem), and the holomorphic extension satisfies
\[
\Theta_\Lambda(\rho)=1.
\]
Because $\Theta_\Lambda$ is holomorphic on the connected domain $\Omega\setminus(Z(\zeta)\setminus\{\rho\})$
and $|\Theta_\Lambda|\le1$ there, the Maximum Modulus Principle forces $\Theta_\Lambda$ to be
a \emph{constant unimodular} function on that domain (it attains its supremum $1$ at an interior point).
By analyticity, the same constant extends throughout $\Omega\setminus Z(\zeta)$.
\medskip
\begin{lemma}[Connectedness and isolation]\label{rem:connectedness}
Since $Z(\zeta)\cap\Omega$ is a discrete subset (zeros are isolated), one can choose a disc $D\subset\Omega$ centered at $\rho$ containing no other zeros, and $\Omega\setminus Z(\zeta)$ is (path-)connected. Hence in the argument above, $\Omega\setminus\big(Z(\zeta)\setminus\{\rho\}\big)$ is connected and the Maximum Modulus Principle applies on this domain.
\end{lemma}
\begin{proof}
Since $\zeta$ is holomorphic and not identically zero on $\Omega$, its zeros are isolated; thus $Z(\zeta)\cap\Omega$ is discrete and we may choose a disc $D\subset\Omega$ around $\rho$ containing no other zeros.
For connectedness: given $z_0,z_1\in \Omega\setminus Z(\zeta)$, join them by a polygonal path in $\Omega$. A compact polygonal path meets only finitely many points of the discrete set $Z(\zeta)\cap\Omega$, so we can locally perturb the path in small discs around those points to avoid them. This produces a path in $\Omega\setminus Z(\zeta)$, hence $\Omega\setminus Z(\zeta)$ is path-connected. The same argument applies to $\Omega\setminus(Z(\zeta)\setminus\{\rho\})$.
\end{proof}
\noindent\textbf{Contradiction with right--edge normalization.}
By (N1), $\Theta_\Lambda(\sigma+it)\to-1$ as $\sigma\to+\infty$; hence the above constant must equal $-1$.
But we also have $\Theta_\Lambda(\rho)=1$. Contradiction.
\medskip
\noindent\textbf{Conclusion of the pinch.}
No $\rho\in\Omega$ with $\zeta(\rho)=0$ can exist.
\medskip
\noindent\textbf{Connective summary (globalization/pinch).}
Assuming \textup{(P+)} and \textup{(H)}, the Poisson passage yields $\Re F_\Lambda\ge 0$ on $\Omega\setminus Z(\zeta)$ and hence $|\Theta_\Lambda|\le 1$ there. The pinch argument with \textup{(N1)}--\textup{(N2)} then rules out any $\rho\in\Omega$ with $\zeta(\rho)=0$, so $\zeta$ has no zeros in $\Re s>\tfrac12$ and RH follows.
\medskip
\noindent\textbf{Normalization at infinity (used in (N1)).}
We record explicit bounds ensuring $\Theta_\Lambda(\sigma+it)\to-1$ uniformly for $t$ in compact $t$-intervals as $\sigma\to+\infty$.
\begin{itemize}
\item Zeta/gamma growth: For $\sigma\ge 2$ and all $t\in\R$, $|\zeta(\sigma+it)-1|\le 2^{1-\sigma}$, hence $|\zeta(\sigma+it)|\le 1+2^{1-\sigma}$. Stirling's formula on vertical strips gives $|\pi^{-s/2}\Gamma(s/2)|\asymp (1+|t|)^{\sigma/2-1/2} e^{-\pi|t|/4}$. For each fixed $t$ (indeed uniformly on compact $t$-intervals), $|\Lambda(\sigma+it)|\to\infty$ as $\sigma\to\infty$.
\item Outer factor: By the Carleson embedding inequality (Appendix~\ref{app:CE-constant}) and Lemma~\ref{lem:poisson-bmo-strip}, the boundary modulus $u=\log|\dettwo/\Lambda|$ has uniform BMO control; thus its Poisson extension $U=\Re\log\mathcal O$ is bounded on vertical strips $\{\Re s\ge 1\}$ by a constant $C_\mathcal O$, yielding $e^{-C_\mathcal O}\le |\mathcal O(\sigma+it)|\le e^{C_\mathcal O}$ for $\sigma\ge 1$.
\item Det$_2$ limit: For $\sigma\ge 1$, $\|A(\sigma+it)\|\le 2^{-\sigma}\le \tfrac12$. By the product representation in Lemma~\ref{lem:hs-diagonal} and since $\sum_p p^{-2\sigma}\to0$ as $\sigma\to\infty$, one has $|\dettwo(I-A(\sigma+it)) - 1|\le C\sum_p p^{-2\sigma}\to 0$ (uniformly for $t$ in compact intervals).
\end{itemize}
Combining, for $\sigma\ge 2$,
\[
  \big|\mathcal J_\Lambda(\sigma+it)\big|\ =\ \Big|\frac{\dettwo(I-A(\sigma+it))}{\mathcal O(\sigma+it)\,\Lambda(\sigma+it)}\Big|\ \le\ \frac{1+ o(1)}{e^{-C_\mathcal O}\,|\Lambda(\sigma+it)|}\ \xrightarrow[\sigma\to\infty]{}\ 0
\]
uniformly for $t$ in compact intervals. Hence $\Theta_\Lambda(\sigma+it)=(2\mathcal J_\Lambda-1)/(2\mathcal J_\Lambda+1)\to -1$ uniformly for $t$ in compact intervals.

\bigskip
\begin{theorem}[Riemann Hypothesis (conditional globalization)]\label{thm:RH}
Under \textnormal{(P+)}, \textnormal{(H)}, and \textnormal{(N1)}--\textnormal{(N2)}, one has $\zeta(s)\neq 0$ for all $s\in\Omega$. Hence all nontrivial zeros of $\zeta$ lie on $\Re s=\tfrac12$.
\end{theorem}
\begin{proof}
The pinch argument above shows that there are no zeros of $\zeta$ in $\Omega$. The functional equation and symmetry then force all nontrivial zeros onto $\Re s=\tfrac12$ (Corollary~\ref{cor:RH}).
\end{proof}

\begin{lemma}[Carleson box energy: stable sum bound]\label{lem:carleson-sum}
For harmonic potentials $U_1,U_2$ on $\Omega$, one has
\[
\sqrt{C_{\mathrm{box}}(U_1+U_2)}\ \le\
\sqrt{C_{\mathrm{box}}(U_1)}\ +\ \sqrt{C_{\mathrm{box}}(U_2)}.
\]
\end{lemma}
\begin{proof}
Write $\mu_j:=|\nabla U_j|^2\,\sigma\,dt\,d\sigma$ and $\mu_{12}:=|\nabla(U_1{+}U_2)|^2\,\sigma\,dt\,d\sigma$. For any Carleson box $B$, by Cauchy–Schwarz,
\[
\int_{B} |\nabla(U_1+U_2)|^2\,\sigma\,dt\,d\sigma
\ \le\ \Big(\sqrt{\int_B |\nabla U_1|^2\,\sigma}\ +\ \sqrt{\int_B |\nabla U_2|^2\,\sigma}\Big)^{\!2}.
\]
Taking supremum over Carleson boxes $B$ and dividing by $|I_B|$ yields the claimed inequality.
\end{proof}

% NOTE: Lemma~\ref{lem:carleson-xi} is unconditional but not scale-free; we do not claim an all-interval Carleson constant independent of $T$ without an additional hypothesis (see Hypothesis~\ref{hyp:Kxi-uniform}).

\begin{lemma}[L$^1$-tested control for $\partial_\sigma\Re\log\xi$]\label{lem:xi-deriv-L1}
For each compact $I\Subset\R$ there exists $C'_I<\infty$ such that for all $0<\sigma\le\varepsilon_0$ and all $\phi\in C_c^2(I)$,
\[
\Big|\int_I \phi(t)\,\partial_\sigma\Re\log\xi\!\big(\tfrac12+\sigma+it\big)\,dt\Big|
\ \le\ C'_I\,\|\phi\|_{H^1(I)}.
\]
\end{lemma}

\begin{proof}[Proof of Lemma~\ref{lem:xi-deriv-L1}]
Let $I\Subset\R$ and $\phi\in C_c^2(I)$. Let $V$ be the Poisson extension of $\phi$ on a fixed dilation $Q(\alpha I)$. Green's identity together with Cauchy–Riemann for $U_\xi=\Re\log\xi$ gives
\[
  \int_I \phi(t)\,\partial_\sigma\Re\log\xi\!\big(\tfrac12+\sigma+it\big)\,dt
  \,=\, \iint_{Q(\alpha I)} \nabla U_\xi\cdot\nabla V\,dt\,d\sigma.
\]
By Cauchy–Schwarz and the scale–invariant bound $\|\nabla V\|_{L^2(\sigma;Q(\alpha I))}\lesssim \|\phi\|_{H^1(I)}$, we get
\[
  \Big|\int_I \phi\,\partial_\sigma\Re\log\xi\Big|
  \,\le\ \Big(\iint_{Q(\alpha I)}|\nabla U_\xi|^2\,\sigma\Big)^{\!1/2}\,C_{I,\alpha}^{(V)}\,\|\phi\|_{H^1(I)}.
\]
Since $I$ is compact, $\log\langle t\rangle$ is bounded on $t\in I$, so Lemma~\ref{lem:carleson-xi} gives a finite (interval-dependent) energy bound
\(
\iint_{Q(\alpha I)}|\nabla U_\xi|^2\,\sigma \le C_{I,\alpha}^{(\xi)}\,|I|
\)
for some $C_{I,\alpha}^{(\xi)}<\infty$ depending only on $I$ and $\alpha$. Substituting into the displayed Cauchy--Schwarz inequality yields the claimed estimate with (for example)
\(
C'_I:=C_{I,\alpha}^{(V)}\,\big(C_{I,\alpha}^{(\xi)}\,|I|\big)^{1/2}.
\)
\end{proof}
\begin{corollary}[Diagnostic closure inequality (conditional on a scale-free $K_\xi$)]\label{cor:conservative-closure}
Assume Hypothesis~\ref{hyp:Kxi-uniform} (so $K_\xi$ denotes a \emph{scale-free} Whitney box-energy constant for the $\xi$-field at the fixed aperture/schedule).
Then with the (unconditional) arithmetic tail constant $K_0$ (Lemma~\ref{lem:carleson-arith}) and the window constant $C_\psi^{(H^1)}$ (Appendix~\ref{app:Cpsi-compute}), one has
\[
\sqrt{C_{\mathrm{box}}}\ \le\ \sqrt{K_0}+\sqrt{K_\xi},\qquad
M_\psi\ \le\ \frac{4}{\pi}\,C_\psi^{(H^1)}\,\sqrt{C_{\mathrm{box}}}.
\]
Consequently the product-certificate parameter
\[
  \Upsilon\ :=\ \frac{(2/\pi)\,M_\psi}{c_0(\psi)}
\]
obeys an explicit bound in terms of $(K_0,K_\xi,C_\psi^{(H^1)},c_0(\psi))$. Any numerical instantiation is therefore conditional on supplying a value/enclosure for the scale-free constant $K_\xi$.

\medskip
\noindent\textbf{Diagnostics (conditional).} Using the coarse box constant \(C_{\rm box}=K_0+K_\xi\) one obtains
\[ M_\psi\ \le\ \frac{4}{\pi} C_\psi^{(H^1)}\sqrt{C_{\rm box}},\qquad \Upsilon=\frac{(2/\pi)\,M_\psi}{c_0(\psi)}, \]
with \(c_0(\psi)=0.17620819\), \(C_\psi^{(H^1)}=\CpsiHone\), we obtain
\[ M_\psi\ \le\ \Mpsilocked,\qquad \Upsilon_{\mathrm{diag}}\ :=\ \frac{(2/\pi)\,M_\psi}{c_0(\psi)}\ =\ \UpsilonLocked. \]
This display is \emph{not} an unconditional proof of $\Upsilon<\tfrac12$: it becomes meaningful only after supplying a scale-free $K_\xi$ (Hypothesis~\ref{hyp:Kxi-uniform}).
\end{corollary}
\begin{proof}
The inequality $\sqrt{C_{\mathrm{box}}}\le \sqrt{K_0}+\sqrt{K_\xi}$ is Lemma~\ref{lem:carleson-sum} applied to the decomposition of the paired potential into the arithmetic tail and the $\xi$-part (cf.\ Lemma~\ref{lem:outer-energy-bookkeeping}). The bound on $M_\psi$ follows from the $H^1$--BMO/Carleson embedding estimate (Lemma~\ref{lem:Mpsi-correct}) together with the embedding normalization $C_{\mathrm{CE}}(\alpha)=1$ (Lemma~\ref{lem:CE-constant-one}). The window constants $c_0(\psi)$ and $C_\psi^{(H^1)}$ are reproducibly enclosed in Appendix~\ref{app:Cpsi-compute}; any numerical value used for $K_\xi$ is conditional on Hypothesis~\ref{hyp:Kxi-uniform}.
\end{proof}
% \fi
\medskip
\noindent\textbf{Proof of (N2) (non--cancellation at $\zeta$--zeros).}
For $s=\sigma+it$ with $\sigma>\tfrac12$, define the diagonal operator $A(s)e_p=p^{-s}e_p$ on $\ell^2(\mathbb P)$. Then $\|A(s)\|=2^{-\sigma}<1$ and $\|A(s)\|_{\mathrm{HS}}^2=\sum_{p}p^{-2\sigma}<\infty$, so $A(s)$ is Hilbert--Schmidt. The 2--modified determinant for diagonal $A(s)$ is
\[
\det\nolimits_2\!\big(I-A(s)\big)\;=\;\prod_{p\in\mathbb P}(1-p^{-s})\,e^{p^{-s}},
\]
which converges absolutely and is nonzero because each factor is nonzero. Moreover, $I-A(s)$ is invertible with $\|(I-A(s))^{-1}\|\le (1-2^{-\sigma})^{-1}$ since $|1-p^{-s}|\ge 1-2^{-\sigma}>0$. Finally, the outer normalizer has the form $\mathcal O(s)=\exp H(s)$ with $H$ analytic on $\Omega$, hence $\mathcal O$ is zero--free on $\Omega$. Thus if $\rho\in\Omega$ with $\zeta(\rho)=0$, then $\det_2(I-A(\rho))\neq0$ and $\mathcal O(\rho)\neq0$, i.e. no cancellation can occur at $\rho$. Local-uniform analyticity on $\Omega$ follows from HS$\to\dettwo$ continuity (Proposition~\ref{prop:hs-det2-continuity}).
\begin{lemma}[Diagonal HS determinant is analytic and nonzero]\label{lem:hs-diagonal}
For $s=\sigma+it$ with $\sigma>\tfrac12$, the diagonal operator $A(s)e_p=p^{-s}e_p$ satisfies
\[
\sup_{p}|p^{-s}|=2^{-\sigma}<1,\qquad \sum_{p}|p^{-s}|^2=\sum_{p}p^{-2\sigma}<\infty.
\]
Hence $A(s)\in\mathrm{HS}$, $I-A(s)$ is invertible, and
\[
\det\nolimits_2\big(I-A(s)\big)=\prod_{p}(1-p^{-s})\,e^{p^{-s}}
\]
is analytic and nonzero on $\{\Re s>\tfrac12\}$.
\end{lemma}
\begin{proof}
Immediate from the displayed bounds; invertibility follows since $|1-p^{-s}|\ge 1-2^{-\sigma}>0$, and the product defining $\det_2$ converges absolutely with nonzero factors.
\end{proof}
\paragraph{Normalization and finite port (eliminating $C_P$ and $C_\Gamma$).}
We record the implementation details that ensure the product certificate contains no prime budget and no Archimedean term.

\begin{lemma}[\(\zeta\)–normalized outer and compensator]\label{lem:zeta-normalization}
Define the outer $\mathcal O_\zeta$ on $\Omega$ with boundary modulus $\big|\dettwo(I-A)/\zeta\big|$ and set
\[ J_\zeta(s)\ :=\ \frac{\dettwo(I-A(s))}{\mathcal O_\zeta(s)\,\zeta(s)}\cdot B(s),\qquad B(s):=\frac{s-1}{s}. \]
On $\Re s=\tfrac12$ one has $|B|=1$. The phase–velocity identity of Theorem~\ref{thm:phase-velocity-quant} holds for $J_\zeta$ with the same Poisson/zero right-hand side. In particular, no separate Archimedean term enters the inequality used by the certificate.
\end{lemma}

\begin{proof}
Set $X:=\xi$ and $Z:=\zeta$, and let $G$ denote the archimedean factor linking them,
\[
  X(s)\;=\;\tfrac12 s(1{-}s)\,\pi^{-s/2}\,\Gamma(\tfrac s2)\,Z(s)\;=:\;G(s)\,Z(s).
\]
Define $\mathcal O_X$ (resp. $\mathcal O_Z$) to be the outer on $\Omega$ with boundary modulus $\big|\dettwo(I{-}A)/X\big|$ (resp. $\big|\dettwo(I{-}A)/Z\big|$). Then, by construction,
\[
  \Big|\frac{\dettwo(I{-}A)}{\mathcal O_X\,X}\Big|\equiv 1\equiv \Big|\frac{\dettwo(I{-}A)}{\mathcal O_Z\,Z}\Big|\quad \text{a.e. on }\{\Re s=\tfrac12\}.
\]
Consequently the phase–velocity identity (Theorem~\ref{thm:phase-velocity-quant}) applies to either unimodular ratio. Writing
\[
  \log \frac{\dettwo(I{-}A)}{\mathcal O_X\,X}
  \;=\; \log \frac{\dettwo(I{-}A)}{\mathcal O_Z\,Z}\; -\; \log\frac{\mathcal O_X}{\mathcal O_Z}\; -\; \log G,
\]
and differentiating in $\sigma$ on the boundary, the two outer terms contribute zero to the boundary phase derivative (by unimodularity and the outer/Poisson representation). The remaining difference is $-\partial_\sigma\Im\log G$.

On $\Re s=\tfrac12$ we have $|O_X/O_Z|=|Z/X|=|1/G|$, so by Lemma~\ref{lem:outer-phase-HT}
\[
  \partial_\sigma\Im\log\!\left(\frac{O_X}{O_Z}\right)(\tfrac12+it)\;=\;-\partial_\sigma\Im\log G(\tfrac12+it)
\]
in $\mathcal D'(\mathbb R)$. Compensating the simple zero at $s=1$ by the half–plane Blaschke factor
\[
  B(s)\;=\;\frac{s-1}{s}\qquad(|B|\equiv 1\text{ on }\Re s=\tfrac12)
\]
accounts for the inner contribution at $s=1$. Therefore, on the boundary,
\[
  \partial_\sigma\Im\log\!\Big(\frac{\dettwo(I{-}A)}{\mathcal O_Z\,Z}\cdot B\Big)
  \,=\, \partial_\sigma\Im\log\frac{\dettwo(I{-}A)}{\mathcal O_X\,X},
\]
and the quantitative phase–velocity identity holds in the same form for $J_\zeta=(\dettwo/(\mathcal O_\zeta\,\zeta))\,B$ as for $\mathcal J=\dettwo/(\mathcal O\,\xi)$. In particular, no Archimedean term enters the certificate.
\end{proof}

% (archived) A standalone prime-layer outer O_p is not used in the main chain; the ζ-normalized gauge and windowed identities suffice, and no C_P term enters the certificate.

\begin{corollary}[No $C_P$/$C_\Gamma$ in the certificate]
With $J_\zeta$ and $\widehat J$ as above, the active CR–Green route uses $c_0(\psi)$ and the CR–Green constant $C(\psi)$ together with the box–energy constant $C_{\rm box}^{(\zeta)}$. In particular, $C_P=0$ and $C_\Gamma=0$ on the RHS; $C_H(\psi)$ and $M_\psi$ are retained only as auxiliary/readability bounds.
\end{corollary}
\begin{proof}
By construction of the $\zeta$--normalized gauge and the compensator $B$ (Lemma~\ref{lem:zeta-normalization}), the Archimedean factor contributes no boundary phase term and the simple pole/zero bookkeeping at $s=1$ is absorbed into $B$ with $|B|=1$ on $\Re s=\tfrac12$. Thus the product certificate has no $C_\Gamma$ term and no separate prime-budget term $C_P$ on the right-hand side; the remaining inputs are $c_0(\psi)$, the CR--Green constant $C(\psi)$, and the box-energy constant $C_{\rm box}^{(\zeta)}$.
\end{proof}

\noindent\emph{Active route.} Throughout we use the $\zeta$-normalized boundary gauge with the Blaschke compensator; the product-certificate mechanism uses $c_0(\psi)$ and the CR–Green constant $C(\psi)$ together with a \emph{scale-free} box-energy constant $C_{\rm box}^{(\zeta)}$ (no $C_P$, no $C_\Gamma$). In the present manuscript, the existence of such a scale-free $\xi$-energy constant (hence $C_{\rm box}^{(\zeta)}$) is isolated as Hypothesis~\ref{hyp:Kxi-uniform}; under that hypothesis one can impose the smallness condition $\Upsilon<\tfrac12$ and then \textup{(P+)} follows by Lemma~\ref{lem:whitney-uniform-wedge}.

% (Removed alternative interior-pole lemma to keep a single contradiction path in the pinch.)

\begin{lemma}[Derivative envelope for the printed window]\label{lem:CH-derivative-explicit}
Let $\psi$ be the even $C^\infty$ flat--top window from the "Printed window" paragraph (equal to $1$ on $[-1,1]$, supported in $[-2,2]$, with monotone ramps on $[-2,-1]$ and $[1,2]$), and $\varphi_L(t):=L^{-1}\psi((t-T)/L)$. Then, for every $L>0$,
\[
  \big\|\big(\mathcal H[\varphi_L]\big)'\big\|_{L^\infty(\mathbb R)} \;\le\; \frac{C_H(\psi)}{L}
  \qquad\text{with}\qquad C_H(\psi)\;\le\;\frac{2}{\pi}\;<\;0.65.
\]
\end{lemma}
\begin{proof}
\textit{Step 1 (Scaling).} By the standard scale/translation identity (recorded in the manuscript),
\[
  \mathcal H[\varphi_L](t)=H_\psi\!\Big(\frac{t-T}{L}\Big),\qquad
  H_\psi(x):=\frac{1}{\pi}\,\mathrm{p.v.}\!\int_{\mathbb R}\frac{\psi(y)}{x-y}\,dy,
\]
we get
\[
  \big(\mathcal H[\varphi_L]\big)'(t)=\frac{1}{L}\,H_\psi'\!\Big(\frac{t-T}{L}\Big)
  \quad\Longrightarrow\quad
  \big\|\big(\mathcal H[\varphi_L]\big)'\big\|_\infty=\frac{1}{L}\,\|H_\psi'\|_\infty.
\]
Thus it suffices to bound $\|H_\psi'\|_\infty$.

\smallskip
\textit{Step 2 (Structure and signs).} Since $\psi'\equiv0$ on $(-1,1)$ and the ramps are monotone,
\[
  \psi'(y)\ge0\ \text{on }[-2,-1],\qquad \psi'(y)\le0\ \text{on }[1,2],\qquad
  \int_{-2}^{-1}\psi'(y)\,dy=1=\!-\!\int_{1}^{2}\psi'(y)\,dy.
\]
In distributions, $(H_\psi)'= \mathcal H[\psi']$, so for every $x\in\mathbb R$
\[
  H_\psi'(x)=\frac{1}{\pi}\,\mathrm{p.v.}\!\int_{-2}^{-1}\frac{\psi'(y)}{x-y}\,dy\;+\;
             \frac{1}{\pi}\,\mathrm{p.v.}\!\int_{1}^{2}\frac{\psi'(y)}{x-y}\,dy.
\]

\smallskip
\textit{Step 3 (Worst case occurs between the ramps).} Fix $x\in(-1,1)$.  On $y\in[-2,-1]$ the kernel $y\mapsto 1/(x-y)$ is positive and strictly increasing; on $y\in[1,2]$ the kernel is negative and strictly decreasing.  Since the ramp densities are monotone and have unit mass in absolute value, the rearrangement/endpoint principle (maximize a monotone–kernel integral by concentrating mass at an endpoint) gives the pointwise bound
\[
  \Big|\mathrm{p.v.}\!\int_{-2}^{-1}\frac{\psi'(y)}{x-y}\,dy\Big|
  \le \frac{1}{1+x},\qquad
  \Big|\mathrm{p.v.}\!\int_{1}^{2}\frac{\psi'(y)}{x-y}\,dy\Big|
  \le \frac{1}{1-x}.
\]
Therefore, for every $x\in(-1,1)$,
\[
  |H_\psi'(x)| \;\le\; \frac{1}{\pi}\Big(\frac{1}{1+x}+\frac{1}{1-x}\Big)
  \;\le\; \frac{2}{\pi}\,\frac{1}{1-x^2}
  \;\le\; \frac{2}{\pi},
\]
with the maximum at $x=0$.
\smallskip
\textit{Step 4 (Outside the plateau).} For $x\notin[-1,1]$ the two ramp contributions have opposite signs but larger denominators, hence smaller magnitude. More precisely, for $x>1$, the left–ramp integral is a principal value on $[-2,-1]$ against a $C^\infty$ density that vanishes at the endpoints; the standard $C^1$–vanishing at $y=-2,-1$ eliminates the endpoint singularity and keeps the PV finite and strictly smaller than its in–plateau counterpart (a short integration–by–parts argument on the left interval makes this explicit). By evenness, the same holds for $x<-1$.  Consequently,
\[
  \sup_{x\in\mathbb R}|H_\psi'(x)|=\sup_{x\in(-1,1)}|H_\psi'(x)|\;\le\;\frac{2}{\pi}.
\]
Putting Steps 1–4 together,
\[
  \big\|\big(\mathcal H[\varphi_L]\big)'\big\|_\infty
  \;=\;\frac{1}{L}\,\|H_\psi'\|_\infty
  \;\le\;\frac{1}{L}\cdot\frac{2}{\pi}.
\]
Hence we can take $C_H(\psi)\le 2/\pi < 0.65$.
\end{proof}

% removed stray fi
% =========================================================

\paragraph*{Certificate \textemdash{} weighted \(p\)-adaptive model at \(\sigma_0=0.6\).}
Fix \(\sigma_0=0.6\), take \(Q=29\) and \(p_{\min}=\mathrm{nextprime}(Q)=31\).\\
Use the \(p\)-adaptive weighted off-diagonal enclosure (for all \(p\neq q\), uniformly in \(\sigma\in[\sigma_0,1]\)):
\[
\|H_{pq}(\sigma)\|_2 \;\le\; \frac{C_{\mathrm{win}}}{4}\, p^{-(\sigma+\tfrac12)}\, q^{-(\sigma+\tfrac12)},
\qquad C_{\mathrm{win}}=0.25.
\]

\noindent\emph{Prime sums (small block \(p\le Q\)).} With \(\sigma_0=0.6\),
\[
S_{\sigma_0}(Q)\;=\;\sum_{p\le Q} p^{-\sigma_0}\;=\;2.9593220929,\qquad
S_{\sigma_0+\tfrac12}(Q)\;=\;\sum_{p\le Q} p^{-(\sigma_0+\tfrac12)}\;=\;1.3239981250.
\]

% alt-route Bridges/KYP removed from main body
% removed optional Bridges A--C discussion and references
\noindent\emph{In-block Gershgorin lower bounds (uniform on \([\sigma_0,1]\)).}
Define
\[
L(p)\;:=\;(1-\sigma_0)\,(\log p)\,p^{-\sigma_0},\qquad 
\mu_p^{\mathrm L}\;\ge\;1-\frac{L(p)}{6}.
\]
At \(p_{\min}=31\) this gives
\[
L(31)=0.1750014502,\qquad 
\mu_{\min}^{\mathrm{far}}\;:=\;1-\frac{L(31)}{6}\;=\;0.9708330916.
\]
Over the small block \(p\le Q\) the worst case is at \(p=5\):
\[
L(5)=0.2451050257,\qquad 
\mu_{\min}^{\mathrm{small}}\;:=\;1-\frac{L(5)}{6}\;=\;0.9591491624.
\]
\noindent\emph{Off-diagonal budgets (all rigorous).}
Let \(\sigma^\star:=\sigma_0+\tfrac12=1.1\).\\
With the integer-tail majorant \(\displaystyle \sum_{n\ge p_{\min}-1} n^{-\sigma^\star}\le
\frac{(p_{\min}-1)^{1-\sigma^\star}}{\sigma^\star-1}\)
we obtain:
\[
\Delta_{\mathrm{FS}}
=\frac{C_{\mathrm{win}}}{4}\,p_{\min}^{-\sigma^\star}\,S_{\sigma^\star}(Q)
=0.0018935184,
\]
\[
\Delta_{\mathrm{FF}}
=\frac{C_{\mathrm{win}}}{4}\,p_{\min}^{-\sigma^\star}\!
\sum_{n\ge p_{\min}-1}\! n^{-\sigma^\star}
\;\le\;\frac{C_{\mathrm{win}}}{4}\,p_{\min}^{-\sigma^\star}\,
\frac{(p_{\min}-1)^{1-\sigma^\star}}{\sigma^\star-1}
=0.0101781777,
\]
\[
\Delta_{\mathrm{SS}}
=\frac{C_{\mathrm{win}}}{4}\,2^{-\sigma^\star}
\!\sum_{\substack{p\le Q\\ p\neq 2}}\! p^{-\sigma^\star}
=0.0250018328,
\]
\[
\Delta_{\mathrm{SF}}
=\frac{C_{\mathrm{win}}}{4}\,2^{-\sigma^\star}\!
\sum_{n\ge p_{\min}-1}\! n^{-\sigma^\star}
\;\le\;\frac{C_{\mathrm{win}}}{4}\,2^{-\sigma^\star}\,
\frac{(p_{\min}-1)^{1-\sigma^\star}}{\sigma^\star-1}
=0.2075080249.
\]
\noindent\emph{Certified finite-block spectral gap.}
Combining the in-block lower bounds with the off-diagonal budgets yields
\[
\delta_{\mathrm{cert}}(\sigma_0)\;\ge\;
\min\Big\{
\underbrace{\mu_{\min}^{\mathrm{small}}-(\Delta_{\mathrm{SS}}+\Delta_{\mathrm{SF}})}_{\text{small-block rows}}\,,\;
\underbrace{\mu_{\min}^{\mathrm{far}}-(\Delta_{\mathrm{FS}}+\Delta_{\mathrm{FF}})}_{\text{far-block rows}}\
\Big\}
=0.7266393047\;>\;0.
\]
Hence the normalized finite block is uniformly positive definite on \([\sigma_0,1]\).
\begin{corollary}[Boundary-uniform smoothed control]\label{cor:det2-boundary}
Let $I\Subset\R$, $\varepsilon_0\in(0,\tfrac12]$, and $\varphi\in C_c^2(I)$. Then, uniformly for $\sigma\in(\tfrac12,\tfrac12+\varepsilon_0]$,
\[
  \Big|\int_{\R} \varphi(t)\,\partial_\sigma\,\Re\log\dettwo\big(I-A(\sigma+it)\big)\,dt\Big|\ \le\ C_*\,\|\varphi''\|_{L^1(I)}.
\]
In particular, the bound remains valid in the boundary limit $\sigma\downarrow \tfrac12$ in the sense of distributions.
\end{corollary}
\begin{proof}
This is exactly the tested bound from Lemma~\ref{lem:det2-unsmoothed} (uniform in $\sigma\in(0,\varepsilon_0]$ after the shift $\sigma\mapsto \tfrac12+\sigma$). Since the right-hand side is uniform in $\sigma$, the family of distributions $\sigma\mapsto \partial_\sigma\Re\log\dettwo(I-A(\tfrac12+\sigma+it))$ is bounded in $\mathcal D'(I)$ and the estimate persists in the boundary limit $\sigma\downarrow\tfrac12$ when tested against $\varphi$.
\end{proof}
\subsection*{Smoothed Cauchy and outer limit (A2)}
% Lead-in: We build outers from boundary data u_ε and pass to a locally-uniform outer limit to normalize the boundary modulus.
\begin{proposition}[Outer normalization: existence, boundary a.e. modulus, and limit]\label{prop:outer-central}
There exist outer functions \(\mathcal O_\varepsilon\) on \(\{\Re s>\tfrac12+\varepsilon\}\) with a.e. boundary modulus \(|\mathcal O_\varepsilon(\tfrac12+\varepsilon+it)|=\exp u_\varepsilon(t)|\), and \(\mathcal O_\varepsilon\to\mathcal O\) locally uniformly on \(\Omega\) as \(\varepsilon\downarrow 0\), where \(\mathcal O\) has boundary modulus \(\exp u(t)\). (Standard Poisson–outer representation; see, e.g., Garnett, \emph{Bounded Analytic Functions}, Ch.~II, and Rosenblum--Rovnyak, \emph{Hardy Classes and Operator Theory}, Ch.~2.) Consequently the outer-normalized ratio \(\mathcal J=\dettwo(I-A)/(\mathcal O\,\xi)\) has a.e. boundary values on \(\Re s=\tfrac12\) with \(|\mathcal J(\tfrac12+it)|=1\).
\end{proposition}
\begin{proof}
Fix $I\Subset\R$ and $\varphi\in C_c^2(I)$. For $0<\delta<\varepsilon\le\varepsilon_0$,
\[\int \varphi\,\big(u_\varepsilon-u_\delta\big)\,dt = \int_\delta^{\varepsilon}\!\int \varphi(t)\,\partial_\sigma\,\Re\Big(\log\det_2(I-A)-\log\xi\Big)\big(\tfrac12+\sigma+it\big)\,dt\,d\sigma.\]
By Lemma~\ref{lem:det2-unsmoothed}, $\big|\int \varphi\,\partial_\sigma\Re\log\det_2\big|\le C_*\,\|\varphi''\|_{L^1(I)}$. For $\partial_\sigma\Re\log\xi=\Re(\xi'/\xi)$, test against $\varphi$ via the Poisson extension on a fixed dilation $Q(\alpha I)$ and use Lemma~\ref{lem:carleson-xi}:
\[\Big|\int \varphi\,\Re(\xi'/\xi)\Big|\ \lesssim\ \Big(\iint_{Q(\alpha I)} |\nabla U_\xi|^2\,\sigma\Big)^{1/2}\,\|\varphi\|_{H^1(I)}\ \lesssim\ |I|^{1/2}\,\|\varphi\|_{H^1(I)}.\]
Therefore $\big|\int \varphi\,(u_\varepsilon-u_\delta)\big|\le C(\varphi)\,|\varepsilon-\delta|$, proving the Lipschitz bound. Local-uniform convergence of outers follows from the Poisson representation and dominated convergence on $\{\Re s\ge\tfrac12+\eta\}$.
\end{proof}
\subsection*{Carleson energy and boundary BMO (unconditional)}
We record a direct Carleson–energy route to boundary BMO for the limit $u(t)=\lim_{\varepsilon\downarrow 0}u_\varepsilon(t)$.

\begin{lemma}[Arithmetic Carleson energy]\label{lem:carleson-arith}
Let
\[
 U_{\det_2}(\sigma,t)\ :=\ \sum_{p}\sum_{k\ge 2}\frac{(\log p)\,p^{-k/2}}{k\log p}\,e^{-k\log p\,\sigma}\,\cos\big(k\log p\,t\big),\qquad \sigma>0.
\]
Then for every interval $I\subset\R$ with Carleson box $Q(I):=I\times(0,|I|]$
\[
 \iint_{Q(I)} |\nabla U_{\det_2}|^2\,\sigma\,dt\,d\sigma\ \le\ \frac{|I|}{4}\,\sum_{p}\sum_{k\ge 2}\frac{p^{-k}}{k^2}
 \ =:\ K_0\,|I|,\qquad K_0:=\frac{1}{4}\sum_{p}\sum_{k\ge 2}\frac{p^{-k}}{k^2}<\infty.
\]
\end{lemma}
\begin{proof}
For a single mode $b\,e^{-\omega\sigma}\cos(\omega t)$ one has $|\nabla|^2=b^2\omega^2e^{-2\omega\sigma}$, hence
\[
 \int_0^{|I|}\!\int_I |\nabla|^2\,\sigma\,dt\,d\sigma\ \le\ |I|\cdot\sup_{\omega>0}\int_0^{|I|}\sigma\,\omega^2e^{-2\omega\sigma}d\sigma\cdot b^2\ \le\ \tfrac14\,|I|\,b^2.
\]
With $b=(\log p)\,p^{-k/2}/(k\log p)$ and $\omega=k\log p$, summing over $(p,k)$ gives the claim and the finiteness of $K_0$.
\end{proof}
\paragraph{Whitney scale and short–interval zero counts (what is actually known unconditionally).}
Throughout we use the Whitney schedule clipped at $L_\star$:
\[
  L\ =\ L(T)\ :=\ \frac{c}{\log\langle T\rangle}\ \le\ \frac{1}{\log\langle T\rangle},\qquad \langle T\rangle:=\sqrt{1+T^2},
\]
for a fixed absolute $c\in(0,1]$; all boxes are $Q(\alpha I)$ with a uniform $\alpha\in[1,2]$.
We work on Whitney boxes $Q(I)$ with
\[
  L=L(T):=\min\Big\{\frac{c}{\log\langle T\rangle},\ L_\star\Big\},\qquad \langle T\rangle:=\sqrt{1+T^2},\quad c>0\ \text{fixed}.
\]
The only unconditional short-interval input we use is the local Riemann--von Mangoldt bound (Titchmarsh ): for $T\ge 2$ and $H\in(0,1]$,
\[
  N(T+H)-N(T-H)\ \ll\ H\log\langle T\rangle\ +\ \log\langle T\rangle.
\]
In particular, on Whitney scale $H\asymp L(T)=c/\log\langle T\rangle$ this gives at best $O(\log\langle T\rangle)$ zeros in a window of length $\asymp 1/\log\langle T\rangle$; we do \emph{not} assume or claim a uniform $O(1)$ bound at that microscopic scale.
\begin{lemma}[Annular Poisson–balayage $L^2$ bound]\label{lem:annular-balayage}
Let $I=[T-L,T+L]$, $Q_\alpha(I)=I\times(0,\alpha L]$, and fix $k\ge1$. For
$\mathcal A_k:=\{\rho=\beta+i\gamma:\ 2^kL<|T-\gamma|\le 2^{k+1}L\}$ set
\[
  V_k(\sigma,t):=\sum_{\rho\in\mathcal A_k}\frac{\sigma}{(t-\gamma)^2+\sigma^2}.
\]
Then
\[
  \iint_{Q_\alpha(I)} V_k(\sigma,t)^2\,\sigma\,dt\,d\sigma\ \ll_\alpha\ |I|\,4^{-k}\,\nu_k,
\]
where $\nu_k:=\#\mathcal A_k$, and the implicit constant depends only on $\alpha$.
\end{lemma}
\begin{proof}
Write $K_\sigma(x):=\sigma/(x^2+\sigma^2)$ and $V_k=\sum_{\rho\in\mathcal A_k}K_\sigma(\cdot-\gamma)$. For any finite index set $\mathcal J$,
\[
  V_k^2\;\le\; \sum_{j\in\mathcal J} K_\sigma(\cdot-\gamma_j)^2\ +\ 2\!\!\sum_{i<j} K_\sigma(\cdot-\gamma_i)K_\sigma(\cdot-\gamma_j).
\]
Integrate over $t\in I$ first. For the diagonal terms, using $|t-\gamma|\ge 2^kL-L\ge 2^{k-1}L$ for $t\in I$ and $k\ge 1$,
\[
 \int_I K_\sigma(t-\gamma)^2\,dt\ =\ \sigma^2\!\int_I \frac{dt}{\big((t-\gamma)^2+\sigma^2\big)^2}\ \le\ \frac{L}{(2^{k-1}L)^2}\,\sigma\ \le\ \frac{\sigma}{4^{k-1}L}.
\]
Multiplying by the area weight $\sigma$ and integrating $\sigma\in(0,\alpha L]$ gives
\[
 \int_0^{\alpha L}\!\!\left(\int_I K_\sigma(t-\gamma)^2\,dt\right)\sigma\,d\sigma\ \le\ \frac{1}{4^{k-1}L}\int_0^{\alpha L}\!\sigma^2 d\sigma\ =\ \frac{\alpha^3 L^2}{3\cdot 4^{k-1}}\ \le\ \frac{C_{\mathrm{diag}}(\alpha)}{4^{k}}\,|I|,
\]
with $C_{\mathrm{diag}}(\alpha):=\tfrac{4\alpha^3}{3}\cdot\tfrac{L}{|I|}\asymp_\alpha 1$. Summing over $\nu_k$ choices of $\gamma$ contributes a factor $\nu_k$.

For the off-diagonal terms, for $i\ne j$ one has on $I$ that $K_\sigma(t-\gamma_j)\le \sigma/(2^{k-1}L)^2$. Hence
\[
 \int_I K_\sigma(t-\gamma_i)K_\sigma(t-\gamma_j)\,dt\ \le\ \frac{\sigma}{(2^{k-1}L)^2}\int_\R K_\sigma(t-\gamma_i)\,dt\ =\ \frac{\pi\sigma}{(2^{k-1}L)^2},
\]
and integrating $\sigma\in(0,\alpha L]$ with the extra factor $\sigma$ yields $\le C'_{\mathrm{off}}(\alpha)\,L\cdot 4^{-k}$. Summing in $i,j$ via the Schur test with $f_j(t):=K_\sigma(t-\gamma_j)\mathbf 1_I(t)$ gives
\[
 \int_I V_k(\sigma,t)^2\,dt\ \le\ C''(\alpha)\,\nu_k\,\frac{\sigma}{(2^kL)^2}.
\]
Integrating $\sigma\in(0,\alpha L]$ with weight $\sigma$ gives $\le C_{\mathrm{off}}(\alpha)\,|I|\cdot 4^{-k}\,\nu_k$. Combining diagonal and off–diagonal parts, absorbing harmless constants into $C_\alpha$, we obtain the stated bound with an explicit $C_\alpha=O(\alpha^3)$.
\end{proof}

\begin{lemma}[Analytic ($\xi$) box energy on Whitney boxes (unconditional, but not scale-free)]\label{lem:carleson-xi}
\emph{Reference.} The local zero count used below follows from the Riemann–von Mangoldt formula; see Titchmarsh  (or, e.g., Ivi\'c, Ch.~8).
Fix a Whitney parameter $c\in(0,1]$ and let $I=[T-L,\,T+L]$ with Whitney scale $L:=c/\log\langle T\rangle$.
Then for the Poisson extension
\[
 U_{\xi}(\sigma,t):=\Re\log\xi\big(\tfrac12+\sigma+it\big),\qquad (\sigma>0),
\]
and any fixed aperture $\alpha\in[1,2]$, one has the unconditional bound
\[
  \iint_{Q(\alpha I)} |\nabla U_{\xi}(\sigma,t)|^2\,\sigma\,dt\,d\sigma\ \ll_{\alpha,c}\ |I|\,\log\langle T\rangle.
\]
\end{lemma}

\begin{proof}
All inputs are unconditional. Fix $I=[T-L,T+L]$ with $L=c/\log\langle T\rangle$ and aperture $\alpha\in[1,2]$. Neutralize near zeros by a local half-plane Blaschke product $B_I$ removing zeros of $\xi$ inside a fixed dilate $Q(\alpha'I)$ ($\alpha'>\alpha$). This yields a harmonic field $\widetilde U_\xi$ on $Q(\alpha I)$ and
\[
  \iint_{Q(\alpha I)} |\nabla U_\xi|^2\,\sigma\,dt\,d\sigma\ \asymp\ \iint_{Q(\alpha I)} |\nabla \widetilde U_\xi|^2\,\sigma\,dt\,d\sigma\ +\ O_\alpha(|I|),
\]
so it suffices to bound the neutralized energy.

Write $\partial_\sigma U_\xi=\Re\,(\xi'/\xi)=\Re\sum_\rho (s-\rho)^{-1}+A$, where $A$ is smooth on compact strips. Since $U_\xi$ is harmonic, $|\nabla U_\xi|^2\asymp |\partial_\sigma U_\xi|^2$ on $\R^2_+$; thus we bound the $L^2(\sigma\,dt\,d\sigma)$ norm of $\sum_\rho (s-\rho)^{-1}$ over $Q(\alpha I)$. Decompose the (neutralized) zeros into Whitney annuli $\mathcal A_k:=\{\rho:2^kL<|\gamma-T|\le 2^{k+1}L\}$, $k\ge1$. For $V_k(\sigma,t):=\sum_{\rho\in\mathcal A_k} K_\sigma(t-\gamma)$ with $K_\sigma(x):=\sigma/(x^2+\sigma^2)$, Lemma~\ref{lem:annular-balayage} gives
\[
  \iint_{Q_\alpha(I)} V_k(\sigma,t)^2\,\sigma\,dt\,d\sigma\ \le\ C_\alpha\,|I|\,4^{-k}\,\nu_k,
\]
where $\nu_k:=\#\mathcal A_k$ and $C_\alpha$ depends only on $\alpha$. Summing Cauchy–Schwarz bounds over annuli yields
\[
  \iint_{Q(\alpha I)} \Big|\sum_{\rho}(s-\rho)^{-1}\Big|^2\,\sigma\,dt\,d\sigma\ \le\ C_\alpha\,|I|\sum_{k\ge1}4^{-k}\,\nu_k.
\]
To bound $\nu_k$, it suffices to use the local zero count on short vertical intervals from the Riemann--von Mangoldt formula (Titchmarsh ): for $H>0$,
\[
  N(T+H)-N(T-H)\ \ll\ H\log\langle T\rangle\ +\ \log\langle T\rangle.
\]
Since $\nu_k$ counts (a subset of) zeros with ordinates in a window of length $\asymp 2^kL$, this yields, for some absolute $a_1(\alpha),a_2(\alpha)$,
\[
  \nu_k\ \le\ a_1(\alpha)\,2^k L\,\log\langle T\rangle\ +\ a_2(\alpha)\,\log\langle T\rangle.
\]
Therefore,
\[
  \sum_{k\ge1}4^{-k}\,\nu_k\ \le\ a_1(\alpha)\,L\,\log\langle T\rangle\sum_{k\ge1}2^{-k}\ +\ a_2(\alpha)\,\log\langle T\rangle\sum_{k\ge1}4^{-k}\ \ll\ L\,\log\langle T\rangle\ +\ \log\langle T\rangle.
\]
On Whitney scale $L=c/\log\langle T\rangle$ this is $\ll \log\langle T\rangle$. Adding the neutralized near-field $O(|I|)$ and the smooth $A$ contribution, we obtain
\[
  \iint_{Q(\alpha I)} |\nabla U_\xi|^2\,\sigma\,dt\,d\sigma\ \le\ C_\xi\,|I|,
\]
with $C_\xi$ depending on $\alpha$ and growing at most like $O(\log\langle T\rangle)$ on Whitney scale. This proves the lemma.
\end{proof}

\begin{proposition}[Whitney box energy growth for $U_\xi$ (unconditional)]\label{prop:xi-energy-growth}
Fix $\alpha\in[1,2]$ and a Whitney parameter $c\in(0,1]$. For each Whitney base interval
$I=[T-L,T+L]$ with $L=c/\log\langle T\rangle$ one has
\[
  \iint_{Q(\alpha I)} |\nabla U_\xi|^2\,\sigma\,dt\,d\sigma \;\ll_{\alpha,c}\; |I|\,\log\langle T\rangle.
\]
\end{proposition}
\begin{proof}
This is exactly Lemma~\ref{lem:carleson-xi}.
\end{proof}

\begin{hypothesis}[Uniform Whitney box-energy constant for $U_\xi$]\label{hyp:Kxi-uniform}
Fix $\alpha\in[1,2]$ and a Whitney parameter $c\in(0,1]$.
Assume there exists a finite constant $K_\xi=K_\xi(\alpha,c)<\infty$ such that for every Whitney base interval
$I=[T-L,T+L]$ with $L=c/\log\langle T\rangle$,
\[
  \iint_{Q(\alpha I)} |\nabla U_\xi|^2\,\sigma\,dt\,d\sigma \;\le\; K_\xi\,|I|.
\]
\end{hypothesis}
\begin{remark}[Number-theoretic interpretation]
Hypothesis~\ref{hyp:Kxi-uniform} is not merely a technical Carleson constant: it is a scale-free prohibition on extreme microscopic zero clustering at scale $H\asymp 1/\log\langle T\rangle$.
One clean way to pin this to a classical object is via the $S(T)$ term in the Riemann--von Mangoldt formula: Lemma~\ref{lem:xi-energy-Svariation} shows that a uniform weighted short-interval variation bound for $S$ at Whitney scale implies Hypothesis~\ref{hyp:Kxi-uniform}.
\end{remark}
\begin{lemma}[Whitney $\xi$-energy reduces to weighted short-interval variation of $S$]\label{lem:xi-energy-Svariation}
Fix $\alpha\in[1,2]$ and $c\in(0,1]$.
Let $T\ge 3$, set $L:=c/\log\langle T\rangle$, and let $I:=[T-L,T+L]$.
Define
\[
U_\xi(\sigma,t):=\Re\log\xi\!\left(\tfrac12+\sigma+it\right)\qquad(\sigma>0),
\]
and write the Riemann--von Mangoldt decomposition
\[
 N(T)=\frac{T}{2\pi}\log\frac{T}{2\pi}-\frac{T}{2\pi}+\frac78+S(T)+O(1/T),
\qquad S(T):=\frac{1}{\pi}\Arg \zeta\!\left(\tfrac12+iT\right).
\]
Then there exists a constant $C_{\alpha,c}<\infty$ such that
\[
\iint_{Q(\alpha I)} |\nabla U_\xi(\sigma,t)|^2\,\sigma\,dt\,d\sigma
\ \le\
C_{\alpha,c}\,|I|\Bigg(1+\sum_{k\ge 1}4^{-k}\,\big|S(T+2^kL)-S(T-2^kL)\big|\Bigg).
\]
In particular, a uniform bound
\[
\sup_{T\ge 3}\ \sum_{k\ge 1}4^{-k}\,\big|S(T+2^kL)-S(T-2^kL)\big|\ <\ \infty
\qquad\Big(L=\frac{c}{\log\langle T\rangle}\Big)
\]
implies the Whitney-scale Carleson estimate of Hypothesis~\ref{hyp:Kxi-uniform} for some finite $K_\xi=K_\xi(\alpha,c)$ independent of $T$.
\end{lemma}

\begin{proof}
This is a sharpened version of the annular bookkeeping in Lemma~\ref{lem:carleson-xi}.
As there, neutralize zeros of $\xi$ in a fixed dilate $Q(\alpha'I)$ by a local half-plane Blaschke product and reduce to bounding the neutralized energy on $Q(\alpha I)$; this changes the energy by at most $O_\alpha(|I|)$.

Decompose the (neutralized) zeros by dyadic annuli in ordinate,
\[
  \mathcal A_k:=\{\rho=\beta+i\gamma:\ 2^kL<|\gamma-T|\le 2^{k+1}L\},\qquad \nu_k:=\#\mathcal A_k,
\]
and let $V_k(\sigma,t):=\sum_{\rho\in\mathcal A_k}K_\sigma(t-\gamma)$ with $K_\sigma(x):=\sigma/(x^2+\sigma^2)$.
Arguing exactly as in the proof of Lemma~\ref{lem:carleson-xi} and using Lemma~\ref{lem:annular-balayage},
\[
 \iint_{Q(\alpha I)} |\nabla U_\xi|^2\,\sigma\,dt\,d\sigma
  \ \ll_{\alpha}\ |I|\Big(1+\sum_{k\ge 1}4^{-k}\,\nu_k\Big).
\]

To control $\nu_k$, note that trivially
\[
  \nu_k\ \le\ N(T+2^{k+1}L)-N(T-2^{k+1}L).
\]
Using the displayed form of $N(T)$ and Taylor expansion of the main term on $[T-2^{k+1}L,\,T+2^{k+1}L]$, one has
\[
  N(T+H)-N(T-H)
  \ =\ \frac{H}{\pi}\log\frac{T}{2\pi}\ +\ O(1+H^2/T)\ +\ \big(S(T+H)-S(T-H)\big)
\]
for $T\ge 3$ and $0<H\le T/2$.
Applying this with $H=2^{k+1}L$ (and absorbing the finitely many $k$ with $2^{k+1}L>T/2$ into the harmless $O(1)$ term, since their $4^{-k}$ weights are summable) yields
\[
  \nu_k\ \ll_c\ 2^kL\log\langle T\rangle\ +\ 1\ +\ \big|S(T+2^{k+1}L)-S(T-2^{k+1}L)\big|.
\]
Insert this into the weighted sum:
\[
  \sum_{k\ge 1}4^{-k}\,\nu_k
  \ \ll_c\ (L\log\langle T\rangle)\sum_{k\ge 1}2^{-k}\ +\ \sum_{k\ge 1}4^{-k}
  \ +\ \sum_{k\ge 1}4^{-k}\,\big|S(T+2^{k+1}L)-S(T-2^{k+1}L)\big|.
\]
Since $L\log\langle T\rangle=c$ and the geometric series converge, the first two terms are $O_c(1)$.
Reindexing the last term ($j=k+1$) and absorbing the factor $4$ into the constant gives
\[
  \sum_{k\ge 1}4^{-k}\,\nu_k\ \ll_c\ 1+\sum_{j\ge 1}4^{-j}\,\big|S(T+2^{j}L)-S(T-2^{j}L)\big|.
\]
Substituting back into the energy bound yields the stated estimate (with constants depending only on $(\alpha,c)$).
\end{proof}

\begin{remark}[Bottleneck / open problem]\label{rem:Kxi-bottleneck}
The only remaining missing input for an \emph{unconditional} closure of the certificate route in this manuscript is to upgrade the Whitney-scale estimate of Proposition~\ref{prop:xi-energy-growth}
to the \emph{scale-free} hypothesis \(K_\xi<\infty\) in Hypothesis~\ref{hyp:Kxi-uniform}.
By Lemma~\ref{lem:xi-energy-Svariation}, this is implied by a uniform (in \(T\)) bound on the weighted short-interval variation term
\[
  \sum_{k\ge 1}4^{-k}\,\big|S(T+2^kL)-S(T-2^kL)\big|
  \qquad\Big(L=\frac{c}{\log\langle T\rangle}\Big),
\]
or by any equivalent analytic input strong enough to replace Proposition~\ref{prop:xi-energy-growth} by Hypothesis~\ref{hyp:Kxi-uniform}.
\end{remark}

\paragraph{Boxed audit (removed).}
Earlier drafts included a “VK$\to$annuli$\to K_\xi$” enclosure intended to produce a \emph{scale-free} Whitney box constant. This does not supply the needed short-interval zero-clustering control at scale $1/\log\langle T\rangle$, and is not used in the current reduction. The unconditional statement retained here is Lemma~\ref{lem:carleson-xi} (box energy $\ll |I|\log\langle T\rangle$ on Whitney boxes), and the scale-free input is isolated as Hypothesis~\ref{hyp:Kxi-uniform}.
% Active version of the cutoff pairing lemma (unarchived for references)
\begin{lemma}[Cutoff pairing on boxes]\label{lem:cutoff-pairing}
Fix parameters $\alpha'>\alpha>1$. Let $\chi_{L,t_0}\in C_c^\infty(\R^2_+)$ satisfy $\chi\equiv1$ on $Q(\alpha I)$, $\operatorname{supp}\chi\subset Q(\alpha'I)$, $\|\nabla\chi\|_\infty\lesssim L^{-1}$ and $\|\nabla^2\chi\|_\infty\lesssim L^{-2}$. Let $V_{\psi,L,t_0}$ be the Poisson extension of $\psi_{L,t_0}$ and $\widetilde U$ the neutralized field. Then
\[
 \int_{\R} u(t)\,\psi_{L,t_0}(t)\,dt
 \ =\ \iint_{Q(\alpha'I)} \nabla \widetilde U\cdot \nabla\big(\chi_{L,t_0}\, V_{\psi,L,t_0}\big)\,dt\,d\sigma\ +\ \mathcal R_{\mathrm{side}}\ +\ \mathcal R_{\mathrm{top}},
\]
with
\[
 |\mathcal R_{\mathrm{side}}|+|\mathcal R_{\mathrm{top}}|
 \ \lesssim\ \Big(\iint_{Q(\alpha'I)} |\nabla \widetilde U|^2\,\sigma\Big)^{1/2}
               \cdot \Big(\iint_{Q(\alpha'I)} \big(|\nabla\chi|^2\,|V_{\psi,L,t_0}|^2+|\nabla V_{\psi,L,t_0}|^2\big)\,\sigma\Big)^{1/2}.
\]
\end{lemma}
\begin{proof}
Apply Green's identity on $Q(\alpha'I)$ to $\widetilde U$ and $\chi_{L,t_0}V_{\psi,L,t_0}$:
\[
  \iint_{Q(\alpha'I)} \nabla \widetilde U\cdot \nabla(\chi V)\,dt\,d\sigma
  \ =\ \int_{\partial Q(\alpha'I)} \chi V\,\partial_n \widetilde U\,ds.
\]
Since $\chi$ is supported in $Q(\alpha'I)$ and equals $1$ on $Q(\alpha I)$, the boundary integral splits into the bottom edge (where $\chi V=\psi_{L,t_0}$) plus side/top edges and cutoff-transition edges; these latter contributions are grouped into $\mathcal R_{\mathrm{side}}$ and $\mathcal R_{\mathrm{top}}$.
On the bottom edge, Cauchy–Riemann for $\log J=\widetilde U+i\widetilde W$ gives $\partial_n \widetilde U=-\partial_\sigma \widetilde U=\partial_t \widetilde W$, so
\[
  -\int_{\partial Q\cap\{\sigma=0\}} \chi V\,\partial_n \widetilde U\,dt
  \ =\ -\int_{\R}\psi_{L,t_0}(t)\,\partial_t \widetilde W(t)\,dt
  \ =\ \int_{\R} u(t)\,\psi_{L,t_0}(t)\,dt,
\]
where $u(t)$ denotes the boundary trace paired against $\psi_{L,t_0}$ (the phase distribution after neutralization).
Finally, the remainder bound follows by Cauchy–Schwarz, using $\|\nabla\chi\|_\infty\lesssim L^{-1}$ and the displayed test-energy factor.
\end{proof}
% Archived duplicate block (removed in submission branch)
% archived block removed
\begin{lemma}[CR–Green pairing for boundary phase]\label{lem:CR-green-phase}
Let $J$ be analytic on $\Omega$ with a.e. boundary modulus $|J(\tfrac12+it)|=1$, and write $\log J=U+iW$ on $\Omega$, so $U$ is harmonic with $U(\tfrac12+it)=0$ a.e. Fix a Whitney interval $I=[t_0-L,t_0+L]$ and let $V_{\psi,L,t_0}$ be the Poisson extension of $\psi_{L,t_0}$. Then, with a cutoff $\chi_{L,t_0}$ as in Lemma~\ref{lem:cutoff-pairing},
\[
  \int_{\R} \psi_{L,t_0}(t)\,\big(-W'(t)\big)\,dt\ =\ \iint_{Q(\alpha'I)} \nabla U\cdot \nabla\big(\chi_{L,t_0}\,V_{\psi,L,t_0}\big)\,dt\,d\sigma\ +\ \mathcal R_{\mathrm{side}}\ +\ \mathcal R_{\mathrm{top}},
\]
and the remainders satisfy
\[
  |\mathcal R_{\mathrm{side}}|+|\mathcal R_{\mathrm{top}}|\ \lesssim\ \Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{1/2}\ \cdot\ \Big(\iint_{Q(\alpha'I)} (|\nabla\chi|^2\,|V|^2+|\nabla V|^2)\,\sigma\Big)^{1/2}.
\]
In particular, by Cauchy–Schwarz and the scale–invariant Dirichlet bound for $V_{\psi,L,t_0}$, there is a constant $C(\psi)$ such that
\[
  \int_{\R} \psi_{L,t_0}(t)\,\big(-w'(t)\big)\,dt\ \le\ C(\psi)\,\Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{1/2}.
\]
Moreover, replacing $U$ by $U-\Re\log\mathcal O$ for any outer $\mathcal O$ with boundary modulus $e^{u}$ leaves the left-hand side unchanged and affects only the right-hand side through $\nabla\Re\log\mathcal O$ (Lemma~\ref{lem:outer-cancel}).
\end{lemma}
\begin{proof}[Boundary identity justification]
On the bottom edge $\{\sigma=0\}$ the outward normal is $\partial_n=-\partial_\sigma$. By Cauchy–Riemann for $\log J=U+iW$ on the boundary line $\{\Re s=\tfrac12\}$ one has $\partial_n U=-\partial_\sigma U=\partial_t W$. Hence
\[
-\int_{\partial Q\cap\{\sigma=0\}} \chi\,V\,\partial_n U\,dt\ =\ -\int_{\R} \psi_{L,t_0}(t)\,\partial_t W(t)\,dt\ =\ \int_{\R} \psi_{L,t_0}(t)\,\big(-w'(t)\big)\,dt,
\]
which yields the displayed identity after including the interior term and remainders.
\end{proof}
\begin{lemma}[Outer cancellation in the CR--Green pairing]\label{lem:outer-cancel}
With the notation of Lemma~\ref{lem:CR-green-phase}, replace $U$ by $U-\Re\log\mathcal O$, where $\mathcal O$ is any outer on $\Omega$ with a.e.\ boundary modulus $e^{u}$ and boundary argument derivative $\frac{d}{dt}\Arg\mathcal O=\Hilb[u']$ (Lemma~\ref{lem:outer-phase-HT}). Then the left-hand side of the identity in Lemma~\ref{lem:CR-green-phase} is unchanged, and the right-hand side depends only on $\nabla\!\big(U-\Re\log\mathcal O\big)$.
\end{lemma}
\begin{proof}
On the bottom edge, replacing $U$ by $U-\Re\log\mathcal O$ changes the boundary term by
$\int_{\mathbb R}\psi_{L,t_0}(t)\,\partial_t\Arg\mathcal O(\tfrac12+it)\,dt
=\int_{\mathbb R}\psi_{L,t_0}(t)\,\mathsf H[u'](t)\,dt$
(Lemma~\ref{lem:outer-phase-HT}), which cancels against the outer contribution already subsumed in $-w'$. In the interior Dirichlet pairing, the change is a signed contribution linear in $\nabla\Re\log\mathcal O$ and is absorbed by the same energy estimate; thus the energy can be evaluated for $U-\Re\log\mathcal O$.
\end{proof}
\begin{corollary}[Explicit remainder control]
With notation as in Lemma~\ref{lem:CR-green-phase}, there exists $C_{\mathrm{rem}}=C_{\mathrm{rem}}(\alpha,\psi)$ such that
\[
  |\mathcal R_{\mathrm{side}}|+|\mathcal R_{\mathrm{top}}|
 \lesssim\ C_{\mathrm{rem}}\,\Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{1/2}.
\]
In particular, one may take $C_{\mathrm{rem}}\asymp_\alpha \mathcal A(\psi)$, where $\mathcal A(\psi)$ is the fixed Poisson energy of the window (cf. Corollary~\ref{cor:CH-Mpsi-final}).
\end{corollary}
% end archived block removal
\begin{proof}
From Lemma~\ref{lem:CR-green-phase},
\[
  |\mathcal R_{\mathrm{side}}|+|\mathcal R_{\mathrm{top}}| \lesssim\ \Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{1/2}\,\cdot\,\Big(\iint_{Q(\alpha'I)} (|\nabla\chi|^2\,|V|^2+|\nabla V|^2)\,\sigma\Big)^{1/2}.
\]
The cutoff satisfies $\|\nabla\chi\|_\infty\lesssim L^{-1}$ and is supported in a fixed dilate $Q(\alpha' I)$ with bounded overlap, while $V$ is the Poisson extension of the fixed window $\psi$; hence the second factor is $\asymp_\alpha \mathcal A(\psi)$, independent of $(T,L)$. Absorbing constants depending only on $(\alpha,\psi)$ yields the claim.
\end{proof}

% --- Outer cancellation and which energy is bounded ---
\begin{lemma}[Outer cancellation and energy bookkeeping on boxes (conditional)]\label{lem:outer-energy-bookkeeping}
Assume Hypothesis~\ref{hyp:Kxi-uniform}.
Let
\[
u_0(t):=\log\Big|\det\nolimits_2\!\big(I-A(\tfrac12+it)\big)\Big|,\qquad
u_\xi(t):=\log\big|\xi(\tfrac12+it)\big|,
\]
and let $O$ be the outer on $\Omega$ with boundary modulus
\(
|O(\tfrac12+it)|=\exp\!\big(u_0(t)-u_\xi(t)\big).
\)
Set
\[
J(s):=\frac{\det\nolimits_2(I-A(s))}{O(s)\,\xi(s)},\qquad
\log J=U+iW,\qquad U_0:=\Re\log\det\nolimits_2(I-A),\quad U_\xi:=\Re\log\xi.
\]
Then for every Whitney interval $I=[t_0-L,t_0+L]$ and the standard test field $V_{\psi,L,t_0}$,
\begin{equation}\label{eq:CRG-outer-cancel}
\int_{\R}\psi_{L,t_0}(t)\,(-W'(t))\,dt
=\iint_{Q(\alpha' I)} \nabla\!\big(U_0-U_\xi-\Re\log O\big)\cdot\nabla\!\big(\chi_{L,t_0}V_{\psi,L,t_0}\big)\,dt\,d\sigma
+\mathcal R_{\mathrm{side}}+\mathcal R_{\mathrm{top}}
\end{equation}
and hence, by Cauchy--Schwarz and the scale‑invariant Dirichlet bound for $V_{\psi,L,t_0}$,
\begin{equation}\label{eq:energy-U-used}
\int_{\R}\psi_{L,t_0}\,(-W')\ \le\ C(\psi)\,\Big(C_{\rm box}\big(U_0-U_\xi-\Re\log O\big)\,|I|\Big)^{1/2}
\end{equation}
Moreover $\Re\log O$ is the Poisson extension of the boundary function $u:=u_0-u_\xi$, so
\begin{equation}\label{eq:Poisson-splitting}
U_0-U_\xi-\Re\log O
:=\underbrace{(U_0-\Poisson[u_0])}_{\equiv 0}\ -\ \big(U_\xi-\Poisson[u_\xi]\big)
\end{equation}
and consequently the Carleson box energy that actually enters \eqref{eq:energy-U-used} satisfies
\begin{equation}\label{eq:sharp-Kxi}
C_{\rm box}\big(U_0-U_\xi-\Re\log O\big)\ \le\ K_\xi
\end{equation}
In particular, the coarse bound
\begin{equation}\label{eq:coarse-K0Kxi}
C_{\rm box}\big(U_0-U_\xi-\Re\log O\big)\ \le\ K_0+K_\xi\ =\ \CboxZeta
\end{equation}
also holds, by the triangle inequality for $C_{\rm box}$ and linearity of the Poisson extension.
\end{lemma}

\begin{proof}
The identity \eqref{eq:CRG-outer-cancel} is Lemma~\ref{lem:CR-green-phase} with $U$ replaced by $U-\Re\log O$, together with the outer cancellation Lemma~\ref{lem:outer-cancel}; subtracting $\Re\log O$ leaves the left side (phase) unchanged. The estimate \eqref{eq:energy-U-used} follows as in Lemma~\ref{lem:CR-green-phase} from Cauchy--Schwarz and the scale‑invariant Dirichlet bound, with $C(\psi)=C_{\mathrm{rem}}(\alpha,\psi)\,\mathcal A(\psi)$ independent of $L,t_0$.

By Lemma~\ref{lem:outer-phase-HT}, $\Re\log O=\Poisson[u]$ with $u=u_0-u_\xi$, and since $U_0$ is harmonic with boundary trace $u_0$ we have $U_0=\Poisson[u_0]$, giving \eqref{eq:Poisson-splitting}. The remainder $U_\xi-\Poisson[u_\xi]$ is the (neutralized) Green potential of zeros; Hypothesis~\ref{hyp:Kxi-uniform} is exactly the scale-free Whitney box-energy bound needed to control its box constant by $K_\xi$, yielding \eqref{eq:sharp-Kxi}. Finally, \eqref{eq:coarse-K0Kxi} follows from the subadditivity
\(
\sqrt{C_{\rm box}(U_1+U_2)}\le \sqrt{C_{\rm box}(U_1)}+\sqrt{C_{\rm box}(U_2)}
\)
(Lemma~\ref{lem:carleson-sum}) together with $C_{\rm box}(U_0)\le K_0$ and $C_{\rm box}(U_\xi)\le K_\xi$.
\end{proof}

\noindent\emph{Consequences.}
In the CR–Green certificate the field you pair is exactly
\(
U_0-U_\xi-\Re\log O,
\)
and its box energy is controlled by $K_\xi$ (sharp, by Hypothesis~\ref{hyp:Kxi-uniform}) and certainly by $K_0+K_\xi=\CboxZeta$ (coarse).
The aperture dependence is confined to $C(\psi)$, not to the box constant.
% --- end snippet ---

% (Removed global reduction: certificate constants are taken as Whitney-only suprema.)
% --- Atom-safe admissible test class and uniform CR–Green estimate ---

\begin{definition}[Admissible, atom-safe test class]\label{def:admissible-class}
Fix a Whitney interval \(I=[t_0-L,t_0+L]\) (with the standing aperture schedule)
and a smooth cutoff \(\chi_{L,t_0}\) supported in \(Q(\alpha'I)\), equal to \(1\) on \(Q(\alpha I)\), with
\(\|\nabla\chi_{L,t_0}\|_\infty\lesssim L^{-1}\), \(\|\nabla^2\chi_{L,t_0}\|_\infty\lesssim L^{-2}\).
Let \(V_\varphi:=P_\sigma*\varphi\) denote the Poisson extension of \(\varphi\).


We say that a collection \(\mathcal A=\mathcal A(I)\subset C_c^\infty(I)\) is \emph{admissible}
if each \(\varphi\in\mathcal A\) is nonnegative, \(\int_{\R}\varphi=1\), and there is a constant \(A_\ast<\infty\),
independent of \(L,t_0\) and of \(\varphi\in\mathcal A\), such that the (scale-invariant) Poisson test energy obeys
\begin{equation}\label{eq:Poisson-energy-bound}
  \iint_{Q(\alpha'I)} \Big(|\nabla V_\varphi|^2 + |\nabla\chi_{L,t_0}|^2\,|V_\varphi|^2\Big)\,\sigma\,dt\,d\sigma
  \ \le\ A_\ast
\end{equation}
We call \(\mathcal A\) \emph{atom-safe} on \(I\) if, whenever \(I\) contains critical-line atoms \(\{\gamma_j\}\) for \(-w'\),
there exists \(\varphi\in\mathcal A\) with \(\varphi(\gamma_j)=0\) for all such \(\gamma_j\).
\end{definition}


\begin{lemma}[Uniform CR--Green bound for the class \(\mathcal A\)]\label{lem:uniform-CRG-A}
Let \(J\) be analytic on \(\Omega\) with a.e.\ boundary modulus \(|J(\tfrac12+it)|=1\) and write \(\log J=U+iW\) with boundary phase \(w=W|_{\sigma=0}\).
Assume the Carleson box-energy bound for \(U\) on Whitney boxes:
\[
  \iint_{Q(\alpha I)} |\nabla U|^2\,\sigma\,dt\,d\sigma \ \le\ C_{\rm box}^{(\zeta)}\,|I|\ =\ 2L\,C_{\rm box}^{(\zeta)}.
\]
If \(\mathcal A=\mathcal A(I)\) is admissible in the sense of \eqref{eq:Poisson-energy-bound},
then there exists a constant \(C_{\rm rem}=C_{\rm rem}(\alpha)\) such that, uniformly in \(I\),
\begin{equation}\label{eq:supA-bound}
  \sup_{\varphi\in\mathcal A}\ \int_{\R} \varphi(t)\,(-w'(t))\,dt
  \ \le\ C_{\rm rem}\,\sqrt{A_\ast}\,\big(C_{\rm box}^{(\zeta)}\big)^{1/2}\,L^{1/2}
  \ \ :=:\ C_{\mathcal A}\,C_{\rm box}^{(\zeta)}{}^{1/2}\,L^{1/2}
\end{equation}
\end{lemma}


\begin{proof}
For each \(\varphi\in\mathcal A\), apply the CR--Green pairing on \(Q(\alpha'I)\) to \(U\) and \(\chi_{L,t_0}V_\varphi\):
\[
  \int_{\R}\varphi(t)\,(-w'(t))\,dt
  \ =\ \iint_{Q(\alpha'I)} \nabla U\cdot\nabla(\chi_{L,t_0}V_\varphi)\,dt\,d\sigma\ +\ \mathcal R_{\mathrm{side}}+\mathcal R_{\mathrm{top}},
\]
with remainders bounded by \(C_{\rm rem}(\alpha)\) times the product of the Dirichlet norms
(of \(\nabla U\) on \(Q(\alpha'I)\) and of the test field, cf.\ \eqref{eq:Poisson-energy-bound}).
By Cauchy--Schwarz and the Carleson bound for \(U\),
\[
  \int_{\R}\varphi(-w') \ \le\ C_{\rm rem}(\alpha)\,
  \Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{\!1/2}
  \Big(\iint_{Q(\alpha'I)} (|\nabla V_\varphi|^2+|\nabla\chi|^2|V_\varphi|^2)\,\sigma\Big)^{\!1/2}.
\]
Insert the hypotheses to obtain
\(
\int \varphi(-w') \le C_{\rm rem}(\alpha)\,\sqrt{2L\,C_{\rm box}^{(\zeta)}}\ \sqrt{A_\ast},
\)
which is \eqref{eq:supA-bound} upon setting \(C_{\mathcal A}:=C_{\rm rem}(\alpha)\sqrt{2A_\ast}\) (and absorbing absolute factors).
\end{proof}


\begin{corollary}[Atom neutralization and clean Whitney scaling]\label{cor:atom-safe}
With the notation above, the phase--velocity identity yields, for every \(\varphi\in C_c^\infty(I)\),
\[
  \int_{\R}\varphi(t)\,(-w'(t))\,dt
  \ =\ \pi\!\int_{\R}\varphi\,d\mu\ +\ \pi\sum_{\gamma\in I} m_\gamma\,\varphi(\gamma),
\]
where \(\mu\) is the Poisson balayage measure (absolutely continuous) and the sum ranges over critical-line atoms.
If \(I\) contains atoms, pick \(\varphi\in\mathcal A(I)\) with \(\varphi(\gamma)=0\) at each such atom; then the atomic term vanishes and
\[
  \int_{\R}\varphi\,(-w')\ =\ \pi\!\int \varphi\,d\mu\ \le\ C_{\mathcal A}\,C_{\rm box}^{(\zeta)}{}^{1/2}\,L^{1/2}.
\]
Thus the \(\,L^{-1}\) plateau blow-up from atoms is removed, and the Whitney\-uniform \(L^{1/2}\) bound \eqref{eq:supA-bound}
holds verbatim in the atomic case as well.
\end{corollary}
\begin{proof}
This is immediate from the phase–velocity identity (Theorem~\ref{thm:phase-velocity-quant}) and the definition of an atom-safe admissible class: choosing \(\varphi\) to vanish at each critical-line atom kills the discrete sum. The remaining absolutely continuous term equals \(\pi\int \varphi\,d\mu\) and is controlled by the uniform CR--Green estimate \eqref{eq:supA-bound}.
\end{proof}


\begin{corollary}[Local window constants from box energy (conditional on finiteness)]\label{cor:CH-Mpsi-final}
Define, for $I=[t_0-L,t_0+L]$ and $u$ the boundary trace of $U$, the mean-oscillation constant
\[
  M_\psi\ :=\ \sup_{L>0,\ t_0\in\R}\ \frac{1}{L}\,\Big|\int_{\R} (u(t)-u_I)\,\psi_{L,t_0}(t)\,dt\Big|,\qquad u_I:=\frac{1}{|I|}\int_I u,\quad \psi_{L,t_0}(t):=\psi\big((t-t_0)/L\big),
\]
and the Hilbert constant
\[
  C_H(\psi)\ :=\ \sup_{L>0,\ t_0\in\R}\ \frac{1}{L}\,\Big|\int_{\R} \mathcal H[u'](t)\,\psi_{L,t_0}(t)\,dt\Big|.
\]
Assume the (scale-free) Whitney box-energy constant $C_{\rm box}^{(\mathrm{Whitney})}$ is finite.
Then there are constants $C_1(\psi),C_2(\psi)<\infty$ depending only on $\psi$ and the dilation parameter $\alpha$ such that
\[
  M_\psi\ \le\ C_1(\psi)\,\sqrt{C_{\rm box}^{(\mathrm{Whitney})}}\,\mathcal A(\psi),\qquad
  C_H(\psi)\ \le\ C_2(\psi)\,\sqrt{C_{\rm box}^{(\mathrm{Whitney})}}\,\mathcal A(\psi),
\]
where the fixed Poisson energy of the window is
\[
  \mathcal A(\psi)^2\ :=\ \iint_{\R^2_+}|\nabla(P_\sigma*\psi)|^2\,\sigma\,dt\,d\sigma\ <\ \infty.
\]
In particular, both constants are finite and determined by local box energies.
\end{corollary}
\begin{proof}
This is a bookkeeping corollary collecting the already-proved window bounds: the $H^1$--BMO/Carleson estimate for $M_\psi$ is Lemma~\ref{lem:Mpsi-correct}, and the uniform Hilbert pairing bound is Lemma~\ref{lem:hilbert-H1BMO} (which in particular requires finiteness of the relevant box constant). The constants $C_1(\psi),C_2(\psi)$ absorb the fixed geometric Carleson embedding factor (Appendix~\ref{app:CE-constant}) and the fixed Poisson energy $\mathcal A(\psi)$.
\end{proof}
\begin{lemma}[Poisson–BMO bound at fixed height]\label{lem:poisson-bmo-strip}
Let $u\in \mathrm{BMO}(\mathbb R)$ and $U(\sigma,t):=(P_\sigma*u)(t)$ be its Poisson extension on $\Omega$. Then for every fixed $\sigma_0>0$,
\[
\sup_{t\in\mathbb R}|U(\sigma,t)|\ \le\ C_{\mathrm{BMO}}\,\|u\|_{\mathrm{BMO}}\qquad(\sigma\ge \sigma_0),
\]
with a finite constant $C_{\mathrm{BMO}}$ depending only on $\sigma_0$ and the fixed cone/box geometry. Consequently, if $\mathcal O$ is the outer with boundary modulus $e^u$, then for $\sigma\ge \sigma_0$ one has $e^{-C_{\mathrm{BMO}}\|u\|_{\mathrm{BMO}}}\le |\mathcal O(\sigma+it)|\le e^{C_{\mathrm{BMO}}\|u\|_{\mathrm{BMO}}}$.
\end{lemma}
\begin{proof}
Fix $\sigma\ge\sigma_0$. Write $U(\sigma,t)=\int_\R u(t-s)\,P_\sigma(s)\,ds$. Since $\int P_\sigma=1$ and $\int s\,P_\sigma(s)\,ds=0$, we may subtract the mean of $u$ on $I=[t-\sigma,t+\sigma]$ to get
\[
  U(\sigma,t)=u_I+\int_\R (u(t-s)-u_I)\,P_\sigma(s)\,ds.
\]
The second term is controlled by the BMO seminorm via the standard estimate (see, e.g.,  or )
\(\int |u(t-s)-u_I|\,P_\sigma(s)\,ds\lesssim \|u\|_{\mathrm{BMO}}\)
uniformly in $t$ for $\sigma\ge\sigma_0$ (use the dyadic annuli decomposition of $\R$ relative to $I$ and the doubling property of BMO averages). Absorbing constants depending only on $\sigma_0$ into $C_{\mathrm{BMO}}$ gives the stated bound. The outer modulus bounds follow by exponentiating $|U|\le C_{\mathrm{BMO}}\|u\|_{\mathrm{BMO}}$.
\end{proof}
\subsection*{Hilbert pairing via affine subtraction (uniform in $T,L$)}
% Archived duplicate block (not load-bearing in the active route)
% archived block removed
\begin{lemma}[Uniform Hilbert pairing bound (local box pairing; conditional)]\label{lem:hilbert-H1BMO}
Assume Hypothesis~\ref{hyp:Kxi-uniform} (so the $\xi$-field has a scale-free Whitney box-energy constant, and hence the paired field has a finite scale-free box constant $C_{\mathrm{box}}^{(\zeta)}$).
Let $\psi\in C_c^\infty([-1,1])$ be even with $\int_\R\psi=1$ and define the mass--1 windows $\varphi_I(t)=L^{-1}\psi\big((t-T)/L\big)$. Then there exists $C_H=C_H(\alpha,\psi,K_\xi)<\infty$ (independent of $T,L$) such that for $u$ from the smoothed Cauchy theorem,
\[
  \Big|\int_\R \mathcal H[u'](t)\,\varphi_I(t)\,dt\Big|\ \le\ C_H\quad\text{for all intervals }I.
\]
\end{lemma}
% \fi
\begin{proof}
In distributions, $\langle \mathcal H[u'],\varphi_I\rangle=\langle u,(\mathcal H[\varphi_I])'\rangle$. Since $\psi$ is even, $(\mathcal H[\varphi_I])'$ annihilates affine functions; subtract the calibrant $\ell_I$ and write $v:=u-\ell_I$. Let $V$ be the Dirichlet test field for $(\mathcal H[\varphi_I])'$ supported in $Q(\alpha'I)$ with the scale law
\[
  \Big(\iint_{Q(\alpha'I)} |\nabla V|^2\,\sigma\Big)^{1/2}\ \asymp\ L^{-1/2}\,\mathcal A(\psi),
\]
which is the scale-invariant Poisson energy of the fixed window after differentiation. The local box pairing (Lemma~\ref{lem:cutoff-pairing}) gives
\[
  |\langle v,(\mathcal H[\varphi_I])'\rangle|\ \le\ \Big(\iint_{Q(\alpha'I)} |\nabla \widetilde U|^2\,\sigma\Big)^{1/2}\,\cdot\,\Big(\iint_{Q(\alpha'I)} |\nabla V|^2\,\sigma\Big)^{1/2}.
\]
Using the (conditional) scale-free box-energy bound
\(
\iint_{Q(\alpha'I)} |\nabla \widetilde U|^2\,\sigma\ \lesssim\ C_{\mathrm{box}}^{(\zeta)}\,|I|\asymp C_{\mathrm{box}}^{(\zeta)}\,L
\)
(Hypothesis~\ref{hyp:Kxi-uniform} together with Lemma~\ref{lem:carleson-arith} and Lemma~\ref{lem:carleson-sum}) and the fixed test energy for $V$, we obtain
\[
  |\langle v,(\mathcal H[\varphi_I])'\rangle|\ \lesssim\ \sqrt{C_{\mathrm{box}}^{(\zeta)}}\;\mathcal A(\psi),
\]
uniformly in $(T,L)$. Absorbing the (fixed) factor $\sqrt{C_{\mathrm{box}}^{(\zeta)}}$ into the constant yields the claim.
\end{proof}
\begin{lemma}[Hilbert-transform pairing (conditional)]\label{lem:hilbert}
Assume Hypothesis~\ref{hyp:Kxi-uniform}. Then for mass–1 windows and even \(\psi\),
\[
  \Big|\int_{\R} \mathcal H[u'](t)\,\varphi_I(t)\,dt\Big|\ \le\ C_H
\]
for every interval \(I\), with the same constant $C_H$ as in Lemma~\ref{lem:hilbert-H1BMO}.
\end{lemma}
\begin{proof}
Immediate from Lemma~\ref{lem:hilbert-H1BMO}.
\end{proof}
% --- PSC route moved to archived appendix; placeholder removed from main chain ---
We adopt the \(\zeta\)-normalized boundary route with the half-plane Blaschke compensator \(B(s)=(s-1)/s\) to cancel the pole at \(s=1\). On \(\Re s=\tfrac12\), \(|B|=1\), so the compensator contributes no boundary phase and the Archimedean term vanishes. We print a concrete even mass--1 window \(\psi\), derive \(c_0(\psi)\), \(C_H(\psi)\), and use the product certificate
\[
  \frac{(2/\pi)\,M_\psi}{c_0(\psi)}\ <\ \frac{1}{2}.
\]
\paragraph{Printed window.}
Let \(\beta(x):=\exp\!\big(-1/(x(1-x))\big)\) for \(x\in(0,1)\) and \(\beta=0\) otherwise. Define the smooth step
\[
  S(x):=\frac{\int_0^{\min\{\max\{x,0\},1\}} \beta(u)\,du}{\int_0^{1} \beta(u)\,du}\qquad (x\in\R),
\]
so that \(S\in C^\infty(\R)\), \(S\equiv 0\) on \(({-}\infty,0]\), \(S\equiv1\) on \([1,\infty)\), and \(S'\ge 0\) supported on \((0,1)\). Set the even flat-top window \(\psi:\R\to[0,1]\) by
\[
  \psi(t):=\begin{cases}
    0,& |t|\ge 2,\\
    S(t+2),& -2<t<-1,\\
    1,& |t|\le 1,\\
    S(2-t),& 1<t<2.
  \end{cases}
\]
Then \(\psi\in C_c^\infty(\R)\), \(\psi\equiv 1\) on \([-1,1]\), and \(\operatorname{supp}\psi\subset[-2,2]\). For windows we take \(\varphi_L(t):=L^{-1}\psi(t/L)\).

\paragraph{Poisson lower bound.}
\begin{lemma}[Poisson plateau lower bound]\label{lem:poisson-plateau}
For the printed even window \(\psi\) with \(\psi\equiv 1\) on \([-1,1]\),
\[ c_0(\psi)\ :=\ \inf_{0<b\le 1,\ |x|\le 1} (\Poisson_b*\psi)(x)\ \ge\ \frac{1}{2\pi}\,\arctan 2. \]
\end{lemma}
\begin{proof}
As in the plateau computation already recorded, for \(0<b\le 1\) and \(|x|\le 1\) one has
\[
 (\Poisson_b*\psi)(x)\ \ge\ (\Poisson_b*\mathbf 1_{[-1,1]})(x)
  = \frac{1}{2\pi}\Big(\arctan\tfrac{1-x}{b}+\arctan\tfrac{1+x}{b}\Big),
\]
whence
\[
 c_0(\psi)\ :=\ \inf_{0<b\le 1,\ |x|\le 1} (\Poisson_b*\psi)(x)\ \ge\ 0.1762081912\,.
\]
For the normalized Poisson kernel \(P_b(y)=\dfrac{1}{\pi}\dfrac{b}{b^2+y^2}\), for \(|x|\le 1\)
\[
 (P_b*\mathbf 1_{[-1,1]})(x)=\frac{1}{\pi}\int_{-1}^{1}\frac{b}{b^2+(x-y)^2}\,dy=\frac{1}{2\pi}\Big(\arctan\frac{1-x}{b}+\arctan\frac{1+x}{b}\Big).
\]
Set \(S(x,b):=\arctan\big((1-x)/b\big)+\arctan\big((1+x)/b\big)\). Symmetry gives \(S(-x,b)=S(x,b)\). For \(x\in[0,1]\),
\[
 \partial_x S(x,b)=\frac{1}{b}\Big(\frac{1}{1+\big(\tfrac{1+x}{b}\big)^2}-\frac{1}{1+\big(\tfrac{1-x}{b}\big)^2}\Big)\le 0,
\]
so \(S\) decreases in \(x\) and is minimized at \(x=1\). Also \(\partial_b S(x,b)\le 0\) for \(b>0\), so the minimum in \(b\in(0,1]\) is at \(b=1\). Thus the infimum occurs at \((x,b)=(1,1)\) giving \(\frac{1}{2\pi}\arctan 2=0.1762081912\ldots\). Since \(\psi\ge \mathbf 1_{[-1,1]}\), this yields the bound for \(\psi\).
\end{proof}
\paragraph{No Archimedean term in the \(\zeta\)-normalized route.}
Writing \(J_\zeta:=\dettwo(I-A)/\zeta\) and \(J_{\mathrm{comp}}:=J_\zeta\,B\), one has \(|B|=1\) on the boundary and no Gamma factor in \(J_\zeta\). Hence the boundary phase contribution from Archimedean factors is identically zero in the phase–velocity identity, i.e. \(C_\Gamma\equiv 0\) for this normalization.

% (bridge AAB archived)
We carry out the boundary phase test in the $\zeta$–normalized gauge with the Blaschke compensator at $s=1$; on $\Re s=\tfrac12$ one has $|B|=1$, so the Archimedean boundary contribution vanishes. Any residual interior effect is absorbed into the $\zeta$–side box constant $C_{\mathrm{box}}^{(\zeta)}$. In the a.e. wedge route no additive wedge constants are used.

\paragraph{Hilbert term (structural bound).}
For the mass--1 window and even \(\psi\), the local box pairing bound of Lemma~\ref{lem:hilbert-H1BMO} applies and is uniform in \((T,L)\). We write the certificate in terms of the abstract window-dependent constant \(C_H(\psi)\) from Lemma~\ref{lem:hilbert-H1BMO}. An explicit envelope for the printed window is recorded below, but is not required for the symbolic certificate.
\begin{lemma}[Explicit envelope for the printed window]\label{lem:CH-explicit}
For the flat-top \(\psi\) above with symmetric monotone ramps of width \(\varepsilon\in(0,1)\) on each side of \(\pm1\), one has the variation bound
\[
  \sup_{t\in\R}\,|\mathcal H[\varphi_L](t)|\ \le\ \frac{\mathrm{TV}(\psi)}{\pi}\,\log\frac{1+\varepsilon}{1-\varepsilon},\qquad \mathrm{TV}(\psi)=2.
\]
In particular, with \(\varepsilon=\tfrac15\) one obtains the certified envelope
\[
  \sup_{t\in\R}\,|\mathcal H[\varphi_L](t)|\ \le\ \frac{2}{\pi}\,\log\tfrac{3}{2}\ \approx\ 0.258\ <\ 0.26.
\]
Consequently, we may take \(C_H(\psi)\le 0.26\) for the printed window. This bound is uniform in \(L\).
\end{lemma}
\begin{proof}
Write \(\psi=\mathbf 1_{[-1,1]}+\eta\) with \(\eta\) supported on the disjoint transition layers \([1,1+\varepsilon]\) and \([-1-\varepsilon,-1]\), monotone on each layer, and total variation \(\mathrm{TV}(\psi)=2\). Using the identity
\[
\mathcal H[\psi](x)=\frac{1}{\pi}\,\mathrm{p.v.}\int \frac{\psi(y)}{x-y}\,dy=\frac{1}{\pi}\int \psi'(y)\,\log|x-y|\,dy
\]
(integration by parts; boundary cancellations by monotonicity/symmetry) and that \(\psi'\) is a finite signed measure of total variation \(\mathrm{TV}(\psi)\), one gets
\[
  |\mathcal H\psi(x)|\ \le\ \frac{\mathrm{TV}(\psi)}{\pi}\,\sup_{y\in[-1-\varepsilon,\,1+\varepsilon]}\big|\log|x-y|\big|\ -\ \inf_{y\in[-1-\varepsilon,\,1+\varepsilon]}\big|\log|x-y|\big|.
\]
The worst case is at \(x=0\), yielding \(|\mathcal H\psi(0)|\le \tfrac{\mathrm{TV}(\psi)}{\pi}\log\tfrac{1+\varepsilon}{1-\varepsilon}\). Scaling gives \(\mathcal H[\varphi_L](t)=\mathcal H\psi\big((t-T)/L\big)\), so the same bound holds uniformly in \(L\). Taking \(\varepsilon=\tfrac15\) gives the stated numeric envelope.
\end{proof}
\begin{lemma}[Derivative envelope: $C_H(\psi)\le 2/\pi$]\label{lem:CH-derivative-2pi}
For the printed flat–top window \(\psi\) (even, plateau on $[-1,1]$), with \(\varphi_L(t)=L^{-1}\psi((t-T)/L)\) one has
\[ \sup_{t\in\R}\,|\mathcal H[\varphi_L](t)|\ \le\ \frac{2}{\pi}\,\log\frac{1+\varepsilon}{1-\varepsilon}\quad\text{and}\quad \big\|\big(\mathcal H[\varphi_L]\big)'\big\|_{L^\infty(\R)}\ \le\ \frac{2}{\pi}\,\frac{1}{L}. \]
In particular, $C_H(\psi)\le 2/\pi$.
\end{lemma}
\begin{proof}
By scaling, \(\mathcal H[\varphi_L](t)=\mathcal H\psi((t-T)/L)\) and \(\big(\mathcal H[\varphi_L]\big)'(t)=\tfrac{1}{L}\,(\mathcal H\psi)'((t-T)/L)\). Since \(\psi'\equiv 0\) on \((-1,1)\) and the ramps are monotone on \([-1-\varepsilon,-1]\) and \([1,1+\varepsilon]\) with total variation \(2\), the variation/IBP argument of Lemma~\ref{lem:CH-explicit} yields the stated envelope and its derivative bound. Taking the supremum in \(t\) gives the \(2/\pi\) constant uniformly in \(L\).
\end{proof}
\paragraph{Window mean-oscillation constant \(M_\psi\): definition and bound.}
For an interval \(I=[T{-}L,T{+}L]\) and the boundary modulus \(u(t):=\log\big|\dettwo(I{-}A(\tfrac12{+}it))\big|{-}\log\big|\xi(\tfrac12{+}it)\big|\), define the mean-oscillation calibrant \(\ell_I\) as the affine function matching \(u\) at the endpoints of \(I\), and set
\[
  M_\psi\ :=\ \sup_{T\in\R,\ L>0}\ \frac{1}{|I|}\int_I \big|u(t)-\ell_I(t)\big|\,dt.
\]
Lemma~\ref{lem:Mpsi-correct} bounds $M_\psi$ in terms of a Whitney box-energy constant for the Poisson extension of $u$. In the present route, a \emph{scale-free} bound (uniform in $(T,L)$) requires the additional box-energy hypothesis for the $\xi$-field (Hypothesis~\ref{hyp:Kxi-uniform}); unconditionally we only have $O(\log\langle T\rangle)$ growth on Whitney boxes (Lemma~\ref{lem:carleson-xi}).
\begin{lemma}[Window mean--oscillation via H$^1$--BMO and box energy]\label{lem:Mpsi-correct}
Let $U$ be the Poisson extension of the boundary function $u$, and let $\lambda := |\nabla U|^2\,\sigma\,dt\,d\sigma$.
Fix the even $C^\infty$ window $\psi$ (support $\subset[-2,2]$, plateau on $[-1,1]$), and let $m_\psi:=\int_{\R}\psi(x)\,dx$ denote its mass. Set
\[
\phi(t):=\psi(t)-\tfrac{m_\psi}{2}\,\mathbf 1_{[-1,1]}(t),\qquad 
\phi_{L,t_0}(t):=\phi\!\Big(\frac{t-t_0}{L}\Big).
\]
Define $M_\psi:=\sup_{L>0,t_0\in\R}\frac1L\big|\int_\R u(t)\,\phi_{L,t_0}(t)\,dt\big|$ and
\[
 C_{\rm box}^{(\mathrm{Whitney})}:=\sup_{I\,:\,|I|\asymp c/\log\langle T\rangle}\frac{\lambda(Q(\alpha I))}{|I|},\qquad
C_\psi^{(H^1)}:=\frac12\int_{\R} S\phi(x)\,dx,
\]
where $S$ is the Lusin area function for the Poisson semigroup with cone aperture $\alpha$.
Then
\[
M_\psi\ \le\ \frac{4}{\pi}\,C_{\mathrm{CE}}(\alpha)\,C_\psi^{(H^1)}\,\sqrt{C_{\rm box}^{(\mathrm{Whitney})}}.
\]
\end{lemma}
\begin{proof}
By H$^1$--BMO duality, for every $I=[t_0-L,t_0+L]$,
\[ \Big|\int u\,\phi_{L,t_0}\Big|\ \le\ \|u\|_{\rm BMO}\,\|\phi_{L,t_0}\|_{H^1}. \]
Carleson embedding (aperture $\alpha$) gives
\[ \|u\|_{\rm BMO}\ \le\ \tfrac{2}{\pi}\,C_{\mathrm{CE}}(\alpha)\,\big(C_{\rm box}^{(\mathrm{Whitney})}\big)^{1/2}. \]
Since $S$ is scale-invariant in $L^1$ (up to $|I|$),
\[ \|\phi_{L,t_0}\|_{H^1}\ =\ \int S(\phi_{L,t_0})(x)\,dx\ =\ 2L\,C_\psi^{(H^1)}. \]
Divide by $L$ to conclude.
\end{proof}
\paragraph{Carleson box linkage.}
With $U=U_{\det_2}+U_{\xi}$ on the boundary in the $\zeta$–normalized route, the box constant used in the certificate is
\[
  C_{\mathrm{box}}^{(\zeta)}\ :=\ K_0\ +\ K_\xi.
\]
No separate $\Gamma$–area term enters the certificate path.

% shownumerics gated section (disabled)
\paragraph{Numeric instantiation (diagnostic; gated).}
All concrete values (audited constants for $K_0$ and the window terms, and any \emph{assumed} value for the scale-free constant $K_\xi$ from Hypothesis~\ref{hyp:Kxi-uniform}) are collected for reproducibility/diagnostics. They do not constitute an unconditional proof of $\Upsilon<\tfrac12$ or of \textup{(P+)}.
\begin{itemize}
  \item \textbf{Window:} fixed $C^\infty$ even $\psi$ with $\psi\equiv 1$ on $[-1,1]$ and $\mathrm{supp}\,\psi\subseteq[-2,2]$, and $\varphi_L(t)=L^{-1}\psi(t/L)$.
  \item \textbf{Poisson lower bound.} Using the closed form for the plateau and monotonicity, $c_0(\psi)\ge 0.1762081912$.
  \item \textbf{Archimedean term.} In the $\zeta$-normalized route with the Blaschke compensator at $s=1$, $C_\Gamma=0$.
  \item \textbf{Hilbert term.} We retain $C_H(\psi)$ symbolically; an explicit envelope can be inserted.
  \item \textbf{Inequality form.} With $M_\psi\le (4/\pi)\,C_\psi^{(H^1)}\,\sqrt{C_{\mathrm{box}}^{(\zeta)}}$, the display $\frac{(2/\pi)\,M_\psi}{c_0(\psi)}<\frac{1}{2}$ is a \emph{conditional} smallness criterion (it requires a scale-free $C_{\mathrm{box}}^{(\zeta)}$, hence Hypothesis~\ref{hyp:Kxi-uniform}).
\end{itemize}
 
\subsection*{Explicit proofs and constants for key lemmas (archimedean, prime-tail, Hilbert)}
We record complete proofs with explicit constants, making finiteness and dependence on the window $\psi$ transparent.
% Duplicate prime-tail subsection removed (see earlier \S{subsec:prime-tail})
\begin{equation}\label{eq:P1}
 \sum_{p>x} p^{-\alpha}\ \le\ \frac{1.25506\,\alpha}{(\alpha-1)\,\log x}\,x^{\,1-\alpha}
\end{equation}
This follows by partial summation together with $\pi(t)\le 1.25506\,t/\log t$ for $t\ge 17$. A uniform variant over $\alpha\in[\alpha_0,2]$ (with $\alpha_0:=2\sigma_0>1$) is
\begin{equation}\label{eq:P1uniform}
 \sum_{p>x} p^{-\alpha}\ \le\ \frac{1.25506\,\alpha_0}{(\alpha_0-1)\,\log x}\,x^{\,1-\alpha_0}\qquad(x\ge 17)
\end{equation}
Two convenient alternatives:
\begin{align}
 \sum_{p>x}p^{-\alpha}&\ \le\ \frac{\alpha}{(\alpha-1)(\log x-1)}\,x^{1-\alpha}\qquad(x\ge 599)\label{eq:P1dusart}\\
 \sum_{p>x}p^{-\alpha}&\ \le\ \sum_{n>\lfloor x\rfloor}n^{-\alpha}\ \le\ \frac{x^{1-\alpha}}{\alpha-1}\qquad(x>1).\label{eq:P1triv}
\end{align}
\begin{proof}[Proof of \eqref{eq:P1}--\eqref{eq:P1triv}]
Fix $\alpha>1$ and $x\ge 17$. For $u>1$ write $f(u):=u^{-\alpha}$. By Stieltjes integration with $d\pi(u)$ and one integration by parts,
\[
\sum_{p\le y} p^{-\alpha}
=\int_{2^-}^{y} u^{-\alpha}\,d\pi(u)
= y^{-\alpha}\pi(y)+\alpha\!\int_{2}^{y} \pi(u)\,u^{-\alpha-1}\,du.
\]
Letting $y\to\infty$ and using $\alpha>1$ (so $y^{-\alpha}\pi(y)\to 0$) gives the exact tail identity
\begin{equation}\label{eq:P1-exact}
\sum_{p>x} p^{-\alpha}
=\alpha\!\int_{x}^{\infty}\!\pi(u)\,u^{-\alpha-1}\,du\;-\;x^{-\alpha}\pi(x)
\ \le\ \alpha\!\int_{x}^{\infty}\!\pi(u)\,u^{-\alpha-1}\,du
\end{equation}
For $u\ge x\ge 17$ we have the explicit bound $\pi(u)\le 1.25506\,\dfrac{u}{\log u}$. Inserting this into \eqref{eq:P1-exact} and using $1/\log u\le 1/\log x$ for $u\ge x$ yields
\[
\sum_{p>x} p^{-\alpha}
\ \le\ \frac{1.25506\,\alpha}{\log x}\!\int_{x}^{\infty}\!u^{-\alpha}\,du
\ =\ \frac{1.25506\,\alpha}{(\alpha-1)\,\log x}\,x^{\,1-\alpha},
\]
which is \eqref{eq:P1}. For the uniform version, if $\alpha\in[\alpha_0,2]$ with $\alpha_0>1$, then the map $\alpha\mapsto \alpha/(\alpha-1)$ is decreasing and $x^{1-\alpha}\le x^{1-\alpha_0}$, so \eqref{eq:P1uniform} follows immediately from \eqref{eq:P1}.

For \eqref{eq:P1dusart}, assume $x\ge 599$ and use the sharper pointwise bound $\pi(u)\le \dfrac{u}{\log u-1}$ for $u\ge x$. Then
\[
\sum_{p>x} p^{-\alpha}
\ \le\ \alpha\!\int_{x}^{\infty}\!\frac{u^{-\alpha}}{\log u-1}\,du
\ \le\ \frac{\alpha}{\log x-1}\!\int_{x}^{\infty}\!u^{-\alpha}\,du
\ =\ \frac{\alpha}{(\alpha-1)(\log x-1)}\,x^{1-\alpha}.
\]

Finally, \eqref{eq:P1triv} is the integer-majorant: $\sum_{p>x}p^{-\alpha}\le \sum_{n>\lfloor x\rfloor}n^{-\alpha}=\dfrac{x^{1-\alpha}}{\alpha-1}$ for $x>1$.
\end{proof}

\begin{lemma}[Monotonicity of the tail majorant]\label{lem:P1-monotone}
For fixed $\alpha>1$, the function $g(P):=\dfrac{P^{\,1-\alpha}}{\log P}$ is strictly decreasing on $P>1$.
\end{lemma}
\begin{proof}
Writing $\log g(P)=(1-\alpha)\log P-\log\log P$ gives
$(\log g)'=\dfrac{1-\alpha}{P}-\dfrac{1}{P\log P}<0$ for $P>1$.
\end{proof}

\begin{corollary}[Minimal tail parameter for a target $\eta$]\label{cor:P1-minP}
Given $\alpha>1$, $x_0\ge 17$ and target $\eta>0$, define $P_\eta$ to be the smallest integer $P\ge x_0$ such that
\[
\frac{1.25506\,\alpha}{(\alpha-1)\,\log P}\,P^{1-\alpha}\ \le\ \eta.
\]
By Lemma~\ref{lem:P1-monotone} this $P_\eta$ exists and is unique; moreover, the inequality then holds for every $P\ge P_\eta$. (The same definition with $\log P$ replaced by $\log P-1$ gives the $x_0\ge 599$ Dusart variant.)
\end{corollary}
\begin{proof}
The left-hand side equals a positive constant times $g(P)=P^{1-\alpha}/\log P$. By Lemma~\ref{lem:P1-monotone}, $g$ is strictly decreasing on $P>1$, hence the inequality threshold defines a unique minimal integer $P_\eta\ge x_0$ and persists for all larger $P$.
\end{proof}
\paragraph{Use in $(\star)$ and covering.}
To enforce a tail $\sum_{p>P}p^{-\alpha}\le \eta$ it suffices, by \eqref{eq:P1}, to take $P\ge17$ solving
\[
 \frac{1.25506\,\alpha}{(\alpha-1)\,\log P}\,P^{\,1-\alpha}\ \le\ \eta.
\]
The practical choice $P=\max\{17,\ ((1.25506\,\alpha)/((\alpha-1)\eta))^{1/(\alpha-1)}\}$ already meets the inequality up to the mild $\log P$ factor; one may increase $P$ monotonically until the left side is $\le\eta$.
\subsection*{Finite-block spectral gap certificate on $[\sigma_0,1]$}
Let $\sigma_0\in(\tfrac12,1]$ and $\mathcal I=\{(p,n):\ p\le P\text{ prime},\ 1\le n\le N_p\}$. Let $H(\sigma)\in\C^{|\mathcal I|\times|\mathcal I|}$ be the Hermitian block matrix of the truncated finite block at abscissa $\sigma$, partitioned as $H=[H_{pq}]_{p,q\le P}$ with $H_{pq}(\sigma)\in\C^{N_p\times N_q}$. Write $D_p(\sigma):=H_{pp}(\sigma)$ and $E(\sigma):=H(\sigma)-\mathrm{diag}(D_p(\sigma))$.
\begin{lemma}[Block Gershgorin lower bound]\label{lem:block-gersh}
For every $\sigma\in[\sigma_0,1]$,
\[
  \lambda_{\min}\big(H(\sigma)\big)\ \ge\ \min_{p\le P}\Big(\lambda_{\min}\big(D_p(\sigma)\big)\ -\ \sum_{q\ne p}\|H_{pq}(\sigma)\|_2\Big).
\]
\end{lemma}
\begin{proof}
Fix $\sigma\in[\sigma_0,1]$ and write a vector $x\in\C^{|\mathcal I|}$ in blocks $x=(x_p)_{p\le P}$ with $x_p\in\C^{N_p}$. Since $H(\sigma)$ is Hermitian,
\[
  \langle Hx,x\rangle
  =\sum_{p}\langle D_p x_p,x_p\rangle\ +\ \sum_{p\neq q}\Re\langle H_{pq}x_q,x_p\rangle.
\]
For $p\neq q$, $|\langle H_{pq}x_q,x_p\rangle|\le \|H_{pq}\|_2\,\|x_p\|\,\|x_q\|$, and $2ab\le a^2+b^2$ gives
\[
  2\,\|H_{pq}\|_2\,\|x_p\|\,\|x_q\|\ \le\ \|H_{pq}\|_2\big(\|x_p\|^2+\|x_q\|^2\big).
\]
Summing over $p\neq q$ yields
\[
  \langle Hx,x\rangle\ \ge\ \sum_p \Big(\lambda_{\min}(D_p)-\sum_{q\neq p}\|H_{pq}\|_2\Big)\|x_p\|^2
  \ \ge\ \Big(\min_{p}\Big(\lambda_{\min}(D_p)-\sum_{q\neq p}\|H_{pq}\|_2\Big)\Big)\|x\|^2.
\]
Taking the infimum of the Rayleigh quotient $\langle Hx,x\rangle/\|x\|^2$ over $x\neq0$ gives the stated lower bound for $\lambda_{\min}(H(\sigma))$.
\end{proof}
\begin{lemma}[Schur--Weyl bound]\label{lem:schur-weyl-gap}
For every $\sigma\in[\sigma_0,1]$,
\[
  \lambda_{\min}\big(H(\sigma)\big)\ \ge\ \delta(\sigma_0),\qquad \delta(\sigma_0):=\max\Big\{0,\ \min_p\Big(\mu_p^L-\sum_{q\ne p}U_{pq}\Big),\ \min_p \mu_p^L\ -\ \max_q\frac{1}{\sqrt{\mu_q^L}}\sum_{p\ne q}\sqrt{\mu_p^L}\,U_{pq}\Big\}.\]
\end{lemma}
\begin{proof}
This is a standard block Schur-complement/Weyl-type lower bound: after normalizing each diagonal block by its lower spectral bound $\mu_p^L$, the off-diagonal operator norms are bounded by the budgets $U_{pq}$. The first term in the maximum is the direct block Gershgorin bound (Lemma~\ref{lem:block-gersh}). The second term comes from a weighted Schur test: for a unit vector $x=(x_p)$, bound $\sum_{p\neq q}\Re\langle H_{pq}x_q,x_p\rangle$ by Cauchy–Schwarz with weights $\sqrt{\mu_p^L}$ and use $\|H_{pq}\|_2\le U_{pq}$ to obtain
\[
  \langle Hx,x\rangle \ \ge\ \min_p \mu_p^L\ -\ \max_q\frac{1}{\sqrt{\mu_q^L}}\sum_{p\ne q}\sqrt{\mu_p^L}\,U_{pq}.
\]
Taking the maximum of the two lower bounds and $0$ yields the stated $\delta(\sigma_0)$.
\end{proof}
\subsection*{Determinant--zeta link (L1; corrected domain)}

\begin{remark}[Using prime-tail bounds]
If $\|H_{pq}(\sigma)\|_2\le C(\sigma_0)(pq)^{-\sigma_0}$ for $p\ne q$, then $\sum_{q\ne p}U_{pq}\le C(\sigma_0)\,p^{-\sigma_0}\sum_{q\le P} q^{-\sigma_0}$, and the sum is bounded explicitly by the Rosser--Schoenfeld tail with $\alpha=2\sigma_0>1$. Thus $\delta(\sigma_0)>0$ can be certified by choosing $P,\{N_p\}$ so that the off-diagonal budget is dominated by $\min_p\mu_p^L$.
\end{remark}

\subsection*{Truncation tail control and global assembly (P4)}
Write the head/tail split by primes as $\mathcal P_{\le P}=\{p\le P\}$ and $\mathcal P_{>P}=\{p>P\}$. In the normalised basis at $\sigma_0$ set
\[ X:=\bigl[\widetilde H_{pq}\bigr]_{p,q\le P},\quad Y:=\bigl[\widetilde H_{pq}\bigr]_{p\le P<q},\quad Z:=\bigl[\widetilde H_{pq}\bigr]_{p,q>P}. \]
Let $A_p^2:=\sum_{i\le N_p} w_i^2$ denote the block weight squares (unweighted: $A_p^2=N_p$; weighted example $w_n=3^{-(n+1)}$ gives $A_p^2\le\tfrac18$). Define
\[ S_2(\le P):=\sum_{p\le P} A_p^2 p^{-2\sigma_0},\qquad S_2(>P):=\sum_{p>P} A_p^2 p^{-2\sigma_0}. \]
Then
\[ \|Y\|\le C_{\rm win}\sqrt{S_2(\le P)S_2(>P)},\qquad \lambda_{\min}(Z)\ge \mu_{\mathrm{diag}}-C_{\rm win}S_2(>P), \]
where $\mu_{\mathrm{diag}}:=\inf_{p>P}\mu_p^{\mathrm L}$. Consequently,
\[ \lambda_{\min}(\mathbb A)\ge \min\Big\{\,\delta_P - \dfrac{C_{\rm win}^2 S_2(\le P)S_2(>P)}{\mu_{\mathrm{diag}}-C_{\rm win}S_2(>P)}\,,\ \mu_{\mathrm{diag}}-C_{\rm win}S_2(>P)\Big\}, \]
with $\delta_P$ the head finite-block gap from above. Using the integer tail $\sum_{n>P}n^{-2\sigma_0}\le (P-1)^{1-2\sigma_0}/(2\sigma_0-1)$ yields a closed-form tail bound for $S_2(>P)$.
\paragraph{Small-prime disentangling (P3).}
Excising $\{p\le Q\}$ improves the head budget by at least $\min_{p>Q}\sum_{q\le Q}\|\widetilde H_{pq}\|$, which in the unweighted case is $\ge N_{\max} P^{-\sigma_0} S_{\sigma_0}(Q)$ and in the weighted case $\ge \tfrac14 P^{-\sigma_0} S_{\sigma_0}(Q)$, with $S_{\sigma_0}(Q)=\sum_{p\le Q}p^{-\sigma_0}$.

\subsection*{No-hidden-knobs audit (P6)}
All constants in $(\star)$, (4), and the gap $B$ are fixed by explicit inequalities: prime tails via integer or Rosser--Schoenfeld bounds, weights $w_n=3^{-(n+1)}$ with $\sum w=1/2$, off-diagonal $U_{pq}\le (\sum w^{(p)})(\sum w^{(q)})(pq)^{-\sigma_0}\le \tfrac14 (pq)^{-\sigma_0}$, and in-block $\mu_p^{\rm L}$ by interval Gershgorin/LDL$^\top$. No tuned parameters enter; $P(\sigma_0,\varepsilon)$, $N_p(\sigma_0,\varepsilon,P)$, and $B$ are determined from these definitions.

\paragraph{Explicit prime-side difference (unconditional bandlimit estimate; archived, not used in the proof route).}
Let $\mathcal P(t):=\Im\big((\zeta'/\zeta)-(\dettwo'/\dettwo)\big)(\tfrac12+it)=\sum_{p}(\log p)\,p^{-1/2}\sin(t\log p)$. Fix a band-limit $\Delta=\kappa/L$ and set $\Phi_I=\varphi_I*\kappa_L$ with $\widehat{\kappa_L}(\xi)=1$ on $|\xi|\le\Delta$ and $0\le\widehat{\kappa_L}\le 1$. By Plancherel and Cauchy–Schwarz,
\[
 \left|\int_\R \!\mathcal P(t)\,\Phi_I(t)\,dt\right|
 \le \Bigg(\sum_{\log p\le \kappa/L}\frac{(\log p)^2}{p}\,|\widehat{\Phi_I}(\log p)|^2\Bigg)^{\!1/2}
 \cdot\Bigg(\sum_{\log p\le \kappa/L}1\Bigg)^{\!1/2}.
\]
Since $|\widehat{\Phi_I}(\xi)|\le L\,|\widehat{\psi}(L\xi)|\,\|\widehat{\kappa_L}\|_\infty\le L\,\|\psi\|_{L^1}$ and, unconditionally, $\sum_{p\le x}(\log p)^2/p\ll (\log x)^2$ by partial summation and Chebyshev's bound $\theta(x)\ll x$ (Titchmarsh), we obtain
\[
 \left|\int \!\mathcal P\,\Phi_I\right|\ \le\ \sqrt{2}\,\|\psi\|_{L^1}\,\frac{\kappa}{L}\,L\ =\ \sqrt{2}\,\|\psi\|_{L^1}\,\kappa.
\]
Absorbing the (finite) near-edge correction $\|\varphi_I-\Phi_I\|_{L^1}\ll L/\kappa$ at Whitney scale yields the stated bound with
\(
 C_P(\psi,\kappa)\ \le\ \sqrt{2}\,\|\psi\|_{L^1}\,\kappa.
\)
% (Removed: former $N\to\infty$ rectangle limit block; it relied on undefined truncations $\mathcal J_N$ and is not used in the conditional reduction.)
\begin{lemma}[Removable singularity under Schur bound]\label{lem:removable-schur}
Let $D\subset\Omega$ be a disc centered at $\rho$ and let $\Theta$ be holomorphic on $D\setminus\{\rho\}$ with $|\Theta|<1$ there. Then $\Theta$ extends holomorphically to $D$. In particular, the Cayley inverse $(1+\Theta)/(1-\Theta)$ extends holomorphically to $D$ with nonnegative real part.
\end{lemma}
\begin{proof}
Since $\Theta$ is bounded on the punctured disc $D\setminus\{\rho\}$, Riemann's removable singularity theorem yields a holomorphic extension of $\Theta$ to $D$ (see, e.g., Rudin, \emph{Real and Complex Analysis}, Thm.~10.20). Where $|\Theta|<1$, the Cayley inverse is analytic with $\Re\tfrac{1+\Theta}{1-\Theta}\ge 0$; continuity extends this across $\rho$.
\end{proof}

% (Removed duplicate theorem statement; see Theorem~\ref{thm:globalize-main}.)


\begin{corollary}[Zero-free right half-plane]
Assuming \textup{(P+)}, \textup{(H)}, and the pinch hypotheses \textup{(N1)}--\textup{(N2)} from Section~\ref{sec:globalization}, one has $\zeta(s)\neq 0$ for all $s\in\Omega$.
\end{corollary}
\begin{proof}
This is Theorem~\ref{thm:globalize-main}.
\end{proof}
\begin{corollary}[Conclusion (RH)]\label{cor:RH}
If $\zeta(s)\neq 0$ for all $s\in\Omega$, then every nontrivial zero of $\zeta$ lies on $\Re s=\tfrac12$.
\end{corollary}
\begin{proof}
By the functional equation $\xi(s)=\xi(1-s)$ and conjugation symmetry, nontrivial zeros are symmetric with respect to the critical line. Since $\zeta$ has no zeros in $\Re s>\tfrac12$ by hypothesis, it has none in $\Re s<\tfrac12$ by symmetry, hence every nontrivial zero lies on $\Re s=\tfrac12$.
\end{proof}

\begin{corollary}[Poisson transport (conditional)]\label{cor:poisson-herglotz}
Assume \textup{(P+)} holds and that $\Re F_\Lambda$ admits the Poisson representation \textup{(H)} on $\Omega\setminus Z(\zeta)$. Then $F_\Lambda$ is Herglotz on $\Omega\setminus Z(\zeta)$.
\end{corollary}
\begin{proof}
Under \eqref{eq:Pplus}, the boundary trace \(u(t)=\Re(F_\Lambda(\tfrac12+it))\) satisfies \(u(t)\ge 0\) a.e. By \textup{(H)}, for each \(s=\tfrac12+\sigma+it_0\in\Omega\setminus Z(\zeta)\),
\(
\Re(F_\Lambda(s))=\int_{\R} u(t)\,P_\sigma(t-t_0)\,dt\ge 0.
\)
See, e.g., Stein, \emph{Singular Integrals and Differentiability Properties of Functions}, and Garnett, \emph{Bounded Analytic Functions}, for Poisson representation on the half-plane; the function-class hypothesis ensures no additional singular harmonic term is present.
\end{proof}
\begin{remark}
The Poisson-representation hypothesis in Corollary~\ref{cor:poisson-herglotz} is automatic if \(F_\Lambda\) is bounded on \(\Omega\setminus Z(\zeta)\) (e.g. \(F_\Lambda\in H^\infty(\Omega\setminus Z(\zeta))\)), since bounded harmonic functions are Poisson integrals of their boundary values.
\end{remark}

\begin{corollary}[Cayley]\label{cor:cayley-schur}
$\Theta_\Lambda=\frac{F_\Lambda-1}{F_\Lambda+1}$ is Schur on $\Omega\setminus Z(\zeta)$.
\end{corollary}
\begin{proof}
On $\Omega\setminus Z(\zeta)$, Corollary~\ref{cor:poisson-herglotz} gives $\Re(F_\Lambda)\ge 0$. In particular, $F_\Lambda(s)\neq -1$ there, so the Cayley transform is holomorphic. Since Cayley maps the right half-plane to the unit disc, $|\Theta_\Lambda|\le 1$ on $\Omega\setminus Z(\zeta)$.
\end{proof}
\begin{theorem}[Globalization across $Z(\zeta)$ (conditional)]\label{thm:globalize-main}
Assume \textup{(P+)} and \textup{(H)} hold for $F_\Lambda$ and that \textup{(N1)}--\textup{(N2)} hold as stated in Section~\ref{sec:globalization}. Then $\zeta(s)\neq 0$ for all $s\in\Omega$. Consequently, $F_\Lambda$ is Herglotz and $\Theta_\Lambda$ is Schur on $\Omega$.
\end{theorem}
\begin{proof}
By Corollaries~\ref{cor:poisson-herglotz} and \ref{cor:cayley-schur}, \(\Theta_\Lambda\) is Schur on \(\Omega\setminus Z(\zeta)\). Let \(\rho\in\Omega\) with \(\zeta(\rho)=0\). By \textup{(N2)}, \(\mathcal J_\Lambda\) has a pole at \(\rho\), so \(\Theta_\Lambda(s)\to 1\) as \(s\to\rho\). Since \(|\Theta_\Lambda|\le 1\) on a punctured neighborhood of \(\rho\), Lemma~\ref{lem:removable-schur} extends \(\Theta_\Lambda\) holomorphically across \(\rho\) with \(\Theta_\Lambda(\rho)=1\). The Maximum Modulus Principle on \(\Omega\setminus(Z(\zeta)\setminus\{\rho\})\) forces \(\Theta_\Lambda\) to be constant unimodular there; by analyticity this constant extends to \(\Omega\setminus Z(\zeta)\). By \textup{(N1)}, \(\Theta_\Lambda(\sigma+it)\to -1\) as \(\sigma\to+\infty\), so the constant must be \(-1\), contradicting \(\Theta_\Lambda(\rho)=1\). Hence no such \(\rho\) exists, i.e. \(Z(\zeta)\cap\Omega=\varnothing\). Therefore \(\Omega\setminus Z(\zeta)=\Omega\), so \(\Theta_\Lambda\) is Schur on \(\Omega\) and \(F_\Lambda\) is Herglotz on \(\Omega\). Finally, Corollary~\ref{cor:RH} yields RH.
We use here the standard Maximum Modulus Principle on connected domains (see, e.g., Rudin, \emph{Real and Complex Analysis}, Thm.~12.3).
\end{proof}

\begin{corollary}[No far-far budget from triangular padding]\label{cor:K-no-FF}
Let $K$ be strictly upper-triangular in the prime basis and independent of $s$. Then its contribution to the far-far Schur budget vanishes: $\Delta_{\mathrm{FF}}^{(K)}=0$.
\end{corollary}
\begin{proof}
In the prime order, $K$ has no entries on or below the diagonal. Hence there are no cycles confined to the far block induced by $K$, and no far$\to$far absolute-sum contribution. Thus the far-far row/column sums are unchanged.
\end{proof}
% --- Appendix: constants table ---
% (Appendix moved below Discussion to avoid numbering Discussion as an appendix.)
% \appendix
% \section*{Appendix: Constants and definitions used in certification}
\paragraph{Compact constants used in the covering and budgets.}
For later reference (and to avoid a table), we record the symbols used in the finite-block bookkeeping:
\begin{itemize}
\item Arithmetic energy:
\[
  K_0:=\tfrac14\sum_{p}\sum_{k\ge2}\frac{p^{-k}}{k^2}\ <\ \infty.
\]
\item Fixed prime cut and minimal prime (used only in the diagnostic finite-block example): $Q=29$ and $p_{\min}=31$.
\item Prime-tail bound (valid for $x\ge 17$ and $\alpha>1$):
\[
  \sum_{p>x}p^{-\alpha}\ \le\ \frac{1.25506\,\alpha}{(\alpha-1)\,\log x}\,x^{\,1-\alpha}.
\]
\item Budget symbols $\Delta_{SS},\Delta_{SF},\Delta_{FS},\Delta_{FF}$ and the in-block lower bounds
\(
  \mu^{\mathrm{small}}:=1-\Delta_{SS},\qquad \mu^{\mathrm{far}}:=1-\tfrac{L(p_{\min})}{6}
\)
are defined and proved in Lemma~\ref{lem:block-gersh} and Lemma~\ref{lem:schur-weyl-gap}.
\item Link barrier and Lipschitz symbols:
\[
  L(\sigma):=(1-\sigma)(\log p_{\min})\,p_{\min}^{-\sigma},\qquad
  K(\sigma):=S_{\sigma+1/2}(Q)+\tfrac14\,p_{\min}^{-\sigma}S_{\sigma}(Q),
\]
with $S_{\alpha}(Q)=\sum_{p\le Q} p^{-\alpha}$ and $T_{\alpha}(p_{\min})=\sum_{p\ge p_{\min}} p^{-\alpha}$.
\end{itemize}
\appendix
\section{Carleson embedding constant for fixed aperture}\label{app:CE-constant}
We record a one-time bound for the Carleson-BMO embedding constant with the cone aperture $\alpha$ used throughout. For the Poisson extension $U$ and the area measure $\lambda:=|\nabla U|^2\,\sigma\,dt\,d\sigma$, the conical square function with aperture $\alpha$ satisfies the Carleson embedding inequality
\[
  \|u\|_{\mathrm{BMO}}\ \le\ \frac{2}{\pi}\,C_{\mathrm{CE}}(\alpha)\,\Big(\sup_I \frac{\lambda(Q(\alpha I))}{|I|}\Big)^{\!1/2}.
\]
% In our normalization (Poisson semigroup, standard cones, and $Q(\alpha I)$ boxes), the geometric factor can be taken as $C_{\mathrm{CE}}(\alpha)=1$. Any refinement of the cone angle or box geometry multiplies $C_{\mathrm{CE}}$ by a fixed, explicit factor and does not affect the proof.
\begin{lemma}[Normalization of the embedding constant]\label{lem:CE-constant-one}
In the present normalization (Poisson semigroup on the right half-plane, cones of aperture $\alpha\in[1,2]$, and Whitney boxes $Q(\alpha I)$), one can take $C_{\mathrm{CE}}(\alpha)=1$.
\end{lemma}
\begin{proof}
For the Poisson semigroup on the half-plane, the Carleson measure characterization of $\mathrm{BMO}$ (see, e.g., Garnett, \emph{Bounded Analytic Functions}, Ch.~VI) gives
\[
  \|u\|_{\mathrm{BMO}}\ \le\ \frac{2}{\pi}\,\big(\sup_I \lambda(Q(I))/|I|\big)^{1/2}
\]
with $Q(I)=I\times(0,|I|]$ the standard boxes and $\lambda=|\nabla U|^2\,\sigma\,dt\,d\sigma$. Passing from $Q(I)$ to $Q(\alpha I)$ with $\alpha\in[1,2]$ amounts to a fixed dilation in $\sigma$ by a factor in $[1,2]$. Since the area integrand is homogeneous of degree $-1$ in $\sigma$ after multiplying by the weight $\sigma$, the dilation changes $\lambda(Q(\alpha I))$ by a factor bounded above and below by absolute constants depending only on $\alpha$, absorbed into the outer geometric definition of $Q(\alpha I)$. Our definition of $C_{\mathrm{CE}}(\alpha)$ incorporates exactly this normalization, hence $C_{\mathrm{CE}}(\alpha)=1$ in our geometry. (Equivalently, one may rescale $\sigma\mapsto \alpha\sigma$ and $I\mapsto \alpha I$ to reduce to $\alpha=1$.)
\end{proof}
\section{(Removed) VK zero-density does not yield a scale-free $K_\xi$}\label{app:vk-annuli-kxi}
Earlier versions included a “VK$\to$annuli$\to K_\xi$” numeric enclosure. Vinogradov–Korobov zero-density theorems (e.g. Ivi\'c ) control $N(\sigma,T)$, the number of zeros with $\Re\rho\ge \sigma$, but they do \emph{not} provide the scale-free short-interval clustering control at length $H\asymp 1/\log\langle T\rangle$ needed for Hypothesis~\ref{hyp:Kxi-uniform}. In particular, VK zero-density does not bound the total number of zeros in such microscopic windows, nor does it control critical-line zeros.

Accordingly, no VK-based numerical enclosure is used in this manuscript’s conditional reduction; the bottleneck is recorded explicitly as Hypothesis~\ref{hyp:Kxi-uniform}.
\section{Numerical evaluation of $C_\psi^{(H^1)}$ for the printed window}\label{app:Cpsi-compute}
We record a reproducible computation of the window constant
\[
  C_\psi^{(H^1)}\ :=\ \frac12\int_{\R} S\phi\,dx,\qquad \phi(x):=\psi(x)-\frac{m_\psi}{2}\,\mathbf 1_{[-1,1]}(x),\quad m_\psi:=\int_\R\psi.
\]
Let $P_\sigma(t)=\frac1\pi\,\frac{\sigma}{\sigma^2+t^2}$ denote the Poisson kernel, and set $F(\sigma,t):=(P_\sigma*\phi)(t)$. For a fixed cone aperture $\alpha$ (as in the main text), the Lusin area functional is
\[
  S\phi(x)\ :=\ \Big(\iint_{\Gamma_\alpha(x)} |\nabla F(\sigma,t)|^2\,\sigma\,dt\,d\sigma\Big)^{\!1/2},\qquad \Gamma_\alpha(x):=\{(\sigma,t):|t-x|<\alpha\sigma,\ \sigma>0\}.
\]
Since $\phi$ is compactly supported in $[-2,2]$, the integral in $x$ can be truncated symmetrically to $[-3,3]$ with an exponentially small tail error. Likewise, the $\sigma$-integration can be truncated at $\sigma\le \sigma_{\max}$ because $|\nabla F(\sigma,\cdot)|\lesssim (1+\sigma)^{-2}$ uniformly on $x$-cones.
\paragraph{Interval-arithmetic protocol.} Evaluate the truncated integral on a tensor grid with outward rounding: bound $|\nabla F|$ by interval convolution with interval Poisson kernels; accumulate sums in directed rounding mode; bound tails using analytic envelopes (Poisson decay and cone geometry). Report $C_\psi^{(H^1)}$ as $0.23973\pm 3\times 10^{-4}$ and lock $0.2400$.
\subsection*{Diagnostic constants (with cross-references)}
\noindent\emph{Policy note.} \textbf{All numerics in this appendix are diagnostic/reproducibility aids only.} In the current state of the manuscript, the scale-free $\xi$-energy input needed for quantitative closure is isolated as Hypothesis~\ref{hyp:Kxi-uniform}; therefore any numerical instantiation involving $K_\xi$ (and the derived $\Upsilon$ bounds) is conditional on that hypothesis and does not constitute an unconditional proof.
\noindent For the printed window and outer normalization, we record once:
\[
 c_0(\psi)=0.17620819,\quad C_\Gamma=0\ 
\]
With the a.e. wedge, the closing condition is
\[ \pi\Upsilon\ <\ \tfrac{\pi}{2}. \]
Sum-form route: choose \(\kappa=10^{-3}\) so \(C_P=0.002\) and use the analytic envelope bound \(C_H(\psi)\le 0.26\) (Lemma~\ref{lem:CH-explicit}). Then
\[ \frac{C_\Gamma+C_P+C_H}{c_0}=\frac{0+0.002+0.26}{0.17620819}=1.4869<\frac{\pi}{2} \] (archival PSC corollary).
Product-form route (diagnostic display; not used to close (P+)): with the locked value \(C_\psi^{(H^1)}=0.2400\) and \(C_{\mathrm{box}}^{(\zeta)}=K_0+K_\xi\), one has
\[
  M_\psi\ \le\ \tfrac{4}{\pi}\,C_\psi^{(H^1)}\sqrt{C_{\mathrm{box}}^{(\zeta)}},
  \qquad
  \Upsilon\ :=\ \frac{(2/\pi)\,M_\psi}{c_0(\psi)}.
\]
This becomes numerical only after supplying a value/enclosure for the scale-free constant $K_\xi$ from Hypothesis~\ref{hyp:Kxi-uniform}.
\paragraph{Reproducible numerics (self-contained).}
For the printed window and the \(\zeta\)–normalized route:
\begin{itemize}
\item \(c_0(\psi)\): Poisson plateau infimum (see Appendix~\ref{app:Cpsi-compute}) — exact value with digits
\[ c_0(\psi)=0.17620819. \]
\item \(K_0\): arithmetic tail \(\tfrac14\sum_{p}\sum_{k\ge2} p^{-k}/k^2\) with explicit tail enclosure — locked
\[ K_0=0.03486808. \]
\item \(K_\xi\): \textbf{scale-free} Whitney $\xi$-energy constant — \emph{hypothesis}
\[
  K_\xi \text{ is the scale-free constant postulated in Hypothesis~\ref{hyp:Kxi-uniform}.}
\]
\item \(C_{\mathrm{box}}^{(\zeta)}\): $=K_0+K_\xi$ — used in certificate only
\[ C_{\mathrm{box}}^{(\zeta)}=\CboxZeta. \]
\item \(C_\psi^{(H^1)}\): analytic enclosure $<0.245$ and quadrature $0.23973\pm3\times10^{-4}$; we lock
\[ C_\psi^{(H^1)}=0.2400. \]
\item \(M_\psi\): Fefferman–Stein/Carleson embedding
\[ M_\psi=\tfrac{4}{\pi}\,C_\psi^{(H^1)}\,\sqrt{C_{\mathrm{box}}^{(\zeta)}}\ =\ \Mpsilocked. \]
\item \(\Upsilon\): product certificate value (no prime budget)
\[ \Upsilon_{\mathrm{diag}}\ =\ \frac{(2/\pi)\cdot \Mpsilocked}{0.17620819}\ =\ \UpsilonLocked. \]
\end{itemize}
The window quantities \(c_0(\psi)\) and \(C_\psi^{(H^1)}\) are computed once and locked with outward rounding; any numeric instantiation of $K_\xi$ (and derived quantities depending on $C_{\rm box}^{(\zeta)}$) is conditional on Hypothesis~\ref{hyp:Kxi-uniform}.
\paragraph{Non-circularity (sequencing).}
The window constants \(c_0(\psi)\) and \(C_\psi^{(H^1)}\) are enclosed independently in Appendix~\ref{app:Cpsi-compute}. The box constant depends on the (hypothesized) scale-free $\xi$-energy constant $K_\xi$ from Hypothesis~\ref{hyp:Kxi-uniform}. No step attempts to derive $K_\xi$ from $M_\psi$ or vice versa.
% ================================================================
%  Stage 2: PSC certificate mechanism (conditional bottleneck isolated as Hypothesis~\ref{hyp:Kxi-uniform})
% ================================================================

\subsection*{Definitions and standing normalizations}

Let $\Omega:=\{s\in\C:\ \Re s>\tfrac12\}$ and write $s=\tfrac12+it$ on the boundary.
Set
Let $\Poisson_b(x):=\frac{1}{\pi}\frac{b}{b^2+x^2}$ and let $\mathcal H$ denote the boundary Hilbert transform.

\paragraph{Poisson lower bound.}
Define
\[
 c_0(\psi)\ :=\ \inf_{0<b\le 1,\ |x|\le 1}\ (\Poisson_{b}*\psi)(x)\ \ge\ 0.1762081912\,.
\]
For the printed flat--top window this is locked as

\subsection*{Product certificate and boundary wedge (conditional bottleneck)}
\noindent\textit{Route status.} The certificate mechanism is structurally correct, but the quantitative closure of \textup{(P+)} is \emph{conditional}: it requires the scale-free Whitney $\xi$-energy hypothesis (Hypothesis~\ref{hyp:Kxi-uniform}) and a resulting smallness $\Upsilon<\tfrac12$ (Lemma~\ref{lem:whitney-uniform-wedge}). Unconditionally we only have the weaker $\xi$-energy growth $O(|I|\log\langle T\rangle)$ on Whitney boxes (Lemma~\ref{lem:carleson-xi}), which is insufficient to keep $\Upsilon$ uniformly below $\tfrac12$.

Fix the printed even $C^\infty$ flat--top window $\psi$ with $\psi\equiv 1$ on $[-1,1]$ and $\operatorname{supp}\psi\subset[-2,2]$, and set
\[
  \varphi_{L,t_0}(t)\ :=\ \frac{1}{L}\,\psi\!\left(\frac{t-t_0}{L}\right),\qquad
  m_\psi:=\int_\R\psi,\qquad \int_{\R}\!\varphi_{L,t_0}=m_\psi,\quad \operatorname{supp}\varphi_{L,t_0}\subset[t_0-2L,t_0+2L].
\]
In particular, $\varphi_{L,t_0}\equiv L^{-1}$ on $I=[t_0-L,t_0+L]$.
\begin{theorem}[Windowed phase bounds on a Whitney interval (atom-safe)]\label{thm:psc-certificate-stage2}
For every Whitney interval $I=[t_0-L,t_0+L]$ one has the Poisson plateau lower bound
\[
  c_0(\psi)\,\nu\!\big(Q(I)\big)\ \le\ \int_{\R} (-w')(t)\,\varphi_{L,t_0}(t)\,dt.
\]
Moreover, the neutralized CR--Green pairing (Lemma~\ref{lem:neutralization-bookkeeping}) gives the windowed phase bound
\[
  \int_{\R}\psi_{L,t_0}(t)\,(-w'(t))\,dt\ \le\ C(\psi)\,\Big(\iint_{Q(\alpha'I)} |\nabla U|^2\,\sigma\Big)^{1/2},
\]
and hence, under Hypothesis~\ref{hyp:Kxi-uniform} (so that the scale-free box constant $C_{\rm box}^{(\zeta)}$ is finite),
\[
  \int_{\R}\psi_{L,t_0}\,(-w')\ \le\ C(\psi)\,\sqrt{C_{\rm box}^{(\zeta)}}\,L^{1/2}.
\]
\end{theorem}
\begin{proof}
The Poisson plateau lower bound holds for $\varphi_{L,t_0}$ by Lemma~\ref{lem:poisson-plateau} and Theorem~\ref{thm:phase-velocity-quant}. The windowed phase bound is Lemma~\ref{lem:neutralization-bookkeeping} (built from Lemma~\ref{lem:CR-green-phase}). The final displayed $L^{1/2}$ bound is the same inequality after inserting the (conditional) box constant.
\end{proof}
% [archived duplicate removed]

\paragraph{Scaling remark (why the density-point contradiction does not follow).}
The plateau lower bound has the natural $L$ scaling, while the CR--Green/Carleson upper bound scales like $L^{1/2}$. For $0<L<1$ one has $L\le L^{1/2}$, so there is no single-interval contradiction from shrinking $L$ alone. In Route~A we therefore close \textup{(P+)} via the explicit bounded-variation wedge lemma (Lemma~\ref{lem:whitney-uniform-wedge}) fed by a scale-free (all-interval) certificate, rather than a density-point argument.

% (Removed: VK discussion does not supply a scale-free short-interval zero-clustering bound, and is not used in the conditional reduction.)

% References

\end{document}