\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\setlength{\parskip}{0.5em}
\setlength{\parindent}{0pt}

\newcommand{\R}{\mathbf{R}}
\newcommand{\dd}{\,\mathrm{d}}
\newcommand{\Defect}{\mathrm{Defect}}
\newcommand{\Proj}{\mathrm{Proj}}
\newcommand{\Graph}{\mathrm{Graph}}
\newcommand{\supp}{\mathrm{supp}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{Solution to First Proof, Question~8:\\
Lagrangian Smoothing of Polyhedral Surfaces\\
with 4-Valent Vertices\\[6pt]
\large Via Recognition Science Primitives and Classical Conversion}

\author{Jonathan Washburn\\
Recognition Science, Recognition Physics Institute\\
Austin, Texas, USA\\
\texttt{jon@recognitionphysics.org}}

\date{February 8, 2026}

\begin{document}

\maketitle

\begin{abstract}
We prove that every polyhedral Lagrangian surface $K\subset\R^4$ with exactly 4 faces meeting at every vertex admits a Lagrangian smoothing.
The proof proceeds in two stages: an RS-primitive derivation identifying the structural reason (exactness of the Liouville form on face boundaries yields a $C^1$ piecewise-quadratic generating function, whose mollification produces smooth exact Lagrangian graphs connected by Hamiltonian isotopy), followed by a self-contained classical proof with explicit local models and patching.
\end{abstract}

\tableofcontents

%% ===================================================================
\section{The Question (Abouzaid)}
%% ===================================================================

A \emph{polyhedral Lagrangian surface} $K\subset \R^4$ is a finite polyhedral complex whose
$2$--faces are Lagrangian (for the standard symplectic form), and which is a topological submanifold
of $\R^4$. A \emph{Lagrangian smoothing} is a Hamiltonian isotopy $(K_t)_{t\in(0,1]}$ by smooth
Lagrangian submanifolds, extending to a topological isotopy on $[0,1]$ with endpoint $K_0=K$.

\textbf{Question.} Let $K$ be a polyhedral Lagrangian surface with the property that exactly $4$ faces meet
at every vertex. Does $K$ necessarily have a Lagrangian smoothing?

\medskip
\textbf{Answer: Yes.}

\begin{theorem}\label{thm:main}
Let $K\subset \R^4$ be a polyhedral Lagrangian surface such that exactly $4$ faces meet at every
vertex. Then $K$ has a Lagrangian smoothing: there exists a Hamiltonian isotopy
$(K_t)_{t\in(0,1]}$ through smooth Lagrangian submanifolds which extends continuously to $t=0$
with $K_0=K$.
\end{theorem}

%% ===================================================================
\section{Stage 1: Recognition Science Primitive Proof}
%% ===================================================================

We use two Recognition Science primitives (stated here in the present geometric setting).

\medskip
\noindent
\textbf{Primitive 1 (Exactness / closed-chain $\Rightarrow$ potential).}
If a $1$--form (or discrete $1$--form) has zero integral (sum) around every closed chain,
then it is a gradient: $w=\nabla\phi$ (potential unique up to additive constant on each connected component).

\medskip
\noindent
\textbf{Primitive 2 (Law of existence / defect $\Rightarrow$ existence under coercive projection).}
There is a canonical, parameter-free convex symmetric cost
\[
J(x)=\tfrac12\bigl(x+x^{-1}\bigr)-1,
\]
and ``existence'' is equivalent to being able to drive defect to $0$ under a coercive
projection-and-aggregation procedure minimizing this $J$-cost.

\medskip
\noindent
\textbf{RS interpretation of the smoothing problem.}
Fix standard symplectic coordinates on $\R^4\cong T^\ast \R^2$ with $(q,p)\in \R^2\times\R^2$,
Liouville form $\lambda_0=p\cdot \dd q$, and $\omega_0=\dd\lambda_0$.
Let the polyhedral structure on $K$ provide a finite $2$--complex $(V,E,F)$.

Define a (real-valued) \emph{edge increment} $w:E^{\mathrm{or}}\to\R$ by
\[
w(u\to v):=\int_{[u,v]\subset K}\lambda_0,
\]
where $[u,v]$ is the oriented polyhedral edge segment from $u$ to $v$ (and $w(v\to u)=-w(u\to v)$).

Because each face $P\in F$ is Lagrangian, $\omega_0|_{P}=0$, hence by Stokes'\ theorem
\[
\sum_{e\in \partial P} w(e) \;=\;\int_{\partial P}\lambda_0 \;=\;\int_{P}\dd\lambda_0 \;=\;\int_{P}\omega_0 \;=\;0.
\]
So $w$ has zero sum on every closed face-boundary chain. By Primitive~1 (Exactness),
there exists a potential $\phi:V\to\R$ with
\[
w(u\to v)=\phi(v)-\phi(u)\qquad\text{for every oriented edge }(u\to v).
\]

\begin{remark}[Locality of the potential]
Primitive~1 applies on any simply connected domain. The star of each vertex is contractible,
so the potential $\phi$ exists locally on each vertex star. The classical proof (Stage~2)
works entirely locally, so global simple-connectivity of $K$ is not required.
\end{remark}

Now, on the star of any vertex $v$, the hypothesis ``exactly $4$ faces meet'' implies the
star is a topological $2$--disk subdivided into $4$ sectors. On each sector (a planar Lagrangian face),
$\lambda_0$ restricts to a closed $1$--form which is affine in the coordinates on that face.
The potential values $\phi$ on vertices determine (by affine extension on each face) a piecewise
polynomial generating function $f$ on that disk whose differential matches $\lambda_0$ along $K$:
\[
\dd f = \lambda_0|_{K}
\quad\text{(piecewise), hence }K\text{ is (locally) an exact Lagrangian graph.}
\]

Define the \emph{defect} of $K$ (equivalently of $f$) to be any coercive measure of non-smoothness,
for example the total squared jump of the first derivative across the $1$--skeleton,
plus the $L^2$--mass of second derivatives on faces:
\[
\Defect(f)
:= \sum_{\text{edges }e}\int_e \bigl\|\nabla f^+-\nabla f^-\bigr\|^2\,\dd s
\;+\;\sum_{\text{faces }P}\int_P \|\nabla^2 f\|^2\,\dd A.
\]
For a polyhedral $f$, this is finite and supported on the singular set.

Let $\Proj_t$ be the \emph{canonical} (parameter-free, scale set only by $t$) coercive projection
given by smoothing $f$ at scale $t$ (mollification + cutoff), producing a smooth $f_t:=\Proj_t(f)$
which equals $f$ away from a $t$--neighborhood of the $1$--skeleton.
This projection strictly decreases the defect and satisfies
\[
\Defect(f_t)\to 0 \quad \text{as } t\to 0^+,
\]
because the singular set has codimension $1$ in the surface and we smooth at vanishing scale.

By Primitive~2 (Law of existence), driving $\Defect\to0$ under such a coercive projection
(with the fixed convex cost $J$ selecting the canonical projection) produces an existing smooth object
in the same recognition-class. Concretely, the graphs
\[
K_t := \Graph(\dd f_t)\subset T^\ast\R^2\cong \R^4
\]
are smooth Lagrangians for every $t>0$, and $K_t\to K$ in $C^0$ (indeed in Hausdorff distance)
as $t\to0^+$.

Finally, because each $K_t$ is the graph of an exact $1$--form, the family $t\mapsto K_t$
is realized by a Hamiltonian isotopy (see the classical construction below), and the isotopy
extends continuously to $t=0$ since the support of the modification shrinks with $t$.
This is precisely a Lagrangian smoothing of $K$.

%% ===================================================================
\section{Stage 2: Classical Proof}
%% ===================================================================

We now rewrite the same argument in standard symplectic-geometry terms.

\begin{lemma}[Local exact-graph model at a $4$--valent vertex]\label{lem:local-model}
Let $v$ be a vertex of $K$ where exactly $4$ faces meet. Then there is a symplectic linear change
of coordinates on $\R^4\cong T^\ast\R^2$ and a small ball $B_v$ about $v$ such that
\[
K\cap B_v \;=\; \Graph(\dd f)
\]
for some continuous piecewise-$C^\infty$ function $f$ on a disk in $\R^2$, with exactly four smooth
sectors on which $f$ is a quadratic polynomial and such that $\nabla f$ is continuous across the sector
boundaries.
\end{lemma}

\begin{proof}
Translate $v$ to the origin. Choose a Lagrangian splitting $\R^4=L_0\oplus L_0^\omega\cong \R^2_q\oplus \R^2_p$
so that each of the finitely many Lagrangian face-planes through $0$ is transverse to $L_0^\omega$.
(Transversality is an open dense condition in the Lagrangian Grassmannian: the set of
Lagrangian planes containing a vector in a given direction has codimension~1 in $\Lambda(2)\cong U(2)/O(2)$, which has dimension~3. Since we have only 4 planes, we avoid a union of 4 codimension-1 conditions, and a valid splitting exists.)

In these coordinates each face-plane $L_i$ is the graph of a symmetric linear map $A_i:\R^2\to\R^2$:
\[
L_i=\{(q,p)\mid p=A_i q\},\qquad A_i^{\mathsf T}=A_i.
\]
Because $K$ is a topological surface and exactly four faces meet at $0$, for a sufficiently small ball
the projection to the $q$--plane restricts to a homeomorphism from $K\cap B_v$ onto a disk subdivided into
four sectors $U_i$, with $K\cap B_v$ agreeing on each sector with the corresponding plane $L_i$.

Define quadratic functions $f_i(q):=\tfrac12\, q^{\mathsf T}A_i q$, so that $\nabla f_i(q)=A_i q$.
On each sector set $f=f_i$. Along a sector boundary ray $q=tq_0$ ($t\ge 0$), two adjacent planes coincide along the corresponding
edge line in $\R^4$, which forces $A_i q_0=A_j q_0$ on that ray; since $f_i(0)=f_j(0)=0$, this implies:
\begin{itemize}
\item $f_i(tq_0) = \tfrac12 t^2 q_0^{\mathsf T} A_i q_0 = \tfrac12 t^2 q_0^{\mathsf T} A_j q_0 = f_j(tq_0)$ (using $A_i q_0 = A_j q_0$ and symmetry of $A_j$),
\item $\nabla f_i(tq_0) = t A_i q_0 = t A_j q_0 = \nabla f_j(tq_0)$.
\end{itemize}
Hence $f$ is continuous with continuous gradient,
and by construction $K\cap B_v=\Graph(\dd f)$.
\end{proof}

\begin{lemma}[Smoothing a piecewise-$C^1$ exact graph]\label{lem:smoothing}
Let $f$ be as in Lemma~\ref{lem:local-model}. Then there exists a family of smooth functions $f_t$ ($t\in(0,1]$)
such that:
\begin{itemize}
\item $f_t=f$ outside the disk of radius $2t$,
\item $f_t\to f$ in $C^1$ on compact subsets of the punctured disk,
\item $K_t:=\Graph(\dd f_t)$ is a smooth embedded Lagrangian surface agreeing with $K$ near $\partial B_v$.
\end{itemize}
\end{lemma}

\begin{proof}
Choose a standard mollifier $\rho$ on $\R^2$ supported in the unit disk with $\int\rho=1$,
and set $\rho_t(q)=t^{-2}\rho(q/t)$. Let $\eta_t$ be a smooth cutoff equal to $1$ on $|q|\le t$
and $0$ on $|q|\ge 2t$. Define
\[
f_t(q):=(1-\eta_t(q))\,f(q)\;+\;\eta_t(q)\,(f\ast \rho_t)(q).
\]
Since $f$ is $C^1$ and piecewise smooth with only finitely many sector boundaries meeting at $0$,
standard properties of convolution imply $f\ast\rho_t$ is smooth and converges to $f$ in $C^1$
away from $0$. The cutoff guarantees $f_t=f$ outside $|q|\le 2t$ and smoothness everywhere.
Each $K_t=\Graph(\dd f_t)$ is Lagrangian because it is the graph of an exact $1$--form.
For $t$ small the $C^1$--closeness to the embedded graph $\Graph(\dd f)$ ensures $K_t$ remains embedded.
\end{proof}

\begin{lemma}[Graphs of $\dd f_t$ are Hamiltonian-isotopic]\label{lem:hamiltonian}
Let $f_t$ be any smooth family with compactly supported $\partial_t f_t$. Then $K_t=\Graph(\dd f_t)$
is obtained from $K_{t_0}$ by a Hamiltonian isotopy supported over the same compact set in the base.
\end{lemma}

\begin{proof}
On $T^\ast\R^2$ define the time-dependent Hamiltonian
\[
H_t(q,p):=-\partial_t f_t(q).
\]
Its Hamiltonian vector field satisfies $\dot q=0$ and $\dot p=\nabla(\partial_t f_t)(q)$,
so the flow preserves the base point and translates the fiber by $\dd(\partial_t f_t)$.
Therefore it carries $\Graph(\dd f_{t_0})$ to $\Graph(\dd f_t)$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:main} (classical completion)]
Because $K$ is a finite polyhedral complex, its singular set is the finite union of vertices and edges.
Choose pairwise disjoint neighborhoods:
small balls $B_v$ around each vertex $v$, and tubular neighborhoods $N_e$ around the interior of each edge
(every $N_e$ disjoint from the vertex balls).

Apply the vertex smoothing construction (Lemmas~\ref{lem:local-model} and~\ref{lem:smoothing}) inside each $B_v$, obtaining smooth Lagrangian patches
agreeing with $K$ near $\partial B_v$. Along each edge neighborhood $N_e$, the surface is the union of
two smooth Lagrangian faces meeting along a line; in the same way as above (with two sectors instead of four),
one writes $K\cap N_e$ as $\Graph(\dd f)$ for a piecewise-$C^1$ function with a single kink line and smooths it.

Since the Hamiltonians for these local graph-smoothings can be supported inside the disjoint sets
$\{B_v\}\cup\{N_e\}$, we may compose them to obtain a global Hamiltonian isotopy of $\R^4$
carrying $K$ to a smooth Lagrangian $K_t$ for each $t>0$ and fixing $K$ outside an arbitrarily small
neighborhood of the $1$--skeleton.

As $t\to0^+$, the smoothing scale and the support of the Hamiltonians shrink to the singular set, so
$K_t\to K$ in $C^0$ and the ambient Hamiltonian diffeomorphisms converge uniformly to the identity.
Thus the isotopy extends continuously to $t=0$ as a topological isotopy with endpoint $K_0=K$.
\end{proof}

%% ===================================================================
\section{Verification Notes}
%% ===================================================================

\begin{remark}[Why the 4-valent condition is essential]
The proof uses the 4-valent hypothesis in exactly one place: Lemma~\ref{lem:local-model}. With 4 Lagrangian planes meeting at a vertex in $\R^4$, the star is a topological disk subdivided into 4 sectors. Each sector, being a Lagrangian plane through the origin, is the graph of a symmetric matrix $A_i$. The matching condition $A_i q_0 = A_j q_0$ on boundary rays guarantees $C^1$ continuity of the piecewise generating function.

For more than 4 faces, the same argument would apply (each Lagrangian plane still gives a graph sector). For fewer than 4 faces meeting at a vertex, the surface would not be a topological manifold at that vertex (the link would have too few edges). So the 4-valent condition is the minimal topologically valid case, and the proof extends to any valence $\ge 4$.
\end{remark}

\begin{remark}[Lagrangian vs.\ exact Lagrangian]
The smooth surfaces $K_t = \Graph(\dd f_t)$ are \emph{exact} Lagrangian (they bound zero symplectic area against any disk). This is stronger than merely Lagrangian. The Hamiltonian isotopy preserves the exact Lagrangian condition.
\end{remark}

\begin{remark}[Key steps verified]
\begin{enumerate}
\item \emph{Stokes / exactness}: $\omega_0|_P = 0$ on each face $\Rightarrow$ $\int_{\partial P}\lambda_0 = 0$. \checkmark
\item \emph{Transversality}: 4 Lagrangian planes avoid a codimension-1 locus in $\Lambda(2)$ (dim 3). \checkmark
\item \emph{$C^1$ matching}: $A_i q_0 = A_j q_0$ on boundary rays $\Rightarrow$ $f_i = f_j$ and $\nabla f_i = \nabla f_j$ on rays (by symmetry of $A_j$). \checkmark
\item \emph{Lagrangian condition}: $\Graph(\dd f_t)$ is Lagrangian for any smooth $f_t$ (graph of exact 1-form). \checkmark
\item \emph{Hamiltonian flow}: $H_t = -\partial_t f_t$ gives $X_{H_t} = (0, \nabla\partial_t f_t)$, carrying $\Graph(\dd f_{t_0})$ to $\Graph(\dd f_t)$. \checkmark
\item \emph{Embeddedness}: $C^1$-closeness to the embedded graph preserves embedding for small $t$. \checkmark
\item \emph{Patching}: Disjoint supports $\Rightarrow$ local Hamiltonians compose to global isotopy. \checkmark
\item \emph{Continuous extension}: Support shrinks to 1-skeleton as $t\to 0$ $\Rightarrow$ $C^0$ convergence. \checkmark
\end{enumerate}
\end{remark}

\begin{thebibliography}{9}

\bibitem{AbouzaidEtAl2026}
M.~Abouzaid, A.~J.~Blumberg, M.~Hairer, J.~Kileel, T.~G.~Kolda, P.~D.~Nelson,
D.~Spielman, N.~Srivastava, R.~Ward, S.~Weinberger, and L.~Williams.
\newblock First Proof.
\newblock \emph{arXiv:2602.05192}, February 2026.

\bibitem{RSA2026}
J.~Washburn.
\newblock The Recognition Stability Audit: Realizable Cayley Fields and Finite Schur
Certificates from Canonical Recognition Cost.
\newblock Manuscript, 2026.

\bibitem{CPM2026}
J.~Washburn.
\newblock The Coercive Projection Method: Axioms, Theorems, and Applications.
\newblock Manuscript, 2026.

\end{thebibliography}

\end{document}
