\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}
\title{Curated Paper List (Suggested 26-Paper Publishing Order)}
\date{Generated Feb 4, 2026}
\begin{document}
\maketitle
\section*{Notes}
\noindent Expected authorship date is inferred from the file's last-modified timestamp.
\section*{Curated Papers}
\begin{enumerate}[leftmargin=*,label=\arabic*.]
\item \textbf{Recognition Geometry}: \texttt{\detokenize{papers/tex/recognition-geometry.tex}} (Published; Authorship: Dec 6, 2025)
\par \textit{Why}: \detokenize{Published anchor; establishes recognition quotient and observable geometry.}

\item \textbf{Recognition Composition Law Primer}: \texttt{\detokenize{papers/tex/Recognition_Composition_Law_Primer.tex}} (Authorship: Jan 2, 2026)
\par \textit{Why}: \detokenize{States the core composition law for recognition costs.}

\item \textbf{Uniqueness of the Canonical Reciprocal Cost}: \texttt{\detokenize{papers/UNIQUENESS OF THE CANONICAL RECIPROCAL COST.tex}} (Authorship: Jan 19, 2026)
\par \textit{Why}: \detokenize{Formal uniqueness proof for J-cost.}

\item \textbf{d'Alembert Inevitability}: \texttt{\detokenize{papers/tex/DAlembert_Inevitability.tex}} (Authorship: Jan 19, 2026)
\par \textit{Why}: \detokenize{Forces d'Alembert structure under consistency constraints.}

\item \textbf{Model-Independent Exclusivity Quotient}: \texttt{\detokenize{papers/tex/Model-Independent-Exclusivity-Quotient.tex}} (Authorship: Jan 16, 2026)
\par \textit{Why}: \detokenize{Exclusivity at the quotient/observable level.}

\item \textbf{The Cost of Existence}: \texttt{\detokenize{The_Cost_of_Existence.tex}} (Authorship: Feb 3, 2026)
\par \textit{Why}: \detokenize{Existence as zero-defect; unity as unique existent.}

\item \textbf{The Law of Inevitable Unity}: \texttt{\detokenize{papers/The_Law_of_Inevitable_Unity.tex}} (Authorship: Jan 4, 2026)
\par \textit{Why}: \detokenize{Unity/identity forced by cost axioms.}

\item \textbf{Logic From Physical Cost}: \texttt{\detokenize{papers/tex/Logic_From_Physical_Cost.tex}} (Authorship: Jan 22, 2026)
\par \textit{Why}: \detokenize{Logic and consistency derived from cost minima.}

\item \textbf{Recognition Science: Foundations}: \texttt{\detokenize{papers/tex/RS-Foundations.tex}} (Authorship: Jan 3, 2026)
\par \textit{Why}: \detokenize{Consolidated foundation overview for physics audience.}

\item \textbf{Eight Axioms Forced}: \texttt{\detokenize{papers/tex/EightAxiomsForced.tex}} (Authorship: Nov 3, 2025)
\par \textit{Why}: \detokenize{Forcing stack for key axioms (discreteness/ledger/etc.).}

\item \textbf{Penrose Golden Ratio and Ledger Structure}: \texttt{\detokenize{papers/tex/Penrose_golden_ratio_and_ledger_structure.tex}} (Authorship: Feb 2, 2026)
\par \textit{Why}: \detokenize{Ledger self-similarity forcing for phi.}

\item \textbf{Geometric Necessity: Recognition Angle}: \texttt{\detokenize{papers/tex/Geometric-Necessity-Recognition-Angle.tex}} (Authorship: Jan 27, 2026)
\par \textit{Why}: \detokenize{Recognition angle forced by structural constraints.}

\item \textbf{Dimensional Rigidity D3}: \texttt{\detokenize{papers/tex/Dimensional_Rigidity_D3.tex}} (Authorship: Feb 3, 2026)
\par \textit{Why}: \detokenize{Independent arguments forcing D=3.}

\item \textbf{The Recognition Operator}: \texttt{\detokenize{papers/root_papers/The_Recognition_Operator.tex}} (Authorship: Jan 8, 2026)
\par \textit{Why}: \detokenize{Dynamics via cost-minimizing operator R-hat.}

\item \textbf{Formalized Derivations T1-T8}: \texttt{\detokenize{papers/tex/Formalized-Derivations-T1-T8.tex}} (Authorship: Jan 3, 2026)
\par \textit{Why}: \detokenize{Derivation chain from the Meta-Principle.}

\item \textbf{CPM Constants Derivation}: \texttt{\detokenize{papers/tex/CPM_Constants_Derivation.tex}} (Authorship: Dec 1, 2025)
\par \textit{Why}: \detokenize{Derives constants from phi/ledger structure.}

\item \textbf{Full First-Principles Mass Derivation}: \texttt{\detokenize{papers/tex/Full_First_Principles_Mass_Derivation.tex}} (Authorship: Jan 31, 2026)
\par \textit{Why}: \detokenize{Charged fermion masses from octave closure.}

\item \textbf{Neutrino Sector No-Go}: \texttt{\detokenize{papers/tex/Neutrino-Sector.tex}} (Authorship: Oct 13, 2025)
\par \textit{Why}: \detokenize{Falsifiable constraint on neutrino sector.}

\item \textbf{Zero-Parameter Quantum Gravity}: \texttt{\detokenize{papers/root_papers/quantum_gravity_B_v4.tex}} (Authorship: Jan 6, 2026)
\par \textit{Why}: \detokenize{Gravity derivation with zero free parameters.}

\item \textbf{Dark Energy (ILG II)}: \texttt{\detokenize{papers/root_papers/Dark_Energy_Paper2_v4.tex}} (Authorship: Jan 7, 2026)
\par \textit{Why}: \detokenize{Cosmology sector within ILG framework.}

\item \textbf{ILG Galaxy Rotation Curves}: \texttt{\detokenize{papers/ILG_Galaxy_Rotation_Curves.tex}} (Authorship: Jan 8, 2026)
\par \textit{Why}: \detokenize{Galaxy rotation as RS/ILG prediction.}

\item \textbf{ILG Validation Synthesis}: \texttt{\detokenize{papers/ILG_Validation_Synthesis.tex}} (Authorship: Jan 8, 2026)
\par \textit{Why}: \detokenize{Empirical synthesis/validation of ILG outputs.}

\item \textbf{Recognition-Riemann Final}: \texttt{\detokenize{papers/tex/Recognition-Riemann-Final.tex}} (Authorship: Dec 24, 2025)
\par \textit{Why}: \detokenize{Riemann application of recognition geometry.}

\item \textbf{Protein Folding as Phase Recognition}: \texttt{\detokenize{papers/tex/protein-dec-6.tex}} (Authorship: Dec 6, 2025)
\par \textit{Why}: \detokenize{Biological application of RS dynamics.}

\item \textbf{Topological Origins of Nuclear Binding Energy}: \texttt{\detokenize{papers/tex/Topological_Origins_Nuclear_Binding_Energy.tex}} (Authorship: Jan 18, 2026)
\par \textit{Why}: \detokenize{Nuclear binding from 8-tick ledger topology.}

\item \textbf{Nuclear Magic Numbers (RS Derivation)}: \texttt{\detokenize{fusion/papers/Nuclear_Magic_Numbers_RS_Derivation.tex}} (Authorship: Jan 25, 2026)
\par \textit{Why}: \detokenize{Magic numbers derived from RS structure.}

\end{enumerate}
\end{document}