% Recognition Computing Architecture — Foundational Patent (Method, System, Medium)
\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{xcolor}

\title{Recognition Computing Architecture\\Foundational Patent (Method, System, Medium)}
\author{Jonathan Washburn}
\date{\today}

\begin{document}
\maketitle

\begin{center}
{\large Foundational Patent Draft}\\[4pt]
{\normalsize Recognition Computing: The Only Zero-Parameter Physical Computing Substrate}
\end{center}

\section*{Field}
This disclosure relates to computer-implemented control and measurement of physical systems. It provides a parameter-free computing architecture---\emph{Recognition Computing}---that factors observables through dimensionless ratios, aggregates them with a uniquely determined convex symmetric ledger kernel, enforces $\varphi$-commensurate neutral timing, gates operation by machine-checkable certificates, and selects an absolute layer by solving simultaneous dimensionless identities. The disclosure covers method, system, and non-transitory computer-readable medium embodiments, as well as hardware primitives including a J-kernel core and timing circuitry.

\section*{Background}
Conventional control/measurement pipelines suffer three endemic problems: (i) \emph{units ambiguity} in raw observables; (ii) \emph{parameter explosion} from ad-hoc objectives and per-problem knobs; and (iii) \emph{calibration debt} due to reliance on external anchors, manual tuning, and drift. Prior art (e.g., ratio-based metrology, Gray-coded timing, statistical audits) addresses fragments but does not provide a \emph{complete, parameter-free} architecture that is units-invariant, self-calibrating, and self-auditing with formal proofs.

\section*{Summary}
Disclosed is a universal, parameter-free physical computing architecture realized as a five-layer stack:
\begin{enumerate}[leftmargin=*]
  \item \textbf{Recognition layer}: form dimensionless recognition ratios $r_i=y_i/y_i^\star$ from observables $y_i$ and declared targets $y_i^\star>0$.
  \item \textbf{Ledger layer}: compute an aggregate cost $L=\sum_i w_i\,F(r_i)$ using the \emph{unique} convex symmetric kernel $F:(0,\infty)\to\mathbb{R}$ determined by five axioms (symmetry $F(x)=F(1/x)$; unit normalization $F(1)=0$; strict convexity on $\mathbb{R}_{>0}$; unit curvature of $F\circ\exp$ at 0; cosh-addition identity). This kernel is equivalently $J(x)=\tfrac12(x+1/x)-1$.
  \item \textbf{Temporal layer}: partition time into $\varphi$-commensurate windows (e.g., eight-tick minimality in $D=3$) and enforce neutrality constraints (mirrored halves with mid-cycle FLIP; $\sum_{i=1}^{8}\Delta S_i=0$).
  \item \textbf{Certificate layer}: execute machine-checkable proofs of invariants (units quotient, route identities, neutrality, gate separation) and gate operation or data ingestion only if a certificate passes.
  \item \textbf{Absolute layer selection}: solve simultaneous dimensionless gate identities to select a unique absolute layer and emit absolute scales/constants without external anchors.
\end{enumerate}
Embodiments include software (SDKs and services), firmware (timing and gating), and hardware (J-kernel core, $\varphi$-scheduler). The architecture produces measurable technical effects: reduced alias/phase error, deterministic reproducibility, fail-closed safety, and reduced compute/energy in audit loops.

\section*{Definitions}
\noindent\textbf{Recognition ratio.} $r_i=y_i/y_i^\star$ with $y_i^\star>0$. Ratios are dimensionless and invariant to unit rescaling.

\noindent\textbf{J-kernel (axiomatized).} The cost kernel $F:(0,\infty)\to\mathbb{R}$ is uniquely determined by:
\begin{enumerate}[label=(A\arabic*),leftmargin=*]
  \item $F(x)=F(1/x)$ for all $x>0$ (symmetry),
  \item $F(1)=0$ (unit normalization),
  \item $F$ is strictly convex on $\mathbb{R}_{>0}$,
  \item $\dfrac{d^2}{dt^2}\big(F(e^t)\big)\big|_{t=0}=1$ (unit curvature in log-domain),
  \item a cosh-addition identity holds, forcing $F(x)=\cosh(\ln x)-1=\tfrac12(x+1/x)-1$.
\end{enumerate}

\noindent\textbf{$\varphi$-commensurate windows.} A control period $T$ is partitioned into windows with durations whose ratios are in $\{\varphi,\varphi^{-1}\}$, with neutrality constraints and Gray-aligned eight-tick minimality.

\noindent\textbf{Certificate.} A hash-addressed, signed bundle of machine-checkable proof results and run metadata that gates operation or data acceptance.

\noindent\textbf{Absolute layer.} The unique layer selected by solving simultaneous dimensionless gate identities (e.g., route-equality and Planck-gate identities), yielding absolute scales/constants internally.

\noindent\textbf{Compliance API.} A controller-agnostic interface exposing window state, gating decisions, ledger aggregates, neutrality flags, and certificate references, with machine-verifiable logs.

\section*{Claims}
\setlist[enumerate]{leftmargin=*,labelsep=0.5em,itemsep=0.6em,topsep=0.4em}

\subsection*{Independent Claims}
\begin{enumerate}
\item \textbf{(Method)} A computer-implemented method of controlling or measuring a physical system, comprising:
\begin{enumerate}[label=(\alph*)]
  \item receiving observables $y_i$ and forming dimensionless recognition ratios $r_i=y_i/y_i^\star$ with declared targets $y_i^\star>0$;
  \item computing a ledger $L=\sum_i w_i\,F(r_i)$ where $F:(0,\infty)\to\mathbb{R}$ is the unique function satisfying: (i) $F(x)=F(1/x)$; (ii) $F(1)=0$; (iii) strict convexity on $\mathbb{R}_{>0}$; (iv) $\tfrac{d^2}{dt^2}(F(e^t))\big|_{t=0}=1$; and (v) a cosh-addition identity; 
  \item partitioning operation into $\varphi$-commensurate time windows and enforcing neutrality constraints over the windows;
  \item executing machine-checkable proofs of invariants and generating a certificate tied to the operation; and
  \item authorizing actuation or data acceptance only upon satisfaction of certificate thresholds and, when applicable, selecting an absolute layer by solving simultaneous dimensionless gate identities.
\end{enumerate}

\item \textbf{(System)} A system for controlling or measuring a physical system, comprising one or more processors, memory, sensors or actuators, and timing circuitry configured to: (i) form recognition ratios $r_i=y_i/y_i^\star$; (ii) compute a ledger $L=\sum_i w_i\,F(r_i)$ where $F$ satisfies (i)--(v) of claim 1; (iii) partition control into $\varphi$-commensurate windows and enforce neutrality constraints; (iv) execute machine-checkable invariant proofs and generate a certificate; and (v) gate operation or data acceptance based on the certificate and, when applicable, select an absolute layer by solving simultaneous dimensionless gate identities.

\item \textbf{(Non-transitory medium)} A non-transitory computer-readable medium storing instructions that, when executed by one or more processors, cause a system to perform the method of claim 1.
\end{enumerate}

\subsection*{Dependent Claims (Method of Claim 1 unless noted)}
\begin{enumerate}[resume]
\item Where $F(x)=\tfrac12(x+1/x)-1$ (J-kernel).
\item Where symmetry is enforced by verifying $F(x)=F(1/x)$ within a numerical tolerance across a sampled set of ratios.
\item Where unit normalization is enforced by verifying $F(1)=0$ within a numerical tolerance.
\item Where strict convexity is enforced by verifying a strict midpoint inequality on logged samples of ratios.
\item Where the unit-curvature condition is enforced by fitting a quadratic to $F(e^t)$ near $t=0$ with curvature coefficient approximately one.
\item Where the cosh-addition identity is verified on structured ratio pairs $(x,y)$.
\item Where the time windows include eight-tick Gray-aligned commits and mirrored halves with a mid-cycle FLIP such that $\sum_{i=1}^{8}\Delta S_i=0$.
\item Where the certificate comprises cryptographic hashes of proof results, run identifiers, tolerance bands, timing telemetry, and acceptance decisions.
\item Where the absolute layer is selected by solving simultaneous identities comprising route-equality and a Planck-gate identity, within declared tolerance bands.
\item Where a compliance API exposes window state, gating decisions, ledger aggregates, neutrality flags, and certificate references with machine-verifiable logs.
\item (System of claim 2) further comprising a hardware J-kernel core with a streaming log-domain unit computing $J(e^u)=\cosh(u)-1$ or equivalently $J(x)=\tfrac12(x+1/x)-1$, a block-sum neutrality accumulator, timing inputs synchronized to $\varphi$-commensurate windows, and a certificate interface.
\item (System of claim 2) further comprising a scheduler that generates $\varphi$-commensurate windows and enforces neutrality with bounded jitter.
\item (Method) further comprising executing axiom-probe tests to detect compliance with (A1)--(A5) by structured stimulus-response experiments on black-box interfaces.
\item (Method) wherein actuation commands are adjusted to minimize a block-sum of the ledger over one or more windows subject to safety constraints and where acceptance is fail-closed if combined-uncertainty thresholds are not satisfied.
\item (Method) wherein selecting the absolute layer emits absolute scales and constants ($\tau_0,\ell_0,c,\hbar,\lambda_{\mathrm{rec}},G$) without reference to external anchors.
\end{enumerate}

\section*{Brief Description of the Drawings}
\begin{enumerate}[leftmargin=*]
  \item Fig. 1: Five-layer architecture diagram (recognition, ledger, temporal, certificate, absolute).
  \item Fig. 2: $\varphi$-commensurate window schedule with eight-tick neutrality and mirrored halves.
  \item Fig. 3: Hardware J-kernel core block diagram and compliance API signals.
  \item Fig. 4: Certificate bundle schema and verification flow.
  \item Fig. 5: Axiom-probe protocol coverage and decision thresholds.
\end{enumerate}

\section*{Detailed Description}
\subsection*{Architecture}
The architecture is implemented as a pipeline. Observables $y_i$ are first normalized to dimensionless recognition ratios $r_i=y_i/y_i^\star$, eliminating unit ambiguity. The ledger layer aggregates $r_i$ via $L=\sum_i w_i\,F(r_i)$ where $F$ is J-kernel, uniquely determined by axioms (A1)--(A5). The temporal layer partitions operation into $\varphi$-commensurate windows and enforces neutrality constraints (e.g., in $D=3$ the minimal spatially complete period is eight). The certificate layer executes machine-checkable proofs of invariants (units quotient, route identities, neutrality, gate separation, policy) and gates operation or data ingestion on pass. Lastly, the absolute layer selects a unique absolute layer by solving simultaneous dimensionless gate identities; this emits absolute scales/constants without external anchors.

\subsection*{Hardware and Firmware Embodiments}
\textbf{J-kernel core.} A streaming unit computes $J(e^u)=\cosh(u)-1$ for log-domain inputs or equivalently $J(x)=\tfrac12(x+1/x)-1$ for positive inputs, with fixed precision and bounded latency. A block-sum accumulator computes window-level neutrality observables and ledger aggregates. Timing inputs synchronize to $\varphi$-commensurate windows.

\textbf{$\varphi$-scheduler.} A firmware timing module partitions each period $T$ into windows whose durations obey $\{\varphi,\varphi^{-1}\}$ ratios, enforces mirrored halves with a mid-cycle FLIP, and guarantees bounded jitter. The scheduler exposes a compliance API and generates signed compliance records.

\textbf{Certificate engine.} A proof runner verifies identities, collects results into a signed, hash-addressed bundle, and exposes acceptance decisions. Certificates include run IDs, invariant values, uncertainty model, proof hashes, timestamps, device metadata, and acceptance verdicts.

\subsection*{Axiom-Probe Detection}
Black-box axiom probes validate compliance with (A1)--(A5): symmetry via $F(x)=F(1/x)$; normalization via $F(1)=0$; strict convexity via midpoint inequality; unit curvature via a local quadratic fit of $F(e^t)$ near $t=0$; cosh-addition identity via $F(xy)+F(x/y)=2F(x)$. Passing all probes establishes functional equivalence to J-kernel via uniqueness; use of any function $F'$ that satisfies the axioms constitutes use of J.

\subsection*{Absolute Layer Selection}
Gate identities, including route-equality and layer policies such as a Planck-gate, are solved simultaneously with tolerance bands to select a unique absolute layer. This step emits absolute scales ($\tau_0,\ell_0$) and constants ($c,\hbar,\lambda_{\mathrm{rec}},G$) internally without external anchors.

\subsection*{Technical Effects}
The architecture delivers measurable improvements: (i) reduction of alias/phase errors via neutrality and ledger minimization; (ii) deterministic, units-invariant reproducibility via recognition ratios and absolute-layer selection; (iii) fail-closed safety via certificate-gated acceptance with combined-uncertainty thresholds; (iv) reduced compute/energy due to faster audit convergence and fail-closed gating.

\subsection*{Industrial Applicability}
The architecture applies to instruments (spectroscopy, timing, imaging), simulators and digital twins, fusion plasma controllers, manufacturing (e.g., perovskite modules), synthetic biology instruments (templating under LISTEN/LOCK/BALANCE), and physically constrained AI training pipelines.

\section*{Examples (Non-limiting)}
\textbf{Metrology.} A spectrometer normalizes measurements, minimizes a block-sum of J over eight-tick windows, verifies route identities and neutrality, and gates acceptance by certificate; absolute-layer selection emits anchor-free scales.

\textbf{Fusion.} A tokamak controller computes ledger from normalized diagnostics and $\varphi$-phased actuator updates; operation is gated by certificate; timing reduces cross-interference.

\textbf{Manufacturing.} A perovskite line enforces window-8 neutrality and IR phase-lock; rate balance ($x\to 1$) at interfaces minimizes drift.

\section*{Best Mode}
Reference implementations include: (i) SDK APIs for recognition ratios, ledger, scheduler, and certificates; (ii) firmware IP cores for $\varphi$-scheduler and J-kernel; (iii) cloud/on-prem certificate engine and registry; (iv) test harnesses for axiom probes.

\section*{Enablement}
Formal proofs (Lean) provide mathematical enablement for the uniqueness of J-kernel and exclusivity of the architecture under zero-parameter constraints. Benchmarks quantify technical effects. Hardware/firmware/software embodiments provide practical enablement.

\section*{Alternative Embodiments}
Window counts other than eight may be used in domains with different dimensional or plant constraints, provided neutrality and $\varphi$-commensurability are preserved. Certificates may be on-device or cloud-verified. The J-kernel may be computed in fixed-point, floating-point, or log-domain forms.

\section*{Legal Notice}
All embodiments, features, and steps described herein may be combined in any operable permutation. Where means-plus-function is implicated, claims are intended to cover corresponding structures and their equivalents.

\end{document}


