\section{Non-circular calibration / hold-out protocol for the structural mass model}
\label{sec:holdout-protocol}

\paragraph{Why this section exists.}
Anil's Section~4 critique is methodological: if we ``test'' a proposed correction term by
rearranging the same measured mass we are testing, the exercise is circular.
This section states (and runs) a strict non-circular procedure for the \emph{structural} part of the mass model.

\paragraph{Important clarification (what this is \emph{not}).}
This hold-out test is \emph{not} the SM single-anchor identity paper (which is purely RG-side).
It is a separate check for the RS \emph{structural ladder} model, whose parameters include sector-global offsets.

\subsection*{Protocol}
We work sector-by-sector (Lepton / UpQuark / DownQuark / Electroweak).
Within each sector we assume the structural-only ladder form
\[
  m_{\mathrm{struct}}(i)
  =
  2^{B(\mathrm{sector})}\,E_{\mathrm{coh}}\,\varphi^{r_0(\mathrm{sector})}\,\varphi^{r_i-8},
\]
where:
\begin{itemize}
  \item \(B(\mathrm{sector})\) is a fixed sector binary weight,
  \item \(E_{\mathrm{coh}}\) and \(\varphi\) are global constants,
  \item \(r_i\) is the rung assigned to the species,
  \item \(r_0(\mathrm{sector})\) is a \emph{single} sector offset (one value per sector).
\end{itemize}

\paragraph{Hold-out rule.}
To avoid circularity, a species' measured mass is never used in its own prediction.
Concretely, for each sector we perform \textbf{leave-one-out} cross-validation:
\begin{itemize}
  \item pick one species to hold out,
  \item fit the sector offset \(r_0(\mathrm{sector})\) using only the \emph{other} species in that sector,
  \item freeze \(r_0(\mathrm{sector})\) and predict the held-out species,
  \item record prediction error.
\end{itemize}

\paragraph{Implementation (reproducible).}
This protocol is implemented by:
\begin{itemize}
  \item \texttt{tools/holdout\_structural\_masses.py}
  \item input data: \texttt{data/masses.json}
  \item outputs:
    \texttt{out/masses/structural\_holdout.csv} and
    \texttt{out/masses/structural\_holdout\_table.tex}.
\end{itemize}

\subsection*{Results (structural-only hold-out)}
\paragraph{What the table means.}
Each row is an \emph{out-of-sample} prediction: the sector offset \(r_0\) was fit from other species,
then used to predict the held-out one.
If the structural-only model were already complete, these errors would be small.

\input{out/masses/structural_holdout_table.tex}

\paragraph{Interpretation.}
The errors are not uniformly small under this strict protocol, which is exactly the point:
\textbf{a sector-wide ladder with one offset is not, by itself, sufficient for high-precision prediction
across all species.}
This is compatible with the broader framework, which includes additional (theory-side) residue structure
and/or transport corrections; but it makes the methodology explicit and prevents accidental self-fitting.


