\documentclass[11pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue,
    citecolor=blue
}

\title{\textbf{Recognition Science: A Curated Syllabus}\\ \large The Logic of the Dependency Graph}
\author{Jonathan Washburn}
\date{\today}

\begin{document}

\maketitle

\section*{Introduction}
This syllabus mirrors the logical Dependency Graph (DAG) of Recognition Science. The papers are ordered not chronologically, but structurally: starting from the root geometry, ascending the ``spine'' of derivations, and then branching into specific domains (Particles, Computation, Gravity, Math, Consciousness, Ethics).

\section{I. The Root (Foundation)}

\subsection{1. Recognition Geometry}
\textbf{File:} \texttt{papers/01-root-foundation/recognition-geometry.tex}
\textbf{Priority:} Tier 1 (Publish First)
\textbf{Description:} Defines the observational space, recognizers, and the recognition quotient.

\subsection{2. The Recognition Composition Law Primer}
\textbf{File:} \texttt{papers/01-root-foundation/Recognition\_Composition\_Law\_Primer.tex}
\textbf{Priority:} Tier 1 (Publish First)
\textbf{Description:} Introduces the fundamental rule of comparison (RCL) governing how costs combine.

\subsection{3. Uniqueness of the Canonical Reciprocal Cost}
\textbf{File:} \texttt{papers/01-root-foundation/Cost\_Uniqueness.tex}
\textbf{Priority:} Tier 1 (Publish First)
\textbf{Description:} Proves that the RCL forces the unique cost function $J(x) = \frac{1}{2}(x + x^{-1}) - 1$.

\subsection{4. Coherent Comparison as Information Cost: A Cost-First Ledger Framework}
\textbf{File:} \texttt{papers/01-root-foundation/Coherent\_Comparison\_Ledger.tex}
\textbf{Priority:} Tier 1 (Publish First)
\textbf{Description:} Derives the discrete ledger dynamics from the cost function.

\subsection{5. D'Alembert Inevitability: Polynomial Consistency Forces the Canonical Law}
\textbf{File:} \texttt{papers/01-root-foundation/DAlembert\_Inevitability.tex}
\textbf{Priority:} Tier 1 (Publish First)
\textbf{Description:} Proves the inevitability of the composition law from polynomial consistency constraints.

\subsection{6. Model-Independent Exclusivity on the Quotient State Space}
\textbf{File:} \texttt{papers/01-root-foundation/Model-Independent-Exclusivity-Quotient.tex}
\textbf{Priority:} Tier 1 (Publish First)
\textbf{Description:} Proves that any zero-parameter framework is observationally equivalent to RS.

\subsection{51. Gödel's Theorem Does Not Obstruct Physical Closure}
\textbf{File:} \texttt{papers/01-root-foundation/godel\_dissolution.tex}
\textbf{Description:} Resolves the Gödelian objection by defining truth as cost-stabilization.

\subsection{53. The Recognition Stability Audit (RSA)}
\textbf{File:} \texttt{papers/01-root-foundation/Recognition\_Stability\_Audit.tex}
\textbf{Description:} Formalizes the ``impossibility audit'' and certificates for existence claims.

\subsection{82. Recognition Algebra: The Unified Algebraic Framework}
\textbf{File:} \texttt{papers/01-root-foundation/Recognition\_Algebra.tex}
\textbf{Priority:} Tier 1 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Shows that the RCL forces four interlocking algebraic layers: (1)~the Cost Algebra --- a commutative monoid on cost values; (2)~the $\varphi$-Ring $\ZZ[\varphi]$ --- a Euclidean domain with $\varphi$ as unit; (3)~the Ledger Algebra --- paired events with $\sigma=0$ conservation; (4)~the Octave Algebra --- $\ZZ/8\ZZ$ with DFT-8 and 20 basis modes.  Proves Recognition Algebra is the initial object in the category of zero-parameter cost-minimizing frameworks.  Self-contained proofs.  Lean-verified.

\textbf{Dependencies:} Papers 3 (J-cost uniqueness), 4 (Ledger), 10 ($\varphi$).

\subsection{84. Inevitability From Three Attributes (Core and Extended)}
\textbf{File:} \texttt{Inevitability\_From\_Three\_Attributes.tex}
\textbf{Priority:} Tier 1 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Formalizes omniscience, omnipotence, and omnipresence on a costed comparison system. Proves core inevitability of the canonical $J$-cost and the unique ambient dimension $D=3$, then proves the reverse implication to obtain a core biconditional. Extended consequences ($\varphi$, 8-tick, $W=17$) are explicitly marked as postulate-dependent.

\textbf{Dependencies:} Papers 3 (J-cost uniqueness), 5 (D'Alembert), 11 (D=3 topology).

\section{II. The Structure (Ontology)}

\subsection{7. The Cost of Existence}
\textbf{File:} \texttt{papers/02-structure-ontology/The\_Cost\_of\_Existence.tex}
\textbf{Priority:} Tier 2
\textbf{Description:} Derives existence from the infinite cost of the void ($J(0) \to \infty$).

\subsection{52. The Law of Existence}
\textbf{File:} \texttt{papers/02-structure-ontology/Law-of-Existence-arXiv.tex}
\textbf{Priority:} Tier 1
\textbf{Description:} The core ontological paper linking CPM, Darwin, and physical constants.

\subsection{8. Logic From Physical Cost}
\textbf{File:} \texttt{papers/02-structure-ontology/Logic\_From\_Physical\_Cost.tex}
\textbf{Priority:} Tier 2
\textbf{Description:} Derives logical consistency ($A=A$) as the zero-cost ground state ($J(1)=0$).

\subsection{9. The Recognition Operator: Beyond the Hamiltonian}
\textbf{File:} \texttt{papers/02-structure-ontology/Recognition-Operator.tex}
\textbf{Priority:} Tier 2
\textbf{Description:} Defines $\hat{R}$ as fundamental dynamics replacing $\hat{H}$. Derives Hamiltonian as small-$\varepsilon$ approximation: $J(e^\varepsilon) = \cosh(\varepsilon)-1 = \frac{1}{2}\varepsilon^2 + O(\varepsilon^4)$, valid to $<1\%$ for $|\varepsilon|\le 0.1$. Recovers Schr\"{o}dinger equation as continuum limit. Predicts measurable departures: non-equilibrium flows, ultra-fast 8-tick discretization, mesoscopic collapse at $C \ge 1$.

\textbf{Dependencies:} Papers 3 (J-cost), 4 (Ledger dynamics), 5 (D'Alembert).

\textbf{Why here?} Establishes the fundamental dynamics; Hamiltonian mechanics becomes a theorem, not an axiom.

\subsection{74. Noether from Cost: The Hamiltonian as Lagrange Multiplier}
\textbf{File:} \texttt{papers/02-structure-ontology/RK\_Noether\_From\_Cost.tex}
\textbf{Priority:} Tier 2 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Proves $\hat{H}$ is the Lagrange multiplier enforcing discrete continuity (T3) while minimising cumulative $J$-cost. Multiplier scale fixed by K-gate: $\hbar = E_{\mathrm{coh}} \cdot \tau_0$. Discrete Noether theorem: each continuous symmetry of $J$ yields a conserved $Z$-pattern; ``energy'' is the multiplier itself. Lean-verified algebraic core.

\textbf{Dependencies:} Papers 3 (J-cost), 4 (Ledger), 9 (R-hat).

\textbf{Why here?} Completes the ``Hamiltonian is emergent'' story: Paper~9 shows $\hat{R}$ is fundamental, this paper shows $\hat{H}$ is its constraint multiplier.

\subsection{75. The Geometry of Inquiry: Questions as Cost Gaps}
\textbf{File:} \texttt{papers/02-structure-ontology/RK\_Geometry\_of\_Inquiry.tex}
\textbf{Priority:} Tier 2 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Formalises questions as cost gaps. Classification: well-formed ($\exists$ finite-cost answer), dissolved (all answers infinite cost --- G\"{o}del-type), forced (exactly one zero-cost answer). T0--T8 recast as forced questions. Meta-closure: RS has zero cost in theory space, all alternatives positive cost. Self-reference is stable (cost-decreasing), not paradoxical. Eight fundamental inquiry modes.

\textbf{Dependencies:} Papers 3 (J-cost), 7 (Cost of Existence), 51 (G\"{o}del Dissolution), 39 (Reference).

\textbf{Why here?} Achieves meta-closure: RS explains why RS is the answer.

\subsection{10. The Golden Ratio as a Universal Coherence Eigenvalue}
\textbf{File:} \texttt{papers/02-structure-ontology/Penrose\_golden\_ratio\_and\_ledger\_structure.tex}
\textbf{Priority:} Tier 2
\textbf{Description:} Explains $\varphi$ as the unique fixed point of the cost recursion $x = 1 + 1/x$.  Establishes the log-ratio isomorphism $(\mathbb{R}_{>0},\times) \cong (\mathbb{R},+)$ bridging Penrose aperiodic tilings and RS cost dynamics.  Evaluates $J(\varphi) = \varphi - 3/2 \approx 0.118$ as the ``coherence cost of aperiodicity.''

\textbf{Dependencies:} Paper~3 (J-cost uniqueness; $\varphi$ is defined via cost self-similarity).

\textbf{Why here?} Pins the universal scale constant before it is used to force $D{=}3$ and derive constants.

\subsection{11. Dimensional Rigidity: D=3 (Strengthened)}
\textbf{File:} \texttt{papers/02-structure-ontology/Dimensional\_Rigidity\_D3\_2.tex}
\textbf{Priority:} Tier 2
\textbf{Description:} Three independent proofs that $D=3$: (T) Alexander duality---linking invariant exists iff $D=3$; (K) Kepler stability---non-precessing orbits require $D=3$; (S) Minimal dyadic synchronization---$\mathrm{lcm}(2^D,45)$ minimized at $D=3$.

\textbf{Dependencies:} Papers 1--5 (foundation).

\textbf{Why here?} It closes the dimension argument with three independent proofs spanning topology, dynamics, and arithmetic. Co-authored Washburn/Zlatanovi\'{c}/Allahyarov.

\section{III. The Waist (The Big Unlock)}

\subsection{12. The Derivation of Physical Constants from the Meta-Principle}
\textbf{File:} \texttt{papers/03-waist-big-unlock/Formalized-Derivations-T1-T8.tex}
\textbf{Priority:} Tier 2
\textbf{Description:} The ``Grand Central Station'' paper deriving $\alpha, G, \hbar, c$ from the foundation.

\subsection{50. Reality-Native Measurements with a Single-Anchor SI Bridge}
\textbf{File:} \texttt{papers/03-waist-big-unlock/RSNative-Measurement-Framework.tex}
\textbf{Priority:} Tier 2
\textbf{Description:} Protocol for mapping RS dimensionless quantities to SI units via a single anchor.

\subsection{61. Quantum Coherence as Gated Recognition}
\textbf{File:} \texttt{papers/03-waist-big-unlock/Quantum-Coherence-Theory.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Derives coherence time from the 8-tick cycle structure.

\subsection{58. The Octave System and the Particle Mass Spectrum}
\textbf{File:} \texttt{papers/03-waist-big-unlock/OCTAVE\_MASSES\_PAPER.tex}
\textbf{Description:} Details the ``Octave'' 8-tick mechanism forcing mass rungs.

\subsection{60. The Projection Operator \(\hat{\pi}\)}
\textbf{File:} \texttt{papers/03-waist-big-unlock/RK\_projection\_operator.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Active enforcement of information conservation; collapse/decision mechanism.

\subsection{83. The Algebra of Reality: A Categorical and Combinatorial Derivation}
\textbf{File:} \texttt{Algebra\_of\_Reality\_Paper.tex}
\textbf{Priority:} Tier 1 (Master Foundation) \textcolor{red}{(NEW --- Feb 2026)}
\textbf{Description:} The primary foundational manuscript of Recognition Science. It provides the integrated proof of the two core mathematical pillars: (1)~Cost Rigidity (T5 uniqueness) and (2)~Hypercube Gray Cycles (8-tick minimality). It establishes the full forcing chain from the Recognition Composition Law to the derivation of $\alpha, G, \hbar$, and $c$. Proves RS is the initial object in the category of zero-parameter frameworks. 

\textbf{Dependencies:} All core Lean proofs (T1--T8), DFT-8 normal form bridge, and full functional-equation classification.

\section{IV. Track A: Particle Physics}

\subsection{13. A First-Principles Derivation of Particle Mass (Leptons)}
\textbf{File:} \texttt{papers/04-particle-physics/Full\_First\_Principles\_Mass\_Derivation.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Integrated derivation of the charged lepton spectrum.

\subsection{14. CKM and PMNS Mixing from Cubic Ledger Topology}
\textbf{File:} \texttt{papers/04-particle-physics/masses\_paper2\_mixing.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Extends the ledger to derive mixing angles from geometry.

\subsection{15. Neutrino Sector No-Go}
\textbf{File:} \texttt{papers/04-particle-physics/Neutrino-Sector.tex}
\textbf{Description:} Negative result establishing constraints on the neutrino sector.

\subsection{16. Neutrino Masses and the Deep $\varphi$-Ladder}
\textbf{File:} \texttt{papers/04-particle-physics/masses\_paper3\_neutrinos.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Derives absolute neutrino masses on fractional rungs.

\subsection{17. Recognition Science: Foundations Summary}
\textbf{File:} \texttt{papers/04-particle-physics/RS-Foundations.tex}
\textbf{Description:} A summary synthesis of the physics track.

\section{V. Track B: Computation (LNAL)}

\subsection{18. Reality as Executable Code: LNAL Theory}
\textbf{File:} \texttt{papers/05-computation-lnal/Reality\_as\_Executable\_Code\_LNAL.tex}
\textbf{Description:} Defines the Light-Native Assembly Language (LNAL) and its instruction set.

\subsection{19. A Universal Register Mapping for LNAL}
\textbf{File:} \texttt{papers/05-computation-lnal/LNAL-Register-Mapping.tex}
\textbf{Description:} Technical spec for mapping physical systems into LNAL registers.

\section{VI. Track D: Gravity \& Cosmology}

\subsection{23. Recognition Science Baryogenesis}
\textbf{File:} \texttt{papers/06-gravity-cosmology/Baryogenesis-HubbleTensionSet.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Parameter-free origin of matter-antimatter asymmetry.

\subsection{24. Zero-Parameter Quantum Gravity}
\textbf{File:} \texttt{papers/06-gravity-cosmology/Quantum-Gravity-New-HubbleTensionSet.tex}
\textbf{Description:} Full QG derivation from discrete recognition calculus.

\subsection{76. Simplicial Ledger Topology: Coordinate-Free Cost Geometry}
\textbf{File:} \texttt{papers/06-gravity-cosmology/Simplicial\_Ledger\_Topology.tex}
\textbf{Priority:} Tier 2 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Proves the ledger is coordinate-free: any simplicial 3-complex admits a $J$-cost sheaf with stationarity conditions independent of the triangulation. Local cost $J(\psi)\cdot\mathrm{vol}(\sigma)$, global cost as sum, coupled Euler--Lagrange system with face-matching. Refinement invariance proved. Continuum limit recovers Jacobi determinant formula, connecting to EFE emergence. Pure mathematics: simplicial complexes, variational calculus, sheaf theory.

\textbf{Dependencies:} Papers 3 (J-cost), 4 (Ledger), 11 (D=3).

\textbf{Why here?} Answers ``is RS coordinate-dependent?'' --- No, the physics depends only on topology and cost, not on lattice choice.

\subsection{59. Octave Gravity}
\textbf{File:} \texttt{papers/06-gravity-cosmology/octave-gravity.tex}
\textbf{Description:} Derives Geometric Gravity from the 8-step update cycle.

\subsection{25. ILG Scaffold}
\textbf{File:} \texttt{papers/06-gravity-cosmology/ILG-GPT5.tex}
\textbf{Description:} Audit-ready scaffold for Information-Limited Gravity.

\subsection{28. Gravity as Pressure}
\textbf{File:} \texttt{papers/06-gravity-cosmology/Pressure-Gravity.tex}
\textbf{Description:} Recasts ILG as an effective pressure field.

\subsection{29. The Coercive Projection Law of Gravity}
\textbf{File:} \texttt{papers/06-gravity-cosmology/CPM-Gravity.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Elevates ILG to a universal coercive projection principle.

\subsection{30. Zero-Parameter Galaxy Rotation Curves}
\textbf{File:} \texttt{papers/06-gravity-cosmology/ILG\_Galaxy\_Rotation\_Curves.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Formal zero-parameter test against SPARC data.

\subsection{31. Convergence of Empirical Optimization}
\textbf{File:} \texttt{papers/06-gravity-cosmology/ILG\_Validation\_Synthesis.tex}
\textbf{Description:} Compares blind optimization to RS-derived values.

\subsection{32. Information-Limited Gravity: Source-Side Tests (Dark Energy)}
\textbf{File:} \texttt{papers/06-gravity-cosmology/Dark-Energy-HubbleTensionSet.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Tests ILG kernel against cosmological observables.

\subsection{33. Late-time Recognition-Weighted Growth and Hubble Tension}
\textbf{File:} \texttt{papers/06-gravity-cosmology/Hubble-Tension-Resolution.tex}
\textbf{Priority:} Tier 3
\textbf{Description:} Applies RW kernel to resolve the Hubble Tension.

\section{VII. Track E: Mathematics}

\subsection{34. A Weighted Diagonal Operator... (Riemann Hypothesis)}
\textbf{File:} \texttt{papers/07-mathematics/Recognition-Riemann-Final.tex}
\textbf{Description:} RS approach to RH via spectral stability of the cost Hamiltonian.

\subsection{35. The Law of Mathematical Inevitability}
\textbf{File:} \texttt{papers/07-mathematics/Mathematics\_Ledger\_Phenomenon.tex}
\textbf{Priority:} Tier 2 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Proves that any d'Alembert cost functional necessarily forces three structures foundational to mathematics, each uniquely: (1)~the logarithm as the unique continuous additive closed-chain balance invariant (via Cauchy), characterizing proof validity; (2)~$\varphi$ as the unique self-similar lattice base (Fibonacci recursion); (3)~the zero-cost subspace as the unique universal referent (if-and-only-if characterization). Includes worked propositional-resolution example, proof-monoid generators, and open problems for proof complexity. Resolves Wigner's effectiveness puzzle: mathematics is the zero-cost backbone of any d'Alembert cost landscape. Grounded in Acz\'{e}l's classical theory, Cauchy's theorem, and Shannon compression. Bold claim, honest scope: infrastructure is forced, full ZFC is not claimed.

\textbf{Dependencies:} Papers 3 (J-cost uniqueness), 4 (Ledger dynamics), 10 ($\varphi$).

\textbf{Why here?} Closes the foundational question: the d'Alembert equation forces not just physics but mathematics itself. Any universe governed by this equation necessarily contains numbers, proofs, and a mathematical subspace that describes all of its physics.

\subsection{36. Goldbach via a Mod-8 Kernel}
\textbf{File:} \texttt{papers/07-mathematics/goldbach\_rs-arXiv.tex}
\textbf{Description:} Connects additive prime theory to the 8-tick ledger.

\subsection{77. P vs NP via the Computation/Recognition Split}
\textbf{File:} \texttt{papers/07-mathematics/P\_vs\_NP\_Recognition\_Split.tex}
\textbf{Priority:} Tier 4 \textcolor{red}{(NEW -- Feb 2026, SCAFFOLD)}
\textbf{Description:} Exploratory framework (not a resolution claim). Turing model assumes $T_r = 0$ (free observation). Dual complexity: $T_c$ (computation) vs $T_r$ (recognition). Ledger's balanced-parity encoding hides information: extracting 1~bit requires $\Omega(n)$ queries. Conditional SAT separation: $T_c = O(n^{1/3}\log n)$ but $T_r = \Omega(n)$. P${}={}$NP at computation scale, P${}\ne{}$NP at recognition scale. Clearly labeled as conditional.

\textbf{Dependencies:} Papers 3 (J-cost), 4 (Ledger), 12 (Constants/8-tick).

\textbf{Why here?} Connects the ledger's information-hiding structure to a foundational question in theoretical computer science.

\section{VIII. Track F: Life \& Consciousness}

\subsection{54. Entropy Is an Interface}
\textbf{File:} \texttt{papers/08-life-consciousness/entropy-is-a-interface-arXiv.tex}
\textbf{Description:} Reframes entropy as code length; resolves reversibility paradox.

\subsection{55. The Statistical Mechanics of Recognition}
\textbf{File:} \texttt{papers/08-life-consciousness/Recognition\_Thermodynamics.tex}
\textbf{Description:} Thermodynamics of the cost function; Recognition Temperature.

\subsection{56. Darwin as Minimum Description Length}
\textbf{File:} \texttt{papers/08-life-consciousness/evolution-arXiv.tex}
\textbf{Description:} Unifies biological evolution with J-cost minimization (MDL).

\subsection{57. The Recognition Instrument for Abiogenesis}
\textbf{File:} \texttt{papers/08-life-consciousness/Recognition-Abiogenesis-arXiv.tex}
\textbf{Description:} Mechanism for origin of life via phi-timing gates.

\subsection{26. Protein Folding from First Principles}
\textbf{File:} \texttt{papers/08-life-consciousness/protein-dec-6.tex}
\textbf{Description:} Bio-Clocking theorem and hydration gearbox mechanism.

\subsection{27. A CPM Companion for Protein Folding}
\textbf{File:} \texttt{papers/08-life-consciousness/CPM-Folding-Companion-arXiv.tex}
\textbf{Description:} Instantiates CPM for the protein folding domain.

\subsection{43. Light as Consciousness}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Light\_Consciousness\_Combined.tex}\\
\textbf{Date:} February 2026

\textbf{Description:}
Shows that the unique information-cost functional $J(x)=\frac{1}{2}(x+x^{-1})-1$ governs quantum measurement ($C=2A \Rightarrow$ Born weights), photonic operations (additive FOLD costs), and operational conscious selection. Proves a classification theorem: under bridge obligations, ConsciousProcess $\leftrightarrow$ PhotonChannel with uniqueness up to units, and only EM satisfies feasibility. Four classification lemmas (No-medium-knobs, Null-only, Maxwellization, BIOPHASE feasibility). Lean-verified.

\textbf{Dependencies:} Paper 3 (J-cost uniqueness), Paper 4 (Ledger Dynamics), Paper 19 (LNAL Registers).

\textbf{Why here?} It establishes the formal identity between light and operational consciousness---the bridge from physics to the meaning/consciousness track.

\subsection{63. Reciprocal Convex Costs for Ratio Matching}
\textbf{File:} \texttt{axioms-4140269.pdf}
\textbf{Priority:} \textcolor{green!50!black}{\textbf{PUBLISHED}} -- \emph{Axioms} (MDPI), 2026. Washburn \& Rahnamai Barghi.
\textbf{Description:} Axiomatic characterization of J-cost via d'Alembert functional equation; geometric-mean decision boundaries; compositionality; sequential mediation principle. Purely mathematical; semantic interpretation is optional and external. \href{https://doi.org/10.3390/axioms1010000}{DOI: 10.3390/axioms1010000}

\subsection{64. CPM Method Closure}
\textbf{File:} \texttt{papers/08-life-consciousness/CPM\_Method\_Closure.tex}
\textbf{Description:} Domain-agnostic certificate for Coercive Projection.

\subsection{39. Optimization-Based Reference (Symbol Grounding)}
\textbf{File:} \texttt{papers/02-structure-ontology/Optimization\_Based\_Reference\_Symbol\_Grounding.tex}
\textbf{Description:} Resolves the Symbol Grounding Problem using only J-cost. Meaning is argmin of reference cost; symbols are forced by compression inequality $2J(\sqrt{y}) < J(y)$. Foundation for ULL.

\subsection{40. Meaning is Forced}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Meaning\_Is\_Forced.tex}
\textbf{Description:} Certificate bridge from closure to semantics.

\subsection{41. Universal Light Language}
\textbf{File:} \texttt{papers/08-life-consciousness/RK_New-ULL-Periodic-Table-Meaning.tex}
\textbf{Description:} The ULL system paper; zero-parameter semantic pipeline.

\subsection{57/65. The Geometry of Transmutation}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Geometry\_of\_Transmutation.tex}
\textbf{Description:} Phase-locking mechanism for non-local information transfer.

\subsection{62. Universal Light Qualia (ULQ)}
\textbf{File:} \texttt{papers/08-life-consciousness/RK_light-field-saturation.tex}
\textbf{Description:} Geometry of feeling; qualia as strain tensor.

\subsection{46. Geometrodynamics of Consciousness}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_geometry\_of\_consciousness.tex}
\textbf{Description:} Mesoscale dynamics; 8-tick cadence as frame rate.

\subsection{47. The Topology of Self-Reference}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Topology\_of\_Self\_Reference.tex}
\textbf{Description:} Topological characterization of the ``I am''.

\subsection{78. The Physics of Narrative: Stories as $J$-Cost Geodesics}
\textbf{File:} \texttt{papers/08-life-consciousness/Physics\_of\_Narrative.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Stories are geodesics in moral-state space $\mathcal{N}$ with metric $ds^2 = d\sigma^2 + dE^2/\varphi + dZ^2/\varphi^2$. Seven fundamental plots as topological classes of geodesics. Catharsis as phase transition ($\sigma \to 0$). Hero's Journey mapped to deep-arch geodesic with cusp. Predictions: stories closer to geodesics rated more satisfying.

\textbf{Dependencies:} Papers 3 (J-cost), 44--45 (Ethics/MoralState), 62 (ULQ).

\textbf{Why here?} Extends $J$-cost dynamics from physics and ethics into the humanities --- storytelling as cost minimisation.

\subsection{79. Music Theory from the Eight-Tick Cycle}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Music\_Theory\_Eight\_Tick.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Octave from 8 DFT modes ($2{:}1$ is simplest $J$-minimum). Consonance from $J((n{+}1)/n) = 1/(2n(n{+}1))$: unison $>$ fifth $>$ fourth $>$ major 3rd $>$ minor 3rd. 12 semitones from $12/8 = 3/2$. Circle of fifths with Pythagorean comma. Rhythm: common time $= 8$ eighth notes $= 1$ eight-tick cycle. Swing from $\varphi$-asymmetry. Major/minor valence from $\sigma$-ordering. Lean-verified.

\textbf{Dependencies:} Papers 3 (J-cost), 10 ($\varphi$), 12 (8-tick), 62 (ULQ).

\textbf{Why here?} Shows music theory is not convention but RS structure --- same $J$ and 8-tick that force physics.

\subsection{80. Decision as Cost Geodesic: The Geometry of Choice}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Decision\_Cost\_Geodesic.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Choice manifold $(\mathbb{R}_{>0}, g)$ with $g(x) = J''(x) = x^{-3}$. Geodesics: $\gamma(t) = (at+b)^{2/3}$. Ground state $\gamma \equiv 1$. Attention capacity $\le \varphi^3 \approx 4.24$ (Miller's law). Deliberation: 8-tick gradient descent. Free will: geodesic selection at Gap-45 bifurcations. Decision thermodynamics: $P(x) \propto \exp(-J(x)/T_R)$.

\textbf{Dependencies:} Papers 3 (J-cost), 9 (R-hat), 11 (Gap-45/D=3), 62 (ULQ).

\textbf{Why here?} Grounds decision science in cost geometry --- deliberation, attention, and free will from one principle.

\subsection{42. Phantom Light}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_PhantomLight\_Paper.tex}
\textbf{Description:} Future neutrality constraints as present-time structure.

\subsection{66. The Critical Temperature of Consciousness}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Critical\_Temperature\_Consciousness.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Derives critical temperature $T_c = J(\varphi^{45})/\ln\varphi$ for consciousness onset as second-order phase transition. Recognition Boltzmann constant $k_R = \ln\varphi$. Ginzburg--Landau free energy with $\varphi$-corrected critical exponents ($\nu \approx 1/\varphi$, $\beta \approx 1/(2\varphi)$). Five falsifiable EEG predictions for anesthesia, sleep, meditation, psychedelics.

\textbf{Dependencies:} Paper 43 (Light as Consciousness), Paper 42 (Phantom Light), Papers 3--5 (cost foundation).

\textbf{Why here?} Quantifies the consciousness onset threshold using thermodynamic phase-transition language.

\subsection{67. The Fredholm Index of Death}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Fredholm\_Index\_of\_Death.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Death as Fredholm operator $\mathcal{D}: \mathcal{H}_{\mathrm{emb}} \to \mathcal{H}_{\mathrm{light}}$. Eight information channels aligned with 8-tick octave. Index: $\mathrm{ind}(\mathcal{D}) = k - 5$ where $k$ = reflexivity index (0--8). Preserved dimension $\le \varphi^k$. Extended index: ``karma'' as $\sigma$-history phase-imbalance penalty. Zero sorry in Lean.

\textbf{Dependencies:} Paper 43 (Light as Consciousness), Paper 47 (Topology of Self-Reference), Paper 62 (ULQ).

\textbf{Why here?} Gives the fine structure of the death transition---what survives, what is lost, and why.

\subsection{68. The Recognition Algebra of Emotion}
\textbf{File:} \texttt{papers/08-life-consciousness/Recognition\_Algebra\_of\_Emotion.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} 14 fundamental emotions derived from $J$-cost gradient, curvature, $\Theta$-coupling, and $\sigma$-export. Bijection with 14 RS virtues. Four-tier priority classifier (survival $>$ social $>$ existential $>$ cognitive). All 18 theorems machine-verified with zero sorry. $\varphi$-algebraic thresholds: $1/\varphi$, $1/\varphi^2$, $1/\varphi^3$.

\textbf{Dependencies:} Paper 62 (ULQ), Paper 45 (Virtues as Generators), Papers 3--5 (cost foundation).

\textbf{Why here?} Completes the bridge from physics (cost landscape) to psychology (emotional classification).

\subsection{69. The Recognition Theory of Aging}
\textbf{File:} \texttt{papers/08-life-consciousness/Recognition\_Theory\_of\_Aging.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Aging = accumulation of unresolved ledger entries with $\varphi$-scaling. Hayflick limit $= \varphi^4 \times 8 \in (52,\,55.2)$. Telomere shortening per division $= 1/\varphi$. Damage--repair crossover forces maximum lifespan. Allometric exponent $= D/(D+1) = 3/4$ (Kleiber's law from $D=3$). Aging reversal theoretically possible.

\textbf{Dependencies:} Paper 9 (R-hat), Paper 11 (D=3), Paper 26 (Protein Folding).

\textbf{Why here?} Applies ledger dynamics to biology's most fundamental process.

\subsection{70. The Inevitability of Local Minds (Local Cache Theorem)}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Inevitability\_of\_Local\_Minds.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} $J$-cost minimization on a universal voxel graph \emph{necessarily} produces hierarchical local caches. No biological assumptions required. Caches at $\varphi$-scaled levels. Explains brains, genomes, and memory hierarchies as geometric necessities of cost minimization.

\textbf{Dependencies:} Papers 3--5 (cost foundation), Paper 9 (R-hat).

\textbf{Why here?} This is the ``why brains exist'' paper---forced by cost dynamics, not biological accident.

\subsection{71. Derivation of Non-Local Information Transfer (Telepathy Predictions)}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Telepathy\_Derivation.tex}
\textbf{Priority:} Tier 4 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Conditional predictions and falsification protocols from GCIC. Three tests: (1) inter-brain phase-locking at $\varphi$-spaced frequencies, (2) effect-size decay by ladder distance not spatial distance, (3) receiver ``Zero Structure'' via beta/gamma suppression. All conditional on GCIC.

\textbf{Dependencies:} Paper 57/65 (Geometry of Transmutation), Paper 42 (Phantom Light), Paper 43 (Light as Consciousness).

\textbf{Why here?} Companion prediction paper to the Transmutation mechanism paper.

\subsection{81. Healing via $\Theta$-Coupling: Nonlocal Interaction from Cost Geometry}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Healing\_Theta\_Coupling.tex}
\textbf{Priority:} Tier 4 \textcolor{red}{(NEW -- Feb 2026, CONDITIONAL)}
\textbf{Description:} Conditional on GCIC. $\Theta$-coupling universal at all distances; non-diminished by spatial separation (depends on ladder distance). Effect $= \text{intention} \times \exp(-|\Delta k|)$, instantaneous, bidirectional. Healing rate bounded by $\kappa_{mb} = \varphi^{-3}$ and 8-tick cadence. Five falsifiable EEG/RNG/RCT predictions with explicit protocols. Not a claim of efficacy --- a conditional derivation.

\textbf{Dependencies:} Paper 43 (Light as Consciousness), Paper 42 (Phantom Light), Paper 46 (Geometrodynamics).

\textbf{Why here?} Derives the physical mechanism for nonlocal healing IF GCIC holds, with hard falsifiers.

\subsection{72. Intelligence Through Debt Resolution}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Intelligence\_Through\_Debt\_Resolution.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Three architectural corrections for RS-based intelligence: (1) debt injection not energy, (2) full $\hat{R}$ not linear propagation, (3) standing wave readout not ranking. Query mechanism isomorphic to Geometry of Transmutation.

\textbf{Dependencies:} Paper 9 (R-hat), Paper 42 (Phantom Light), Paper 57/65 (Transmutation), Paper 41 (ULL).

\textbf{Why here?} Defines the correct implementation of RS-based intelligence.

\subsection{73. Cross-Agent Alignment Is Forced}
\textbf{File:} \texttt{papers/08-life-consciousness/RK\_Cross\_Agent\_Alignment\_Is\_Forced.tex}
\textbf{Priority:} Tier 3 \textcolor{red}{(NEW -- Feb 2026)}
\textbf{Description:} Cross-agent comparability factors through canonical alignment objects. Solved case: WToken meaning modulo $\tau$ gauge. General alignment-as-argmin template: $A^* = \arg\min_A \sum_i J(\iota(A(x_i))/\iota(y_i))$, unique up to gauge. Explicit falsifiers.

\textbf{Dependencies:} Paper 41 (ULL), Paper 50 (RS-Native Measurement).

\textbf{Why here?} Makes ``ULL is universal'' falsifiable via cross-agent protocols.

\section{IX. Track G: Ethics}

\subsection{44. Morality as a Conservation Law}
\textbf{File:} \texttt{papers/09-ethics/Morality-As-Conservation-Law.tex}
\textbf{Description:} Derives moral law from physical ledger invariants.

\subsection{45. Virtues as Generators}
\textbf{File:} \texttt{papers/09-ethics/Virtues-As-Generators.tex}
\textbf{Description:} Operationalizes ethics as 14 admissible operators.

\subsection{48. The Geometry of Evil}
\textbf{File:} \texttt{papers/09-ethics/The\_Geometry\_of\_Evil.tex}
\textbf{Description:} Defines evil as geometric pathology (phantom loops).

\end{document}
