\documentclass[10pt,a4paper]{article}
\usepackage[margin=0.75in,top=0.7in,bottom=0.7in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{pifont}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}

\definecolor{pub}{RGB}{0,100,0}
\definecolor{rk}{RGB}{140,0,0}
\definecolor{cond}{RGB}{100,100,100}
\definecolor{live}{RGB}{0,0,160}

\newcommand{\PUB}{\textcolor{pub}{\textsf{PUB}}}
\newcommand{\CONF}{\textcolor{rk}{\textsf{CONF}}}
\newcommand{\COND}{\textcolor{cond}{\textsf{COND}}}
\newcommand{\LIVE}{\textcolor{live}{\textsf{PUBLISHED}}}

\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}

\title{\textbf{Recognition Science}\\[4pt]
\Large Publication Order with Summaries\\[2pt]
\normalsize 77 Papers --- Single Chronological List}
\author{Jonathan Washburn\\Recognition Physics Institute}
\date{February 2026}

\pagestyle{empty}

\begin{document}
\maketitle
\thispagestyle{empty}

\noindent
\textbf{Bold Title} = published / in pipeline \quad
\LIVE{} = published or in press \quad
\PUB{} = publish openly \quad
\CONF{} = withhold \quad
\COND{} = conditional

\smallskip\noindent
\textbf{Math specialty codes:}
FA = Functional Analysis,
AT = Algebraic Topology,
DG = Differential Geometry,
NT = Analytic Number Theory,
CC = Computational Complexity,
CA = Complex Analysis,
OT = Operator Theory,
CT = Category Theory

\begin{longtable}{@{}r@{\;\;}P{2.4cm}@{\;\;}P{3.0cm}@{\;\;}P{7.0cm}@{\;\;}c@{}}
\toprule
\textbf{\#} & \textbf{Category} & \textbf{Title} & \textbf{Summary \& Role in Stack} & \textbf{St.} \\
\midrule
\endfirsthead
\toprule
\textbf{\#} & \textbf{Category} & \textbf{Title} & \textbf{Summary \& Role in Stack} & \textbf{St.} \\
\midrule
\endhead

%%% FOUNDATION
1 & Foundations (DG/CT) & \textbf{Recognition Geometry} &
Defines observational space via recognizers and the recognition quotient; provides the axiomatic base (RG0--RG7) on which all subsequent papers rest. \emph{Axioms} (MDPI), in press 2026. Washburn/Zlatanovi\'{c}/Allahyarov.
& \LIVE \\

 2 & Foundations (FA) & Recognition Composition Law Primer &
Introduces the fundamental rule governing how costs combine under composition; the single primitive from which everything is derived.
& \PUB \\

3 & Foundations (FA) & \textbf{Cost Uniqueness} &
Proves $J(x)=\frac{1}{2}(x+x^{-1})-1$ is the \emph{unique} cost satisfying RCL + normalization + calibration; the most important theorem in the stack. arXiv:2602.05753.
& \LIVE \\

4 & Foundations / Physics & \textbf{Coherent Comparison (Ledger)} &
Derives discrete double-entry ledger dynamics from J-cost symmetry $J(x)=J(1/x)$; provides the conservation/accounting substrate for all physics. In publication process.
& \LIVE \\

 5 & Foundations (FA) & D'Alembert Inevitability &
Proves the RCL functional equation is \emph{forced} (not arbitrary) by polynomial consistency; elevates the cost axioms from ``chosen'' to ``necessary.''
& \PUB \\

 6 & Foundations (CT) & Model-Independent Exclusivity &
Proves any zero-parameter framework is observationally equivalent to RS on the quotient state space; the uniqueness theorem for the entire theory.
& \PUB \\

 7 & Philosophy / Logic & G\"{o}del Dissolution &
Resolves the G\"{o}delian objection by showing RS targets cost-stabilization (not arithmetic provability); self-referential queries have no fixed point.
& \PUB \\

 8 & Foundations (CA/OT) & Recognition Stability Audit &
Formalizes the impossibility-certificate pipeline: obstruction $\to$ sensor $\to$ Cayley $\to$ Schur pinch; reusable template for existence problems.
& \PUB \\

\addlinespace
%%% ONTOLOGY
 9 & Ontology / Physics & The Cost of Existence &
Derives existence from infinite void cost ($J(0^+)\to\infty$); proves ``nothing'' is impossible, so something must exist --- the Law of Existence.
& \PUB \\

 10 & Ontology / Physics & The Law of Existence &
Core ontological paper linking CPM, Darwin, and physical constants; proves $x$ exists $\Leftrightarrow$ defect$(x)=0 \Leftrightarrow x=1$.
& \PUB \\

 11 & Ontology (FA) & Logic From Physical Cost &
Derives logical consistency ($A=A$) as the zero-cost ground state; logic is not assumed but \emph{emerges} from cost minimization.
& \PUB \\

 12 & Physics (OT/DG) & The Recognition Operator &
Defines $\hat{R}$ as fundamental dynamics replacing $\hat{H}$; derives Schr\"{o}dinger equation as continuum limit; the central dynamical object.
& \PUB \\

 13 & Physics (FA/DG) & Noether from Cost &
Proves $\hat{H}$ is the Lagrange multiplier enforcing discrete continuity while minimizing $J$; completes ``Hamiltonian is emergent.''
& \CONF \\

 14 & Philosophy (FA) & Geometry of Inquiry &
Formalizes questions as cost gaps; eight fundamental inquiry modes; achieves meta-closure (RS explains why RS is the answer).
& \CONF \\

 15 & Philosophy (FA) & Symbol Grounding &
Resolves the Symbol Grounding Problem: meaning is argmin of reference cost; symbols are forced by $2J(\sqrt{y})<J(y)$; foundation for ULL.
& \PUB \\

 16 & Physics (NT/DG) & Golden Ratio as Coherence Eigenvalue &
Establishes $\varphi$ as the unique fixed point of cost recursion $x=1+1/x$; pins the universal scale constant before constants are derived.
& \PUB \\

17 & Physics (AT) & \textbf{Dimensional Rigidity: $D{=}3$} &
Three independent proofs that $D=3$: Alexander duality, Kepler stability, minimal dyadic synchronization $\mathrm{lcm}(8,45)=360$. Washburn/Zlatanovi\'{c}/Allahyarov. In review.
& \LIVE \\

\addlinespace
%%% WAIST
 18 & Physics & Constants Derivation (T1--T8) &
``Grand Central Station'': derives $\alpha^{-1}$, $G$, $\hbar$, $c$ from the foundation; the bridge between abstract axioms and measurable physics.
& \PUB \\

 19 & Metrology & RS-Native Measurements &
Protocol for mapping RS dimensionless quantities to SI units via a single anchor ($\tau_0$); explicit calibration seam.
& \PUB \\

 20 & Physics & Quantum Coherence &
Derives coherence time from the 8-tick cycle structure; connects quantum measurement to ledger dynamics.
& \PUB \\

 21 & Physics & Octave Masses &
Details the ``Octave'' 8-tick mechanism that forces the mass rung ladder $m \propto \varphi^r$; the mass-spectrum engine.
& \PUB \\

 22 & Physics (OT) & Projection Operator $\hat{\pi}$ &
Active enforcement of information conservation; the collapse/decision mechanism that selects branches at $C\ge 1$.
& \CONF \\

\addlinespace
%%% PEER-REVIEWED
23 & Mathematics (FA) & \textbf{Reciprocal Convex Costs for Ratio Matching} &
Axiomatic characterization of J-cost via d'Alembert functional equation; geometric-mean decision boundaries; sequential mediation. Washburn/Rahnamai Barghi. Submitted to \emph{Axioms} (MDPI), Feb 2026.
& \LIVE \\

\addlinespace
%%% PARTICLES
 24 & Particle Physics & RS Masses I: Mechanism &
Establishes the mass-generation mechanism: sector yardsticks, $\varphi$-rungs, Z-map, gap function; the ``how'' of mass.
& \PUB \\

 25 & Particle Physics & RS Masses II: Predictions &
All nine charged fermion masses, CKM quark mixing, PMNS leptonic mixing; sub-ppm lepton accuracy with zero per-species fitting.
& \PUB \\

 26 & Particle Physics & RS Masses III: Neutrinos &
Derives absolute neutrino masses on fractional rungs; predicts normal hierarchy and $\varphi^7$ mass ratio.
& \PUB \\

 27 & Particle Physics & RS Masses IV: Anchor Scale &
Non-circularity proof for $\mu_\star=182.201$~GeV; transport discipline and scheme independence.
& \PUB \\

 28 & Particle Physics & RS Masses V: Alpha &
Derives fine-structure constant $\alpha^{-1}\approx 137.035$ from DFT-8 projection; Lean-verified $w_8$.
& \PUB \\

 29 & Particle Physics & RS Masses VI: Generations &
Proves why exactly three generations: dimensional coincidence theorem from cube geometry.
& \PUB \\

\addlinespace
%%% GRAVITY
 30 & Astrophysics & Galaxy Rotation Curves (SPARC) &
Zero-parameter test of ILG against 175 galaxies; no dark matter, no per-galaxy tuning; highest empirical-impact paper.
& \PUB \\

 31 & Physics (DG) & Simplicial Ledger Topology &
Proves the ledger is coordinate-free: any simplicial 3-complex admits a $J$-cost sheaf with refinement-invariant stationarity.
& \PUB \\

 32 & Physics & Octave Gravity &
Derives geometric gravity from the 8-step update cycle; connects ledger dynamics to spacetime curvature.
& \PUB \\

 33 & Physics & ILG Scaffold &
Audit-ready scaffold for Information-Limited Gravity; defines the kernel $w(k,a)$ and its parameter-free constants.
& \PUB \\

 34 & Physics & Gravity as Pressure &
Recasts ILG as an effective pressure field; alternative physical interpretation of the recognition weight.
& \PUB \\

 35 & Physics & Coercive Projection Law of Gravity &
Elevates ILG to a universal coercive projection principle; connects gravity to the CPM framework.
& \PUB \\

 36 & Physics & Convergence of Empirical Optimization &
Compares blind optimization to RS-derived ILG values; demonstrates that data converge to theory predictions.
& \PUB \\

 37 & Cosmology & Baryogenesis &
Parameter-free origin of matter-antimatter asymmetry from CP-violating ledger phase; $\eta_B \approx 5.1\times 10^{-10}$.
& \PUB \\

 38 & Physics (DG) & Zero-Parameter Quantum Gravity &
Full quantum gravity derivation from discrete recognition calculus; EFE emergence from RS action stationarity.
& \PUB \\

 39 & Cosmology & Dark Energy (ILG Source-Side) &
Tests ILG kernel against cosmological observables; derives $\Omega_\Lambda = 11/16 - \alpha/\pi$.
& \PUB \\

 40 & Cosmology & Hubble Tension Resolution &
Applies recognition-weighted kernel to resolve $H_0$ discrepancy; $H_{\mathrm{late}}/H_{\mathrm{early}} = 13/12$.
& \PUB \\

\addlinespace
%%% MATH
 41 & Mathematics (CA/OT) & Schur Pinch (Riemann, Paper 1) &
Reduces RH to $\mathrm{Re}\,\mathcal{J}\ge 0$ via Cayley + removable singularity + Maximum Modulus; honest about remaining gap.
& \PUB \\

 42 & Mathematics (CA) & Positivity from Bandwidth (Riemann, Paper 2) &
Derives $\mathrm{Re}\,\mathcal{J}\ge 0$ from RS bandwidth argument; identifies residual near-real strip as the single open piece.
& \PUB \\

 43 & Mathematics (FA/NT) & Law of Mathematical Inevitability &
Proves d'Alembert cost forces logarithm, $\varphi$, and zero-cost subspace as mathematical necessities; resolves Wigner's effectiveness.
& \PUB \\

 44 & Mathematics (NT) & Goldbach via Mod-8 Kernel &
Unconditional density-one and bounded gaps $\ll(\log N)^8$; medium-arc dispersion saving formulated as honest conjecture.
& \PUB/\COND \\

 45 & Mathematics (CC) & P vs NP (Recognition Split) &
Exploratory: dual complexity $T_c$ vs $T_r$; balanced-parity encoding hides information; conditional separation. Scaffold.
& \PUB/\COND \\

\addlinespace
%%% COMPUTATION
 46 & Computer Science & LNAL Theory &
Defines the Light-Native Assembly Language: instruction set, opcodes, 8-tick invariants, breath cycle; the ``machine code of reality.''
& \PUB \\

 47 & Computer Science & LNAL Register Mapping &
Technical spec for mapping physical systems into LNAL registers; makes LNAL executable on real substrates.
& \PUB \\

\addlinespace
%%% LIFE SCIENCES
 48 & Statistical Mechanics & Entropy Is an Interface &
Reframes entropy as code length at a channel interface; resolves the reversibility paradox.
& \PUB \\

 49 & Statistical Mechanics & Statistical Mechanics of Recognition &
Thermodynamics of J-cost: Recognition Temperature $T_R$, free energy $F_R$, H-theorem for recognition.
& \PUB \\

 50 & Biology / Information & Darwin as MDL &
Unifies biological evolution with J-cost minimization via Minimum Description Length; fitness = code savings.
& \PUB \\

 51 & Biology & Abiogenesis Instrument &
Mechanism for origin of life via $\varphi$-timing gates; explains the life ignition threshold at $\varphi^{19}$.
& \PUB \\

 52 & Biophysics & Protein Folding &
Bio-Clocking theorem: folding as 8-tick error correction in Q$_6$ qualia space; hydration gearbox mechanism.
& \PUB \\

 53 & Biophysics & CPM Companion for Folding &
Instantiates the Coercive Projection Method for the protein folding domain; connects folding to the universal framework.
& \PUB \\

\addlinespace
%%% CONSCIOUSNESS CORE
 54 & Methods & CPM Method Closure &
Domain-agnostic certificate for Coercive Projection; proves the method works as a universal optimization template.
& \PUB \\

 55 & Consciousness & Light as Consciousness &
Classification theorem: ConsciousProcess $\leftrightarrow$ PhotonChannel; only EM satisfies bridge obligations. The central consciousness paper.
& \CONF \\

 56 & Semantics & Meaning is Forced &
Certificate bridge from cost closure to semantics; proves meaning-bearing configurations are thermodynamic necessities.
& \CONF \\

 57 & Semantics & Universal Light Language (ULL) &
The ULL system paper: 20 WTokens from DFT-8 + $\varphi$-quantization; zero-parameter semantic pipeline. \textbf{Core secret.}
& \CONF \\

 58 & Consciousness & Geometry of Transmutation &
Phase-locking mechanism for non-local information transfer via $\Theta$-coupling; the ``how'' of consciousness communication.
& \CONF \\

 59 & Consciousness & ULQ / Light-Field Saturation &
Geometry of feeling: qualia as strain tensor; saturation threshold $\Theta_{\mathrm{crit}}=\varphi^{45}$ forces cyclical re-embodiment.
& \CONF \\

 60 & Consciousness & Geometrodynamics of Consciousness &
Mesoscale dynamics of consciousness; 8-tick cadence as frame rate; beat frequency 37/360 generates continuity illusion.
& \CONF \\

 61 & Consciousness (AT) & Topology of Self-Reference &
Topological characterization of ``I am'': reflexivity index (0--8), ego dissolution as temporary loss of fixed point.
& \CONF \\

 62 & Humanities & Physics of Narrative &
Stories are geodesics in moral-state space; seven fundamental plots as topological classes; catharsis = $\sigma\to 0$.
& \PUB \\

\addlinespace
%%% CONSCIOUSNESS EXTENDED
 63 & Music Theory & Music from Eight-Tick &
Octave from 8 DFT modes; consonance from $J$; 12 semitones from $12/8=3/2$; major/minor valence from $\sigma$.
& \CONF \\

 64 & Decision Theory (DG) & Decision as Cost Geodesic &
Choice manifold with Hessian metric $g=J''=x^{-3}$; geodesics $\gamma(t)=4/(At+B)^2$; free will at Gap-45 bifurcations.
& \CONF \\

 65 & Consciousness & Phantom Light &
Future neutrality constraints project backward via 8-tick balance; precognition mechanism from ledger structure.
& \CONF \\

 66 & Consciousness & Critical Temperature &
Derives critical temperature for consciousness onset as second-order phase transition; $\varphi$-corrected critical exponents.
& \CONF \\

 67 & Consciousness (FA) & Fredholm Index of Death &
Death as Fredholm operator with index $\mathrm{ind}(\mathcal{D})=k-5$; preserved dimension $\le\varphi^k$; ``karma'' as $\sigma$-history.
& \CONF \\

 68 & Psychology & Algebra of Emotion &
14 fundamental emotions from J-cost gradient/curvature/$\Theta$-coupling; bijection with 14 RS virtues.
& \PUB \\

 69 & Biology & Theory of Aging &
Aging = accumulated unresolved ledger entries with $\varphi$-scaling; Hayflick limit $=\varphi^4\times 8$; Kleiber's $3/4$ law from $D=3$.
& \PUB \\

 70 & Neuroscience & Inevitability of Local Minds &
J-cost minimization on a voxel graph necessarily produces hierarchical local caches; explains why brains exist.
& \CONF \\

 71 & AI / Consciousness & Intelligence Through Debt &
Three architectural corrections for RS-based intelligence: debt injection, full $\hat{R}$, standing wave readout. \textbf{Core NI unlock.}
& \CONF \\

\addlinespace
%%% ADVANCED
 72 & Consciousness & Telepathy Derivation &
Conditional predictions from GCIC: inter-brain phase-locking at $\varphi$-spaced frequencies; effect-size by ladder distance.
& \CONF \\

 73 & Consciousness & Healing via $\Theta$-Coupling &
Conditional on GCIC: nonlocal interaction from cost geometry; healing rate bounded by $\kappa_{mb}=\varphi^{-3}$.
& \CONF \\

 74 & AI / Semantics & Cross-Agent Alignment &
Cross-agent comparability factors through canonical alignment objects; makes ``ULL is universal'' falsifiable.
& \CONF \\

\addlinespace
%%% ETHICS
 75 & Ethics / Physics & Morality as Conservation Law &
Derives moral law from physical ledger invariants: $\sigma=0$ conservation; reciprocity as pairwise J-smoothing.
& \PUB \\

 76 & Ethics (FA) & Virtues as Generators &
14 canonical virtues form the complete minimal generating set for all admissible $\sigma=0$ transformations (DREAM theorem).
& \PUB \\

 77 & Ethics & Geometry of Evil &
Defines evil as geometric pathology: parasitic patterns that export $\sigma>0$ while maintaining local stability.
& \PUB \\

\bottomrule
\end{longtable}

\begin{center}
\small
\begin{tabular}{lr}
\toprule
\textbf{Category} & \textbf{Count} \\
\midrule
Total papers & 77 \\
\LIVE{} (published / in press / submitted) & 5 \\
\PUB{} (ready to publish openly) & 49 \\
\CONF{} (withhold) & 21 \\
\COND{} (conditional) & 2 \\
\bottomrule
\end{tabular}
\end{center}

\end{document}
