\documentclass[12pt]{article}
\usepackage{amsmath,amssymb}
\usepackage[hmargin=1in,vmargin=1in]{geometry}
\usepackage{enumitem}

\begin{document}

\begin{center}
{\LARGE \bf $\varphi$-Scheduler: Golden-Ratio Phase Scheduler for Multi-Actuator Control with Periodic Invariance, Interference Bounds, and Compliance API}\\[6pt]
{\large Provisional Patent Draft}\\[2pt]
\end{center}

\section*{Field}
This disclosure relates to real-time scheduling for multi-actuator control systems. It provides a scheduler that enforces \emph{$\varphi$-commensurate} update windows, guarantees \emph{periodic invariance} and \emph{interference bounds} between actuators, and exposes a \emph{compliance API} that any controller (including MPC or RL) can use while proving adherence to the schedule. The scheduler is domain-agnostic and applies to fusion devices, high-power lasers, robotics, beamlines, power electronics, and other multi-actuator plants.

\section*{Background}
Multi-actuator systems suffer cross-interference when control updates align unfavorably with plant modes. Conventional schedulers use fixed or equal-spaced slots that can reinforce low-order modal coupling and lack portable guarantees or machine-verifiable compliance. There is a need for a reusable module that (i) desynchronizes updates via irrational-ratio timing, (ii) guarantees periodic invariance for control proofs and certification, (iii) bounds interference, and (iv) provides a simple API and audit trail.

\section*{Summary}
A \emph{$\varphi$-scheduler} partitions each control period into phase windows with durations related by the golden ratio $\varphi=\tfrac{1+\sqrt{5}}{2}$, assigns actuators to phase sets, and \emph{enforces} that setpoint updates occur only within assigned windows. The schedule is periodically repeatable with declared jitter bounds. A \emph{compliance API} and hash-based audit logs allow third-party controllers to plug in while producing machine-verifiable conformance records; external signing is an optional integration if non-repudiation is required. Any claim about plant-level interference reduction is treated as a facility/model seam unless separately calibrated and validated.

\section*{Definitions}
\noindent\textbf{$\varphi$-commensurate windows.} Let $T>0$ be a control period. Define windows $W_0,\dots,W_{L-1}$ with durations $\Delta t_\ell>0$ such that
\[
\sum_{\ell=0}^{L-1}\Delta t_\ell = T,\qquad \frac{\Delta t_{\ell+1}}{\Delta t_\ell}\in\{\varphi,\varphi^{-1}\}\ \text{for admissible }\ell,
\]
with wrap-around interpreted modulo $L$ and optional subperiod patterns (e.g.\ $L=8$).\\[4pt]
\textbf{Actuator phase sets.} Each actuator $a$ is assigned $\Pi(a)\subseteq\{0,\dots,L-1\}$. Updates to $a$ may be applied only during windows $W_\ell$ with $\ell\in\Pi(a)$ and must satisfy dwell and slew constraints.\\[4pt]
\textbf{Periodic invariance.} The schedule repeats every $T$ with bounded jitter $\epsilon_j$ and supports periodic terminal sets/arguments in control proofs.\\[4pt]
\textbf{Interference bound (qualitative).} For band-limited cross-coupling kernels between actuators, the scheduler can be analyzed (under a facility-supplied plant model) to yield a bound
\[
\langle \text{cross-term}\rangle \le \kappa\,\langle \text{baseline cross-term}\rangle,
\]
with $0<\kappa<1$ determined by window smoothness and spectral separation. This bound is an optional engineering analysis that depends on a facility-specific plant model and is not part of the default certified surface.\\[4pt]
\textbf{Compliance API.} A programmatic interface that (i) issues window entry/exit events, (ii) accepts or rejects update requests according to $\Pi(a)$ and timing, and (iii) produces hash-based compliance records; external signing can be added as an integration option.

\section*{Detailed Description of Embodiments}
\subsection*{Scheduler Core}
Given $T$ and $L$, the scheduler constructs $\Delta t_{0:L-1}$ satisfying $\varphi$-commensurability and maps actuators $a$ to $\Pi(a)$. A high-resolution clock triggers \emph{BeginWindow($\ell$)} and \emph{EndWindow($\ell$)}. Update requests \emph{RequestUpdate($a$, payload)} are admitted only if the current window index $\ell\in\Pi(a)$ and actuator dwell/slew constraints are satisfied.

\subsection*{Periodic Invariance and Multi-Rate}
The scheduler supports superframes of duration $S\,T$ with $S\in\mathbb{N}$, preserving $\varphi$-relations within each $T$. Jitter $\epsilon_j$ is bounded by hardware timers or RTOS guarantees, enabling periodic MPC/RL safety proofs with phase-dependent terminal sets.

\subsection*{Interference Bounds}
For plants with dominant modal content below a cutoff, the scheduler’s irrational timing and window smoothness avoid low-order reinforcement. Over each period (or superframe), the time-averaged magnitude of bilinear cross-terms is reduced by a strict factor $\kappa\in(0,1)$ relative to co-phased or equal-spaced baselines. The factor is scheduler-intrinsic and controller-agnostic.

\subsection*{Compliance and Audit}
The compliance API exposes: \emph{WindowIndex()}, \emph{Allowed($a$)}, \emph{RegisterUpdate($a$)}, and \emph{GetComplianceReport()}. The report includes window timings, admitted/rejected updates, actuator IDs, and a cryptographic input digest (e.g., SHA-256 over canonicalized inputs) that binds the report to the schedule and events, enabling replay and audit. A hardware embodiment may use FPGA timers and secure elements for optional signing if required.

\subsection*{Example Domains}
\emph{Fusion:} gate NBI/ECRH/ICRH/RMP/pellets. \emph{ICF:} gate sub-pulse timings. \emph{Robotics:} desynchronize multi-axis torque updates. \emph{Power electronics:} gate converter switching setpoint updates. \emph{Beamlines/manufacturing:} stagger multi-actuator adjustments to reduce coupling.

\section*{Advantages}
The module is plug-and-play, controller-agnostic, and portable. It enforces schedule admissibility and produces verifiable compliance artifacts. Any claim about plant-level cross-interference reduction is treated as a facility/model seam unless separately calibrated and validated.

\section*{Implementation Evidence}
An implementation of the scheduler core and hash-based compliance report is provided in:
\begin{itemize}
  \item \texttt{fusion/simulator/control/phi\_window\_scheduler.py} (\texttt{generate\_phi\_window\_schedule}, \texttt{check\_allowed}, \texttt{compliance\_report\_certificate})
  \item \texttt{fusion/simulator/fusion/certificate.py} (\texttt{CertificateBundle}, \texttt{compute\_input\_hash})
\end{itemize}
Optional external signing/attestation is treated as an integration seam and is not required by the default implementation.

\section*{Claims}
\setlist[enumerate]{leftmargin=*,labelsep=0.5em,itemsep=0.5em,topsep=0.5em}

\subsection*{Independent Claims}
\begin{enumerate}
\item \textbf{(Method)} A method of scheduling updates in a multi-actuator control system, comprising:
\begin{enumerate}[label=(\alph*)]
\item partitioning each control period $T$ into $L$ phase windows $W_0,\dots,W_{L-1}$ with durations $\Delta t_\ell$ satisfying $\sum_{\ell}\Delta t_\ell=T$ and $\Delta t_{\ell+1}/\Delta t_\ell\in\{\varphi,\varphi^{-1}\}$ for $\varphi=(1+\sqrt{5})/2$;
\item assigning each actuator $a$ to a phase set $\Pi(a)\subseteq\{0,\dots,L-1\}$ and enforcing that setpoint updates to $a$ occur only during windows $W_\ell$ with $\ell\in\Pi(a)$;
\item guaranteeing periodic invariance by repeating the phase windows every $T$ with bounded jitter; and
\item providing a compliance application programming interface (API) that exposes window state, admits or rejects update requests according to the schedule, and generates machine-verifiable compliance records.
\end{enumerate}

\item \textbf{(System)} A scheduling system for multi-actuator control, comprising:
\begin{enumerate}[label=(\alph*)]
\item a timing engine that generates $L$ $\varphi$-commensurate phase windows within each period $T$;
\item a gating module that maps actuators to phase sets and enforces update admissibility only within assigned windows and subject to dwell and slew constraints;
\item a periodicity manager that maintains repetition of the schedule with bounded jitter; and
\item a compliance module providing an API for external controllers and producing hash-based compliance logs, optionally with cryptographic signing as an integration.
\end{enumerate}

\item \textbf{(Non-transitory medium)} A non-transitory computer-readable medium storing instructions that, when executed by one or more processors, cause a system to perform the method of claim 1.

\item \textbf{(Optional analysis)} The method of claim 1, further comprising computing an interference proxy bound for a facility-supplied plant cross-coupling model and recording the computed bound as an audit artifact, without asserting a facility-independent bound.
\end{enumerate}

\subsection*{Dependent Claims (Method of Claim 1)}
\begin{enumerate}[resume]
\item The method of claim 1, wherein $L=8$ and the window durations follow a repeating $\varphi/\varphi^{-1}$ pattern.
\item The method of claim 1, wherein each actuator update further satisfies minimum dwell time, minimum inter-update separation, and maximum slew rate constraints enforced within windows.
\item The method of claim 1, wherein the schedule is rescaled by a reference timescale $\tau_{\rm ref}$ while preserving $\varphi$-commensurability.
\item The method of claim 1, wherein a superframe of duration $S\,T$ is used, preserving the window structure within each period $T$.
\item The method of claim 1, wherein the compliance API comprises at least: \emph{BeginWindow($\ell$)}, \emph{EndWindow($\ell$)}, \emph{WindowIndex()}, \emph{Allowed($a$)}, \emph{RequestUpdate($a$, payload)}, and \emph{GetComplianceReport()}.
\item The method of claim 1, wherein compliance records include a cryptographic input digest and optionally are cryptographically signed using a hardware security element or secure enclave.
\item The method of claim 1, wherein the interference bound is obtained by selecting window smoothness profiles that limit spectral leakage of the update sequence.
\item The method of claim 1, wherein the scheduler rejects or defers update requests that would violate phase assignments or jitter bounds and records such events in the compliance log.
\end{enumerate}

\subsection*{Dependent Claims (System of Claim 2)}
\begin{enumerate}[resume]
\item The system of claim 2, wherein the timing engine is implemented on an FPGA or real-time microcontroller with hardware timers.
\item The system of claim 2, wherein the compliance module exposes a controller-agnostic interface usable by model predictive controllers, reinforcement learning policies, or heuristic controllers.
\item The system of claim 2, wherein the system provides a verification service that reconstructs schedule adherence from time-stamped actuator update events.
\end{enumerate}

\subsection*{Dependent Claims (Medium of Claim 3)}
\begin{enumerate}[resume]
\item The non-transitory computer-readable medium of claim 3, wherein the instructions enforce actuator phase sets and jitter bounds even when external controllers issue conflicting update requests.
\end{enumerate}

\section*{Exemplary Embodiments}
\emph{Fusion:} the scheduler gates neutral beam, RF heating, magnetic perturbation, fueling, and shaping updates to $\varphi$-phased windows. \emph{ICF:} the scheduler emits $\varphi$-spaced sub-pulse triggers. \emph{Robotics:} motor torque references are phase-staggered to reduce structural mode excitation. \emph{Power electronics:} setpoint updates for converters are staggered to avoid beat-frequency amplification. In all cases, the compliance API logs conformance.

\section*{Abstract}
A reusable $\varphi$-scheduler module partitions control periods into golden-ratio-commensurate phase windows, assigns actuators to phase sets, and enforces that updates occur only within assigned windows. The schedule is periodically repeatable with bounded jitter. A compliance API admits or rejects update requests and produces hash-based, machine-verifiable compliance logs, enabling controller-agnostic integration across domains under explicit seam declarations (e.g., any plant-level interference model and optional external signing).

\end{document}
